### 2023-02-08 Article/article.tex
first commit

\documentclass{article}
% \documentclass[review]{elsarticle}
\usepackage{epsfig}
\usepackage{amsmath}
\usepackage{mathtools}
\usepackage{algorithm}
\usepackage[algo2e,noend]{algorithm2e}
\usepackage[compatible,noend]{algpseudocode}
\usepackage{enumitem}
\usepackage{xspace}
\usepackage[table,xcdraw]{xcolor}
\usepackage{amsfonts}
\usepackage{pifont}
\usepackage{bbding}
\usepackage[english]{babel}
\usepackage{adjustbox}
\usepackage{tikz}
\usetikzlibrary{automata, topaths, calc, positioning, shapes, backgrounds, fit, matrix}
\usepackage{caption}
\usepackage{tabularx}
\usepackage[T1]{fontenc}
\usepackage{natbib}
\usepackage{datatool}
\usepackage{array}
\usepackage{pgfplotstable}
\pgfplotsset{width=7cm,compat=newest}
\usepackage{booktabs}
\usepackage{longtable}
\usepackage{pdflscape}
\usepackage{afterpage}
\usepackage{capt-of}
\usepackage{multirow}
\usepackage{float}
\usepackage[text={15cm,21cm},centering]{geometry}
\usepackage{setspace}
\usepackage{array}
\allowdisplaybreaks
\usepackage{lineno}
\usepackage{tkz-graph}
\usepackage{breqn}
\tikzset{
base/.style = {shape=rectangle, 
                      anchor=center,minimum size=5mm, 
                     align=center,
                     top color=white, ,inner sep=0ex,
                     minimum size=5mm
                     },
    LC11/.style =   {  base,   
                    text width=3em, 
                    bottom color=brown!100,
                    },
    UC11/.style = {     base,    
                    text width=5em, 
                    bottom color=brown!100,
                    align = right,
                },
    LC12/.style =   {  base,   
                    text width=1em, 
                    bottom color=orange!100,
                    },
    UC12/.style = {     base,    
                    text width=2.5em, 
                    bottom color=orange!100,
                },
    LC21/.style =   {  base,   
                    text width=1.5em, 
                    bottom color=brown!100,
                    },
    UC21/.style = {     base,    
                    text width=3.5em, 
                    bottom color=brown!100,
                },
    LC22/.style =   {  base,   
                    text width=4em, 
                    bottom color=blue!100,
                    },
    UC22/.style = {     base,    
                    text width=7em, 
                    bottom color=blue!100,
                    align = left,
                },
    UC221/.style = {     base,    
                    text width=3.5em, 
                    bottom color=blue!100,
                    % color=blue!100,
                    outer xsep=0em,
                    % align = right,
                },
                UC231/.style =   {  base,   
                    text width=1.5em, 
                    bottom color=blue!100,
                    },
    LC23/.style =   {  base,   
                    text width=1.5em, 
                    bottom color=orange!100,
                    },
    UC23/.style = {     base,    
                    text width=2.em, 
                    bottom color=orange!100,
                },
  env/.style      = {base, font=\ttfamily\normalsize},
label/.style   = {base,font=\ttfamily\normalsize, text width = 8em, 
  inner sep=0em, rounded corners=1mm,bottom color=gray!80,},
  dummy/.style    = {circle,draw}
}

\newtheorem{theorem}{Theorem}[section]
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{proposition}[theorem]{Proposition}
\newtheorem{corollary}[theorem]{Corollary}
\newtheorem{definition}[theorem]{Definition}
\newcommand{\pcmaxc}{P$|\mbox{\em cont}|$C$_{\max}$\xspace}
\newcommand{\alns}{adaptive large neighborhood search\xspace}
\newcommand{\Alns}{Adaptive large neighborhood search\xspace}
\newcommand{\F}{$\mathcal{F}$\xspace}
\newcommand{\B}{$\mathcal{B}$\xspace}
\newcommand{\C}{$\mathcal{C}$\xspace}
\newcommand{\Lb}{$\mathcal{L}$\xspace}
\newcommand{\D}{$\mathcal{D}$\xspace}

\renewcommand{\algorithmicforall}{\textbf{for each}}
\SetKwFor{ForEach}{for each}{}{}%
\renewcommand{\algorithmiccomment}[2][1\linewidth]{%
\leavevmode\hfill\makebox[#1][l]{/* \textbf{~#2} */}}
\renewcommand{\algorithmicrequire}{\textbf{Input:}}

\renewcommand{\tabcolsep}{4pt}

\modulolinenumbers[1]
\linenumbers

\oddsidemargin=0.5in
\topmargin=-0.25in
\textwidth=5.5in
\textheight=8.25in

%\journal{European Journal of Operational Research}

%\begin{frontmatter}

\title{A GRASP algorithm for the concrete delivery problem. }
\author{Ousmane Ali$^{(1)}$, Jean-Fran\c cois C\^ot\'e$^{(2)}$, Leandro C.~Coelho$^{(3)}$\\
 $(1)$ {\tt nassoma-wattara-ousmane.ali.1@ulaval.ca}\\
 $(2)$ {\tt Jean-Francois.Cote@fsa.ulaval.ca}\\
 $(3)$ {\tt Leandro.Coelho@fsa.ulaval.ca}\\
}

% \date{\today}

\begin{document}
\maketitle
\begin{abstract}
    %We study the daily scheduling of concrete delivery. Truck drivers are dispatched based on assignment priority and remaining working time over the horizon planning. A customer may be served from several available production plants. Due to the heterogeneity of the fleet, the number of visits to a customer location is not known ahead of time. The objective is to schedule drivers while minimizing travel times, waiting duration, idle time, and overtime. We solve this problem using a greedy randomized adaptive search procedure. Computational experiments with real data show the performance of our methods over the approach used by a local company. 
\end{abstract}

% \begin{keyword}
\noindent{\textbf{Keywords:} Vehicle routing; scheduling; ready-mixed concrete; concrete delivery
% \end{keyword}

%\end{frontmatter}

\section{Introduction}
\label{sec:intro}

Construction projects involve a huge movement of equipment, people, and materials. Among the latter, concrete is one of the most used. It is a perishable product with many factors affecting its quality \citep{sinha_quality_2021}. It comes in two types: ready-mixed concrete (RMC) and site-mixed concrete (SMC). As its name implies, SMC is produced on the spot using raw materials (water, aggregates, and cement) stored on the construction site, while RMC is manufactured in a batch plant and delivered to the construction site. SMC can avoid delays caused by road traffic, but has a slower and more difficult production process, requires storage for mixing materials and equipment, and is suitable for low amounts of concrete. It is quite the opposite for the RMC, which has better quality and benefits from lower production cost \citep{muresan_comparing}. Nevertheless, the batch plant manager must employ a fleet of high-cost revolving drum trucks (concrete mixers) to dispatch the ready-mixed concrete. Hence, the key to reaping the RMC benefits is to ensure efficient and prompt delivery on the construction site.

Concrete delivery under the form of RMC is subject to many operational constraints that make it a challenging problem met in Operations Research and addressed by the Concrete Delivery Problem (CDP). In this paper, we study a variant of the Concrete Delivery Problem to schedule the daily production and dispatching of RMC for a company located in the province of Quebec, Canada. This company produces RMC in multiple batching plants with different production rates, using a fleet of concrete mixers of various capacities. Each plant has its specific fleet, but a truck can load elsewhere if deemed necessary, except it must return to its home plant at the end of the day. They own two types of trucks with different capacities but can call on an external fleet when needed. They serve a construction site from any batch plant, with the first delivery starting at the time specified by the customer.  Loading and unloading a concrete mixer depends on the truck capacity, a plant's loading, and a construction site's unloading rate, respectively. The problem is like the one addressed in \cite {schmid2009hybrid, schmid2010hybridization}, except for the plant's production rate and the lack of unloading instrumentation in our case. Also, as per union rules, they assign a truck driver to a delivery task based on his seniority. The company uses a centralized dispatcher system to schedule all daily orders, but this system has issues satisfying all daily demands without using an external fleet.

According to \cite{blazewicz2019handbook}, the concrete delivery problem combines vehicle routing with scheduling issues to plan routes to deliver concrete from batch plants (depots) to customers' construction sites. Ready-mixed concrete is an on-demand product with a short life cycle from production through end use. It cannot be stored and cannot stay too long in a truck, or it will harden. Hence, concrete mixers must deliver RMC at the planned construction site shortly after its production. RMC production and delivery is then an example of a Just-In-Time (JIT) production system in construction \citep{tommelein1999just}. A customer quantity requirement is often greater than the truck size and must be fulfilled by multiple deliveries. In that sense, CDP is like the Vehicle Routing Problem (VRP) with split delivery (VRPSD) \citep{archetti2008split}, except that the same truck may visit a customer more than once.

Since concrete hardens quickly, in case of multiple deliveries, back-to-back deliveries must be continuous or at least close in time to avoid the problem of cold joint that negatively affects the quality of the concrete. Customers request to be served within a specific time window, complicating truck loading schedules when a plant can only load one truck at a time. Likewise, often only one truck can unload at a time at a customer location, sometimes leading to concrete mixers queuing and waiting their turn to deliver. Furthermore, with trucks of varied sizes, loading, travel, and unloading times that may be uncertain, the Concrete Delivery Problem is a complex problem to solve and has been proven by \cite{asbach2009analysis} to be NP-hard. With a single batching plant, the objective is to maximize the service level (serve all customers and avoid delays between subsequent deliveries and truck queuing). With additional production centers to choose from, a minimization of the total travel time is added to the objective. %However, these objectives can be conflictive.   
% We go a step further to add human constraints to the CDP, overtime, minimum and maximum hour of work,

The remainder of the paper is organized as follows. Section \ref{lit_review} presents an overview of the CDP-related literature.  Section \ref{desc_form} provides a formal description of the problem. Then in Section \ref{method}, we describe the Greedy Randomized Search Procedure (GRASP) algorithm and the constructive heuristics we design to solve this variant of the CDP. Computational experiments and our conclusions follow in Sections \ref{comp_exp} and \ref{concl}, respectively.

\section{Literature review}
\label{lit_review}

A text mining approach for reviewing the ready-mixed concrete literature \citep{maghrebi2015text} showed that concrete technology and material science are the main core of research in this area. The first academic research on concrete batching and delivery began in the late 1990s. \cite{tommelein1999just} describe RMC as a prototypical example of a Just-In-Time production system in construction and identify two practices occurring when delivering it. An alternative is for the customer to haul the product from the batch plant with their concrete mixer. The other approach is where the batch plant delivers the concrete to the customer's location. This latter approach is the one studied in all related papers found in the literature.

Several works on the CDP mainly focussed on Simulation related methods, either standalone \citep{zayed2001simulation, wang2001scheduling, tian_simulation_based_2010, panas_simulation_based_2013, galic2016simulation}, etc. or hybridized with optimization methods \citep{feng2004optimizing, lu2005optimized, feng_integrating_2006}, etc. to schedule and dispatch concrete production and delivery. \cite{wang2001scheduling} developed a simulation model to reveal the effect and value of the concrete mixers' inter-arrival time on the productivity of hired unloading equipment on site. With a combination of Genetic Algorithm (GA) and simulation process, \cite{feng2004optimizing} tried to minimize the total waiting time for trucks at a customer site. Loading trucks with identical capacities is carried out at the same batch plant. Loading and unloading durations are fixed. GA is used to find the most efficient and effective loading sequence of RMC trucks to be assigned to different construction sites. The simulation process determines the loading, arrival, departure, and waiting time of trucks and thus evaluates the cost of each dispatching sequence. They evaluated the method using data from a batch plant in Taiwan with up to nine customers served. \cite{mayteekrieangkrai2015optimized} solved the same problem with the same data using a bee algorithm (BA) and found better solutions than the GA. \cite{lu2005optimized} used the same combination of GA and simulation to decide on the optimal number of concrete mixers to be deployed together with an optimal schedule for batching and delivering concrete. Their objective is to minimize the site crew idle times due to late concrete deliveries plus truck queuing time. The particularity of this setting is that mortar useful for unloading pump lubrication must be delivered on-site before concrete. Therefore, mortar batching and delivery are also modeled in the simulation. Finding the best RMC truck size was also the purpose of the discrete-event simulation model proposed by \cite{panas_simulation_based_2013}.

Besides simulation-based methods, we find in the literature works using metaheuristics \citep{faria2006distributed, misir2011selection, maghrebi2016sequential, yang2022concrete}, exact methods \citep{yan2007optimal, kinable2014concrete}, matheuristics \citep{schmid2009hybrid, schmid2010hybridization}, Benders Decomposition \citep{maghrebi2014benders}, Column Generation (CG) \citep{maghrebi2014solving, maghrebi2016column}, Lagrangian relaxation \citep{narayanan2015using}, and machine learning approach \citep{graham2006modeling, maghrebi2014exploring, maghrebi2016matching}.
\cite{matsatsinis2004towards} designed a decision support system (DSS) for the dynamic routing of both concrete and pumps that may be necessary for some sites to help the concrete unloading. Three plants are available, but vehicles fulfilling the same order must all load at the same plant. An order that cannot be executed may be postponed for the next day. The routing of the pumps is modeled as a multi-depot VRP with time windows. \cite{naso2007genetic} proposed a sequential GA method combined with constructive heuristics to solve a more general variant of the CDP. In this problem, besides batching concrete delivered to a customer site, the plant's production schedule must include orders picked up by the customers themselves. The algorithm first schedules the plant loading operations before scheduling truck job deliveries. A non-linear model minimizing transportation costs, waiting times, outsourced costs, and overtime work is also developed. The authors ran experiments using real-world instances of a concrete supply chain in the Netherlands. They found a reduction in the number of requests redirected to external companies. \cite{yan2007optimal} also emphasized overtime considerations in their paper scheduling RMC for one batching plant with two loading docks. Overtime wages are paid for factory and construction site operations after 4 PM. A mixed integer programming (MIP) model on a time-space network is developed to minimize travel times and operating costs at both normal and overtime working hours at the plant and the construction sites. Real data consisting of 3 days of operation is tested using a two-stage algorithm. First, they solved the MIP relaxation with CPLEX. Then they simplify the original model by fixing some decision variables before solving it. This algorithm seems to improve the actual plant operation by 10\%.
A time-space network is the principal component of the real-time DSS developed by \cite{durbin2008or} to solve a dynamic CDP every five minutes. The DSS can receive new orders, schedule them on the fly, and deal with unexpected events such as plant closures, truck breakdowns, delays in transportation times, etc. Combined with a Tabu Search (TS) heuristic to warm start CPLEX, the model is performant enough to solve instances with up to 1,500 loads per day with up to 250 trucks. The authors consider the case of a customer who places two orders, with the first being completed before the second starts. The real-time planning and monitoring of CDP are studied more in detail by \cite{garza2021dynamic} in his thesis. Another variant of the CDP is modeled by \cite{schmid2009hybrid} as an integer multicommodity network flow (MCNF) problem on a time-space network. In this paper, concrete is delivered using a heterogeneous fleet of vehicles, and each plant can load an illimited number of trucks simultaneously. Among the trucks, some have specialized equipment and must arrive first at certain construction sites to assist concrete-mixers unloading. The objective is to fulfill all orders, minimize the travel cost, and avoid delays between two consecutive unloading operations for an order. The model is typically solved using a matheuristic algorithm that combines the MCNF with a variable neighborhood search (VNS) heuristic. The method can solve large problem instances with more than 60 orders per day quickly without encountering any memory issues. The same problem is addressed by \cite{schmid2010hybridization}. The authors proposed a MIP model combined with a VNS and a very large neighborhood search (VLNS) to develop two matheuristics approaches. Comparisons between both matheuristics and a standalone VNS show that the former methods are much better and suitable for solving larger problem instances. These methods also provide better solutions for small to medium instances than the matheuristic used in \cite{schmid2009hybrid}. A pure VNS approach with the same problem but without the use of instrumentation has been applied by \cite{payr2009optimizing}.

Regarding objectives, most authors have been interested in minimizing travel time and delays between consecutive deliveries. Some authors, however, were more interested in maximizing only customer satisfaction. We find these situations in the works of \cite{durbin2008or, kinable2014concrete, kinable2014logic, sulaman2017simulated}.
\cite{kinable2014concrete} introduce a general MIP and constraint programming (CP) models of the CDP reflecting the main constraints commonly found in all CDP works: time lag and no overlapping between consecutive deliveries, covering of all customers' demands, delivery time window, and heterogeneous fleet. However, the model did not include constraints limiting the time that concrete may reside in a truck. The authors also propose a constructive heuristic that schedules the visits to the customers one by one according to the start time of the visit and the truck capacity. The procedure is invoked multiple times for different permutations of the customer's order which is determined using the steepest descent (SD) local search procedure. One of the paper's main contributions is the creation of the first public test instances for the CDP with up to 50 customers, four batching plants, and 20 concrete mixers. They found the CP model to be highly effective in finding high-quality solutions in relatively little time or improving existing schedules, while the MIP model can be used to compute bounds, as it seems ineffective in solving large problem instances. Finally, the heuristic often yields good solutions in less than a second. A detailed analysis of the MIP model presented in cite{kinable2014concrete}and of two more compact models can be found in the thesis of \cite{hernandez_lopez_study_2020}. In \cite{kinable2014logic}, we find an attempt to solve the previous problem with a Logic Based Benders' approach. \cite{sulaman2017simulated} expand upon the SD heuristic proposed in \cite{kinable2014concrete}, proposing a Simulated Annealing (SA) with a Time-Slot Heuristic (TH). TH mechanism is to look for a slot between existing visits of a truck to schedule a new delivery instead of assigning it to the time slot strictly after the truck's latest assigned delivery. The goal is to reduce the large time gaps that can be present in a schedule created with SD due to ignoring the intermediate available time slots. Experimental results indicated that SATH outperforms SD in speed and solution quality. A generalization of the MIP model of \cite{kinable2014concrete} is addressed in \cite{asbach2009analysis}. This model simultaneously minimizes the total sum of travel costs and the penalty costs for customers with unfulfilled demand. A customer can request that all concrete deliveries come from the same plant or a subset of plants and that a delivery truck belongs to a subset of the vehicle fleet. The MIP model is used in a local search scheme as a black-box solver to reoptimize an incumbent solution in which a neighborhood operator has unfixed some variables.

%\cite{tzanetos2023systematic} present a systematic search and mapping review of the concrete delivery problem. this paper fills that gap by (a) performing a comprehensive search on the CDP, and (b) mapping out and categorizing the existing literature. Specifically, this paper provides an overview of the various methods of the literature, categorizing the problem formulations based on the literature’s different concepts,

\section{Problem description}
\label{desc_form}
The problem studied within this paper is about the distribution of ready-mixed concrete from a Canadian company operating in the greater Montreal area. Concrete order from a customer arrives at a central center and is assigned to one of the batching plants from where the product will be produced and delivered to the customer. We define an instance of our problem on a set of batching plants, a set of customer orders, and a set of drivers.

% \subsection*{Batching plants}

The company owns and operates a total of eight concrete batching plants located in different geographical areas. The loading bays in each plant can only accommodate one truck at a time, resulting in trucks lining up. Let $\mathcal{B}$ be the set of batching plants. The plants are heterogeneous, as each plant $b$ has its hourly loading rate $\tau^l_b$ which influences the loading time duration. After loading, a driver takes $\alpha_b$ minutes to adjust the concrete in his truck before departing to the customer site. Any batch plant can serve a construction site if the traveled time between them is less than the concrete lifespan $\delta$. Each plant has its assigned fleet, but it can get a vehicle from another plant or even hire an external fleet. If $q_i$ is the quantity of RMC loaded to customer $i$ from plant $b$, we compute the loading time duration $LD^b_i$ as

\begin{equation}
    \label{eq:LD}
    {LD}^b_i = \frac{q_i}{\tau^l_b}.
\end{equation}

% \subsection*{Customer orders}

A customer $i$ requests one or several types of concrete to be delivered to his construction site on a particular day for a total amount of $q_i$. When placing an order, he specifies the number of types of RMC $P_i \geq 1$, the quantity $q^p_i$ of concrete of each type $p$, the arrival time $a_i$ of the first concrete mixer, and the unloading rate $\tau^u_i$ at his site. $\mathcal{C}$ is the set of construction sites (customers) with a planned delivery for the day.

We refer to order $o^p_i$ as the request of customer $i$ to receive a type $p$ of concrete. If the amount of concrete exceeds any truck capacity, several deliveries are scheduled to satisfy an order. $n^p_i \geq 1$ is the number of deliveries needed to satisfy $o^p_i$. Let $d^p_{ij}$ be the $j_{th}$ visit with load $q^p_{ij}$ for order $o^p_i$. We represent the delivery of order $o^p_i$ by the visits to the ordered set of nodes $\mathcal{D}^p_i= (d^p_{i0},d^p_{i1},\cdots, d^p_{in^p_i})$. All quantities of a type of concrete must be thoroughly delivered before moving on to another type. The orders $o^p_i$ $(1 \leq p \leq P_i$) of a customer $i$ are not ordered. An order can start at any time, provided it is completed before delivering another. We represent the delivery of customer $i$ by the set $\mathcal{D}_i= (\mathcal{D}^{p^0}_i, \mathcal{D}^{p^1}_i,\cdots,\mathcal{D}^{p^{P_i}})$, where $p^l$ is the $lth$ order delivered. We define for the first node of $\mathcal{D}^{p^0}_i$ the time window $ \left\lbrack a_i, b_i  \right\rbrack$ to ensure that the first delivery of $i$ must be due at $a_i$. Once a plant is chosen to produce a customer's first order, it must be the provider of all subsequent orders of this customer.

To prevent cold joint in concrete, subsequent deliveries must be in just in time, or at least close in time. We define a maximum time lag $\gamma_i$ beyond which no next unloading operation should be allowed. The construction site unloading rate and the quantity to unload give the time necessary to discharge a truckload. The unloading time duration of $q^p_{ij}$ is
\begin{equation}
    \label{eq:UD}
    {UD}_{d^p_{ij}} = \frac{q^p_{ij}}{\tau^u_i}
\end{equation}

% \subsection*{Drivers}

%A driver is an important entity here, as we must incorporate some union constraints into our model. 
The company has two types of trucks with capacities of 8 and 12 $m^3$. A driver $k$ is assigned to a plant from where he starts and ends his shift day, and drives one concrete mixer of capacity $Q_k$. The set of drivers is $K=\cup_{b \in \mathcal{B}}K_b $, $K_b$ being the set of drivers scheduled to start their shift at the batching plant $b$.

A driver must work between four and eight hours in normal time, plus optional overtime of two hours at most, during a scheduled day. A driver mainly loads RMC at his home plant but can still drive at other plants and load there if needed. Batching plant produces concrete on-demand with recipes specific to customers, which means that truck cannot hold orders for more than one client, even if it has spare capacity. Thus, a driver must refill at a plant between two consecutive deliveries. After unloading the RMC, a driver takes $\beta_k$ minutes to clean the concrete mixer before traveling to his next loading plant. When assigning a driver to a delivery, priority is given to the employees with the highest seniority.

For an order $o^p_i$, the number of deliveries $|D^p_i|$ is not known a priori, as we use a fleet with various capacities. However, we can compute the lower and upper bounds of $|D^p_i|$ using the capacities of the highest ($Q_{max}$) and smallest ($Q_{min}$) available trucks. The lower (upper) bound $\left\lceil \frac{q^p_i}{Q_{max}} \right\rceil$ $\left(\left\lceil \frac{q^p_i}{Q_{min}} \right\rceil \right)$ is the number of deliveries needed if we only use trucks of capacity $Q_{max}$ ($Q_{min}$).

% \subsection*{Model}

A solution to our problem is a set of decisions about truck loading schedules, driver assignments for different deliveries, and truck arrival times at construction sites for unloading. For a batching plant, the decision is to determine which driver should be loaded, when, and for which construction site. For a driver, the decision is the sequence of loading depots and delivering sites. And for a construction site, the arrival time of all daily scheduled deliveries.

Let the binary variable $x^{bk}_{j}$ be $1$ if a driver $k$ loads at plant $b$ to serve the delivery node $j$. % $y_i=1$ if $i$ is completely served.
Let $SLT^{bk}_{j}$ and $ELT^{bk}_{j}$ be the start and end loading time of driver $k$ at plant $b$. $SUT_{j}$, $EUT_{j}$ be the start and end unloading time. $LTS^{bk}_{j}=\left[ SLT^{bk}_{j}, ELT^{bk}_{j} \right] $, and $UTS^{bk}_{j}=\left[ SUT^{bk}_{j}, EUT^{bk}_{j} \right]$ stand for the loading  and unloading operations timeslots. $AT_{j}$ is the arrival time of a driver at the node location.

\begin{alignat}{2}
    \label{eq:ELT}  ELT^{bk}_{j} &= SLT^{bk}_{j}+ LD^b_{j}\\
    \label{eq:S}         EUT_{j} & = SUT_{j} + UD_{j}
\end{alignat}

For each delivery node $j$, we define an expected delivery time $EDT_{j}$ of unloading operation, the driver waiting duration $W^k_{j}$, and the node waiting time $W_{j}$.

\begin{alignat}{3}
    \label{eq:ES} EDT_{d^{p^l}_{ij}} &=
          \begin{cases}
            a_i & \text{if $j=0$, $l=0$}\\
            E_{d^{p^l}_{ij-1}} & \text{ $j \geq 1$}\\
            E_{d^{p^{l-1}}_{in^{p^{l-1}}_i}}  & 
          \end{cases} &  \\
    \label{eq:Wd} W^k_{d^p_{ij}} &= \max\{0, SUT_{d^p_{ij}} - AT_{d^p_{ij}}\} & \\
    \label{eq:Wc} W_{d^p_{ij}} & = \max\{0, AT_{d^p_{ij}} - EDT_{d^p_{ij}}- \lambda_i \} &
\end{alignat}

% Parler de la reduction des W

We keep track of the working time of each driver $k$ with the variable $WT_k$. Parameters $N_t$ and $O_t$ are the normal and overtime working hours. $M_t$ is the minimum working time for a day. $\Gamma_1$ and $\Gamma_2$ are the incurred penalties when a driver works less than $M_t$ and more than $N_t$.
\begin{equation}
    \forall k \in K, WT_k \leq O_t 
\end{equation}

The objective is to minimize the traveled distance,
 of delivered concrete to the satisfied customers as in

\begin{equation}
    \min \smashoperator{\sum_{i,j}}{t_{ij}x_{ij}}
\end{equation}

The cost to travel between two nodes ($i$, $j$) is the travel time $t_{ij}$. COST...

\paragraph*{}
Let us consider the small instance illustrated in Figures~\ref{fig_Example} and \ref{fig:ganttExample}. Two batching plants $B_1$ and $B_2$ with three drivers serve two construction sites. Customer $C_1$ requests one order of $q^1_1=12$ m$^3$. Customer $C_2$ requests three orders of $q^1_2=3$, $q^2_2=11$, and $q^3_2=1$ m$^3$ of three different types of concrete. $B_1$ has two drivers ($D_1$, $D_2$) using concrete mixers of capacity 8 m$^3$, and $B_2$ has $D_3$ of capacity 12 m$^3$. Figure~\ref{fig_Example} depicts the concrete mixers flow in the network. We schedule $D_1$ and $D_2$ to respectively deliver 8 and 4 m$^3$ to $C_1$. We select $o^2_2$ as the first order of $C_2$ to deliver. $D_3$ travels to $B_1$ to load $q^2_2$, then visits $C_2$ before returning to his home plant. Meanwhile, after their first trip,  $D_1$ and $D_2$ deliver the remaining orders of $C_2$.

Figure~\ref{fig:ganttExample} shows the sequence of loading and unloading operations. The first deliveries of $C_1$ and $C_2$ are timely and there is no gap between the two deliveries received by $C_1$. There is a delay between the end of the first delivery and the start of the second delivery of $C_2$, however, it is less than the maximal time lag of 20 minutes defined for this example.

\begin{figure}[!ht]
    % \caption{Solution of an instance with two plants, two construction sites, four orders, and three drivers.}
    \vspace*{-10mm}

    \centering
    \begin{adjustbox}{max width=0.85\textwidth}

  \begin{tikzpicture}
    \SetUpEdge[lw = 1.2pt,
        labelcolor = white,
        % labelstyle = {draw}
    ]
    \tikzstyle{LabelStyle}=[fill=white]

    \SetGraphUnit{5}
    \GraphInit[vstyle=Normal]
    \tikzset{VertexStyle/.style = {shape = circle,text = black,minimum size = 24 pt,draw = black}}
    \Vertex[L={$B_1$}]{D1}
    \tikzset{VertexStyle/.style = {shape = circle,text = black,minimum size = 10 pt,draw = blue}}
    \tikzset{VertexStyle/.style = {shape = rectangle,text = black,minimum size = 10 pt,draw = blue}}
    \Vertex[x=2,y=5,L={11 m$^3$}]{C21}
    \EA[unit=2.0,L={3 m$^3$}](C21){C22}
    \EA[unit=2.0,L={1 m$^3$}](C22){C23}
    \tikzset{VertexStyle/.style = {shape = rectangle,text = black,minimum size = 10 pt,draw = blue}}
    \SOEA[L={$12m^3$}](D1){C1}
    \tikzset{VertexStyle/.style = {shape = circle,text = black,minimum size = 24 pt,draw = black}}
    \SOEA[L={$B_2$}](C23){D2}
    % %  \SetVertexArt
    \tikzset{VertexStyle/.style = {shape = circle,text = black,minimum size = 10 pt}}
    \SO[unit=0.8,L={$\tau^u_1=$30 $m^3/h$, 07:00 AM}](C1){tau1}
    \SO[unit=0.5,L={\textbf{$C_1$} }](tau1){o1}

    \NO[unit=0.9,L={\textbf{$C_2$}}](C22){C2}
    \SOEA[unit=0.4,L={\textcolor{white}{  } $\tau^u_2=30$ $m^3/h$, 07:25 AM}](C2){tau2}

    \tikzset{VertexStyle/.style = {shape = rectangle,text = black,minimum size = 10 pt}}
    \SOWE[unit=1.,L={$\tau^l_1=$ 70 $m^3/h$}](D1){cap1}
    \SO[unit=0.5,L={\textbf{\textcolor{brown}{D1: 8 m$^3$}}}](cap1){driver11}
    \SO[unit=0.5,L={\textbf{\textcolor{orange}{D2: 8 m$^3$}}}](driver11){driver12}
    \SO[unit=1,L={$\tau^l_2= 87$ $m^3/h$}](D2){cap2}
    \SO[unit=0.5,L={\textbf{\textcolor{blue}{D3: 12 m$^3$}}}](cap2){driver2}
    %%%%%%% Edge
    \tikzset{EdgeStyle/.style={->}}
    \Edge[](C21)(C22)
    \Edge[](C22)(C23)
    \tikzset{EdgeStyle/.style={<->,bend left=0}}
    \Edge[label=$22$](D1)(C1)
    \tikzset{EdgeStyle/.style={<->}}
    \Edge[label=$19$](D2)(C23)
    \Edge[label=$34$](D2)(C1)
    \tikzset{EdgeStyle/.style={->}}
    \tikzset{EdgeStyle/.append style = {bend left = 8}}
    \Edge[label=$17$](D1)(D2)
    \Edge[label=$18$](D2)(D1)
    \tikzset{EdgeStyle/.append style = {bend left = 65}}
    \Edge[label=$6$](D1)(C21)
    \tikzset{EdgeStyle/.append style = {bend right = 50}}
    \Edge[label=$5$](C21)(D1)
    \tikzset{EdgeStyle/.style = {->,bend left = 20,draw=brown}}
    \Edge[label=$D1$](D1)(C1)
    \Edge[label=$D1$](C1)(D1)
    \tikzset{EdgeStyle/.style = {->,bend left = 37,draw=orange}}
    \Edge[label=$D2$](C1)(D1)
    \Edge[label=$D2$](D1)(C1)

    \tikzset{EdgeStyle/.style = {->,bend left= 20,draw=brown}}
    \Edge[label=$D1$](D1)(C22)
    \tikzset{EdgeStyle/.style = {->,bend left= 0,draw=brown}}
    \Edge[label=$D1$](C22)(D1)

    \tikzset{EdgeStyle/.style = {->,bend left = 0,draw=orange}}
    \Edge[label=$D2$](C23)(D1)
    \tikzset{EdgeStyle/.style = {->,bend left = -15,draw=orange}}
    \Edge[label=$D2$](D1)(C23)

    \tikzset{EdgeStyle/.style = {->,,draw=blue}}
    \Edge[label=$D3$](D2)(D1)
    \tikzset{EdgeStyle/.style = {->,bend left = 15,draw=blue}}
    \Edge[label=$D3$](D1)(C21)
    \tikzset{EdgeStyle/.style = {->,bend left = 80,draw=blue}}
    \Edge[label=$D3$](C21)(D2)
    \end{tikzpicture}
    \end{adjustbox}
    \vspace*{-10mm}
    % }
    \caption{Solution of an instance with two plants, two construction sites, four orders, and three drivers.}
    \label{fig_Example}

\end{figure}

\begin{figure}[!htb]
    \centering

    \vspace*{-0mm}
    \begin{adjustbox}{max width=0.85\textwidth}
        \begin{tikzpicture}[-latex]
            \matrix (chart)
            [
            matrix of nodes, nodes in empty cells,
            column sep      = 0em,
            row sep         = 3ex,
            column 1/.style = {nodes={label}},    column 2/.style = {},       column 3/.style = {nodes={env}},      column 4/.style = {nodes={env}},         column 5/.style = {nodes={env}},         column 6/.style = {nodes={env}},           column 7/.style = {nodes={env}},           column 8/.style = {nodes={env}},           column 9/.style = {nodes={env}},           column 10/.style = {nodes={env}},         column 11/.style = {nodes={env}},           column 12/.style = {nodes={env}},    column 13/.style = {nodes={env}}
            ]
            {
            $C_2-D_2$ (1 m$^3$) &
            &   &   &   &   &   &   &   &   &   &   &   &            |[LC23]| L &            &       &     |[UC23]| U \\
            $C_2-D_1$ (3 m$^3$) &           &      &        &        &           &            &            &
            &
            &
            &
            |[LC21]| L &
            &
            &
            &
            |[UC21]| U &
            \\
            $C_2-D_3$ (11 m$^3$) &
            &
            &
            &
            &
            &
            |[LC22]| L &
            &
            &
            |[UC22]| &
            |[UC22]|U  &
            |[UC221]|  &
            &
            &
            &
            &
            \\
            $C_1-D_2$ (4 m$^3$) &
            &
            &
            &
            |[LC12]| L &
            &
            &
            &
            |[UC12]| U &
            &
            &
            &
            &
            &
            &
            &
            \\
            $C_1-D_1$ (8 m$^3$) &
            &
            |[LC11]| L &
            &
            &
            &
            |[UC11]|U &
            |[UC11]| &
            &
            &
            &
            &
            &
            &
            &
            &
            \\
            \vspace*{-10em}
            Start time &&
            06:21 &
            &
            06:40 &
            &
            07:00 &
            &
            07:16 &  07:25 &
            &
            07:43 &
            &
            07:51 &
            &
            08:02 & 08:08 \\
            };
            % \node[fit={(chart-3-10) (chart-3-12)},style=UC22]{U};
            \draw (chart-5-3) edge (chart-5-7);
            % \draw (chart-5-8.3) edge (chart-2-12);
            \draw  (chart-4-5) edge (chart-4-9);
            % \draw  (chart-4-9) edge (chart-1-14);
            \draw  (chart-3-7) edge (chart-3-10);
            \draw  (chart-2-12) edge (chart-2-16);
            \draw  (chart-1-14) edge (chart-1-17);

        \end{tikzpicture}
    \end{adjustbox}
    \caption{Schedule of the instance of Figure~\ref{fig_Example}.}
    \label{fig:ganttExample}
\end{figure}

\section{Constructive heuristics and GRASP}
\label{method}

We now present the solution approach we implement to solve this variant of the Concrete Delivery Problem. Our method consists in constructing feasible solutions for the CDP with randomized heuristics and iteratively calling these heuristics in the greedy randomized adaptive search procedure (GRASP) metaheuristic.

In this section, we represent the scheduling of a delivery node $d^p_{ij}$ as a pair of loading $l_{d^p_{ij}}$ and unloading $u_{d^p_{ij}}$ tasks. We then refer to a solution $S$ as a set of loading and unloading tasks performed by a set of drivers during a particular day.
\begin{alignat*}{2}
    l_{d^p_{ij}} &= \left\lbrace (b,k,q^p_{ij},LTS^{bk}_{d^p_{ij}}), b \in \mathcal{B}, k \in K \right\rbrace \\
    u_{d^p_{ij}} & =  (k,A_{{d^p_{ij}}}, UTS_{{d^p_{ij}}}), k \in K  \} \\
    S &=\left\lbrace \cup _{1 \leq p \leq n^p_i} \cup _{1 \leq j \leq |D^p_i|} (l_{d^p_{ij}},u_{d^p_{ij}}), i \in \mathcal{C} \right\rbrace
\end{alignat*}

\subsection{GRASP algorithm  }

The GRASP algorithm is an iterative suite of constructive and local search algorithms introduced by \cite{feo1995greedy}. For each iteration, a feasible solution $S$ is built using a greedy randomized algorithm. Then a local search algorithm investigates the neighborhood of  $S$ to find a local optimum. The pseudo-code of Algorithm~\ref{alg_grasp} shows that the procedure returns the best overall solution (for a minimization problem) after some stopping conditions (time limit or a maximal number of iterations) are met.

{\setstretch{1.0}
    {\small
        \begin{algorithm}[hpt]
            \caption{Pseudo-code of the GRASP algorithm }
            \label{alg_grasp}
            \DontPrintSemicolon
            \LinesNumbered
            \setcounter{AlgoLine}{0}
            \KwIn{ \textit{H}: List of constructive randomized heuristics }
            $S^* \leftarrow \emptyset$    \hspace{2mm}       $Cost(S^*) \leftarrow \infty$

            \While{Conditions not met}{
                $Cost(S') \leftarrow \infty$

                \ForEach { $h \in H$ }{

                    $S \leftarrow h(S)$

                    $S \leftarrow LocalSearch(S)$

                    \uIf{$Cost(S)< Cost(S')$}{
                        $S' \leftarrow S$

                        \uIf{$Cost(S')< Cost(S^*)$}{
                            $S^* \leftarrow S'$
                        }
                    }

                }
            }

            \KwRet{$S^{*}$}
        \end{algorithm}}
}

\subsection{Greedy randomized insertion algorithm}

As referred to in \cite{resende2019greedy}, the general outline of a greedy randomized algorithm used in the GRASP framework works as described in Algorithm~\ref{alg:greedyIns}. At each iteration, let's consider the set $CL$ of candidate delivery nodes that are not yet scheduled. We note that, for our problem, not all delivery nodes are candidates for insertion at each iteration, since deliveries of the same order are ordered, any order of a customer may be chosen, and it must be satisfied before switching to another order, and the number of visits to a customer is not known beforehand. We first initialize $CL$ with the first delivery of a random order for each customer. We create the set $CL'$ with elements of $CL$ we deem promising for a better solution. We evaluate the increase in the cost function when incorporating each $d \in CL'$ into the incumbent solution $s$ and create a restricted candidate list $RCL$ formed by nodes whose incremental costs are less than a defined threshold. From $RCL$, we randomly select the next delivery node to be incorporated into the incumbent solution and we determine its loading and unloading schedule with the procedure described in Algorithm~\ref{alg_ScheduleTask}. Next, we update $S$ and the remaining quantities of the order and customer. Then we add to $CL$ the next delivery node candidate. If an order $o^p$ is not yet fulfilled after the visiting $d^p_{ij}$, the candidate is the next delivery node $d^p_{ij+1}$. Otherwise, if the customer has remaining orders, the candidate is the first delivery node $d^{p'}_{i0}$ of another randomly selected order $p'$.

The greedy aspect of the GRASP is the creation of the $RCL$ set, the probabilistic aspect is the random selection in $RCL$, and the adaptive aspect is the update of $CL$ and the reevaluating of the incremental costs. We use the additional set $CL'$ because we noticed that restricting $CL$ to delivery nodes with the same demand, delivery due time, or overlapping unloading timeslots helps us intensify the search. Using $CL'$ also helps improve the algorithm complexity by reducing the number of nodes to evaluate. To diversify the search we simply remove the filter component. We thus obtain four greedy randomized insertion heuristics for our GRASP framework.

 {\setstretch{1}
%  {\small
     \begin{algorithm}[hbt]         
         \caption{Greedy randomized insertion algorithm }
         \label{alg:greedyIns}
         \DontPrintSemicolon
         \LinesNumbered
         \setcounter{AlgoLine}{0}
         \KwIn{  $S$: empty solution $S$}

         $CL \leftarrow \emptyset$

         \ForEach(customer i){}{
                Select a random order $o^p_i$

                $CL \leftarrow CL \cup \{d^p_{i0}\}$
        }

         \While{$CL \neq \emptyset$}{

            $CL' \leftarrow \text{Filter (CL)} $

            \ForEach( $d \in CL'$){}{

            $C(d) \leftarrow$ incremental cost of inserting $d$

            }
            $C_{min} \leftarrow min\left\lbrace C(d), \text{ } d \in CL' \right\rbrace $, $C_{max} \leftarrow max\left\lbrace C(d), \text{ } d \in CL' \right\rbrace $

            $RCL \leftarrow \left\lbrace d \in CL, \text{ } C(d) \leq C_{min} + \alpha (C_{max}-C_{min}) \right\rbrace $

            Select random delivery node $d^p_{ij}$ from $RCL$

            $l^p_{ij}$, $u^p_{ij}$ = ScheduleTasks($d^p_{ij},S$)

            $S \leftarrow S \cup (l^p_{ij}$, $u^p_{ij})  $

            $q_i \leftarrow q_i-q^p_{ij}$;  $q^p_i \leftarrow q^p_i-q^p_{ij}$

            \uIf {$q^p_i \neq 0$ }{

            $CL \leftarrow CL \cup \{d^p_{ij+1}\}$

            }
            \uElseIf{ $q_i \neq 0$}{

            Select another order $o^{p'}_i$

            $CL \leftarrow CL \cup \{d^{p'}_{i0}\} $

            }

            Update $CL$
         }
        \KwRet{$S$}
     \end{algorithm}
%  }
 }

 Algorithm~\ref{alg_ScheduleTask} takes as parameters a delivery node $d^p_{ij}$ and a partial solution and looks for a plant and a driver available for loading and unloading operations. For each driver $k$ and plant $b$ we simulate the loading and unloading operations while checking the concrete lifespan constraints, and the plant and driver availability. For each node, we first determine the start loading time $SLT$ by subtracting $t_{ij}$, $\alpha_b$, and the loading duration $LD^b_{d^p_{ij}}$ to the expected delivery time $EDT$. $EDT$ is either the arrival time $a_i$ for the first delivery of a customer or the end of the unloading of the precedent delivery. At line $15$, we ensure with the procedure \textit{ FindDriverLoadingSlot} that the loading starts only if the driver is present at the depot. The presence of the driver does not guarantee that the loading dock is available, thus we use the procedure \textit{FindDepotLoadingSlot} at line $16$ to ensure that. Once we know $SLT$, we can easily deduce the arrival time, start unloading time, driver waiting time, and client waiting time. We also keep track of the driver's work duration. The algorithm returns the best loading and unloading operations with the least cost.

 We keep track of all timeslots for a plant when its loading dock is busy, and for a driver when he starts loading until the end of the unloading service. Within \textit{FindDepotLoadingSlot} (\textit{FindDriverLoadingSlot}), we iterate through the timeslots of a depot (driver) to schedule the current operation at a free timeslot. These allow us to insert the scheduling of a node before existing schedules.

{\setstretch{1}
    {\small
        % \vspace{-4mm}
        \begin{algorithm}[htb]
            \caption{Schedule loading and unloading tasks }
            \label{alg_ScheduleTask}
            % \DontPrintSemicolon
            \LinesNumbered
            \setcounter{AlgoLine}{0}
            \KwIn{Partial solution $S$, delivery node $d^p_{ij}$}

            \SetKwFunction{FMain}{ScheduleTask}
            \SetKwProg{Fn}{Function}{}{}
            \Fn{\FMain{$d^p_{ij},S$}}
            {
            % \SetNlSkip{2em}
            % \SetAlgoNlRelativeSize{-1}
            $n_i$: last node of $i$ visited

            $\mathcal{LS}:$ Partial solutions list

            $bestSol$ Cost($bestSol)= +\infty$

            \ForEach(driver $k$){}
            {
                $n_k$: current location of $k$ (plant or delivery node)

                \ForEach(plant $b$){}
                {

                    \uIf(continue){$t_{bi} >= \delta $}{      }

                    $EDT=SUT=a_i$

                \uIf {$n_i \neq null$ }{

                    $SUT= rand( EUt_{n_i}-\gamma/3,
                        EUt_{n_i})$

                    $EDT= EUT_{n_i}$

                }

            $q^p_{ij}=min\{Q_k,q^p_i \}$ // $q^p_i$ \textit{remaining demand of order} $o^p_i$

            $SLT$ = $EDT - t_{bi} - \alpha_b - LD^b_{d^p_{ij}} $

            $SLT = FindDriverLoadingSlot(b, n_k, SLT, LD^b_{d^p_{ij}})$

            $SLT = FindDepotLoadingSlot(b, SLT, LD^b_{d^p_{ij}})$

            $AT =  SLT + LD^b_{d^p{ij}} + \alpha_b + t_{bi}$

            $SUT = max\{AT, EDT\}$

            $W^k_{d}= max\{0,SUT - AT\}$

            $W_{d}= max\{0,AT - EDT - \lambda_i\}$

            $EUT = SUT + UD_{d^p_{ij}} $

            $WT = WT_k + (EUT + \beta_k - SLT) + t_{n_kb} + t_{j,k}$

            $TC_{k,j} = t_{n_k,b} + t_{b,j}  - t_{pos_k,k}$

            \uIf{$Cost(bestSol) < Cost(S) + TC_{k,j} $}{
                $bestSol = (j,b,k,SLt) \cup (j,b,k,SUt) $

                $Cost(bestSol) = Cost(S) + TC_{k,j}$
            }
            }
            }
            \KwRet{$bestSol$}
            }
        \end{algorithm}
    }
}

\subsection*{ Local Search}

\section{ Experimental results}
\label{comp_exp}

\subsection{Data sets}

\section{Conclusion}
\label{concl}

\vspace{0.1in}

\vspace{1.5cm} \noindent \textbf{Acknowledgments}

Financial support for this work was provided by the Canadian Natural Sciences and Engineering Research Council (NSERC) under grants 2015-04893 and 2019-00094. This support is gratefully acknowledged.

\bibliographystyle{plainnat}
\bibliography{References}

\end{document}

### 2023-02-08 Presentation/Automne 2020/ciment_quebec.tex
first commit

\documentclass[9pt,xcolor={dvipsnames}]{beamer}

\usepackage{fontspec}
\setmainfont{Times New Roman}

\usepackage{appendixnumberbeamer}
\usepackage{booktabs}
\usepackage[scale=2]{ccicons}

\usepackage{pgfplots}
\usepgfplotslibrary{dateplot}
\usepackage{algorithm}
\usepackage{algorithmic}
\usepackage{natbib}
\usepackage{appendixnumberbeamer}
\usepackage[english]{babel}
\usepackage[utf8x]{inputenc}
\usepackage{lmodern} 
\usepackage{times}
\usepackage[T1]{fontenc}
\usepackage{graphicx}
\usepackage{epstopdf}					  
\usepackage[export]{adjustbox}
\usepackage{adjustbox}
\usepackage{colortbl}
\usepackage{booktabs}                      % tabelle
\usepackage{tabularx}
\usepackage{etoolbox}
\usepackage[autostyle]{csquotes}		   % per le citazioni
\usepackage{eqnarray,amsmath,amssymb,amsthm}
\usepackage{xcolor}
\usepackage{multirow}
\usepackage[export]{adjustbox}
\usepackage{pgf,tikz,pgfplots}
\usetikzlibrary{automata,topaths,calc,positioning,shapes,backgrounds, arrows}
\usepackage{mdframed}
\usepackage{fancybox}
\usepackage{hyperref}
\usepackage{xspace}
\usepackage[absolute,overlay]{textpos}
\usepackage{csquotes}
\usepackage{amsmath}
\usepackage{ulem}
\usepackage[most]{tcolorbox}
\usepackage{mathtools}
\fontfamily{lmss}\selectfont

\newcommand{\themename}{\textbf{\textsc{metropolis}}\xspace}

\setbeamercovered{dynamic}

\hypersetup{
			pdftitle={Real time optimization},
			%pdfsubject={},
			pdfauthor={ Ousmane Ali},
			pdfpagemode=FullScreen, % full screen
}

\mode<presentation>{
%\usetheme[progressbar=frametitle]{Boadilla}
\usetheme{Frankfurt}
%\useoutertheme{smoothbars}
\usecolortheme{whale}

\definecolor{vert}{rgb}{30,144,255}%{125,5,25}
%\usecolortheme[named=vert]{structure}

\definecolor{mSybilaRed}{HTML}{990000}
\usefonttheme{professionalfonts} 
\setbeamercovered{invisible}

\setbeamercolor{block title default}{fg=red, bg=Red!50}
%\setbeamercolor{frametitle}{bg=Blue!75}
%\setbeamercolor{block body default}{fg=gray,bg=Blue!50}
\setbeamercolor{postit1}{fg=white,bg=Blue}
\setbeamercolor{background canvas}{bg=white, fg=black}
\iffalse
\setbeamercolor{title separator}{
  fg=mSybilaRed
}

\setbeamercolor{progress bar}{%
  fg=mSybilaRed,
  bg=mSybilaRed!90!black!30
}

\setbeamercolor{progress bar in section page}{
  use=progress bar,
  parent=progress bar
}

\setbeamercolor{alerted text}{%
  fg=mSybilaRed
}
\fi
}
\title{Real time optimization of a FTL transportation system}
%\subtitle{Examen de Doctorat en Sciences de l'Administration}

\titlegraphic{\hfill\includegraphics[height=1.5cm]{cirrelt.jpg}}

\date{14 Mai 2019}
\author{\textbf{N. W. Ousmane~Ali, Jean-Fran\c cois C\^ot\'e, Leandro C.~Coelho}}

\institute{\large {Université Laval}  }

\setbeamertemplate{footline}
{
  \leavevmode
  \hbox{
  \begin{beamercolorbox}[wd=.15\paperwidth,ht=2.25ex,dp=1ex,center]{title in head/foot}
    \usebeamerfont{author in head/foot}{Ali et al.}%\insertshortauthor
  \end{beamercolorbox}

  \begin{beamercolorbox}[wd=.7\paperwidth,ht=2.25ex,dp=1ex,center]{author in head/foot}
    \usebeamerfont{author in head/foot}\insertshorttitle
  \end{beamercolorbox}

  \begin{beamercolorbox}[wd=.15\paperwidth,ht=2.25ex,dp=1ex,center]{title in head/foot}
    \insertframenumber{} / \inserttotalframenumber
  \end{beamercolorbox}
  }
}

\setlength{\abovecaptionskip}{1pt plus 0pt minus 4pt}

\begin{document}

\maketitle

\begin{frame}{Outline}
  \setbeamertemplate{section in toc}[sections numbered]
  \tableofcontents[hideallsubsections]
\end{frame}

\section{Introduction}

\begin{frame}{Introduction}
       \begin{block}{Delivery is a complex problem addressed under several perspectives}
		\begin{itemize}
			\item 	Time windows assignment \citep{agatz2011time}
			\item  	City logistics traffic \citep{Coelho2016}
			\item 	Loading of the truck \citep{cote2014exact}
			\item 	Routing of the vehicles  \citep{Toth2014}
			\item   Customer satisfaction
			\begin{itemize}[label=$\cdot$]

			\pause

			\item<1->  {Delivery within desired time window  \citep{desaulniers2014chapter}}
			\item<3->  {Reduction of service time \citep{PUREZA2012636},  \citep{schneider2016vehicle}}
			\item<4-5> \only<4>{Additional services \citep{HOJABRI201887}} \only<5>{\alert{Installation service}}
			\end{itemize}
		\end{itemize}
        \end{block}
\end{frame}

\begin{frame}{Introduction}
         \begin{block}{Logistic companies use the following installation strategies}
		\begin{itemize}[<+->]
			\item<1->   Installation/assembly is performed by the deliverymen
			\item<1->   Installation/assembly is performed after delivery by a dedicated crew
			\item <2-3> \only<2>{Installation/assembly can be performed by either a deliveryman or an installer}\only<3> {\alert{\textbf{Installation/assembly can be performed by either a deliveryman or an installer}}}
	 	\end{itemize}
        \end{block}
\end{frame}

\section{Problem definition}

\begin{frame}[shrink=5]{What is the  DIRPTW?}
	\only<1>{
	\begin{block}{The DIRPTW is a VRPTW variant with synchronization constraints}

		\begin{itemize}
			\item Installation by either deliverymen or installers
			\item Deliverymen and installers have different installation times
			\item Deliverymen have higher installation times
			\item Installer vehicles are faster, smaller, and cheaper than delivery vehicles
		\end{itemize}
	\end{block}}
	%\pause

	\begin{block}{Constraints}
		\begin{enumerate}
		\setlength{\parskip}{0pt} 
 		\setlength{\itemsep}{0pt plus 1pt}
			\item Each customer is served by exactly one delivery vehicle
			\item A deliveryman must serve a customer within its time window
			\item An installer must serve a customer within its time window after the delivery
			\item Each vehicle can wait at the customer until the time window opens 
			\item Each installation is either performed by a deliveryman or an installer
		\end{enumerate}
	\end{block}

\end{frame}

\begin{frame}[shrink=5]{Definition of the DIRPTW}
%\begin{center}
\begin{figure}[]
\begin{tikzpicture}
	[del/.style={circle,draw=black!50,thick, inner sep=0pt,minimum size=6mm},
	install/.style={circle,draw=black!50,fill=red!50,thick, inner sep=0pt,minimum size=4mm},
	depot/.style={rectangle,draw=black!50,fill=green!50,thick, inner sep=0pt,minimum size=6mm},
	pre/.style={<-,shorten <=1pt,thick,line width = 2pt },
	post/.style={->,shorten >=1pt,thick,line width = 2pt},
	pre1/.style={<-,shorten <=1pt,semithick},
	post1/.style={->,shorten >=1pt,semithick},
	pre2/.style={<-,shorten <=1pt,dashed,thick},
	post2/.style={->,shorten >=1pt,dashed,thick}]

	% first route
	\node [depot](Dpt) {Depot};
	\node [del] (d5) 	[below=2mm of Dpt,xshift=-20mm]{5}
		edge [pre] (Dpt);
	\node [del] (d6) 	[above left=10mm of d5]{6}
		edge [pre] (d5);
	\node [install] (i6) 	[above  =2.0mm of d6]{6}
	edge [pre] (d6);
	\node [del] (d7) 	[above right =10mm of i6]{7}
		edge [pre] (i6)
		edge [post] (Dpt);
	% Second route
	\node [del] (d1) 	[above=10mm of Dpt,xshift=20mm]{1}
		edge [pre1] (Dpt);
	\node [del] (d2) 	[below right=15mm of d1]{2}
		edge [pre1] (d1);
	\node [del] (d3) 	[below left =20mm of d2,xshift=10mm]{3}
	edge [pre1] (d2);
	\node [del] (d4) [left =10mm of d3]{4}
		edge [pre1] (d3)
		edge [post1] (Dpt);

	% Third route
	\node [install] (i1) 	[above=0.5mm of d1]{1}
		edge [pre2] (Dpt);
	\node [install] (i2) 	[above =.5mm of d2,xshift=-1mm]{2}
		edge [pre2] (i1);
	\node [install] (i4) 	[below right = 2mm of d4,xshift=-2.5mm]{4}
	edge [pre2] (i2);
	\node [install] (i5) [below right =0.5mm of d5]{5}
		edge [pre2] (i4)
		edge [post2] (Dpt);

		%Legende
	\node [install] (inode) 	[ below right=5mm of i1,xshift=25mm,label=right: Installation node ]{$n$} ;
	\node [del] (dnode) 	[below =5pt of inode,label=right: Delivery node]{$n$} ;
	\draw [post]($ (dnode) + (-0.2,-0.6) $) -- ($ (dnode) + (0.8,-0.6) $)
	node [right,align=center]
	 { Delivery route 1};
	 \draw [post1]($ (dnode) + (-.2,-1.1) $) -- ($ (dnode) + (0.8,-1.1) $)
	node [right,align=center]
	 { Delivery route 2 };
	 \draw [post2]($ (dnode) + (-.2,-1.6) $) -- ($ (dnode) + (0.8,-1.6) $)
	node [right,align=center]
	 { Installation route };
\end{tikzpicture}
\caption{Solution of a MCDIRP with seven customers}
\label{fig:fig1}
	\end{figure}
%\end{center}
\end{frame}

\section{Mathematical formulation}
\begin{frame}{}

\begin{block}{Variables}
		\begin{itemize}
	 		\item $x_{ijk}$  equal to 1 if a deliveryman travels from $i$ to $j$ using vehicle $k$, and 0 otherwise
	 		\item $y_{ijk}$ equal to 1 if an installer travels from $i$ to $j$ using vehicle $k$, and 0 otherwise 
	  		\item $z_{i}$ equal to 1 if a deliverymen installs at customer $i$ location, and 0 otherwise
	  	\end{itemize}
	\end{block}

	\begin{block}{Subsets}
		\begin{itemize}
	 		\item $D$ is the set of delivery nodes
	 		\item $I$ is the set of installation nodes
	 		\item  $D^I$ is the subset of delivery nodes that require an installation ($D^I \subseteq D$)
	  		\item $n(i)$ is the delivery node associated with the  installation node $i$
	  		\item  $l(j)$ is the installation node associated with the delivery node $j$
  		\end{itemize}
	\end{block}

\end{frame}
\begin{frame}[shrink=20]{Three-index formulation (1)}
\begin{figure}[]
 \includegraphics[scale=0.9]{equation1.PNG}
\end{figure}
\end{frame}

\begin{frame}[shrink=20]{Three-index formulation (2)}
\begin{figure}[]
 \includegraphics[scale=0.8]{equation2.PNG}
\end{figure}
\end{frame}

\iffalse
\begin{frame}[shrink=20]{Three-index formulation (1)}
\begin{align}
\label{model:obj} \min  &  \sum_{k \in K^{D}}\sum_{(i,j) \in A^D} c_{ijk}x_{ijk} + \sum_{k \in K^{I}}\sum_{(i,j) \in A^I} \hspace{-2mm}c_{ijk}y_{ijk} 
\end{align}
\begin{tcolorbox}[colback=red!5!white,colframe=red!75!black]
Usage of vehicles
\vspace{-2mm}
\end{tcolorbox}

\begin{align}
\label{c20} & \sum_{j \in \delta^{+}(0)}x_{0jk}\ \leq 1 & \hspace{-40mm}  k \in K^{D},j \in D \hspace{10mm}   \\
\label{c21} & \sum_{j \in \delta^{+}(0)}y_{0jk}\ \leq 1 & \hspace{-40mm}  k \in K^{I},j \in I \hspace{10mm} 
\end{align}
\begin{align}
\label{c22} & \sum_{k \in K^{D}}\sum_{j \in \delta^{+}(i)}x_{ijk}\ = 1 & \hspace{-40mm}   i \in D \hspace{10mm}  \\
\label{c23} & \sum_{j \in \delta^{-}(i)}x_{jik} - \sum_{j \in \delta^{+}(i)}x_{ijk} = 0 &\hspace{-40mm}   i \in N ,\ k \in K^{D} \hspace{10mm} \\
\label{c24} & \sum_{j \in \delta^{-}(i)}y_{jik} - \sum_{j \in \delta^{+}(i)}y_{ijk} = 0 & \hspace{-40mm}  i \in I ,\ k \in K^{I} \hspace{10mm}  \\
\label{c25} & \sum_{i \in D}q_{i} \sum_{j \in \delta^{+}(i)}x_{ijk}\ \leq Q & \hspace{-40mm}   k \in K^{D} \hspace{10mm}  \\
\label{c27} & \sum_{k \in K^{I}}\sum_{j \in \delta^{+}(i)}y_{ijk}\ = 1 - z_{i} & \hspace{-40mm}  i \in I \hspace{10mm}
\end{align}

\end{frame}

\begin{frame}[shrink=20]{Three-index formulation (2)}
\begin{align}
\label{c30} & S_j \geq \hspace{-1mm} S_i +  s_i^D + \sum_{k \in K^{D}} t_{ijk} x_{ijk} - M\left(1-\sum_{k \in K^{D}} x_{ijk}\right) & \hspace{-15mm}  (i,j) \in A^D, i \not\in D^I \hspace{7mm} \\
\label{c31} & S_j \hspace{-1mm} \geq \hspace{-1mm} S_i \hspace{-1mm}+ s^D_i \hspace{-1mm}+ \hspace{-3mm}\sum_{k \in K^{D}}\hspace{-1mm}(s_{ik}^I + t_{ijk} )x_{ijk} \hspace{-1mm}- \hspace{-1mm}M\hspace{-1mm}\left(\hspace{-1mm}2-\hspace{-3mm}\sum_{k \in K^{D}} \hspace{-2mm}x_{ijk}-z_{l(i)}\hspace{-1mm}\right)  &  \hspace{-4mm}   (i,j) \in A^D\hspace{-1mm}, i \in D^I \hspace{2mm}  \\
\label{c32} & S_j \geq \hspace{-1mm} S_i +  \sum_{k \in K^{I}}(s_{ik}^I + t_{ijk} )y_{ijk} - M\left(1-\sum_{k \in K^{I}} y_{ijk}\right)  & \hspace{-40mm}  (i,j) \in A^I \hspace{10mm} \\
\label{c33} & S_i \geq S_{n(i)} + s_{n(i)} & \hspace{-40mm}   i \in I \hspace{10mm}  \\
\label{c34}  & a_i \leq S_i \leq b_{i} & \hspace{-40mm}  i \in N \hspace{10mm} \\
\label{c36} & x_{ijk} \in \{0,1\} & \hspace{-50mm}  (i,j) \in A , k \in K^D  \hspace{10mm} \\
\label{c37} & y_{ijk} \in \{0,1\} & \hspace{-50mm} (i,j) \in A , k \in K^I  \hspace{10mm} \\
\label{c38} & z_{i} \in \{0,1\} & \hspace{-40mm}  i \in I. \hspace{10mm} 
\end{align}
\end{frame}
\fi

\section{ALNS algorithm}
\subsection{Operators}
\begin{frame}{ Adaptive Large Neighorhood Search}
	\begin{block}{First introduced in \citet{RopkePisinger} for the pickup and delivery routing problem}
	\begin{itemize}
			\item Initial solution
	 		\item {Destroy operators}
	 			\only<2>{
	 			\begin{itemize}
	 				\item Random delivery removal
			 		\item Related delivery removal	
	  			\end{itemize}
	  			}
	 		\item Repair operators	
	 		\only<2>{
	 			\begin{itemize}
	 				\item Greedy sequential insertion
			 		\item Regret-$k$ insertion
	  			\end{itemize}
	  			}
		\end{itemize}
	\end{block}

\end{frame}
\subsection{Preprocessing procedure}
\begin{frame}{Preprocessing procedure for insertion}
	\begin{block}{Forward time slack \citep{cordeau2004improved}}
		\begin{equation}
			\label{eq:slack1}
	 		F_i= \min_{i \leq k \leq \gamma+1}{ \sum_{i < p \leq k}{ W_{p} + (b_k-B_k)}\hspace{0.1cm}} 
	 	\end{equation}
	 	\begin{itemize}
	 		\item  $\gamma$: number of nodes in a route 
	 		\item $A_i$: arrival time at node $i$
	 		\item $B_i$: beginning service time at node $i$
	 		\item $W_{i}$: waiting time at node $i$
	 	\end{itemize}
	\end{block}

	\begin{block}{Addition of the waiting time at node $i$}
		\begin{equation}
		\label{eq:slack}
	 F_i= \min_{i \leq k \leq \gamma+1}{ \sum_{i \leq p \leq k}{ W_{p} + (b_k-B_k)} \hspace{0.1cm} }
	 \end{equation}
	 $W_{i}= \max{\lbrace 0, a_{i} - A_{i} \rbrace} $ for a delivery node $i$ \\
	  $W_{l(i)}= \max{\lbrace 0, B_{i} + s_i - A_{l(i)} \rbrace} $ for its installation node $l(i)$
	 \end{block}	
\end{frame}

\begin{frame}{Preprocessing procedure for insertion}
\begin{block}{Delivery and installation on different routes}
		\begin{itemize}[]
	 		\item compute the arrival time at all delivery nodes of the route
			\item compute the arrival time at all installation nodes of the route
			\item compute $F_{l(i)}$ using equation~(\ref{eq:slack})
	 		\item compute $F_{i}$ as:
	  	\end{itemize}
	\begin{align*}		
	%\label{eq:slackdel}
	 \hspace{0mm} F_{i} = \min \left\lbrace  \hspace{-1mm} \min_{i < k \leq \gamma+1}{ \sum_{i < p \leq k}{ \hspace{-2mm}W_{p} + (b_k-B_k )}}, \hspace{1mm} b_i-B_i - s_i,\hspace{1mm} F_{l(i)}-W_{l(i)} \right\rbrace  + W_i  
	 \end{align*}
\end{block}
\end{frame}

\begin{frame}{Illustration of the forward time slack computation}
\begin{figure}[]
 \resizebox{1.1\linewidth}{!}{
\begin{tikzpicture}
	[depot/.style={circle,draw=black!50,thick, inner sep=0pt,minimum size=6mm},
	install/.style={circle,draw=black!50,fill=black!20,thick, inner sep=0pt,minimum size=8mm},
	del/.style={circle,draw=black!50,fill=black!50,thick, inner sep=1pt,minimum size=8mm},
	st/.style={<->,shorten <=1pt,thick,line width = 1.5pt },
	pre/.style={<-,shorten <=1pt,thick,line width = 2pt },
	post/.style={->,shorten >=1pt,semithick,line width = 1.pt},
	pre1/.style={<-,shorten <=1pt,semithick},
	post1/.style={->,shorten >=1pt,semithick},
	pre2/.style={<-,shorten <=1pt,dashed,thick},
	post2/.style={->,shorten >=1pt,dashed,thick}
	,slack/.style={diamond,draw=black!50,thick, inner sep=1pt,minimum size=1mm}
	,strike/.style={strike out,draw=black!50,thick, inner sep=1pt,minimum size=1mm}, scale = 0.6
]
	% first route
	 \node (d1)[above= 0.mm,xshift= 0mm,yshift=-0mm]{ \small{D1}};
	 \node [depot](Dpt1) [right= 2 mm of d1,label={[xshift=.0cm, yshift=0.3cm][0, 20]},label={[xshift=.0cm, yshift=0.001cm] \footnotesize{0}}]{Depot};
	\node [slack](s1)[below right = 0.4mm of Dpt1,xshift=1mm,yshift=0.5mm]{2};
	%\node [strike](s1)[below left = 0.4mm of Dpt1,xshift=1mm,yshift=0.5mm]{5};
	\node [del] (d1) [right=15mm of Dpt1,label={[xshift=.0cm, yshift=0.3cm][0, 12], 3},label={[xshift=-.cm, yshift=-0.1cm] \footnotesize{2, 2}}]{1*};
	\node [slack](s2)[below right = 0.2mm of d1,xshift=1mm,yshift=1.2mm]{2};
	%\node [strike](s2)[below left = 0.2mm of d1,xshift=1mm,yshift=1.2mm]{5};
	 %\draw [st]($ (d1) + (-0.6,-0.75) $) -- ($ (d1) + (0.6,-0.75) $)
	 node [above= 0.mm,xshift= -5mm,yshift=-1.1mm]{ \scriptsize{1}};
		\path (Dpt1) edge[post]  node [ pos=0.5, sloped, above] {2} (d1);
	\node [del] (d2) [right =20mm of d1,label={[xshift=.0cm, yshift=0.3cm][9, 12], 2},label={[xshift=-.cm, yshift=-0.1cm] \footnotesize{8, 9}}]{2};
	\node [slack](s3)[below right = 0.5mm of d2,xshift=1mm,yshift=1.2mm]{4};
	 %\draw [st]($ (d2) + (-0.6,-0.75) $) -- ($ (d2) + (0.6,-0.75) $)
	 node [above= 0.mm,xshift= -5mm,yshift=-1.1mm]{ \scriptsize{1}};
	\path (d1) edge  [post] node [ pos=0.5, sloped, above] {3} (d2);
	\node [depot] (Dpt11) [ right =20mm of d2,label={[xshift=.0cm, yshift=0.3cm][0, 20]},label={[xshift=.0cm, yshift=0.001cm] \footnotesize{14}}]{Depot};
	\node [slack](s4)[below right = 0.5mm of Dpt11]{6};
	\path (d2) edge  [post] node [ pos=0.5, sloped, above] {3} (Dpt11);

	% 2nd road
	\node (d2)[below = 15 mm of d1,xshift= -33mm,yshift=-3mm]{ \small{D2}};
		\node [depot](Dpt2) [below= 15mm of Dpt1,label={[xshift=.0cm, yshift=0.3cm][0, 20]},label={[xshift=.0cm, yshift=0.001cm] \footnotesize{0}}]{Depot};
		\node [slack](s21)[below right = 0.4mm of Dpt2,xshift=1mm,yshift=0.5mm]{3};
		%\node [strike](s21)[below left = 0.4mm of Dpt2,xshift=1mm,yshift=0.5mm]{8};
	\node [del] (d3) [right=32.5mm of Dpt2,label={[xshift=.0cm, yshift=0.3cm][4, 10], 1},label={[xshift=-.cm, yshift=-0.1cm] \footnotesize{3, 4}}]{3*};
	\node [slack](s32)[below right = 0.5mm of d3,xshift=1mm,yshift=1.2mm]{3};
	%\node [strike](s32)[below left = 0.5mm of d3,xshift=1mm,yshift=1.2mm]{8};
	 %\draw [st]($ (d3) + (-0.6,-0.75) $) -- ($ (d3) + (0.6,-0.75) $)
	 node [above= 0.mm,xshift= -5mm,yshift=-1.mm]{ \scriptsize{1}};
		\path (Dpt2) edge[post]  node [ pos=0.5, sloped, above] {3} (d3);
	\node [depot] (Dpt22) [ right =32.5mm of d3,label={[xshift=.0cm, yshift=0.3cm][0, 20]},label={[xshift=.0cm, yshift=0.001cm] \footnotesize{8}}]{Depot};
	\node [slack](s322)[below right = 0.5mm of Dpt22,xshift=1mm,yshift=1.2mm]{12};
	\path (d3) edge  [post] node [ pos=0.5, sloped, above] {3} (Dpt22);

	%3rd road (installer)
		\node (I1)[below = 15 mm of d2,xshift= -0mm,yshift=-5mm]{ \small{I1}};
		\node [depot](Dpt3) [below= 15 mm of Dpt2,label={[xshift=.0cm, yshift=0.3cm][0, 20]},label={[xshift=.0cm, yshift=0.001cm] \footnotesize{0}}]{Depot};
		\node [slack](sDpt3)[below right = 0.5mm of Dpt3,xshift=1mm,yshift=1.2mm]{2};
	\node [install] (i1) [right=20mm of Dpt3,label={[xshift=.0cm, yshift=0.3cm][4, 10], 2},label={[xshift=-.0cm, yshift=-0.1cm] \footnotesize{3, 5}}]{3};
	\node [slack](si1)[below right = 0.5mm of i1,xshift=1mm,yshift=1.2mm]{4};
	 %\draw [st]($ (i1) + (-0.6,-0.75) $) -- ($ (i1) + (0.6,-0.75) $)
	 node [above= 0.mm,xshift= -5mm,yshift=-1.1mm]{ \scriptsize{2}};
		\path (Dpt3) edge[post2]  node [ pos=0.5, sloped, above] {3} (i1);

	\node [install] (i3) [right =20mm of i1,label={[xshift=.0cm, yshift=0.3cm][0, 12], 4},label={[xshift=-.0cm, yshift=-0.1cm] \footnotesize{10, 10}}]{1};
	\node [slack](si3)[below right = 0.5mm of i3,xshift=1mm,yshift=1.2mm]{2};
	% \draw [st]($ (i3) + (-0.6,-0.75) $) -- ($ (i3) + (0.6,-0.75) $)
	 node [above= 0.mm,xshift= -5mm,yshift=-1.1mm]{ \scriptsize{1}};
	\path (i1) edge  [post2] node [ pos=0.5, sloped, above] {3} (i3);
	\node [depot] (Dpt31) [ right =15mm of i3,label={[xshift=.0cm, yshift=0.3cm][0, 20]},label={[xshift=.0cm, yshift=0.001cm] \footnotesize{16}}]{Depot};
	\node [slack](sDpt31)[below right = 0.5mm of Dpt31,xshift=1mm,yshift=1.2mm]{4};
	\path (i3) edge  [post2] node [ pos=0.5, sloped, above] {2} (Dpt31);

		%Legende

	 \node [depot] (inode) [  right =15mm of Dpt11,xshift=-02mm,yshift=-5mm,label={[xshift=2mm, yshift=0.3cm]\tiny{[Time window], service time}},label={[xshift=1mm, yshift=0.001cm] \tiny{Arrival time, Beginning time}}]{\scriptsize{$Node$}} ;
	  %\draw [st]($ (inode) + (-0.5,-0.6) $) -- ($ (inode) + (0.5,-0.6) $)
	 node [above= 0.mm,xshift= -5mm,yshift=-0.5mm]{ \tiny{Service time}};
	 \node [slack](slack)[below  = 3.5mm of inode,xshift=-1mm,yshift=0.1mm,scale=0.7]{\footnotesize{slack}};
	 %\node [strike](sinode)[below  = 2mm of slack,xshift=-1mm,yshift=0.1mm,scale=0.8,label=right: \small{Wrong value of slack} $n$]{\footnotesize{Slack}};
	 \node [del] (dnode) [below =3 mm of slack,scale=0.8,label=right: \scriptsize{Delivery node }]{$n$} ;
	 \node [install] (insnode) [below =3 mm of dnode,scale=0.8,label=right: \scriptsize{Installation node}]{$n$} ;
	  \node  (n) [below right =4mm of insnode,text width=4cm,xshift= -0.8cm,scale=1]{$n^*$: \scriptsize{Node with \\ installation}} ;
	% \draw [post]($ (dnode) + (-0.2,-0.6) $) -- ($ (dnode) + (0.8,-0.6) $)
\end{tikzpicture}}
%\caption{Computation of the forward time slacks}
\label{fig:slack}
	\end{figure}
%\end{center}
\end{frame}

\section{Computational experiments}

\subsection{Instances}
\begin{frame}{Instances generation}
\begin{block}{Modification of Solomon instances \citep{Solomon1987} for the  VRPTW}
 \begin{itemize}
	 \setlength{\parskip}{0pt} 
 	\setlength{\itemsep}{0pt plus 1pt}
	 \item Customers: 15, 25, 50, 100
	 \item Speeds of each delivery/installation vehicle: 1/2
	 \item Costs of the delivery/installation vehicles: 2/1, 5/1 and 10/1
	 \item Efficiency of the deliverymen when installing: [30\%, 50\%], [60\%, 80\%]
	 \item Percentage of customers with installation: 20\%, 50\% and 75\%
 \end{itemize} 
\end{block}
\end{frame}

\subsection{Parameters tuning}

\begin{frame}{ALNS parameters tuning}
\begin{block}{Number of iterations}
	\begin{figure}[]

		\label{fig:fig2}
			\pgfplotsset{every tick label/.append style={font=\tiny},width=7.15cm,compat=1.15}
			\begin{tikzpicture}

			\begin{axis}[
				%x tick label style={
			%rotate = 90,
			%/pgf/number format/1000 sep=},
			%symbolic x coords={25000,50000,100000,200000,400000},
    			xtick=data,
			ylabel=Gap (\%),
			%enlargelimits=0.1,
			legend style={at={(0.23,1.15)},
			anchor=north,legend columns=1},
			ybar,
			%ybar = 15pt,	
			bar width=10pt,
			axis y line*=left,
  			ymin=0, ymax=2,
 			xlabel= \# of iterations,
			%enlargelimits=true
			%enlarge x limits=false, try min ticks={10} 
		]
		\addplot[black]   coordinates{
		(25000,	1.89)
		(50000,	1.6)
		(100000,1.37)
		(200000, 0.75)
		(400000	, 0.77)
		};
		\legend{Gap ( \%)}
	\end{axis}
	\begin{axis}[
  		axis y line*=right,
		  axis x line=none,
		  ymin=0	, ymax=4500,
		  ylabel= Time (s),
			legend style={at={(0.7,1.15)},
		anchor=north,legend columns=-1},
		%nodes near coords
		]

		\addplot[smooth,mark=*,]
		  coordinates{
			(25000,	253.61)
			(50000,	502.55)
			(100000, 1030.79)
			(200000, 2039.79)
			(400000	, 3440.62)
			};
		 \legend{Time (s)}

	\end{axis}

	\end{tikzpicture}
	%\caption{Effect of the number of iterations on the performance of the ALNS}
	\end{figure}
	\end{block}
\end{frame}

\iffalse
\begin{frame}{ALNS parameters tuning}
	\begin{block}{Operators available}
	\begin{itemize}
 \setlength{\parskip}{0pt} 
 \setlength{\itemsep}{0pt plus 1pt}
	\item Destroy operators: random and related removal
	\item Repair operators: greedy sequential (seq), regret-$3$, regret-$m$, regret-$3n$, regret-$mn$
\end{itemize}
	\end{block}
	\begin{block}{Operators selected within  $2^{7}$ combinations}
	\begin{itemize}
 \setlength{\parskip}{0pt} 
 \setlength{\itemsep}{0pt plus 1pt}
	\item Apply a resolution $III$ fractional factorial design on the five repair heuristics to reduce the number of combinations to $2^{3}$ from $2^5$
	\item Add one of the removal operators to each combination of insertion operators
	\item  Add to each combination of insertion operators the two available removal operators
\end{itemize}
\end{block}
\end{frame}
\fi

\begin{frame}{ALNS parameters tuning}
	\begin{block}{Operators available}
		\begin{itemize}
 \setlength{\parskip}{0pt} 
 \setlength{\itemsep}{0pt plus 1pt}
			\item Destroy operators: random and related removal
			\item Repair operators: greedy sequential (seq), regret-$3$, regret-$m$, regret-$3n$, regret-$mn$
		\end{itemize}
	\end{block}

\end{frame}

\begin{frame}<1>[label=zooms]
\frametitle{{Operators selection}}

\fontfamily{lmss}\selectfont{
	\begin{table}[]
\caption{\small{Quality of the solution when combining the operators}}
\vspace*{-2mm}
\label{tab:combo1}
\resizebox{\textwidth}{!}{
\begin{tabular}{|l|c|c|c|}
\hline
\textbf{Combination of operators}    & \textbf{Gap (\%)} & \textbf{Time  (s)} & \textbf{\% of improvement} \\ \hline
seq, regret-3, regret-$m$, regret-3$n$, regret-$$mn$$, related       & -1.20 & 643.7             & 86.9                      \\ \hline
seq, regret-3, regret-$m$, regret-$3n$, regret-$mn$, random, related & -1.08             & 648.5             & 85.2                      \\ \hline
seq, regret-$m$, regret-$mn$, random                                 & -0.98             & 690.6             & 84.3                      \\ \hline
seq, regret-3, regret-$m$, regret-$3n$, regret-$mn$, random & -0.96 & 649.5   & 83.5 \\ \hline
seq, regret-$m$, regret-$mn$, random, related                        & -1.10              & 680.8             & 82.6                  \\ \hline
seq, regret-3, regret-$3n$, related  & -0.62 & 531.1 & 82.6                      \\ \hline
seq, regret-$m$, regret-$mn$, related                                & -0.96             & 667.4             & 80.9                      \\ \hline
seq, regret-3, regret-$3n$, random, related                          & -0.59             & 535.9             & 79.1                      \\ \hline
regret-3, regret-$mn$, related                                       & -0.80              & 708.4             & 78.3                      \\ \hline
regret-3, regret-$m$, related                                        & -0.80              & 708.6             & 78.2                      \\ \hline
regret-$3n$, regret-$mn$, related                                    & -0.80              & 712.2              & 78.3                      \\ \hline
seq, random   & -0.36             & 132.3             & 78.3                      \\ \hline
seq, related         & -0.35             & 131.1             & 78.3                     \\ \hline
seq, regret-3, regret-$3n$, random                                   & -0.49             & 534.2             & 77.4                      \\ \hline
regret-$m$, regret-$3n$, related  & -0.71             & 707.8             & 75.6                      \\ \hline
regret-3, regret-$mn$, random                                        & -0.61             & 727.5             & 75.6                      \\ \hline
regret-3, regret-$m$, random  & -0.60              & 732.0                & 75.6                      \\ \hline
regret-3, regret-$m$, random, related    & -0.82             & 711.8             & 74.8                     \\ \hline
regret-$3n$, regret-$mn$, random, related                            & -0.82             & 712.6              & 74.8                      \\ \hline
regret-3, regret-$mn$, random, related     & -0.82             & 714.8             & 74.8                      \\ \hline
regret-$3n$, regret-$mn$, random   & -0.58             & 737.2             & 74.8                      \\ \hline
regret-$m$, regret-$3n$, random   & -0.87             & 728.5             & 73.9                      \\ \hline
regret-$m$, regret-$3n$, random, related                             & -0.80              & 720.2             & 73.9                      \\ \hline
seq, random, related                                                 & -0.54             & 133.4              & 73.9   \\ \hline 
\end{tabular} }
\end{table}
}
\end{frame}

\subsection{Results}

\begin{frame}{Results}
\fontfamily{lmss}\selectfont{
\begin{table}[]
\centering
\caption{Results for all instances}
\label{alns1}
\resizebox{\textwidth}{!}{%
\begin{tabular}{@{}p{0.7cm}lp{0.9cm}p{0.9cm}llclp{0.9cm}p{0.9cm}llcl@{}}
\hline
\multicolumn{1}{l}{} &  & \multicolumn{6}{c}{{[}30\%, 50\%{]}} & \multicolumn{6}{c}{{[}60\%, 80\%{]}} \\ \cmidrule(l){3-8} \cmidrule(l){9-14}
Del cost & \%Inst & Best(\%) & Avg(\%) & Del & Inst & Time(s) & Gap & Best(\%) & Avg(\%) & Del & Inst & Time(s) & Gap \\
\hline
\\[-1em]
\multirow{3}{*}{2} & 20\% & 0.0 & 0.2 & 3.3 & 0.2 & 51.3 & 16.5 & 0.0 & 0.2 & 3.2 & 0.1 & 51.7 & 15.6 \\
 & 50\% & 0.1 & 0.4 & 3.5 & 0.4 & 126.8 & 19.5 & 0.0 & 0.2 & 3.5 & 0.2 & 128.7 & 18.5 \\
 & 75\% & 0.0 & 0.4 & 3.8 & 0.6 & 190.2 & 22.6 & 0.0 & 0.4 & 3.7 & 0.3 & 190.2 & 21.3 \\[0.5em]
\multirow{3}{*}{5} & 20\% & 0.0 & 0.2 & 3.3 & 0.3 & 50.5 & 16.6 & 0.0 & 0.2 & 3.2 & 0.2 & 52.4 & 15.7 \\
 & 50\% & 0.0 & 0.3 & 3.4 & 0.6 & 128.5 & 19.3 & 0.0 & 0.3 & 3.3 & 0.4 & 129.1 & 19.0 \\
 & 75\% & 0.0 & 0.4 & 3.5 & 0.9 & 200.3 & 21.6 & 0.0 & 0.5 & 3.5 & 0.7 & 203.4 & 21.5 \\[0.5em]
\multirow{3}{*}{10} & 20\% & 0.0 & 0.2 & 3.2 & 0.4 & 51.0 & 17.1 & 0.0 & 0.2 & 3.1 & 0.2 & 52.5 & 16.0 \\
 & 50\% & 0.0 & 0.3 & 3.3 & 0.9 & 131.2 & 18.9 & 0.0 & 0.2 & 3.2 & 0.6 & 127.4 & 18.5 \\
 & 75\% & 0.0 & 0.4 & 3.4 & 1.3 & 200.7 & 21.2 & 0.0 & 0.4 & 3.3 & 0.9 & 197.2 & 21.0 \\
 \cmidrule{1-14}
\multicolumn{2}{c}{Average} & 0.0 & 0.3 & 3.4 & 0.6 & 125.6 & 19.2 & 0.0 & 0.3 & 3.3 & 0.4 & 125.8 & 18.6 \\ \bottomrule
\end{tabular}%
}
\end{table}
}
\end{frame}

\begin{frame}{Results}

\fontfamily{lmss}\selectfont{
\begin{table}[]
\centering
\caption{Results for all optimal instances}
\label{tab:my-table}
\resizebox{1.\textwidth}{!}{%
 \begin{tabular}{@{}cp{1.68cm}|cccc|cccc|cccc@{}}
	\cmidrule(r){3-14} &  & \multicolumn{4}{c|}{2}            & \multicolumn{4}{c|}{5}& \multicolumn{4}{c}{10}\\
\hline
\% of inst & \multicolumn{1}{c|}{Perf} & Opt & Best (\%) & Time & Gap & Opt & Best (\%) & Time & Gap & Opt & Best (\%) & Time & Gap \\
\hline
\multirow{2}{*}{20} & {[}30\%, 50\%{]} & 30    & 0.1 & 9    & 37  & 29    & 0.1 & 5.9  & 37.6        & 26    & 0.1 & 8.5  & 36.7        \\
  & {[}60\%, 80\%{]} & 30    & 0.1 & 5.3  & 33.7        & 30    & 0.0   & 9.6  & 34.7        & 28    & 0.0   & 8.8  & 34.6        \\
 \hline
\multirow{2}{*}{50} & {[}30\%, 50\%{]} & 27    & 0.2 & 10   & 43.8        & 24    & 0.2 & 13.9 & 41.8        & 22    & 0.1 & 13.5 & 39.2        \\
                    & {[}60\%, 80\%{]} & 25    & 0.1 & 8.1  & 39.2        & 21    & 0.2 & 7.2  & 38.8        & 23    & 0.1 & 8.4  & 39          \\
                        \hline
\multirow{2}{*}{75} & {[}30\%, 50\%{]} & 18    & 0.0   & 13.4 & 45.1        & 18    & 0.0   & 14.5 & 43.5        & 16    & 0.0   & 24.6 & 41.1        \\
                    & {[}60\%, 80\%{]} & 22    & 0.0   & 14.5 & 44.5        & 17    & 0.0   & 11.2 & 43          & 18    & 0.0   & 13.2 & 42.4        \\
                    \hline
 \multicolumn{2}{c|} {Solved}           & \multicolumn{4}{c|}{152}          & \multicolumn{4}{c|}{139}          & \multicolumn{4}{c}{133}       \\
 \bottomrule 
\end{tabular}%
}
\end{table}
}
\end{frame}

\iffalse
\begin{frame}{Flexibility of the mathematical model}
\vspace*{-5mm}

\begin{block}{ } 
	 \begin{itemize}[]
	 	\item $z_i$ = 0
	 		\begin{itemize}
				\item Installation made only by installers	
				\item Equivalent to the VRPMS of \cite{HOJABRI201887}
			\end{itemize}
		\item $z_i$ = 1
			\begin{itemize}
		 		\item Installation made by deliverymen
				\item Equivalent to the VRPTWDST of \cite{schneider2016vehicle}
			\end{itemize}
	\end{itemize}
\end{block}

\end{frame}
\fi
\begin{frame}{Results}
\vspace*{-2mm}
 \[ \sum_{k \in K^{I}}\sum_{j \in \delta^{+}(i)}y_{ijk}\ = \only<1>{1 - z_{i}} \only<2>{0} \only<3>{1} \only<4>{1 - z_{i}}  \hspace{10mm} i \in I \hspace{3mm} \only<2>{\text{\alert{Deliverymen make all installations}}} \only<3>{\text{\alert{Deliverymen never install}}} \only<4>{\text{\alert{Both can install}}}\]

\fontfamily{lmss}\selectfont{

 Comparison when deliverymen efficiency are within {[}30\%, 50\%{]}

	\begin{table}[!htbp]
\centering
\vspace*{-3mm}
\label{bus_set1}
\resizebox{\textwidth}{!}{%
\begin{tabular}{@{}p{0.5cm}p{0.5cm}lllp{0.5cm}lllp{0.5cm}lll@{}}
\toprule
 &  \multicolumn{4}{c}{20\%} & \multicolumn{4}{c}{50\%} & \multicolumn{4}{c}{75\%} \\ \cmidrule(l){2-5} \cmidrule(l){6-9} \cmidrule(l){10-13}
Del cost &  Opt & $z_i=1$ & $z_i=0$ & \small DIRPTW & Opt& $z_i=1$ & $z_i=0$ & \small DIRPTW &  Opt& $z_i=1$ & $z_i=0$ & \small DIRPTW \\
\hline
\\[-0.8em]
  \multicolumn{1}{l}{2} &28 &909.6 & 908.6 & \textbf{864.2} &26 &722.6 & 803.3 & \textbf{696.1} &17& 896.1 & 1106.9 & \textbf{855.2} \\[0.3em]

 \multicolumn{1}{l}{5} &29& 2247.8 & 2014.7 & \textbf{1972.4} & 24&2481.6 & 2107.0 & \textbf{2020.0} &28& 2107.9 & 1860.2 & \textbf{1709.8} \\[0.3em]

 \multicolumn{1}{l}{10} &26& 4507.6 & 3875.3 & \textbf{3838.4}& 22& 5008.6 & 3985.3 & \textbf{3899.5} &16& 5887.6 & 4174.9 & \textbf{4051.2} \\ \bottomrule
\end{tabular}%
}
\end{table}
}
Comparison when deliverymen efficiency are within {[}60\%, 80\%{]}
\fontfamily{lmss}\selectfont{
	\begin{table}[!htbp]
\centering
\vspace*{-3mm}
\label{bus_set2}
\resizebox{\textwidth}{!}{%
\begin{tabular}{@{}p{0.5cm}p{0.3cm}lllp{0.3cm}lllp{0.3cm}lll@{}}
\toprule
  &  \multicolumn{4}{c}{20\%} & \multicolumn{4}{c}{50\%} & \multicolumn{4}{c}{75\%} \\ \cmidrule(l){2-5} \cmidrule(l){6-9} \cmidrule(l){10-13}
Del cost &  Opt & $z_i=1$ & $z_i=0$ & \small DIRPTW & Opt& $z_i=1$ & $z_i=0$ & \small DIRPTW &  Opt& $z_i=1$ & $z_i=0$ & \small DIRPTW \\
\hline
\\[-0.8em]
\multicolumn{1}{l}{2} &27& {635.0} & 719.9 & \textbf{633.4} &24& \textbf{619.6} & 794.6 & 640.3 &21 &\textbf{913.2} & 1268.8 & 920.7 \\[0.3em]
\multicolumn{1}{l}{5} &30 &2256.3 & 2275.2 & \textbf{2182.8} &21& 1589.9 & 1724.3 & \textbf{1574.6} & 17&1446.8 & 1630.2 & \textbf{1424.9} \\[0.3em]
\multicolumn{1}{l}{10} &28& 4066.8 & 3868.1 & \textbf{3795.7} &23& 3052.3 & 3112.8 & \textbf{2972.6} &18&3429.5 & 3299.0 & \textbf{3182.9} \\ \bottomrule
\end{tabular}%
}
\end{table}
}

\end{frame}

\begin{frame}{Managerial insights}
\begin{block}{}
	\begin{itemize}
		\item The proposed strategy is better than the traditional strategies 
		\item  No need for installation fleet in the following cases: 
			\begin{itemize}
				\item A small number of customers with very few having an installation regardless of the cost or the performance of the deliveryman 
				\item Low deliveryman cost, higher efficiency, medium size of customers
			\end{itemize}
		\item Use a distinct installation fleet is better when:
			\begin{itemize}
				\item[•] Deliverymen efficiency are low and the cost range from medium to high
				\item Deliverymen efficiency and cost are both high
			\end{itemize}
	\end{itemize}
\end{block}{}
\end{frame}

\begin{frame}[allowframebreaks]{Performance on the VRPMS instances set}
	\begin{block}{ \cite{HOJABRI201887} studied the delivery and installation of large items}
		\begin{itemize}
			\item Two distinct fleets are used
			\item A deliveryman and its installer arrival must be synchronized at a certain number of customer locations
			\item The installer must visit the customer within \textbf{ten} units of time after the start of the delivery service
		\end{itemize}
	\end{block}

\end{frame}

\begin{frame}{Results on the VRPMS instances set}
\fontfamily{lmss}\selectfont{
\textbf{Performance on the VRPMS instances}
\begin{table}[]
	%\caption{\small Performance on the VRPMS instances}
	\label{vrpms}
\resizebox{0.95\textwidth}{!}{
	\begin{tabular}{@{}cccccccccc}
	\toprule
	\multicolumn{1}{c}{} & \multicolumn{4}{c}{\cite{HOJABRI201887}}                                                    & \multicolumn{5}{c}{ALNS}  \\
 \cmidrule(l){2-5}   \cmidrule(l){6-10}
& \multirow{1}{*}{Best(\%)} & \multirow{1}{*}{Del}     & \multirow{1}{*}{Inst}    & \multirow{1}{*}{   Time (s)}       & Best(\%) & Avg (\%) & \multirow{1}{*}{Del} & \multirow{1}{*}{Inst} & \multirow{1}{*}{Time (s)} \\
\hline
\\[-1em] 
 \multicolumn{1}{r}{Average}     & \multicolumn{1}{c}{6.9}  & \multicolumn{1}{c}{8.4}  & \multicolumn{1}{c}{3.7} & \multicolumn{1}{c}{63681.0}   & \multicolumn{1}{c}{0.0}  & \multicolumn{1}{c}{0.3} & \multicolumn{1}{c}{7.6} & \multicolumn{1}{c}{3.4} & \multicolumn{1}{c}{170.7} \\
                  \hline
\end{tabular}}
\end{table}

\pause
%\begin{frame}{Results on the VRPMS instances set}

\textbf{Results for instances solved optimally with CPLEX}
\begin{table}[H]
%\caption{\small Results for instances solved optimally on VRPMS instances set}
\vspace*{-1mm}
\label{vrpmsExact}
\resizebox{1\textwidth}{!}{
\begin{tabular}{m{0.4cm}|cccccccccccccc}
\toprule
\multicolumn{1}{l}{}   & \multicolumn{4}{c}{\cite{HOJABRI201887}} & \multicolumn{5}{c}{ALNS}                                                                                                                       & \multicolumn{5}{c}{Exact method}                                                                                               \\ \cmidrule(r){2-5}\cmidrule(r){6-10}\cmidrule(r){11-15}
%\small
\multirow{1}{*}{} & \multirow{1}{*}{\small {Best (\%)}} & \multirow{1}{*}{Del}     & \multirow{1}{*}{Inst}    & \multirow{1}{*}{Time (s)}       & \multicolumn{1}{c}{Best (\%)}   & Avg (\%) & \multirow{1}{*}{Del} & \multirow{1}{*}{Inst} & \multicolumn{1}{c}{\multirow{1}{*}{Time (s)}} & \multicolumn{1}{c}{\multirow{1}{*}{Opt}} & \multirow{1}{*}{\small{Gap(\%) }} & \multirow{1}{*}{Del} & \multirow{1}{*}{Inst} & \multirow{1}{*}{Time (s)}   \\ \cline{1-15}

\\[-0.8em]            
\multicolumn{1}{r}{Average}  & 3.0  & 9.1  & 4.3   & 33319.8  & 0.2    & 0.3  & 8.5 & 4.1 & 72.7 & 20.3 & 34.2  & 8.5   & 4.1   & 505.7   \\ \hline         
\multicolumn{1}{r|}{Solved}  &\multicolumn{4}{c}{44} &\multicolumn{5}{c}{122}  & \multicolumn{5}{c}{139}  \\ \hline         
\end{tabular}}
\end{table}
}
\end{frame}

\begin{frame}[allowframebreaks]{Performance on the VRPTWDST instances set}
\vspace*{10mm}
	\begin{block}{ \cite{schneider2016vehicle} }
		\begin{itemize}
			\item Drivers have different efficiency with their customers
			\item Drivers have specific travel and service times in order to model their familiarity with the customers to visit
			\item No synchronization constraints
		\end{itemize}
	\end{block}

%\end{frame}
%\begin{frame}[allowframebreaks]{Results on the VRPTWDST instances set}
\newpage
\vspace*{-5mm}
\fontfamily{lmss}\selectfont{

\vspace*{-5mm}

\begin{table}[]
\centering
\caption{Lower and upper bounds for the VRPTWDST for the secondary objective of minimizing travelled distance}
\label{tab:random}
\resizebox{0.8\textwidth}{!}{%
\begin{tabular}{lcllcllll}
\toprule
  & \multicolumn{3}{c}{ \cite{schneider2016vehicle}}
  & \multicolumn{3}{c}{ALNS} & \multicolumn{2}{c}{Exact} \\
 \cmidrule(l){2-4} \cmidrule(l){5-7} \cmidrule(l){8-9}
  & Best (\%) & Opt & Time & Best (\%)& Opt & Time &  Gap(\%)& Opt \\
\hline
  C1 & 0.0 & 28/90 & 6.9 & 0.0 & 28/90 & 33.7 & 22.0 & 28/90 \\
 C2 & 0.0 & 24/80 & 10.5 & 0.0 & 24/80 & 46.8 & 4.1 & 24/80 \\
 R1 & 2.4 & 0/120 & 94.7 & 0.2 & 4/120 & 37.6 & 28.7 & 8/120 \\
 R2 & 6.6 & 0/110 & 45.3 & 0.1 & 0/110 & 73.2 & 19.6 & 0/110 \\
 RC1 & 3.8 & 0/80 & 53.1 & 0.3 & 0/80 & 32.6 & 40.2 & 0/80 \\
 RC2 & 10.7 & 0/80 & 67.4 & 0.1 & 0/80 & 54.1 & 28.6 & 0/80 \\[0.5em]
\hline
\multicolumn{1}{l}{Average} & 3.9 & - & 52.8 & 0.1 & - & 46.9 & 24.0 & - \\
 \hline
\multicolumn{1}{l}{Solved} & \multicolumn{3}{c}{52/560} &\multicolumn{3}{c}{56/560} &\multicolumn{2}{c}{60/560} \\
\hline
\end{tabular}%
}
\end{table}}

\end{frame}

\section{Conclusion}
\begin{frame}{Conclusion}

	\begin{itemize}
		\item Novel distribution strategy proposed
		\item Insights about existing distribution and installation strategies
		\item An exact and heuristic algorithm to solve the DIRPTW
		\item New best-know and optimal solutions for the VRPMS and the VRPTWDST
	\end{itemize}
	\begin{block}{Applications of the DIRPTW}
		Choose between a specialist and a generalist in the home care staff scheduling problem
	\end{block}
\end{frame}

\appendix

\begin{frame}[allowframebreaks]{References}

  \bibliography{biblio}
  \bibliographystyle{plainnat}

\end{frame}

\end{document}

### 2023-02-08 Presentation/Automne 2020/presentation_CQ.tex
first commit

\documentclass[9pt,xcolor={dvipsnames}]{beamer}
\usepackage{algorithm}
\usepackage{algorithmic}
\usepackage{natbib}
\usepackage{tikz}
\usepackage[overload]{empheq}
\usepackage{empheq}
\usepackage{array, makecell}
\usepackage{tabularx}
\usepackage{caption}
\usepackage{rotating}
\usepackage{natbib}
\mode<presentation>{
	\usetheme{Warsaw}			%  http://www.hartwork.org/beamer-theme-matrix/
	\usecolortheme[named=Blue]{structure}
	\useoutertheme{shadow}
	\usefonttheme{professionalfonts}

	%\useoutertheme{sidebar}
	%\setbeamercolor*{sidebar}{fg=RoyalBlue,bg=!75!white}
	\setbeamercovered{transparent}
	\setbeamercolor{block title example}{fg=white,bg=Blue}
	\setbeamercolor{block body example}{fg=gray,bg=Blue!10}
	\setbeamercolor{postit1}{fg=white,bg=Blue}
	\setbeamercolor{postit2}{fg=gray,bg=Blue}
	\setbeamertemplate{itemize items}[default]
	\setbeamertemplate{enumerate items}[default]
	\setbeamertemplate{sections/subsections in toc}[sections numbered]
}

\makeatletter
\long\def\beamer@@ssection*#1{\beamer@section[]{}}
\makeatother
\newcommand{\RN}[1]{%
	\textup{\uppercase\expandafter{\romannumeral#1}}%
}
\usepackage{pgfpages}		% http://mathoverflow.net/questions/5893/beamer-printout
\mode<handout>{
	\usetheme{default}
	\setbeamercolor{background canvas}{bg=Gray!5}
	\pgfpagesuselayout{4 on 1}[a4paper,portrait,border shrink=2.5mm]		% 4 slide per pagina
}

%\usetheme{Boadila}
\usetheme{Warsaw}			%  http://www.hartwork.org/beamer-theme-matrix/
%\usecolortheme[named=RoyalBlue]{structure}
%\useoutertheme{shadow}
%\usefonttheme{professionalfonts} 			
%
%\usepackage{pgfpages}		% http://mathoverflow.net/questions/5893/beamer-printout
%\mode<handout>{
%  \usetheme{default}
%  \setbeamercolor{background canvas}{bg=Black!5}
%  \pgfpagesuselayout{4 on 1}[a4paper,portrait,border shrink=2.5mm]		% 4 slide per pagina
%}

\usepackage[english]{babel}
\usepackage[utf8x]{inputenc}
\usepackage{lmodern} 
\usepackage{times}
\usepackage[T1]{fontenc}
\usepackage{graphicx}
\usepackage{epstopdf}					   % immagini formato .eps
\graphicspath{{images/}}
\usepackage[export]{adjustbox}
\usepackage{adjustbox}
\usepackage{colortbl}
\usepackage{booktabs}                      % tabelle
\usepackage{tabularx}
\usepackage{etoolbox}
\usepackage[autostyle]{csquotes}		   % per le citazioni
\usepackage{eqnarray,amsmath,amssymb,amsthm}
\usepackage{xcolor}
\usepackage{multirow}
\usepackage[export]{adjustbox}
\usepackage{amsmath}
\usepackage{tikz}
\usetikzlibrary{shapes,arrows}
\tikzstyle{square}=[draw]
\usepackage{pgfplots}
\pgfplotsset{compat = newest}
\newcolumntype{P}[1]{>{\centering\arraybackslash}p{#1}}
\newcolumntype{M}[1]{>{\centering\arraybackslash}m{#1}}
\newcolumntype{C}[1]{>{\centering\arraybackslash}c{#1}}
\setbeamercovered{dynamic}

\hypersetup{
			pdftitle={Real-time optimization},
			%pdfsubject={},
			%pdfauthor={Ousmane Ali},
			pdfpagemode=FullScreen, % full screen
}

\usepackage[absolute,overlay]{textpos}
\setlength{\TPHorizModule}{1mm}
\setlength{\TPVertModule}{1mm}

%\begin{left}
  %\includegraphics[scale=0.1]{image/LavalLogo.jpg}\\[\bigskipamount]
%\end{left}
\begin{document}

\begin{center}
\vspace*{0.1cm}
 \includegraphics[scale=0.15]{cirrelt.png}\\[\bigskipamount]
\end{center}

%\vspace{-1.2pt}
\vspace*{-2.2cm}
\title[Concrete transport optimization]{Real-time vehicle scheduling of a FTL transportation system}
\author[Ousmane Ali]{\textbf{Ousmane Ali\\ Jean-Fran\c{c}ois C\^ot{\'e} \\ Leandro C. Coelho}}

\date{}

\vspace{-0.5cm}
%\logo{\vspace{0.85cm}\includegraphics[height=0.8cm]{images/logo-universite-laval}} % logo in ogni pagina
%\titlegraphic{\vspace{0.85cm}\includegraphics[height=0.8cm, angle=0, right]{images/cirrelt}} % immagine nella title page

\begin{frame}[noframenumbering]{}

\begin{center}
\vspace{-0.43651pt}
%\includegraphics[height=1.4cm, angle=0]{images/logo-polimi}
\titlepage

{\vspace*{-3cm}
	\texorpdfstring{
		\vspace{0.5cm}
		Université Laval \\
		\vspace{0.2cm}
		Département d'Opérations et Systèmes de Décision \\
		\centering
		\vspace*{0.3cm}
		December 7th, 2020
		\vspace*{1.5cm}
	}
	{}
}
\vspace{-3cm}
\end{center}

\end{frame}
\logo{}

\setbeamertemplate{footline}{%		% numerazione pagine
\vskip-4pt%
  \leavevmode%
  \hbox{\begin{beamercolorbox}[wd=.5\paperwidth,ht=2.5ex,dp=1.125ex,leftskip=.2cm plus1fill,rightskip=.2cm]{author in head/foot}%
    \usebeamerfont{author in head/foot}\insertshortauthor
  \end{beamercolorbox}%
  \begin{beamercolorbox}[wd=.5\paperwidth,ht=2.5ex,dp=1.125ex,leftskip=.3cm,rightskip=.3cm plus1fil]{title in head/foot}%
    \usebeamerfont{title in head/foot}\insertshorttitle
    ~~~\raggedleft\insertframenumber/\inserttotalframenumber
  \end{beamercolorbox}}%
  \vskip0pt%
}

\begin{frame}{Outline}
	\tableofcontents[hidesubsections]
\end{frame}

\section{Introduction}

\subsection{Context of the problem}
\begin{frame}{Cement  industry in Canada}
\setbeamercolor{block title}{use=structure,fg=white,bg=red!75!black}
\begin{block}{ }
\begin{itemize}
% \item 14 factories.
\item 13 million of tonnes of cement. 
\item 1.6 billion \$ of production. \footnote{\label{1} 2014}
\item Wide variety of products:   cement, concrete (UNISOLANT, UNIGEL, UNIPLAN, AGRIMIX, UNIFLOW)

\end{itemize}
\vspace{0.1cm}
\begin{center}
	\includegraphics[scale=0.2]{ciment_quebec.jpg}
	\includegraphics[scale=0.2]{Lafarge-logo.jpg}
	\includegraphics[scale=0.2]{st-marys.jpg}
	\includegraphics[scale=0.2]{mc-innis.png}
\end{center}
\end{block}

\end{frame}

\begin{frame}{Issues for concrete deliveries}
\setbeamercolor{block title}{use=structure,fg=white,bg=red!75!black}
\begin{block}{ }
\includegraphics[scale=0.5]{Unibeton.jpg}
\begin{itemize}
\item Hundred thousands deliveries per year.
\item Specialized workforce to drive the concrete-mixer.
	\includegraphics[scale=0.5]{betonniere.jpg}
\vspace{0.1cm}
\item Highly dynamic, perishable and seasonal demand.
\item Restrictions on deliverymen weekly work time.
\item High operating costs.
\item Weather dependent activities.
\end{itemize}
\end{block}
\end{frame}

\section{Problem description}

\begin{frame}{Flowchart}

\includegraphics[scale=0.7]{flowchart.png}

\end{frame}

\begin{frame}{Problem description}

\begin{block}{Constraints (1)}
\begin{itemize}
\item Loading operation
	\begin{itemize}
		\item One vehicle at a time
		\item Loading time dependent of the type of concrete
		\item Additional cooling operation for some product.
	\end{itemize}
\item Transit time dependent of the road traffic.

\item Delivery at due time (synchronization with other services at the customer location).

\item Demand of a customer may be delivered at different periods in the same day per the customer requirements.

\item  Concrete must be delivered at most 3 hours after loading.

\item Unload concrete-mixer one at a time.

\item Deliverymen must work at least 40 hours weekly.

\item Deliverymen may have additional constraints related to the maximal daily working time.

\end{itemize}
\end{block}

\end{frame}

\begin{frame}{Problem description}

\begin{block}{Constraints (2)}
\begin{itemize}

\item Concrete-mixer with different sizes.

\item Demands are known two or three days before the delivery, therefore the planning of the deliverymen is highly dynamic

\end{itemize}
\end{block}

\setbeamercolor{block title}{use=structure,fg=white,bg=red!75!black}
\begin{block}{Objectives}
\begin{itemize}
\item Minimize the fleet utilization.
\item Plan deliveries of each day according to the available deliverymen and concrete-mixers.
\item Plan the deliverymen weekly schedule.

\end{itemize}
\end{block}

\end{frame}

\begin{frame}{Literature review}

\begin{block}{Relevant problems}
\begin{itemize}
\item The single depot vehicle scheduling problem with length of path restrictions. \citep{raff1983routing}
\item The single depot vehicle scheduling problem with multiple vehicle types. \citep{raff1983routing}
\item The Tractor-trailer routing and scheduling with full load. \citep{raff1983routing}
\item Real-time dispatching problem \citep{brown1981real}
\item Dynamic vehicle routing problem \citep{liao2004tabu}
\end{itemize}
\end{block}

\end{frame}

\section{Solution approaches}

\subsection{Simulation}
\begin{frame}{Simulation of the current system}

\begin{block}{}
\begin{itemize}
\item Model the current system
\item Simulate the system with a discrete-event simulation software (SIMIO)
\end{itemize}
\end{block}
\end{frame}

\subsection{Integer programming formulation}

\begin{frame}{Mathematical formulation}

\begin{block}{}
\begin{itemize}
\item Propose a mathematical model for the problem
\item Solve this model with small to medium instances. 
\item Compare the solution obtained with the current state of the system.
\item Adjust the model if required.
\end{itemize}
\end{block}
\end{frame}

\subsection{Heuristic solution approach}

\begin{frame}{Heuristic algorithm}

\begin{block}{}
\begin{itemize}
\item Design heuristic (metaheuristic) algorithm(s)
\item Compare the solution obtained with the current state of the system.
\item Adjust the algorithm(s) if required.
\item Assess the algorithm(s) performance.
\end{itemize}
\end{block}
\end{frame}

\section{Expected results}
\begin{frame}{Expected results}

\setbeamercolor{block title}{use=structure,fg=white,bg=red!75!black}
\begin{block}{Reduction of the concrete-mixers utilization}
\textbf{Objective}: Reduce the number of yearly delivery trips. %by 1\%
\end{block}

\setbeamercolor{block title}{use=structure,fg=white,bg=red!75!black}
\begin{block}{Dynamic dispatching of the deliverymen}
\textbf{Objective}:  Implement a software to dispatch in real-time deliveries to workers.
\end{block}

\begin{block}{Contribution to the OR literature}
\textbf{Objective}: Develop quick and efficient algorithms for real-time vehicle scheduling problems.
\end{block}
\end{frame}

\section{Conclusion}
\begin{frame}{Highlights}

\begin{itemize}
\item Describe a complex problem arising in the concrete transportation
\vspace{0.2cm}
\item Present some solution approaches.
\vspace{0.2cm}
\item Present expected results
\vspace{0.2cm}
\item This problem can be applied to other truck load transportation system such as transportation of vehicles from plants to dealers. 
\end{itemize}

\end{frame}

\section*{Bibliographie}
\vspace{0.2cm}
%\bibliographystyle{elsarticle-harv}
\bibliographystyle{plainnat}
\bibliography{biblio}

\end{document}

### 2023-02-08 Data/instances/script/data.py

@@ -1,0 +1,29 @@
+
+
+class Node:
+    def __init__(self):
+        self.id = -1
+        self.nbr = -1
+        self.coord = (0, 0)
+        self.adresse = ""
+        self.code_postal = ""
+        self.demand = 0
+        self.ltw = 0
+        self.utw = 0
+        self.type = -1
+
+
+class Driver:
+    def __init__(self):
+        self.id = -1
+        self.rank = -1
+        self.capacity = 0
+        self.driver_type = -1
+        self.truck_capacity = 0
+        self.factory_nbr = -1
+        self.location_id = -1
+        self.ltw = 0
+        self.utw = 0
+
+    def __str__(self):
+        return f"Driver {self.id} {self.rank}"

### 2023-02-08 Data/instances/script/driver.py

@@ -1,0 +1,18 @@
+"""Class driver."""
+class Driver:
+    def __init__(self):
+        self.id = -1
+        self.rank = -1
+        self.capacity = 0
+        self.driver_type = -1
+        self.truck_capacity = 0
+        self.factory_nbr = -1
+        self.depot_nbr = -1
+        self.location_id = -1
+        self.ltw = 0
+        self.utw = 0
+        self.nbr = -1
+        self.description = ""
+
+    def __str__(self):
+        return f"Driver Nbr {self.nbr} Rank {self.rank} {self.description} depot n {self.depot_nbr} fac {self.factory_nbr} de capacité {self.capacity}\n"

### 2023-02-08 Data/instances/script/drivers_schedule.py

@@ -1,0 +1,148 @@
+
+#%%
+import pandas as pd
+import os
+import datetime
+from collections import Counter
+#%%
+from driver import Driver
+
+#%%
+TICKET_FOLDER = "/".join([os.path.dirname(__file__), "tickets/"])
+INSTANCE_FOLDER = "/".join([os.path.dirname(__file__), "instances/"])
+#%%
+""" Load dataframes """
+df_tickets = pd.read_excel(f"{os.path.dirname(__file__)}/DTICKETHISTAB_BETON.xlsx")
+#%%
+df_orders = pd.read_excel(f"{os.path.dirname(__file__)}/DORDERSTAB_BETON.xlsx",sheet_name='data')
+#%%
+df_employee = pd.read_excel(f"{os.path.dirname(__file__)}/EMPLOYETAB.xlsx")
+df_employee.drop(['COMP_NBR', 'emp_sch_specificite'], axis=1, inplace=True)
+df_employee.set_index('EMPL_NBR',inplace=True)
+#%%
+df_depots = pd.read_excel(
+    f"{os.path.dirname(__file__)}/adresses_complete.xlsx")
+df_depots = df_depots[df_depots['Location_type'] != 'customer']
+df_depots.dropna(axis=1,  inplace=True)
+df_depots['NBR'] = df_depots['JOB_DESC'].apply(
+    lambda x: x if len(x.split(' ')) == 1 else int(x.split(' ')[1]))
+df_depots.set_index('NBR',inplace=True)
+#%%
+# Depot location
+df_depots_loc = pd.read_excel(
+    f"{os.path.dirname(__file__)}/LOCATIONTAB.xlsx")
+df_depots_loc['LOC_NBR'] = df_depots_loc['LOC_NBR'].apply(lambda x: int(x))
+df_depots_loc.set_index('LOC_NBR',inplace=True)
+#%%
+# Les tickets contiennent les séquences des opérations de distribution
+# Un ticket est la livraison de béton sur un chantier par un livreur
+
+# Je dois déterminer la séquence de livraison par jour.
+df_tickets['DATE'] = df_tickets['TICKET_DATE'].apply(lambda x:datetime.date(x.year,x.month,x.day))
+df_tickets['WEEK'] = df_tickets['DATE'].apply(lambda x:datetime.date(x.year,x.month,x.day).isocalendar()[1])
[…]

### 2023-02-08 Data/instances/script/drivers_schedule2.py

@@ -1,0 +1,160 @@
+
+#%%
+import pandas as pd
+import os
+import datetime
+from collections import Counter
+#%%
+from driver import Driver
+
+#%%
+TICKET_FOLDER = "/".join([os.path.dirname(__file__), "tickets/"])
+INSTANCE_FOLDER = "/".join([os.path.dirname(__file__), "instance2/"])
+#%%
+""" Load dataframes """
+df_tickets = pd.read_excel(f"{os.path.dirname(__file__)}/DTICKETHISTAB_BETON.xlsx")
+#%%
+df_orders = pd.read_excel(f"{os.path.dirname(__file__)}/DORDERSTAB_BETON.xlsx",sheet_name='data')
+#%%
+df_employee = pd.read_excel(f"{os.path.dirname(__file__)}/EMPLOYETAB.xlsx")
+df_employee.drop(['COMP_NBR', 'emp_sch_specificite'], axis=1, inplace=True)
+df_employee.set_index('EMPL_NBR',inplace=True)
+#%%
+df_depots = pd.read_excel(
+    f"{os.path.dirname(__file__)}/adresses_complete.xlsx")
+df_depots = df_depots[df_depots['Location_type'] != 'customer']
+df_depots.dropna(axis=1,  inplace=True)
+df_depots['NBR'] = df_depots['JOB_DESC'].apply(
+    lambda x: x if len(x.split(' ')) == 1 else int(x.split(' ')[1]))
+df_depots.set_index('NBR',inplace=True)
+#%%
+# Depot location
+df_depots_loc = pd.read_excel(
+    f"{os.path.dirname(__file__)}/LOCATIONTAB.xlsx")
+df_depots_loc['LOC_NBR'] = df_depots_loc['LOC_NBR'].apply(lambda x: int(x))
+df_depots_loc.set_index('LOC_NBR',inplace=True)
+#%%
+# Les tickets contiennent les séquences des opérations de distribution
+# Un ticket est la livraison de béton sur un chantier par un livreur
+
+# Je dois déterminer la séquence de livraison par jour.
+df_tickets['DATE'] = df_tickets['TICKET_DATE'].apply(lambda x:datetime.date(x.year,x.month,x.day))
+df_tickets['WEEK'] = df_tickets['DATE'].apply(lambda x:datetime.date(x.year,x.month,x.day).isocalendar()[1])
+df_tickets['DAY'] = df_tickets['DATE'].apply(lambda x:datetime.date(x.year,x.month,x.day).toordinal())
[…]

### 2023-02-08 Data/instances/script/instance2.py

@@ -1,0 +1,231 @@
+
+#%%
+import pandas as pd
+import os
+import datetime
+from collections import Counter
+#%%
+from driver import Driver
+
+#%%
+TICKET_FOLDER = "/".join([os.path.dirname(__file__), "tickets/"])
+INSTANCE_FOLDER = "/".join([os.path.dirname(__file__), "instance2/"])
+#%%
+""" Load dataframes """
+df_tickets = pd.read_excel(f"{os.path.dirname(__file__)}/DTICKETHISTAB_BETON.xlsx")
+#%%
+df_orders = pd.read_excel(f"{os.path.dirname(__file__)}/DORDERSTAB_BETON.xlsx",sheet_name='data')
+#%%
+df_employee = pd.read_excel(f"{os.path.dirname(__file__)}/EMPLOYETAB.xlsx")
+df_employee.drop(['COMP_NBR', 'emp_sch_specificite'], axis=1, inplace=True)
+df_employee.set_index('EMPL_NBR',inplace=True)
+#%%
+df_depots = pd.read_excel(
+    f"{os.path.dirname(__file__)}/adresses_complete.xlsx")
+df_depots = df_depots[df_depots['Location_type'] != 'customer']
+df_depots.dropna(axis=1,  inplace=True)
+df_depots['NBR'] = df_depots['JOB_DESC'].apply(
+    lambda x: x if len(x.split(' ')) == 1 else int(x.split(' ')[1]))
+df_depots.set_index('NBR',inplace=True)
+#%%
+# Depot location
+df_depots_loc = pd.read_excel(
+    f"{os.path.dirname(__file__)}/LOCATIONTAB.xlsx")
+df_depots_loc['LOC_NBR'] = df_depots_loc['LOC_NBR'].apply(lambda x: int(x))
+df_depots_loc.set_index('LOC_NBR',inplace=True)
+#%%
+# Les tickets contiennent les séquences des opérations de distribution
+# Un ticket est la livraison de béton sur un chantier par un livreur
+
+# Je dois déterminer la séquence de livraison par jour.
+df_tickets['DATE'] = df_tickets['TICKET_DATE'].apply(lambda x:datetime.date(x.year,x.month,x.day))
+df_orders['DATE'] = df_orders['SHIPDATE'].apply(lambda x:datetime.date(x.year,x.month,x.day))
+#%% Clean data
+df_tickets.dropna(axis=0,inplace=True)
[…]

### 2023-02-08 Data/instances/script/instance_horaire.py

@@ -1,0 +1,219 @@
+
+#%%
+import pandas as pd
+import os
+import datetime
+from collections import Counter
+#%%
+from driver import Driver
+
+#%%
+TICKET_FOLDER = "/".join([os.path.dirname(__file__), "tickets/"])
+INSTANCE_FOLDER = "/".join([os.path.dirname(__file__), "instances/"])
+#%%
+""" Load dataframes """
+df_tickets = pd.read_excel(f"{os.path.dirname(__file__)}/DTICKETHISTAB_BETON.xlsx")
+#%%
+df_orders = pd.read_excel(f"{os.path.dirname(__file__)}/DORDERSTAB_BETON.xlsx",sheet_name='data')
+#%%
+df_employee = pd.read_excel(f"{os.path.dirname(__file__)}/EMPLOYETAB.xlsx")
+df_employee.drop(['COMP_NBR', 'emp_sch_specificite'], axis=1, inplace=True)
+df_employee.set_index('EMPL_NBR',inplace=True)
+#%%
+df_depots = pd.read_excel(
+    f"{os.path.dirname(__file__)}/adresses_complete.xlsx")
+df_depots = df_depots[df_depots['Location_type'] != 'customer']
+df_depots.dropna(axis=1,  inplace=True)
+df_depots['NBR'] = df_depots['JOB_DESC'].apply(
+    lambda x: x if len(x.split(' ')) == 1 else int(x.split(' ')[1]))
+df_depots.set_index('NBR',inplace=True)
+#%%
+# Depot location
+df_depots_loc = pd.read_excel(
+    f"{os.path.dirname(__file__)}/LOCATIONTAB.xlsx")
+df_depots_loc['LOC_NBR'] = df_depots_loc['LOC_NBR'].apply(lambda x: int(x))
+df_depots_loc.set_index('LOC_NBR',inplace=True)
+#%%
+# Les tickets contiennent les séquences des opérations de distribution
+# Un ticket est la livraison de béton sur un chantier par un livreur
+
+# Je dois déterminer la séquence de livraison par jour.
+df_tickets['DATE'] = df_tickets['TICKET_DATE'].apply(lambda x:datetime.date(x.year,x.month,x.day))
+df_tickets['WEEK'] = df_tickets['DATE'].apply(lambda x:datetime.date(x.year,x.month,x.day).isocalendar()[1])
[…]

### 2023-02-08 Data/instances/script/instances.py

@@ -1,0 +1,112 @@
+
+"""Create instances."""
+
+#%%
+import pandas as pd
+import numpy as np
+import os
+from driver import Driver
+from nodes import Node
+
+INSTANCE_FOLDER = "/".join([os.path.dirname(__file__), "instances/"])
+
+#%%
+# Load dataframes
+df_nodes = pd.read_excel(f"{os.path.dirname(__file__)}/DORDERSTAB_BETON.xlsx")
+df_depots = pd.read_excel(
+    f"{os.path.dirname(__file__)}/adresses_complete.xlsx")
+df_depots = df_depots[df_depots['Location_type'] != 'customer']
+df_driver = pd.read_excel(f"{os.path.dirname(__file__)}/EMPLOYETAB.xlsx")
+df_driver.drop(['COMP_NBR', 'emp_sch_specificite'], axis=1, inplace=True)
+df_tickets = pd.read_excel(
+    f"{os.path.dirname(__file__)}/DTICKETHISTAB.xlsx")
+#%%
+
+df_nodes.columns
+
+#%%
+print(f" {len(df_nodes)} ordres")
+print(f" {len(df_driver)} chauffeurs")
+print(f" {len(df_depots)} depots")
+# df_nodes.head()
+#%%
+# Load depots nodes
+depots = df_depots.copy()
+depots.dropna(axis=1, how='all', inplace=True)
+depots['JOB_DESC'] = depots['JOB_DESC'].apply(
+     lambda x: x if len(x.split(' ')) == 1 else int(x.split(' ')[1]))
+list_depots = []
+depots_loc = []
+for i in range(len(depots)):
+    d = Node(id=i)
+    d.adresse = depots.loc[i:,('ORDER_DESC')]
+    d.location_id = depots.iloc[i]['Location ID']
+    d.location_type = depots.iloc[i]['Location_type']
+    d.coord = (depots.iloc[i]['LONGITUDE_X'], depots.iloc[i]['LATITUDE_Y'])
+    d.depot_loc = depots.iloc[i]['JOB_DESC']
+
+    depots_loc.append(d.depot_loc)
+    list_depots.append(d)
+
+#%%
+df_driver_subset = pd.merge(
+    df_driver, depots, how='inner', left_on='emp_usi_numero', right_on='JOB_DESC')
+print(f" {len(df_driver_subset)} Vs {len(df_driver)} au total")
+
+#%% Load Drivers
+# take only drivers with ORDER_LOC in depots
[…]

### 2023-02-08 Data/instances/script/matrices.py

@@ -1,0 +1,97 @@
+import os
+import pandas as pd
+import numpy as np
+import requests
+
+filepath = "adresses_complete.xlsx"
+API_KEY = "AIzaSyAnV5rCE2wBHFyrwoHCD-7kEXGYb3s2g60"
+# https://maps.googleapis.com/maps/api/distancematrix/json?origins=Seattle&destinations=San+Francisco&key=YOUR_API_KEY
+url = "https://maps.googleapis.com/maps/api/distancematrix/json?"
+
+NOTVALUE = -9
+#Use google maps API pour trouver les distances
+
+def get_dist_time(vers, de):
+    """
+    return distance and time values
+    """
+    try:
+        r = requests.get(f"{url}origins={de}&destinations={vers}&key={API_KEY}")
+        # print(r.json())
+        dist = r.json()['rows'][0]['elements'][0]['distance']['value']
+        time = r.json()['rows'][0]['elements'][0]['duration']['value']
+    except (KeyError, TypeError, IndexError):
+        return NOTVALUE, NOTVALUE
+    return dist, time
+
+
+def get_matrices(nodes_test):
+    """
+        Create distance and time matrices
+    """
+    dim = len(nodes_test)
+    matrices_dist = [np.arange(-1, dim) for i in range(dim+1)]
+    matrices_time = [np.arange(-1, dim) for i in range(dim+1)]
+    for i in range(dim):
+        print(f"ligne {i+1}/{dim} ")
+        ligne = nodes_test.iloc[i]
+        de = f"{ligne['LATITUDE_Y']},{ligne['LONGITUDE_X']}"
+        de1 = '+'.join(ligne['ORDER_DESC'].split(" ")) + "+" + ligne['JOBCITY']
+
+        matrices_dist[i+1][0] = i
+        matrices_dist[i+1][i+1] = 0
+        for j in range(i+1, dim):
+            col = nodes_test.iloc[j]
+            vers = f"{col['LATITUDE_Y']},{col['LONGITUDE_X']}"
+            vers1 = '+'.join(col['ORDER_DESC'].split(" ")
+                             ) + "+" + col['JOBCITY']
+            dist1, time1 = get_dist_time(de1, vers1)
+            dist, time = get_dist_time(de, vers)
+
+            if dist1 == NOTVALUE and dist == NOTVALUE:
+                dist2 = NOTVALUE
+                time2 = NOTVALUE
[…]

### 2023-02-08 Data/instances/script/ticket.py

@@ -1,0 +1,175 @@
+"""Ticket."""
+#%%
+import pandas as pd
+import os
+import datetime
+from collections import Counter
+#%%
+from driver import Driver
+#%%
+def time_to_minute(stamps):
+
+    return round(stamps.hour*60 + stamps.minute + stamps.second/60,2)
+
+#%%
+TICKET_FOLDER = "/".join([os.path.dirname(__file__), "tickets/"])
+INSTANCE_FOLDER = "/".join([os.path.dirname(__file__), "instances/"])
+#%%
+""" Load dataframes """
+df_tickets = pd.read_excel(f"{os.path.dirname(__file__)}/DTICKETHISTAB_BETON.xlsx")
+#%%
+df_orders = pd.read_excel(f"{os.path.dirname(__file__)}/DORDERSTAB_BETON.xlsx")
+#%%
+df_employee = pd.read_excel(f"{os.path.dirname(__file__)}/EMPLOYETAB.xlsx")
+df_employee.drop(['COMP_NBR', 'emp_sch_specificite'], axis=1, inplace=True)
+df_employee.set_index('EMPL_NBR',inplace=True)
+#%%
+df_depots = pd.read_excel(
+    f"{os.path.dirname(__file__)}/adresses_complete.xlsx")
+df_depots = df_depots[df_depots['Location_type'] != 'customer']
+df_depots.dropna(axis=1,  inplace=True)
+df_depots['NBR'] = df_depots['JOB_DESC'].apply(
+    lambda x: x if len(x.split(' ')) == 1 else int(x.split(' ')[1]))
+df_depots.set_index('NBR',inplace=True)
+#%%
+# Depot location
+df_depots_loc = pd.read_excel(
+    f"{os.path.dirname(__file__)}/LOCATIONTAB.xlsx")
+df_depots_loc['LOC_NBR'] = df_depots_loc['LOC_NBR'].apply(lambda x: int(x))
+df_depots_loc.set_index('LOC_NBR',inplace=True)
+#%%
+# Les tickets contiennent les séquences des opérations de distribution
+# Un ticket est la livraison de béton sur un chantier par un livreur
+
+# Je dois déterminer la séquence de livraison par jour.
+df_tickets['DATE'] = df_tickets['TICKET_DATE'].apply(lambda x:datetime.date(x.year,x.month,x.day))
+df_orders['DATE'] = df_orders['SHIPDATE'].apply(lambda x:datetime.date(x.year,x.month,x.day))
[…]

### 2023-02-08 modele/conflict_refiner.py

@@ -1,0 +1,23 @@
+import cplex
+
+# Create a new modeler object
+modeler = cplex.Cplex()
+
+# Read the .lp file
+modeler.read("filename.lp")
+
+# Extract the model from the modeler
+model = modeler.get_model()
+
+# Set up the conflict refiner
+conflict = model.conflict()
+conflict.set_level(2)  # Set the level of detail to 2
+
+# Solve the model
+model.solve()
+
+# Check if the model is infeasible
+if model.solution.is_infeasible():
+  # Identify the minimal conflicting set of constraints or variables
+  minimal_conflict = conflict.refine()
+  print("Minimal conflicting set:", minimal_conflict)

### 2023-02-08 Article/article.tex
version  envoyée en janvier 2023

@@
 label/.style   = {base,font=\ttfamily\normalsize, text width = 8em, 
-  inner sep=0em, rounded corners=1mm,bottom color=gray!80,},
+  inner sep=0em, rounded corners=1mm,bottom color=red!80,},
   dummy/.style    = {circle,draw}
@@
 \begin{abstract}
-    %We study the daily scheduling of concrete delivery.
-Truck drivers are dispatched based on assignment priority and remaining working time over the horizon planning.
-A customer may be served from several available production plants.
-Due to the heterogeneity of the fleet, the number of visits to a customer location is not known ahead of time.
-The objective is to schedule drivers while minimizing travel times, waiting duration, idle time, and overtime.
-We solve this problem using a greedy randomized adaptive search procedure.
-Computational experiments with real data show the performance of our methods over the approach used by a local company.
+   The concrete delivery problem (CDP) is a combinatorial optimization problem that involves scheduling the delivery of ready-mixed concrete to construction sites while balancing the conflicting goals of minimizing transportation costs and maximizing customer satisfaction.
+This paper proposes an exact formulation and a heuristic approach based on the Greedy Randomized Adaptive Search Procedure (GRASP) to tackle a new variant of the CDP that incorporates realistic side constraints, such as drivers' working shifts, minimum working time for each driver, and overtime penalties.
+Additionally, the problem also considers the possibility of customers requiring multiple types of concrete to be delivered within the same time window.
+The performance of the proposed heuristic is evaluated on real-world and generated CDP instances, and it is compared to another variant to demonstrate its effectiveness.
+
 \end{abstract}
 
-% \begin{keyword}
-\noindent{\textbf{Keywords:} Vehicle routing; scheduling; ready-mixed concrete; concrete delivery
-% \end{keyword}
+% \begin{keywords}
+\noindent{\textbf{Keywords:} Vehicle scheduling;  concrete delivery; GRASP; ready-mixed concrete } 
+% \end{keywords}
 
@@
 \label{sec:intro}
+Concrete is a widely used building material in construction projects.
+It is a perishable product that can be affected by many different factors that impact its quality \citep{sinha_quality_2021} which is crucial for the durability and strength of the final construction.
 
-Construction projects involve a huge movement of equipment, people, and materials.
-Among the latter, concrete is one of the most used.
-It is a perishable product with many factors affecting its quality \citep{sinha_quality_2021}.
-It comes in two types: ready-mixed concrete (RMC) and site-mixed concrete (SMC).
-As its name implies, SMC is produced on the spot using raw materials (water, aggregates, and cement) stored on the construction site, while RMC is manufactured in a batch plant and delivered to the construction site.
-SMC can avoid delays caused by road traffic, but has a slower and more difficult production process, requires storage for mixing materials and equipment, and is suitable for low amounts of concrete.
-It is quite the opposite for the RMC, which has better quality and benefits from lower production cost \citep{muresan_comparing}.
-Nevertheless, the batch plant manager must employ a fleet of high-cost revolving drum trucks (concrete mixers) to dispatch the ready-mixed concrete.
-Hence, the key to reaping the RMC benefits is to ensure efficient and prompt delivery on the construction site.
+Concrete comes in two types: ready-mixed concrete (RMC) and site-mixed concrete (SMC).
+RMC is manufactured in a batch plant and delivered to the construction site, while SMC is produced on-site using raw materials stored on the construction site.
+Site-mixed concrete can avoid delays caused by road traffic, but it has a slower and more difficult production process, requires storage for mixing materials and equipment, and is suitable for low amounts of concrete.
+On the other hand, RMC has better quality and benefits from lower production costs \citep{muresan_comparing}.
+However, to take advantage of these benefits, the batch plant manager must ensure efficient and prompt delivery on the construction site, which may require a fleet of high-cost revolving drum trucks (concrete mixers) to dispatch the ready-mixed concrete.
 
@@
 In this paper, we study a variant of the Concrete Delivery Problem to schedule the daily production and dispatching of RMC for a company located in the province of Quebec, Canada.
-This company produces RMC in multiple batching plants with different production rates, using a fleet of concrete mixers of various capacities.
-Each plant has its specific fleet, but a truck can load elsewhere if deemed necessary, except it must return to its home plant at the end of the day.
-They own two types of trucks with different capacities but can call on an external fleet when needed.
-They serve a construction site from any batch plant, with the first delivery starting at the time specified by the customer.
-Loading and unloading a concrete mixer depends on the truck capacity, a plant's loading, and a construction site's unloading rate, respectively.
-The problem is like the one addressed in \cite {schmid2009hybrid, schmid2010hybridization}, except for the plant's production rate and the lack of unloading instrumentation in our case.
-Also, as per union rules, they assign a truck driver to a delivery task based on his seniority.
+This company operates multiple batching plants with varying production rates, using a fleet of concrete mixers with different capacities.
+Each plant has its own fleet of trucks, however, the trucks are able to move between plants if necessary.
+The trucks must return to their home plant at the end of the day.
+The company owns two types of trucks with different capacities but also has the option to call on an external fleet when needed.
+They serve construction sites from any of their batch plants, with the first delivery starting at the time specified by the customer.
+The loading and unloading of a concrete mixer depend on the truck capacity, the loading rate at the plant, and the unloading rate at the construction site.
+Drivers are allocated based on their daily work schedules.
+A customer may request several types of concrete to be delivered within the same time window, with no required sequence for the orders, but an order can only start after the completion of the previous order.
+This constraint generalizes the linked order constraints of \cite{durbin2008or} where some customers place two orders for the same day and request that they are linked (the second order begins only after the first order is completed).
+The setting of this study is similar to a variant of the Concrete Delivery Problem previously studied in \cite{schmid2009hybrid, schmid2010hybridization}.
+However, it includes the plant's production rate and driver shift schedule and does not consider unloading instrumentation in contrast to the previous research.
 The company uses a centralized dispatcher system to schedule all daily orders, but this system has issues satisfying all daily demands without using an external fleet.
+% Also, as per union rules, they assign a truck driver to a delivery task based on his seniority.
 
@@
 Hence, concrete mixers must deliver RMC at the planned construction site shortly after its production.
-RMC production and delivery is then an example of a Just-In-Time (JIT) production system in construction \citep{tommelein1999just}.
+RMC production and delivery is described by \cite{tommelein1999just} as an example of a Just-In-Time (JIT) production system in construction.
 A customer quantity requirement is often greater than the truck size and must be fulfilled by multiple deliveries.
 In that sense, CDP is like the Vehicle Routing Problem (VRP) with split delivery (VRPSD) \citep{archetti2008split}, except that the same truck may visit a customer more than once.
+Concrete hardens quickly, so when there are multiple deliveries, they must be done back-to-back or at least close in time to avoid the problem of cold joint, which can reduce the strength and durability of the concrete.
+Customers request to be served within a specific time window, which can complicate truck-loading schedules when a plant can only load one truck at a time.
+Similarly, only one truck can unload at a time at a customer location, sometimes leading to concrete mixers queuing and waiting their turn to deliver.
+Furthermore, with trucks of varied sizes, loading, travel, and unloading times that may be uncertain, the Concrete Delivery Problem is a complex problem to solve and has been proven to be NP-hard \citep{asbach2009analysis,kinable2014concrete}.
 
-Since concrete hardens quickly, in case of multiple deliveries, back-to-back deliveries must be continuous or at least close in time to avoid the problem of cold joint that negatively affects the quality of the concrete.
-Customers request to be served within a specific time window, complicating truck loading schedules when a plant can only load one truck at a time.
-Likewise, often only one truck can unload at a time at a customer location, sometimes leading to concrete mixers queuing and waiting their turn to deliver.
-Furthermore, with trucks of varied sizes, loading, travel, and unloading times that may be uncertain, the Concrete Delivery Problem is a complex problem to solve and has been proven by \cite{asbach2009analysis} to be NP-hard.
-With a single batching plant, the objective is to maximize the service level (serve all customers and avoid delays between subsequent deliveries and truck queuing).
-With additional production centers to choose from, a minimization of the total travel time is added to the objective.
-%However, these objectives can be conflictive.
-% We go a step further to add human constraints to the CDP, overtime, minimum and maximum hour of work,
+In this paper, we propose a mathematical model and a Greedy Randomized Search Procedure (GRASP) heuristic solution to solve a new variant for the CDP.
+Our model takes into consideration the working shifts of the drivers and the scheduling of multiple orders within the same time window at a construction site.
+To the best of our knowledge, this paper is the first that deals with these specific constraints in the context of the CDP.
 
-The remainder of the paper is organized as follows.
-Section \ref{lit_review} presents an overview of the CDP-related literature.
-Section \ref{desc_form} provides a formal description of the problem.
-Then in Section \ref{method}, we describe the Greedy Randomized Search Procedure (GRASP) algorithm and the constructive heuristics we design to solve this variant of the CDP.
-Computational experiments and our conclusions follow in Sections \ref{comp_exp} and \ref{concl}, respectively.
+The rest of the paper is structured as follows: Section \ref{lit_review} provides a literature review of previous work related to the CDP.
+Section \ref{desc_form} provides a formal description and mathematical model of the problem.
+In Section \ref{method}, we describe the Greedy Randomized Search Procedure algorithm and the constructive heuristics developed to solve this variant of the CDP.
+Section \ref{comp_exp}  presents computational experiments to evaluate the proposed approach, and finally, the conclusions are presented in Section \ref{concl}.
 
@@
 
-A text mining approach for reviewing the ready-mixed concrete literature \citep{maghrebi2015text} showed that concrete technology and material science are the main core of research in this area.
-The first academic research on concrete batching and delivery began in the late 1990s.
-\cite{tommelein1999just} describe RMC as a prototypical example of a Just-In-Time production system in construction and identify two practices occurring when delivering it.
-An alternative is for the customer to haul the product from the batch plant with their concrete mixer.
-The other approach is where the batch plant delivers the concrete to the customer's location.
-This latter approach is the one studied in all related papers found in the literature.
+A text mining approach by \cite{maghrebi2015text} for reviewing the ready-mixed concrete literature showed that concrete technology and material science are the main core of research in this area.
+Academic research on concrete batching and delivery began in the late 1990s.
+\cite{tommelein1999just} described ready-mixed concrete (RMC) as a prototypical example of a Just-In-Time production system in construction and identified two practices for delivering it.
+One approach is for the customer to haul the product from the batch plant with their concrete mixer, while the other approach is for the batch plant to deliver the concrete directly to the customer's location.
+This latter approach is the one that has been studied in all related papers found in the literature.
 
-Several works on the CDP mainly focussed on Simulation related methods, either standalone \citep{zayed2001simulation, wang2001scheduling, tian_simulation_based_2010, panas_simulation_based_2013, galic2016simulation}, etc.
-or hybridized with optimization methods \citep{feng2004optimizing, lu2005optimized, feng_integrating_2006}, etc.
-to schedule and dispatch concrete production and delivery.
+Several works to schedule and dispatch concrete production and delivery have mainly focused on simulation-related methods.
+These methods can be standalone, such as those used by \cite{zayed2001simulation, wang2001scheduling, tian_simulation_based_2010, panas_simulation_based_2013} and \cite{galic2016simulation}, among others.
+Alternatively, these methods can be hybridized with optimization techniques, such as those used by \cite{feng2004optimizing, lu2005optimized, feng_integrating_2006}, etc.
 \cite{wang2001scheduling} developed a simulation model to reveal the effect and value of the concrete mixers' inter-arrival time on the productivity of hired unloading equipment on site.
-With a combination of Genetic Algorithm (GA) and simulation process, \cite{feng2004optimizing} tried to minimize the total waiting time for trucks at a customer site.
-Loading trucks with identical capacities is carried out at the same batch plant.
-Loading and unloading durations are fixed.
-GA is used to find the most efficient and effective loading sequence of RMC trucks to be assigned to different construction sites.
-The simulation process determines the loading, arrival, departure, and waiting time of trucks and thus evaluates the cost of each dispatching sequence.
-They evaluated the method using data from a batch plant in Taiwan with up to nine customers served.
-\cite{mayteekrieangkrai2015optimized} solved the same problem with the same data using a bee algorithm (BA) and found better solutions than the GA.
-\cite{lu2005optimized} used the same combination of GA and simulation to decide on the optimal number of concrete mixers to be deployed together with an optimal schedule for batching and delivering concrete.
-Their objective is to minimize the site crew idle times due to late concrete deliveries plus truck queuing time.
-The particularity of this setting is that mortar useful for unloading pump lubrication must be delivered on-site before concrete.
-Therefore, mortar batching and delivery are also modeled in the simulation.
+\cite{feng2004optimizing} used a combination of Genetic Algorithm (GA) and simulation process to minimize the total waiting time for trucks at a customer site.
+The study focused on loading trucks with identical capacities at the same batch plant, with fixed loading and unloading durations.
+The GA was used to find the best loading sequence of RMC trucks to be assigned to different construction sites.
+The simulation process determined the loading, arrival, departure, and waiting time of trucks and thus evaluated the cost of each dispatching sequence.
+They evaluated their method using data from a batch plant in Taiwan with up to nine customers served.
+\cite{mayteekrieangkrai2015optimized} addressed the same problem with the same data using a bee algorithm (BA) and found better solutions than the GA.
+\cite{lu2005optimized} used the same combination of GA and simulation to determine the optimal number of concrete mixers to be deployed and an optimal schedule for batching and delivering concrete.
+Their objective was to minimize the idle time of the site crew due to late concrete deliveries and truck queuing time.
+In this setting, it was also necessary to deliver a batch of mortar on-site to lubricate the unloading pump before the concrete delivery.
+As such, the simulation model also included the batching and delivery of mortar.
 Finding the best RMC truck size was also the purpose of the discrete-event simulation model proposed by \cite{panas_simulation_based_2013}.
 
-Besides simulation-based methods, we find in the literature works using metaheuristics \citep{faria2006distributed, misir2011selection, maghrebi2016sequential, yang2022concrete}, exact methods \citep{yan2007optimal, kinable2014concrete}, matheuristics \citep{schmid2009hybrid, schmid2010hybridization}, Benders Decomposition \citep{maghrebi2014benders}, Column Generation (CG) \citep{maghrebi2014solving, maghrebi2016column}, Lagrangian relaxation \citep{narayanan2015using}, and machine learning approach \citep{graham2006modeling, maghrebi2014exploring, maghrebi2016matching}.
-\cite{matsatsinis2004towards} designed a decision support system (DSS) for the dynamic routing of both concrete and pumps that may be necessary for some sites to help the concrete unloading.
-Three plants are available, but vehicles fulfilling the same order must all load at the same plant.
-An order that cannot be executed may be postponed for the next day.
-The routing of the pumps is modeled as a multi-depot VRP with time windows.
-\cite{naso2007genetic} proposed a sequential GA method combined with constructive heuristics to solve a more general variant of the CDP.
-In this problem, besides batching concrete delivered to a customer site, the plant's production schedule must include orders picked up by the customers themselves.
-The algorithm first schedules the plant loading operations before scheduling truck job deliveries.
-A non-linear model minimizing transportation costs, waiting times, outsourced costs, and overtime work is also developed.
-The authors ran experiments using real-world instances of a concrete supply chain in the Netherlands.
-They found a reduction in the number of requests redirected to external companies.
-\cite{yan2007optimal} also emphasized overtime considerations in their paper scheduling RMC for one batching plant with two loading docks.
-Overtime wages are paid for factory and construction site operations after 4 PM.
-A mixed integer programming (MIP) model on a time-space network is developed to minimize travel times and operating costs at both normal and overtime working hours at the plant and the construction sites.
-Real data consisting of 3 days of operation is tested using a two-stage algorithm.
+In addition to simulation-based methods, several other approaches have been used in the literature to solve the CDP.
+These include metaheuristics \citep{faria2006distributed, misir2011selection, maghrebi2016sequential, yang2022concrete}, exact methods \citep{yan2007optimal, asbach2009analysis, kinable2014concrete}, matheuristics \citep{schmid2009hybrid, schmid2010hybridization}, Benders Decomposition \citep{maghrebi2014benders}, Column Generation (CG) \citep{maghrebi2014solving, maghrebi2016column}, Lagrangian relaxation \citep{narayanan2015using}, and machine learning approaches such as those used by \cite{graham2006modeling, maghrebi2014exploring, maghrebi2016matching}.
+\cite{matsatsinis2004towards} designed a decision support system (DSS) for the dynamic routing of both concrete and pumps that may be necessary for some construction sites to aid in the unloading of concrete.
+The DSS considered the availability of three plants but stipulated that vehicles fulfilling the same order must all load at the same plant.
+Orders that could not be executed immediately could be postponed for the next day.
+The routing of the pumps was modelled as a multi-depot vehicle routing problem with time windows.
+\cite{naso2007genetic} proposed a sequential GA method combined with constructive heuristics to solve another variant of the CDP.
+In this problem, the plant's production schedule must account for orders for concrete to be delivered to a customer site and orders that customers must pick up themselves.
+The algorithm first schedules the plant loading operations before scheduling truck deliveries.
+The authors also developed a non-linear model that minimizes transportation costs, waiting times, outsourced costs, and overtime work.
+They ran experiments using real-world instances of a concrete supply chain in the Netherlands and found a reduction in the number of requests redirected to external companies.
+\cite{yan2007optimal} also considered overtime considerations in their paper, which focused on scheduling RMC for one batching plant with two loading docks.
+The study took into account that overtime wages are paid for factory and construction site operations after 4 PM.
+They developed a mixed integer programming (MIP) model on a time-space network to minimize travel times and operating costs at both normal and overtime working hours at the plant and the construction sites.
+They tested the model using real-data consisting of 3 days of operation using a two-stage algorithm.
 First, they solved the MIP relaxation with CPLEX.
-Then they simplify the original model by fixing some decision variables before solving it.
-This algorithm seems to improve the actual plant operation by 10\%.
-A time-space network is the principal component of the real-time DSS developed by \cite{durbin2008or} to solve a dynamic CDP every five minutes.
-The DSS can receive new orders, schedule them on the fly, and deal with unexpected events such as plant closures, truck breakdowns, delays in transportation times, etc.
-Combined with a Tabu Search (TS) heuristic to warm start CPLEX, the model is performant enough to solve instances with up to 1,500 loads per day with up to 250 trucks.
-The authors consider the case of a customer who places two orders, with the first being completed before the second starts.
-The real-time planning and monitoring of CDP are studied more in detail by \cite{garza2021dynamic} in his thesis.
+Then, they simplified the original model by fixing some decision variables before solving it.
+The algorithm was found to improve the actual plant operation by 10\%.
+A time-space network is the key component of the real-time decision support system (DSS) developed by \cite{durbin2008or} to solve a dynamic CDP problem every five minutes.
+The DSS is able to receive new orders, schedule them on the fly, and handle unexpected events such as plant closures, truck breakdowns, and delays in transportation times.
+The authors combined the DSS with a Tabu Search (TS) heuristic to warm start CPLEX, which made the model performant enough to solve instances with up to 1,500 loads per day with up to 250 trucks.
+The DSS also considers the case of a customer who places two orders, with the first being completed before the second starts.
+Further insights on the real-time planning and monitoring of CDP are available in \cite{garza2021dynamic}.
 Another variant of the CDP is modeled by \cite{schmid2009hybrid} as an integer multicommodity network flow (MCNF) problem on a time-space network.
-In this paper, concrete is delivered using a heterogeneous fleet of vehicles, and each plant can load an illimited number of trucks simultaneously.
-Among the trucks, some have specialized equipment and must arrive first at certain construction sites to assist concrete-mixers unloading.
+In this paper, concrete is delivered using a heterogeneous fleet of vehicles, and each plant can load an unlimited number of trucks simultaneously.
+Some of the trucks have specialized equipment and must arrive first at certain construction sites to assist in unloading the concrete.
 The objective is to fulfill all orders, minimize the travel cost, and avoid delays between two consecutive unloading operations for an order.
 The model is typically solved using a matheuristic algorithm that combines the MCNF with a variable neighborhood search (VNS) heuristic.
-The method can solve large problem instances with more than 60 orders per day quickly without encountering any memory issues.
-The same problem is addressed by \cite{schmid2010hybridization}.
-The authors proposed a MIP model combined with a VNS and a very large neighborhood search (VLNS) to develop two matheuristics approaches.
+The method can quickly solve large problem instances with more than 60 orders per day without encountering any memory issues.
+The same problem is addressed by \cite{schmid2010hybridization} who proposed a mixed integer programming (MIP) model combined with a VNS and a very large neighborhood search (VLNS) to develop two matheuristics approaches.
 Comparisons between both matheuristics and a standalone VNS show that the former methods are much better and suitable for solving larger problem instances.
@@
 
-Regarding objectives, most authors have been interested in minimizing travel time and delays between consecutive deliveries.
-Some authors, however, were more interested in maximizing only customer satisfaction.
+Regarding objectives, most authors have focused on minimizing travel time and delays between consecutive deliveries.
+However, some authors have been more interested in maximizing customer satisfaction alone.
 We find these situations in the works of \cite{durbin2008or, kinable2014concrete, kinable2014logic, sulaman2017simulated}.
@@
 However, the model did not include constraints limiting the time that concrete may reside in a truck.
-The authors also propose a constructive heuristic that schedules the visits to the customers one by one according to the start time of the visit and the truck capacity.
+The authors propose a constructive heuristic that schedules the visits to the customers one by one according to the start time of the visit and the truck capacity.
 The procedure is invoked multiple times for different permutations of the customer's order which is determined using the steepest descent (SD) local search procedure.
@@
 A detailed analysis of the MIP model presented in cite{kinable2014concrete}and of two more compact models can be found in the thesis of \cite{hernandez_lopez_study_2020}.
-In \cite{kinable2014logic}, we find an attempt to solve the previous problem with a Logic Based Benders' approach.
+In \cite{kinable2014logic}, we found an attempt to solve the previous problem with a Logic Based Benders' approach.
 \cite{sulaman2017simulated} expand upon the SD heuristic proposed in \cite{kinable2014concrete}, proposing a Simulated Annealing (SA) with a Time-Slot Heuristic (TH).
@@
 The MIP model is used in a local search scheme as a black-box solver to reoptimize an incumbent solution in which a neighborhood operator has unfixed some variables.
-
-%\cite{tzanetos2023systematic} present a systematic search and mapping review of the concrete delivery problem.
-this paper fills that gap by (a) performing a comprehensive search on the CDP, and (b) mapping out and categorizing the existing literature.
-Specifically, this paper provides an overview of the various methods of the literature, categorizing the problem formulations based on the literature’s different concepts,
+\cite{tzanetos2023systematic} provides an overview of the various methods used in the literature to address the Concrete Delivery Problem and categorizes the problem formulations based on the different concepts used in the literature, such as Integer Programming (IP), time-space network, and job shop model.
+They also discussed the consistency between industry needs and existing constraints and provided insights into the datasets corresponding to real-world cases, identifying the necessary data for practitioners.
 
@@
 \label{desc_form}
-The problem studied within this paper is about the distribution of ready-mixed concrete from a Canadian company operating in the greater Montreal area.
-Concrete order from a customer arrives at a central center and is assigned to one of the batching plants from where the product will be produced and delivered to the customer.
-We define an instance of our problem on a set of batching plants, a set of customer orders, and a set of drivers.
+The focus of this paper is the distribution of ready-mixed concrete from a Canadian company that operates in the greater Montreal area.
+When a customer places an order, it is received at a central center and then assigned to one of the company's batching plants.
+These plants produce the concrete and then deliver it to the customer.
+The problem we are examining involves a set of batching plants, a set of customer orders, and a set of concrete-mixer drivers.
 
@@
 
-The company owns and operates a total of eight concrete batching plants located in different geographical areas.
-The loading bays in each plant can only accommodate one truck at a time, resulting in trucks lining up.
+The company owns a total of eight batching plants, located in various geographical regions.
+Each plant has a loading bay that can only accommodate one truck at a time, which leads to trucks lining up.
 Let $\mathcal{B}$ be the set of batching plants.
-The plants are heterogeneous, as each plant $b$ has its hourly loading rate $\tau^l_b$ which influences the loading time duration.
-After loading, a driver takes $\alpha_b$ minutes to adjust the concrete in his truck before departing to the customer site.
-Any batch plant can serve a construction site if the traveled time between them is less than the concrete lifespan $\delta$.
-Each plant has its assigned fleet, but it can get a vehicle from another plant or even hire an external fleet.
-If $q_i$ is the quantity of RMC loaded to customer $i$ from plant $b$, we compute the loading time duration $LD^b_i$ as
-
-\begin{equation}
-    \label{eq:LD}
-    {LD}^b_i = \frac{q_i}{\tau^l_b}.
-\end{equation}
+The plants are heterogeneous, as each plant has its own hourly loading rate, represented by $\tau^l_b$, which affects the duration of the loading process.
+After the concrete has been loaded, the driver spends $\alpha_b$ minutes adjusting the concrete in the truck before heading to the customer site.
+A batching plant can serve a construction site as long as the travel time between the two is less than the concrete's lifespan, which is represented by $\Delta$.
+Each plant has its own assigned fleet of trucks, but it can also borrow trucks from other plants or hire external fleets if necessary.
 
@@
 
-A customer $i$ requests one or several types of concrete to be delivered to his construction site on a particular day for a total amount of $q_i$.
-When placing an order, he specifies the number of types of RMC $P_i \geq 1$, the quantity $q^p_i$ of concrete of each type $p$, the arrival time $a_i$ of the first concrete mixer, and the unloading rate $\tau^u_i$ at his site.
-$\mathcal{C}$ is the set of construction sites (customers) with a planned delivery for the day.
-
-We refer to order $o^p_i$ as the request of customer $i$ to receive a type $p$ of concrete.
-If the amount of concrete exceeds any truck capacity, several deliveries are scheduled to satisfy an order.
-$n^p_i \geq 1$ is the number of deliveries needed to satisfy $o^p_i$.
-Let $d^p_{ij}$ be the $j_{th}$ visit with load $q^p_{ij}$ for order $o^p_i$.
-We represent the delivery of order $o^p_i$ by the visits to the ordered set of nodes $\mathcal{D}^p_i= (d^p_{i0},d^p_{i1},\cdots, d^p_{in^p_i})$.
-All quantities of a type of concrete must be thoroughly delivered before moving on to another type.
-The orders $o^p_i$ $(1 \leq p \leq P_i$) of a customer $i$ are not ordered.
-An order can start at any time, provided it is completed before delivering another.
-We represent the delivery of customer $i$ by the set $\mathcal{D}_i= (\mathcal{D}^{p^0}_i, \mathcal{D}^{p^1}_i,\cdots,\mathcal{D}^{p^{P_i}})$, where $p^l$ is the $lth$ order delivered.
-We define for the first node of $\mathcal{D}^{p^0}_i$ the time window $ \left\lbrack a_i, b_i  \right\rbrack$ to ensure that the first delivery of $i$ must be due at $a_i$.
-Once a plant is chosen to produce a customer's first order, it must be the provider of all subsequent orders of this customer.
-
-To prevent cold joint in concrete, subsequent deliveries must be in just in time, or at least close in time.
+A customer $i$ requests a total of $q_i$ of one or several types of concrete to be delivered to his construction site on a particular day.
+We call an order a request for a specific type of concrete.
+When placing an order $o$, the customer $i$ specifies the required quantity, $q_o$, the desired arrival time of the first concrete mixer, $a_i$, and the unloading rate at their construction site, $\tau^u_i$.
+If the order requires more concrete than a single truck can carry, multiple deliveries are scheduled.
+Let $O_i$ be the unordered set of all orders requested by customer $i$.
+Each element of $O_i$ must be thoroughly delivered before moving on to another order.
+One element $o$ of $O_i$ must have its first delivery start at $a_i$, while the others can start any time after all deliveries of $o$.
+Once a plant is chosen to produce a customer's first delivery, it must be the provider of all subsequent deliveries of this customer.
+To prevent cold joint issues with the concrete, subsequent deliveries must be made just in time, or at least close in time.
 We define a maximum time lag $\gamma_i$ beyond which no next unloading operation should be allowed.
 The construction site unloading rate and the quantity to unload give the time necessary to discharge a truckload.
-The unloading time duration of $q^p_{ij}$ is
-\begin{equation}
-    \label{eq:UD}
-    {UD}_{d^p_{ij}} = \frac{q^p_{ij}}{\tau^u_i}
-\end{equation}
+$\mathcal{C}$ is the set of construction sites (customers) with a planned delivery for the day.
+$\mathcal{O}$ is the set of all requested orders.
 
@@
 
-%A driver is an important entity here, as we must incorporate some union constraints into our model. 
-The company has two types of trucks with capacities of 8 and 12 $m^3$.
-A driver $k$ is assigned to a plant from where he starts and ends his shift day, and drives one concrete mixer of capacity $Q_k$.
-The set of drivers is $K=\cup_{b \in \mathcal{B}}K_b $, $K_b$ being the set of drivers scheduled to start their shift at the batching plant $b$.
+The company has two types of concrete mixer trucks with capacities of 8 and 12 cubic meters.
+Each driver $k$ is assigned to a specific batching plant and is responsible for driving a truck with a specific capacity $Q_k$.
+The set of drivers is represented by $K =\cup_{b \in \mathcal{B}}K_b $, $K_b$ being the set of drivers scheduled to start their shift at the batching plant $b$.
+$t_{ij}$ is the known time to travel between any two locations $i$ and $j$.
+A scheduled driver $k$ is required to begin their shift at $H_k$, work for a minimum of $M_T$ hours and a maximum of $N_T$ hours during regular working hours, with the possibility of overtime of up to $O_T$ hours.
+%Let the continuous variable $WT_k$ be the working time of each driver $k$ during a day ($WT_k \leq O_T$).
+$\Gamma_1$ and $\Gamma_2$ are the incurred penalties when a driver works less than $M_t$ and more than $N_t$.
 
-A driver must work between four and eight hours in normal time, plus optional overtime of two hours at most, during a scheduled day.
-A driver mainly loads RMC at his home plant but can still drive at other plants and load there if needed.
-Batching plant produces concrete on-demand with recipes specific to customers, which means that truck cannot hold orders for more than one client, even if it has spare capacity.
-Thus, a driver must refill at a plant between two consecutive deliveries.
-After unloading the RMC, a driver takes $\beta_k$ minutes to clean the concrete mixer before traveling to his next loading plant.
-When assigning a driver to a delivery, priority is given to the employees with the highest seniority.
-
-For an order $o^p_i$, the number of deliveries $|D^p_i|$ is not known a priori, as we use a fleet with various capacities.
-However, we can compute the lower and upper bounds of $|D^p_i|$ using the capacities of the highest ($Q_{max}$) and smallest ($Q_{min}$) available trucks.
-The lower (upper) bound $\left\lceil \frac{q^p_i}{Q_{max}} \right\rceil$ $\left(\left\lceil \frac{q^p_i}{Q_{min}} \right\rceil \right)$ is the number of deliveries needed if we only use trucks of capacity $Q_{max}$ ($Q_{min}$).
+A driver typically loads RMC at their assigned batching plant, but may also be required to drive to and load at other plants if needed.
+Batching plant produces concrete on-demand with recipes specific to each customer.
+This means that a truck cannot hold orders for more than one customer, even if there is spare capacity.
+To fulfill multiple orders, a driver must refill at a plant between deliveries.
+After unloading the RMC, a driver takes $\beta_k$ minutes to clean the concrete mixer before travelling to his next loading plant.
+%When assigning drivers to deliveries, the company prioritizes employees with the highest seniority.
 
@@
 
-A solution to our problem is a set of decisions about truck loading schedules, driver assignments for different deliveries, and truck arrival times at construction sites for unloading.
-For a batching plant, the decision is to determine which driver should be loaded, when, and for which construction site.
-For a driver, the decision is the sequence of loading depots and delivering sites.
-And for a construction site, the arrival time of all daily scheduled deliveries.
+A solution to the problem involves making decisions about truck loading schedules, driver assignments to different deliveries, and truck arrival times at construction sites for unloading.
+For a batching plant, the decision involves choosing which driver to load, when to load them, and which construction site they should deliver to.
+For a driver, the decision involves determining the sequence of loading depots and delivery sites.
+And for a construction site, the decision involves determining the arrival times of all scheduled deliveries for the day.
 
-Let the binary variable $x^{bk}_{j}$ be $1$ if a driver $k$ loads at plant $b$ to serve the delivery node $j$. % $y_i=1$ if $i$ is completely served.
+We now present a formal definition of our problem.
+Let $n_o$ be the number of deliveries needed to fulfill order $o$.
+$n_o$ is not known in advance, as we use a fleet of trucks with various capacities.
+However, we can compute its lower ($n_o^{min}$) and upper ($n_o^{max}$) bounds using the capacities of the highest ($Q_{max}$) and smallest ($Q_{min}$) available trucks.
+The lower bound is the number of deliveries needed if we only use trucks of capacity $Q_{max}$, while the upper bound is the number of deliveries needed if we only use trucks of capacity $Q_{min}$.
+
+\begin{alignat}{3}
+    \label{mod:c1}
+       \left\lceil \frac{q_o}{Q_{max}} \right\rceil \leq n_o \leq \left\lceil \frac{q_o}{Q_{min}} \right\rceil  & \text{ } & 
+ \forall  o \in \mathcal{O}.
+\end{alignat}
+
+Let $d^j_{o}$ be the $j_{th}$ visit with load $q^j_{o}$ for order $o$.
+We represent the delivery of order $o$ by the visits to the ordered set of nodes $\mathcal{D}_o= \left(d^0_{o},d^1_{o},\cdots, d^{n_o}_{o}\right)$.
+We represent the deliveries of customer $i$ by the ordered set $\mathcal{D}_i= (\mathcal{D}_{o_1}, \mathcal{D}_{o_2},\cdots,\mathcal{D}_{o_{|O_i|}})$, where $o_r$ is the $rth$ order delivered.
+We will refer to $d \in \mathcal{D}_i$ ($d \in \mathcal{D}_o$) as the $dth$ potential delivery of customer $i$ (order $o$).
+$\mathcal{D}=\bigcup_{i\in \mathcal{C}} \mathcal{D}_i$ is the union of all delivery nodes.
+
+We represent each plant $b$ by the ordered set $L_b=\{L_{b,1}, L_{b,2},\cdots, L_{b,n(b)}\}$ of loading docks nodes, with $n(b)$ being the number of delivery nodes at a maximal travel time of $\Delta$ of $b$.
+We will refer to $L_{b,l}$ or  $l \in L_b$  as the $lth$ potential loading operation at $b$.
+$\mathcal{L}=\bigcup_{b\in B} L_b$ is the union of all loading docks.
+
+Each driver leaves and returns to his home plant each day.
+We represent a driver $k$ home plant's with a starting $s_k$ and ending $e_k$ depot.
+$S$ and $E$ are respectively the sets of starting and ending depot.
+
+We define our problem on a complete directed graph where $V=\{ S \cup \mathcal{L} \cup \mathcal{D} \cup E\}$ is the set of nodes.
+The arc sets are $A =  \{(i,j,k) \hspace*{1mm} \vert \hspace*{1mm} i, j \in V \hspace{1mm} k\in K \}$, and $A^D = \{(i,j) \hspace*{1mm} \vert \hspace*{1mm} i, j \in \mathcal{D} \hspace{1mm} \}$.
+$A$ corresponds to allowed movements of drivers from node $i$ to node $j$. For each driver $k$, the allowed movements are the following:
+\begin{itemize}
+    \item From the starting depot $s_k$ to a loading dock $l \in \mathcal{L}$ or to the ending depot $e_k$.
+    \item From a loading dock $l \in \mathcal{L}$ to a delivery node $d \in \mathcal{D}$.
+    \item From a delivery node  $d \in \mathcal{D}$ to a loading dock $l \in \mathcal{L}$ or to the ending depot $e_k$.
+\end{itemize}
+
+For a customer $c$, arcs in $A^D$ link consecutive delivery nodes of the same order   $\lbrace (i,j)\in \mathcal{D}_o, o \in \mathcal{O}_c, i < j  \rbrace$, and pair of delivery nodes of two different orders $ \lbrace (i,d^{0}_{o_2}),  i \in \mathcal{D}_{o_1}, i \geq n^{min}_{o_1}, o_1, o_2 \in \mathcal{O}_c, o_1 \neq o_2 \rbrace $.
+
+$\delta^{+}(i) = \{(i, j,k) \in A \}$, and $\delta^{-}(i) = \{(j, i,k) \in A \}$ are the outcoming and incoming arc sets of any node $i \in V$.
+Similarly, $\delta^{+}_D(i) = \{(i, j)  \in A^D\}$, and $\delta^{-}_D(i) = \{(j, i) \in A^D \}$ are the outcoming and incoming arc sets of delivery node $i \in \mathcal{D}$.
+
+Let the binary variable $x^{k}_{ij}$ be $1$ if driver $k$ travels from node $i$ to $j$.
+Binary variable $y_i$ equals $1$ if customer $i$ is completely served.
+$v_i$ is the start of the loading (unloading) operation at node $i \in L \cup D$.
+$w_i$ is the end of the loading (unloading) operation at node $i \in L\cup D$.
+The binary variable $u_{i,j}$ equals $1$ if delivery node $j$ is served just after delivery $i$.
+Binary variable $\sigma_{ib}$ equals 1 if customer $i$ orders' are loaded from plant $b$.
+% Binary variable $u_{o_{1},o_{2}}$ equals $1$ if order $o_{2}$ is served directly after order $o_{1}$.
+Binary variables $z_{o,d}$ ($z_{i,d}$) indicates whether or not delivery $d \in  \mathcal{D}_o$ ($d \in  \mathcal{D}_i$) is supposed to be executed.
+
+A driver $k$ leaves his starting node exactly once a day and travels to either a loading dock or his ending node.
+When loading at a loading dock $i$, he takes some time to adjust the concrete in the truck before travelling to delivery node $j$ to unload.
+The loading operation duration depends on the plant loading rate and the quantity $q^k_j$ of RMC to load towards $j$.
+After completing his last unloading operation of the day, he returns to his end node.
+
+\begin{alignat}{3}
+\label{mod:c2} \sum_{j \in \delta^{+}(s_k) }{x^k_{s_kj}} = 1 & \text{ } &\forall k \in K \\
+\label{mod:c3}   v_j \geq  w_i + \alpha_{i} + t_{ij} - M(1-x^{k}_{ij}) & \text{ }& \forall i \in \mathcal{L}, j \in \mathcal{D},  k \in K \\
+\label{mod:c4}  w_{i} \geq v_{i}  + \frac{q^k_j}{\tau^l_b} -M(1- x^{k}_{ij}) &  \hspace{5mm} &\forall  b \in \mathcal{B},  i \in L_{b}, j \in \mathcal{D}, k \in K \\
+\label{mod:c5} \sum_{j \in \delta^{-}(e_k)}{x^k_{je_k}} = 1 & \text{ }   &  \forall k \in K
+\end{alignat}
+
+Constraints (\ref{mod:c4}) ensure that two trucks cannot be loaded at the same time in a plant.
+
+\begin{alignat}{3}
+    \label{mod:c6}       w_{l-1} \leq v_{l} & \hspace{5mm} &\forall  b \in B,  l \in L_{b}, l \geq 1 
+\end{alignat}
+
+After completing a delivery  at node $i$, a driver must clean the truck, and travel to the next loading depot or end node.
+The unloading service depends on the construction site's rate and the received quantity.
+Unloading operations must finish $\Delta$ minutes after the begin of loading.
+\begin{alignat}{3}
+    \label{mod:c7}     v_j \geq  w_i + \beta_{k} + t_{ij} - M\left(1-x^{k}_{ij}\right) & \text{ }& \forall i \in \mathcal{D}, j \in \mathcal{L} \cup \{e_k\},   k \in K \\
+    \label{mod:c8}          w_{j} \geq v_{j}  + {  \frac{q^k_j}{\tau^u_c} -M\left(1- x^{k}_{ij}\right) } & \text{ }&  \forall c \in  \mathcal{C}, j \in \mathcal{D}_{c}, i \in \mathcal{L},  k \in K \\
+    \label{mod:c31}          w_{j} \leq v_{i}  + \Delta + M\left(1- x^{k}_{ij}\right)&  \text{ }&  \forall c \in  \mathcal{C}, j \in \mathcal{D}_{c}, i \in \mathcal{L},  k \in K
+    % \label{mod:c31}      w_{d-1}  \leq v_{d} & \hspace{5mm} & \forall i \in \mathcal{C}, o \in O_i, d \in \mathcal{D}_{o}, d \geq 1 \\
+    % \label{mod:c10}      z_{o,d} \leq z_{o,d-1} & \hspace{5mm} & \forall i \in \mathcal{C}, o \in O_i, d \in \mathcal{D}_{o}, d \geq 1\\
+    % \label{mod:c11}  v_{d} - w_{d-1} \leq \gamma_i   & \hspace{5mm}&  \forall i \in \mathcal{C}, o \in O_i, d \in \mathcal{D}_{o}, d \geq 1
+\end{alignat}
+
+The first service of any customer $i$ has to start at the due time $a_i$.
+Let $g_i$ be the gap between the due time and the start of first service at $i$.
+This first service may be performed at the first delivery node of any order $\in O_i$.
+Constraints (\ref{mod:c11}--\ref{mod:c13}) are used to find the delivery sequence of the first delivery node of all orders $\in \mathcal{O}_i$.
+
+\begin{alignat}{3}
+    \label{mod:c9}   v_{d^0_{o}} \geq a_i & & \forall  i \in \mathcal{C}, \forall o \in \mathcal{O}_i\\
+    \label{mod:c10}  g_{i} \geq v_{d} - a_i - M\left(  \sum_{ j \in \delta^{-}_{D}(d)  }{u_{j,d}}\right) & & \forall  i \in \mathcal{C}, \forall o_1 \in \mathcal{O}_i, d=d^0_{o_1} \\
+    \label{mod:c11}   \sum_{ \substack{o_1 \in \mathcal{O}_i }}{ \sum_{ j \in \delta^{-}_{D}(d^0_{o_1})}}{u_{j,d^0_{o_1} } } = |\mathcal{O}_i|-1 & \text{ }& \forall  i \in \mathcal{C},|\mathcal{O}_i| > 1 \\
+    \label{mod:c12}   \sum_{ \substack{o_1 \in \mathcal{O}_i }}\sum_{ j \in \delta^{+}_{D}(d^0_{o_1})}{u_{d^0_{o_1},j } } = |\mathcal{O}_i|-1 & \text{ }& \forall  i \in \mathcal{C}, |\mathcal{O}_i| > 1 \\
+    \label{mod:c13}  \hspace{3mm} \smashoperator{ \sum_{ j \in \delta^{+}_{D}(d^0_{o_1})}}{u_{d^0_{o_1},j } } \leq 1, \hspace{2mm} \smashoperator{ \sum_{ j \in \delta^{-}_{D}(d^0_{o_1})}}{u_{j,d^0_{o_1} } } \leq 1, \hspace{2mm} \smashoperator{ \sum_{ j \in \delta^{+}_{D}(d^0_{o_1})}}{u_{d^0_{o_1},j } } +\smashoperator{ \sum_{ j \in \delta^{-}_{D}(d^0_{o_1})}}{u_{j,d^0_{o_1} } } \geq1         & \text{ }& \forall  i \in \mathcal{C}, \forall o_1 \in \mathcal{O}_i
+\end{alignat}
+
+For a customer, an order must be completed before starting another.
+Constraints (\ref{mod:c14}--\ref{mod:c15}) enforce the precedence constraints between the last delivery of an order and the first delivery of the subsequent order.
+
+\begin{alignat}{3}
+  \label{mod:c14} & v_{d^0_{o_2}} \geq w_j - M(1- u_{j,d^0_{o_2}}) & \text{} & \forall i \in \mathcal{C}, \forall o_1, o_2 \in O_i, j \in \mathcal{D}_{o_1}, (j,d^0_{o_2}) \in \delta_{D}^{-}(d^0_{o_2})  \\
+  \label{mod:c15} &  v_{d^0_{o_2}}  \leq  w_j + \gamma_i + M(1- u_{j,d^0_{o_2}}) & \hspace{1mm} & \forall i \in \mathcal{C}, \forall o_1, o_2 \in O_i, j \in \mathcal{D}_{o_1}, (j,d^0_{o_2}) \in \delta_{D}^{-}(d^0_{o_2})
+%   \label{mod:c12} & v_{j} \geq w_d - M(1- u_{d,j}) & \text{} & \forall i \in \mathcal{C}, \forall o_1 \in O_i, d= d_{o_1,0}, j \in \mathcal{D}_{i}, j \neq  d\\
+%   \label{mod:c112} &  v_{j}  \leq  w_d + \gamma_i + M(1- u_{d,j}) & \text{ } & \forall i \in \mathcal{C}, j \in \mathcal{D}_{i}, o_1 \in O_i, d= d_{o_1,0},    j \neq  d\\
+%   \label{mod:c112} &  \smashoperator{\sum \limits_{j \in \mathcal{D}_i, j\neq d}}{u_{d,j}} \leq 1 & \text{ } & \forall i \in \mathcal{C}, o \in O_i,  d = d_{o,0} \\
+%   \label{mod:c113} &  \smashoperator{\sum \limits_{j \in \mathcal{D}_i, j\neq d}}{u_{j,d}} \leq 1 & \text{ } & \forall i \in \mathcal{C}, o \in O_i,  d = d_{o,0} \\
+%   \label{mod:c111} &  \smashoperator{\sum \limits_{j \in \mathcal{D}_i, j\neq d}}{u_{d,j}}+ \smashoperator{\sum_{j \in \mathcal{D}_i, j\neq d}}{u_{j,d}} \geq 1 & \text{ } & \forall i \in \mathcal{C}, o \in O_i,  d = d_{o,0}
+\end{alignat}
+
+Constraints (\ref{mod:c29}--\ref{mod:c30}) enforce the precedence constraints between two consecutive delivery nodes of the same order.
+
+\begin{alignat}{3}
+\label{mod:c29}   &   u_{j-1,j} \geq u_{j,j+1} & \hspace{5mm} & \forall i \in \mathcal{C}, \forall o \in O_i,  j \in \mathcal{D}_{o}, 1 \leq j \leq n_o-1\\
+  \label{mod:c30} &  u_{j-1,j} \geq \sum_{l \in \mathcal{L}}{x_{lj}} & \text{} & \forall i \in \mathcal{C}, \forall o \in O_i, j \in \mathcal{D}_{o}, j \geq 1 
+\end{alignat}
+
+Constraints (\ref{mod:c17}--\ref{mod:c18}) state that it is not possible for two trucks serving consecutive deliveries to be unloaded at the same time and there is a maximum time lag between them.
+
+\begin{alignat}{3}
+  \label{mod:c17} & v_{j} \geq w_{j-1} - M(1- u_{j-1,j}) & \text{} & \forall i \in \mathcal{C}, \forall o \in O_i, j \in \mathcal{D}_{o}, j \geq 1 \\
+  \label{mod:c18} &  v_{j}  \leq  w_{j-1} + \gamma_i + M(1- u_{j-1,j}) & \text{ } & \forall i \in \mathcal{C}, \forall o \in O_i, j \in \mathcal{D}_{o}, j \geq 1
+\end{alignat}
+
+With constraints (\ref{mod:c19}) and (\ref{mod:c20}), we respectively impose that the cumulative capacity of all concrete mixers serving an order and a customer may exceed the required quantities.
+
+\begin{alignat}{1}
+    \label{mod:c19} \sum_{d \in D_o}\sum_{k \in K}{q^k} \sum_{j \in \mathcal{L}}{x^k_{jd}}\geq q_o, \forall o \in \mathcal{O} \\
+    \label{mod:c20} \sum_{d \in D_i}\sum_{k \in K}{q^k} \sum_{j \in \mathcal{L}}{x^k_{jd}}\geq q_i, \forall i \in \mathcal{C}
+\end{alignat}
+
+Constraints (\ref{mod:c21}) and (\ref{mod:c22}) impose that a loading or delivery node may only be visited once by one driver. Constraints (\ref{mod:c23}) and (\ref{mod:c24}) are degree constraints.
+
+\begin{alignat}{3}
+\label{mod:c21} &  \sum_{k \in K}\sum_{j \in \mathcal{D}}{x^{k}_{ij} \leq 1} & \text{ } & i \in \mathcal{L}\\
+  \label{mod:c22}  &  \sum_{k \in K}\sum_{i \in \mathcal{L}}{x^{k}_{ij} \leq 1} & \text{ } & j \in \mathcal{D} \\
+  \label{mod:c23}  &  \sum_{k \in K}\sum_{i \in \mathcal{L}}{x^{k}_{ij} } =  \sum_{k \in K}\sum_{i \in \mathcal{L}}{x^{k}_{ji} } & \text{ } & j \in \mathcal{D} \\
+  \label{mod:c24}  &  \sum_{k \in K}\sum_{i \in \mathcal{D}}{x^{k}_{ij} } =  \sum_{k \in K}\sum_{i \in \mathcal{D}}{x^{k}_{ji} } & \text{ } & j \in \mathcal{L}
+\end{alignat}
+
+ Constraints (\ref{mod:c25}) and (\ref{mod:c26}) impose that all orders of customer $i$ must be loaded from the same plant.
+
+\begin{alignat}{2}
+    \label{mod:c25} & \sum_{k \in K} \sum_{l \in \mathcal{L}_b} \sum_{d \in \mathcal{D}}{x^k_{ld}} \leq M*\sigma_{i, b} & \forall i \in \mathcal{C}, b \in \mathcal{B} \\
+    \label{mod:c26} & \sum_{b \in \mathcal{B}} \sigma_{i, b} \leq 1 & \text{ } \forall i \in \mathcal{C}, b \in \mathcal{B} 
+\end{alignat}
+
+Let $w^1_k$ be the binary variables indicating if the driver $k$ has reached the minimum number of hours worked in the day.
+$w^2_k$ is a continuous variable indicating the gap between the driver working time and the normal working time.
+
+\begin{alignat}{3}
+    \label{mod:c27}  M_T*w^1_k \leq v_{e_k} - H_k & \hspace{5mm} &\forall k \in K \\
+    \label{mod:c28}  w^2_k \geq (v_{e_k} - H_k) - N_T & \hspace{5mm} &\forall k \in K.
+\end{alignat}
+
+The objective function (\ref{mod:obj}) minimizes the travelled distance, the penalty costs incurred when customers' demands are not fully satisfied, the lateness of each customer first delivery, the number of drivers who work under the minimum working time and the total overtime cost when drivers work beyond their scheduled hours.
+%The objective function (\ref{mod:obj}) is designed to minimize multiple factors that impact the overall cost and efficiency of the concrete delivery process.
+These factors include: The total traveled distance, which represents the fuel consumption and wear and tear on the delivery vehicles.
+The penalty costs $\beta_1$, which are incurred when customers' demands are not fully satisfied.The lateness of each customer's first delivery, which can impact customer satisfaction and lead to additional penalty costs.The number of drivers who work under the minimum working time, which can lead to increased labor costs and decreased productivity.The total overtime cost, which is incurred when drivers work beyond their scheduled hours.All these factors are taken into consideration in the objective function to minimize the cost and maximize the satisfaction of the customers.
+\begin{align}
+\label{mod:obj}
+\hspace{-20mm}
+    \begin{split}
+      \min  \smashoperator{\sum_{k \in K}} \sum_{ \substack{i \in \mathcal{L}} } \sum_{ j \in \delta^{+}{(i)} } {t_{ij}x^{k}_{ij}} + 
+      \sum_{k \in K} \sum_{ \substack{i \in \mathcal{D}}} \sum_{j \in \delta^{+}{(i)} } {t_{ij}x^{k}_{ij}} + 
+      \smashoperator{\sum_{ i \in C}}{ \beta_1 \left(1-y_i\right)} \hspace{-0mm} \\ +
+      \beta_2 \smashoperator{ \sum_{i \in \mathcal{C} } } { g_i  } 
+    %   + \beta_3 \smashoperator{\sum_{i \in \mathcal{C}}}{  {\sum_{d \in D_i}}{\left(v_d - w_{d-1} - \gamma_i\right)} } 
+      +  \sum_{k \in K}{\beta_3* w^{1}_k } + \beta_4 *{ w^{2}_k }
+    \end{split}
+\end{align}
+
+\iffalse
+graphe: G=(V,A)
+arc between origin depot, depot
+arc between depot delivery node
+arc between delivery node -> depot
+arc between delivery node -> origin depot
+
 Let $SLT^{bk}_{j}$ and $ELT^{bk}_{j}$ be the start and end loading time of driver $k$ at plant $b$.
@@
 
-% Parler de la reduction des W
-
-We keep track of the working time of each driver $k$ with the variable $WT_k$.
-Parameters $N_t$ and $O_t$ are the normal and overtime working hours.
-$M_t$ is the minimum working time for a day.
-$\Gamma_1$ and $\Gamma_2$ are the incurred penalties when a driver works less than $M_t$ and more than $N_t$.
-\begin{equation}
-    \forall k \in K, WT_k \leq O_t 
-\end{equation}
-
-The objective is to minimize the traveled distance,
- of delivered concrete to the satisfied customers as in
-
-\begin{equation}
-    \min \smashoperator{\sum_{i,j}}{t_{ij}x_{ij}}
-\end{equation}
-
-The cost to travel between two nodes ($i$, $j$) is the travel time $t_{ij}$. COST...
+\fi
 
 \paragraph*{}
-Let us consider the small instance illustrated in Figures~\ref{fig_Example} and \ref{fig:ganttExample}.
+Let us consider the solution of a small instance of our problem illustrated in Figures~\ref{fig_Example} and \ref{fig:ganttExample}.
 Two batching plants $B_1$ and $B_2$ with three drivers serve two construction sites.
@@
 
-We now present the solution approach we implement to solve this variant of the Concrete Delivery Problem.
+We now present the heuristic approach we implement to solve this variant of the Concrete Delivery Problem.
 Our method consists in constructing feasible solutions for the CDP with randomized heuristics and iteratively calling these heuristics in the greedy randomized adaptive search procedure (GRASP) metaheuristic.
 
-In this section, we represent the scheduling of a delivery node $d^p_{ij}$ as a pair of loading $l_{d^p_{ij}}$ and unloading $u_{d^p_{ij}}$ tasks.
+In this section, we represent the scheduling of a delivery node $d$ as a pair of loading $L_{d}$ and unloading $U_{d}$ tasks.
 We then refer to a solution $S$ as a set of loading and unloading tasks performed by a set of drivers during a particular day.
 \begin{alignat*}{2}
-    l_{d^p_{ij}} &= \left\lbrace (b,k,q^p_{ij},LTS^{bk}_{d^p_{ij}}), b \in \mathcal{B}, k \in K \right\rbrace \\
-    u_{d^p_{ij}} & =  (k,A_{{d^p_{ij}}}, UTS_{{d^p_{ij}}}), k \in K  \} \\
-    S &=\left\lbrace \cup _{1 \leq p \leq n^p_i} \cup _{1 \leq j \leq |D^p_i|} (l_{d^p_{ij}},u_{d^p_{ij}}), i \in \mathcal{C} \right\rbrace
+    L_{d} &= \left\lbrace (b,k,q_d,v_b), b \in \mathcal{B}, k \in K \right\rbrace \\
+    U_{d} & =  (k,v_d, w_d), k \in K  \} \\
+    S &=\left\lbrace \cup _{1 \leq j \leq n_o}  (L_{d^j_{o}},U_{d^j_{o}}), \forall o \in \mathcal{O} \right\rbrace
 \end{alignat*}
@@
 We create the set $CL'$ with elements of $CL$ we deem promising for a better solution.
-We evaluate the increase in the cost function when incorporating each $d \in CL'$ into the incumbent solution $s$ and create a restricted candidate list $RCL$ formed by nodes whose incremental costs are less than a defined threshold.
+We evaluate the increase in the cost function when incorporating each $d \in CL'$ into the incumbent solution $S$ and create a restricted candidate list $RCL$ formed by nodes whose incremental costs are less than a defined threshold.
 From $RCL$, we randomly select the next delivery node to be incorporated into the incumbent solution and we determine its loading and unloading schedule with the procedure described in Algorithm~\ref{alg_ScheduleTask}.
@@
 Then we add to $CL$ the next delivery node candidate.
-If an order $o^p$ is not yet fulfilled after the visiting $d^p_{ij}$, the candidate is the next delivery node $d^p_{ij+1}$.
-Otherwise, if the customer has remaining orders, the candidate is the first delivery node $d^{p'}_{i0}$ of another randomly selected order $p'$.
+If an order $o$ is not yet fulfilled after visiting $d^j_{o}$, the candidate is the next delivery node $d^{j+1}_{o}$.
+Otherwise, if the customer has remaining orders, the candidate is the first delivery node $d^{0}_{o'}$ of another randomly selected order $o'$.
 
@@
          \ForEach(customer i){}{
-                Select a random order $o^p_i$
+                Select a random order $o_1 \in \mathcal{O}_i$
 
-                $CL \leftarrow CL \cup \{d^p_{i0}\}$
+                $CL \leftarrow CL \cup \{d^0_{o_1}\}$
         }
@@
 
-            Select random delivery node $d^p_{ij}$ from $RCL$
+            Select random delivery node $d^j_o$ of customer $i$ from $RCL$
 
-            $l^p_{ij}$, $u^p_{ij}$ = ScheduleTasks($d^p_{ij},S$)
+            $L_{d^j_o}$, $U_{d^j_o}$ = ScheduleTasks($d^j_o,S$)
 
-            $S \leftarrow S \cup (l^p_{ij}$, $u^p_{ij})  $
+            $S \leftarrow S \cup (L_{d^j_o}$, $U_{d^j_o})  $
 
-            $q_i \leftarrow q_i-q^p_{ij}$;  $q^p_i \leftarrow q^p_i-q^p_{ij}$
+            $q_i \leftarrow q_i-q_{d^j_o}$;  $q_{o} \leftarrow q_{o}-q_{d^j_o}$
 
-            \uIf {$q^p_i \neq 0$ }{
+            \uIf {$q_{o} \neq 0$ }{
 
-            $CL \leftarrow CL \cup \{d^p_{ij+1}\}$
+            $CL \leftarrow CL \cup \{d^{j+1}_{o}\}$
 
@@
 
-            Select another order $o^{p'}_i$
+            Select another order $o' \in \mathcal{O}_i$
 
-            $CL \leftarrow CL \cup \{d^{p'}_{i0}\} $
+            $CL \leftarrow CL \cup \{d^{0}_{o'}\} $
 
@@
 
- Algorithm~\ref{alg_ScheduleTask} takes as parameters a delivery node $d^p_{ij}$ and a partial solution and looks for a plant and a driver available for loading and unloading operations.
+ Algorithm~\ref{alg_ScheduleTask} takes as parameters a delivery node $d^j_{o}$ of the customer $i$ and a partial solution and looks for a plant and a driver available for loading and unloading operations.
 For each driver $k$ and plant $b$ we simulate the loading and unloading operations while checking the concrete lifespan constraints, and the plant and driver availability.
-For each node, we first determine the start loading time $SLT$ by subtracting $t_{ij}$, $\alpha_b$, and the loading duration $LD^b_{d^p_{ij}}$ to the expected delivery time $EDT$.
+For each node, we first determine the start loading time $SLT$ by subtracting $t_{bi}$, $\alpha_b$, and the loading duration $LD^b_{d^j_{o}}$ to the expected delivery time $EDT$.
 $EDT$ is either the arrival time $a_i$ for the first delivery of a customer or the end of the unloading of the precedent delivery.
@@
 The presence of the driver does not guarantee that the loading dock is available, thus we use the procedure \textit{FindDepotLoadingSlot} at line $16$ to ensure that.
-Once we know $SLT$, we can easily deduce the arrival time, start unloading time, driver waiting time, and client waiting time.
+Once we know $SLT$, we can easily deduce the arrival time ($AT$), start unloading time ($SUT$), driver waiting time ($W^k_{d^j_o}$), and client waiting time ($W_{d^j_o}$).
 We also keep track of the driver's work duration.
 The algorithm returns the best loading and unloading operations with the least cost.
-
  We keep track of all timeslots for a plant when its loading dock is busy, and for a driver when he starts loading until the end of the unloading service.
 Within \textit{FindDepotLoadingSlot} (\textit{FindDriverLoadingSlot}), we iterate through the timeslots of a depot (driver) to schedule the current operation at a free timeslot.
-These allow us to insert the scheduling of a node before existing schedules.
+These allow us to insert the scheduling of a node between existing schedules.
 
@@
         % \vspace{-4mm}
-        \begin{algorithm}[htb]
+        \begin{algorithm}[H]
             \caption{Schedule loading and unloading tasks }
@@
             \setcounter{AlgoLine}{0}
-            \KwIn{Partial solution $S$, delivery node $d^p_{ij}$}
+            \KwIn{Partial solution $S$, delivery node $d^j_{o}$}
 
@@
             \SetKwProg{Fn}{Function}{}{}
-            \Fn{\FMain{$d^p_{ij},S$}}
+            \Fn{\FMain{$d^j_{o},S$}}
             {
@@
             % \SetAlgoNlRelativeSize{-1}
-            $n_i$: last node of $i$ visited
+            $n_i$: last node visited for the customer $i$
 
@@
 
-            $q^p_{ij}=min\{Q_k,q^p_i \}$ // $q^p_i$ \textit{remaining demand of order} $o^p_i$
+            $q^j_{o}=min\{Q_k,q_o \}$ // $q_o$ \textit{remaining demand of order} $o$
 
-            $SLT$ = $EDT - t_{bi} - \alpha_b - LD^b_{d^p_{ij}} $
+            $SLT$ = $EDT - t_{bi} - \alpha_b - LD^b_{d^j_{o}} $
 
-            $SLT = FindDriverLoadingSlot(b, n_k, SLT, LD^b_{d^p_{ij}})$
+            $SLT = FindDriverLoadingSlot(b, n_k, SLT, LD^b_{d^j_{o}})$
 
-            $SLT = FindDepotLoadingSlot(b, SLT, LD^b_{d^p_{ij}})$
+            $SLT = FindDepotLoadingSlot(b, SLT, LD^b_{d^j_{o}})$
 
-            $AT =  SLT + LD^b_{d^p{ij}} + \alpha_b + t_{bi}$
+            $AT =  SLT + LD^b_{d^j{o}} + \alpha_b + t_{bi}$
 
@@
 
-            $W^k_{d}= max\{0,SUT - AT\}$
+            $W^k_{d^j_o}= max\{0,SUT - AT\}$
 
-            $W_{d}= max\{0,AT - EDT - \lambda_i\}$
+            $W_{d^j_o}= max\{0,AT - EDT - \lambda_i\}$
 
@@
 
-            $WT = WT_k + (EUT + \beta_k - SLT) + t_{n_kb} + t_{j,k}$
+            $WT = WT_k + (EUT + \beta_k - SLT) + t_{n_kb} + t_{i,k}$
 
-            $TC_{k,j} = t_{n_k,b} + t_{b,j}  - t_{pos_k,k}$
+            $TC = t_{n_k,b} + t_{b,i}$  %- t_{pos_k,k}$
 
-            \uIf{$Cost(bestSol) < Cost(S) + TC_{k,j} $}{
-                $bestSol = (j,b,k,SLt) \cup (j,b,k,SUt) $
+            \uIf{$Cost(bestSol) < Cost(S) + TC $}{
+                $bestSol = (j,b,k,SLT) \cup (j,b,k,SUT) $
 
-                $Cost(bestSol) = Cost(S) + TC_{k,j}$
+                $Cost(bestSol) = Cost(S) + TC$
             }

### 2023-05-26 code/heur/AllInsertionOperator.h
Save

@@ -1,0 +1,80 @@
+
+
+#ifndef _ALL_INSERT_CDP_OPERATOR_
+#define _ALL_INSERT_CDP_OPERATOR_
+
+#include "../InsertOperator.h"
+#include "CustInsertion.h"
+#include "CustInsertionBacktrack.h"
+#include "PriorityQueueInsertion.h"
+
+/**
+ * k=0, sort by increasing early TW
+ */
+class AllInsertionOperator : public InsertOperator<Customer, Driver> {
+public:
+    AllInsertionOperator(CustInsertion *insert_opt, int k)
+            : cust_insert_opt(insert_opt), _k(k),
+              InsertOperator<Customer, Driver>("AllInsertionOperator"), cust_insert_bt_opt(nullptr),prio_insert_opt(
+                    nullptr) {
+        cust_insert_opt->SetK(_k);
+    }
+
+    AllInsertionOperator(CustInsertionBacktrack *insert_opt_bt, int k)
+            : cust_insert_bt_opt(insert_opt_bt), _k(k),
+              InsertOperator<Customer, Driver>("AllInsertionOperatorBacktrack"), cust_insert_opt(nullptr),prio_insert_opt(
+                    nullptr) {
+        cust_insert_bt_opt->SetK(_k);
+    }
+    AllInsertionOperator(PriorityQueueInsertion *insert_opt_prio, int k)
+            : prio_insert_opt(insert_opt_prio), _k(k),
+              InsertOperator<Customer, Driver>("AllInsertionOperatorPrioQueue"), cust_insert_opt(nullptr),cust_insert_bt_opt(
+                    nullptr) {
+        prio_insert_opt->SetK(_k);
+    }
+
+    AllInsertionOperator(CustInsertion *insert_opt, int k, std::string const &name)
+            : cust_insert_opt(insert_opt), _k(k), InsertOperator<Customer, Driver>(name),prio_insert_opt(nullptr) ,cust_insert_bt_opt(nullptr){
+        cust_insert_opt->SetK(_k);
+        cust_insert_opt->name = name;
+    }
+
+    AllInsertionOperator(CustInsertionBacktrack *insert_opt_bt, int k, std::string const &name)
+            : cust_insert_bt_opt(insert_opt_bt), _k(k), cust_insert_opt(nullptr), InsertOperator<Customer, Driver>(name),prio_insert_opt(nullptr)  {
+        cust_insert_bt_opt->SetK(_k);
+        cust_insert_bt_opt->name = name;
+    }
[…]

### 2023-05-26 code/heur/CustInsertion.h
Save

@@ -13,11 +13,9 @@ class CustInsertion : public InsertOperator<Customer, Driver> {
 
 public:
     CustInsertion(Data &prob, InsRmvBuilder &insrmv)
-            : _data(prob), _insrmv(insrmv),  customersListId(0),
-              customersList(0),  _k(0),removedList(0),removedListId(0),
+            : _data(prob), _insrmv(insrmv),
+              customersList(0),  _k(0),removedList(0),
               name("CustInsertion"),listMoves(0) {
-        customersListId = std::vector<int>();
-        removedListId = std::vector<int>();
         customersList = std::vector<Customer *>();
         removedList = std::vector<Customer *>();
         listMoves = std::vector<ListMoveVrp>();
@@ -33,10 +31,11 @@ public:
     std::string name;
     std::vector<ListMoveVrp>  listMoves;
     void Insert(Sol &s, std::vector<int> const &list);
-    void Insert2(Sol &s, std::vector<Customer *> &list,std::vector<int> &list_ID);
 private:
-    void Insert(Sol &s,std::vector<Customer *> &list, std::vector<int> &list_ID);
-    void Sort(Sol &s,std::vector<Customer *> &list, std::vector<int> &list_ID, int k);
+    void InsertWithBactrack(Sol &s,std::vector<Customer *> &list);
+    static void Sort(Sol &s,std::vector<Customer *> &list, int k);
+    static void Sort(Sol &s,std::vector<int> &list_ID, int k);
+    void Insert2(Sol &s, std::vector<Customer *> &list,std::vector<int> &list_ID);
     void Init()
     {
         for (int i = 0; i < _data.GetDepotCount(); i++)
@@ -51,8 +50,6 @@ private:
     Data &_data;
     std::vector<Depot *> depotList;
     std::vector<Driver *> driversList;
-    std::vector<int> customersListId;
-    std::vector<int> removedListId;
     std::vector<Customer*> customersList;
     std::vector<Customer *> removedList;
 };

### 2023-05-26 code/Solution.h
Save

@@ -2,6 +2,7 @@
 
 #ifndef CODE_SOLUTION_H
 #define CODE_SOLUTION_H
+
 #include "Data.h"
 #include "Cost.h"
 #include "TimeSlot.h"
@@ -14,12 +15,12 @@
 #include <iostream>
 
 
-
-
 class Sol {
 public:
-    Sol()=default;
+    Sol() = default;
+
     explicit Sol(Data *data);
+
     bool isFeasible{false};
     int iter{-1};
     double duration_time{0};
@@ -37,19 +38,23 @@ public:
     std::vector<int> RoutesLength;
     std::vector<int> orderCapRestante;
     std::vector<int> clientCapRestante;
-    std::vector<int>  shiftDuration;
-    std::vector<int>  shiftDurationCost;
-    std::vector<int>  DepotSize;
-    std::vector<int>  ArrivalTime;
-    std::vector<int>  StartServiceTime;
-    std::vector<int>  EndServiceTime;
-    std::vector<int>  ExpectedServiceTime;
-    std::vector<int>  DepartureTime;
-    std::vector<int>  DeliveryLoad;
-    std::vector<int>  WaitingTime;
-    std::vector<int>  UnassignedIndex;
-    std::vector<int>  OrderVisitCount;
-    std::vector<int>  sumServiceTime;
+    std::vector<int> shiftDuration;
+    std::vector<int> shiftDurationCost;
+    std::vector<int> DepotSize;
+    std::vector<int> ArrivalTime;
+    std::vector<int> StartServiceTime;
+    std::vector<int> EndServiceTime;
+    std::vector<int> ExpectedServiceTime;
+    std::vector<int> DepartureTime;
+    std::vector<int> DeliveryLoad;
+    std::vector<int> WaitingTime;
+    std::vector<int> ClientWaitingTime;
+    std::vector<int> TruckWaitingTime;
+    std::vector<int> OrderLateDelivery;
+    std::vector<int> NodeLateDelivery;
+    std::vector<int> UnassignedIndex;
+    std::vector<int> OrderVisitCount;
+    std::vector<int> sumServiceTime;
     std::vector<bool> VisitFlags;
     std::vector<bool> VisitFlagCost;
     std::vector<std::vector<int>> DriverVisitCount;
@@ -65,6 +70,7 @@ public:
 
 
     static std::vector<int> FailureCause;
+    static std::vector<int> CustomerConflict;
     static std::vector<std::set<int>> TabuFleet;
[…]

### 2023-05-26 code/Data.cpp
Save

@@ -373,6 +373,9 @@ void Data::AddDock(const Dock &n){
 Depot *Data::GetDepot(int i) {
     return dynamic_pointer_cast<Depot>( _nodes[_depots[i]]).get();
 }
+Depot *Data::GetDepot(Node *n ) {
+    return GetDepot(n->depotID);
+}
 
 Customer * Data::GetCustomer(int i) const {
     return dynamic_pointer_cast<Customer>( _nodes[_customers[i]]).get();
@@ -474,7 +477,6 @@ void Data::AddDockNodes(){
         AddDock(d1);
     }
 }
-
 
 void Data::LoadMatrices(std::vector<std::vector<double>> &array, const string &matrix_filename,const int rate)
 {

### 2023-05-26 code/Data.h

@@ -58,6 +58,7 @@ public:
     void AddDriver(Driver &d);
     Driver * GetDriver(int id);
     Depot * GetDepot(int i);
+    Depot * GetDepot(Node *n);
     Node * GetNode(int i){ return _nodes[i].get();}
     Customer * GetCustomer(int i) const;
     Order * GetOrder(Customer *c, int index);
@@ -94,6 +95,12 @@ public:
             return d->serviceDuration;
 
         return std::ceil(double(60 * demand) / Parameters::UNLOADING_RATE);
+    }
+    static int UnloadingTime( int demand, int serviceDuration) {
+        if(Parameters::KINABLE)
+            return serviceDuration;
+
+        return std::ceil(double(demand) / Parameters::UNLOADING_RATE);
     }
     static int CleaningTime( Delivery *del,Driver *d) {
         return Parameters::CLEANING_DURATION;
@@ -259,7 +266,7 @@ private:
     std::vector<std::vector<int>> _index_orders;
     std::vector<std::shared_ptr<Node>> _nodes;
     std::string time_mat_file, distance_mat_file;
-    static void LoadMatrices(std::vector<std::vector<double>> &array, const std::string &matrix_filename,const int rate);
+    static void LoadMatrices(std::vector<std::vector<double>> &array, const std::string &matrix_filename,int rate);
 
     std::vector<std::vector<double>> _distances;
     std::vector<std::vector<double>> _times;

### 2023-05-26 code/Solution.cpp
Save

@@ -1,9 +1,11 @@
 
 #include "Solution.h"
 #include <fstream>
+
 using namespace std;
 
 std::vector<int> Sol::FailureCause = std::vector<int>();
+std::vector<int> Sol::CustomerConflict = std::vector<int>();
 std::vector<int> Sol::FailureCount = std::vector<int>();
 std::vector<int> Sol::minDelay = std::vector<int>();
 std::vector<int> Sol::StartBefore = std::vector<int>();
@@ -11,42 +13,46 @@ std::vector<int> Sol::pullVisit = std::vector<int>();
 std::vector<int> Sol::pushVisit = std::vector<int>();
 std::vector<std::set<int>> Sol::TabuFleet = std::vector<std::set<int>>();
 std::map<std::tuple<int, int>, int> Sol::nodeMaxStartService =
-        std::map<std::tuple<int, int>, int>();
+        std::map < std::tuple<int, int>,
+int>();
 
 
-Sol::Sol(Data *data): _data(data), _last_cost(false), DriverAssignTo(data->GetNodeCount(),nullptr),
-RoutesLength(data->GetDriverCount(), 0),
-                      orderCapRestante(data->GetOrderCount(),0),
-                      shiftDuration(data->GetDriverCount(), 0),
-                      shiftDurationCost(data->GetDriverCount(), 0),
-                      DriverVisitCount(data->GetDriverCount()),
-                      DepotSize(data->GetDepotCount(), 0),
-                      EndServiceTime(data->GetNodeCount(),0),
-                      ExpectedServiceTime(data->GetNodeCount(),0),
-                      StartServiceTime(data->GetNodeCount(),0),
-                      clientCapRestante(data->GetCustomerCount(),0),
-                      driverWorkingIntervals(data->GetDriverCount()),
-                      nodeServiceIntervals(data->GetNodeCount()),
-                      depotLoadingIntervals(data->GetDepotCount()),
-                      ArrivalTime(data->GetNodeCount(),0),
-                      DepartureTime(data->GetNodeCount(),0),
-                      DeliveryLoad(data->GetNodeCount(),0), WaitingTime(data->GetNodeCount(),0),
-                      Unassigned(data->GetNodeCount(),nullptr),
-                      VisitFlags(data->GetNodeCount(),false),
[…]

### 2023-05-26 code/UpdateCost.cpp
Save

@@ -2,6 +2,7 @@
 #include <iostream>
 
 using namespace std;
+
 void Sol::Update() {
 
     updateCost = Cost();
@@ -10,47 +11,18 @@ void Sol::Update() {
     driverUsed.clear();
     BuildFromDepotSetIntervall();
     isFeasible = true;
-    for (int i = 0; i < GetDriverCount(); i++)
-    {
-        Driver *d = GetDriver(i);
-        d->Init();
-        shiftDuration[d->id] = 0;
-        driverWorkingIntervals[d->id].clear();
-        std::fill(DriverVisitCount[i].begin(), DriverVisitCount[i].end(), 0);
-    }
-    for (int i = 0; i < GetCustomerCount(); i++)
-    {
-        Customer *c = GetCustomer(i);
-        clientCapRestante[c->custID] = c->demand;
-        clientDriverUsed[c->custID].clear();
-        OrderVisitCount[c->custID] = 0;
-        sumServiceTime[c->custID]=0;
-    }
-    for (int i = 0; i < GetOrderCount(); i++)
-    {
-        Order *o = GetOrder(i);
-        orderCapRestante[o->orderID] = o->demand;
-        Delivery *del = GetDelivery(o, 0);
-        ExpectedServiceTime[del->id] = EarlyTW(del);
-        //TODO del->expected_del_time = c->early_tw;
-    }
 
-    for (int i = 0; i < GetDepotCount(); i++)
-    {
-        Depot *dep = GetDepot(i);
-        depotLoadingIntervals[dep->depotID].clear();
-    }
+    InitDrivers();
+    InitCustomers();
+    InitDepots();
+
     nodeServiceIntervals =
             std::vector<TimeSlot>(GetNodeCount(), TimeSlot(0, 1440));
     UpdateForward();
 
-    for (int i = 0; i < GetDriverCount(); i++)
-    {
-        Driver *d = GetDriver(i);
-        if (RoutesLength[d->id] < EPS)
-            continue;
+    for (auto id: driverUsed) {
+        Driver *d = GetDriver(id);
         Node *lastOfDriver = DriverPrev[d->EndNodeID];
-
         shiftDuration[d->id] +=
                 Travel(lastOfDriver->distID, d->distID);
         d->shiftDuration = shiftDuration[d->id] - d->start_shift_time;
@@ -61,34 +33,28 @@ void Sol::Update() {
[…]

### 2023-05-26 code/TimeSlot.cpp
Save

@@ -80,6 +80,6 @@ bool TimeSlot::operator!=(const TimeSlot &rhs) const {
 }
 
 std::ostream &operator<<(std::ostream &os, const TimeSlot &slot) {
-    os <<  slot.typeNoeud << ":[" << slot.lower << ":" << slot.upper << "] "<< ":"<< slot.nodeID;
+    os <<  slot.typeNoeud <<slot.nodeID <<  ":[" << slot.lower << ":" << slot.upper << "]-";
     return os;
 }

### 2023-05-26 code/heur/Solver.h
Save

@@ -8,6 +8,7 @@
 #include "../Combinatorial.h"
 #include "../InsertOperator.h"
 #include "../BestSolutionList.h"
+#include "AllInsertionOperator.h"
 #include "RechercheLocale.h"
 #include "../GRASP.h"
 #include <vector>
@@ -18,15 +19,7 @@
 class Solver {
 public:
     explicit Solver(Data *data):_data(data){
-        Sol::FailureCause.resize(_data->GetNodeCount(), Parameters::FAILURECAUSE::NONE);
-        Sol::FailureCount.resize(_data->GetNodeCount(), 0);
-        Sol::StartBefore.resize(_data->GetNodeCount(), 0);
-        Sol::pullVisit.resize(_data->GetNodeCount(), 0);
-        Sol::pushVisit.resize(_data->GetNodeCount(), 0);
-        Sol::TabuFleet.resize(_data->GetNodeCount());
         CDPSolver::disjointClients.resize(_data->GetCustomerCount(), std::set < int > ());
-        Sol::minDelay.resize(_data->GetNodeCount(), 0);
-        TimeSlot::myData = *_data;
     }
     void run();
     static void SolveGrasp(Sol &s,Data &dat,std::vector<std::set<int>> const & linkedClientSlot);
@@ -41,7 +34,10 @@ private:
     void run2(Sol &s,const std::set<int>& feasible);
     void TestRecursive(Sol &cur, Delivery *del,int cap);
     void TestRecursive2(Sol &cur, Delivery *del,std::unordered_map<std::string, Sol, MyHashFunction> * umap);
+    void Test0();
+    void SaveResults(Sol &s);
     Data *_data;
+
 };
 
 

### 2023-05-26 Article/article.tex
correct model

@@
 \newcommand{\pcmaxc}{P$|\mbox{\em cont}|$C$_{\max}$\xspace}
-\newcommand{\alns}{adaptive large neighborhood search\xspace}
-\newcommand{\Alns}{Adaptive large neighborhood search\xspace}
+
+\newcommand{\GRASP}{Greedy randomized adaptive search procedure\xspace}
 \newcommand{\F}{$\mathcal{F}$\xspace}
@@
 \newcommand{\C}{$\mathcal{C}$\xspace}
-\newcommand{\Lb}{$\mathcal{L}$\xspace}
+\newcommand{\LB}{$\mathcal{L}$\xspace}
+\newcommand{\Lb}{$\mathcal{L}_b$\xspace}
 \newcommand{\D}{$\mathcal{D}$\xspace}
+ \newcommand*{\red}{\textcolor{red}}
 
@@
 Concrete is a widely used building material in construction projects.
-It is a perishable product that can be affected by many different factors that impact its quality \citep{sinha_quality_2021} which is crucial for the durability and strength of the final construction.
-
+Its perishable nature is affected by many factors that impact its quality \citep{sinha_quality_2021}, which is crucial for the durability and strength of the final construction.
 Concrete comes in two types: ready-mixed concrete (RMC) and site-mixed concrete (SMC).
 RMC is manufactured in a batch plant and delivered to the construction site, while SMC is produced on-site using raw materials stored on the construction site.
-Site-mixed concrete can avoid delays caused by road traffic, but it has a slower and more difficult production process, requires storage for mixing materials and equipment, and is suitable for low amounts of concrete.
+Using SMC can avoid delays caused by road traffic, but it has a slower and more difficult production process, requires storage for mixing materials and equipment, and is suitable for low amounts of concrete.
 On the other hand, RMC has better quality and benefits from lower production costs \citep{muresan_comparing}.
-However, to take advantage of these benefits, the batch plant manager must ensure efficient and prompt delivery on the construction site, which may require a fleet of high-cost revolving drum trucks (concrete mixers) to dispatch the ready-mixed concrete.
+However, to take advantage of these benefits, the batch plant manager must ensure efficient and prompt delivery on the construction site, which may require a fleet of high-cost revolving drum trucks (concrete mixers) to dispatch the RMC.
 
-Concrete delivery under the form of RMC is subject to many operational constraints that make it a challenging problem met in Operations Research and addressed by the Concrete Delivery Problem (CDP).
-In this paper, we study a variant of the Concrete Delivery Problem to schedule the daily production and dispatching of RMC for a company located in the province of Quebec, Canada.
-This company operates multiple batching plants with varying production rates, using a fleet of concrete mixers with different capacities.
-Each plant has its own fleet of trucks, however, the trucks are able to move between plants if necessary.
+Concrete delivery under the form of RMC is subject to many operational constraints that make the Concrete Delivery Problem (CDP) very challenging.
+In this paper, we study a variant of the CDP to schedule the daily production and dispatching of RMC for a company located in the province of Quebec, Canada.
+This company operates multiple batching plants with varying production rates, using a fleet of concrete mixers of different capacities.
+Each plant has its own fleet of trucks; however, under certain conditions, the trucks can move between plants if necessary.
 The trucks must return to their home plant at the end of the day.
-The company owns two types of trucks with different capacities but also has the option to call on an external fleet when needed.
-They serve construction sites from any of their batch plants, with the first delivery starting at the time specified by the customer.
+The company owns two types of trucks with different capacities and can call on an external fleet when needed.
+They serve construction sites from any of their plants, with the first delivery starting at the time specified by the customer.
 The loading and unloading of a concrete mixer depend on the truck capacity, the loading rate at the plant, and the unloading rate at the construction site.
@@
 This constraint generalizes the linked order constraints of \cite{durbin2008or} where some customers place two orders for the same day and request that they are linked (the second order begins only after the first order is completed).
-The setting of this study is similar to a variant of the Concrete Delivery Problem previously studied in \cite{schmid2009hybrid, schmid2010hybridization}.
-However, it includes the plant's production rate and driver shift schedule and does not consider unloading instrumentation in contrast to the previous research.
+The setting of our study is similar to a variant of the CDP previously studied in \cite{schmid2009hybrid, schmid2010hybridization}.
+However, we include the plant's production rate and driver shift schedule.
 The company uses a centralized dispatcher system to schedule all daily orders, but this system has issues satisfying all daily demands without using an external fleet.
@@
 
-According to \cite{blazewicz2019handbook}, the concrete delivery problem combines vehicle routing with scheduling issues to plan routes to deliver concrete from batch plants (depots) to customers' construction sites.
-Ready-mixed concrete is an on-demand product with a short life cycle from production through end use.
-It cannot be stored and cannot stay too long in a truck, or it will harden.
+According to \cite{blazewicz2019handbook}, the CDP combines vehicle routing with scheduling issues to plan routes to deliver concrete from batch plants (depots) to customers' construction sites.
+RMC is an on-demand product with a short life cycle from production through end use.
+It cannot be stored and cannot stay too long on a truck, or it will harden.
 Hence, concrete mixers must deliver RMC at the planned construction site shortly after its production.
-RMC production and delivery is described by \cite{tommelein1999just} as an example of a Just-In-Time (JIT) production system in construction.
+\cite{tommelein1999just} describes RMC production and delivery as an example of a just-in-time (JIT) production system in construction.
 A customer quantity requirement is often greater than the truck size and must be fulfilled by multiple deliveries.
-In that sense, CDP is like the Vehicle Routing Problem (VRP) with split delivery (VRPSD) \citep{archetti2008split}, except that the same truck may visit a customer more than once.
-Concrete hardens quickly, so when there are multiple deliveries, they must be done back-to-back or at least close in time to avoid the problem of cold joint, which can reduce the strength and durability of the concrete.
+In that sense, CDP is similar to the vehicle routing problem with split delivery \citep{archetti2008split}, except that the same truck may visit a customer more than once.
+Concrete hardens quickly, so multiple deliveries must be done back-to-back or at least close in time to avoid the problem of cold joint, which can reduce the strength and durability of the concrete.
 Customers request to be served within a specific time window, which can complicate truck-loading schedules when a plant can only load one truck at a time.
 Similarly, only one truck can unload at a time at a customer location, sometimes leading to concrete mixers queuing and waiting their turn to deliver.
-Furthermore, with trucks of varied sizes, loading, travel, and unloading times that may be uncertain, the Concrete Delivery Problem is a complex problem to solve and has been proven to be NP-hard \citep{asbach2009analysis,kinable2014concrete}.
+Furthermore, with trucks of varied sizes, loading, travel, and unloading times that may be uncertain, the CDP is a complex and challenging problem.
 
-In this paper, we propose a mathematical model and a Greedy Randomized Search Procedure (GRASP) heuristic solution to solve a new variant for the CDP.
+In this paper, we propose a mathematical model and a Greedy Randomized Search Procedure (GRASP) heuristic solution to solve a new variant of the CDP.
 Our model takes into consideration the working shifts of the drivers and the scheduling of multiple orders within the same time window at a construction site.
@@
 Section \ref{desc_form} provides a formal description and mathematical model of the problem.
-In Section \ref{method}, we describe the Greedy Randomized Search Procedure algorithm and the constructive heuristics developed to solve this variant of the CDP.
+In Section \ref{grasp_method}, we describe the GRASP algorithm and the constructive heuristics developed to solve this variant of the CDP.
 Section \ref{comp_exp}  presents computational experiments to evaluate the proposed approach, and finally, the conclusions are presented in Section \ref{concl}.
@@
 
-A text mining approach by \cite{maghrebi2015text} for reviewing the ready-mixed concrete literature showed that concrete technology and material science are the main core of research in this area.
-Academic research on concrete batching and delivery began in the late 1990s.
-\cite{tommelein1999just} described ready-mixed concrete (RMC) as a prototypical example of a Just-In-Time production system in construction and identified two practices for delivering it.
-One approach is for the customer to haul the product from the batch plant with their concrete mixer, while the other approach is for the batch plant to deliver the concrete directly to the customer's location.
+Academic Research on concrete batching and delivery began in the late 1990s.
+\cite{tommelein1999just} described RMC as a prototypical example of a JIT production system in construction and identified two practices for delivering it.
+One approach is for the customer to haul the product from the batch plant with their concrete mixer, while the other is for the batch plant to deliver the concrete directly to the customer's location.
 This latter approach is the one that has been studied in all related papers found in the literature.
-
 Several works to schedule and dispatch concrete production and delivery have mainly focused on simulation-related methods.
 These methods can be standalone, such as those used by \cite{zayed2001simulation, wang2001scheduling, tian_simulation_based_2010, panas_simulation_based_2013} and \cite{galic2016simulation}, among others.
-Alternatively, these methods can be hybridized with optimization techniques, such as those used by \cite{feng2004optimizing, lu2005optimized, feng_integrating_2006}, etc.
+Alternatively, these methods can be hybridized with optimization techniques, such as those used by \cite{feng2004optimizing, lu2005optimized}, and \cite{feng_integrating_2006}.
 \cite{wang2001scheduling} developed a simulation model to reveal the effect and value of the concrete mixers' inter-arrival time on the productivity of hired unloading equipment on site.
-\cite{feng2004optimizing} used a combination of Genetic Algorithm (GA) and simulation process to minimize the total waiting time for trucks at a customer site.
+\cite{feng2004optimizing} used a combination of genetic algorithm (GA) and simulation process to minimize the total waiting time for trucks at a customer site.
 The study focused on loading trucks with identical capacities at the same batch plant, with fixed loading and unloading durations.
@@
 As such, the simulation model also included the batching and delivery of mortar.
-Finding the best RMC truck size was also the purpose of the discrete-event simulation model proposed by \cite{panas_simulation_based_2013}.
+Finding the best RMC fleet size was also the purpose of the discrete-event simulation model proposed by \cite{panas_simulation_based_2013}.
 
 In addition to simulation-based methods, several other approaches have been used in the literature to solve the CDP.
-These include metaheuristics \citep{faria2006distributed, misir2011selection, maghrebi2016sequential, yang2022concrete}, exact methods \citep{yan2007optimal, asbach2009analysis, kinable2014concrete}, matheuristics \citep{schmid2009hybrid, schmid2010hybridization}, Benders Decomposition \citep{maghrebi2014benders}, Column Generation (CG) \citep{maghrebi2014solving, maghrebi2016column}, Lagrangian relaxation \citep{narayanan2015using}, and machine learning approaches such as those used by \cite{graham2006modeling, maghrebi2014exploring, maghrebi2016matching}.
+These include metaheuristics \citep{faria2006distributed, misir2011selection, maghrebi2016sequential, yang2022concrete}, exact methods \citep{yan2007optimal, asbach2009analysis, kinable2014concrete}, matheuristics \citep{schmid2009hybrid, schmid2010hybridization}, Benders Decomposition \citep{maghrebi2014benders}, column generation (CG) \citep{maghrebi2014solving, maghrebi2016column}, Lagrangian relaxation \citep{narayanan2015using}, and machine learning approaches such as those used by \cite{graham2006modeling, maghrebi2014exploring, maghrebi2016matching}.
 \cite{matsatsinis2004towards} designed a decision support system (DSS) for the dynamic routing of both concrete and pumps that may be necessary for some construction sites to aid in the unloading of concrete.
@@
 Orders that could not be executed immediately could be postponed for the next day.
-The routing of the pumps was modelled as a multi-depot vehicle routing problem with time windows.
+The routing of the pumps was modeled as a multi-depot vehicle routing problem with time windows.
 \cite{naso2007genetic} proposed a sequential GA method combined with constructive heuristics to solve another variant of the CDP.
@@
 The authors also developed a non-linear model that minimizes transportation costs, waiting times, outsourced costs, and overtime work.
-They ran experiments using real-world instances of a concrete supply chain in the Netherlands and found a reduction in the number of requests redirected to external companies.
+They ran experiments using real-world instances of a concrete supply chain in the Netherlands and found a reduction in the number of outsourced requests.
 \cite{yan2007optimal} also considered overtime considerations in their paper, which focused on scheduling RMC for one batching plant with two loading docks.
 The study took into account that overtime wages are paid for factory and construction site operations after 4 PM.
-They developed a mixed integer programming (MIP) model on a time-space network to minimize travel times and operating costs at both normal and overtime working hours at the plant and the construction sites.
-They tested the model using real-data consisting of 3 days of operation using a two-stage algorithm.
+They developed a mixed-integer programming (MIP) model on a time-space network to minimize travel times and operating costs at both normal and overtime working hours at the plant and the construction sites.
+They tested the model using real data consisting of three days of operation using a two-stage algorithm.
 First, they solved the MIP relaxation with CPLEX.
@@
 The algorithm was found to improve the actual plant operation by 10\%.
-A time-space network is the key component of the real-time decision support system (DSS) developed by \cite{durbin2008or} to solve a dynamic CDP problem every five minutes.
+A time-space network is the key component of the real-time DSS developed by \cite{durbin2008or} to solve a dynamic CDP every five minutes.
 The DSS is able to receive new orders, schedule them on the fly, and handle unexpected events such as plant closures, truck breakdowns, and delays in transportation times.
-The authors combined the DSS with a Tabu Search (TS) heuristic to warm start CPLEX, which made the model performant enough to solve instances with up to 1,500 loads per day with up to 250 trucks.
+The authors combined the DSS with a tabu search (TS) heuristic to warm start CPLEX, which made the model performant enough to solve instances with up to 1,500 loads per day with up to 250 trucks.
 The DSS also considers the case of a customer who places two orders, with the first being completed before the second starts.
@@
 The method can quickly solve large problem instances with more than 60 orders per day without encountering any memory issues.
-The same problem is addressed by \cite{schmid2010hybridization} who proposed a mixed integer programming (MIP) model combined with a VNS and a very large neighborhood search (VLNS) to develop two matheuristics approaches.
+The same problem is addressed by \cite{schmid2010hybridization}, who proposed a MIP model combined with a VNS and a very large neighborhood search (VLNS) to develop two matheuristics approaches.
 Comparisons between both matheuristics and a standalone VNS show that the former methods are much better and suitable for solving larger problem instances.
@@
 One of the paper's main contributions is the creation of the first public test instances for the CDP with up to 50 customers, four batching plants, and 20 concrete mixers.
-They found the CP model to be highly effective in finding high-quality solutions in relatively little time or improving existing schedules, while the MIP model can be used to compute bounds, as it seems ineffective in solving large problem instances.
+They found the CP model to be highly effective in finding high-quality solutions in a relatively short time or improving existing schedules, while the MIP model can be used to compute bounds, as it seems ineffective in solving large problem instances.
 Finally, the heuristic often yields good solutions in less than a second.
-A detailed analysis of the MIP model presented in cite{kinable2014concrete}and of two more compact models can be found in the thesis of \cite{hernandez_lopez_study_2020}.
-In \cite{kinable2014logic}, we found an attempt to solve the previous problem with a Logic Based Benders' approach.
-\cite{sulaman2017simulated} expand upon the SD heuristic proposed in \cite{kinable2014concrete}, proposing a Simulated Annealing (SA) with a Time-Slot Heuristic (TH).
-TH mechanism is to look for a slot between existing visits of a truck to schedule a new delivery instead of assigning it to the time slot strictly after the truck's latest assigned delivery.
+A detailed analysis of the MIP model presented in \cite{kinable2014concrete} and of two more compact models can be found in the thesis of \cite{hernandez_lopez_study_2020}.
+In \cite{kinable2014logic}, we found an attempt to solve the previous problem with a logic-based Benders' approach.
+\cite{sulaman2017simulated} expand upon the SD heuristic proposed in \cite{kinable2014concrete}, proposing a simulated annealing (SA) combined with a time-slot Heuristic (SATH).
+This method looks for a slot between existing visits of a truck to schedule a new delivery instead of assigning it to the time slot strictly after the truck's latest assigned delivery.
 The goal is to reduce the large time gaps that can be present in a schedule created with SD due to ignoring the intermediate available time slots.
@@
 The MIP model is used in a local search scheme as a black-box solver to reoptimize an incumbent solution in which a neighborhood operator has unfixed some variables.
-\cite{tzanetos2023systematic} provides an overview of the various methods used in the literature to address the Concrete Delivery Problem and categorizes the problem formulations based on the different concepts used in the literature, such as Integer Programming (IP), time-space network, and job shop model.
+\cite{tzanetos2023systematic} provide an overview of the various methods used in the literature to address the CDP and categorizes the problem formulations based on the different concepts used in the literature.
 They also discussed the consistency between industry needs and existing constraints and provided insights into the datasets corresponding to real-world cases, identifying the necessary data for practitioners.
@@
 \label{desc_form}
-The focus of this paper is the distribution of ready-mixed concrete from a Canadian company that operates in the greater Montreal area.
-When a customer places an order, it is received at a central center and then assigned to one of the company's batching plants.
+The focus of this paper is on the distribution of RMC from a Canadian company that operates in the greater Montreal area.
+When a customer places an order, it is received at a control center and assigned to one of the company's batching plants.
 These plants produce the concrete and then deliver it to the customer.
-The problem we are examining involves a set of batching plants, a set of customer orders, and a set of concrete-mixer drivers.
-
-% \subsection*{Batching plants}
-
-The company owns a total of eight batching plants, located in various geographical regions.
-Each plant has a loading bay that can only accommodate one truck at a time, which leads to trucks lining up.
-Let $\mathcal{B}$ be the set of batching plants.
-The plants are heterogeneous, as each plant has its own hourly loading rate, represented by $\tau^l_b$, which affects the duration of the loading process.
-After the concrete has been loaded, the driver spends $\alpha_b$ minutes adjusting the concrete in the truck before heading to the customer site.
-A batching plant can serve a construction site as long as the travel time between the two is less than the concrete's lifespan, which is represented by $\Delta$.
-Each plant has its own assigned fleet of trucks, but it can also borrow trucks from other plants or hire external fleets if necessary.
+The problem we are examining involves a set of customer orders, a set of concrete-mixer drivers, and a set of batching plants.
 
@@
 
-A customer $i$ requests a total of $q_i$ of one or several types of concrete to be delivered to his construction site on a particular day.
-We call an order a request for a specific type of concrete.
-When placing an order $o$, the customer $i$ specifies the required quantity, $q_o$, the desired arrival time of the first concrete mixer, $a_i$, and the unloading rate at their construction site, $\tau^u_i$.
+A customer $i$ requests one or more types of concrete to be delivered to their  construction site on a particular day, starting at time $a_i$.
+We call an order $o$ a request for a specific type of concrete.
+$q_i$ is the sum of the demands $q_o$ of each order $o$ placed, $a_i$ is the desired arrival time of the first concrete mixer, and $\tau^u_i$ is the unloading rate for customer $i$.
 If the order requires more concrete than a single truck can carry, multiple deliveries are scheduled.
-Let $O_i$ be the unordered set of all orders requested by customer $i$.
-Each element of $O_i$ must be thoroughly delivered before moving on to another order.
-One element $o$ of $O_i$ must have its first delivery start at $a_i$, while the others can start any time after all deliveries of $o$.
-Once a plant is chosen to produce a customer's first delivery, it must be the provider of all subsequent deliveries of this customer.
-To prevent cold joint issues with the concrete, subsequent deliveries must be made just in time, or at least close in time.
-We define a maximum time lag $\gamma_i$ beyond which no next unloading operation should be allowed.
-The construction site unloading rate and the quantity to unload give the time necessary to discharge a truckload.
-$\mathcal{C}$ is the set of construction sites (customers) with a planned delivery for the day.
-$\mathcal{O}$ is the set of all requested orders.
+
+Let $O_i$ be the set of all orders requested by customer $i$.
+Each element of $O_i$ must be fully delivered before moving on to another order.
+Exactly one order $o$ of $O_i$ must have its first delivery start at $a_i$, while the others can start at any time after $o$ is completed.
+Once a plant is selected to produce a customer's first delivery, it must be the supplier of all subsequent deliveries from that customer.
+To avoid cold joint problems with the concrete, the subsequent deliveries must be made in close succession.
+We define a maximum time delay $\gamma_i$ after which no more deliveries are allowed.
+The customer unloading rate and the quantity to be unloaded give the time required to unload a truckload.
+Let $\mathcal{C}$ be the set of construction sites (customers) with a planned delivery for the day, and $\mathcal{O}=\{{O_i, i \in \mathcal{C}}\}$ be the set of all requested orders for all customers.
 
@@
 The company has two types of concrete mixer trucks with capacities of 8 and 12 cubic meters.
-Each driver $k$ is assigned to a specific batching plant and is responsible for driving a truck with a specific capacity $Q_k$.
-The set of drivers is represented by $K =\cup_{b \in \mathcal{B}}K_b $, $K_b$ being the set of drivers scheduled to start their shift at the batching plant $b$.
+Each driver $k$ is assigned to a particular batch plant and is responsible for driving a truck with capacity $Q_k$.
+The set of drivers is represented by $K =\cup_{b \in \mathcal{B}}K_b $, where $K_b$ is the set of drivers scheduled to start their shift at batch plant $b$.
 $t_{ij}$ is the known time to travel between any two locations $i$ and $j$.
-A scheduled driver $k$ is required to begin their shift at $H_k$, work for a minimum of $M_T$ hours and a maximum of $N_T$ hours during regular working hours, with the possibility of overtime of up to $O_T$ hours.
-%Let the continuous variable $WT_k$ be the working time of each driver $k$ during a day ($WT_k \leq O_T$).
-$\Gamma_1$ and $\Gamma_2$ are the incurred penalties when a driver works less than $M_t$ and more than $N_t$.
+A scheduled driver $k$ is required to start his shift at $H_k$, work a minimum of $M_T$ hours and a maximum of $N_T$ hours during regular working hours, with the possibility of overtime of up to $O_T$ hours.
+$\beta_3$ and $\beta_4$ are the penalties incurred if a driver works less than $M_t$ and more than $N_t$.
 
-A driver typically loads RMC at their assigned batching plant, but may also be required to drive to and load at other plants if needed.
-Batching plant produces concrete on-demand with recipes specific to each customer.
-This means that a truck cannot hold orders for more than one customer, even if there is spare capacity.
-To fulfill multiple orders, a driver must refill at a plant between deliveries.
-After unloading the RMC, a driver takes $\beta_k$ minutes to clean the concrete mixer before travelling to his next loading plant.
+A driver typically loads RMC at his assigned batch plant but may be required to drive to and load at other plants if needed.
+The batch plant produces concrete on demand using recipes specific to each order.
+This means that a truck can only haul RMC for one order, even if there is spare capacity.
+After unloading the RMC,  driver $k$ takes $Ct_k$ minutes to clean the concrete mixer before proceeding.
 %When assigning drivers to deliveries, the company prioritizes employees with the highest seniority.
 
+Let $n_o$ be the number of deliveries needed to fulfill the order $o$.
+$n_o$ is not known in advance because we use a fleet of trucks with different capacities.
+However, we can compute its lower ($n_o^{min}$) and upper ($n_o^{max}$) bounds using the capacities of the largest ($Q_{max}$) and smallest ($Q_{min}$) available trucks.
+\begin{alignat}{3}
+    \label{mod:c0}
+       n_o^{min} = \left\lceil \frac{q_o}{Q_{max}} \right\rceil \leq n_o \leq n_o^{max} = \left\lceil \frac{q_o}{Q_{min}} \right\rceil  & \text{ } & 
+ \forall  o \in \mathcal{O}.
+\end{alignat}
+
+Let $d^j_{o}$ be the $j^{th}$ visit with load $q^j_{o}$ for order $o$.
+We represent the fulfillment of order $o$ by the visits to the ordered set of delivery nodes $\mathcal{D}_o= \left(d^0_{o},d^1_{o},\cdots, d^{n_o}_{o}\right)$.
+The deliveries of customer $i$ are the ordered set $\mathcal{D}_i= (\mathcal{D}_{o_1}, \mathcal{D}_{o_2},\cdots,\mathcal{D}_{o_{|O_i|}})$, where $o_r$ is the $r^{th}$ delivered order.
+We will refer to $d \in \mathcal{D}_i$ ($d \in \mathcal{D}_o$) as the $d^{th}$ potential delivery of customer $i$ (order $o$).
+$\mathcal{D}=\bigcup_{i\in \mathcal{C}} \mathcal{D}_i$ is the union of all delivery nodes.
+
+% \subsection*{Batching plants}
+
+The company owns a total of eight batching plants.
+Each plant has a loading dock that can accommodate only one truck at a time, which leads to trucks lining up.
+Let $\mathcal{B}$ be the set of batching plants.
+The plants are heterogeneous, as each plant $b$ has its own hourly loading rate, represented by $\tau^l_b$, which affects the duration of the loading process.
+After loading the concrete, the driver spends $\alpha_b$ minutes adjusting the concrete in the truck before heading to the customer site.
+Each plant has its own assigned fleet of trucks, but it can borrow trucks from other plants.
+% or hire external vehicles if necessary.
+ A batching plant can serve a construction site as long as the travel time between the two is less than the concrete's lifespan, which is represented by $\Delta$.
+Let $l_{bj}$ be the loading dock node associated with delivery node $j$ at plant $b$.
+We represent each plant $b$ by the set $\mathcal{L}_b = \{l_{bj}, t_{bj}\leq  \Delta, j \in \mathcal{D} \}$  of loading docks nodes.
+$\mathcal{L}=\bigcup_{b\in B} \mathcal{L}_b$ is the union of all loading docks.
+
 % \subsection*{Model}
-
-A solution to the problem involves making decisions about truck loading schedules, driver assignments to different deliveries, and truck arrival times at construction sites for unloading.
-For a batching plant, the decision involves choosing which driver to load, when to load them, and which construction site they should deliver to.
+A solution to the problem involves decisions about truck loading schedules, driver assignments to different deliveries, and truck arrival times at construction sites for unloading.
+For a batching plant, the decision involves choosing which driver to \textcolor{red}{load}, when to load them, and which construction site they should deliver to.
 For a driver, the decision involves determining the sequence of loading depots and delivery sites.
@@
 
-We now present a formal definition of our problem.
-Let $n_o$ be the number of deliveries needed to fulfill order $o$.
-$n_o$ is not known in advance, as we use a fleet of trucks with various capacities.
-However, we can compute its lower ($n_o^{min}$) and upper ($n_o^{max}$) bounds using the capacities of the highest ($Q_{max}$) and smallest ($Q_{min}$) available trucks.
-The lower bound is the number of deliveries needed if we only use trucks of capacity $Q_{max}$, while the upper bound is the number of deliveries needed if we only use trucks of capacity $Q_{min}$.
-
-\begin{alignat}{3}
-    \label{mod:c1}
-       \left\lceil \frac{q_o}{Q_{max}} \right\rceil \leq n_o \leq \left\lceil \frac{q_o}{Q_{min}} \right\rceil  & \text{ } & 
- \forall  o \in \mathcal{O}.
-\end{alignat}
-
-Let $d^j_{o}$ be the $j_{th}$ visit with load $q^j_{o}$ for order $o$.
-We represent the delivery of order $o$ by the visits to the ordered set of nodes $\mathcal{D}_o= \left(d^0_{o},d^1_{o},\cdots, d^{n_o}_{o}\right)$.
-We represent the deliveries of customer $i$ by the ordered set $\mathcal{D}_i= (\mathcal{D}_{o_1}, \mathcal{D}_{o_2},\cdots,\mathcal{D}_{o_{|O_i|}})$, where $o_r$ is the $rth$ order delivered.
-We will refer to $d \in \mathcal{D}_i$ ($d \in \mathcal{D}_o$) as the $dth$ potential delivery of customer $i$ (order $o$).
-$\mathcal{D}=\bigcup_{i\in \mathcal{C}} \mathcal{D}_i$ is the union of all delivery nodes.
-
-We represent each plant $b$ by the ordered set $L_b=\{L_{b,1}, L_{b,2},\cdots, L_{b,n(b)}\}$ of loading docks nodes, with $n(b)$ being the number of delivery nodes at a maximal travel time of $\Delta$ of $b$.
-We will refer to $L_{b,l}$ or  $l \in L_b$  as the $lth$ potential loading operation at $b$.
-$\mathcal{L}=\bigcup_{b\in B} L_b$ is the union of all loading docks.
-
-Each driver leaves and returns to his home plant each day.
-We represent a driver $k$ home plant's with a starting $s_k$ and ending $e_k$ depot.
-$S$ and $E$ are respectively the sets of starting and ending depot.
+Each driver leaves and returns to their home plant every day.
+We represent the home plant of a driver $k$ with a starting depot $s_k$ and an ending depot $e_k$.
+$S$ and $E$ are the sets of starting and ending depots, respectively.
 
 We define our problem on a complete directed graph where $V=\{ S \cup \mathcal{L} \cup \mathcal{D} \cup E\}$ is the set of nodes.
-The arc sets are $A =  \{(i,j,k) \hspace*{1mm} \vert \hspace*{1mm} i, j \in V \hspace{1mm} k\in K \}$, and $A^D = \{(i,j) \hspace*{1mm} \vert \hspace*{1mm} i, j \in \mathcal{D} \hspace{1mm} \}$.
+The arc sets are $A =  \{(i,j,k) \hspace*{1mm} \vert \hspace*{1mm} i, j \in V \hspace{1mm} k\in K \}$, $A^D = \{(i,j) \hspace*{1mm} \vert \hspace*{1mm} i, j \in \mathcal{D} \hspace{1mm} \}$, and $A^L = \{(i,j) \hspace*{1mm} \vert \hspace*{1mm} i, j \in \mathcal{L} \hspace{1mm} \}$.
 $A$ corresponds to allowed movements of drivers from node $i$ to node $j$. For each driver $k$, the allowed movements are the following:
+
 \begin{itemize}
@@
 For a customer $c$, arcs in $A^D$ link consecutive delivery nodes of the same order   $\lbrace (i,j)\in \mathcal{D}_o, o \in \mathcal{O}_c, i < j  \rbrace$, and pair of delivery nodes of two different orders $ \lbrace (i,d^{0}_{o_2}),  i \in \mathcal{D}_{o_1}, i \geq n^{min}_{o_1}, o_1, o_2 \in \mathcal{O}_c, o_1 \neq o_2 \rbrace $.
+Arcs in $A^L$ link all pairs of loading docks of the same batching plant.
 
-$\delta^{+}(i) = \{(i, j,k) \in A \}$, and $\delta^{-}(i) = \{(j, i,k) \in A \}$ are the outcoming and incoming arc sets of any node $i \in V$.
-Similarly, $\delta^{+}_D(i) = \{(i, j)  \in A^D\}$, and $\delta^{-}_D(i) = \{(j, i) \in A^D \}$ are the outcoming and incoming arc sets of delivery node $i \in \mathcal{D}$.
+We define $\delta^{+}(i) = \{(i, j,k) \in A \}$ and $\delta^{-}(i) = \{(j, i,k) \in A \}$ as the outcoming and incoming arc sets of node $i \in V$; $\delta^{+}_D(i) = \{(i, j)  \in A^D\}$ and $\delta^{-}_D(i) = \{(j, i) \in A^D \}$ are the outcoming and incoming arc sets of delivery node $i \in \mathcal{D}$.
+Similarly, $\delta^{+}_L(i) = \{(i, j)  \in A^L\}$ and $\delta^{-}_L(i) = \{(j, i) \in A^L \}$ are the outcoming and incoming arc sets of loading node $i \in \mathcal{L}$.
 
 Let the binary variable $x^{k}_{ij}$ be $1$ if driver $k$ travels from node $i$ to $j$.
-Binary variable $y_i$ equals $1$ if customer $i$ is completely served.
-$v_i$ is the start of the loading (unloading) operation at node $i \in L \cup D$.
-$w_i$ is the end of the loading (unloading) operation at node $i \in L\cup D$.
-The binary variable $u_{i,j}$ equals $1$ if delivery node $j$ is served just after delivery $i$.
-Binary variable $\sigma_{ib}$ equals 1 if customer $i$ orders' are loaded from plant $b$.
+Binary variable $y_o$ is $1$ when order $o$ is completely served; $v_i$ and $w_i$ are the start and end of the loading (unloading) operation at node $i \in \mathcal{L} \cup \mathcal{D}$.
+The binary variable $u_{ij}$ is $1$ if node $j$ is served just after $i$, the service being either an unloading or a loading operation.
+Binary variable $\sigma_{ib}$ is equal to 1 if the orders of customer $i$ are loaded by plant $b$.
+Variable $q^k_j$ is the quantity to be loaded towards $j$ with vehicle $k$.
+Let $w^1_k$ be a continuous variable indicating the difference between the driver's work time and the minimum number of hours to be worked in a day, and $w^2_k$ indicating the difference between the driver's work time and the normal work time.
+Let $g_i$ be the time between due date and first service start for customer $i$.
 % Binary variable $u_{o_{1},o_{2}}$ equals $1$ if order $o_{2}$ is served directly after order $o_{1}$.
 Binary variables $z_{o,d}$ ($z_{i,d}$) indicates whether or not delivery $d \in  \mathcal{D}_o$ ($d \in  \mathcal{D}_i$) is supposed to be executed.
+$w_i$ is the end of the loading (unloading) operation at node $i \in L\cup D$.
 
-A driver $k$ leaves his starting node exactly once a day and travels to either a loading dock or his ending node.
-When loading at a loading dock $i$, he takes some time to adjust the concrete in the truck before travelling to delivery node $j$ to unload.
-The loading operation duration depends on the plant loading rate and the quantity $q^k_j$ of RMC to load towards $j$.
-After completing his last unloading operation of the day, he returns to his end node.
+The objective function minimizes the travel cost, the penalty costs incurred when customer demands are not fully met, the lateness of each customer's first delivery, the cost of drivers working less than the minimum hours, and the total overtime cost of drivers working beyond their scheduled hours.
+The mathematical model of this variant of the CDP is as follows:
 
-\begin{alignat}{3}
-\label{mod:c2} \sum_{j \in \delta^{+}(s_k) }{x^k_{s_kj}} = 1 & \text{ } &\forall k \in K \\
-\label{mod:c3}   v_j \geq  w_i + \alpha_{i} + t_{ij} - M(1-x^{k}_{ij}) & \text{ }& \forall i \in \mathcal{L}, j \in \mathcal{D},  k \in K \\
-\label{mod:c4}  w_{i} \geq v_{i}  + \frac{q^k_j}{\tau^l_b} -M(1- x^{k}_{ij}) &  \hspace{5mm} &\forall  b \in \mathcal{B},  i \in L_{b}, j \in \mathcal{D}, k \in K \\
-\label{mod:c5} \sum_{j \in \delta^{-}(e_k)}{x^k_{je_k}} = 1 & \text{ }   &  \forall k \in K
-\end{alignat}
-
-Constraints (\ref{mod:c4}) ensure that two trucks cannot be loaded at the same time in a plant.
-
-\begin{alignat}{3}
-    \label{mod:c6}       w_{l-1} \leq v_{l} & \hspace{5mm} &\forall  b \in B,  l \in L_{b}, l \geq 1 
-\end{alignat}
-
-After completing a delivery  at node $i$, a driver must clean the truck, and travel to the next loading depot or end node.
-The unloading service depends on the construction site's rate and the received quantity.
-Unloading operations must finish $\Delta$ minutes after the begin of loading.
-\begin{alignat}{3}
-    \label{mod:c7}     v_j \geq  w_i + \beta_{k} + t_{ij} - M\left(1-x^{k}_{ij}\right) & \text{ }& \forall i \in \mathcal{D}, j \in \mathcal{L} \cup \{e_k\},   k \in K \\
-    \label{mod:c8}          w_{j} \geq v_{j}  + {  \frac{q^k_j}{\tau^u_c} -M\left(1- x^{k}_{ij}\right) } & \text{ }&  \forall c \in  \mathcal{C}, j \in \mathcal{D}_{c}, i \in \mathcal{L},  k \in K \\
-    \label{mod:c31}          w_{j} \leq v_{i}  + \Delta + M\left(1- x^{k}_{ij}\right)&  \text{ }&  \forall c \in  \mathcal{C}, j \in \mathcal{D}_{c}, i \in \mathcal{L},  k \in K
-    % \label{mod:c31}      w_{d-1}  \leq v_{d} & \hspace{5mm} & \forall i \in \mathcal{C}, o \in O_i, d \in \mathcal{D}_{o}, d \geq 1 \\
-    % \label{mod:c10}      z_{o,d} \leq z_{o,d-1} & \hspace{5mm} & \forall i \in \mathcal{C}, o \in O_i, d \in \mathcal{D}_{o}, d \geq 1\\
-    % \label{mod:c11}  v_{d} - w_{d-1} \leq \gamma_i   & \hspace{5mm}&  \forall i \in \mathcal{C}, o \in O_i, d \in \mathcal{D}_{o}, d \geq 1
-\end{alignat}
-
-The first service of any customer $i$ has to start at the due time $a_i$.
-Let $g_i$ be the gap between the due time and the start of first service at $i$.
-This first service may be performed at the first delivery node of any order $\in O_i$.
-Constraints (\ref{mod:c11}--\ref{mod:c13}) are used to find the delivery sequence of the first delivery node of all orders $\in \mathcal{O}_i$.
-
-\begin{alignat}{3}
-    \label{mod:c9}   v_{d^0_{o}} \geq a_i & & \forall  i \in \mathcal{C}, \forall o \in \mathcal{O}_i\\
-    \label{mod:c10}  g_{i} \geq v_{d} - a_i - M\left(  \sum_{ j \in \delta^{-}_{D}(d)  }{u_{j,d}}\right) & & \forall  i \in \mathcal{C}, \forall o_1 \in \mathcal{O}_i, d=d^0_{o_1} \\
-    \label{mod:c11}   \sum_{ \substack{o_1 \in \mathcal{O}_i }}{ \sum_{ j \in \delta^{-}_{D}(d^0_{o_1})}}{u_{j,d^0_{o_1} } } = |\mathcal{O}_i|-1 & \text{ }& \forall  i \in \mathcal{C},|\mathcal{O}_i| > 1 \\
-    \label{mod:c12}   \sum_{ \substack{o_1 \in \mathcal{O}_i }}\sum_{ j \in \delta^{+}_{D}(d^0_{o_1})}{u_{d^0_{o_1},j } } = |\mathcal{O}_i|-1 & \text{ }& \forall  i \in \mathcal{C}, |\mathcal{O}_i| > 1 \\
-    \label{mod:c13}  \hspace{3mm} \smashoperator{ \sum_{ j \in \delta^{+}_{D}(d^0_{o_1})}}{u_{d^0_{o_1},j } } \leq 1, \hspace{2mm} \smashoperator{ \sum_{ j \in \delta^{-}_{D}(d^0_{o_1})}}{u_{j,d^0_{o_1} } } \leq 1, \hspace{2mm} \smashoperator{ \sum_{ j \in \delta^{+}_{D}(d^0_{o_1})}}{u_{d^0_{o_1},j } } +\smashoperator{ \sum_{ j \in \delta^{-}_{D}(d^0_{o_1})}}{u_{j,d^0_{o_1} } } \geq1         & \text{ }& \forall  i \in \mathcal{C}, \forall o_1 \in \mathcal{O}_i
-\end{alignat}
-
-For a customer, an order must be completed before starting another.
-Constraints (\ref{mod:c14}--\ref{mod:c15}) enforce the precedence constraints between the last delivery of an order and the first delivery of the subsequent order.
-
-\begin{alignat}{3}
-  \label{mod:c14} & v_{d^0_{o_2}} \geq w_j - M(1- u_{j,d^0_{o_2}}) & \text{} & \forall i \in \mathcal{C}, \forall o_1, o_2 \in O_i, j \in \mathcal{D}_{o_1}, (j,d^0_{o_2}) \in \delta_{D}^{-}(d^0_{o_2})  \\
-  \label{mod:c15} &  v_{d^0_{o_2}}  \leq  w_j + \gamma_i + M(1- u_{j,d^0_{o_2}}) & \hspace{1mm} & \forall i \in \mathcal{C}, \forall o_1, o_2 \in O_i, j \in \mathcal{D}_{o_1}, (j,d^0_{o_2}) \in \delta_{D}^{-}(d^0_{o_2})
-%   \label{mod:c12} & v_{j} \geq w_d - M(1- u_{d,j}) & \text{} & \forall i \in \mathcal{C}, \forall o_1 \in O_i, d= d_{o_1,0}, j \in \mathcal{D}_{i}, j \neq  d\\
-%   \label{mod:c112} &  v_{j}  \leq  w_d + \gamma_i + M(1- u_{d,j}) & \text{ } & \forall i \in \mathcal{C}, j \in \mathcal{D}_{i}, o_1 \in O_i, d= d_{o_1,0},    j \neq  d\\
-%   \label{mod:c112} &  \smashoperator{\sum \limits_{j \in \mathcal{D}_i, j\neq d}}{u_{d,j}} \leq 1 & \text{ } & \forall i \in \mathcal{C}, o \in O_i,  d = d_{o,0} \\
-%   \label{mod:c113} &  \smashoperator{\sum \limits_{j \in \mathcal{D}_i, j\neq d}}{u_{j,d}} \leq 1 & \text{ } & \forall i \in \mathcal{C}, o \in O_i,  d = d_{o,0} \\
-%   \label{mod:c111} &  \smashoperator{\sum \limits_{j \in \mathcal{D}_i, j\neq d}}{u_{d,j}}+ \smashoperator{\sum_{j \in \mathcal{D}_i, j\neq d}}{u_{j,d}} \geq 1 & \text{ } & \forall i \in \mathcal{C}, o \in O_i,  d = d_{o,0}
-\end{alignat}
-
-Constraints (\ref{mod:c29}--\ref{mod:c30}) enforce the precedence constraints between two consecutive delivery nodes of the same order.
-
-\begin{alignat}{3}
-\label{mod:c29}   &   u_{j-1,j} \geq u_{j,j+1} & \hspace{5mm} & \forall i \in \mathcal{C}, \forall o \in O_i,  j \in \mathcal{D}_{o}, 1 \leq j \leq n_o-1\\
-  \label{mod:c30} &  u_{j-1,j} \geq \sum_{l \in \mathcal{L}}{x_{lj}} & \text{} & \forall i \in \mathcal{C}, \forall o \in O_i, j \in \mathcal{D}_{o}, j \geq 1 
-\end{alignat}
-
-Constraints (\ref{mod:c17}--\ref{mod:c18}) state that it is not possible for two trucks serving consecutive deliveries to be unloaded at the same time and there is a maximum time lag between them.
-
-\begin{alignat}{3}
-  \label{mod:c17} & v_{j} \geq w_{j-1} - M(1- u_{j-1,j}) & \text{} & \forall i \in \mathcal{C}, \forall o \in O_i, j \in \mathcal{D}_{o}, j \geq 1 \\
-  \label{mod:c18} &  v_{j}  \leq  w_{j-1} + \gamma_i + M(1- u_{j-1,j}) & \text{ } & \forall i \in \mathcal{C}, \forall o \in O_i, j \in \mathcal{D}_{o}, j \geq 1
-\end{alignat}
-
-With constraints (\ref{mod:c19}) and (\ref{mod:c20}), we respectively impose that the cumulative capacity of all concrete mixers serving an order and a customer may exceed the required quantities.
-
-\begin{alignat}{1}
-    \label{mod:c19} \sum_{d \in D_o}\sum_{k \in K}{q^k} \sum_{j \in \mathcal{L}}{x^k_{jd}}\geq q_o, \forall o \in \mathcal{O} \\
-    \label{mod:c20} \sum_{d \in D_i}\sum_{k \in K}{q^k} \sum_{j \in \mathcal{L}}{x^k_{jd}}\geq q_i, \forall i \in \mathcal{C}
-\end{alignat}
-
-Constraints (\ref{mod:c21}) and (\ref{mod:c22}) impose that a loading or delivery node may only be visited once by one driver. Constraints (\ref{mod:c23}) and (\ref{mod:c24}) are degree constraints.
-
-\begin{alignat}{3}
-\label{mod:c21} &  \sum_{k \in K}\sum_{j \in \mathcal{D}}{x^{k}_{ij} \leq 1} & \text{ } & i \in \mathcal{L}\\
-  \label{mod:c22}  &  \sum_{k \in K}\sum_{i \in \mathcal{L}}{x^{k}_{ij} \leq 1} & \text{ } & j \in \mathcal{D} \\
-  \label{mod:c23}  &  \sum_{k \in K}\sum_{i \in \mathcal{L}}{x^{k}_{ij} } =  \sum_{k \in K}\sum_{i \in \mathcal{L}}{x^{k}_{ji} } & \text{ } & j \in \mathcal{D} \\
-  \label{mod:c24}  &  \sum_{k \in K}\sum_{i \in \mathcal{D}}{x^{k}_{ij} } =  \sum_{k \in K}\sum_{i \in \mathcal{D}}{x^{k}_{ji} } & \text{ } & j \in \mathcal{L}
-\end{alignat}
-
- Constraints (\ref{mod:c25}) and (\ref{mod:c26}) impose that all orders of customer $i$ must be loaded from the same plant.
-
-\begin{alignat}{2}
-    \label{mod:c25} & \sum_{k \in K} \sum_{l \in \mathcal{L}_b} \sum_{d \in \mathcal{D}}{x^k_{ld}} \leq M*\sigma_{i, b} & \forall i \in \mathcal{C}, b \in \mathcal{B} \\
-    \label{mod:c26} & \sum_{b \in \mathcal{B}} \sigma_{i, b} \leq 1 & \text{ } \forall i \in \mathcal{C}, b \in \mathcal{B} 
-\end{alignat}
-
-Let $w^1_k$ be the binary variables indicating if the driver $k$ has reached the minimum number of hours worked in the day.
-$w^2_k$ is a continuous variable indicating the gap between the driver working time and the normal working time.
-
-\begin{alignat}{3}
-    \label{mod:c27}  M_T*w^1_k \leq v_{e_k} - H_k & \hspace{5mm} &\forall k \in K \\
-    \label{mod:c28}  w^2_k \geq (v_{e_k} - H_k) - N_T & \hspace{5mm} &\forall k \in K.
-\end{alignat}
-
-The objective function (\ref{mod:obj}) minimizes the travelled distance, the penalty costs incurred when customers' demands are not fully satisfied, the lateness of each customer first delivery, the number of drivers who work under the minimum working time and the total overtime cost when drivers work beyond their scheduled hours.
 %The objective function (\ref{mod:obj}) is designed to minimize multiple factors that impact the overall cost and efficiency of the concrete delivery process.
@@
 The penalty costs $\beta_1$, which are incurred when customers' demands are not fully satisfied.The lateness of each customer's first delivery, which can impact customer satisfaction and lead to additional penalty costs.The number of drivers who work under the minimum working time, which can lead to increased labor costs and decreased productivity.The total overtime cost, which is incurred when drivers work beyond their scheduled hours.All these factors are taken into consideration in the objective function to minimize the cost and maximize the satisfaction of the customers.
-\begin{align}
-\label{mod:obj}
-\hspace{-20mm}
-    \begin{split}
-      \min  \smashoperator{\sum_{k \in K}} \sum_{ \substack{i \in \mathcal{L}} } \sum_{ j \in \delta^{+}{(i)} } {t_{ij}x^{k}_{ij}} + 
-      \sum_{k \in K} \sum_{ \substack{i \in \mathcal{D}}} \sum_{j \in \delta^{+}{(i)} } {t_{ij}x^{k}_{ij}} + 
-      \smashoperator{\sum_{ i \in C}}{ \beta_1 \left(1-y_i\right)} \hspace{-0mm} \\ +
-      \beta_2 \smashoperator{ \sum_{i \in \mathcal{C} } } { g_i  } 
-    %   + \beta_3 \smashoperator{\sum_{i \in \mathcal{C}}}{  {\sum_{d \in D_i}}{\left(v_d - w_{d-1} - \gamma_i\right)} } 
-      +  \sum_{k \in K}{\beta_3* w^{1}_k } + \beta_4 *{ w^{2}_k }
-    \end{split}
-\end{align}
 
-\iffalse
-graphe: G=(V,A)
-arc between origin depot, depot
-arc between depot delivery node
-arc between delivery node -> depot
-arc between delivery node -> origin depot
-
-Let $SLT^{bk}_{j}$ and $ELT^{bk}_{j}$ be the start and end loading time of driver $k$ at plant $b$.
-$SUT_{j}$, $EUT_{j}$ be the start and end unloading time.
-$LTS^{bk}_{j}=\left[ SLT^{bk}_{j}, ELT^{bk}_{j} \right] $, and $UTS^{bk}_{j}=\left[ SUT^{bk}_{j}, EUT^{bk}_{j} \right]$ stand for the loading  and unloading operations timeslots.
-$AT_{j}$ is the arrival time of a driver at the node location.
-
-\begin{alignat}{2}
-    \label{eq:ELT}  ELT^{bk}_{j} &= SLT^{bk}_{j}+ LD^b_{j}\\
-    \label{eq:S}         EUT_{j} & = SUT_{j} + UD_{j}
+\begin{alignat}{3}
+\label{mod:obj}   \min &  \smashoperator{ \sum_{ \substack{(i,j,k) \in A} }} {t_{ij}x^{k}_{ij}} + \beta_1\smashoperator{\sum_{ o \in \mathcal{O}}}{  \left(1-y_o\right)} \hspace{-0mm}  + \beta_2 \smashoperator{ \sum_{i \in \mathcal{C} } } { g_i  } +  \sum_{k \in K}{\beta_3 w^{1}_k } + \beta_4 { w^{2}_k }&\\
+    \label{mod:c1} &\sum_{j \in \delta^{+}(s_k) }{x^k_{s_kj}} = 1   &  \forall k \in K  \\
+    \label{mod:c2} &\sum_{j \in \delta^{-}(e_k)}{x^k_{je_k}} = 1   &  \forall k \in K\\
+    \label{mod:c3} & v_j \geq  w_i + \alpha_{i} + t_{ij} - M\left(1-x^{k}_{ij}\right)  & \hspace{-20mm} \forall i \in \mathcal{L}, j \in \delta^{+}(i),  k \in K \\
+    \label{mod:c4}  & v_j \geq  w_i + Ct_{k} + t_{ij} - M\left(1-x^{k}_{ij}\right) & \hspace{-20mm} \forall i \in \mathcal{D}, j \in \delta^{+}(i), k \in K \\
+    \label{mod:c5} & w_{i} \geq v_{i}  + \frac{q^k_j}{\tau^l_b} -M\left(1- x^{k}_{ij}\right) &  \hspace{-20mm} \forall  b \in \mathcal{B},  i \in \mathcal{L}_{b},  j \in \delta^{+}(i), k \in K \\
+    \label{mod:c6} &          w_{j} \geq v_{j}  + {  \frac{q^k_j}{\tau^u_c} -M\left(1- x^{k}_{ij}\right) } & \hspace{-20mm}  \forall c \in  \mathcal{C}, j \in \mathcal{D}_{c}, i \in \delta^{-}(j),  k \in K \\
+    \label{mod:c7} &         w_{j} \leq v_{i}  + \Delta + M\left(1- x^{k}_{ij}\right) & \hspace{-20mm}   j \in \mathcal{D},  i \in \delta^{-}(j),  k \in K \\
+    \label{mod:c8}  &  v_{d^0_{o}} \geq a_i  & \hspace{-20mm} \forall  i \in \mathcal{C}, \forall o \in \mathcal{O}_i\\
+    \label{mod:c9} & g_{i} \geq v_{d^0_{o_1}} - a_i - M\left(  \sum_{ j \in \delta^{-}_{D}(d^0_{o_1})  }{u_{j,d^0_{o_1}}}\right)  & \hspace{-20mm}  \forall  i \in \mathcal{C}, \forall o_1 \in \mathcal{O}_i, d=d^0_{o_1} \\
+     \label{mod:c101} & v_{d^0_{o_1}} \geq w_j - M\left(1- u_{j,d^0_{o_1}}\right)  & \hspace{-40mm}    \forall o_1 \in \mathcal{O}_i,  j \in \delta_{D}^{-}(d^0_{o_1})  \\
+  \label{mod:c10} &  v_{d^0_{o_1}}  \leq  w_j + \gamma_i + M\left(1- u_{j,d^0_{o_1}}\right) &  \hspace{-20mm}  \forall  o_1 \in \mathcal{O},  j \in \delta_{D}^{-}(d^0_{o_1}) \\
+    \label{mod:c11} &  \sum_{ \substack{o_1 \in \mathcal{O}_i }}{ \sum_{ j \in \delta^{-}_{D}(d^0_{o_1})}}{u_{j,d^0_{o_1} } } = |\mathcal{O}_i|-1 & \hspace{-20mm} \forall  i \in \mathcal{C},|\mathcal{O}_i| > 1 \\
+    \label{mod:c12}&   \sum_{ \substack{o_1 \in \mathcal{O}_i }}\sum_{ j \in \delta^{+}_{D}(d^0_{o_1})}{u_{d^0_{o_1},j } } = |\mathcal{O}_i|-1 & \hspace{-40mm} \forall  i \in \mathcal{C}, |\mathcal{O}_i| > 1 \\
+    \label{mod:c13} & \hspace{3mm} \smashoperator{ \sum_{ j \in \delta^{+}_{D}(d^0_{o_1})}}{u_{d^0_{o_1},j } } \leq 1 & \hspace{-20mm}  \forall o_1 \in \mathcal{O} \\
+    \label{mod:c14} & \hspace{2mm} \smashoperator{ \sum_{ j \in \delta^{-}_{D}(d^0_{o_1})}}{u_{j,d^0_{o_1} } } \leq 1 & \hspace{-20mm} \forall o_1 \in \mathcal{O} \\
+    \label{mod:c15} & \hspace{2mm} \smashoperator{ \sum_{ j \in \delta^{+}_{D}(d^0_{o_1})}}{u_{d^0_{o_1},j } } +\smashoperator{ \sum_{ j \in \delta^{-}_{D}(d^0_{o_1})}}{u_{j,d^0_{o_1} } } \geq 1         & \hspace{-20mm} \forall o_1 \in \mathcal{O} \\
+    \label{mod:c16} & v_{j} \geq w_{j-1} - M\left(1- u_{j-1,j}\right) & \hspace{-20mm}  \forall o \in \mathcal{O}, j \in \mathcal{D}_{o}, j \geq 1 \\
+  \label{mod:c17} &  v_{j}  \leq  w_{j-1} + \gamma_i + M\left(1- u_{j-1,j}\right) & \hspace{-20mm} \forall i \in \mathcal{C}, \forall o \in O_i, j \in \mathcal{D}_{o}, j \geq 1 \\
+   \label{mod:c18}   &   u_{j-1,j} \geq u_{j,j+1} & \hspace{-35mm} \forall o_1 \in\mathcal{O},  j \in \mathcal{D}_{o_1}, 1 \leq j \leq n_{o_{1}}-1\\
+  \label{mod:c19} &  u_{j-1,j} \geq \sum_{l \in \mathcal{L}}{x_{lj}} &\hspace{-20mm}  \forall o_1 \in \mathcal{O}, j \in \mathcal{D}_{o_1}, j \geq 1 \\
+    \label{mod:c20} & v_{j} \geq w_{i} - M\left(1- u_{i,j}\right) & \hspace{-20mm}  \forall i \in \mathcal{L}, j \in \delta^{+}_{L}(i) \\
+   \label{mod:c21} & \hspace{3mm} \smashoperator{ \sum_{ j \in \delta^{+}_{L}(i)}}{u_{i,j } } \leq 1 & \hspace{-20mm}  \forall i \in \mathcal{L} \\
+    \label{mod:c22} & \hspace{2mm} \smashoperator{ \sum_{ j \in \delta^{-}_{L}(i)}}{u_{j,i } } \leq 1 & \hspace{-20mm} \forall i \in \mathcal{L} \\
+    \label{mod:c23} & \hspace{2mm} \smashoperator{ \sum_{ j \in \delta^{-}_{L}(i)}}{u_{j,i } } + \smashoperator{ \sum_{ j \in \delta^{+}_{L}(i)}}{u_{i,j } } \geq \sum_{k \in K}\sum_{j \in  \delta^{-}(i) }{x^k_{ji}} & \hspace{-20mm} \forall i \in \mathcal{L} \\
+    \label{mod:c24} & \sum_{k \in K}\sum_{j \in \mathcal{D}_o}{q^k_j} = q_o & \hspace{-20mm}  \forall o \in \mathcal{O} \\
+    \label{mod:c25} &  \sum_{k \in K}\sum_{j \in \delta^{+}(i)}{x^{k}_{ij} \leq 1} &  i \in \mathcal{L} \cup \mathcal{D} \\
+  \label{mod:c26}  &  \sum_{k \in K}\sum_{j \in \delta^{+}(i)} {x^{k}_{ij} } =  \sum_{k \in K}\sum_{j \in \delta^{-}(i)}{x^{k}_{ji} } &  i \in \mathcal{L} \cup \mathcal{D} \\
+  \label{mod:c27} & \sum_{k \in K}  \sum_{i \in \mathcal{D}_c} \sum_{j \in \delta^{-}(i) }  {x^k_{ji}} \leq M\sigma_{c, b} & \forall c \in \mathcal{C}, b \in \mathcal{B} \\
+    \label{mod:c28} & \sum_{b \in \mathcal{B}} \sigma_{c, b} \leq 1 & \hspace{-20mm}\forall c \in \mathcal{C}, b \in \mathcal{B} \\
+    \label{mod:c29} & w^1_k \geq  M_T + H_k - v_{e_k}  & \hspace{-20mm} \forall k \in K \\
+    \label{mod:c30} & w^2_k \geq (v_{e_k} - H_k) - N_T & \hspace{-20mm} \forall k \in K \\
+    \label{mod:c31}  & w_{e_k} \leq H_k + O_T  & \hspace{-00mm}  k \in K \hspace{0mm} \\
+    \label{mod:c32} &  0 \leq q^k_j \leq Q^k  & \hspace{-20mm}  j \in \mathcal{D}, k \in K \\
+    \label{mod:c33} &  x^k_{ij} \in \{0,1\} & \hspace{-20mm}  (i,j) \in A , k \in K  \hspace{0mm} \\
+    \label{mod:c34} &  u_{ij} \in \{0,1\} & \hspace{-20mm}  (i,j) \in A^D , k \in K  \hspace{0mm} \\
+    \label{mod:c35} &  \sigma_{ib} \in \{0,1\} & \hspace{-20mm}  i \in \mathcal{C}, b \in \mathcal{B} \\
+    \label{mod:c36} &  y_o \in \{0,1\} & \hspace{-20mm}  o \in \mathcal{O} \\
+    \label{mod:c37} &  v_i \geq 0,  w_i \geq 0 & \hspace{-20mm}  i \in V \\\
+    \label{mod:c38} &  w^1_{k} \geq 0,  w^2_k \geq 0  & \hspace{-20mm}  i \in V \\
+    \label{mod:cend} &  g_i \geq 0   & \hspace{-20mm}  i \in \mathcal{C}.
 \end{alignat}
 
-For each delivery node $j$, we define an expected delivery time $EDT_{j}$ of unloading operation, the driver waiting duration $W^k_{j}$, and the node waiting time $W_{j}$.
+$\beta_1$ to $\beta_4$ are the penalty coefficients of each component of the objective function (\ref{mod:obj}).
+Constraints (\ref{mod:c1}) and (\ref{mod:c2}) state that driver $k$ leaves his start node exactly once a day and returns to his end node after completing his last unloading operation of the day.
+Constraints (\ref{mod:c3}) and (\ref{mod:c4}) set a driver to take some time after loading or unloading to adjust the concrete or clean the truck before moving on to the next node.
+The duration of the loading operation depends on the plant's loading rate and the amount $q^k_j$ of RMC loaded (\ref{mod:c5}).
+Similarly, the unloading service depends on the site's rate and $q^k_j$ (\ref{mod:c6}).
+Unloading operations must end at most $\Delta$ minutes after loading begins (\ref{mod:c7}).
 
-\begin{alignat}{3}
-    \label{eq:ES} EDT_{d^{p^l}_{ij}} &=
-          \begin{cases}
-            a_i & \text{if $j=0$, $l=0$}\\
-            E_{d^{p^l}_{ij-1}} & \text{ $j \geq 1$}\\
-            E_{d^{p^{l-1}}_{in^{p^{l-1}}_i}}  & 
-          \end{cases} &  \\
-    \label{eq:Wd} W^k_{d^p_{ij}} &= \max\{0, SUT_{d^p_{ij}} - AT_{d^p_{ij}}\} & \\
-    \label{eq:Wc} W_{d^p_{ij}} & = \max\{0, AT_{d^p_{ij}} - EDT_{d^p_{ij}}- \lambda_i \} &
-\end{alignat}
+Constraints (\ref{mod:c8}) - - (\ref{mod:c10}) ensure that the first service of any customer $i$ must start at the due time $a_i$.
+This first service may be performed at the first delivery node of any order $\in O_i$, and one order must be completed before another is started.
+It also enforces the precedence constraints between the last delivery of an order and the first delivery of the following order.
+Constraints (\ref{mod:c11}) - - (\ref{mod:c15}) find the delivery sequence of all orders $\in \mathcal{O}_i$.
+Constraints (\ref{mod:c16}) - - (\ref{mod:c19}) prohibit two trucks serving consecutive deliveries of the same order from unloading at the same time and enforce a maximum time delay between them.
+Similarly, constraints (\ref{mod:c20}) - - (\ref{mod:c23}) ensure that two trucks cannot be loaded at the same time at a plant.
+Constraints (\ref{mod:c24}) require that the cumulative load of all concrete mixers serving an order must equal the required quantities.
+Constraints (\ref{mod:c25}) require that a driver can only visit a loading/delivery node once.
+Constraints (\ref{mod:c26}) are degree constraints.
+Constraints (\ref{mod:c27}) and (\ref{mod:c28}) require that all orders from customer $i$ must be loaded from the same plant.
+Constraints (\ref{mod:c29}--\ref{mod:c30}) calculate the difference between a driver's hours of service and the minimum and normal hours of service.
+Finally, constraints (\ref{mod:c31}) -- (\ref{mod:cend}) define the nature and bounds of the variables.
 
-\fi
-
-\paragraph*{}
-Let us consider the solution of a small instance of our problem illustrated in Figures~\ref{fig_Example} and \ref{fig:ganttExample}.
+% \paragraph*{}
+Let us consider the solution of a small instance of our problem shown in Figures~\ref{fig_Example} and \ref{fig:ganttExample}.
 Two batching plants $B_1$ and $B_2$ with three drivers serve two construction sites.
@@
 Customer $C_2$ requests three orders of $q^1_2=3$, $q^2_2=11$, and $q^3_2=1$ m$^3$ of three different types of concrete.
-$B_1$ has two drivers ($D_1$, $D_2$) using concrete mixers of capacity 8 m$^3$, and $B_2$ has $D_3$ of capacity 12 m$^3$.
-Figure~\ref{fig_Example} depicts the concrete mixers flow in the network.
-We schedule $D_1$ and $D_2$ to respectively deliver 8 and 4 m$^3$ to $C_1$.
-We select $o^2_2$ as the first order of $C_2$ to deliver.
+$B_1$ has two drivers ($D_1$, $D_2$) using concrete mixers of capacity 8 m$^3$, and $B_2$ has $D_3$ with a capacity 12 m$^3$.
+Figure~\ref{fig_Example} shows the flow of concrete mixers flow in the network.
+We schedule $D_1$ and $D_2$ to deliver 8 and 4 m$^3$ to $C_1$, respectively.
+We select $o^2_2$ as the first order of $C_2$.
 $D_3$ travels to $B_1$ to load $q^2_2$, then visits $C_2$ before returning to his home plant.
@@
 The first deliveries of $C_1$ and $C_2$ are timely and there is no gap between the two deliveries received by $C_1$.
-There is a delay between the end of the first delivery and the start of the second delivery of $C_2$, however, it is less than the maximal time lag of 20 minutes defined for this example.
+There is a delay between the end of the first delivery and the start of the second delivery of $C_2$, however, it is less than the maximal time delay of 20 minutes defined for this example.
 
@@
 \section{Constructive heuristics and GRASP}
-\label{method}
+\label{grasp_method}
 
 We now present the heuristic approach we implement to solve this variant of the Concrete Delivery Problem.
-Our method consists in constructing feasible solutions for the CDP with randomized heuristics and iteratively calling these heuristics in the greedy randomized adaptive search procedure (GRASP) metaheuristic.
+Our method consists of constructing feasible solutions to the CDP with randomized heuristics and iteratively invoking these heuristics in the GRASP metaheuristic.
 
-In this section, we represent the scheduling of a delivery node $d$ as a pair of loading $L_{d}$ and unloading $U_{d}$ tasks.
-We then refer to a solution $S$ as a set of loading and unloading tasks performed by a set of drivers during a particular day.
-\begin{alignat*}{2}
-    L_{d} &= \left\lbrace (b,k,q_d,v_b), b \in \mathcal{B}, k \in K \right\rbrace \\
-    U_{d} & =  (k,v_d, w_d), k \in K  \} \\
-    S &=\left\lbrace \cup _{1 \leq j \leq n_o}  (L_{d^j_{o}},U_{d^j_{o}}), \forall o \in \mathcal{O} \right\rbrace
-\end{alignat*}
+In this section, we represent the scheduling of a delivery node $d$ as a pair of loading ($L_{d}$) and unloading ($U_{d}$) tasks.
+$L_{d} = \left(b,k,q_d,v_b\right) $ indicates that driver $k$ starts loading $q_d$ $m^3$ of RMC at plant $b$ at time $v_b$.
+$U_{d} = \left(k,v_d, w_d\right)$ indicates that driver $k$ is serving $d$ between $v_d$ and $w_d$.
+We then call a solution $ S =\left\lbrace \cup _{1 \leq j \leq n_o} (L_{d^j_{o}}, U_{d^j_{o}}), \forall o \in \mathcal{O} \right\rbrace$ the set of all loading and unloading tasks performed on a given day.
 
@@
 The GRASP algorithm is an iterative suite of constructive and local search algorithms introduced by \cite{feo1995greedy}.
-For each iteration, a feasible solution $S$ is built using a greedy randomized algorithm.
-Then a local search algorithm investigates the neighborhood of  $S$ to find a local optimum.
-The pseudo-code of Algorithm~\ref{alg_grasp} shows that the procedure returns the best overall solution (for a minimization problem) after some stopping conditions (time limit or a maximal number of iterations) are met.
+A feasible solution $S$ is constructed using a greedy randomized algorithm at each iteration.
+Then a local search algorithm explores the neighborhood of $S$ to find a local optimum.
+The pseudocode of the algorithm~\ref{alg_grasp} shows that the procedure returns the best overall solution (for a minimization problem) after some stopping conditions (time limit or a maximum number of iterations) are met.
+In our implementation, we introduce at line 13 an additional local search step using the solutions found by all heuristics in the current iteration.
 
@@
             \KwIn{ \textit{H}: List of constructive randomized heuristics }
-            $S^* \leftarrow \emptyset$    \hspace{2mm}       $Cost(S^*) \leftarrow \infty$
+            $S^* \leftarrow \emptyset$    \hspace{2mm} $T \leftarrow \emptyset$       $Cost(S^*) \leftarrow \infty$
 
@@
                 $Cost(S') \leftarrow \infty$
+
+                $T \leftarrow \emptyset$
 
@@
                     $S \leftarrow LocalSearch(S)$
+
+                    $T \leftarrow T \cup \{S\}$
 
@@
                 }
+                $S^* \leftarrow LocalSearch(T)$
             }
@@
 
-As referred to in \cite{resende2019greedy}, the general outline of a greedy randomized algorithm used in the GRASP framework works as described in Algorithm~\ref{alg:greedyIns}.
-At each iteration, let's consider the set $CL$ of candidate delivery nodes that are not yet scheduled.
-We note that, for our problem, not all delivery nodes are candidates for insertion at each iteration, since deliveries of the same order are ordered, any order of a customer may be chosen, and it must be satisfied before switching to another order, and the number of visits to a customer is not known beforehand.
-We first initialize $CL$ with the first delivery of a random order for each customer.
-We create the set $CL'$ with elements of $CL$ we deem promising for a better solution.
-We evaluate the increase in the cost function when incorporating each $d \in CL'$ into the incumbent solution $S$ and create a restricted candidate list $RCL$ formed by nodes whose incremental costs are less than a defined threshold.
-From $RCL$, we randomly select the next delivery node to be incorporated into the incumbent solution and we determine its loading and unloading schedule with the procedure described in Algorithm~\ref{alg_ScheduleTask}.
-Next, we update $S$ and the remaining quantities of the order and customer.
-Then we add to $CL$ the next delivery node candidate.
+As mentioned in \cite{resende2019greedy}, the general outline of a greedy randomized algorithm used in the GRASP framework works as described in Algorithm~\ref{alg:greedyIns}.
+Let's consider the set $CL$ of candidate delivery nodes that have not yet been scheduled.
+We first initialize $CL$ with the first delivery node of a randomly selected order, for each customer.
+
+At each iteration, we create the set $CL'$ with elements of $CL$ from a given neighborhood.
+We then evaluate the increase in the cost function of incorporating each $d \in CL'$ into the incumbent solution $S$ and construct a restricted candidate list $RCL$ consisting of nodes whose incremental cost is less than a given threshold.
+From $RCL$, we randomly select the next delivery node to be incorporated into the incumbent solution and determine its loading and unloading schedule using the procedure described in Algorithm~\ref{alg_ScheduleTask}.
+Next, we update $S$ and the remaining order and customer quantities.
+Then we add the next candidate delivery node to $CL$.
 If an order $o$ is not yet fulfilled after visiting $d^j_{o}$, the candidate is the next delivery node $d^{j+1}_{o}$.
 Otherwise, if the customer has remaining orders, the candidate is the first delivery node $d^{0}_{o'}$ of another randomly selected order $o'$.
+Note that for our problem, not all unscheduled delivery nodes are candidates at each iteration, since deliveries of the same order are ordered, any order of a customer can be chosen and must be satisfied before switching to another order, and the number of visits to a customer is not known in advance.
 
-The greedy aspect of the GRASP is the creation of the $RCL$ set, the probabilistic aspect is the random selection in $RCL$, and the adaptive aspect is the update of $CL$ and the reevaluating of the incremental costs.
-We use the additional set $CL'$ because we noticed that restricting $CL$ to delivery nodes with the same demand, delivery due time, or overlapping unloading timeslots helps us intensify the search.
-Using $CL'$ also helps improve the algorithm complexity by reducing the number of nodes to evaluate.
-To diversify the search we simply remove the filter component.
-We thus obtain four greedy randomized insertion heuristics for our GRASP framework.
+Consider two customers whose services are required to start at the same time or in the same time slot.
+We could either finish scheduling one before the other, or schedule both at the same time.
+Obviously, each of these choices would result in a different solution.
+Similarly, in some cases, it would be better to schedule a customer that has a high demand and therefore requires more resources first.
+That's why we added the filtering part to impose some criteria on the algorithm.
+We can limit $CL'$ to delivery nodes with the same demand, the same due date, or from the same customer.
+We can also decide if we want to schedule customers with the highest (or lowest) demand or the earliest (latest) due time first.
+In fact, this helps us to intensify the search in specific neighborhoods of the solution.
+To diversify the search, we simply remove the filter component.
+The greedy aspect of GRASP is the creation of the set $RCL$, the probabilistic aspect is the random selection in $RCL$, and the adaptive aspect is the updating of $CL$ and the re-evaluation of the incremental costs.
+
+%Also, using $CL'$ helps reduce the complexity of the algorithm by reducing the number of nodes to evaluate.  %We thus obtain four greedy randomized insertion heuristics for our GRASP framework.
 
@@
 
-            $RCL \leftarrow \left\lbrace d \in CL, \text{ } C(d) \leq C_{min} + \alpha (C_{max}-C_{min}) \right\rbrace $
+            $RCL \leftarrow \left\lbrace d \in CL', \text{ } C(d) \leq C_{min} + \alpha (C_{max}-C_{min}) \right\rbrace $
 
@@
  Algorithm~\ref{alg_ScheduleTask} takes as parameters a delivery node $d^j_{o}$ of the customer $i$ and a partial solution and looks for a plant and a driver available for loading and unloading operations.
-For each driver $k$ and plant $b$ we simulate the loading and unloading operations while checking the concrete lifespan constraints, and the plant and driver availability.
+For each driver $k$ and plant $b$ we simulate the loading and unloading operations while checking the concrete lifespan and cold joint constraints, and the plant and driver availability.
 For each node, we first determine the start loading time $SLT$ by subtracting $t_{bi}$, $\alpha_b$, and the loading duration $LD^b_{d^j_{o}}$ to the expected delivery time $EDT$.
@@
 
-\subsection*{ Local Search}
+\subsection{ Local Search}
+
+    % We implement two local search operators for the CDP. The first one is the relocate move commonly used for vehicle routing problems.
+
+   % At the end of each iteration, we expect each customer to have a feasible schedule respecting all constraints.
+For each customer, we combine each scheduling in order to construct a new and possibly better solution.
 
@@
 
+% In this section, we present and discuss the results obtained in the computational experiments of the GRASP heuristic described in Section~\ref{grasp_method}. The GRASP algorithm was coded in C++.
+% and the branch-and-bound (B\&B) algorithm applied to the model described in Section~\ref{model}.
+The different algorithms were coded in C++ and run on a processor Intel $2.4$ GHz Gold $6148$ Skylake.
+For the exact method, we have used IBM Concert Technology and CPLEX $12.10$ as MIP solver.
+A single thread was used with a time limit of one hour.
+% No benchmark instances for the \dirp exist so we have generated the dataset for the experiments as follows.
+
+\subsection{Generation of instances}
+% Due to the lack of publicly available benchmark data for the CDP, two data sets are created.
+Both sets, as well as code to generate new instances, are available online at [17].
+A summary of the instances is provided in Table 4.DataSetAcontains instances with 10–20 customers,and2–5 vehicles,whereasData Set B contains instances with up to 50 customers and 20 vehicles.
+The demands of customers are always divisible by 5 and are selected at random from the interval [10–75].
+In a similar fashion, the capacities of the vehicles are selected between [10–25].
+It is unrealistic to have a unique capacity for each vehicle, and therefore the number of different vehicles is bounded by the number of vehicle classes.Theprocessing time of a vehicle kAK is set proportional to its capacity, i.e.
+pk ¼qk for all kAK.
+The width of the time window of a customer i AC is computed by λiqi,whereλi is a scalar uniformly selected from the interval ½1:1; 2:1.
+Each instance has up to 4 production stations.
+The locations of the start and end depot, construction sites, and production depots, lay in the euclidean plane.
+The travel time between two locations equals the euclidean distance, rounded upwards.
+For simplicity, the start and end depot are the same.
+The distance between each station and the depot is uniformly selected from the interval [1–30].
+Furthermore, for each customer, there exists a production station within a distance of [1–25].Finally,thetimelagγ is fixed to 5 for all deliveries.
 \subsection{Data sets}
+
+\subsubsection{ Results for the generated instances}
+
+\subsubsection{ Results for the real-world datasets}
+\subsubsection{ Results for the instances of \cite{kinable2014concrete}}
+
+We run our algorithm on the dataset provided in \cite{kinable2014concrete}.
 

### 2023-05-26 code/Cost.h
Save

@@ -80,16 +80,16 @@ public:
     }
     double travelCost;
     double satisfiedCost;
-    double waitingCost;
-    double undeliveredCost;
+    int waitingCost;
+    int undeliveredCost;
     double totalCost;
-    double truckWaitingCost;
-    double clientWaitingCost;
-    double firstDeliveryCost;
+    int truckWaitingCost;
+    int clientWaitingCost;
+    int firstDeliveryCost;
     int firstDeliveryCount;
-    double lateDeliveryCost;
+    int lateDeliveryCost;
     int lateDeliveryCount;
-    double overTimeCost;
+    int overTimeCost;
     int driverUsed;
     bool isFeasible;
     int waste;

### 2023-05-26 code/heur/CustInsertion.cpp
Save

@@ -7,171 +7,192 @@
 using namespace std;
 
 void CustInsertion::Insert(Sol &s) {
+    Sol::InitStructure(s.GetData());
+    removedList.clear();
     s.Update();
+    _insrmv.FillStructures(s,customersList,driversList);
 
-    if (not s.availableDrivers.empty()) {
-        driversList.clear();
-        for (int i: s.availableDrivers) {
-            driversList.push_back(s.GetDriver(i));
-        }
-    } else {
-        driversList.clear();
-        Init();
-    }
-    if (not s.keyCustomers.empty()) {
-        for (int i: s.keyCustomers) {
-            Customer *c = s.GetCustomer(i);
-            if (s.isClientSatisfied(c))
-                continue;
-            customersList.emplace_back(c);
-            customersListId.emplace_back(c->custID);
-        }
-    } else {
-        for (int i = 0; i < s.GetCustomerCount(); i++) {
-            Customer *c = s.GetCustomer(i);
-            customersList.emplace_back(c);
-            customersListId.emplace_back(c->custID);
-        }
-    }
-    Insert(s, customersList, customersListId);
+    InsertWithBactrack(s, customersList);
     bool sortie = false;
     while (!removedList.empty() && !sortie) {
         double demand = s.updateCost.satisfiedCost;
-        // cout << s.updateCost << endl;
         customersList = removedList;
-        customersListId = removedListId;
-        Insert(s, customersList, removedListId);
-        // cout << s.updateCost << endl;
+        InsertWithBactrack(s, customersList);
         sortie = (s.updateCost.satisfiedCost == demand);
     }
     s.Update();
 }
 
-
-void CustInsertion::Insert(Sol &s, std::vector<Customer *> &list,
-                           std::vector<int> &list_ID) {
-    if(list.size()>1){
-        Sort(s, list, list_ID, CustInsertion::_k);
+void CustInsertion::InsertWithBactrack(Sol &s, std::vector<Customer *>  &list) {
+    if (list.size() > 1) {
+        Sort(s, list, CustInsertion::_k);
     }
-//    Prompt::print(list);
     removedList.clear();
-    removedListId.clear();
     listMoves.clear();
     listMoves.resize(s.GetDeliveryCount());
-//TODO    std::shuffle(driversList.begin(), driversList.end(), Parameters::RANDOM_GEN);
-    for (auto c: list) {
-//        cout<<*c<<"--"<<std::endl;
[…]

### 2023-05-26 code/heur/CustInsertion.h

@@ -31,8 +31,8 @@ public:
     std::string name;
     std::vector<ListMoveVrp>  listMoves;
     void Insert(Sol &s, std::vector<int> const &list);
-private:
     void InsertWithBactrack(Sol &s,std::vector<Customer *> &list);
+private:
     static void Sort(Sol &s,std::vector<Customer *> &list, int k);
     static void Sort(Sol &s,std::vector<int> &list_ID, int k);
     void Insert2(Sol &s, std::vector<Customer *> &list,std::vector<int> &list_ID);

### 2023-05-26 code/heur/CustInsertionBacktrack.cpp
Save

@@ -1,0 +1,397 @@
+
+#include "CustInsertionBacktrack.h"
+#include "../Prompt.h"
+#include <iostream>
+#include "../Combinatorial.h"
+
+using namespace std;
+
+void CustInsertionBacktrack::Insert(Sol &s) {
+    Sol::InitStructure(s.GetData());
+    customersList.clear();
+    customersListId.clear();
+    s.Update();
+    if (not s.availableDrivers.empty()) {
+        driversList.clear();
+        for (int i: s.availableDrivers) {
+            driversList.push_back(s.GetDriver(i));
+        }
+    } else {
+        driversList.clear();
+        Init();
+    }
+    if (not s.keyCustomers.empty()) {
+        for (int i: s.keyCustomers) {
+            Customer *c = s.GetCustomer(i);
+            if (s.isClientSatisfied(c))
+                continue;
+            customersList.emplace_back(c);
+            customersListId.emplace_back(c->custID);
+        }
+    } else {
+        for (int i = 0; i < s.GetCustomerCount(); i++) {
+            Customer *c = s.GetCustomer(i);
+            customersList.emplace_back(c);
+            customersListId.emplace_back(c->custID);
+        }
+    }
+    Insert(s, customersList, customersListId);
+    bool sortie = false;
+    while (!removedList.empty() && !sortie) {
+        double demand = s.updateCost.satisfiedCost;
+        // cout << s.updateCost << endl;
+        customersList = removedList;
+        customersListId = removedListId;
+        Insert(s, customersList, customersListId);
+        // cout << s.updateCost << endl;
+        sortie = (s.updateCost.satisfiedCost == demand);
+    }
+    s.Update();
+}
+
+void CustInsertionBacktrack::Insert(Sol &s, std::vector<Customer *> &list,
+                           std::vector<int> &list_ID) {
+    if (list.size() > 1) {
+        Sort(s, list, list_ID, CustInsertionBacktrack::_k);
+//        Prompt::print(list,",");
+//        Prompt::print({to_string(list.size())});
+//        Parameters::SHOW = true;
+    }
+    removedList.clear();
+    removedListId.clear();
+    listMoves.clear();
+    listMoves.resize(s.GetDeliveryCount());
+    std::shuffle(driversList.begin(), driversList.end(), Parameters::RANDOM_GEN);
+    Customer *c = list[0];
+//    c->early_tw=144;
+     for (int i = 0; i < list.size();) {
[…]

### 2023-05-26 code/heur/CustInsertionOperator.h
Save

@@ -5,6 +5,7 @@
 
 #include "../InsertOperator.h"
 #include "CustInsertion.h"
+#include "CustInsertionBacktrack.h"
 
 /**
  * k=0, sort by increasing early TW
@@ -15,28 +16,51 @@
  * k=5, sort by decreasing demand
  * k=5, shuffle customer
  */
-class CustInsertionOperator: public InsertOperator<Customer, Driver>
-{
-	public:	
-    CustInsertionOperator(CustInsertion * insert_opt, int k)
-            : _insert_opt(insert_opt),_k(k),
-            InsertOperator<Customer, Driver>("CustInsertionOperator"){
+class CustInsertionOperator : public InsertOperator<Customer, Driver> {
+public:
+    CustInsertionOperator(CustInsertion *insert_opt, int k)
+            : _insert_opt(insert_opt), _k(k),
+              InsertOperator<Customer, Driver>("CustInsertionOperator"), _insert_opt_bt(nullptr) {
         _insert_opt->SetK(_k);
-            }
-    CustInsertionOperator(CustInsertion * insert_opt, int k,std::string const & name)
-            : _insert_opt(insert_opt),_k(k),InsertOperator<Customer, Driver>(name){
-        _insert_opt->SetK(_k);
-        _insert_opt->name=name;
     }
-		void Insert(Sol & s) override
-		{
-			_insert_opt->SetK(_k);
-			_insert_opt->Insert(s);
-		}
-	
-	private:
-		int _k;
-		CustInsertion *_insert_opt;
+
+    CustInsertionOperator(CustInsertionBacktrack *insert_opt_bt, int k)
+            : _insert_opt_bt(insert_opt_bt), _k(k),
+              InsertOperator<Customer, Driver>("CustInsertionOperatorBacktrack"), _insert_opt(nullptr) {
+        _insert_opt_bt->SetK(_k);
+    }
+
+    CustInsertionOperator(CustInsertion *insert_opt, int k, std::string const &name)
+            : _insert_opt(insert_opt), _k(k), InsertOperator<Customer, Driver>(name) {
+        _insert_opt->SetK(_k);
+        _insert_opt->name = name;
+    }
+
+    CustInsertionOperator(CustInsertionBacktrack *insert_opt_bt, int k, std::string const &name)
+            : _insert_opt_bt(insert_opt_bt), _k(k), _insert_opt(nullptr), InsertOperator<Customer, Driver>(name) {
[…]

### 2023-05-26 code/Data.cpp
Save

@@ -11,6 +11,8 @@ void Data::Load(){
     instance_name = filesystem::path(input).stem();
     problem_name = filesystem::path(input).parent_path().filename();
     sol_output = path + "/solution/" + problem_name + "/" + instance_name + ".csv";
+    result_file = path + "/result/" + problem_name + "/results.csv";
+
     if(Parameters::KINABLE){
         Parameters::ADJUSTMENT_DURATION=0;
         Parameters::CLEANING_DURATION = 0;

### 2023-05-26 code/Data.h

@@ -16,7 +16,7 @@ public:
     Data() : input(), output(), nbDepots(0), nbCustomers(0), nbOrders(0), nbDrivers(0), _depots(0), _customers(0), driverCapacities(),
               maxDriverCap(0), minDriverCap(0), _deliveries(0), _drivers(0), _orders_deliveries(0), _distances(0), _times(0), _docks(0),
               _index_orders(0),path(std::filesystem::current_path().parent_path()),instance_name(),
-              problem_name(),sol_output()
+              problem_name(),sol_output(),result_file()
     {
         time_mat_file= "../../matrices/Matrice_de_temps_all.csv";
         distance_mat_file= "../../matrices/Matrice_de_distance_all.csv";
@@ -27,6 +27,7 @@ public:
     std::string output;
     std::string sol_output;
     std::string instance_name;
+    std::string result_file;
     std::string problem_name;
     std::string path;
     int nbDepots,nbCustomers, nbOrders, nbDrivers, maxDriverCap, minDriverCap;

### 2023-05-26 code/InsRmvMethod.h
Save

@@ -47,6 +47,7 @@ public:
 
   virtual void repairSolution(Sol &s) = 0;
 
+
   virtual void SetServiceParams(Sol &s, NodeT *n, DriverT *d,
                                 int demand) = 0;
 

### 2023-05-26 code/heur/InsRmvBuilder.h

@@ -45,12 +45,24 @@ public:
     {
         insrmv.repairSolution(s);
     }
+    virtual void repairSolution(Sol &s,Order *o)
+    {
+        insrmv.repairSolution(s, o);
+    }
+    virtual void repairSolution(Sol &s,Customer *c)
+    {
+        insrmv.repairSolution(s,c);
+    }
 
     virtual Move<Delivery, Driver, MoveVrp>
     GetCost(Sol &s,  Delivery *n, Driver *d, Cost &solcost,
             double demand, ListMove<Delivery, Driver, MoveVrp> *temp_moves)
     {
         return insrmv.GetCost(s,  n, d, solcost, demand, temp_moves);
+    }
+    virtual void FillStructures( Sol &s, std::vector<Customer *> & customersList,
+                                           std::vector<Driver *> & driversList){
+        insrmv.FillStructures(  s, customersList,driversList);
     }
 
     InsRmvMethodFast insrmv;

### 2023-05-26 code/heur/InsRmvBuilder3.h

@@ -7,41 +7,35 @@
 #include <vector>
 #include <iostream>
 
-class InsRmvBuilder3 : public InsRmvBuilder
-{
+class InsRmvBuilder3 : public InsRmvBuilder {
 public:
 
-    explicit InsRmvBuilder3(Data &prob)
-    {
+    explicit InsRmvBuilder3(Data &prob) {
         insrmv = InsRmvMethodFast(prob);
     }
+
     void GetBestInsertion(Sol &s, const std::vector<int> &listId, const std::vector<Driver *> &driversList,
-                          Move<Delivery, Driver, MoveVrp> &best) override
-    {
+                          Move<Delivery, Driver, MoveVrp> &best) override {
         const int NB_ITER = 1;
         int NBR_ITER_WITHOUT_SOL = 0;
-        while (NBR_ITER_WITHOUT_SOL < NB_ITER)
-        {
+        while (NBR_ITER_WITHOUT_SOL < NB_ITER) {
             ListMoveVrp l_moves;
-            for (int delID : listId)
-            {
+            for (int delID: listId) {
                 Delivery *del = s.GetDelivery(delID);
-                assert(del!= nullptr);
-                for (Driver *d : driversList)
-                {
-                    if( std::find(Sol::TabuFleet[del->id].begin(),Sol::TabuFleet[del->id].end(), d->id )!=Sol::TabuFleet[del->id].end())
+                assert(del != nullptr);
+                for (Driver *d: driversList) {
+                    if (std::find(Sol::TabuFleet[del->id].begin(), Sol::TabuFleet[del->id].end(), d->id) !=
+                        Sol::TabuFleet[del->id].end())
                         continue;
                     insrmv.cancel = false;
                     Move<Delivery, Driver, MoveVrp> m;
                     InsertCost(s, del, d, m);
-                    if (m.IsFeasible)
-                    {
+                    if (m.IsFeasible) {
                         l_moves.Add(m);
                     }
                 }
             }
-            if (l_moves.Count() > 0)
-            {
+            if (l_moves.Count() > 0) {
 //                l_moves.Show();
 //                std::cout<<"********\n";
                 best = l_moves.GetRandom();
@@ -50,9 +44,7 @@ public:
 //                best.Show();
 //                exit(1);
                 NBR_ITER_WITHOUT_SOL = NB_ITER;
-            }
-            else
-            {
[…]

### 2023-05-26 code/heur/InsRmvMethodFast.cpp

@@ -1,5 +1,6 @@
 
 #include "InsRmvMethodFast.h"
+#include "../Prompt.h"
 
 using namespace std;
 
@@ -82,25 +83,23 @@ void InsRmvMethodFast::ApplyInsertMove(Sol &s, Move<Delivery, Driver, MoveVrp> &
     assert(m.arrival_dock <= m.arrival_del + s.Travel(m.move.dock, m.n));
 }
 
-Move<Delivery, Driver, MoveVrp> InsRmvMethodFast::GetCost(
-        Sol &s, Delivery *n, Driver *d, Cost &solcost, int demand,
-        ListMove<Delivery, Driver, MoveVrp> *temp_moves) {
-    Order *o = s.GetOrder(n->orderID);
+Move<Delivery, Driver, MoveVrp> InsRmvMethodFast::GetCost(Sol &s, Delivery *n, Driver *d, Cost &solcost, int demand,
+                                                          ListMove<Delivery, Driver, MoveVrp> *temp_moves) {
     Customer *c = s.GetCustomer(n->custID);
     Dock *dock = s.GetDock(n->dockID);
-    int ELT = expected_del_time - s.Travel(dock, n) - ADJUSTMENT_DURATION -
-              LOAD_DURATION;
+    Depot *dep = s.GetDepot(dock->depotID);
+    int ELT = expected_del_time - s.Travel(dock, n) - ADJUSTMENT_DURATION - LOAD_DURATION;
     auto *prec_del_of_cust = dynamic_cast<Delivery *>( s.CustomerPrev[n->EndNodeID]);
     if (Parameters::SHOW) {
-        cout << "Try insertion with " << d->id << endl;
-        s.ShowSlot(d);
+//        Prompt::print({"Try insertion with", to_string(d->id)});
+//        s.ShowSlot(d);
+//        s.Show(d);
     }
     if (prec_del_of_cust != nullptr) {
         Dock *prec_dock = s.GetDock(prec_del_of_cust->dockID);
         ELT = std::max(ELT, s.EndServiceTime[prec_dock->id]);
     }
     TimeSlot loadSlot = TimeSlot(ELT, ELT + LOAD_DURATION);
-
     Node *prev = s.GetNode(d->StartNodeID);
     while (prev->type != Parameters::END_LINK) {
         assert(d == s.DriverAssignTo[prev->id]);
@@ -109,17 +108,17 @@ Move<Delivery, Driver, MoveVrp> InsRmvMethodFast::GetCost(
[…]

### 2023-05-26 code/heur/InsRmvMethodFast.h

@@ -32,6 +32,8 @@ public:
     void SetServiceParams(Sol &s,  Delivery *n, Driver *d, int demand) override;
 
     void repairSolution(Sol &s) override{}
+    void repairSolution(Sol &s,Order *o);
+    void repairSolution(Sol &s,Customer *c);
 
     void InsertCost(Sol &s, Delivery *n, Driver *d,  ListMove<Delivery, Driver, MoveVrp> *temp_moves){}
     Move<Delivery, Driver, MoveVrp> GetCost(
@@ -46,6 +48,8 @@ public:
     int expected_del_time{0};
     int real_del_time{0};
     int max_arrival_Time{0};
+    void FillStructures(Sol &s,std::vector<Customer *> &customersList,
+    std::vector<Driver *> & driversList);
 private:
     Data *_data;
 

### 2023-05-26 code/heur/ListMoveVrp.cpp
Save

@@ -5,15 +5,18 @@
 using namespace std;
 
 bool ListMoveVrpSorter::operator()(Move<Delivery, Driver, MoveVrp> &m1, Move<Delivery, Driver, MoveVrp> &m2) {
+    if(m1.FailureCause!=Parameters::FAILURECAUSE::NONE) return false;
     return m1 <= m2;
 }
 
-
 bool ListMoveVrp::Sort1(Move<Delivery, Driver, MoveVrp> &m1, Move<Delivery, Driver, MoveVrp> &m2) {
+    if(m1.FailureCause!=Parameters::FAILURECAUSE::NONE) return false;
+    if(m1.DeltaCost.lateDeliveryCost<m2.DeltaCost.lateDeliveryCost)
+        return true;
     if (m1.n == m2.n) {
-        if (m1.nbDriver < m2.nbDriver)
+        if (m1.allDriver < m2.allDriver)
             return true;
-        else if (m1.nbDriver == m2.nbDriver) {
+        else if (m1.allDriver == m2.allDriver) {
 
             if (m1.waste == 0) return true;
             else if (m2.waste == 0) return false;
@@ -50,6 +53,10 @@ bool ListMoveVrp::Sort1(Move<Delivery, Driver, MoveVrp> &m1, Move<Delivery, Driv
 }
 
 bool ListMoveVrp::Sort2(Move<Delivery, Driver, MoveVrp> &m1, Move<Delivery, Driver, MoveVrp> &m2) {
+
+    if(m1.FailureCause!=-1) return false;
+    if(m1.DeltaCost.lateDeliveryCost<m2.DeltaCost.lateDeliveryCost)
+        return true;
     if (m1.n == m2.n) {
         if (m1.waste == 0) return true;
         else if (m2.waste == 0) return false;
@@ -58,7 +65,7 @@ bool ListMoveVrp::Sort2(Move<Delivery, Driver, MoveVrp> &m1, Move<Delivery, Driv
             return true;
         } else if (m1.arrival_del == m2.arrival_del) {
 
-            if (m1.nbDriver < m2.nbDriver)
+            if (m1.allDriver < m2.allDriver)
                 return true;
 
 //            if(m1.DeltaCost.travelCost < m2.DeltaCost.travelCost) return true;
@@ -77,7 +84,7 @@ bool ListMoveVrp::Sort2(Move<Delivery, Driver, MoveVrp> &m1, Move<Delivery, Driv
                 return false;
             }
         } else {
[…]

### 2023-05-26 code/heur/ListMoveVrp.h

@@ -23,20 +23,22 @@ class ListMoveVrp : public ListMove<Delivery, Driver, MoveVrp> {
 public:
   ListMoveVrp();
   void Clear();
-  void Add(Move<Delivery, Driver, MoveVrp> &m);
-  int Count();
-  void Insert(int i, Move<Delivery, Driver, MoveVrp> &m);
-  Move<Delivery, Driver, MoveVrp> &Get(int i);
-  Move<Delivery, Driver, MoveVrp>Extract();
+  void Add(Move<Delivery, Driver, MoveVrp> &m) override;
+  int Count() override;
+  void Insert(int i, Move<Delivery, Driver, MoveVrp> &m) override;
+  void Insert( ListMoveVrp &to_insert);
+  Move<Delivery, Driver, MoveVrp> &Get(int i) override;
+  Move<Delivery, Driver, MoveVrp>Extract() override;
 
-  void Sort();
-  Move<Delivery, Driver, MoveVrp> &GetRandom();
-  void Resize(int n);
-  void partial_Sort(int k);
-  void Show();
+  void Sort() override;
+  Move<Delivery, Driver, MoveVrp> &GetRandom() override;
+  void Resize(int n) override;
+  void partial_Sort(int k) override;
+  void Show() override;
   std::vector<Move<Delivery, Driver, MoveVrp>> _moves;
 
 private:
+    std::unordered_map<std::string, Move<Delivery, Driver, MoveVrp>  , MoveHashFunction<Delivery, Driver, MoveVrp> > moveMap;
   ListMoveVrpSorter move_sorter;
   static bool Sort1(Move<Delivery, Driver, MoveVrp> &m1,Move<Delivery, Driver, MoveVrp> &m2);
   static bool Sort2(Move<Delivery, Driver, MoveVrp> &m1,Move<Delivery, Driver, MoveVrp> &m2);

### 2023-05-26 code/Move.h
Save

@@ -6,52 +6,81 @@
 
 #ifndef _MOVE_H_
 #define _MOVE_H_
+
 #include "Solution.h"
 #include "lib/mathfunc.h"
-#include <cstdlib>
+//#include <cstdlib>
+#include <sstream>
 
-template <class NodeT, class DriverT, class MoveT> class Move {
+template<class NodeT, class DriverT, class MoveT>
+class Move {
 public:
-  NodeT *n;
-  DriverT *to;
-  double demand;
-  Cost DeltaCost;
-  double DeltaDistance;
-  int waste;
-  int arrival_dock, arrival_del;
-  bool IsFeasible;
-  int nbDriver{0};
-  MoveT move;
+    NodeT *n;
+    DriverT *to;
+    int demand;
+    Cost DeltaCost;
+    double DeltaDistance;
+    int waste;
+    int arrival_dock, arrival_del;
+    bool IsFeasible;
+    int FailureCause;
+    int clientDriver{0};
+    int allDriver{0};
+    MoveT move;
 
-  Move()
-      : n(nullptr), to(nullptr), IsFeasible(false), demand(0),
-         DeltaDistance(INFINITY),
-        DeltaCost(false), arrival_dock(0), arrival_del(0),
-        waste(0) {}
+    Move()
+            : n(nullptr), to(nullptr), IsFeasible(false), demand(0),
+              DeltaDistance(INFINITY),
+              DeltaCost(false), arrival_dock(0), arrival_del(0),
+              waste(0), FailureCause(Parameters::FAILURECAUSE::NONE) {}
 
-  bool operator<(const Move<NodeT, DriverT, MoveT> &m) const {
-        if(this->n == m.n){
+    Move(NodeT *n, DriverT *to) : n(n), to(to), IsFeasible(false), demand(0),
+                                  DeltaDistance(INFINITY),
+                                  DeltaCost(false), arrival_dock(0), arrival_del(0),
+                                  waste(0), FailureCause(Parameters::FAILURECAUSE::NONE) {}
+
+    bool operator<(const Move<NodeT, DriverT, MoveT> &m) const {
+        if (this->n == m.n) {
             return (this->arrival_del <= m.arrival_del);
         }
         return (DeltaCost < m.DeltaCost);
-  }
+    }
 
-  bool operator<=(const Move<NodeT, DriverT, MoveT> &m)const {
-    return (DeltaCost <= m.DeltaCost);
-  }
[…]

### 2023-05-26 code/heur/MoveVrp.h

@@ -25,6 +25,7 @@ public:
 	double DeltaDistance;
 	int arrival_time;
 	bool IsFeasible;
+    int prevMaxTime;
 };
 
 #endif

### 2023-05-26 code/Parameters.cpp
Save

@@ -22,8 +22,9 @@ std::mt19937 Parameters::RANDOM_GEN =
 int Parameters::LOADING_DURATION=5;
 int Parameters::TIME_BTW_DELIVERY=5;
 bool Parameters::PENALTY_COST = false;
-bool Parameters::KINABLE= true;
+bool Parameters::KINABLE = true;
 bool Parameters::SHOW = false;
+int Parameters::DRIVER_USE = Parameters::MINIMIZEDRIVER::SOLUTION;
 int Parameters::SORT_TYPE= Parameters::SORT::TWO;
  std::chrono::steady_clock::time_point Parameters::START_TIME= std::chrono::high_resolution_clock::now() ;
 

### 2023-05-26 code/Parameters.h

@@ -31,11 +31,15 @@ public:
     static bool PENALTY_COST;
     static bool KINABLE;
     static bool SHOW;
+    static int DRIVER_USE;
     static int SORT_TYPE;
     static std::chrono::steady_clock::time_point START_TIME;
     static void ShowTime(){
         auto elapsed_time = std::chrono::duration_cast<std::chrono::milliseconds>(std::chrono::high_resolution_clock::now() - Parameters::START_TIME ).count();
         std::cout << "\nTemps " << elapsed_time / 1000.0 << " s "<<(elapsed_time / 1000.0)/60<<" min"<< std::endl;
+    }
+    static long GetElapsedTime(){
+        return  std::chrono::duration_cast<std::chrono::milliseconds>(std::chrono::high_resolution_clock::now() - Parameters::START_TIME ).count();
     }
     enum TypeNode
     {
@@ -60,6 +64,11 @@ public:
         TWO,
         THREE,
         FOUR
+    };
+    enum MINIMIZEDRIVER{
+        SOLUTION,
+        CLIENT,
+        HYBRID
     };
 
 };

### 2023-05-26 code/heur/PriorityQueueInsertion.cpp
Save

@@ -9,52 +9,21 @@ using namespace std;
  */
 void PriorityQueueInsertion::Insert(Sol &s) {
     _data = *s.GetData();
+    Sol::InitStructure(s.GetData());
     s.Update();
-
-    if (not s.availableDrivers.empty())
-    {
-        driversList.clear();
-        for (int i : s.availableDrivers)
-        {
-            driversList.push_back(s.GetDriver(i));
-        }
-    }
-    else{
-        driversList.clear();
-        for (int i = 0; i < _data.GetDriverCount(); i++)
-            driversList.push_back(_data.GetDriver(i));
-    }
-    customersList.clear();
     removedList.clear();
-    if (!s.keyCustomers.empty()) {
-        for (int i: s.keyCustomers) {
-            Customer *c = s.GetCustomer(i);
-            if (s.isClientSatisfied(c))
-                continue;
-            customersList.emplace_back(c);
-        }
-    } else {
-        for (int i = 0; i < s.GetCustomerCount(); i++) {
-            Customer *c = s.GetCustomer(i);
-            if (s.isClientSatisfied(c))
-                continue;
-            customersList.emplace_back(c);
-        }
-    }
-
-    InsertByTW(s);
+    _insrmv.FillStructures(s,customersList,driversList);
+    InsertByRules(s);
 }
 
-void PriorityQueueInsertion::InsertByTW(Sol &s) {
+void PriorityQueueInsertion::InsertByRules(Sol &s) {
 
     Insert(s, customersList);
     bool sortie = false;
     while (!removedList.empty() && !sortie) {
         double demand = s.updateCost.satisfiedCost;
-        // cout << s.updateCost << endl;
         customersList = removedList;
         Insert(s, customersList);
-        // cout << s.updateCost << endl;
         sortie = (s.updateCost.satisfiedCost == demand);
     }
     s.Update();
@@ -64,7 +33,7 @@ void PriorityQueueInsertion::Insert(Sol &s, std::vector<Customer *> &list) {
     priority_file.Vider();
     FillQueue(s, list);
     removedList.clear();
-    std::shuffle(driversList.begin(), driversList.end(), Parameters::RANDOM_GEN);
+//    std::shuffle(driversList.begin(), driversList.end(), Parameters::RANDOM_GEN);
 
     while (!priority_file.EstVide()) {
         Move<Delivery, Driver, MoveVrp> best;
[…]

### 2023-05-26 code/heur/PriorityQueueInsertion.h

@@ -35,7 +35,7 @@ public:
   std::string name;
 
 private:
-  void InsertByTW(Sol& s);
+  void InsertByRules(Sol& s);
   void Insert(Sol &s,std::vector<Customer *> &list);
   void DecreaseQueue(Sol &s,Delivery *del, Delivery *next_del, Customer *c);
   void Init() {
@@ -58,7 +58,6 @@ private:
   std::vector<Depot *> depotList;
   std::vector<Customer *> customersList;
   std::vector<Customer *> removedList;
-
   std::vector<Driver *> driversList;
 };
 

### 2023-05-26 code/Solution.cpp
Save

@@ -5,7 +5,7 @@
 using namespace std;
 
 std::vector<int> Sol::FailureCause = std::vector<int>();
-std::vector<int> Sol::CustomerConflict = std::vector<int>();
+std::vector<std::set<int>> Sol::CustomerConflict = std::vector<std::set<int>>();
 std::vector<int> Sol::FailureCount = std::vector<int>();
 std::vector<int> Sol::minDelay = std::vector<int>();
 std::vector<int> Sol::StartBefore = std::vector<int>();
@@ -62,9 +62,17 @@ Sol::Sol(Data *data) : _data(data), _last_cost(false), DriverAssignTo(data->GetN
 }
 
 void Sol::InitCustomers() {
-    for (int i = 0; i < _data->GetCustomerCount(); i++) {
-        InitCustomer(_data->GetCustomer(i));
+    updateCost.undeliveredCost = 0;
+    if (keyCustomers.empty()) {
+        for (int i = 0; i < _data->GetCustomerCount(); i++) {
+            InitCustomer(_data->GetCustomer(i));
+        }
+    } else {
+        for (auto id: keyCustomers) {
+            InitCustomer(_data->GetCustomer(id));
+        }
     }
+
 }
 
 void Sol::InitCustomer(Customer *c) {
@@ -73,12 +81,26 @@ void Sol::InitCustomer(Customer *c) {
     OrderVisitCount[c->custID] = 0;
     sumServiceTime[c->custID] = 0;
     WaitingTime[c->id] = 0;
-}
-void Sol::InitDrivers() {
-    for (int i = 0; i < GetDriverCount(); i++) {
-        InitDriver(GetDriver(i));
+    updateCost.undeliveredCost += c->demand;
+    for(auto o: GetOrders(c))
+    {
+        InitOrder(o);
     }
 }
+
+void Sol::InitDrivers() {
+    if (availableDrivers.empty()) {
+        for (int i = 0; i < GetDriverCount(); i++) {
+            InitDriver(GetDriver(i));
+        }
+    }
+    else{
+        for(auto id:availableDrivers){
+            InitDriver(GetDriver(id));
+        }
+    }
+}
+
 void Sol::InitDriver(Driver *d) {
     d->Init();
     shiftDuration[d->id] = 0;
@@ -88,9 +110,10 @@ void Sol::InitDriver(Driver *d) {
 
 void Sol::InitDepots() {
     for (int i = 0; i < GetDepotCount(); i++) {
[…]

### 2023-05-26 code/Solution.h

@@ -70,7 +70,7 @@ public:
 
 
     static std::vector<int> FailureCause;
-    static std::vector<int> CustomerConflict;
+    static std::vector<std::set<int>> CustomerConflict;
     static std::vector<std::set<int>> TabuFleet;
     static std::vector<int> FailureCount;
     static std::vector<int> minDelay;
@@ -434,7 +434,7 @@ public:
 
     static void InitStructure(Data *dat){
         Sol::CustomerConflict.clear();
-        Sol::CustomerConflict.resize(dat->GetCustomerCount(),-1);
+        Sol::CustomerConflict.resize(dat->GetCustomerCount());
         Sol::FailureCause.clear();
         Sol::FailureCause.resize(dat->GetNodeCount(), Parameters::FAILURECAUSE::NONE);
         Sol::FailureCount.clear();

### 2023-05-26 code/ListMove.h
Save

@@ -20,6 +20,8 @@ public:
   virtual void Insert(int i, Move<NodeT, DriverT, MoveT> &m) = 0;
   virtual Move<NodeT, DriverT, MoveT> &Get(int i) = 0;
   virtual Move<NodeT, DriverT, MoveT> Extract() = 0;
+  void Insert( ListMove<NodeT, DriverT, MoveT> &to_insert);
+
   virtual void Sort() = 0;
   virtual Move<NodeT, DriverT, MoveT> &GetRandom() = 0;
   virtual void Resize(int n) = 0;

### 2023-05-26 code/GRASP.cpp
Save

@@ -1,33 +1,29 @@
 #include "GRASP.h"
 #include "Driver.h"
 #include "Node.h"
-#include <cassert>
 
 using namespace std;
-template <class NodeT, class DriverT>
-void GRASP<NodeT, DriverT>::Optimize(Sol &s,bool first_improvement)
-{
-	Optimize(s, nullptr,nullptr,nullptr, false);
-}
-template <class NodeT, class DriverT>
-void GRASP<NodeT, DriverT>::Optimize(
-	Sol &s, ISolutionList<NodeT, DriverT> *best_sol_list)
-{
-	cout << "Not implemented\n";
-	exit(1);
+
+template<class NodeT, class DriverT>
+void GRASP<NodeT, DriverT>::Optimize(Sol &s, bool first_improvement) {
+    Optimize(s, nullptr, nullptr, nullptr, first_improvement);
 }
 
-template <class NodeT, class DriverT>
+template<class NodeT, class DriverT>
 void GRASP<NodeT, DriverT>::Optimize(
-	Sol &s, ISolutionList<NodeT, DriverT> *best_sol_list, unordered_map<string, Sol, MyHashFunction> *Mymap,
-    RechercheLocale * loc_search,bool first_improvement)
-{
-	Sol best;
-	std::vector<Node *> list_cust;
+        Sol &s, ISolutionList<NodeT, DriverT> *best_sol_list) {
+    Optimize(s, best_sol_list, nullptr, nullptr, true);
+}
 
-	std::vector<BestSolutionList<Node, Driver>> tempVectSolutions;
+template<class NodeT, class DriverT>
+void GRASP<NodeT, DriverT>::Optimize(
+        Sol &s, ISolutionList<NodeT, DriverT> *best_sol_list, unordered_map<string, Sol, MyHashFunction> *Mymap,
+        RechercheLocale *loc_search, bool first_improvement) {
+    Sol best;
+    std::vector<Node *> list_cust;
+    std::vector<BestSolutionList<Node, Driver>> tempVectSolutions;
 
-	list_cust.resize(s.GetCustomerCount());
+    list_cust.resize(s.GetCustomerCount());
 // TODO	for (int i = 0; i < s.GetCustomerCount(); i++)
 //	{
 //		Node *c = s.GetCustomer(i);
@@ -35,40 +31,40 @@ void GRASP<NodeT, DriverT>::Optimize(
[…]

### 2023-05-26 code/GRASP.h

@@ -49,7 +49,7 @@ template <class NodeT, class DriverT>
 class GRASP : public IAlgorithm<NodeT, DriverT> {
 public:
   GRASP() : _data(nullptr) {
-    _iterator_count = 25000;
+    _iterator_count = 10;
     _temperature = 0.9896;
     _pourcentage_max = 0.4;
     _pourcentage_min = 0.1;

### 2023-05-26 code/heur/main_heur.cpp
Save

@@ -3,7 +3,6 @@
 #include "../Prompt.h"
 #include "../TimeSlot.h"
 #include "Solver.h"
-#include "../model/SolverExact.h"
 using namespace std;
 int main(int argc, const char **argv) {
     std::cout << "Ciment Quebec" << std::endl;
@@ -13,8 +12,6 @@ int main(int argc, const char **argv) {
     }
     Data data;
     data.input = argv[1];
-
-
     cout<<data.input<<endl;
     auto start_time = std::chrono::high_resolution_clock::now();
 
@@ -22,9 +19,8 @@ int main(int argc, const char **argv) {
     Solver solver(&data);
     solver.run();
 
-    auto elapsed_time = std::chrono::duration_cast<std::chrono::milliseconds>(std::chrono::high_resolution_clock::now() - start_time).count();
-    cout << "\nTemps " << elapsed_time / 1000.0 << " s "<<(elapsed_time / 1000.0)/60<<" min"<< std::endl;
-//    data.ShowData();
+    auto elapsed_time = std::chrono::duration_cast<std::chrono::milliseconds>(std::chrono::high_resolution_clock::now() - Parameters::START_TIME ).count();
 
+    Parameters::ShowTime();
     return 0;
 }

### 2023-05-26 code/heur/CDPSolver.cpp
Save

@@ -1,7 +1,9 @@
 
 #include "CDPSolver.h"
 #include "CustInsertion.h"
+#include "CustInsertionBacktrack.h"
 #include "CustInsertionOperator.h"
+#include "AllInsertionOperator.h"
 #include "DriverInsertion.h"
 #include "InsRmvMethodFast.h"
 #include "InsRmvBuilder1.h"
@@ -58,7 +60,12 @@ std::set<int> CDPSolver::EliminateCustomer(Data &data, const int iter) {
         std::vector<int> temp{i};
         Data dat = data.copyCustomersData(temp);
         Sol s(&dat);
-        CDPSolver::SolveInstance(s, dat, iter);
+        s.keyCustomers = {0};
+        InsRmvBuilder3 builder3(dat);
+        CustInsertionBacktrack custIns_BT3(dat, builder3);
+        custIns_BT3.Insert(s);
+
+//        CDPSolver::SolveInstance(s, dat, iter);
         if (s.isFeasible) {
             cout << i << "-" << std::flush;
             feasibleClients.insert(i);
@@ -75,15 +82,6 @@ std::set<int> CDPSolver::EliminateCustomer(Data &data, const int iter) {
 
 void CDPSolver::SolveInstance(Sol &s, Data &dat, int iter) {
 
-    Sol::FailureCause.resize(dat.GetNodeCount(), Parameters::FAILURECAUSE::NONE);
-    Sol::FailureCount.resize(dat.GetNodeCount(), 0);
-//    Sol::minDelay.resize(dat.GetNodeCount(),0);
-    Sol::StartBefore.resize(dat.GetNodeCount(), 0);
-    Sol::pullVisit.resize(dat.GetNodeCount(), 0);
-    Sol::pushVisit.resize(dat.GetNodeCount(), 0);
-    Sol::TabuFleet.resize(dat.GetNodeCount());
-
-    TimeSlot::myData = dat;
     InsRmvMethodFast insrmv(dat);
     InsRmvBuilder1 builder1(dat);
     InsRmvBuilder2 builder2(dat);
@@ -91,6 +89,7 @@ void CDPSolver::SolveInstance(Sol &s, Data &dat, int iter) {
     CustInsertion custIns1(dat, builder1);
     CustInsertion custIns2(dat, builder2);
     CustInsertion custIns3(dat, builder3);
+    CustInsertionBacktrack custIns_BT3(dat, builder3);
 
[…]

### 2023-05-26 code/heur/Solver.cpp
Save

@@ -4,7 +4,6 @@
 #include <istream>
 #include "CustInsertion.h"
 #include "CustInsertionOperator.h"
-#include "../Combinatorial.h"
 #include "InsRmvMethodFast.h"
 #include "InsRmvBuilder1.h"
 #include "InsRmvBuilder2.h"
@@ -12,27 +11,104 @@
 
 using namespace std;
 
-void Solver::run() {
-
+void Solver::run(){
     feasibleClients = CDPSolver::EliminateCustomer(*_data, 1);
-    std::vector<Customer *> listCust = _data->GetCustomers();
-    std::vector<int> listCustId(_data->GetCustomerCount());
-    std::iota(listCustId.begin(), listCustId.end(), 0);
+    vector<TimeSlot> listInt;
+    for (int i: feasibleClients) {
+        Customer *c = _data->GetCustomer(i);
+        listInt.emplace_back(c->early_tw, c->late_tw, c->custID);
+    }
+    vector<set<int>> linkedClientSlot;
+    vector<set<int>> linkedClientDemand;
+    vector<set<int>> linkedClients;
+    vector<set<int>> linkedClientsInf;
+    vector<set<int>> linkedClientSup;
+    Data dat=*_data;
+    CDPSolver::findCorrelation(dat, listInt, linkedClientSlot, linkedClientDemand, linkedClients, linkedClientsInf,
+                               linkedClientSup);
+    Sol s(&dat);
+    s.keyCustomers=feasibleClients;
+    SolveGrasp(s,dat,linkedClientSlot);
+    SaveResults(s);
+    s.ShowCustomer();
+}
+void Solver::SaveResults(Sol &s){
+    printf("Instance:%s\n", _data->instance_name.c_str());
 
-    SortNode<Customer, Driver>::radixSortMinWidthTW(listCust, _data->maxLateTW - _data->minEarlyTW);
-    Prompt::print(listCust);
-//    Test();
+    printf("Nombre clients %d \t", _data->nbCustomers);
+    printf("Nombre ordre %d \t", _data->nbOrders);
+    printf("Total Cost:%.2lf \t", s.GetLastCost().satisfiedCost);
 
-    Sol s1(_data);
+    if ( _data->result_file.c_str() != nullptr) {
+        FILE *f = fopen(_data->result_file.c_str(), "a");
+        if (f != nullptr) {
+            time_t now = time(nullptr);
+            tm *ltm = localtime(&now);
+            fprintf(f, "%d-", 1900 + ltm->tm_year);
[…]

### 2023-05-27 code/Combinatorial.cpp
save

@@ -1,6 +1,7 @@
 
 #include "Combinatorial.h"
 #include <functional>
+#include <cmath>
 
 using namespace std;
 void Combinatorial::find_combinations(vector<int> &nums, int k, vector<int> &combination, set<vector<int>> &result, set

### 2023-05-27 code/Parameters.cpp

@@ -26,6 +26,4 @@ bool Parameters::KINABLE = true;
 bool Parameters::SHOW = false;
 int Parameters::DRIVER_USE = Parameters::MINIMIZEDRIVER::SOLUTION;
 int Parameters::SORT_TYPE= Parameters::SORT::TWO;
- std::chrono::steady_clock::time_point Parameters::START_TIME= std::chrono::high_resolution_clock::now() ;
-
-
+std::chrono::time_point<std::chrono::steady_clock> Parameters::START_TIME= std::chrono::steady_clock::now();

### 2023-05-27 code/Parameters.h

@@ -33,13 +33,16 @@ public:
     static bool SHOW;
     static int DRIVER_USE;
     static int SORT_TYPE;
-    static std::chrono::steady_clock::time_point START_TIME;
+    static std::chrono::time_point<std::chrono::steady_clock> START_TIME;
+    
+       
     static void ShowTime(){
-        auto elapsed_time = std::chrono::duration_cast<std::chrono::milliseconds>(std::chrono::high_resolution_clock::now() - Parameters::START_TIME ).count();
+        auto elapsed_time = std::chrono::duration_cast<std::chrono::milliseconds>(std::chrono::steady_clock::now()- Parameters::START_TIME ).count();
         std::cout << "\nTemps " << elapsed_time / 1000.0 << " s "<<(elapsed_time / 1000.0)/60<<" min"<< std::endl;
     }
     static long GetElapsedTime(){
-        return  std::chrono::duration_cast<std::chrono::milliseconds>(std::chrono::high_resolution_clock::now() - Parameters::START_TIME ).count();
+        return  std::chrono::duration_cast<std::chrono::milliseconds>(std::chrono::steady_clock::now()
+         - Parameters::START_TIME ).count();
     }
     enum TypeNode
     {

### 2023-05-27 code/heur/ListMoveVrp.h

@@ -13,7 +13,7 @@
 #include "../Node.h"
 #include "MoveVrp.h"
 #include <vector>
-
+#include <unordered_map>
 struct ListMoveVrpSorter {
   bool operator()(Move<Delivery, Driver, MoveVrp> &m1,
                   Move<Delivery, Driver, MoveVrp> &m2) ;

### 2023-05-27 code/heur/main_heur.cpp

@@ -19,7 +19,7 @@ int main(int argc, const char **argv) {
     Solver solver(&data);
     solver.run();
 
-    auto elapsed_time = std::chrono::duration_cast<std::chrono::milliseconds>(std::chrono::high_resolution_clock::now() - Parameters::START_TIME ).count();
+    // auto elapsed_time = std::chrono::duration_cast<std::chrono::milliseconds>(std::chrono::high_resolution_clock::now() - Parameters::START_TIME ).count();
 
     Parameters::ShowTime();
     return 0;

### 2023-05-27 code/heur/CDPSolver.cpp
Save

@@ -97,11 +97,11 @@ void CDPSolver::SolveInstance(Sol &s, Data &dat, int iter) {
     vector<InsertOperator<Customer, Driver> *> grasp_heuristics;
 
     vector<pair<int, string>> custInfo = {
-//            {0, "Cust Sort Early TW"},
-//            {1, "Cust Sort Greater D"},
+            {0, "Cust Sort Early TW"},
+            {1, "Cust Sort Greater D"},
 //            {2, "Cust Sort Late TW "},
 //            {3, "Cust Sort Min Width TW"},
-//            {4,"Cust Random"},
+            {4,"Cust Random"},
 //            {5, "Cust Sort Kinable"},
 //            {-1, "Cust customized Sort"}
     };
@@ -145,10 +145,10 @@ void CDPSolver::SolveInstance(Sol &s, Data &dat, int iter) {
             Sol cur(&dat);
             cur.keyCustomers = s.keyCustomers;
             cur.availableDrivers = s.availableDrivers;
-            Prompt::print({"Iter", to_string(i), heur.name});
+//            Prompt::print({"Iter", to_string(i), heur.name});
             heur.Insert(cur);
             if (cur < s) {
-                cout << "current best cost " << cur.GetCost().satisfiedCost << " with " << heur.name << endl;
+//                cout << "current best cost " << cur.GetCost().satisfiedCost << " with " << heur.name << endl;
                 s = cur;
                 if (s.hasScheduled(s.keyCustomers))
                     break;

### 2023-08-02 codeCQ/heur/CustInsertion.cpp
Save CQ, release driver after 2h of overtime

@@ -77,7 +77,7 @@ bool CustInsertion::Insert(Sol &s, Customer *c, Order *cur_order) {
     int depth = 0;
 //    if (c->custID == 86)
 //        Parameters::SHOW = true;
-    Prompt::log(s.depotLoadingIntervals[c->depotID]);
+    Prompt::log(s.depotLoadingIntervals[cur_order->depotID]);
     for (int j = 0; j < s.GetDeliveryCount(cur_order);) {
         Delivery *del = s.GetDelivery(cur_order, j);
         if (del == nullptr) break;

### 2023-08-02 codeCQ/heur/InsRmvBuilder3.h

@@ -21,9 +21,9 @@ public:
             Delivery *del = s.GetDelivery(delID);
             assert(del != nullptr);
             for (Driver *d: driversList) {
-                if (std::find(Sol::TabuFleet[del->id].begin(), Sol::TabuFleet[del->id].end(), d->id) !=
-                    Sol::TabuFleet[del->id].end())
+                if(d->overTime >= Parameters::MAX_OVERTIME){
                     continue;
+                }
                 insrmv.cancel = false;
                 Move<Delivery, Driver, MoveVrp> m;
                 InsertCost(s, del, d, &l_moves);
@@ -40,11 +40,10 @@ public:
 
         for (int delID: listId) {
             Delivery *del = s.GetDelivery(delID);
-            assert(del != nullptr);
             for (Driver *d: driversList) {
-                if (std::find(Sol::TabuFleet[del->id].begin(), Sol::TabuFleet[del->id].end(), d->id) !=
-                    Sol::TabuFleet[del->id].end())
+                if(d->overTime >= Parameters::MAX_OVERTIME){
                     continue;
+                }
                 insrmv.cancel = false;
                 Move<Delivery, Driver, MoveVrp> m;
                 InsertCost(s, del, d, m);
@@ -68,10 +67,6 @@ public:
                     std::min(d->capacity, s.orderCapRestante[n->orderID]) : Sol::FixLoad[n->delID];
 
         if (mo.demand > d->capacity) return;
-        if (mo.demand <= 0) {
-            Prompt::print({std::to_string(Sol::FixLoad[n->delID]), std::to_string(mo.demand)});
-        }
-        assert(mo.demand > 0);
 
         Dock *dock = s.GetDock(n->dockID);
         Node *prev = s.CustomerPrev[n->EndNodeID];
@@ -102,11 +97,6 @@ public:
                     std::min(d->capacity, s.orderCapRestante[n->orderID]) : Sol::FixLoad[n->delID];
 
         if (mo.demand > d->capacity) return;
-        if (mo.demand <= 0) {
-            Prompt::print({std::to_string(Sol::FixLoad[n->delID]), std::to_string(mo.demand)
-                          });
-        }
-        assert(mo.demand > 0);
 
         Dock *dock = s.GetDock(n->dockID);
         Cost newcost(false);
@@ -138,13 +128,7 @@ public:
             insrmv.expected_del_time = s.EarlyTW(n);
[…]

### 2023-08-02 codeCQ/heur/InsRmvMethodFast.cpp

@@ -14,6 +14,9 @@ InsRmvMethodFast::GetBestInsertion(Sol &s, const std::vector<int> &listId, const
         Delivery *del = s.GetDelivery(delID);
         assert(del != nullptr);
         for (Driver *d: driversList) {
+            if(d->overTime >= Parameters::MAX_OVERTIME){
+                continue;
+            }
             cancel = false;
             Move<Delivery, Driver, MoveVrp> m;
             InsertCost(s, del, d, m);
@@ -92,7 +95,6 @@ void InsRmvMethodFast::InsertCost(Sol &s, Delivery *n, Driver *d,
     }
     mo = GetCost(s, n, d, newcost, mo.demand, temp_moves);
 }
-
 
 void InsRmvMethodFast::SetServiceParams(Sol &s, Delivery *n, Driver *d, double demand) {
     Depot *dep = s.GetDepot(n->depotID);
@@ -329,6 +331,13 @@ Move<Delivery, Driver, MoveVrp> InsRmvMethodFast::GetCost(Sol &s, Delivery *n, D
             continue;
         }
         node_arr += UNLOADING_DURATION;
+        double overtime = Sol::GetOvertimeCost(node_arr + Parameters::CLEANING_DURATION + s.Travel(n->distID, d->distID)
+                                               - d->start_shift_time);
+        if ( overtime  > Parameters::MAX_OVERTIME )
+        {
+            prev = s.DriverNext[prev->id];
+            continue;
+        }
 
         if (node_arr > s.LateTW(n)) {
             if (prev->type == Parameters::DELIVERY) {
@@ -455,6 +464,8 @@ Move<Delivery, Driver, MoveVrp> InsRmvMethodFast::GetCost(Sol &s, Delivery *n, D
         newcost.overTimeCost =
                 newcost.overTimeCost - d->overTime + Sol::GetOvertimeCost(node_arr + s.Travel(n->distID, d->distID)
                                                                           - d->start_shift_time);
+
+
         newcost.undeliveredCost -= Parameters::UNDELIVERY_PENALTY * m.demand;
 
         if (Parameters::SHOW) {

### 2023-08-02 codeCQ/Data.cpp
Save Cq, MAx_OVERTIME

@@ -135,7 +135,6 @@ void Data::LoadInstance() {
                       d.capacity >> d.start_shift_time;
             d.id = i;
             if (d.rank == -1) d.rank = MAX_RANK;
-//            d.start_shift_time = 260;
             Depot *dep = GetDepot(d.depotID);
             d.distID = dep->distID;
 
@@ -178,7 +177,6 @@ void Data::LoadCQInstance() {
                       d.capacity >> d.start_shift_time;
             d.id = i;
             if (d.rank == -1) d.rank = MAX_RANK;
-            d.start_shift_time = 260;
             Node d1;
             d1.id = GetNodeCount();
             d1.no = d1.id + 1;
@@ -216,7 +214,7 @@ void Data::LoadCQInstance() {
         for (int i = 0; i < this->nbCustomers; ++i) {
             Customer n;
             n.c = 'C';
-            inputFile >> n.custID >> n.orderNbr >> dummy >> n.distID >> n.depotID >> n.early_tw >> n.demand
+            inputFile >> n.custID >> n.orderNbr >> dummy >> n.distID >> dummy >> n.early_tw >> n.demand
                       >> n.nbOrder;
             n.custID = i;
             n.constID = n.custID;
@@ -260,7 +258,7 @@ void Data::LoadCQInstance() {
 
         for (int i = 0; i < this->nbOrders; ++i) {
             Order o;
-            inputFile >> o.orderID >> o.custID >> o.demand >> dummy;
+            inputFile >> o.orderID >> o.custID >> o.demand >>  o.depotID;
             o.orderID = GetOrderCount();
             AddOrder(o);
         }
@@ -543,6 +541,10 @@ Depot *Data::GetDepot(int i) {
 }
 
 Depot *Data::GetDepot(Node *n) {
+    if(n->type == Parameters::CUSTOMER){
+        printf(" not defined for customer\n");
+        exit(1);
+    }
     return GetDepot(n->depotID);
 }
 
@@ -622,7 +624,7 @@ void Data::AddDeliveryNodes(Customer *c) {
 
 void Data::AddDeliveryNodes(Order *o, int nb) {
     Customer *c = GetCustomer(o->custID);
-    assert(c->depotID != -1);
+    assert(o->depotID != -1);
     for (int j = 0; j < nb; ++j) {
         Delivery del;
         del.id = GetNodeCount();
@@ -637,6 +639,7 @@ void Data::AddDeliveryNodes(Order *o, int nb) {
[…]

### 2023-08-02 codeCQ/Data.h

@@ -159,13 +159,25 @@ public:
     }
 
     double Travel(Customer *from, Customer *to) {
-        Depot *dep_to = GetDepot(to->depotID);
-        return Travel(from, dep_to) + Travel(dep_to, to);
+        double _min = INFINI;
+        for (auto to_order: GetOrders(to)) {
+            Depot *to_dep = GetDepot(to_order->depotID);
+            if (_min > Travel(from->distID, to_dep->distID) + Travel(to_dep->distID, to->distID)) {
+                _min = Travel(from->distID, to_dep->distID) + Travel(to_dep->distID, to->distID);
+            }
+        }
+        return _min;
     }
 
     double Distance(Customer *from, Customer *to) {
-        Depot *dep_to = GetDepot(to->depotID);
-        return Distance(from, dep_to) + Distance(dep_to, to);
+        double _min = INFINI;
+        for (auto to_order: GetOrders(to)) {
+            Depot *to_dep = GetDepot(to_order->depotID);
+            if (_min > Distance(from->distID, to_dep->distID) + Distance(to_dep->distID, to->distID)) {
+                _min = Distance(from->distID, to_dep->distID) + Distance(to_dep->distID, to->distID);
+            }
+        }
+        return _min;
     }
 
     double Travel(int from, int to) {

### 2023-08-02 codeCQ/Node.h

@@ -115,18 +115,19 @@ public:
     int service_duration;
     int nbDelMin, nbDelMax;
     int orderID;
+    int depotID;
     double demand;
     int early_tw, late_tw;
 
     Order() : orderID(-1), demand(0), custID(-1), nbDelMin(0), nbDelMax(0), service_duration(0),
-              early_tw(0), late_tw(0) {}
+              early_tw(0), late_tw(0),depotID(-1) {}
 
 public:
     friend std::ostream &operator<<(std::ostream &os, const Order &order) {
         os << "ooooooooooooooooooooooooooooooooooooooooooooooooooooooooo\n";
         os << "Order: demand: " << order.demand << " custID: " << order.custID << " nbDelMin: " << order.nbDelMin
            << " nbDelMax: " << order.nbDelMax << " orderID: " << order.orderID << " [" << order.early_tw << "-"
-           << order.late_tw << "]  |\n";
+           << order.late_tw << "]  | depotID" <<order.depotID << "\n";
         os << "ooooooooooooooooooooooooooooooooooooooooooooooooooooooooo";
         return os;
     }

### 2023-08-02 codeCQ/Parameters.cpp

@@ -6,6 +6,8 @@ long Parameters::RUNTIME = 300;
 int Parameters::MAX_LATE_TW = 1600;//1440;
 int Parameters::TW_WIDTH = 60;
 int Parameters::MAX_WORKING_TIME = 600;
+int Parameters::MAX_OVERTIME = 120;
+
 
 double Parameters::UNLOADING_RATE = 30; //30 m3/h
 

### 2023-08-02 Article/article.tex
save article

@@
 Additionally, the problem also considers the possibility of customers requiring multiple types of concrete to be delivered within the same time window.
-The performance of the proposed heuristic is evaluated on real-world and generated CDP instances, and it is compared to another variant to demonstrate its effectiveness.
+The performance of the proposed heuristic is evaluated on real-world instances, and it is compared to another variant to demonstrate its effectiveness.
 
@@
 In this paper, we study a variant of the CDP to schedule the daily production and dispatching of RMC for a company located in the province of Quebec, Canada.
-This company operates multiple batching plants with varying production rates, using a fleet of concrete mixers of different capacities.
+This company operates multiple batch plants with varying production rates, using a fleet of concrete mixers of different capacities.
 Each plant has its own fleet of trucks; however, under certain conditions, the trucks can move between plants if necessary.
@@
 
-Academic Research on concrete batching and delivery began in the late 1990s.
+Academic Research on concrete batch and delivery began in the late 1990s.
 \cite{tommelein1999just} described RMC as a prototypical example of a JIT production system in construction and identified two practices for delivering it.
@@
 \cite{mayteekrieangkrai2015optimized} addressed the same problem with the same data using a bee algorithm (BA) and found better solutions than the GA.
-\cite{lu2005optimized} used the same combination of GA and simulation to determine the optimal number of concrete mixers to be deployed and an optimal schedule for batching and delivering concrete.
+\cite{lu2005optimized} used the same combination of GA and simulation to determine the optimal number of concrete mixers to be deployed and an optimal schedule for batch and delivering concrete.
 Their objective was to minimize the idle time of the site crew due to late concrete deliveries and truck queuing time.
 In this setting, it was also necessary to deliver a batch of mortar on-site to lubricate the unloading pump before the concrete delivery.
-As such, the simulation model also included the batching and delivery of mortar.
+As such, the simulation model also included the batch and delivery of mortar.
 Finding the best RMC fleet size was also the purpose of the discrete-event simulation model proposed by \cite{panas_simulation_based_2013}.
@@
 They ran experiments using real-world instances of a concrete supply chain in the Netherlands and found a reduction in the number of outsourced requests.
-\cite{yan2007optimal} also considered overtime considerations in their paper, which focused on scheduling RMC for one batching plant with two loading docks.
+\cite{yan2007optimal} also considered overtime considerations in their paper, which focused on scheduling RMC for one batch plant with two loading docks.
 The study took into account that overtime wages are paid for factory and construction site operations after 4 PM.
@@
 The procedure is invoked multiple times for different permutations of the customer's order which is determined using the steepest descent (SD) local search procedure.
-One of the paper's main contributions is the creation of the first public test instances for the CDP with up to 50 customers, four batching plants, and 20 concrete mixers.
+One of the paper's main contributions is the creation of the first public test instances for the CDP with up to 50 customers, four batch plants, and 20 concrete mixers.
 They found the CP model to be highly effective in finding high-quality solutions in a relatively short time or improving existing schedules, while the MIP model can be used to compute bounds, as it seems ineffective in solving large problem instances.
@@
 The focus of this paper is on the distribution of RMC from a Canadian company that operates in the greater Montreal area.
-When a customer places an order, it is received at a control center and assigned to one of the company's batching plants.
+When a customer places an order, it is received at a control center and immediately assigned to one of the company's batch plants.
 These plants produce the concrete and then deliver it to the customer.
-The problem we are examining involves a set of customer orders, a set of concrete-mixer drivers, and a set of batching plants.
+The problem we are examining involves a set of customer orders, a set of concrete-mixer drivers, and a set of batch plants.
 
@@
 
-A customer $i$ requests one or more types of concrete to be delivered to their  construction site on a particular day, starting at time $a_i$.
+A customer $i$ requests one or more types of concrete to be delivered to their construction site on a specific day, with the delivery service starting at the due time $a_i$.
 We call an order $o$ a request for a specific type of concrete.
@@
 Each element of $O_i$ must be fully delivered before moving on to another order.
-Exactly one order $o$ of $O_i$ must have its first delivery start at $a_i$, while the others can start at any time after $o$ is completed.
-Once a plant is selected to produce a customer's first delivery, it must be the supplier of all subsequent deliveries from that customer.
-To avoid cold joint problems with the concrete, the subsequent deliveries must be made in close succession.
-We define a maximum time delay $\gamma_i$ after which no more deliveries are allowed.
+Exactly one order $o \in O_i$ must have its first delivery start at $a_i$, while the others can start at most $\gamma^2$ time after $o$ is completed.
+%Once a plant is selected to produce a customer's first delivery, it must be the supplier of all subsequent deliveries from that customer.
+To avoid cold joint problems with the concrete, subsequent deliveries of the same order must be made in close succession.
+We define a maximum time delay $\gamma^1$ after which no more deliveries are allowed.
 The customer unloading rate and the quantity to be unloaded give the time required to unload a truckload.
@@
 A scheduled driver $k$ is required to start his shift at $H_k$, work a minimum of $M_T$ hours and a maximum of $N_T$ hours during regular working hours, with the possibility of overtime of up to $O_T$ hours.
-$\beta_3$ and $\beta_4$ are the penalties incurred if a driver works less than $M_t$ and more than $N_t$.
+$\beta_3$ and $\beta_4$ are the penalties for a driver working less than $M_T$ and more than $N_T$.
 
@@
 
-The company owns a total of eight batching plants.
-Each plant has a loading dock that can accommodate only one truck at a time, which leads to trucks lining up.
-Let $\mathcal{B}$ be the set of batching plants.
+Each of the company's plants has a single loading dock that can accommodate only one truck at a time.
+As a result, trucks often form a queue while waiting for their turn at the loading dock.
+Let $\mathcal{B}$ be the set of batch plants.
 The plants are heterogeneous, as each plant $b$ has its own hourly loading rate, represented by $\tau^l_b$, which affects the duration of the loading process.
@@
 Each plant has its own assigned fleet of trucks, but it can borrow trucks from other plants.
-% or hire external vehicles if necessary.
- A batching plant can serve a construction site as long as the travel time between the two is less than the concrete's lifespan, which is represented by $\Delta$.
-Let $l_{bj}$ be the loading dock node associated with delivery node $j$ at plant $b$.
-We represent each plant $b$ by the set $\mathcal{L}_b = \{l_{bj}, t_{bj}\leq  \Delta, j \in \mathcal{D} \}$  of loading docks nodes.
-$\mathcal{L}=\bigcup_{b\in B} \mathcal{L}_b$ is the union of all loading docks.
+Let $l_{j}$ be the loading dock node associated with delivery node $j$.
+After loading RMC at $l_j$, it must be fully delivered to $j$ at most before $\Delta$ time, which is the concrete lifetime.
+We denote $\mathcal{L}_b$ as the set of loading dock nodes of plant $b$ and $\mathcal{L}$ as the set of all loading docks.
 
@@
 A solution to the problem involves decisions about truck loading schedules, driver assignments to different deliveries, and truck arrival times at construction sites for unloading.
-For a batching plant, the decision involves choosing which driver to \textcolor{red}{load}, when to load them, and which construction site they should deliver to.
-For a driver, the decision involves determining the sequence of loading depots and delivery sites.
+For a batch plant, the decision involves choosing which driver to load, when to load them, how much to load, and which construction site to deliver.
+For a driver, the decision is to determine the sequence of loading depots and delivery sites.
 And for a construction site, the decision involves determining the arrival times of all scheduled deliveries for the day.
@@
 For a customer $c$, arcs in $A^D$ link consecutive delivery nodes of the same order   $\lbrace (i,j)\in \mathcal{D}_o, o \in \mathcal{O}_c, i < j  \rbrace$, and pair of delivery nodes of two different orders $ \lbrace (i,d^{0}_{o_2}),  i \in \mathcal{D}_{o_1}, i \geq n^{min}_{o_1}, o_1, o_2 \in \mathcal{O}_c, o_1 \neq o_2 \rbrace $.
-Arcs in $A^L$ link all pairs of loading docks of the same batching plant.
+Arcs in $A^L$ link all pairs of loading docks of the same batch plant.
 
@@
 Let the binary variable $x^{k}_{ij}$ be $1$ if driver $k$ travels from node $i$ to $j$.
-Binary variable $y_o$ is $1$ when order $o$ is completely served; $v_i$ and $w_i$ are the start and end of the loading (unloading) operation at node $i \in \mathcal{L} \cup \mathcal{D}$.
-The binary variable $u_{ij}$ is $1$ if node $j$ is served just after $i$, the service being either an unloading or a loading operation.
-Binary variable $\sigma_{ib}$ is equal to 1 if the orders of customer $i$ are loaded by plant $b$.
+Binary variable $y_o$ is $1$ when order $o$ is completely served; $v_i$ and $w_i$ are the start and end of the loading (unloading) operation at node $i \in \mathcal{L}\cup \mathcal{D}$.
+Binary variable $u_{ij}$ is $1$ if node $j$ is served just after $i$, the service being either an unloading or a loading operation.
+% Binary variable $\sigma_{ib}$ is equal to 1 if the orders of customer $i$ are loaded by plant $b$.
 Variable $q^k_j$ is the quantity to be loaded towards $j$ with vehicle $k$.
@@
 \begin{alignat}{3}
-\label{mod:obj}   \min &  \smashoperator{ \sum_{ \substack{(i,j,k) \in A} }} {t_{ij}x^{k}_{ij}} + \beta_1\smashoperator{\sum_{ o \in \mathcal{O}}}{  \left(1-y_o\right)} \hspace{-0mm}  + \beta_2 \smashoperator{ \sum_{i \in \mathcal{C} } } { g_i  } +  \sum_{k \in K}{\beta_3 w^{1}_k } + \beta_4 { w^{2}_k }&\\
-    \label{mod:c1} &\sum_{j \in \delta^{+}(s_k) }{x^k_{s_kj}} = 1   &  \forall k \in K  \\
-    \label{mod:c2} &\sum_{j \in \delta^{-}(e_k)}{x^k_{je_k}} = 1   &  \forall k \in K\\
-    \label{mod:c3} & v_j \geq  w_i + \alpha_{i} + t_{ij} - M\left(1-x^{k}_{ij}\right)  & \hspace{-20mm} \forall i \in \mathcal{L}, j \in \delta^{+}(i),  k \in K \\
-    \label{mod:c4}  & v_j \geq  w_i + Ct_{k} + t_{ij} - M\left(1-x^{k}_{ij}\right) & \hspace{-20mm} \forall i \in \mathcal{D}, j \in \delta^{+}(i), k \in K \\
+    \label{mod:obj}   \min & \smashoperator{ \sum_{ \substack{(i,j,k) \in A} }} {t_{ij}x^{k}_{ij}} + \beta_1\smashoperator{\sum_{ o \in \mathcal{O}}}{  \left(1-y_o\right)} \hspace{-0mm}  + \beta_2 \smashoperator{ \sum_{i \in \mathcal{C} } } { g_i  } +  \sum_{k \in K}{\beta_3 w^{1}_k } + \beta_4 { w^{2}_k } &                                                                                                  \\
+    \label{mod:c1}         & \sum_{j \in \delta^{+}(i) }{x^k_{ij}} =1                                                                                                                                                                                                                                               & \forall i \in S, k \in K                                                                         \\
+    \label{mod:c2}         & \sum_{j \in \delta^{-}(i)}{x^k_{ji}} = 1                                                                                                                                                                                                                                               & \forall i \in E, k \in K                                                                         \\
+    \label{mod:c03}        & v_j \geq  w_i  - M\left(1-x^{k}_{ij}\right)                                                                                                                                                                                                                                            & \hspace{-20mm}  i \in S, j \in \delta^{+}(i),  k \in K                                           \\
+    \label{mod:c3}         & v_j \geq  w_i + \alpha_{b} + t_{ij} - M\left(1-x^{k}_{ij}\right)                                                                                                                                                                                                                       & \hspace{-20mm} \forall b \in \mathcal{B}, i \in \mathcal{L}_b, j \in \delta^{+}(i),  k \in K     \\
+    \label{mod:c4}         & v_j \geq  w_i + Ct_{k} + t_{ij} - M\left(1-x^{k}_{ij}\right)                                                                                                                                                                                                                           & \hspace{-20mm} \forall i \in \mathcal{D}  , j \in \delta^{+}(i), k \in K                         \\
     \label{mod:c5}         & w_{i} \geq v_{i}  + \frac{q^k_j}{\tau^l_b} -M\left(1- x^{k}_{ij}\right)                                                                                                                                                                                                                & \hspace{-20mm} \forall  b \in \mathcal{B},  i \in \mathcal{L}_{b},  j \in \delta^{+}(i), k \in K \\
@@
     \label{mod:c7}         & w_{j} \leq v_{i}  + \Delta + M\left(1- x^{k}_{ij}\right)                                                                                                                                                                                                                               & \hspace{-20mm}   j \in \mathcal{D},  i \in \delta^{-}(j),  k \in K                               \\
-    \label{mod:c8}  &  v_{d^0_{o}} \geq a_i  & \hspace{-20mm} \forall  i \in \mathcal{C}, \forall o \in \mathcal{O}_i\\
-    \label{mod:c9} & g_{i} \geq v_{d^0_{o_1}} - a_i - M\left(  \sum_{ j \in \delta^{-}_{D}(d^0_{o_1})  }{u_{j,d^0_{o_1}}}\right)  & \hspace{-20mm}  \forall  i \in \mathcal{C}, \forall o_1 \in \mathcal{O}_i, d=d^0_{o_1} \\
-     \label{mod:c101} & v_{d^0_{o_1}} \geq w_j - M\left(1- u_{j,d^0_{o_1}}\right)  & \hspace{-40mm}    \forall o_1 \in \mathcal{O}_i,  j \in \delta_{D}^{-}(d^0_{o_1})  \\
-  \label{mod:c10} &  v_{d^0_{o_1}}  \leq  w_j + \gamma_i + M\left(1- u_{j,d^0_{o_1}}\right) &  \hspace{-20mm}  \forall  o_1 \in \mathcal{O},  j \in \delta_{D}^{-}(d^0_{o_1}) \\
+    \label{mod:c8}         & v_{d^0_{o}} \geq a_i                                                                                                                                                                                                                                                                   & \hspace{-20mm} \forall  i \in \mathcal{C}, \forall o \in \mathcal{O}_i                           \\
+    \label{mod:c9}         & g_{i} \geq v_{d^0_{o_1}} - a_i - M\left(  \sum_{ j \in \delta^{-}_{D}(d^0_{o_1})  }{u_{jd^0_{o_1}}}\right)                                                                                                                                                                             & \hspace{-20mm}  \forall  i \in \mathcal{C}, \forall o_1 \in \mathcal{O}_i                        \\
+    \label{mod:c101}       & v_{d^0_{o_1}} \geq w_j - M\left(1- u_{jd^0_{o_1}}\right)                                                                                                                                                                                                                               & \hspace{-40mm}    \forall o_1 \in \mathcal{O}_i,  j \in \delta_{D}^{-}(d^0_{o_1})                \\
+    \label{mod:c10}        & v_{d^0_{o_1}}  \leq  w_j + \gamma^2 + M\left(1- u_{j,d^0_{o_1}}\right)                                                                                                                                                                                                                 & \hspace{-20mm}  \forall  o_1 \in \mathcal{O},  j \in \delta_{D}^{-}(d^0_{o_1})                   \\
     \label{mod:c11}        & \sum_{ \substack{o_1 \in \mathcal{O}_i }}{ \sum_{ j \in \delta^{-}_{D}(d^0_{o_1})}}{u_{j,d^0_{o_1} } } = |\mathcal{O}_i|-1                                                                                                                                                             & \hspace{-20mm} \forall  i \in \mathcal{C},|\mathcal{O}_i| > 1                                    \\
-    \label{mod:c12}&   \sum_{ \substack{o_1 \in \mathcal{O}_i }}\sum_{ j \in \delta^{+}_{D}(d^0_{o_1})}{u_{d^0_{o_1},j } } = |\mathcal{O}_i|-1 & \hspace{-40mm} \forall  i \in \mathcal{C}, |\mathcal{O}_i| > 1 \\
+    \label{mod:c12}        & \sum_{ \substack{o_1 \in \mathcal{O}_i }}\sum_{ j \in \delta^{+}_{D}(d^0_{o_1})}{u_{d^0_{o_1},j } } = |\mathcal{O}_i|-1                                                                                                                                                                & \hspace{-40mm} \forall  i \in \mathcal{C}, |\mathcal{O}_i| > 1                                   \\
     \label{mod:c13}        & \hspace{3mm} \smashoperator{ \sum_{ j \in \delta^{+}_{D}(d^0_{o_1})}}{u_{d^0_{o_1},j } } \leq 1                                                                                                                                                                                        & \hspace{-20mm}  \forall o_1 \in \mathcal{O}                                                      \\
@@
     \label{mod:c16}        & v_{j} \geq w_{j-1} - M\left(1- u_{j-1,j}\right)                                                                                                                                                                                                                                        & \hspace{-20mm}  \forall o \in \mathcal{O}, j \in \mathcal{D}_{o}, j \geq 1                       \\
-  \label{mod:c17} &  v_{j}  \leq  w_{j-1} + \gamma_i + M\left(1- u_{j-1,j}\right) & \hspace{-20mm} \forall i \in \mathcal{C}, \forall o \in O_i, j \in \mathcal{D}_{o}, j \geq 1 \\
-   \label{mod:c18}   &   u_{j-1,j} \geq u_{j,j+1} & \hspace{-35mm} \forall o_1 \in\mathcal{O},  j \in \mathcal{D}_{o_1}, 1 \leq j \leq n_{o_{1}}-1\\
-  \label{mod:c19} &  u_{j-1,j} \geq \sum_{l \in \mathcal{L}}{x_{lj}} &\hspace{-20mm}  \forall o_1 \in \mathcal{O}, j \in \mathcal{D}_{o_1}, j \geq 1 \\
+    \label{mod:c17}        & v_{j}  \leq  w_{j-1} + \gamma^1 + M\left(1- u_{j-1,j}\right)                                                                                                                                                                                                                           & \hspace{-20mm} \forall i \in \mathcal{C}, \forall o \in O_i, j \in \mathcal{D}_{o}, j \geq 1     \\
+    \label{mod:c18}        & u_{j-1,j} \geq u_{j,j+1}                                                                                                                                                                                                                                                               & \hspace{-35mm} \forall o_1 \in\mathcal{O},  j \in \mathcal{D}_{o_1}, 1 \leq j \leq n_{o_{1}}-1   \\
+    \label{mod:c19}        & u_{j-1,j} \geq \sum_{l \in \mathcal{L}}{x_{lj}}                                                                                                                                                                                                                                        & \hspace{-20mm}  \forall o_1 \in \mathcal{O}, j \in \mathcal{D}_{o_1}, j \geq 1                   \\
     \label{mod:c20}        & v_{j} \geq w_{i} - M\left(1- u_{i,j}\right)                                                                                                                                                                                                                                            & \hspace{-20mm}  \forall i \in \mathcal{L}, j \in \delta^{+}_{L}(i)                               \\
@@
     \label{mod:c22}        & \hspace{2mm} \smashoperator{ \sum_{ j \in \delta^{-}_{L}(i)}}{u_{j,i } } \leq 1                                                                                                                                                                                                        & \hspace{-20mm} \forall i \in \mathcal{L}                                                         \\
+    \label{mod:c230}       & \hspace{2mm} \smashoperator{ \sum_{ j \in \delta^{-}_{L}(i)}}{u_{j,i } }  \geq \sum_{k \in K}\sum_{j \in  \delta^{-}(i) }{x^k_{ji}}                                                                                                                                                    & \hspace{-20mm} \forall i \in \mathcal{L}                                                         \\
+    \label{mod:c231}       & \hspace{2mm} \smashoperator{ \sum_{ j \in \delta^{-}_{L}(i)}}{u_{j,i } } \geq \smashoperator{ \sum_{ j \in \delta^{-}_{L}(i+1)}}{u_{j,i+1 } }                                                                                                                                          & \hspace{-20mm} \forall i \in \mathcal{L}                                                         \\
     \label{mod:c23}        & \hspace{2mm} \smashoperator{ \sum_{ j \in \delta^{-}_{L}(i)}}{u_{j,i } } + \smashoperator{ \sum_{ j \in \delta^{+}_{L}(i)}}{u_{i,j } } \geq \sum_{k \in K}\sum_{j \in  \delta^{-}(i) }{x^k_{ji}}                                                                                       & \hspace{-20mm} \forall i \in \mathcal{L}                                                         \\
     \label{mod:c24}        & \sum_{k \in K}\sum_{j \in \mathcal{D}_o}{q^k_j} = q_o                                                                                                                                                                                                                                  & \hspace{-20mm}  \forall o \in \mathcal{O}                                                        \\
+    \label{mod:c240}       & q^k_j \leq \sum_{i \in  \delta^{-}(j)}{Q^k x^k_{ij}}                                                                                                                                                                                                                                   & \hspace{-20mm}  \forall j \in \mathcal{D}, k \in K                                               \\
     \label{mod:c25}        & \sum_{k \in K}\sum_{j \in \delta^{+}(i)}{x^{k}_{ij} \leq 1}                                                                                                                                                                                                                            & i \in \mathcal{L} \cup \mathcal{D}                                                               \\
     \label{mod:c26}        & \sum_{k \in K}\sum_{j \in \delta^{+}(i)} {x^{k}_{ij} } =  \sum_{k \in K}\sum_{j \in \delta^{-}(i)}{x^{k}_{ji} }                                                                                                                                                                        & i \in \mathcal{L} \cup \mathcal{D}                                                               \\
-  \label{mod:c27} & \sum_{k \in K}  \sum_{i \in \mathcal{D}_c} \sum_{j \in \delta^{-}(i) }  {x^k_{ji}} \leq M\sigma_{c, b} & \forall c \in \mathcal{C}, b \in \mathcal{B} \\
-    \label{mod:c28} & \sum_{b \in \mathcal{B}} \sigma_{c, b} \leq 1 & \hspace{-20mm}\forall c \in \mathcal{C}, b \in \mathcal{B} \\
+    % \label{mod:c27}        & \sum_{k \in K}  \sum_{i \in \mathcal{D}_c} \sum_{j \in \delta^{-}(i) }  {x^k_{ji}} \leq M\sigma_{c, b}     & \forall c \in \mathcal{C}, b \in \mathcal{B}                                                     \\
+    % \label{mod:c28}        & \sum_{b \in \mathcal{B}} \sigma_{c, b} \leq 1    & \hspace{-20mm}\forall c \in \mathcal{C}, b \in \mathcal{B}                                       \\
     \label{mod:c29}        & w^1_k \geq  M_T + H_k - v_{e_k}                                                                                                                                                                                                                                                        & \hspace{-20mm} \forall k \in K                                                                   \\
@@
     \label{mod:c34}        & u_{ij} \in \{0,1\}                                                                                                                                                                                                                                                                     & \hspace{-20mm}  (i,j) \in A^D , k \in K  \hspace{0mm}                                            \\
-    \label{mod:c35} &  \sigma_{ib} \in \{0,1\} & \hspace{-20mm}  i \in \mathcal{C}, b \in \mathcal{B} \\
+    % \label{mod:c35}        & \sigma_{ib} \in \{0,1\}    & \hspace{-20mm}  i \in \mathcal{C}, b \in \mathcal{B}                                             \\
     \label{mod:c36}        & y_o \in \{0,1\}                                                                                                                                                                                                                                                                        & \hspace{-20mm}  o \in \mathcal{O}                                                                \\
-    \label{mod:c37} &  v_i \geq 0,  w_i \geq 0 & \hspace{-20mm}  i \in V \\\
+    \label{mod:c37}        & v_i \geq 0,  w_i \geq 0                                                                                                                                                                                                                                                                & \hspace{-20mm}  i \in V                                                                          \\
     \label{mod:c38}        & w^1_{k} \geq 0,  w^2_k \geq 0                                                                                                                                                                                                                                                          & \hspace{-20mm}  i \in V                                                                          \\
@@
 $\beta_1$ to $\beta_4$ are the penalty coefficients of each component of the objective function (\ref{mod:obj}).
-Constraints (\ref{mod:c1}) and (\ref{mod:c2}) state that driver $k$ leaves his start node exactly once a day and returns to his end node after completing his last unloading operation of the day.
+Constraints (\ref{mod:c1}) and (\ref{mod:c2}) state that driver $k$ leaves his start node exactly once a day and returns to his end node.
+A driver cannot serve a node before the start of his shift with constraints (\ref{mod:c03}).
 Constraints (\ref{mod:c3}) and (\ref{mod:c4}) set a driver to take some time after loading or unloading to adjust the concrete or clean the truck before moving on to the next node.
@@
 Similarly, constraints (\ref{mod:c20}) - - (\ref{mod:c23}) ensure that two trucks cannot be loaded at the same time at a plant.
-Constraints (\ref{mod:c24}) require that the cumulative load of all concrete mixers serving an order must equal the required quantities.
+Constraints (\ref{mod:c24}) require that the cumulative load of all concrete mixers serving an order must equal the required quantities, and (\ref{mod:c240}) bound a driver load.
 Constraints (\ref{mod:c25}) require that a driver can only visit a loading/delivery node once.
 Constraints (\ref{mod:c26}) are degree constraints.
-Constraints (\ref{mod:c27}) and (\ref{mod:c28}) require that all orders from customer $i$ must be loaded from the same plant.
-Constraints (\ref{mod:c29}--\ref{mod:c30}) calculate the difference between a driver's hours of service and the minimum and normal hours of service.
-Finally, constraints (\ref{mod:c31}) -- (\ref{mod:cend}) define the nature and bounds of the variables.
+Constraints (\ref{mod:c29}) - - (\ref{mod:c30}) calculate the difference between a driver's hours of service and the minimum and normal hours of service.
+Finally, constraints (\ref{mod:c31}) - - (\ref{mod:cend}) define the nature and bounds of the variables.
 
@@
 Let us consider the solution of a small instance of our problem shown in Figures~\ref{fig_Example} and \ref{fig:ganttExample}.
-Two batching plants $B_1$ and $B_2$ with three drivers serve two construction sites.
+Two batch plants $B_1$ and $B_2$ with three drivers serve two construction sites.
 Customer $C_1$ requests one order of $q^1_1=12$ m$^3$.
@@
 
-\begin{figure}[!ht]
+\begin{figure}[H]
     % \caption{Solution of an instance with two plants, two construction sites, four orders, and three drivers.}
@@
 
-\begin{figure}[!htb]
+\begin{figure}[H]
     \centering
@@
 
-We now present the heuristic approach we implement to solve this variant of the Concrete Delivery Problem.
-Our method consists of constructing feasible solutions to the CDP with randomized heuristics and iteratively invoking these heuristics in the GRASP metaheuristic.
-
-In this section, we represent the scheduling of a delivery node $d$ as a pair of loading ($L_{d}$) and unloading ($U_{d}$) tasks.
-$L_{d} = \left(b,k,q_d,v_b\right) $ indicates that driver $k$ starts loading $q_d$ $m^3$ of RMC at plant $b$ at time $v_b$.
-$U_{d} = \left(k,v_d, w_d\right)$ indicates that driver $k$ is serving $d$ between $v_d$ and $w_d$.
-We then call a solution $ S =\left\lbrace \cup _{1 \leq j \leq n_o} (L_{d^j_{o}}, U_{d^j_{o}}), \forall o \in \mathcal{O} \right\rbrace$ the set of all loading and unloading tasks performed on a given day.
+In this section, we present the heuristic approach we implement to solve this variant of the Concrete Delivery Problem.
+Our method consists of constructing feasible solutions to the CDP with randomized greedy heuristics and iteratively invoking these heuristics in the GRASP metaheuristic.
+From now on, we represent the scheduling of a delivery node $d$ as the pair  ($L^k_{d}$,$U^k_{d}$) of loading  and unloading  tasks.
+$L^k_{d} = \left(b,q_d,v_b,w_b\right)$ indicates that driver $k$ starts loading $q_d$ $m^3$ of RMC at plant $b$ between $\left[v_b,w_b\right]$.
+$U^k_{d} = \left(k,v_d, w_d\right)$ indicates that driver $k$ serves $d$ between $v_d$ and $w_d$.
+We then call a solution $ S =\left\lbrace \cup _{1 \leq j \leq |\mathcal{D}|} (L^k_{j}, U^k_{j}), k \in K \right\rbrace$ the set of all loading and unloading tasks performed on a given day.
 
@@
 
-The GRASP algorithm is an iterative suite of constructive and local search algorithms introduced by \cite{feo1995greedy}.
-A feasible solution $S$ is constructed using a greedy randomized algorithm at each iteration.
-Then a local search algorithm explores the neighborhood of $S$ to find a local optimum.
-The pseudocode of the algorithm~\ref{alg_grasp} shows that the procedure returns the best overall solution (for a minimization problem) after some stopping conditions (time limit or a maximum number of iterations) are met.
-In our implementation, we introduce at line 13 an additional local search step using the solutions found by all heuristics in the current iteration.
+The GRASP algorithm, proposed by \cite{feo1989probabilistic}, is a two-phase iterative suite consisting of constructive and local search algorithms.
+In the construction phase, a feasible solution $S$ is generated using a greedy randomized algorithm.
+Then, a local search algorithm explores the neighborhood of $S$ to identify a local optimum.
+Algorithm~\ref{alg_grasp} outlines the pseudocode of the GRASP algorithm.
+To avoid redundant computations, a hash table $H$ is used to ensure that the local search is not applied multiple times to the same solution.
+The procedure stops and returns the best overall solution when certain stopping conditions, such as reaching a time limit or a maximum number of iterations, are met.
+In our implementation, we also add a path relinking procedure that uses information about previous solutions to improve the search process.
 
@@
                 \setcounter{AlgoLine}{0}
-            \KwIn{ \textit{H}: List of constructive randomized heuristics }
+                % \KwIn{ \textit{H}: List of constructive greedy randomized heuristics }
+                \KwIn{ $\phi$: number of iterations before path relinking }
                 $S^* \leftarrow \emptyset$    \hspace{2mm} $T \leftarrow \emptyset$       $Cost(S^*) \leftarrow \infty$
 
+                Initialize an empty hash table $H$
+
+                $counter \leftarrow  0$
+
                 \While{Conditions not met}{
-                $Cost(S') \leftarrow \infty$
+
+                    $S \leftarrow GreedyRandomizedHeuristic()$
+
+                    \uIf{ $S \notin H$}{
+                        $S \leftarrow LocalSearch(S)$
+
+                        Add $S$ to $H$
+                    }
+
+                    $T \leftarrow T \cup \{S\}$
+
+                    \uIf{$Cost(S)< Cost(S^*)$}{
+                        $S^* \leftarrow S$
+                    }
+
+                    \uIf{$counter \text{ } mod \text{ } \phi == 0$}{
+                        $S^* \leftarrow PathRelinking(T)$
 
                         $T \leftarrow \emptyset$
-
-                \ForEach { $h \in H$ }{
-
-                    $S \leftarrow h(S)$
-
-                    $S \leftarrow LocalSearch(S)$
-
-                    $T \leftarrow T \cup \{S\}$
-
-                    \uIf{$Cost(S)< Cost(S')$}{
-                        $S' \leftarrow S$
-
-                        \uIf{$Cost(S')< Cost(S^*)$}{
-                            $S^* \leftarrow S'$
-                        }
-                    }
-
-                }
-                $S^* \leftarrow LocalSearch(T)$
+                    }
+
+                    $counter \leftarrow counter + 1$
+
                 }
@@
 
-\subsection{Greedy randomized insertion algorithm}
-
-As mentioned in \cite{resende2019greedy}, the general outline of a greedy randomized algorithm used in the GRASP framework works as described in Algorithm~\ref{alg:greedyIns}.
-Let's consider the set $CL$ of candidate delivery nodes that have not yet been scheduled.
-We first initialize $CL$ with the first delivery node of a randomly selected order, for each customer.
-
-At each iteration, we create the set $CL'$ with elements of $CL$ from a given neighborhood.
-We then evaluate the increase in the cost function of incorporating each $d \in CL'$ into the incumbent solution $S$ and construct a restricted candidate list $RCL$ consisting of nodes whose incremental cost is less than a given threshold.
-From $RCL$, we randomly select the next delivery node to be incorporated into the incumbent solution and determine its loading and unloading schedule using the procedure described in Algorithm~\ref{alg_ScheduleTask}.
-Next, we update $S$ and the remaining order and customer quantities.
-Then we add the next candidate delivery node to $CL$.
-If an order $o$ is not yet fulfilled after visiting $d^j_{o}$, the candidate is the next delivery node $d^{j+1}_{o}$.
-Otherwise, if the customer has remaining orders, the candidate is the first delivery node $d^{0}_{o'}$ of another randomly selected order $o'$.
-Note that for our problem, not all unscheduled delivery nodes are candidates at each iteration, since deliveries of the same order are ordered, any order of a customer can be chosen and must be satisfied before switching to another order, and the number of visits to a customer is not known in advance.
-
-Consider two customers whose services are required to start at the same time or in the same time slot.
-We could either finish scheduling one before the other, or schedule both at the same time.
-Obviously, each of these choices would result in a different solution.
-Similarly, in some cases, it would be better to schedule a customer that has a high demand and therefore requires more resources first.
-That's why we added the filtering part to impose some criteria on the algorithm.
-We can limit $CL'$ to delivery nodes with the same demand, the same due date, or from the same customer.
-We can also decide if we want to schedule customers with the highest (or lowest) demand or the earliest (latest) due time first.
-In fact, this helps us to intensify the search in specific neighborhoods of the solution.
-To diversify the search, we simply remove the filter component.
-The greedy aspect of GRASP is the creation of the set $RCL$, the probabilistic aspect is the random selection in $RCL$, and the adaptive aspect is the updating of $CL$ and the re-evaluation of the incremental costs.
-
-%Also, using $CL'$ helps reduce the complexity of the algorithm by reducing the number of nodes to evaluate.  %We thus obtain four greedy randomized insertion heuristics for our GRASP framework.
+For the construction phase, we have developed two greedy randomized heuristics.
+The first heuristic, called customer-based insertion, schedules customers one at a time.
+On the other hand, the second heuristic, called delivery-based insertion, does not necessarily complete the scheduling of one customer before starting to plan another customer's schedule.
+% The greedy aspect of GRASP is the creation of the set $RCL$, the probabilistic aspect is the random selection in $RCL$, and the adaptive aspect is the updating of $CL$ and the re-evaluation of the incremental costs.
+
+\subsection{Customer-based insertion heuristic}
+
+The customer-based insertion heuristic builds a solution from scratch by scheduling all of a customer's orders one at a time.
+As shown in Algorithm~\ref{alg:greedyCustIns}, for each customer, we randomly select the first order to schedule.
+To schedule an order, we iterate over each of its delivery nodes, find a feasible pair of task, and add it to the current solution.
+For a delivery node $d$, we check if the list of operations $M[d]$ is empty.
+If it is, for each available driver $k$ we simulate the loading and unloading operations $(L^k_d,U^k_d)$ with the procedure $SimulateVisit$, and if $k$ can serve $d$, we add $k$ to a candidate list $CL$ and $(L^k_d,U^k_d)$ to $M[d]$.
+
+The procedure $SimulateVisit$ takes as parameters a partial solution, a delivery node $d$, a driver $k$, and a numeric value $\lambda$.
+Its goal is to find the earliest time at which the delivery must start at $d$ and the best position within the route of $k$.
+When planning the visit of customer $i$ to node $d$ of order $o$ with driver $k$, we first compute his expected delivery time $v_d$ as follows:
+
+$ v_d = \left\{
+    \begin{array}{rl}
+        a_i + \lambda                                                   & \text{for the first visited node} \\
+        w_j + rand(\lambda, \gamma^1), j \in \mathcal{D}_o,  u_{jd}=1   &                                   \\
+        w_j + rand(\lambda, \gamma^2), j \notin \mathcal{D}_o, u_{jd}=1 &
+    \end{array}
+    \right.$
+
+Then we determine the start loading time $v_l$ of the corresponding dock node $l$ by subtracting the travel time, adjustment time, and loading duration from $v_d$.
+Starting from $v_l$ we check when $k$ will be available at an idle loading bay and adjust $w_l$ accordingly.
+Changing $w_l$ can cause the cold-joint constraints to be violated, especially for a customer's subsequent deliveries, and thus make the schedule infeasible.
+Once we have $v_l$, we can check for the best position to insert the visit of $d$ into the route of $k$.
+
+We extract from $CL$ the restricted candidate list ($RCL$) of drivers for whom the cost of visiting $d$ falls within a specified range based on the parameter $\theta$.
+If $RCL$ is empty and a visit to $d$ cannot be scheduled due to cold-joint constraints, we use the $DelayPreviousDelivery$ procedure to determine if there is a previous delivery of the same order that we can delay the start of service to make a visit to $d$ feasible.
+This procedure returns the index of the node and $\lambda$, which is the value of the delay.
+$\lambda$ is also the same parameter given to $SimulateVisit$.
+If there is such a previous node, we backtrack to it and restart the scheduling.
+
+If $RCL$ is not empty, we select a random driver from it, insert its corresponding load and unload tasks into $S$, and update the remaining quantities for the order and customer demands.
+The algorithm returns at the end the constructed solution $S$.
 
@@
         %  {\small
-     \begin{algorithm}[hbt]         
-         \caption{Greedy randomized insertion algorithm }
-         \label{alg:greedyIns}
+        \begin{algorithm}[htbp]
+            \caption{Customer-Based insertion algorithm }
+            \label{alg:greedyCustIns}
+            \DontPrintSemicolon
+            \LinesNumbered
+            \setcounter{AlgoLine}{0}
+            \KwIn{  $S$: empty solution $S$, $K$: vehicles set \hspace{15mm} \\$\Pi$: custom sorted customers list }
+
+            $M \leftarrow \emptyset $ \tcp{List of pair of loading and unloading tasks for all delivery nodes}
+            $CL \leftarrow \emptyset$ \tcp{Candidate driver list all delivery nodes}
+
+            \ForEach(customer $i \in \Pi$){}{
+
+            Shuffle $ \mathcal{O}_i$
+
+            \ForEach{order $o_1 \in \mathcal{O}_i$}{
+
+            $delay \leftarrow 0$
+
+            \For{ $j=0$ \KwTo $n^{max}_{o_1}-1$    }{
+
+            Delivery $d =\mathcal{D}^j_{o_1}$
+
+            \uIf{$M[d] == \emptyset $}{
+
+            \ForEach{ $k \in K$  }{
+
+            $(L^k_{d}$, $U^k_{d})$ $\leftarrow $ SimulateVisit($S$,$d$,$k$,$\lambda$)
+
+            $Cost(d,k)$ $\leftarrow $ Cost of ($L^k_{d}$, $U^k_{d}$)
+
+            Add $(L^k_{d}, U^k_{d}$) to $M[d]$
+
+            $CL[d] \leftarrow \{k\}$
+            }
+            }
+
+            $C_{min} \leftarrow min\left\lbrace C(x), \text{ } x \in M[d] \right\rbrace $, $C_{max} \leftarrow max\left\lbrace C(x), \text{ } x \in M[d] \right\rbrace $
+
+            $RCL \leftarrow \left\lbrace k \in CL[d], \text{ } C(d,k) \leq C_{min} + \theta (C_{max}-C_{min}) \right\rbrace $
+
+            \uIf{$RCL == \emptyset $}{
+
+                $(ind,delay)$ = $DelayPreviousDelivery$($S$,$d$)
+
+                \uIf{j is none}{
+
+                    break \tcp{continue with next customer}
+                }
+                $j \leftarrow ind$, $\lambda \leftarrow delay$
+
+                continue \tcp{backtrack to $ind$ position }
+
+            }
+            Select from $RCL$ a random driver $k$ to visit $d$ with load $q^k_d$
+
+            $M[d]\leftarrow M[d] \backslash (L^k_{d}, U^k_{d})$, $CL[d]\leftarrow CL[d] \backslash \{k\} $
+
+            $S \leftarrow S \cup (L^k_{d}$, $U^k_{d})  $
+
+            $q_i \leftarrow q_i-q^k_{d}$;  $q_{o_1} \leftarrow q_{o_1}-q^k_{d}$
+
+            $\lambda \leftarrow 0$
+
+            \uIf {$q_{o_1} \neq 0$ }{
+                continue
+
+            }
+            \uElseIf{ $q_i \neq 0$}{
+                break
+
+            }
+            }
+            }
+            }
+
+            \KwRet{$S$}
+        \end{algorithm}
+        %  }
+    }
+
+Consider the scenario where two customers require their services to start at the same time or within the same time window.
+In this case, we can either finish scheduling one customer before the other, or schedule both simultaneously.
+If there are sufficient resources available, either option would be feasible.
+However, when resources are limited, the customer-based insertion heuristic may leave some customers unscheduled.
+To address this limitation, we introduce the delivery-based insertion heuristic.
+
+\subsection{Delivery-based insertion heuristic}
+
+The delivery-based insertion heuristic constructs a feasible solution from scratch by scheduling a delivery node one at a time.
+The pseudocode for this heuristic is outlined in Algorithm~\ref{alg:delbasedIns}.
+It starts with an empty solution $S$ and initially fills $CL$ with the first delivery of a randomly selected order for each customer.
+Within each iteration of the $while$ loop, we filter $CL$ to obtain a reduced candidate list $CL'$ of deliveries from a given neighborhood.
+With this filtering, we can force the algorithm to schedule nodes with certain criteria first, which could be nodes from customers with the same demand, due date, or geographic area.
+We can also decide whether to schedule customers with the highest (or lowest) demand or the earliest (latest) due date first.
+
+For each delivery $d$ in $CL'$, the algorithm determines the best available driver $k$ and adds the corresponding loading and unloading task $(L^k_d, U^k_d)$ to a list of tasks.
+This is done using the $SimulateVisit$ procedure described earlier.
+Next, the algorithm constructs a $RCL$ from the list of tasks.
+If it is empty, the algorithm proceeds to the next iteration.
+Otherwise, it randomly selects a pair of tasks from the RCL, inserts them into $S$, and updates the remaining order and customer quantities.
+If an order $o$ is not fulfilled after visiting $d^j_{o}$, we add its next delivery node $d^{j+1}_{o}$ to $CL$.
+Otherwise, if the customer has remaining orders, we select the first delivery node $d^{0}_{o'}$ of another randomly selected order $o'$ and add it to $CL$.
+
+    % The greedy aspect of GRASP is the creation of the set $RCL$, the probabilistic aspect is the random selection in $RCL$, and the adaptive aspect is the updating of $CL$ and the re-evaluation of the incremental costs.
+
+    {\setstretch{1}
+        %  {\small
+        \begin{algorithm}[H]
+            \caption{Delivery-based insertion algorithm }
+            \label{alg:delbasedIns}
             \DontPrintSemicolon
@@
             \ForEach(customer i){}{
+
             Select a random order $o_1 \in \mathcal{O}_i$
@@
 
+            $Tasks \leftarrow \emptyset $ \tcp{List of loading and unloading tasks}
+
             \ForEach( $d \in CL'$){}{
 
-            $C(d) \leftarrow$ incremental cost of inserting $d$
-
-            }
-            $C_{min} \leftarrow min\left\lbrace C(d), \text{ } d \in CL' \right\rbrace $, $C_{max} \leftarrow max\left\lbrace C(d), \text{ } d \in CL' \right\rbrace $
-
-            $RCL \leftarrow \left\lbrace d \in CL', \text{ } C(d) \leq C_{min} + \alpha (C_{max}-C_{min}) \right\rbrace $
-
-            Select random delivery node $d^j_o$ of customer $i$ from $RCL$
-
-            $L_{d^j_o}$, $U_{d^j_o}$ = ScheduleTasks($d^j_o,S$)
-
-            $S \leftarrow S \cup (L_{d^j_o}$, $U_{d^j_o})  $
-
-            $q_i \leftarrow q_i-q_{d^j_o}$;  $q_{o} \leftarrow q_{o}-q_{d^j_o}$
+                % $C(d) \leftarrow$ incremental cost of inserting $d$
+
+                $Tasks \leftarrow (L^k_d,U^k_d)$ \tcp{$k$ is the best available driver to visit $d$}
+
+            }
+            $C_{min} \leftarrow \min\left\lbrace Cost(x), \text{ } x \in Tasks \right\rbrace $, $C_{max} \leftarrow \max\left\lbrace Cost(x), \text{ } x \in Tasks \right\rbrace $
+
+            $RCL \leftarrow \left\lbrace x \in Tasks, \text{ } Cost(x) \leq C_{min} + \theta (C_{max}-C_{min}) \right\rbrace $
+
+            \uIf{$RCL \leftarrow \emptyset$}{
+                continue
+            }
+
+            Select random operations $(L^k_{d^{j}_o},U^k_{d^{j}_o})$ of customer $i$ (order $o$) from $RCL$
+
+            $ S \leftarrow S \cup (L^k_{{d^{j}_o}}, U^k_{{d^{j}_o}})$
+
+            $q_i \leftarrow q_i-q^k_{d^{j}_o}$;  $q_{o} \leftarrow q_{o}-q^k_{d^{j}_o}$
 
@@
 
- Algorithm~\ref{alg_ScheduleTask} takes as parameters a delivery node $d^j_{o}$ of the customer $i$ and a partial solution and looks for a plant and a driver available for loading and unloading operations.
-For each driver $k$ and plant $b$ we simulate the loading and unloading operations while checking the concrete lifespan and cold joint constraints, and the plant and driver availability.
-For each node, we first determine the start loading time $SLT$ by subtracting $t_{bi}$, $\alpha_b$, and the loading duration $LD^b_{d^j_{o}}$ to the expected delivery time $EDT$.
-$EDT$ is either the arrival time $a_i$ for the first delivery of a customer or the end of the unloading of the precedent delivery.
-At line $15$, we ensure with the procedure \textit{ FindDriverLoadingSlot} that the loading starts only if the driver is present at the depot.
-The presence of the driver does not guarantee that the loading dock is available, thus we use the procedure \textit{FindDepotLoadingSlot} at line $16$ to ensure that.
-Once we know $SLT$, we can easily deduce the arrival time ($AT$), start unloading time ($SUT$), driver waiting time ($W^k_{d^j_o}$), and client waiting time ($W_{d^j_o}$).
-We also keep track of the driver's work duration.
-The algorithm returns the best loading and unloading operations with the least cost.
- We keep track of all timeslots for a plant when its loading dock is busy, and for a driver when he starts loading until the end of the unloading service.
-Within \textit{FindDepotLoadingSlot} (\textit{FindDriverLoadingSlot}), we iterate through the timeslots of a depot (driver) to schedule the current operation at a free timeslot.
-These allow us to insert the scheduling of a node between existing schedules.
-
-{\setstretch{1}
-    {\small
-        % \vspace{-4mm}
-        \begin{algorithm}[H]
-            \caption{Schedule loading and unloading tasks }
-            \label{alg_ScheduleTask}
-            % \DontPrintSemicolon
-            \LinesNumbered
-            \setcounter{AlgoLine}{0}
-            \KwIn{Partial solution $S$, delivery node $d^j_{o}$}
-
-            \SetKwFunction{FMain}{ScheduleTask}
-            \SetKwProg{Fn}{Function}{}{}
-            \Fn{\FMain{$d^j_{o},S$}}
-            {
-            % \SetNlSkip{2em}
-            % \SetAlgoNlRelativeSize{-1}
-            $n_i$: last node visited for the customer $i$
-
-            $\mathcal{LS}:$ Partial solutions list
-
-            $bestSol$ Cost($bestSol)= +\infty$
-
-            \ForEach(driver $k$){}
-            {
-                $n_k$: current location of $k$ (plant or delivery node)
-
-                \ForEach(plant $b$){}
-                {
-
-                    \uIf(continue){$t_{bi} >= \delta $}{      }
-
-                    $EDT=SUT=a_i$
-
-                \uIf {$n_i \neq null$ }{
-
-                    $SUT= rand( EUt_{n_i}-\gamma/3,
-                        EUt_{n_i})$
-
-                    $EDT= EUT_{n_i}$
-
-                }
-
-            $q^j_{o}=min\{Q_k,q_o \}$ // $q_o$ \textit{remaining demand of order} $o$
-
-            $SLT$ = $EDT - t_{bi} - \alpha_b - LD^b_{d^j_{o}} $
-
-            $SLT = FindDriverLoadingSlot(b, n_k, SLT, LD^b_{d^j_{o}})$
-
-            $SLT = FindDepotLoadingSlot(b, SLT, LD^b_{d^j_{o}})$
-
-            $AT =  SLT + LD^b_{d^j{o}} + \alpha_b + t_{bi}$
-
-            $SUT = max\{AT, EDT\}$
-
-            $W^k_{d^j_o}= max\{0,SUT - AT\}$
-
-            $W_{d^j_o}= max\{0,AT - EDT - \lambda_i\}$
-
-            $EUT = SUT + UD_{d^p_{ij}} $
-
-            $WT = WT_k + (EUT + \beta_k - SLT) + t_{n_kb} + t_{i,k}$
-
-            $TC = t_{n_k,b} + t_{b,i}$  %- t_{pos_k,k}$
-
-            \uIf{$Cost(bestSol) < Cost(S) + TC $}{
-                $bestSol = (j,b,k,SLT) \cup (j,b,k,SUT) $
-
-                $Cost(bestSol) = Cost(S) + TC$
-            }
-            }
-            }
-            \KwRet{$bestSol$}
-            }
-        \end{algorithm}
-    }
-}
-
 \subsection{ Local Search}
 
-    % We implement two local search operators for the CDP. The first one is the relocate move commonly used for vehicle routing problems.
-
-   % At the end of each iteration, we expect each customer to have a feasible schedule respecting all constraints.
-For each customer, we combine each scheduling in order to construct a new and possibly better solution.
+We use four local search operators to improve the solution obtained after the construction phase.
+
+\subsubsection*{Swap load operator}
+
+The first operator exchanges loads between two deliveries of the same order. This move affects the loading operations and can potentially improve the number of scheduled orders.
+
+\subsubsection*{Swap driver operator}
+
+This operator investigates different driver assignments by swapping the drivers assigned to two delivery nodes in the current solution.
+
+\subsubsection*{Relocate load operator}
+
+This operator moves a delivery load that was originally split between two drivers to a single driver, if that driver's capacity allows.
+%   By consolidating the load into a single vehicle, we can potentially optimize the routing and reduce the total travel distance.
+This operator helps in better utilizing the available capacity of the vehicles and improving the overall efficiency of the solution.
+
+\subsubsection*{Remove and reschedule operator}
+
+This operator takes a customer $i$ not scheduled in the current solution.
+It then identifies a set $C$ of scheduled customer's with the same time slot as $i$.
+Elements of $C$ are removed and attempted to be rescheduled along with $i$.
+
+% The application of these local search operators is performed iteratively, allowing for multiple rounds of neighborhood exploration and solution refinement.
+Each operator is applied sequentially, and the resulting solution is evaluated to determine if an improvement has been achieved.
+The process continues until no further improvements can be made, indicating that a locally optimal solution has been reached.
+
+\subsection{Path relinking}
+
+After the construction and local search steps, it may happen that some customer orders cannot be scheduled within a solution.
+The goal of the path relinking procedure is to use each customer's scheduling information from existing solutions and combine them effectively to construct a solution that schedules all customers.
+This algorithm iterates through each element of a pool of solutions and attempts to insert the customer's deliveries into a newly constructed solution.
 
@@
 
-% In this section, we present and discuss the results obtained in the computational experiments of the GRASP heuristic described in Section~\ref{grasp_method}. The GRASP algorithm was coded in C++.
+In this section, we present and discuss the results obtained in the computational experiments of the GRASP heuristic described in Section~\ref{grasp_method}.
+The GRASP algorithm was coded in C++.
+We run our experiments on dataset extracted from data provid by our industrial partner.
 % and the branch-and-bound (B\&B) algorithm applied to the model described in Section~\ref{model}.
@@
 \subsection{Generation of instances}
-% Due to the lack of publicly available benchmark data for the CDP, two data sets are created.
-Both sets, as well as code to generate new instances, are available online at [17].
-A summary of the instances is provided in Table 4.DataSetAcontains instances with 10–20 customers,and2–5 vehicles,whereasData Set B contains instances with up to 50 customers and 20 vehicles.
-The demands of customers are always divisible by 5 and are selected at random from the interval [10–75].
-In a similar fashion, the capacities of the vehicles are selected between [10–25].
-It is unrealistic to have a unique capacity for each vehicle, and therefore the number of different vehicles is bounded by the number of vehicle classes.Theprocessing time of a vehicle kAK is set proportional to its capacity, i.e.
-pk ¼qk for all kAK.
-The width of the time window of a customer i AC is computed by λiqi,whereλi is a scalar uniformly selected from the interval ½1:1; 2:1.
-Each instance has up to 4 production stations.
-The locations of the start and end depot, construction sites, and production depots, lay in the euclidean plane.
-The travel time between two locations equals the euclidean distance, rounded upwards.
-For simplicity, the start and end depot are the same.
-The distance between each station and the depot is uniformly selected from the interval [1–30].
-Furthermore, for each customer, there exists a production station within a distance of [1–25].Finally,thetimelagγ is fixed to 5 for all deliveries.
-\subsection{Data sets}
+
+In order to evaluate the effectiveness and performance of our GRASP algorithm, we obtained a dataset from our partner that represents the delivery operations of 8 production centers for a period of 42 days.
+
+We generate 42 instances, each corresponding to a day operation. An instances has informations about the number of production centers, customers, orders, and drivers.
+For each depot, we have its ID and loading capacity.
+For a customer, we have the ID of the assigned depot, his demand, due time and number of orders.
+Then for each order we have his demand.
+For the driver, we have the capacity and shift start time.
+
+\begin{itemize}
+    \item   8 production centers located in greater Montreal
+    \item   42 days of delivery operations from Fall 2020
+    \item   From 7 to 139 orders each day
+    \item   Up to 137 drivers
+    \item   2 types of trucks of capacity 8 and 12 m$^3$
+\end{itemize}
 
@@
 
-\subsubsection{ Results for the real-world datasets}
 \subsubsection{ Results for the instances of \cite{kinable2014concrete}}
 
-We run our algorithm on the dataset provided in \cite{kinable2014concrete}.
+\cite{kinable2014concrete} provided a dataset for their variant of the $CDP$.
+as remainder, their objective is to maximize the total load served each day, and there is no time allocated for loading operations at a plant.
+While the original objective of the dataset may not align directly with our specific objective, we can still utilize it to analyze and compare the performance of our algorithm under different circumstances.
+
+The results are below.
+
+\subsubsection{Comparisons with the real operations}
 

### 2023-08-02 Data/instances/createInstance/create_instance.py

@@ -1,0 +1,209 @@
+
+import pandas as pd
+import os
+import math
+from driver import Driver
+import datetime
+from collections import Counter
+
+INSTANCE_FOLDER = "/".join([os.path.dirname(os.getcwd()), "createInstance/instances/"])
+INSTANCE_FOLDER_CQ = "/".join([os.path.dirname(os.getcwd()), "createInstance/cq/"])
+DATA_FOLDER = "/".join([os.path.dirname(os.getcwd()), "createInstance/data/"])
+
+Instances_with_extern_fleet   = set()  
+Instances_with_intern_fleet = set ()
+
+
+def create_instance_CQ(Object,jour)->None:
+    day = jour.day
+    if day < 10:
+        day = f"0{day}"
+    print(f"Instance {jour}")
+    list_ordres = list(set( Object.df_orders[Object.df_orders['DATE']==jour][Object.df_orders['JOB_DESC']!='PICKUP']['ORDER_ID']))
+    tickets = Object.df_tickets[Object.df_tickets['ORDER_ID'].isin(list_ordres)]
+    ordres = Object.df_orders[Object.df_orders['ORDER_ID'].isin(list_ordres)]
+    sortedMap = dict()
+    totalDemand = 0
+    for ord in list_ordres:
+        ordered_prod =  list(set(ordres[ordres['ORDER_ID']==ord]['PROD_NBR']))
+        delivered_prod =  list(set(tickets[tickets['ORDER_ID']==ord]['PROD_NBR']) & set(ordered_prod))
+        if len(delivered_prod) == 0:
+            continue
+        cur =  tickets[tickets['PROD_NBR'].isin(delivered_prod)][tickets['ORDER_ID']==ord]
+        # print(ord,' ',ordered_prod, ' ',delivered_prod,' ',cur['QUANTITY'].sum())
+        totalDemand = totalDemand + cur['QUANTITY'].sum()
+        sortedMap[ord] = (cur['QUANTITY'].sum(), ordres[ordres['ORDER_ID']==ord]['SHIPTIME (min)'].min(),len(delivered_prod))
+
+    sortedMap = sorted(sortedMap.items(), key=lambda x:(x[1][0],x[1][1],x[1][2]))
+    # print(sortedMap)
+    print(totalDemand)
+    drivers = list(dict.fromkeys(tickets['DRIVER_NBR']))
+    dict_driver = {}
+    for driver in drivers:
[…]

### 2023-08-05 codeCQ/heur/Solver.cpp
Save CQ

@@ -112,28 +112,28 @@ void Solver::SolveGrasp(Sol &s, Data &dat, vector<set<int>> const &linkedClientS
     AllInsertionOperator d2(&driverIns3, driverInfo[2].first, "Builder 3 " + driverInfo[2].second);
 
 
-    grasp.AddInsertOperatorVrp(&c0_1);
+    // grasp.AddInsertOperatorVrp(&c0_1);
     grasp.AddInsertOperatorVrp(&c0_3);
-    grasp.AddInsertOperatorVrp(&c1_1);
-    grasp.AddInsertOperatorVrp(&c1_3);
-    grasp.AddInsertOperatorVrp(&c2_1);
-    grasp.AddInsertOperatorVrp(&c2_3);
-    grasp.AddInsertOperatorVrp(&c3_1);
-    grasp.AddInsertOperatorVrp(&c3_3);
-    grasp.AddInsertOperatorVrp(&c4_1);
-    grasp.AddInsertOperatorVrp(&c4_3);
-    grasp.AddInsertOperatorVrp(&c5_1);
-    grasp.AddInsertOperatorVrp(&c5_3);
+    // grasp.AddInsertOperatorVrp(&c1_1);
+    // grasp.AddInsertOperatorVrp(&c1_3);
+    // grasp.AddInsertOperatorVrp(&c2_1);
+    // grasp.AddInsertOperatorVrp(&c2_3);
+    // grasp.AddInsertOperatorVrp(&c3_1);
+    // grasp.AddInsertOperatorVrp(&c3_3);
+    // grasp.AddInsertOperatorVrp(&c4_1);
+    // grasp.AddInsertOperatorVrp(&c4_3);
+    // grasp.AddInsertOperatorVrp(&c5_1);
+    // grasp.AddInsertOperatorVrp(&c5_3);
 //    grasp.AddInsertOperatorVrp(&cFixed);
 
-    grasp.AddInsertOperatorVrp(&p01);
-    grasp.AddInsertOperatorVrp(&p03);
-    grasp.AddInsertOperatorVrp(&p11);
-    grasp.AddInsertOperatorVrp(&p13);
-    grasp.AddInsertOperatorVrp(&p23);
-    grasp.AddInsertOperatorVrp(&p33);
-    grasp.AddInsertOperatorVrp(&p43);
-    grasp.AddInsertOperatorVrp(&p53);
+    // grasp.AddInsertOperatorVrp(&p01);
[…]

### 2023-08-05 Article/article.tex
save CQ

@@
 \usepackage[T1]{fontenc}
+\usepackage{graphicx}
 \usepackage{natbib}
@@
 In \cite{kinable2014logic}, we found an attempt to solve the previous problem with a logic-based Benders' approach.
-\cite{sulaman2017simulated} expand upon the SD heuristic proposed in \cite{kinable2014concrete}, proposing a simulated annealing (SA) combined with a time-slot Heuristic (SATH).
+\cite{sulaman2017simulated} expand upon the SD-heuristic proposed in \cite{kinable2014concrete}, proposing a simulated annealing (SA) combined with a time-slot Heuristic (SATH).
 This method looks for a slot between existing visits of a truck to schedule a new delivery instead of assigning it to the time slot strictly after the truck's latest assigned delivery.
@@
 
-In this section, we present and discuss the results obtained in the computational experiments of the GRASP heuristic described in Section~\ref{grasp_method}.
-The GRASP algorithm was coded in C++.
-We run our experiments on dataset extracted from data provid by our industrial partner.
-% and the branch-and-bound (B\&B) algorithm applied to the model described in Section~\ref{model}.
-The different algorithms were coded in C++ and run on a processor Intel $2.4$ GHz Gold $6148$ Skylake.
-For the exact method, we have used IBM Concert Technology and CPLEX $12.10$ as MIP solver.
-A single thread was used with a time limit of one hour.
-% No benchmark instances for the \dirp exist so we have generated the dataset for the experiments as follows.
+In this section, we present and discuss the results of the computational experiments of the GRASP heuristic described in section~\ref{grasp_method}.
+The GRASP algorithm is coded in C++.
+We run our experiments on a dataset extracted from data provided by our industrial partner.
 
@@
 
-In order to evaluate the effectiveness and performance of our GRASP algorithm, we obtained a dataset from our partner that represents the delivery operations of 8 production centers for a period of 42 days.
-
-We generate 42 instances, each corresponding to a day operation. An instances has informations about the number of production centers, customers, orders, and drivers.
-For each depot, we have its ID and loading capacity.
-For a customer, we have the ID of the assigned depot, his demand, due time and number of orders.
-Then for each order we have his demand.
-For the driver, we have the capacity and shift start time.
+The dataset used in this section was obtained from our partner and contains records representing delivery operations from up to 8 production centers over a period of 50 days.
+We generated 50 instances, with each instance corresponding to a daily operation.
+The instances include informations such as each driver capacity, associated depot, and shift start time.
+For each customer, we have the due time, total demand, and the number of orders (type of concrete) received.
+For each order, we specified its demand and associated production plant.
+Finally, we include loading capacities for each production center.
+The name of an instance has the format $C\_k\_n\_o\_z$. $k$ is the number of available drivers, $n$ the number of customers, $o$ the total number of daily orders, and $z$ the number of plants.
+
+The dataset is subdivised in a small, medium and large sets, according to the number of drivers and total demands served. We summarized both sets in Table~\ref*{}
+or 
+The dataset is subdivised in a small, medium and large sets, according to the number of orders served. We summarized both sets in Table~\ref*{}
+
+% This dataset allows us to evaluate the effectiveness and performance of our GRASP algorithm in optimizing delivery operations in a real-world setting.
 
@@
 
+- resultats avec les instances (8 au total) où uniquement la flotte de Unibéton a éét utilisée. (utiliser les mêmes pour la comparaison avec les données réelles )
+distance, demande servie, first, underwork, overtime.
+- resultats avec les autres instances
+- distance, demande servie, first delivery.
+
+% df = df[df['Name'].isin([20201010,20201021,20201107,20201017,20201031,20201109,20201026,20201121])]
+
 \subsubsection{ Results for the instances of \cite{kinable2014concrete}}
 
-\cite{kinable2014concrete} provided a dataset for their variant of the $CDP$.
-as remainder, their objective is to maximize the total load served each day, and there is no time allocated for loading operations at a plant.
-While the original objective of the dataset may not align directly with our specific objective, we can still utilize it to analyze and compare the performance of our algorithm under different circumstances.
-
-The results are below.
+\cite{kinable2014concrete} provided a benchmark dataset for their variant of the $CDP$.
+Although their primary objective is to maximize the total load delivered each day, without considering loading operations at a plant and ensuring deliveries within specified time windows, we can still use this dataset to analyze and compare the performance of our algorithm under different conditions.
+Table \ref*{tab:cdp_instances} shows the performance of the GRASP heuristic along with the other methods (CP, MIP, SD-heuristic, and Hybrid) used in \cite{kinable2014concrete}.
+For each method, the table shows the average execution time in milliseconds and the corresponding gap between the obtained solution and the upper bound provided by \cite{kinable2014concrete}.
+Overall, the GRASP algorithm shows promise in providing near-optimal solutions in a relatively short time for this benchmark of the CDP.
+It outperformed all methods except CP, achieving an average distance of 5.7\% (13.8\%) with a runtime of 13,186 (574,983) milliseconds for Set A (Set B).
+The SD-heuristic appears to be much faster, but GRASP provides a better balance between solution quality and computation time.
+It should also be noted that the exact MIP, CP, and Hybrid are initialized by the results of the SD-heuristic.
+
+\begin{table}[htbp]
+    \centering
+    \caption{Performance of the GRASP heuristic on the CDP instances}
+    \label{tab:cdp_instances}
+    \resizebox{1\textwidth}{!}{
+    \begin{tabular}{@{}cccccccccccc@{}}
+    \toprule
+     &  & \multicolumn{2}{c}{CP} & \multicolumn{2}{c}{MIP} & \multicolumn{2}{c}{SD-heuristic} & \multicolumn{2}{c}{Hybrid} & \multicolumn{2}{c}{GRASP} \\ \cmidrule(l){3-10} \cmidrule(l){11-12} 
+     &  & Gap (\%) & Time (ms) & Gap (\%) & Time (ms) & Gap (\%) & Time (ms) & Gap (\%) & Time (ms) & Gap (\%) & Time (ms) \\ \cmidrule(l){3-10} \cmidrule(l){11-12} 
+    \multirow{2}{*}{Set A} & Avg & 4.2 & 197,012 & 7.3 & 13,405,798 & 9.1 & 23 & 7.0 & 100,765 & 5.7 & 13,186 \\ \cmidrule(l){2-10} \cmidrule(l){11-12} 
+     & Opt & \multicolumn{2}{c}{40/64} & \multicolumn{2}{c}{37/64} & \multicolumn{2}{c}{30/64} & \multicolumn{2}{c}{35/64} & \multicolumn{2}{c}{35/64} \\ \cmidrule(l){3-10} \cmidrule(l){11-12} 
+     \multirow{2}{*}{Set B} & Avg & 12.1 & 357,313 & - & - & 16.3 & 1,331 & - & - & 13.8 & 574,983\\ \cmidrule(l){2-10} \cmidrule(l){11-12} 
+     & Opt & \multicolumn{2}{c}{55/128} & \multicolumn{2}{c}{-} & \multicolumn{2}{c}{40/128} & \multicolumn{2}{c}{-} & \multicolumn{2}{c}{44/128} \\ \bottomrule
+    \end{tabular}
+    }
+    \end{table}
 

### 2023-08-07 codeCQ/Parameters.cpp
Save CQ

@@ -21,7 +21,7 @@ int Parameters::SORT_TYPE = Parameters::SORT::TWO;
 std::chrono::time_point<std::chrono::steady_clock> Parameters::START_TIME = std::chrono::steady_clock::now();
 
 
-int Parameters::INTRA_ORDER_DELIVERY = 10;
+int Parameters::INTRA_ORDER_DELIVERY = 20;
 int Parameters::INTER_ORDER_DELIVERY = 25;
 int Parameters::MIN_WORKING_TIME = 180;
 int Parameters::NORMAL_WORKING_TIME = 480;
@@ -36,7 +36,7 @@ int Parameters::FIRST_DEL_PENALTY = 1;//;1000;
 int Parameters::UNDERWORK_PENALTY = 1;//;100;
 int Parameters::OVERTIME_PENALTY = 1;//200;
 int Parameters::LATE_ARRIVAL_PENALTY = 10;
-int Parameters::FIRST_DEL_MAX_ARRIVAL = 20;
+int Parameters::FIRST_DEL_MAX_ARRIVAL = 60;
 
 int Parameters::LOAD_INSERTION = Parameters::DEPOTINSERTION::FORWARD;
 

### 2023-08-07 codeCQ/UpdateCost.cpp

@@ -199,6 +199,7 @@ void Sol::Update(Depot *dep, Dock *dock, Delivery *del) {
     } else {
         if (WaitingTime[del->id] < 0) {
             lateCustomers.insert(c->custID);
+            updateCost.maxLateness = std::max(-WaitingTime[del->id] ,updateCost.maxLateness );
         }
     }
 
@@ -359,13 +360,17 @@ void Sol::GetCost(Depot *dep, Dock *dock, Delivery *del, Cost &cur_cost) {
 
     double arr_node = ArrivalTime[del->id];
 
+    double lateness = cur_cost.firstDeliveryCost;
     Sol::SetTimingCost(del, arr_node, real_del_time, EarlyTW(del),
                        cur_cost);
-
+    lateness = cur_cost.firstDeliveryCost - lateness;
+    if(CustomerPrev[del->id]->type != Parameters::TypeNode::DELIVERY){
+        cur_cost.maxLateness = std::max(lateness,cur_cost.maxLateness);
+    }
     arr_node = StartServiceTime[del->id];
     arr_node += UNLOADING_DURATION;
 
-    cur_cost.lateDeliveryCost += Sol::GetLateDeliveryCost(del, arr_node, LateTW(del));
+    cur_cost.lateDeliveryCost += Sol::GetLateDeliveryCost(del, arr_node, LateTW(del));;
     arr_node += Data::CleaningTime(del, d);
 
     shiftDurationCost[d->id] = arr_node + Travel(del->distID, d->distID) - d->start_shift_time;

### 2023-08-07 codeCQ/Solution.cpp
Save CQ, solve bug with swapLoad

@@ -182,7 +182,9 @@ void Sol::UnassignCustomer(Customer *c) {
     std::vector<Order *> orders = GetOrders(c);
     for (auto o1: orders) {
         UnassignOrder(o1);
+        orderCapRestante[o1->orderID]=o1->demand;
     }
+    clientCapRestante[c->custID] = c->demand;
 }
 
 void Sol::UnassignOrder(Order *o) {

### 2023-08-07 codeCQ/Solution.h

@@ -317,7 +317,7 @@ public:
 
     bool isOrderSatisfied(int ordId) const { return (orderCapRestante[ordId] <= 0); }
 
-    bool isClientSatisfied(Customer *c) const { return (clientCapRestante[c->custID] <= 0); }
+    bool isClientSatisfied(Customer *c) const { return isClientSatisfied(c->custID); }
 
     bool isSatisfied(Customer *c) const { return (clientCapRestante[c->custID] <= 0); }
 
@@ -356,6 +356,7 @@ public:
     void UpdateDemand(Customer *c, Order *o, double load) {
         orderCapRestante[o->orderID] -= load;
         clientCapRestante[c->custID] -= load;
+        assert(clientCapRestante[c->custID] >= 0);
         orderLoads[o->orderID].insert(load);
     }
 
@@ -565,7 +566,7 @@ public:
     static bool FindForwardSlot(std::set<TimeSlot> const &SlotSet, TimeSlot &slot, const double duration);
     static bool FindBackwardSlot(std::set<TimeSlot> const &SlotSet, TimeSlot &slot, const double duration);
 
-
+    bool isUpdated{false};
 private:
     Data *_data{};
     Cost _last_cost;

### 2023-08-07 codeCQ/heur/CDPSolver.cpp

@@ -121,6 +121,7 @@ void CDPSolver::SolveInstance(Sol &s, Data &dat, int iter) {
 }
 
 void CDPSolver::BuildOnSolution(Sol &s, Data &dat, int iter) {
+    s.isUpdated = true;
     CDPSolver::SolveCDP(s, dat, iter, false);
 }
 

### 2023-08-07 codeCQ/heur/CustInsertion.cpp

@@ -9,7 +9,10 @@ using namespace std;
 void CustInsertion::Insert(Sol &s) {
     Sol::InitStructure(s.GetData());
     removedList.clear();
-    s.Update();
+    if(!s.isUpdated){
+        s.Update();
+    }
+    s.isUpdated = false;
     _insrmv.FillStructures(s, customersList, driversList);
     assert(!customersList.empty());
     const int insertion_type = Parameters::LOAD_INSERTION;

### 2023-08-07 codeCQ/heur/InsRmvBuilder1.h

@@ -49,7 +49,7 @@ public:
         mo.n = n;
         mo.to = d;
         mo.demand = (Sol::FixLoad[n->delID] == -1) ?
-                    std::min(d->capacity, s.orderCapRestante[n->orderID]) : Sol::FixLoad[n->delID];
+                    std::min(d->capacity, s.orderCapRestante[n->orderID]) : std::min(s.orderCapRestante[n->orderID],Sol::FixLoad[n->delID]);
 
         if (mo.demand > d->capacity) return;
 
@@ -102,7 +102,7 @@ public:
         mo.n = n;
         mo.to = d;
         mo.demand = (Sol::FixLoad[n->delID] == -1) ?
-                    std::min(d->capacity, s.orderCapRestante[n->orderID]) : Sol::FixLoad[n->delID];
+                    std::min(d->capacity, s.orderCapRestante[n->orderID]) : std::min(s.orderCapRestante[n->orderID],Sol::FixLoad[n->delID]);
 
         if (mo.demand > d->capacity) return;
 

### 2023-08-07 codeCQ/heur/InsRmvBuilder3.h

@@ -76,7 +76,7 @@ public:
         mo.n = n;
         mo.to = d;
         mo.demand = (Sol::FixLoad[n->delID] == -1) ?
-                    std::min(d->capacity, s.orderCapRestante[n->orderID]) : Sol::FixLoad[n->delID];
+                    std::min(d->capacity, s.orderCapRestante[n->orderID]) : std::min(s.orderCapRestante[n->orderID],Sol::FixLoad[n->delID]);
 
         if (mo.demand > d->capacity) return;
 
@@ -106,7 +106,7 @@ public:
         mo.n = n;
         mo.to = d;
         mo.demand = (Sol::FixLoad[n->delID] == -1) ?
-                    std::min(d->capacity, s.orderCapRestante[n->orderID]) : Sol::FixLoad[n->delID];
+                    std::min(d->capacity, s.orderCapRestante[n->orderID]) : std::min(s.orderCapRestante[n->orderID],Sol::FixLoad[n->delID]);
 
         if (mo.demand > d->capacity) return;
 

### 2023-08-07 codeCQ/heur/InsRmvMethodFast.cpp

@@ -37,7 +37,7 @@ void InsRmvMethodFast::InsertCost(Sol &s, Delivery *n, Driver *d,
     mo.n = n;
     mo.to = d;
     mo.demand = (Sol::FixLoad[n->delID] == -1) ?
-                std::min(d->capacity, s.orderCapRestante[n->orderID]) : Sol::FixLoad[n->delID];
+                std::min(d->capacity, s.orderCapRestante[n->orderID]) : std::min(s.orderCapRestante[n->orderID],Sol::FixLoad[n->delID]);
 
     if (mo.demand > d->capacity) return;
     if (mo.demand <= 0) {
@@ -71,13 +71,10 @@ void InsRmvMethodFast::InsertCost(Sol &s, Delivery *n, Driver *d,
     mo.n = n;
     mo.to = d;
     mo.demand = (Sol::FixLoad[n->delID] == -1) ?
-                std::min(d->capacity, s.orderCapRestante[n->orderID]) : Sol::FixLoad[n->delID];
+                std::min(d->capacity, s.orderCapRestante[n->orderID]) : std::min(s.orderCapRestante[n->orderID],Sol::FixLoad[n->delID]);
 
     if (mo.demand > d->capacity) return;
-    if (mo.demand <= 0) {
-        Prompt::print({std::to_string(Sol::FixLoad[n->delID]), std::to_string(mo.demand)
-                      });
-    }
+
     assert(mo.demand > 0);
 
     Dock *dock = s.GetDock(n->dockID);

### 2023-08-07 codeCQ/heur/PriorityQueueInsertion.cpp

@@ -13,7 +13,10 @@ using namespace std;
 void PriorityQueueInsertion::Insert(Sol &s) {
     _data = *s.GetData();
     Sol::InitStructure(s.GetData());
-    s.Update();
+    if(!s.isUpdated){
+        s.Update();
+    }
+    s.isUpdated = false;
     removedList.clear();
     _insrmv.FillStructures(s, customersList, driversList);
     InsertByRules(s);

### 2023-08-07 Article/article.tex
Save article

@@
 
-In this paper, we propose a mathematical model and a Greedy Randomized Search Procedure (GRASP) heuristic solution to solve a new variant of the CDP.
+In this paper, we propose a mathematical model and a Greedy Randomized Search Procedure (GRASP) heuristic to solve a new variant of the CDP.
 Our model takes into consideration the working shifts of the drivers and the scheduling of multiple orders within the same time window at a construction site.
@@
 In \cite{kinable2014logic}, we found an attempt to solve the previous problem with a logic-based Benders' approach.
-\cite{sulaman2017simulated} expand upon the SD-heuristic proposed in \cite{kinable2014concrete}, proposing a simulated annealing (SA) combined with a time-slot Heuristic (SATH).
+%\cite{sulaman2017simulated} expand upon the SD-heuristic proposed in \cite{kinable2014concrete}, proposing a simulated annealing (SA) combined with a time-slot Heuristic (SATH).
 This method looks for a slot between existing visits of a truck to schedule a new delivery instead of assigning it to the time slot strictly after the truck's latest assigned delivery.
@@
 Exactly one order $o \in O_i$ must have its first delivery start at $a_i$, while the others can start at most $\gamma^2$ time after $o$ is completed.
-%Once a plant is selected to produce a customer's first delivery, it must be the supplier of all subsequent deliveries from that customer.
+A plant is assigned to an order, and it must be the supplier of all subsequent deliveries of that order.
 To avoid cold joint problems with the concrete, subsequent deliveries of the same order must be made in close succession.
 We define a maximum time delay $\gamma^1$ after which no more deliveries are allowed.
-The customer unloading rate and the quantity to be unloaded give the time required to unload a truckload.
+The customer's unloading rate and the quantity to be unloaded give the time required to unload a truckload.
 Let $\mathcal{C}$ be the set of construction sites (customers) with a planned delivery for the day, and $\mathcal{O}=\{{O_i, i \in \mathcal{C}}\}$ be the set of all requested orders for all customers.
@@
 Constraints (\ref{mod:c11}) - - (\ref{mod:c15}) find the delivery sequence of all orders $\in \mathcal{O}_i$.
-Constraints (\ref{mod:c16}) - - (\ref{mod:c19}) prohibit two trucks serving consecutive deliveries of the same order from unloading at the same time and enforce a maximum time delay between them.
+With constraints (\ref{mod:c16}) - - (\ref{mod:c19}), two trucks cannot unload at the same time for consecutive deliveries of the same order.
+Additionally, these constraints impose a maximum time delay between the two trucks.
 Similarly, constraints (\ref{mod:c20}) - - (\ref{mod:c23}) ensure that two trucks cannot be loaded at the same time at a plant.
@@
 
-\begin{figure}[H]
+\begin{figure}[htbp]
     % \caption{Solution of an instance with two plants, two construction sites, four orders, and three drivers.}
@@
 
-\begin{figure}[H]
+\begin{figure}[htbp]
     \centering
@@
 
+- resultats avec les instances où toute la demande est satisfaite. Montrer la distance parcourue, le temps de première livraison, overtime, underwork, temps d'execution.
+
 - resultats avec les instances (8 au total) où uniquement la flotte de Unibéton a éét utilisée. (utiliser les mêmes pour la comparaison avec les données réelles )
@@
 % df = df[df['Name'].isin([20201010,20201021,20201107,20201017,20201031,20201109,20201026,20201121])]
+
+\begin{table}
+    \resizebox*{\textwidth}{!}{
+
+\begin{tabular}{lrrrrrrr}
+    \toprule
+    {} &  satisfiedCost &  driverUsed &  travelCost &  firstDeliveryCost &  underWorkCost &  overTimeCost &  runtime \\
+    Name     &                &             &             &                    &                &               &          \\
+    \midrule
+    20201121 &          226.5 &          13 &     1243.55 &               0.51 &           0.00 &          0.00 &     5.26 \\
+    20201010 &          267.0 &          13 &      745.10 &               0.00 &           0.00 &          0.00 &     5.32 \\
+    20201107 &          333.5 &          18 &      777.25 &               0.00 &         200.16 &          0.00 &     5.19 \\
+    20201031 &          375.0 &          15 &     1017.64 &               0.00 &         216.06 &         28.10 &     5.42 \\
+    20201114 &          388.0 &          19 &     1743.64 &               0.00 &         149.48 &          0.00 &     5.42 \\
+    20201128 &          776.0 &          31 &     2278.18 &               0.00 &           4.19 &        438.14 &     5.16 \\
+    20201024 &          937.5 &          35 &     4403.57 &               0.00 &          58.30 &       1005.24 &     5.72 \\
+    20201021 &         1160.5 &          73 &     6107.22 &               0.00 &        2535.40 &        185.02 &     5.28 \\
+    20201013 &         1565.0 &         102 &     9214.74 &              19.53 &        2955.35 &         95.74 &     5.32 \\
+    20201026 &         1839.5 &         116 &    13585.60 &              44.44 &        1826.70 &        227.56 &     5.41 \\
+    20201007 &         2512.5 &         127 &    15166.00 &               5.09 &         877.68 &        574.76 &     5.67 \\
+    \bottomrule
+    \end{tabular}
+    }
+\end{table}
 
@@
 The SD-heuristic appears to be much faster, but GRASP provides a better balance between solution quality and computation time.
-It should also be noted that the exact MIP, CP, and Hybrid are initialized by the results of the SD-heuristic.
+It should also be noted that the exact MIP, CP, and Hybrid are initialized with the results of the SD-heuristic.
 
@@
 
+- rouler les instances sans shift des livreurs. comparer la distance parcourue, et demande livrée.
+- rouler sans les livreurs externes. regarder la demande servie par jour.
+
 \section{Conclusion}

### 2023-08-08 Article/article.tex
save article

@@
 
-The procedure $SimulateVisit$ takes as parameters a partial solution, a delivery node $d$, a driver $k$, and a numeric value $\lambda$.
-Its goal is to find the earliest time at which the delivery must start at $d$ and the best position within the route of $k$.
-When planning the visit of customer $i$ to node $d$ of order $o$ with driver $k$, we first compute his expected delivery time $v_d$ as follows:
+The $SimulateVisit$ procedure takes a partial solution, a delivery node $d$, a driver $k$ and a numeric value $\lambda$ as parameters.
+Its goal is to find the earliest time when an unloading service can start at $d$ and the best position within the route of $k$.
+When planning to visit node $d$ of order $o$ and customer $i$ with driver $k$, we first compute his expected delivery time $v_d$ as follows:
 
@@
         %  {\small
-        \begin{algorithm}[htbp]
+        \begin{algorithm}[]
             \caption{Customer-Based insertion algorithm }
@@
 
-            $delay \leftarrow 0$
+            $\lambda \leftarrow 0$
 
@@
 If it is empty, the algorithm proceeds to the next iteration.
-Otherwise, it randomly selects a pair of tasks from the RCL, inserts them into $S$, and updates the remaining order and customer quantities.
+Otherwise, it randomly selects a pair of tasks from the $RCL$, inserts them into $S$, and updates the remaining order and customer quantities.
 If an order $o$ is not fulfilled after visiting $d^j_{o}$, we add its next delivery node $d^{j+1}_{o}$ to $CL$.
@@
 
-\subsection{ Local Search}
-
-We use four local search operators to improve the solution obtained after the construction phase.
-
-\subsubsection*{Swap load operator}
-
-The first operator exchanges loads between two deliveries of the same order. This move affects the loading operations and can potentially improve the number of scheduled orders.
-
-\subsubsection*{Swap driver operator}
-
-This operator investigates different driver assignments by swapping the drivers assigned to two delivery nodes in the current solution.
-
-\subsubsection*{Relocate load operator}
-
-This operator moves a delivery load that was originally split between two drivers to a single driver, if that driver's capacity allows.
-%   By consolidating the load into a single vehicle, we can potentially optimize the routing and reduce the total travel distance.
-This operator helps in better utilizing the available capacity of the vehicles and improving the overall efficiency of the solution.
-
-\subsubsection*{Remove and reschedule operator}
-
-This operator takes a customer $i$ not scheduled in the current solution.
-It then identifies a set $C$ of scheduled customer's with the same time slot as $i$.
-Elements of $C$ are removed and attempted to be rescheduled along with $i$.
-
-% The application of these local search operators is performed iteratively, allowing for multiple rounds of neighborhood exploration and solution refinement.
-Each operator is applied sequentially, and the resulting solution is evaluated to determine if an improvement has been achieved.
-The process continues until no further improvements can be made, indicating that a locally optimal solution has been reached.
+\subsection{Local Search}
+
+After the construction phase, the local search phase iteratively invokes five operators using a first-improvement strategy.
+The first operator (\textit{Swap Load}) exchanges loads between two deliveries of the same order.
+The \textit{Swap Driver} operator explores alternative driver assignments by swapping drivers assigned to two delivery nodes.
+The \textit{Relocate} operator removes a delivery node and places it in a different location.
+The \textit{Consolidate Load} operator assigns an order request initially split between two drivers to a single driver, if the driver's capacity allows.
+The \textit{Remove and Reschedule} operator selects an unscheduled customer, denoted $i$, within the current solution.
+It then identifies a set $C$ containing scheduled customers with the same time slot as $i$.
+All elements of $C$ are removed and rescheduled along with $i$.
+This local search process continues until no further improvements can be made or a timeout is reached.
 
@@
 
-After the construction and local search steps, it may happen that some customer orders cannot be scheduled within a solution.
+After the construction and local search steps, it is possible that some customers remain unscheduled in one solution while being accommodated in another.
 The goal of the path relinking procedure is to use each customer's scheduling information from existing solutions and combine them effectively to construct a solution that schedules all customers.
-This algorithm iterates through each element of a pool of solutions and attempts to insert the customer's deliveries into a newly constructed solution.
-
-\section{ Experimental results}
+This algorithm iterates through each element of a pool of solutions and attempts to insert each customer's schedule into a newly constructed solution.
+
+\section{Computational results}
 \label{comp_exp}

### 2023-08-09 Article/article.tex
Save article

@@
 
-\section{Computational results}
+\section{Computational experiments}
 \label{comp_exp}
@@
 The GRASP algorithm is coded in C++.
-We run our experiments on a dataset extracted from data provided by our industrial partner.
+We run our experiments on a dataset extracted from delivery operations records provided by our industry partner and on the benchmark data used in \cite{kinable2014concrete}.
 
 \subsection{Generation of instances}
-
-The dataset used in this section was obtained from our partner and contains records representing delivery operations from up to 8 production centers over a period of 50 days.
-We generated 50 instances, with each instance corresponding to a daily operation.
-The instances include informations such as each driver capacity, associated depot, and shift start time.
-For each customer, we have the due time, total demand, and the number of orders (type of concrete) received.
-For each order, we specified its demand and associated production plant.
-Finally, we include loading capacities for each production center.
-The name of an instance has the format $C\_k\_n\_o\_z$. $k$ is the number of available drivers, $n$ the number of customers, $o$ the total number of daily orders, and $z$ the number of plants.
-
-The dataset is subdivised in a small, medium and large sets, according to the number of drivers and total demands served. We summarized both sets in Table~\ref*{}
-or 
-The dataset is subdivised in a small, medium and large sets, according to the number of orders served. We summarized both sets in Table~\ref*{}
-
-% This dataset allows us to evaluate the effectiveness and performance of our GRASP algorithm in optimizing delivery operations in a real-world setting.
-
-\begin{itemize}
-    \item   8 production centers located in greater Montreal
-    \item   42 days of delivery operations from Fall 2020
-    \item   From 7 to 139 orders each day
-    \item   Up to 137 drivers
-    \item   2 types of trucks of capacity 8 and 12 m$^3$
-\end{itemize}
+\newcommand{\nbInstance}{36}
+The dataset used in this section was obtained from our partner and contains records representing delivery operations from up to 8 production centers over a period of $\nbInstance$ days.
+We have created $\nbInstance$ instances, where each instance corresponds to a daily operation.
+An instance contains information such as each driver's capacity, associated batch plant, and shift start time.
+For each customer, we have the due time, the total demand, and the number of orders (type of concrete) received.
+For each order, we have its demand and associated production plant.
+Finally, we have the loading capacity for each plant.
+The name of an instance has the format $C\_k\_n\_o\_z$, where $k$ is the number of available drivers, $n$ is the number of customers, $o$ is the total number of daily orders, and $z$ is the number of plants.
+The dataset is divided into small, medium, and large sets according to the total number of daily orders.
+We have summarized the data set in the table~\ref{tab:instances}.
+
+\begin{table}[htpb]
+    \caption{Instances summary}
+    \label{tab:instances}
+    \resizebox*{\textwidth}{!}{
+    \begin{tabular}{lcccccc}
+        \toprule
+        {} &  \# &           Load  &   \#Order & \#Client &  \#Driver & \#Depot \\
+          &       &         (m$^3$)        &         &         &          &        \\
+        \midrule
+        Small  &     8 &    226.5-957.5 &     7-15 &    4-11 &    13-39 &    1-3 \\
+        Medium &    14 &  1,164.5-2,971.0 &    44-98 &   41-89 &   76-140 &    6-8 \\
+        Large  &    14 &  3,078.5-3,953.5 &  107-137 &  92-128 &  133-153 &    8-8 \\
+        \bottomrule
+        \end{tabular}  
+    }
+\end{table}
 
@@
 
-\begin{tabular}{lrrrrrrr}
+    \begin{tabular}{llrrrrrrrr}
         \toprule
-    {} &  satisfiedCost &  driverUsed &  travelCost &  firstDeliveryCost &  underWorkCost &  overTimeCost &  runtime \\
-    Name     &                &             &             &                    &                &               &          \\
+              &                 &  satisfiedCost &  travelCost &  firstDeliveryCost &  maxfirstDelivery &  underWorkCost &  overTimeCost &  driverUsed &  runtime \\
+        size & Instance &                &             &                    &                   &                &               &             &          \\
         \midrule
-    20201121 &          226.5 &          13 &     1243.55 &               0.51 &           0.00 &          0.00 &     5.26 \\
-    20201010 &          267.0 &          13 &      745.10 &               0.00 &           0.00 &          0.00 &     5.32 \\
-    20201107 &          333.5 &          18 &      777.25 &               0.00 &         200.16 &          0.00 &     5.19 \\
-    20201031 &          375.0 &          15 &     1017.64 &               0.00 &         216.06 &         28.10 &     5.42 \\
-    20201114 &          388.0 &          19 &     1743.64 &               0.00 &         149.48 &          0.00 &     5.42 \\
-    20201128 &          776.0 &          31 &     2278.18 &               0.00 &           4.19 &        438.14 &     5.16 \\
-    20201024 &          937.5 &          35 &     4403.57 &               0.00 &          58.30 &       1005.24 &     5.72 \\
-    20201021 &         1160.5 &          73 &     6107.22 &               0.00 &        2535.40 &        185.02 &     5.28 \\
-    20201013 &         1565.0 &         102 &     9214.74 &              19.53 &        2955.35 &         95.74 &     5.32 \\
-    20201026 &         1839.5 &         116 &    13585.60 &              44.44 &        1826.70 &        227.56 &     5.41 \\
-    20201007 &         2512.5 &         127 &    15166.00 &               5.09 &         877.68 &        574.76 &     5.67 \\
+        small & C\_15\_4\_7\_2 &          375.0 &     1013.11 &               0.00 &              0.00 &          56.79 &        244.61 &          15 &  3009.54 \\
+              & C\_19\_7\_8\_2 &          388.0 &     1714.74 &               0.00 &              0.00 &          97.56 &        283.83 &          19 &  3020.89 \\
+              & C\_13\_5\_8\_1 &          267.0 &      745.10 &               0.00 &              0.00 &           0.00 &          0.00 &          13 &  3027.39 \\
+              & C\_31\_8\_11\_3 &          776.0 &     2297.20 &               0.00 &              0.00 &         627.37 &        539.96 &          30 &  3033.85 \\
+              & C\_39\_10\_11\_3 &          937.5 &     4077.20 &               0.00 &              0.00 &         720.00 &        866.98 &          31 &  3020.75 \\
+              & C\_18\_6\_11\_2 &          333.5 &      753.26 &               0.00 &              0.00 &         653.78 &          0.00 &          17 &  3004.51 \\
+              & C\_13\_11\_12\_1 &          226.5 &     1243.55 &               0.00 &              0.00 &          66.01 &          0.00 &          13 &  3022.06 \\
+              & C\_35\_11\_15\_3 &          613.5 &     3724.29 &              12.41 &             12.41 &         246.04 &        454.51 &          28 &  3035.23 \\
+        Medium & C\_76\_41\_44\_6 &         1160.5 &     5715.53 &               1.00 &              1.00 &        3566.49 &        771.23 &          69 &  3021.92 \\
+              & C\_104\_42\_47\_8 &         1565.0 &     9114.62 &              23.00 &             16.24 &        4180.95 &        671.94 &          94 &  3047.99 \\
+              & C\_98\_63\_70\_7 &         1746.5 &    13888.90 &              19.86 &             12.40 &        1086.60 &       1477.76 &          94 &  3025.72 \\
+              & C\_118\_57\_70\_6 &         2327.5 &    15066.80 &              83.79 &             16.44 &        1522.11 &       1892.64 &         117 &  3020.70 \\
+              & C\_130\_68\_74\_7 &         2595.0 &    19133.50 &             251.77 &             54.40 &        1523.19 &       2924.28 &         128 &  3053.96 \\
+              & C\_117\_67\_78\_8 &         1839.5 &    13946.30 &              44.68 &              9.98 &        1578.90 &        639.28 &         116 &  3007.88 \\
+              & C\_128\_66\_80\_8 &         2512.5 &    15177.80 &               2.00 &              2.00 &        1874.83 &        965.65 &         125 &  3007.70 \\
+              & C\_120\_71\_82\_8 &         2060.0 &    12461.20 &              68.38 &             20.06 &        3119.66 &       1478.63 &         107 &  3028.24 \\
+              & C\_140\_79\_83\_7 &         2425.0 &    18811.40 &             252.94 &             41.65 &        2193.63 &       1963.05 &         136 &  3025.83 \\
+              & C\_129\_78\_85\_8 &         2685.5 &    16927.30 &             202.26 &             50.12 &        1492.50 &       2794.21 &         128 &  3030.32 \\
+              & C\_133\_80\_89\_8 &         2893.5 &    17572.10 &             187.42 &             49.24 &        1803.22 &       1306.94 &         130 &  3040.74 \\
+              & C\_136\_85\_97\_8 &         2673.0 &    16832.10 &             154.97 &             50.21 &        3507.40 &       1958.58 &         132 &  3028.15 \\
+              & C\_137\_89\_97\_7 &         2939.5 &    21934.10 &             697.18 &             59.10 &        1220.15 &       3038.33 &         137 &  3025.45 \\
+              & C\_135\_84\_98\_7 &         2971.0 &    22399.30 &             867.12 &             59.44 &         550.77 &       2887.51 &         133 &  3137.11 \\
+        Large & C\_143\_92\_107\_8 &         3684.5 &    22700.20 &             147.42 &             48.31 &         899.82 &       3702.01 &         142 &  3019.52 \\
+              & C\_141\_98\_108\_8 &         3541.0 &    20550.10 &             566.69 &             57.49 &        1438.11 &       2910.97 &         139 &  3011.84 \\
+              & C\_133\_98\_109\_8 &         3078.5 &    23102.40 &             193.01 &             49.44 &         450.57 &       3931.83 &         132 &  3018.05 \\
+              & C\_133\_101\_119\_8 &         3229.0 &    18978.90 &             467.94 &             59.47 &        1084.78 &       2334.16 &         129 &  3179.81 \\
+              & C\_145\_108\_122\_8 &         3670.5 &    23401.60 &             668.15 &             59.77 &        1438.60 &       3569.62 &         143 &  3056.51 \\
+              & C\_144\_101\_123\_8 &         3437.5 &    24560.70 &             451.36 &             56.61 &        1663.43 &       4359.64 &         142 &  3039.02 \\
+              & C\_153\_99\_123\_8 &         3513.0 &    20696.60 &             839.56 &             52.60 &        2465.87 &       2892.55 &         149 &  3072.58 \\
+              & C\_146\_114\_129\_8 &         3499.5 &    21991.50 &             898.10 &             55.22 &        1750.60 &       4943.03 &         142 &  3011.54 \\
+              & C\_142\_112\_129\_8 &         3471.0 &    21526.10 &             252.62 &             58.09 &         624.87 &       3360.64 &         137 &  3018.76 \\
+              & C\_153\_112\_131\_8 &         3953.5 &    25455.30 &             868.81 &             49.15 &         500.68 &       3775.57 &         148 &  3005.96 \\
+              & C\_144\_114\_132\_8 &         3401.5 &    23259.00 &            1092.82 &             57.99 &         634.53 &       2705.26 &         140 &  3046.75 \\
+              & C\_143\_114\_136\_8 &         3350.5 &    21760.00 &            1131.39 &             59.58 &        1284.13 &       2360.19 &         141 &  3056.56 \\
+              & C\_153\_128\_137\_8 &         3946.5 &    25822.20 &             742.75 &             57.10 &         899.92 &       5101.58 &         150 &  3060.79 \\
+              & C\_141\_115\_137\_8 &         3739.5 &    24670.70 &             750.04 &             55.88 &         518.63 &       5156.41 &         137 &  3021.26 \\
         \bottomrule
         \end{tabular}
+
     }

### 2023-08-15 solution/cq/gantt.py
save

@@ -1,0 +1,90 @@
+import pandas as pd
+import numpy as np
+import plotly.express as px
+
+# Transform minutes to format"hh:mm:ss:"
+def time_to_date(date,t):
+    date=str(date)
+    if len(date.split('-'))==1:
+        date="2022-05-16"
+    else:
+        date=date[:4]+"-"+date[4:6]+"-"+date[6:]
+    min=int(t%60)
+    hour=int(t//60)
+    if min<10:
+        min=f"0{min}"
+    new_date=f"{date} {hour}:{min}:00"
+    if hour>23: 
+        print(new_date,' ',t)
+    return new_date
+
+
+def show(df):
+    df['task'] = df.apply(lambda x: f" del {x.Del}",axis=1)
+    fig = px.timeline(df, x_start="start", x_end="end", y="Del",title="All deliveries",color="task")
+    # fig = px.timeline(df, x_start="startLoad", x_end="endLoad", y="task",title="All deliveries",color="Cust")
+    fig.update_yaxes(autorange="reversed") # otherwise tasks are listed from the bottom up
+    fig.show()
+
+# Creer des diagrammes de Gantt
+# Par ordre: afficher les heures d'arrivée et de départ des différents noeuds
+# J'ai besoin de la date, des heures d'arrivée et de départ.
+def showGanttDel(df):
+    df['task'] = df.apply(lambda x: f" del {x.Del}_{x.Order}_{x.Cust}",axis=1)
+    fig = px.timeline(df, x_start="startUnload", x_end="endUnload", y="task",title="All deliveries",color="Cust")
+    # fig.update_yaxes(autorange="reversed") # otherwise tasks are listed from the bottom up
+    fig.show()
+
+def showGanttDock(df):
+    df['task'] = df.apply(lambda x: f" dock {x.Del}_{x.Order}_{x.Cust}",axis=1)
+    fig = px.timeline(df, x_start="startLoad", x_end="endLoad", y="task",title="All Loading",color="Cust")
+    # fig.update_yaxes(autorange="reversed") # otherwise tasks are listed from the bottom up
+    fig.show()
+
+# Creer des diagrammes de Gantt
+# Par ordre: afficher les heures d'arrivée et de départ des différents noeuds
+# J'ai besoin de la date, des heures d'arrivée et de départ.
+def showGanttOrder(df):
+    orderList=np.unique(df['ordreId'])
[…]

### 2023-08-15 Article/article.tex
Save article

--- 2023-08-08 Article/article.tex
@@
 This means that a truck can only haul RMC for one order, even if there is spare capacity.
-After unloading the RMC,  driver $k$ takes $Ct_k$ minutes to clean the concrete mixer before proceeding.
+After unloading the RMC,  driver $k$ takes $\rho$ minutes to clean the concrete mixer before proceeding.
 %When assigning drivers to deliveries, the company prioritizes employees with the highest seniority.
@@
 Let $g_i$ be the time between due date and first service start for customer $i$.
-% Binary variable $u_{o_{1},o_{2}}$ equals $1$ if order $o_{2}$ is served directly after order $o_{1}$.
-Binary variables $z_{o,d}$ ($z_{i,d}$) indicates whether or not delivery $d \in  \mathcal{D}_o$ ($d \in  \mathcal{D}_i$) is supposed to be executed.
-$w_i$ is the end of the loading (unloading) operation at node $i \in L\cup D$.
-
-The objective function minimizes the travel cost, the penalty costs incurred when customer demands are not fully met, the lateness of each customer's first delivery, the cost of drivers working less than the minimum hours, and the total overtime cost of drivers working beyond their scheduled hours.
+
+The objective function minimizes total travel cost ($TC$), penalties associated with unfulfilled orders ($PUO$), first delivery delays ($FDD$), driver underutilization costs ($DUC$), and driver overtime costs ($DOC$).
+Together, these components drive the optimization process to find a solution that efficiently balances travel costs, customer satisfaction, on-time delivery, driver utilization, and scheduling constraints.
+
+% The objective function minimizes the total travel cost, the penalty costs incurred when customer demands are not fully met, the lateness of each customer's first delivery, the cost of drivers working less than the minimum hours, and the total overtime cost of drivers working beyond their scheduled hours.
 The mathematical model of this variant of the CDP is as follows:
 
-%The objective function (\ref{mod:obj}) is designed to minimize multiple factors that impact the overall cost and efficiency of the concrete delivery process.
-These factors include: The total traveled distance, which represents the fuel consumption and wear and tear on the delivery vehicles.
-The penalty costs $\beta_1$, which are incurred when customers' demands are not fully satisfied.The lateness of each customer's first delivery, which can impact customer satisfaction and lead to additional penalty costs.The number of drivers who work under the minimum working time, which can lead to increased labor costs and decreased productivity.The total overtime cost, which is incurred when drivers work beyond their scheduled hours.All these factors are taken into consideration in the objective function to minimize the cost and maximize the satisfaction of the customers.
-
 \begin{alignat}{3}
-    \label{mod:obj}   \min & \smashoperator{ \sum_{ \substack{(i,j,k) \in A} }} {t_{ij}x^{k}_{ij}} + \beta_1\smashoperator{\sum_{ o \in \mathcal{O}}}{  \left(1-y_o\right)} \hspace{-0mm}  + \beta_2 \smashoperator{ \sum_{i \in \mathcal{C} } } { g_i  } +  \sum_{k \in K}{\beta_3 w^{1}_k } + \beta_4 { w^{2}_k } &                                                                                                  \\
+    \label{mod:obj}   \min & \smashoperator{ \sum_{ \substack{(i,j,k) \in A} }} {t_{ij}x^{k}_{ij}} + \beta_1\smashoperator{\sum_{ o \in \mathcal{O}}}{  \left(1-y_o\right)} \hspace{-0mm}  + \beta_2 \smashoperator{ \sum_{i \in \mathcal{C} } } { g_i  } +  \sum_{k \in K}\left({\beta_3 w^{1}_k  + \beta_4  w^{2}_k}\right)  &                         \\
     \label{mod:c1}         & \sum_{j \in \delta^{+}(i) }{x^k_{ij}} =1       & \forall i \in S, k \in K                                                                         \\
@@
     \label{mod:c3}         & v_j \geq  w_i + \alpha_{b} + t_{ij} - M\left(1-x^{k}_{ij}\right)                                                                                                                                                                                                                       & \hspace{-20mm} \forall b \in \mathcal{B}, i \in \mathcal{L}_b, j \in \delta^{+}(i),  k \in K     \\
-    \label{mod:c4}         & v_j \geq  w_i + Ct_{k} + t_{ij} - M\left(1-x^{k}_{ij}\right)                                                                                                                                                                                                                           & \hspace{-20mm} \forall i \in \mathcal{D}  , j \in \delta^{+}(i), k \in K                         \\
+    \label{mod:c4}         & v_j \geq  w_i + \rho + t_{ij} - M\left(1-x^{k}_{ij}\right)                                                                                                                                                                                                                           & \hspace{-20mm} \forall i \in \mathcal{D}  , j \in \delta^{+}(i), k \in K                         \\
     \label{mod:c5}         & w_{i} \geq v_{i}  + \frac{q^k_j}{\tau^l_b} -M\left(1- x^{k}_{ij}\right)                                                                                                                                                                                                                & \hspace{-20mm} \forall  b \in \mathcal{B},  i \in \mathcal{L}_{b},  j \in \delta^{+}(i), k \in K \\
@@
 
-\section{Computational results}
+\section{Computational experiments}
 \label{comp_exp}
 
-In this section, we present and discuss the results of the computational experiments of the GRASP heuristic described in section~\ref{grasp_method}.
-The GRASP algorithm is coded in C++.
-We run our experiments on a dataset extracted from data provided by our industrial partner.
+In this section, we present and discuss the results of computational experiments conducted using the GRASP heuristic described in section~\ref{grasp_method}.
+The implementation of the GRASP algorithm is coded in C++.
+We run our experiments on the benchmark data used in \cite{kinable2014concrete}, and on a dataset extracted from delivery operations records provided by our industry partner.
 
 \subsection{Generation of instances}
-
-The dataset used in this section was obtained from our partner and contains records representing delivery operations from up to 8 production centers over a period of 50 days.
-We generated 50 instances, with each instance corresponding to a daily operation.
-The instances include informations such as each driver capacity, associated depot, and shift start time.
-For each customer, we have the due time, total demand, and the number of orders (type of concrete) received.
-For each order, we specified its demand and associated production plant.
-Finally, we include loading capacities for each production center.
-The name of an instance has the format $C\_k\_n\_o\_z$. $k$ is the number of available drivers, $n$ the number of customers, $o$ the total number of daily orders, and $z$ the number of plants.
-
-The dataset is subdivised in a small, medium and large sets, according to the number of drivers and total demands served. We summarized both sets in Table~\ref*{}
-or 
-The dataset is subdivised in a small, medium and large sets, according to the number of orders served. We summarized both sets in Table~\ref*{}
-
-% This dataset allows us to evaluate the effectiveness and performance of our GRASP algorithm in optimizing delivery operations in a real-world setting.
-
-\begin{itemize}
-    \item   8 production centers located in greater Montreal
-    \item   42 days of delivery operations from Fall 2020
-    \item   From 7 to 139 orders each day
-    \item   Up to 137 drivers
-    \item   2 types of trucks of capacity 8 and 12 m$^3$
-\end{itemize}
-
-\subsubsection{ Results for the generated instances}
-
-- resultats avec les instances où toute la demande est satisfaite. Montrer la distance parcourue, le temps de première livraison, overtime, underwork, temps d'execution.
-
-- resultats avec les instances (8 au total) où uniquement la flotte de Unibéton a éét utilisée. (utiliser les mêmes pour la comparaison avec les données réelles )
-distance, demande servie, first, underwork, overtime.
-- resultats avec les autres instances
-- distance, demande servie, first delivery.
-
-% df = df[df['Name'].isin([20201010,20201021,20201107,20201017,20201031,20201109,20201026,20201121])]
-
-\begin{table}
-    \resizebox*{\textwidth}{!}{
-
-\begin{tabular}{lrrrrrrr}
+\newcommand{\nbInstance}{36}
+The dataset used in this section was obtained from our partner and contains records representing delivery operations from up to 8 production centers over a period of $\nbInstance$ days.
+We have created $\nbInstance$ instances, where each instance corresponds to a daily operation.
+An instance contains information such as each driver's capacity, associated batch plant, and shift start time.
+For each customer, we have the due time, the total demand, and the number of orders (type of concrete) received.
+For each order, we have its demand and associated production plant.
+Finally, we have the loading capacity for each plant.
+The name of an instance has the format $C\_k\_n\_o\_z$, where $k$ is the number of available drivers, $n$ is the number of customers, $o$ is the total number of daily orders, and $z$ is the number of plants.
+The dataset is divided into small, medium, and large sets according to the total number of daily orders.
+We have summarized the dataset in the Table~\ref{tab:instances}.
+
+\begin{table}[htpb]
+    \centering
+    \caption{Instances summary}
+    \label{tab:instances}
+    \resizebox*{0.8\textwidth}{!}{
+    \begin{tabular}{lcccccc}
         \toprule
-    {} &  satisfiedCost &  driverUsed &  travelCost &  firstDeliveryCost &  underWorkCost &  overTimeCost &  runtime \\
-    Name     &                &             &             &                    &                &               &          \\
+        {} &  \# &         Demand  &   \#Order & \#Client &  \#Driver & \#Depot \\
+          &       &         (m$^3$)        &         &         &          &        \\
         \midrule
-    20201121 &          226.5 &          13 &     1243.55 &               0.51 &           0.00 &          0.00 &     5.26 \\
-    20201010 &          267.0 &          13 &      745.10 &               0.00 &           0.00 &          0.00 &     5.32 \\
-    20201107 &          333.5 &          18 &      777.25 &               0.00 &         200.16 &          0.00 &     5.19 \\
-    20201031 &          375.0 &          15 &     1017.64 &               0.00 &         216.06 &         28.10 &     5.42 \\
-    20201114 &          388.0 &          19 &     1743.64 &               0.00 &         149.48 &          0.00 &     5.42 \\
-    20201128 &          776.0 &          31 &     2278.18 &               0.00 &           4.19 &        438.14 &     5.16 \\
-    20201024 &          937.5 &          35 &     4403.57 &               0.00 &          58.30 &       1005.24 &     5.72 \\
-    20201021 &         1160.5 &          73 &     6107.22 &               0.00 &        2535.40 &        185.02 &     5.28 \\
-    20201013 &         1565.0 &         102 &     9214.74 &              19.53 &        2955.35 &         95.74 &     5.32 \\
-    20201026 &         1839.5 &         116 &    13585.60 &              44.44 &        1826.70 &        227.56 &     5.41 \\
-    20201007 &         2512.5 &         127 &    15166.00 &               5.09 &         877.68 &        574.76 &     5.67 \\
+        Small  &     8 &    226.5-937.5 &     7-14 &    4-11 &    13-35 &    1-3 \\
+        Medium &    14 &  1,160.5-2,971.0 &    43-98 &   40-89 &   76-137 &    6-8 \\
+        Large  &    14 &  3,078.5-3,953.5 &  107-136 &  92-127 &  129-150 &    8 \\
         \bottomrule
@@
 
-\subsubsection{ Results for the instances of \cite{kinable2014concrete}}
+\subsubsection{Results for the instances of \cite{kinable2014concrete}}
 
 \cite{kinable2014concrete} provided a benchmark dataset for their variant of the $CDP$.
-Although their primary objective is to maximize the total load delivered each day, without considering loading operations at a plant and ensuring deliveries within specified time windows, we can still use this dataset to analyze and compare the performance of our algorithm under different conditions.
+Although their primary objective is to maximize the total load delivered each day, without considering loading operations at a plant and ensuring deliveries within specified time windows, we still use this dataset to analyze and compare the performance of our algorithm under different conditions.
 Table \ref*{tab:cdp_instances} shows the performance of the GRASP heuristic along with the other methods (CP, MIP, SD-heuristic, and Hybrid) used in \cite{kinable2014concrete}.
 For each method, the table shows the average execution time in milliseconds and the corresponding gap between the obtained solution and the upper bound provided by \cite{kinable2014concrete}.
-Overall, the GRASP algorithm shows promise in providing near-optimal solutions in a relatively short time for this benchmark of the CDP.
-It outperformed all methods except CP, achieving an average distance of 5.7\% (13.8\%) with a runtime of 13,186 (574,983) milliseconds for Set A (Set B).
-The SD-heuristic appears to be much faster, but GRASP provides a better balance between solution quality and computation time.
-It should also be noted that the exact MIP, CP, and Hybrid are initialized with the results of the SD-heuristic.
 
@@
     \centering
-    \caption{Performance of the GRASP heuristic on the CDP instances}
+    \caption{Performance of the GRASP heuristic on the CDP benchmark instances}
     \label{tab:cdp_instances}
@@
 
+Overall, the GRASP algorithm shows promise in providing near-optimal solutions in a relatively short time for this benchmark of the CDP, making it a suitable candidate for further analysis and real-world application.
+It outperformed all methods except CP, achieving an average distance of 5.7\% (13.8\%) with a runtime of 13,186 (574,983) milliseconds for Set A (Set B).
+The SD-heuristic appears to be much faster, but GRASP provides a better balance between solution quality and computation time.
+It should also be noted that the exact MIP, CP, and Hybrid are initialized with the results of the SD-heuristic.
+
+\subsubsection{Results for the generated instances}
+
+The computational results of our instances solved with the GRASP heuristic are summarized in the tables below.
+The stopping criterion is set to 20,000 iterations.
+The parameters used in our experiments are listed in Table~\ref*{tab:problem_parameters}.
+
+\begin{table}[htpb]
+    \centering
+    \caption{Parameters of the real-world instances}
+    \label{tab:problem_parameters}
+%    \resizebox*{\textwidth}{!}{
+    \begin{tabularx}{\textwidth}{Xr}
+        \toprule
+        \multicolumn{1}{c}{Parameter} & \multicolumn{1}{c}{Value} \\ \midrule
+        Unloading duration ($\tau_c^u$) & 2 min$/m^3$ \\
+        Cleaning duration ($\rho$) & 10 min \\
+        Adjustment duration ($\alpha$) & 10 min \\
+        Max travel time ($\Delta$) & 120 min \\
+        Min working time ($M_T$) & 180 min \\
+        Max working time ($N_T$) & 480 min \\
+        Max overtime duration ($O_T$) & 120 min \\
+        Max delay between two consecutive deliveries of the same order ($\gamma^1$) & 20 min \\
+        Max delay between two consecutive deliveries of different orders ($\gamma^2)$ & 25 min \\
+    %    $\beta_1$ & 10000  \\
+    %    $\beta_2$ & 10000  \\
+    %    $\beta_3$ & 1000  \\
+    %    $\beta_4$ & 100  \\
+        \bottomrule
+    \end{tabularx}
+%    }
+\end{table}
+
+Table \ref*{tab:runtime} shows the percentage of concrete delivered by varying the execution times of the algorithm.
+For small instances, all requests are delivered in less than five minutes, while for medium and large instances, over 99\% of requests are delivered.
+For these instances, it takes more than ten minutes to ensure complete delivery of all orders.
+
+\begin{table}[htbp]
+    \centering
+    \caption{RMC delivered within different runtimes}
+    \label{tab:runtime}
+    \resizebox*{0.5\textwidth}{!}{
+        \begin{tabular}{llcccc}
+            \toprule
+            &  &      \multicolumn{4}{c}{Runtime (min)} \\
+            \cmidrule{3-6}
+            & &      5   & 10  &  30  & 60  \\
+            %  & {} &         &         &        &        \\
+            \midrule
+            \multirow{1}{*}{Small} & \% Load &  100.00 &  100.00 &  100.0 &  100.0 \\
+                %   & \# instances &    8 &    8 &    8 &    8 \\
+            \cmidrule{2-6}
+                  \multirow{1}{*}{Medium} & \% Load &   99.86 &   99.86 &  100.0 &  100.0 \\
+                %   & \# instances &   13 &   13 &   14 &   14 \\
+                  \cmidrule{2-6}
+                  \multirow{1}{*}{Large} & \% Load &   99.79 &   99.93 &  100.0 &  100.0 \\
+                %   & \# instances &   12 &   13 &   14 &   14 \\
+            \bottomrule
+            \end{tabular} 
+    }
+
+\end{table}
+
+Table \ref*{tab:results_grasp_ls} shows the detailed results for each instance.
+The runtime is set to 3600 seconds.
+The table reports the requirements of each instance, the components of the objective function, the number of drivers used, and the execution time in seconds.
+%  the travel cost ($TC$), the penalties associated with unfulfilled orders ($PUO$), the sum of the first delivery delays ($FDD$) for each customer, the sum of the driver underutilization costs ($DUC$) and driver overtime costs ($DOC$).
+
+All daily orders are served with our algorithm, using less drivers than the available fleet.
+For the small size instances, all first deliveries are timely, except for one instance.
+The sum of the driver underutilization costs ($DUC$) is high for medium and large instances.
+This indicates underutilization of the scheduled fleet.
+Some drivers are not scheduled at all or work very little.
+The sum of the driver overtime costs ($DOC$) is also high for these instances.
+This can be explained by the fact that the algorithm prefers to put certain drivers on overtime rather than use a driver who has to travel to another plant for a delivery, which would increase the travel cost ($TC$).
+
+$FDD$ is the sum of the delays of the first deliveries.
+To get an idea of the actuel delay per customer, we report $mFDD$, which is the maximum delay for all first deliveries.
+This tells us that we have a maximum delay that varies between 0 and 12 minutes for small instances, 1 and 60 minutes for medium instances, and 50 to 60 minutes for large instances.
+
+\begin{table}[htbp]
+    \caption{Results with all instances}
+    \label{tab:results_grasp_ls}
+    \resizebox*{\textwidth}{!}{
+        \begin{tabular}{llccccccccc}
+        \toprule
+        &   &  Demand &  $TC$ & $PUO$&  $FDD$ & $mFDD$ &  $DUC$ &  $DOC$ &  \#Drivers &  Time \\
+        & \multicolumn{1}{c}{Instance} &  (m$^3$)  &(min)  &  & (min) &(min) &(min)    & (min) & & (s)\\
+        \midrule
+            \multirow{9}{*}{Small} & C\_13\_11\_12\_1 &          226.5 &     1243.55 &  0.0 &               0.00 &              0.00 &          66.01 &          0.00 &          13 &  3022.06 \\
+                  & C\_13\_5\_8\_1 &          267.0 &      745.10 &  0.0 &               0.00 &              0.00 &           0.00 &          0.00 &          13 &  3027.39 \\
+                  & C\_18\_6\_11\_2 &          333.5 &      753.26 &  0.0 &               0.00 &              0.00 &         653.78 &          0.00 &          17 &  3004.51 \\
+                  & C\_15\_4\_7\_2 &          375.0 &     1013.11 &  0.0 &               0.00 &              0.00 &          56.79 &        244.61 &          15 &  3009.54 \\
+                  & C\_19\_7\_8\_2 &          388.0 &     1714.74 &  0.0 &               0.00 &              0.00 &          97.56 &        283.83 &          19 &  3020.89 \\
+                  & C\_29\_10\_14\_3 &          613.5 &     3724.29 &  0.0 &              12.41 &             12.41 &         246.04 &        454.51 &          28 &  3035.23 \\
+                  & C\_31\_8\_11\_3 &          776.0 &     2297.20 &  0.0 &               0.00 &              0.00 &         627.37 &        539.96 &          30 &  3033.85 \\
+                  & C\_35\_9\_10\_3 &          937.5 &     4077.20 &  0.0 &               0.00 &              0.00 &         720.00 &        866.98 &          31 &  3020.75 \\
+                  \cmidrule{2-11}
+                  \multirow{14}{*}{Medium} & C\_76\_40\_43\_6 &         1160.5 &     5715.53 &  0.0 &               1.00 &              1.00 &        3566.49 &        771.23 &          69 &  3021.92 \\
+                  & C\_104\_42\_47\_8 &         1565.0 &     9114.62 &  0.0 &              23.00 &             16.24 &        4180.95 &        671.94 &          94 &  3047.99 \\
+                  & C\_94\_63\_70\_7 &         1746.5 &    13888.90 &  0.0 &              19.86 &             12.40 &        1086.60 &       1477.76 &          94 &  3025.72 \\
+                  & C\_116\_67\_78\_8 &         1839.5 &    13946.30 &  0.0 &              44.68 &              9.98 &        1578.90 &        639.28 &         116 &  3007.88 \\
+                  & C\_116\_71\_82\_8 &         2060.0 &    12461.20 &  0.0 &              68.38 &             20.06 &        3119.66 &       1478.63 &         107 &  3028.24 \\
+                  & C\_117\_57\_70\_6 &         2327.5 &    15066.80 &  0.0 &              83.79 &             16.44 &        1522.11 &       1892.64 &         117 &  3020.70 \\
+                  & C\_137\_79\_83\_7 &         2425.0 &    18811.40 &  0.0 &             252.94 &             41.65 &        2193.63 &       1963.05 &         136 &  3025.83 \\
+                  & C\_127\_66\_80\_8 &         2512.5 &    15177.80 &  0.0 &               2.00 &              2.00 &        1874.83 &        965.65 &         125 &  3007.70 \\
+                  & C\_128\_68\_74\_7 &         2595.0 &    19133.50 &  0.0 &             251.77 &             54.40 &        1523.19 &       2924.28 &         128 &  3053.96 \\
+                  & C\_136\_85\_97\_8 &         2673.0 &    16832.10 &  0.0 &             154.97 &             50.21 &        3507.40 &       1958.58 &         132 &  3028.15 \\
+                  & C\_128\_78\_85\_8 &         2685.5 &    16927.30 &  0.0 &             202.26 &             50.12 &        1492.50 &       2794.21 &         128 &  3030.32 \\
+                  & C\_131\_77\_85\_8 &         2893.5 &    17572.10 &  0.0 &             187.42 &             49.24 &        1803.22 &       1306.94 &         130 &  3040.74 \\
+                  & C\_137\_89\_97\_7 &         2939.5 &    21934.10 &  0.0 &             697.18 &             59.10 &        1220.15 &       3038.33 &         137 &  3025.45 \\
+                  & C\_133\_84\_98\_7 &         2971.0 &    22399.30 &  0.0 &             867.12 &             59.44 &         550.77 &       2887.51 &         133 &  3137.11 \\
+                  \cmidrule{2-11}
+                  \multirow{14}{*}{Large} & C\_132\_98\_109\_8 &         3078.5 &    23102.40 &  0.0 &             193.01 &             49.44 &         450.57 &       3931.83 &         132 &  3018.05 \\
+                  & C\_129\_101\_119\_8 &         3229.0 &    18978.90 &  0.0 &             467.94 &             59.47 &        1084.78 &       2334.16 &         129 &  3179.81 \\
+                  & C\_141\_114\_136\_8 &         3350.5 &    21760.00 &  0.0 &            1131.39 &             59.58 &        1284.13 &       2360.19 &         141 &  3056.56 \\
+                  & C\_140\_114\_132\_8 &         3401.5 &    23259.00 &  0.0 &            1092.82 &             57.99 &         634.53 &       2705.26 &         140 &  3046.75 \\
+                  & C\_143\_101\_123\_8 &         3437.5 &    24560.70 &  0.0 &             451.36 &             56.61 &        1663.43 &       4359.64 &         142 &  3039.02 \\
+                  & C\_137\_112\_129\_8 &         3471.0 &    21526.10 &  0.0 &             252.62 &             58.09 &         624.87 &       3360.64 &         137 &  3018.76 \\
+                  & C\_142\_114\_129\_8 &         3499.5 &    21991.50 &  0.0 &             898.10 &             55.22 &        1750.60 &       4943.03 &         142 &  3011.54 \\
+                  & C\_149\_98\_122\_8 &         3513.0 &    20696.60 &  0.0 &             839.56 &             52.60 &        2465.87 &       2892.55 &         149 &  3072.58 \\
+                  & C\_139\_98\_108\_8 &         3541.0 &    20550.10 &  0.0 &             566.69 &             57.49 &        1438.11 &       2910.97 &         139 &  3011.84 \\
+                  & C\_144\_108\_122\_8 &         3670.5 &    23401.60 &  0.0 &             668.15 &             59.77 &        1438.60 &       3569.62 &         143 &  3056.51 \\
+                  & C\_142\_92\_107\_8 &         3684.5 &    22700.20 &  0.0 &             147.42 &             48.31 &         899.82 &       3702.01 &         142 &  3019.52 \\
+                  & C\_138\_114\_136\_8 &         3739.5 &    24670.70 &  0.0 &             750.04 &             55.88 &         518.63 &       5156.41 &         137 &  3021.26 \\
+                  & C\_150\_127\_136\_8 &         3946.5 &    25822.20 &  0.0 &             742.75 &             57.10 &         899.92 &       5101.58 &         150 &  3060.79 \\
+                  & C\_148\_112\_131\_8 &         3953.5 &    25455.30 &  0.0 &             868.81 &             49.15 &         500.68 &       3775.57 &         148 &  3005.96 \\
+            \bottomrule
+            \end{tabular}
+    }
+\end{table}
+
+We tried to reduce the maximum delay of the first delivery but found that some orders couldn't be fulfilled because of driver shifts and overtime constraints.
+We decided to remove these constraints and rerun our computation.
+We found that for small instances, the maximum delay dropped to zero, and for medium and large instances, it was reduced by almost 50\%.
+This suggests that when we input a good work schedule, our algorithm works effectively to ensure that all orders are filled and that delays for first deliveries and travel costs are minimized.
+In Table~\ref*{tab:lift_shift_constraints}, we can see that the average travel time decreases for all instances except the large ones.
+That's because the goals of minimizing travel cost and not delaying the first delivery are sometimes in conflict: to ensure that the first delivery is not late, drivers assigned to one plant often have to travel to another.
+And since all plants are used in large cases, this can increase travel times.
+
+%  Therefore, we can use our algorithm for fleet sizing or as a submodule of a driver scheduling algorithm.
+
+\begin{table}[htbp]
+    \centering
+    \caption{Comparaison with and without shift and overtime constraints}
+    \label{tab:lift_shift_constraints}
+    \resizebox*{0.8\textwidth}{!}{
+        \begin{tabular}{lcccccccc}
+            \toprule
+        {} & \multicolumn{4}{c}{Default parameters} &\multicolumn{4}{c}{$M_T=0$, $N_T,O_T=\infty$, $H_k=0$} \\
+        \cmidrule(l){2-5} \cmidrule(l){6-9}
+        {} &   Load & $TC$ &   $FDD$ &  $mFDD$ &   Load &      $TC$ &   $FDD$ &   $mFDD$ \\
+               &($m^3$)  & (min)   &(min)   &(min)  &($m^3$)   &  (min) &(min)  & (min)   \\
+            \midrule
+            Small  &   489.62 &   1946.06 &    1.55 &       1.55 &   489.62 &   1832.90 &    0.00 &       0.00 \\
+            Medium &  2313.86 &  15641.50 &  204.03 &      31.59 &  2313.86 &  14607.66 &   74.76 &      16.56 \\
+            Large  &  3536.86 &  22748.24 &  647.90 &      55.48 &  3536.86 &  22293.04 &  210.14 &      29.00 \\
+            \bottomrule
+            \end{tabular}
+
+    }
+
+\end{table}
+
+The influence of the parameter $\gamma^1$ on the solution quality is demonstrated in Figure~\ref*{fig:gamma1_influence}.
+Tests were conducted by varying $\gamma^1$ values between 10 and 25 minutes, without any driver shifts or overtime constraints.
+In small instances, all customers are served regardless of the value of $\gamma^1$.
+For medium (large) instances, all customers are served with $\lambda^1$ starting at 15 (20) minutes.
+The delay of the initial delivery decreases as the value of $\gamma^1$ increases.
+The results indicate that if $\gamma^1$ is high, the plant has the flexibility to delay ongoing deliveries to make room for new ones or make customers wait until the drivers are available.
+By increasing the maximum delay between two successive deliveries from 10 to 20 minutes, the first delivery delay is reduced by over 50\% for the medium and large size instances.
+The average travel time remains constant for all instances, regardless of the value of $\gamma^1$.
+
+\begin{figure}[htbp]
+    \centering
+    \includegraphics[width=0.8\textwidth]{gamma1.pdf} 
+    \caption{Influence of the maximum time delay between consecutive deliveries of the same order ($\gamma^1$)}
+    \label{fig:gamma1_influence}
+\end{figure}
+
+We checked how well the local search part of the GRASP algorithm works by running tests using only constructive heuristics.
+The results are presented in Table~\ref*{tab:grasp_ls_nols}.
+For medium and large instances, the use of local search was necessary to service all customers within 60 minutes.
+And without local search, most objective metrics experienced a significant increase, except for travel costs.
+
+\begin{table}[htbp]
+    \centering
+    \caption{Influence of the local search step on the  GRASP algorithm}
+    \label{tab:grasp_ls_nols}
+    \resizebox*{1\textwidth}{!}{
+    \begin{tabular}{lccccccccccccc}
+        \toprule
+           & &\multicolumn{5}{c}{GRASP$\backslash$LS} & \multicolumn{5}{c}{GRASP}   \\
+           \cmidrule(l){2-14}
+        {size} &Load &&$TC$&   $FDD$ &  $DUC$ &  $DOC$ & Time&&  $TC$ &  $FDD$&  $DUC$ &  $DOC$&Time \\
+        {} & (m$^3$)&\%Served&(min)&   (min) &  (min) &  (min) & (min)&\%Served&  (min) &  (min)&  (min) &  (min)& (min) \\
+        \cmidrule(l){2-2} \cmidrule(l){3-8} \cmidrule(l){9-14}
+        Small  & 3917.0& 100 &       1998.75 &   64.60 &           231.63 &          215.43 &21.24& 100&      1946.06 &    1.55 &    229.40 &          298.74 &50.34\\
+        Medium & 32394.0& 99.6&     13144.05 &   2528.33 &  2496.47 & 1086.12 &49.41&100&  13564.08 &   75.77 &   2250.66 &         1347.59&50.6 \\
+        Large  & 49516.0& 99.2& 21395.23 & 4885.11 &  1250.14 &  3543.16 & 50.20&100&     21784.95 &  471.62 & 1168.73 &  3640.04&50.74 \\
+        \bottomrule
+        \end{tabular}
+    }  
+\end{table}
+
 \subsubsection{Comparisons with the real operations}
 
+- resultats avec les instances (8 au total) où uniquement la flotte de Unibéton a éét utilisée.
+enlver les contraintes de temps. reduire le gamma1
+
+(utiliser les mêmes pour la comparaison avec les données réelles )
+distance, demande servie, first, underwork, overtime.
+- resultats avec les autres instances
+- distance, demande servie, first delivery.
 - rouler les instances sans shift des livreurs. comparer la distance parcourue, et demande livrée.
@@
 
+Il serait intéressant de voir l'effet des paramètres lambda 1 et lambda 2 différents par clients et /ou ordre. Également le temps de service différent par client.
+
+l'algorithme serait intéressant à utiliser dans le cadre d'un
+
+%  Therefore, we can use our algorithm for fleet sizing or as a submodule of a driver scheduling algorithm.
+
 \vspace{0.1in}

### 2023-08-15 Article/versions/04-03-23/article_04_03_23.tex
Save

--- 2023-08-08 Article/article.tex
@@
 \usepackage[T1]{fontenc}
-\usepackage{graphicx}
 \usepackage{natbib}
@@
 Additionally, the problem also considers the possibility of customers requiring multiple types of concrete to be delivered within the same time window.
-The performance of the proposed heuristic is evaluated on real-world instances, and it is compared to another variant to demonstrate its effectiveness.
+The performance of the proposed heuristic is evaluated on real-world and generated CDP instances, and it is compared to another variant to demonstrate its effectiveness.
 
@@
 In this paper, we study a variant of the CDP to schedule the daily production and dispatching of RMC for a company located in the province of Quebec, Canada.
-This company operates multiple batch plants with varying production rates, using a fleet of concrete mixers of different capacities.
+This company operates multiple batching plants with varying production rates, using a fleet of concrete mixers of different capacities.
 Each plant has its own fleet of trucks; however, under certain conditions, the trucks can move between plants if necessary.
@@
 
-In this paper, we propose a mathematical model and a Greedy Randomized Search Procedure (GRASP) heuristic to solve a new variant of the CDP.
+In this paper, we propose a mathematical model and a Greedy Randomized Search Procedure (GRASP) heuristic solution to solve a new variant of the CDP.
 Our model takes into consideration the working shifts of the drivers and the scheduling of multiple orders within the same time window at a construction site.
@@
 
-Academic Research on concrete batch and delivery began in the late 1990s.
+Academic Research on concrete batching and delivery began in the late 1990s.
 \cite{tommelein1999just} described RMC as a prototypical example of a JIT production system in construction and identified two practices for delivering it.
@@
 \cite{mayteekrieangkrai2015optimized} addressed the same problem with the same data using a bee algorithm (BA) and found better solutions than the GA.
-\cite{lu2005optimized} used the same combination of GA and simulation to determine the optimal number of concrete mixers to be deployed and an optimal schedule for batch and delivering concrete.
+\cite{lu2005optimized} used the same combination of GA and simulation to determine the optimal number of concrete mixers to be deployed and an optimal schedule for batching and delivering concrete.
 Their objective was to minimize the idle time of the site crew due to late concrete deliveries and truck queuing time.
 In this setting, it was also necessary to deliver a batch of mortar on-site to lubricate the unloading pump before the concrete delivery.
-As such, the simulation model also included the batch and delivery of mortar.
+As such, the simulation model also included the batching and delivery of mortar.
 Finding the best RMC fleet size was also the purpose of the discrete-event simulation model proposed by \cite{panas_simulation_based_2013}.
@@
 They ran experiments using real-world instances of a concrete supply chain in the Netherlands and found a reduction in the number of outsourced requests.
-\cite{yan2007optimal} also considered overtime considerations in their paper, which focused on scheduling RMC for one batch plant with two loading docks.
+\cite{yan2007optimal} also considered overtime considerations in their paper, which focused on scheduling RMC for one batching plant with two loading docks.
 The study took into account that overtime wages are paid for factory and construction site operations after 4 PM.
@@
 The procedure is invoked multiple times for different permutations of the customer's order which is determined using the steepest descent (SD) local search procedure.
-One of the paper's main contributions is the creation of the first public test instances for the CDP with up to 50 customers, four batch plants, and 20 concrete mixers.
+One of the paper's main contributions is the creation of the first public test instances for the CDP with up to 50 customers, four batching plants, and 20 concrete mixers.
 They found the CP model to be highly effective in finding high-quality solutions in a relatively short time or improving existing schedules, while the MIP model can be used to compute bounds, as it seems ineffective in solving large problem instances.
@@
 In \cite{kinable2014logic}, we found an attempt to solve the previous problem with a logic-based Benders' approach.
-%\cite{sulaman2017simulated} expand upon the SD-heuristic proposed in \cite{kinable2014concrete}, proposing a simulated annealing (SA) combined with a time-slot Heuristic (SATH).
+\cite{sulaman2017simulated} expand upon the SD heuristic proposed in \cite{kinable2014concrete}, proposing a simulated annealing (SA) combined with a time-slot Heuristic (SATH).
 This method looks for a slot between existing visits of a truck to schedule a new delivery instead of assigning it to the time slot strictly after the truck's latest assigned delivery.
@@
 \label{desc_form}
-The focus of this paper is on the distribution of RMC from a Canadian company that operates in the greater Montreal area.
-When a customer places an order, it is received at a control center and immediately assigned to one of the company's batch plants.
+The focus of this paper is the distribution of RMC from a Canadian company that operates in the greater Montreal area.
+When a customer places an order, it is received at a central center and assigned to one of the company's batching plants.
 These plants produce the concrete and then deliver it to the customer.
-The problem we are examining involves a set of customer orders, a set of concrete-mixer drivers, and a set of batch plants.
+The problem we are examining involves a set of customer orders, a set of concrete-mixer drivers, and a set of batching plants.
 
@@
 
-A customer $i$ requests one or more types of concrete to be delivered to their construction site on a specific day, with the delivery service starting at the due time $a_i$.
+A customer $i$ requests one or more types of concrete to be delivered to his construction site on a particular day, starting at time $a_i$.
 We call an order $o$ a request for a specific type of concrete.
-$q_i$ is the sum of the demands $q_o$ of each order $o$ placed, $a_i$ is the desired arrival time of the first concrete mixer, and $\tau^u_i$ is the unloading rate for customer $i$.
+$q_i$ is the sum of the demand $q_o$ of each order $o$ placed, $a_i$ is the desired arrival time of the first concrete mixer, and $\tau^u_i$ is the unloading rate for customer $i$.
 If the order requires more concrete than a single truck can carry, multiple deliveries are scheduled.
@@
 Each element of $O_i$ must be fully delivered before moving on to another order.
-Exactly one order $o \in O_i$ must have its first delivery start at $a_i$, while the others can start at most $\gamma^2$ time after $o$ is completed.
-A plant is assigned to an order, and it must be the supplier of all subsequent deliveries of that order.
-To avoid cold joint problems with the concrete, subsequent deliveries of the same order must be made in close succession.
-We define a maximum time delay $\gamma^1$ after which no more deliveries are allowed.
-The customer's unloading rate and the quantity to be unloaded give the time required to unload a truckload.
+Exactly one order $o$ of $O_i$ must have its first delivery start at $a_i$, while the others can start at any time after $o$ is completed.
+Once a plant is selected to produce a customer's first delivery, it must be the supplier of all subsequent deliveries from that customer.
+To avoid cold joint problems with the concrete, the subsequent deliveries must be made in close succession.
+We define a maximum time delay $\gamma_i$ after which no more deliveries will be allowed.
+The customer unloading rate and the quantity to be unloaded give the time required to unload a truckload.
 Let $\mathcal{C}$ be the set of construction sites (customers) with a planned delivery for the day, and $\mathcal{O}=\{{O_i, i \in \mathcal{C}}\}$ be the set of all requested orders for all customers.
@@
 A scheduled driver $k$ is required to start his shift at $H_k$, work a minimum of $M_T$ hours and a maximum of $N_T$ hours during regular working hours, with the possibility of overtime of up to $O_T$ hours.
-$\beta_3$ and $\beta_4$ are the penalties for a driver working less than $M_T$ and more than $N_T$.
+$\beta_3$ and $\beta_4$ are the penalties incurred if a driver works less than $M_t$ and more than $N_t$.
 
@@
 This means that a truck can only haul RMC for one order, even if there is spare capacity.
-After unloading the RMC,  driver $k$ takes $Ct_k$ minutes to clean the concrete mixer before proceeding.
+To fill multiple orders, a driver must restock at a plant between deliveries.
+After unloading the RMC, a driver takes $\beta_k$ minutes to clean the concrete mixer before proceeding.
 %When assigning drivers to deliveries, the company prioritizes employees with the highest seniority.
@@
 However, we can compute its lower ($n_o^{min}$) and upper ($n_o^{max}$) bounds using the capacities of the largest ($Q_{max}$) and smallest ($Q_{min}$) available trucks.
+The lower bound is the number of deliveries required if we only use trucks with capacity $Q_{max}$, while the upper bound is the number of deliveries required if we only use trucks with capacity $Q_{min}$.
 \begin{alignat}{3}
@@
 
-Let $d^j_{o}$ be the $j^{th}$ visit with load $q^j_{o}$ for order $o$.
+Let $d^j_{o}$ be the $j_{th}$ visit with load $q^j_{o}$ for order $o$.
 We represent the fulfillment of order $o$ by the visits to the ordered set of delivery nodes $\mathcal{D}_o= \left(d^0_{o},d^1_{o},\cdots, d^{n_o}_{o}\right)$.
-The deliveries of customer $i$ are the ordered set $\mathcal{D}_i= (\mathcal{D}_{o_1}, \mathcal{D}_{o_2},\cdots,\mathcal{D}_{o_{|O_i|}})$, where $o_r$ is the $r^{th}$ delivered order.
-We will refer to $d \in \mathcal{D}_i$ ($d \in \mathcal{D}_o$) as the $d^{th}$ potential delivery of customer $i$ (order $o$).
+The deliveries of customer $i$ are the ordered set $\mathcal{D}_i= (\mathcal{D}_{o_1}, \mathcal{D}_{o_2},\cdots,\mathcal{D}_{o_{|O_i|}})$, where $o_r$ is the $rth$ delivered order.
+We will refer to $d \in \mathcal{D}_i$ ($d \in \mathcal{D}_o$) as the $dth$ potential delivery of customer $i$ (order $o$).
 $\mathcal{D}=\bigcup_{i\in \mathcal{C}} \mathcal{D}_i$ is the union of all delivery nodes.
@@
 
-Each of the company's plants has a single loading dock that can accommodate only one truck at a time.
-As a result, trucks often form a queue while waiting for their turn at the loading dock.
-Let $\mathcal{B}$ be the set of batch plants.
+The company owns a total of eight batching plants located in various geographical regions.
+Each plant has a loading dock that can only accommodate only one truck at a time, which leads to trucks lining up.
+Let $\mathcal{B}$ be the set of batching plants.
 The plants are heterogeneous, as each plant $b$ has its own hourly loading rate, represented by $\tau^l_b$, which affects the duration of the loading process.
 After loading the concrete, the driver spends $\alpha_b$ minutes adjusting the concrete in the truck before heading to the customer site.
-Each plant has its own assigned fleet of trucks, but it can borrow trucks from other plants.
-Let $l_{j}$ be the loading dock node associated with delivery node $j$.
-After loading RMC at $l_j$, it must be fully delivered to $j$ at most before $\Delta$ time, which is the concrete lifetime.
-We denote $\mathcal{L}_b$ as the set of loading dock nodes of plant $b$ and $\mathcal{L}$ as the set of all loading docks.
+Each plant has its own assigned fleet of trucks, but it can borrow trucks from other plants or hire external vehicles if necessary.
+A batching plant can serve a construction site as long as the travel time between the two is less than the concrete's lifespan, which is represented by $\Delta$.
+Let $l_{b,j}$ be the loading dock node associated with delivery node $j$ at plant $b$.
+We represent each plant $b$ by the set $\mathcal{L}_b = \{l_{b,j}, t_{bj}\leq  \Delta, j \in \mathcal{D} \}$  of loading docks nodes.
+$\mathcal{L}=\bigcup_{b\in B} \mathcal{L}_b$ is the union of all loading docks.
 
 % \subsection*{Model}
-A solution to the problem involves decisions about truck loading schedules, driver assignments to different deliveries, and truck arrival times at construction sites for unloading.
-For a batch plant, the decision involves choosing which driver to load, when to load them, how much to load, and which construction site to deliver.
-For a driver, the decision is to determine the sequence of loading depots and delivery sites.
+A solution to the problem involves making decisions about truck loading schedules, driver assignments to different deliveries, and truck arrival times at construction sites for unloading.
+For a batching plant, the decision involves choosing which driver to load, when to load them, and which construction site they should deliver to.
+For a driver, the decision involves determining the sequence of loading depots and delivery sites.
 And for a construction site, the decision involves determining the arrival times of all scheduled deliveries for the day.
 
-Each driver leaves and returns to their home plant every day.
+Each driver leaves and returns to his home plant every day.
 We represent the home plant of a driver $k$ with a starting depot $s_k$ and an ending depot $e_k$.
@@
 $A$ corresponds to allowed movements of drivers from node $i$ to node $j$. For each driver $k$, the allowed movements are the following:
-
 \begin{itemize}
@@
 For a customer $c$, arcs in $A^D$ link consecutive delivery nodes of the same order   $\lbrace (i,j)\in \mathcal{D}_o, o \in \mathcal{O}_c, i < j  \rbrace$, and pair of delivery nodes of two different orders $ \lbrace (i,d^{0}_{o_2}),  i \in \mathcal{D}_{o_1}, i \geq n^{min}_{o_1}, o_1, o_2 \in \mathcal{O}_c, o_1 \neq o_2 \rbrace $.
-Arcs in $A^L$ link all pairs of loading docks of the same batch plant.
-
-We define $\delta^{+}(i) = \{(i, j,k) \in A \}$ and $\delta^{-}(i) = \{(j, i,k) \in A \}$ as the outcoming and incoming arc sets of node $i \in V$; $\delta^{+}_D(i) = \{(i, j)  \in A^D\}$ and $\delta^{-}_D(i) = \{(j, i) \in A^D \}$ are the outcoming and incoming arc sets of delivery node $i \in \mathcal{D}$.
-Similarly, $\delta^{+}_L(i) = \{(i, j)  \in A^L\}$ and $\delta^{-}_L(i) = \{(j, i) \in A^L \}$ are the outcoming and incoming arc sets of loading node $i \in \mathcal{L}$.
-
-Let the binary variable $x^{k}_{ij}$ be $1$ if driver $k$ travels from node $i$ to $j$.
-Binary variable $y_o$ is $1$ when order $o$ is completely served; $v_i$ and $w_i$ are the start and end of the loading (unloading) operation at node $i \in \mathcal{L}\cup \mathcal{D}$.
-Binary variable $u_{ij}$ is $1$ if node $j$ is served just after $i$, the service being either an unloading or a loading operation.
-% Binary variable $\sigma_{ib}$ is equal to 1 if the orders of customer $i$ are loaded by plant $b$.
+Arcs in $A^L$ link all pairs of loading docks of the same batching plant.
+
+$\delta^{+}(i) = \{(i, j,k) \in A \}$, and $\delta^{-}(i) = \{(j, i,k) \in A \}$ are the outcoming and incoming arc sets of any node $i \in V$.
+$\delta^{+}_D(i) = \{(i, j)  \in A^D\}$, and $\delta^{-}_D(i) = \{(j, i) \in A^D \}$ are the outcoming and incoming arc sets of delivery node $i \in \mathcal{D}$.
+Similarly, $\delta^{+}_L(i) = \{(i, j)  \in A^L\}$, and $\delta^{-}_L(i) = \{(j, i) \in A^L \}$ are the outcoming and incoming arc sets of loading node $i \in \mathcal{D}$.
+
+Let the binary variable $x^{k}_{ij}$ be $1$ if the driver $k$ travels from node $i$ to $j$.
+The binary variable $y_o$ is $1$ when the order $o$ is completely served.
+$v_i$ and $w_i$ are the start and end of the loading (unloading) operation at node $i \in \mathcal{L} \cup \mathcal{D}$.
+The binary variable $u_{i,j}$ is $1$ if node $j$ is served just after $i$, the service being either an unloading or a loading operation.
+The binary variable $\sigma_{ib}$ is equal to 1 if the orders of customer $i$ are loaded by plant $b$.
 Variable $q^k_j$ is the quantity to be loaded towards $j$ with vehicle $k$.
-Let $w^1_k$ be a continuous variable indicating the difference between the driver's work time and the minimum number of hours to be worked in a day, and $w^2_k$ indicating the difference between the driver's work time and the normal work time.
+Let $w^1_k$ be a continuous variable indicating the difference between the driver's work time and the minimum number of hours to be worked in a day.
+$w^2_k$ is a continuous variable indicating the difference between the driver's work time and the normal work time.
 Let $g_i$ be the time between due date and first service start for customer $i$.
@@
 \begin{alignat}{3}
-    \label{mod:obj}   \min & \smashoperator{ \sum_{ \substack{(i,j,k) \in A} }} {t_{ij}x^{k}_{ij}} + \beta_1\smashoperator{\sum_{ o \in \mathcal{O}}}{  \left(1-y_o\right)} \hspace{-0mm}  + \beta_2 \smashoperator{ \sum_{i \in \mathcal{C} } } { g_i  } +  \sum_{k \in K}{\beta_3 w^{1}_k } + \beta_4 { w^{2}_k } &                                                                                                  \\
-    \label{mod:c1}         & \sum_{j \in \delta^{+}(i) }{x^k_{ij}} =1                                                                                                                                                                                                                                               & \forall i \in S, k \in K                                                                         \\
-    \label{mod:c2}         & \sum_{j \in \delta^{-}(i)}{x^k_{ji}} = 1                                                                                                                                                                                                                                               & \forall i \in E, k \in K                                                                         \\
-    \label{mod:c03}        & v_j \geq  w_i  - M\left(1-x^{k}_{ij}\right)                                                                                                                                                                                                                                            & \hspace{-20mm}  i \in S, j \in \delta^{+}(i),  k \in K                                           \\
-    \label{mod:c3}         & v_j \geq  w_i + \alpha_{b} + t_{ij} - M\left(1-x^{k}_{ij}\right)                                                                                                                                                                                                                       & \hspace{-20mm} \forall b \in \mathcal{B}, i \in \mathcal{L}_b, j \in \delta^{+}(i),  k \in K     \\
-    \label{mod:c4}         & v_j \geq  w_i + Ct_{k} + t_{ij} - M\left(1-x^{k}_{ij}\right)                                                                                                                                                                                                                           & \hspace{-20mm} \forall i \in \mathcal{D}  , j \in \delta^{+}(i), k \in K                         \\
+\label{mod:obj}   \min &  \smashoperator{ \sum_{ \substack{(i,j,k) \in A} }} {t_{ij}x^{k}_{ij}} + \beta_1\smashoperator{\sum_{ o \in \mathcal{O}}}{  \left(1-y_o\right)} \hspace{-0mm}  + \beta_2 \smashoperator{ \sum_{i \in \mathcal{C} } } { g_i  } +  \sum_{k \in K}{\beta_3* w^{1}_k } + \beta_4 *{ w^{2}_k }&\\
+    \label{mod:c1} &\sum_{j \in \delta^{+}(s_k) }{x^k_{s_kj}} = 1   &  \forall k \in K  \\
+    \label{mod:c2} &\sum_{j \in \delta^{-}(e_k)}{x^k_{je_k}} = 1   &  \forall k \in K\\
+    \label{mod:c3} & v_j \geq  w_i + \alpha_{i} + t_{ij} - M\left(1-x^{k}_{ij}\right)  & \hspace{-20mm} \forall i \in \mathcal{L}, j \in \delta^{+}(i),  k \in K \\
+    \label{mod:c4}  & v_j \geq  w_i + \beta_{k} + t_{ij} - M\left(1-x^{k}_{ij}\right) & \hspace{-20mm} \forall i \in \mathcal{D}, j \in \delta^{+}(i), k \in K \\
     \label{mod:c5} & w_{i} \geq v_{i}  + \frac{q^k_j}{\tau^l_b} -M\left(1- x^{k}_{ij}\right) &  \hspace{-20mm} \forall  b \in \mathcal{B},  i \in \mathcal{L}_{b},  j \in \delta^{+}(i), k \in K \\
@@
     \label{mod:c7} &         w_{j} \leq v_{i}  + \Delta + M\left(1- x^{k}_{ij}\right) & \hspace{-20mm}   j \in \mathcal{D},  i \in \delta^{-}(j),  k \in K \\
-    \label{mod:c8}         & v_{d^0_{o}} \geq a_i                                                                                                                                                                                                                                                                   & \hspace{-20mm} \forall  i \in \mathcal{C}, \forall o \in \mathcal{O}_i                           \\
-    \label{mod:c9}         & g_{i} \geq v_{d^0_{o_1}} - a_i - M\left(  \sum_{ j \in \delta^{-}_{D}(d^0_{o_1})  }{u_{jd^0_{o_1}}}\right)                                                                                                                                                                             & \hspace{-20mm}  \forall  i \in \mathcal{C}, \forall o_1 \in \mathcal{O}_i                        \\
-    \label{mod:c101}       & v_{d^0_{o_1}} \geq w_j - M\left(1- u_{jd^0_{o_1}}\right)                                                                                                                                                                                                                               & \hspace{-40mm}    \forall o_1 \in \mathcal{O}_i,  j \in \delta_{D}^{-}(d^0_{o_1})                \\
-    \label{mod:c10}        & v_{d^0_{o_1}}  \leq  w_j + \gamma^2 + M\left(1- u_{j,d^0_{o_1}}\right)                                                                                                                                                                                                                 & \hspace{-20mm}  \forall  o_1 \in \mathcal{O},  j \in \delta_{D}^{-}(d^0_{o_1})                   \\
+    \label{mod:c8}  &  v_{d^0_{o}} \geq a_i  & \hspace{-20mm} \forall  i \in \mathcal{C}, \forall o \in \mathcal{O}_i\\
+    \label{mod:c9} & g_{i} \geq v_{d^0_{o_1}} - a_i - M\left(  \sum_{ j \in \delta^{-}_{D}(d^0_{o_1})  }{u_{j,d^0_{o_1}}}\right)  & \hspace{-20mm}  \forall  i \in \mathcal{C}, \forall o_1 \in \mathcal{O}_i, d=d^0_{o_1} \\
+     \label{mod:c101} & v_{d^0_{o_1}} \geq w_j - M\left(1- u_{j,d^0_{o_1}}\right)  & \hspace{-40mm}    \forall o_1 \in \mathcal{O}_i,  j \in \delta_{D}^{-}(d^0_{o_1})  \\
+  \label{mod:c10} &  v_{d^0_{o_1}}  \leq  w_j + \gamma_i + M\left(1- u_{j,d^0_{o_1}}\right) &  \hspace{-20mm}  \forall  o_1 \in \mathcal{O},  j \in \delta_{D}^{-}(d^0_{o_1}) \\
     \label{mod:c11} &  \sum_{ \substack{o_1 \in \mathcal{O}_i }}{ \sum_{ j \in \delta^{-}_{D}(d^0_{o_1})}}{u_{j,d^0_{o_1} } } = |\mathcal{O}_i|-1 & \hspace{-20mm} \forall  i \in \mathcal{C},|\mathcal{O}_i| > 1 \\
-    \label{mod:c12}        & \sum_{ \substack{o_1 \in \mathcal{O}_i }}\sum_{ j \in \delta^{+}_{D}(d^0_{o_1})}{u_{d^0_{o_1},j } } = |\mathcal{O}_i|-1                                                                                                                                                                & \hspace{-40mm} \forall  i \in \mathcal{C}, |\mathcal{O}_i| > 1                                   \\
+    \label{mod:c12}&   \sum_{ \substack{o_1 \in \mathcal{O}_i }}\sum_{ j \in \delta^{+}_{D}(d^0_{o_1})}{u_{d^0_{o_1},j } } = |\mathcal{O}_i|-1 & \hspace{-40mm} \forall  i \in \mathcal{C}, |\mathcal{O}_i| > 1 \\
     \label{mod:c13} & \hspace{3mm} \smashoperator{ \sum_{ j \in \delta^{+}_{D}(d^0_{o_1})}}{u_{d^0_{o_1},j } } \leq 1 & \hspace{-20mm}  \forall o_1 \in \mathcal{O} \\
@@
     \label{mod:c16} & v_{j} \geq w_{j-1} - M\left(1- u_{j-1,j}\right) & \hspace{-20mm}  \forall o \in \mathcal{O}, j \in \mathcal{D}_{o}, j \geq 1 \\
-    \label{mod:c17}        & v_{j}  \leq  w_{j-1} + \gamma^1 + M\left(1- u_{j-1,j}\right)                                                                                                                                                                                                                           & \hspace{-20mm} \forall i \in \mathcal{C}, \forall o \in O_i, j \in \mathcal{D}_{o}, j \geq 1     \\
-    \label{mod:c18}        & u_{j-1,j} \geq u_{j,j+1}                                                                                                                                                                                                                                                               & \hspace{-35mm} \forall o_1 \in\mathcal{O},  j \in \mathcal{D}_{o_1}, 1 \leq j \leq n_{o_{1}}-1   \\
-    \label{mod:c19}        & u_{j-1,j} \geq \sum_{l \in \mathcal{L}}{x_{lj}}                                                                                                                                                                                                                                        & \hspace{-20mm}  \forall o_1 \in \mathcal{O}, j \in \mathcal{D}_{o_1}, j \geq 1                   \\
+  \label{mod:c17} &  v_{j}  \leq  w_{j-1} + \gamma_i + M\left(1- u_{j-1,j}\right) & \hspace{-20mm} \forall i \in \mathcal{C}, \forall o \in O_i, j \in \mathcal{D}_{o}, j \geq 1 \\
+   \label{mod:c18}   &   u_{j-1,j} \geq u_{j,j+1} & \hspace{-35mm} \forall o_1 \in\mathcal{O},  j \in \mathcal{D}_{o_1}, 1 \leq j \leq n_{o_{1}}-1\\
+  \label{mod:c19} &  u_{j-1,j} \geq \sum_{l \in \mathcal{L}}{x_{lj}} &\hspace{-20mm}  \forall o_1 \in \mathcal{O}, j \in \mathcal{D}_{o_1}, j \geq 1 \\
     \label{mod:c20} & v_{j} \geq w_{i} - M\left(1- u_{i,j}\right) & \hspace{-20mm}  \forall i \in \mathcal{L}, j \in \delta^{+}_{L}(i) \\
@@
     \label{mod:c22} & \hspace{2mm} \smashoperator{ \sum_{ j \in \delta^{-}_{L}(i)}}{u_{j,i } } \leq 1 & \hspace{-20mm} \forall i \in \mathcal{L} \\
-    \label{mod:c230}       & \hspace{2mm} \smashoperator{ \sum_{ j \in \delta^{-}_{L}(i)}}{u_{j,i } }  \geq \sum_{k \in K}\sum_{j \in  \delta^{-}(i) }{x^k_{ji}}                                                                                                                                                    & \hspace{-20mm} \forall i \in \mathcal{L}                                                         \\
-    \label{mod:c231}       & \hspace{2mm} \smashoperator{ \sum_{ j \in \delta^{-}_{L}(i)}}{u_{j,i } } \geq \smashoperator{ \sum_{ j \in \delta^{-}_{L}(i+1)}}{u_{j,i+1 } }                                                                                                                                          & \hspace{-20mm} \forall i \in \mathcal{L}                                                         \\
     \label{mod:c23} & \hspace{2mm} \smashoperator{ \sum_{ j \in \delta^{-}_{L}(i)}}{u_{j,i } } + \smashoperator{ \sum_{ j \in \delta^{+}_{L}(i)}}{u_{i,j } } \geq \sum_{k \in K}\sum_{j \in  \delta^{-}(i) }{x^k_{ji}} & \hspace{-20mm} \forall i \in \mathcal{L} \\
     \label{mod:c24} & \sum_{k \in K}\sum_{j \in \mathcal{D}_o}{q^k_j} = q_o & \hspace{-20mm}  \forall o \in \mathcal{O} \\
-    \label{mod:c240}       & q^k_j \leq \sum_{i \in  \delta^{-}(j)}{Q^k x^k_{ij}}                                                                                                                                                                                                                                   & \hspace{-20mm}  \forall j \in \mathcal{D}, k \in K                                               \\
     \label{mod:c25} &  \sum_{k \in K}\sum_{j \in \delta^{+}(i)}{x^{k}_{ij} \leq 1} &  i \in \mathcal{L} \cup \mathcal{D} \\
   \label{mod:c26}  &  \sum_{k \in K}\sum_{j \in \delta^{+}(i)} {x^{k}_{ij} } =  \sum_{k \in K}\sum_{j \in \delta^{-}(i)}{x^{k}_{ji} } &  i \in \mathcal{L} \cup \mathcal{D} \\
-    % \label{mod:c27}        & \sum_{k \in K}  \sum_{i \in \mathcal{D}_c} \sum_{j \in \delta^{-}(i) }  {x^k_{ji}} \leq M\sigma_{c, b}     & \forall c \in \mathcal{C}, b \in \mathcal{B}                                                     \\
-    % \label{mod:c28}        & \sum_{b \in \mathcal{B}} \sigma_{c, b} \leq 1    & \hspace{-20mm}\forall c \in \mathcal{C}, b \in \mathcal{B}                                       \\
+  \label{mod:c27} & \sum_{k \in K}  \sum_{i \in \mathcal{D}_c} \sum_{j \in \delta^{-}(i) }  {x^k_{ji}} \leq M*\sigma_{c, b} & \forall c \in \mathcal{C}, b \in \mathcal{B} \\
+    \label{mod:c28} & \sum_{b \in \mathcal{B}} \sigma_{c, b} \leq 1 & \hspace{-20mm}\forall c \in \mathcal{C}, b \in \mathcal{B} \\
     \label{mod:c29} & w^1_k \geq  M_T + H_k - v_{e_k}  & \hspace{-20mm} \forall k \in K \\
@@
     \label{mod:c34} &  u_{ij} \in \{0,1\} & \hspace{-20mm}  (i,j) \in A^D , k \in K  \hspace{0mm} \\
-    % \label{mod:c35}        & \sigma_{ib} \in \{0,1\}    & \hspace{-20mm}  i \in \mathcal{C}, b \in \mathcal{B}                                             \\
+    \label{mod:c35} &  \sigma_{ib} \in \{0,1\} & \hspace{-20mm}  i \in \mathcal{C}, b \in \mathcal{B} \\
     \label{mod:c36} &  y_o \in \{0,1\} & \hspace{-20mm}  o \in \mathcal{O} \\
-    \label{mod:c37}        & v_i \geq 0,  w_i \geq 0                                                                                                                                                                                                                                                                & \hspace{-20mm}  i \in V                                                                          \\
+    \label{mod:c37} &  v_i \geq 0,  w_i \geq 0 & \hspace{-20mm}  i \in V \\\
     \label{mod:c38} &  w^1_{k} \geq 0,  w^2_k \geq 0  & \hspace{-20mm}  i \in V \\
@@
 $\beta_1$ to $\beta_4$ are the penalty coefficients of each component of the objective function (\ref{mod:obj}).
-Constraints (\ref{mod:c1}) and (\ref{mod:c2}) state that driver $k$ leaves his start node exactly once a day and returns to his end node.
-A driver cannot serve a node before the start of his shift with constraints (\ref{mod:c03}).
-Constraints (\ref{mod:c3}) and (\ref{mod:c4}) set a driver to take some time after loading or unloading to adjust the concrete or clean the truck before moving on to the next node.
-The duration of the loading operation depends on the plant's loading rate and the amount $q^k_j$ of RMC loaded (\ref{mod:c5}).
-Similarly, the unloading service depends on the site's rate and $q^k_j$ (\ref{mod:c6}).
-Unloading operations must end at most $\Delta$ minutes after loading begins (\ref{mod:c7}).
-
-Constraints (\ref{mod:c8}) - - (\ref{mod:c10}) ensure that the first service of any customer $i$ must start at the due time $a_i$.
+Constraints (\ref{mod:c1}--\ref{mod:c2}) state that a driver $k$ leaves his start node exactly once a day and returns to his end node after completing his last unloading operation of the day.
+Constraints (\ref{mod:c3}--\ref{mod:c4}) set a driver to take some time after loading or unloading to adjust the concrete or clean the truck before moving on to the next node.
+The duration of the loading operation depends on the plant's loading rate and the amount $q^k_j$ of RMC loaded (Eq.
+\ref{mod:c5}).
+Similarly, the unloading service depends on the site's rate and $q^k_j$ (Eq.
+\ref{mod:c6}).
+Unloading operations must end $\Delta$ minutes after loading begins (Eq.
+\ref{mod:c7}).
+
+Constraints (\ref{mod:c8}--\ref{mod:c10}) ensure that the first service of any customer $i$ must start at the due time $a_i$.
 This first service may be performed at the first delivery node of any order $\in O_i$, and one order must be completed before another is started.
 It also enforces the precedence constraints between the last delivery of an order and the first delivery of the following order.
-Constraints (\ref{mod:c11}) - - (\ref{mod:c15}) find the delivery sequence of all orders $\in \mathcal{O}_i$.
-With constraints (\ref{mod:c16}) - - (\ref{mod:c19}), two trucks cannot unload at the same time for consecutive deliveries of the same order.
-Additionally, these constraints impose a maximum time delay between the two trucks.
-Similarly, constraints (\ref{mod:c20}) - - (\ref{mod:c23}) ensure that two trucks cannot be loaded at the same time at a plant.
-Constraints (\ref{mod:c24}) require that the cumulative load of all concrete mixers serving an order must equal the required quantities, and (\ref{mod:c240}) bound a driver load.
+Constraints (\ref{mod:c11}--\ref{mod:c15}) find the delivery sequence of all orders $\in \mathcal{O}_i$.
+Constraints (\ref{mod:c16}--\ref{mod:c19}) prohibit two trucks serving consecutive deliveries of the same order from unloading at the same time and enforce a maximum time delay between them.
+Similarly, constraints (\ref{mod:c20}--\ref{mod:c23}) ensure that two trucks cannot be loaded at the same time at a plant.
+Constraints (\ref{mod:c24}) require that the cumulative load of all concrete mixers serving an order must equal the required quantities.
 Constraints (\ref{mod:c25}) require that a driver can only visit a loading/delivery node once.
 Constraints (\ref{mod:c26}) are degree constraints.
-Constraints (\ref{mod:c29}) - - (\ref{mod:c30}) calculate the difference between a driver's hours of service and the minimum and normal hours of service.
-Finally, constraints (\ref{mod:c31}) - - (\ref{mod:cend}) define the nature and bounds of the variables.
+Constraints (\ref{mod:c27}) and (\ref{mod:c28}) require that all orders from customer $i$ must be loaded from the same plant.
+Constraints (\ref{mod:c29}--\ref{mod:c30}) calculate the difference between a driver's hours of service and the minimum and normal hours of service.
+Finally, constraints (\ref{mod:c31})--(\ref{mod:cend}) define the nature and bounds of the variables.
 
@@
 Let us consider the solution of a small instance of our problem shown in Figures~\ref{fig_Example} and \ref{fig:ganttExample}.
-Two batch plants $B_1$ and $B_2$ with three drivers serve two construction sites.
+Two batching plants $B_1$ and $B_2$ with three drivers serve two construction sites.
 Customer $C_1$ requests one order of $q^1_1=12$ m$^3$.
@@
 We schedule $D_1$ and $D_2$ to deliver 8 and 4 m$^3$ to $C_1$, respectively.
-We select $o^2_2$ as the first order of $C_2$.
+We select $o^2_2$ as the first order of $C_2$ to deliver.
 $D_3$ travels to $B_1$ to load $q^2_2$, then visits $C_2$ before returning to his home plant.
@@
 
-\begin{figure}[htbp]
+\begin{figure}[!ht]
     % \caption{Solution of an instance with two plants, two construction sites, four orders, and three drivers.}
@@
 
-\begin{figure}[htbp]
+\begin{figure}[!htb]
     \centering
@@
 
-In this section, we present the heuristic approach we implement to solve this variant of the Concrete Delivery Problem.
-Our method consists of constructing feasible solutions to the CDP with randomized greedy heuristics and iteratively invoking these heuristics in the GRASP metaheuristic.
-From now on, we represent the scheduling of a delivery node $d$ as the pair  ($L^k_{d}$,$U^k_{d}$) of loading  and unloading  tasks.
-$L^k_{d} = \left(b,q_d,v_b,w_b\right)$ indicates that driver $k$ starts loading $q_d$ $m^3$ of RMC at plant $b$ between $\left[v_b,w_b\right]$.
-$U^k_{d} = \left(k,v_d, w_d\right)$ indicates that driver $k$ serves $d$ between $v_d$ and $w_d$.
-We then call a solution $ S =\left\lbrace \cup _{1 \leq j \leq |\mathcal{D}|} (L^k_{j}, U^k_{j}), k \in K \right\rbrace$ the set of all loading and unloading tasks performed on a given day.
+We now present the heuristic approach we implement to solve this variant of the Concrete Delivery Problem.
+Our method consists of constructing feasible solutions to the CDP with randomized heuristics and iteratively invoking these heuristics in the GRASP metaheuristic.
+
+In this section, we represent the scheduling of a delivery node $d$ as a pair of loading ($L_{d}$) and unloading ($U_{d}$) tasks.
+$L_{d} = \left(b,k,q_d,v_b\right) $ indicates that driver $k$ starts loading $q_d$ $m^3$ of RMC at plant $b$ at time $v_b$.
+$U_{d} = \left(k,v_d, w_d\right)$ indicates that driver $k$ is serving $d$ between $v_d$ and $w_d$.
+We then call a solution $ S =\left\lbrace \cup _{1 \leq j \leq n_o} (L_{d^j_{o}}, U_{d^j_{o}}), \forall o \in \mathcal{O} \right\rbrace$ the set of all loading and unloading tasks performed on a given day.
 
@@
 
-The GRASP algorithm, proposed by \cite{feo1989probabilistic}, is a two-phase iterative suite consisting of constructive and local search algorithms.
-In the construction phase, a feasible solution $S$ is generated using a greedy randomized algorithm.
-Then, a local search algorithm explores the neighborhood of $S$ to identify a local optimum.
-Algorithm~\ref{alg_grasp} outlines the pseudocode of the GRASP algorithm.
-To avoid redundant computations, a hash table $H$ is used to ensure that the local search is not applied multiple times to the same solution.
-The procedure stops and returns the best overall solution when certain stopping conditions, such as reaching a time limit or a maximum number of iterations, are met.
-In our implementation, we also add a path relinking procedure that uses information about previous solutions to improve the search process.
+The GRASP algorithm is an iterative suite of constructive and local search algorithms introduced by \cite{feo1995greedy}.
+A feasible solution $S$ is constructed using a greedy randomized algorithm at each iteration.
+Then a local search algorithm explores the neighborhood of $S$ to find a local optimum.
+The pseudocode of the algorithm~\ref{alg_grasp} shows that the procedure returns the best overall solution (for a minimization problem) after some stopping conditions (time limit or a maximum number of iterations) are met.
+In our implementation, we introduce at line 13 an additional local search step using the solutions found by all heuristics in the current iteration.
 
@@
             \setcounter{AlgoLine}{0}
-                % \KwIn{ \textit{H}: List of constructive greedy randomized heuristics }
-                \KwIn{ $\phi$: number of iterations before path relinking }
+            \KwIn{ \textit{H}: List of constructive randomized heuristics }
             $S^* \leftarrow \emptyset$    \hspace{2mm} $T \leftarrow \emptyset$       $Cost(S^*) \leftarrow \infty$
 
-                Initialize an empty hash table $H$
-
-                $counter \leftarrow  0$
-
             \While{Conditions not met}{
-
-                    $S \leftarrow GreedyRandomizedHeuristic()$
-
-                    \uIf{ $S \notin H$}{
+                $Cost(S') \leftarrow \infty$
+
+                $T \leftarrow \emptyset$
+
+                \ForEach { $h \in H$ }{
+
+                    $S \leftarrow h(S)$
+
                     $S \leftarrow LocalSearch(S)$
 
-                        Add $S$ to $H$
-                    }
-
                     $T \leftarrow T \cup \{S\}$
 
-                    \uIf{$Cost(S)< Cost(S^*)$}{
-                        $S^* \leftarrow S$
-                    }
-
-                    \uIf{$counter \text{ } mod \text{ } \phi == 0$}{
-                        $S^* \leftarrow PathRelinking(T)$
-
-                        $T \leftarrow \emptyset$
-                    }
-
-                    $counter \leftarrow counter + 1$
-
+                    \uIf{$Cost(S)< Cost(S')$}{
+                        $S' \leftarrow S$
+
+                        \uIf{$Cost(S')< Cost(S^*)$}{
+                            $S^* \leftarrow S'$
+                        }
+                    }
+
+                }
+                $S^* \leftarrow LocalSearch(T)$
             }
@@
 
-For the construction phase, we have developed two greedy randomized heuristics.
-The first heuristic, called customer-based insertion, schedules customers one at a time.
-On the other hand, the second heuristic, called delivery-based insertion, does not necessarily complete the scheduling of one customer before starting to plan another customer's schedule.
-% The greedy aspect of GRASP is the creation of the set $RCL$, the probabilistic aspect is the random selection in $RCL$, and the adaptive aspect is the updating of $CL$ and the re-evaluation of the incremental costs.
-
-\subsection{Customer-based insertion heuristic}
-
-The customer-based insertion heuristic builds a solution from scratch by scheduling all of a customer's orders one at a time.
-As shown in Algorithm~\ref{alg:greedyCustIns}, for each customer, we randomly select the first order to schedule.
-To schedule an order, we iterate over each of its delivery nodes, find a feasible pair of task, and add it to the current solution.
-For a delivery node $d$, we check if the list of operations $M[d]$ is empty.
-If it is, for each available driver $k$ we simulate the loading and unloading operations $(L^k_d,U^k_d)$ with the procedure $SimulateVisit$, and if $k$ can serve $d$, we add $k$ to a candidate list $CL$ and $(L^k_d,U^k_d)$ to $M[d]$.
-
-The $SimulateVisit$ procedure takes a partial solution, a delivery node $d$, a driver $k$ and a numeric value $\lambda$ as parameters.
-Its goal is to find the earliest time when an unloading service can start at $d$ and the best position within the route of $k$.
-When planning to visit node $d$ of order $o$ and customer $i$ with driver $k$, we first compute his expected delivery time $v_d$ as follows:
-
-$ v_d = \left\{
-    \begin{array}{rl}
-        a_i + \lambda                                                   & \text{for the first visited node} \\
-        w_j + rand(\lambda, \gamma^1), j \in \mathcal{D}_o,  u_{jd}=1   &                                   \\
-        w_j + rand(\lambda, \gamma^2), j \notin \mathcal{D}_o, u_{jd}=1 &
-    \end{array}
-    \right.$
-
-Then we determine the start loading time $v_l$ of the corresponding dock node $l$ by subtracting the travel time, adjustment time, and loading duration from $v_d$.
-Starting from $v_l$ we check when $k$ will be available at an idle loading bay and adjust $w_l$ accordingly.
-Changing $w_l$ can cause the cold-joint constraints to be violated, especially for a customer's subsequent deliveries, and thus make the schedule infeasible.
-Once we have $v_l$, we can check for the best position to insert the visit of $d$ into the route of $k$.
-
-We extract from $CL$ the restricted candidate list ($RCL$) of drivers for whom the cost of visiting $d$ falls within a specified range based on the parameter $\theta$.
-If $RCL$ is empty and a visit to $d$ cannot be scheduled due to cold-joint constraints, we use the $DelayPreviousDelivery$ procedure to determine if there is a previous delivery of the same order that we can delay the start of service to make a visit to $d$ feasible.
-This procedure returns the index of the node and $\lambda$, which is the value of the delay.
-$\lambda$ is also the same parameter given to $SimulateVisit$.
-If there is such a previous node, we backtrack to it and restart the scheduling.
-
-If $RCL$ is not empty, we select a random driver from it, insert its corresponding load and unload tasks into $S$, and update the remaining quantities for the order and customer demands.
-The algorithm returns at the end the constructed solution $S$.
+\subsection{Greedy randomized insertion algorithm}
+
+As mentioned in \cite{resende2019greedy}, the general outline of a greedy randomized algorithm used in the GRASP framework works as described in Algorithm~\ref{alg:greedyIns}.
+Let's consider the set $CL$ of candidate delivery nodes that have not yet been scheduled.
+We first initialize $CL$ with the first delivery node of a randomly selected order, for each customer.
+
+At each iteration, we create the set $CL'$ with elements of $CL$ from a given neighborhood.
+We then evaluate the increase in the cost function of incorporating each $d \in CL'$ into the incumbent solution $S$ and construct a restricted candidate list $RCL$ consisting of nodes whose incremental cost is less than a given threshold.
+From $RCL$, we randomly select the next delivery node to be incorporated into the incumbent solution and determine its loading and unloading schedule using the procedure described in Algorithm~\ref{alg_ScheduleTask}.
+Next, we update $S$ and the remaining order and customer quantities.
+Then we add the next candidate delivery node to $CL$.
+If an order $o$ is not yet fulfilled after visiting $d^j_{o}$, the candidate is the next delivery node $d^{j+1}_{o}$.
+Otherwise, if the customer has remaining orders, the candidate is the first delivery node $d^{0}_{o'}$ of another randomly selected order $o'$.
+Note that for our problem, not all unscheduled delivery nodes are candidates at each iteration, since deliveries of the same order are ordered, any order of a customer can be chosen and must be satisfied before switching to another order, and the number of visits to a customer is not known in advance.
+
+Consider two customers whose services are required to start at the same time or in the same time slot.
+We could either finish scheduling one before the other, or schedule both at the same time.
+Obviously, each of these choices would result in a different solution.
+Similarly, in some cases, it would be better to schedule a customer that has a high demand and therefore requires more resources first.
+That's why we added the filtering part to impose some criteria on the algorithm.
+We can limit $CL'$ to delivery nodes with the same demand, the same due date, or from the same customer.
+We can also decide if we want to schedule customers with the highest (or lowest) demand or the earliest (latest) due time first.
+In fact, this helps us to intensify the search in specific neighborhoods of the solution.
+To diversify the search, we simply remove the filter component.
+The greedy aspect of GRASP is the creation of the set $RCL$, the probabilistic aspect is the random selection in $RCL$, and the adaptive aspect is the updating of $CL$ and the re-evaluation of the incremental costs.
+
+%Also, using $CL'$ helps reduce the complexity of the algorithm by reducing the number of nodes to evaluate.  %We thus obtain four greedy randomized insertion heuristics for our GRASP framework.
 
@@
 %  {\small
-        \begin{algorithm}[]
-            \caption{Customer-Based insertion algorithm }
-            \label{alg:greedyCustIns}
-            \DontPrintSemicolon
-            \LinesNumbered
-            \setcounter{AlgoLine}{0}
-            \KwIn{  $S$: empty solution $S$, $K$: vehicles set \hspace{15mm} \\$\Pi$: custom sorted customers list }
-
-            $M \leftarrow \emptyset $ \tcp{List of pair of loading and unloading tasks for all delivery nodes}
-            $CL \leftarrow \emptyset$ \tcp{Candidate driver list all delivery nodes}
-
-            \ForEach(customer $i \in \Pi$){}{
-
-            Shuffle $ \mathcal{O}_i$
-
-            \ForEach{order $o_1 \in \mathcal{O}_i$}{
-
-            $\lambda \leftarrow 0$
-
-            \For{ $j=0$ \KwTo $n^{max}_{o_1}-1$    }{
-
-            Delivery $d =\mathcal{D}^j_{o_1}$
-
-            \uIf{$M[d] == \emptyset $}{
-
-            \ForEach{ $k \in K$  }{
-
-            $(L^k_{d}$, $U^k_{d})$ $\leftarrow $ SimulateVisit($S$,$d$,$k$,$\lambda$)
-
-            $Cost(d,k)$ $\leftarrow $ Cost of ($L^k_{d}$, $U^k_{d}$)
-
-            Add $(L^k_{d}, U^k_{d}$) to $M[d]$
-
-            $CL[d] \leftarrow \{k\}$
-            }
-            }
-
-            $C_{min} \leftarrow min\left\lbrace C(x), \text{ } x \in M[d] \right\rbrace $, $C_{max} \leftarrow max\left\lbrace C(x), \text{ } x \in M[d] \right\rbrace $
-
-            $RCL \leftarrow \left\lbrace k \in CL[d], \text{ } C(d,k) \leq C_{min} + \theta (C_{max}-C_{min}) \right\rbrace $
-
-            \uIf{$RCL == \emptyset $}{
-
-                $(ind,delay)$ = $DelayPreviousDelivery$($S$,$d$)
-
-                \uIf{j is none}{
-
-                    break \tcp{continue with next customer}
-                }
-                $j \leftarrow ind$, $\lambda \leftarrow delay$
-
-                continue \tcp{backtrack to $ind$ position }
-
-            }
-            Select from $RCL$ a random driver $k$ to visit $d$ with load $q^k_d$
-
-            $M[d]\leftarrow M[d] \backslash (L^k_{d}, U^k_{d})$, $CL[d]\leftarrow CL[d] \backslash \{k\} $
-
-            $S \leftarrow S \cup (L^k_{d}$, $U^k_{d})  $
-
-            $q_i \leftarrow q_i-q^k_{d}$;  $q_{o_1} \leftarrow q_{o_1}-q^k_{d}$
-
-            $\lambda \leftarrow 0$
-
-            \uIf {$q_{o_1} \neq 0$ }{
-                continue
-
-            }
-            \uElseIf{ $q_i \neq 0$}{
-                break
-
-            }
-            }
-            }
-            }
-
-            \KwRet{$S$}
-        \end{algorithm}
-        %  }
-    }
-
-Consider the scenario where two customers require their services to start at the same time or within the same time window.
-In this case, we can either finish scheduling one customer before the other, or schedule both simultaneously.
-If there are sufficient resources available, either option would be feasible.
-However, when resources are limited, the customer-based insertion heuristic may leave some customers unscheduled.
-To address this limitation, we introduce the delivery-based insertion heuristic.
-
-\subsection{Delivery-based insertion heuristic}
-
-The delivery-based insertion heuristic constructs a feasible solution from scratch by scheduling a delivery node one at a time.
-The pseudocode for this heuristic is outlined in Algorithm~\ref{alg:delbasedIns}.
-It starts with an empty solution $S$ and initially fills $CL$ with the first delivery of a randomly selected order for each customer.
-Within each iteration of the $while$ loop, we filter $CL$ to obtain a reduced candidate list $CL'$ of deliveries from a given neighborhood.
-With this filtering, we can force the algorithm to schedule nodes with certain criteria first, which could be nodes from customers with the same demand, due date, or geographic area.
-We can also decide whether to schedule customers with the highest (or lowest) demand or the earliest (latest) due date first.
-
-For each delivery $d$ in $CL'$, the algorithm determines the best available driver $k$ and adds the corresponding loading and unloading task $(L^k_d, U^k_d)$ to a list of tasks.
-This is done using the $SimulateVisit$ procedure described earlier.
-Next, the algorithm constructs a $RCL$ from the list of tasks.
-If it is empty, the algorithm proceeds to the next iteration.
-Otherwise, it randomly selects a pair of tasks from the $RCL$, inserts them into $S$, and updates the remaining order and customer quantities.
-If an order $o$ is not fulfilled after visiting $d^j_{o}$, we add its next delivery node $d^{j+1}_{o}$ to $CL$.
-Otherwise, if the customer has remaining orders, we select the first delivery node $d^{0}_{o'}$ of another randomly selected order $o'$ and add it to $CL$.
-
-    % The greedy aspect of GRASP is the creation of the set $RCL$, the probabilistic aspect is the random selection in $RCL$, and the adaptive aspect is the updating of $CL$ and the re-evaluation of the incremental costs.
-
-    {\setstretch{1}
-        %  {\small
-        \begin{algorithm}[H]
-            \caption{Delivery-based insertion algorithm }
-            \label{alg:delbasedIns}
+     \begin{algorithm}[hbt]         
+         \caption{Greedy randomized insertion algorithm }
+         \label{alg:greedyIns}
          \DontPrintSemicolon
@@
          \ForEach(customer i){}{
-
                 Select a random order $o_1 \in \mathcal{O}_i$
@@
 
-            $Tasks \leftarrow \emptyset $ \tcp{List of loading and unloading tasks}
-
             \ForEach( $d \in CL'$){}{
 
-                % $C(d) \leftarrow$ incremental cost of inserting $d$
-
-                $Tasks \leftarrow (L^k_d,U^k_d)$ \tcp{$k$ is the best available driver to visit $d$}
-
-            }
-            $C_{min} \leftarrow \min\left\lbrace Cost(x), \text{ } x \in Tasks \right\rbrace $, $C_{max} \leftarrow \max\left\lbrace Cost(x), \text{ } x \in Tasks \right\rbrace $
-
-            $RCL \leftarrow \left\lbrace x \in Tasks, \text{ } Cost(x) \leq C_{min} + \theta (C_{max}-C_{min}) \right\rbrace $
-
-            \uIf{$RCL \leftarrow \emptyset$}{
-                continue
-            }
-
-            Select random operations $(L^k_{d^{j}_o},U^k_{d^{j}_o})$ of customer $i$ (order $o$) from $RCL$
-
-            $ S \leftarrow S \cup (L^k_{{d^{j}_o}}, U^k_{{d^{j}_o}})$
-
-            $q_i \leftarrow q_i-q^k_{d^{j}_o}$;  $q_{o} \leftarrow q_{o}-q^k_{d^{j}_o}$
+            $C(d) \leftarrow$ incremental cost of inserting $d$
+
+            }
+            $C_{min} \leftarrow min\left\lbrace C(d), \text{ } d \in CL' \right\rbrace $, $C_{max} \leftarrow max\left\lbrace C(d), \text{ } d \in CL' \right\rbrace $
+
+            $RCL \leftarrow \left\lbrace d \in CL', \text{ } C(d) \leq C_{min} + \alpha (C_{max}-C_{min}) \right\rbrace $
+
+            Select random delivery node $d^j_o$ of customer $i$ from $RCL$
+
+            $L_{d^j_o}$, $U_{d^j_o}$ = ScheduleTasks($d^j_o,S$)
+
+            $S \leftarrow S \cup (L_{d^j_o}$, $U_{d^j_o})  $
+
+            $q_i \leftarrow q_i-q_{d^j_o}$;  $q_{o} \leftarrow q_{o}-q_{d^j_o}$
 
@@
 
-\subsection{Local Search}
-
-After the construction phase, the local search phase iteratively invokes five operators using a first-improvement strategy.
-The first operator (\textit{Swap Load}) exchanges loads between two deliveries of the same order.
-The \textit{Swap Driver} operator explores alternative driver assignments by swapping drivers assigned to two delivery nodes.
-The \textit{Relocate} operator removes a delivery node and places it in a different location.
-The \textit{Consolidate Load} operator assigns an order request initially split between two drivers to a single driver, if the driver's capacity allows.
-The \textit{Remove and Reschedule} operator selects an unscheduled customer, denoted $i$, within the current solution.
-It then identifies a set $C$ containing scheduled customers with the same time slot as $i$.
-All elements of $C$ are removed and rescheduled along with $i$.
-This local search process continues until no further improvements can be made or a timeout is reached.
-
-\subsection{Path relinking}
-
-After the construction and local search steps, it is possible that some customers remain unscheduled in one solution while being accommodated in another.
-The goal of the path relinking procedure is to use each customer's scheduling information from existing solutions and combine them effectively to construct a solution that schedules all customers.
-This algorithm iterates through each element of a pool of solutions and attempts to insert each customer's schedule into a newly constructed solution.
-
-\section{Computational results}
+ Algorithm~\ref{alg_ScheduleTask} takes as parameters a delivery node $d^j_{o}$ of the customer $i$ and a partial solution and looks for a plant and a driver available for loading and unloading operations.
+For each driver $k$ and plant $b$ we simulate the loading and unloading operations while checking the concrete lifespan and cold joint constraints, and the plant and driver availability.
+For each node, we first determine the start loading time $SLT$ by subtracting $t_{bi}$, $\alpha_b$, and the loading duration $LD^b_{d^j_{o}}$ to the expected delivery time $EDT$.
+$EDT$ is either the arrival time $a_i$ for the first delivery of a customer or the end of the unloading of the precedent delivery.
+At line $15$, we ensure with the procedure \textit{ FindDriverLoadingSlot} that the loading starts only if the driver is present at the depot.
+The presence of the driver does not guarantee that the loading dock is available, thus we use the procedure \textit{FindDepotLoadingSlot} at line $16$ to ensure that.
+Once we know $SLT$, we can easily deduce the arrival time ($AT$), start unloading time ($SUT$), driver waiting time ($W^k_{d^j_o}$), and client waiting time ($W_{d^j_o}$).
+We also keep track of the driver's work duration.
+The algorithm returns the best loading and unloading operations with the least cost.
+ We keep track of all timeslots for a plant when its loading dock is busy, and for a driver when he starts loading until the end of the unloading service.
+Within \textit{FindDepotLoadingSlot} (\textit{FindDriverLoadingSlot}), we iterate through the timeslots of a depot (driver) to schedule the current operation at a free timeslot.
+These allow us to insert the scheduling of a node between existing schedules.
+
+{\setstretch{1}
+    {\small
+        % \vspace{-4mm}
+        \begin{algorithm}[H]
+            \caption{Schedule loading and unloading tasks }
+            \label{alg_ScheduleTask}
+            % \DontPrintSemicolon
+            \LinesNumbered
+            \setcounter{AlgoLine}{0}
+            \KwIn{Partial solution $S$, delivery node $d^j_{o}$}
+
+            \SetKwFunction{FMain}{ScheduleTask}
+            \SetKwProg{Fn}{Function}{}{}
+            \Fn{\FMain{$d^j_{o},S$}}
+            {
+            % \SetNlSkip{2em}
+            % \SetAlgoNlRelativeSize{-1}
+            $n_i$: last node visited for the customer $i$
+
+            $\mathcal{LS}:$ Partial solutions list
+
+            $bestSol$ Cost($bestSol)= +\infty$
+
+            \ForEach(driver $k$){}
+            {
+                $n_k$: current location of $k$ (plant or delivery node)
+
+                \ForEach(plant $b$){}
+                {
+
+                    \uIf(continue){$t_{bi} >= \delta $}{      }
+
+                    $EDT=SUT=a_i$
+
+                \uIf {$n_i \neq null$ }{
+
+                    $SUT= rand( EUt_{n_i}-\gamma/3,
+                        EUt_{n_i})$
+
+                    $EDT= EUT_{n_i}$
+
+                }
+
+            $q^j_{o}=min\{Q_k,q_o \}$ // $q_o$ \textit{remaining demand of order} $o$
+
+            $SLT$ = $EDT - t_{bi} - \alpha_b - LD^b_{d^j_{o}} $
+
+            $SLT = FindDriverLoadingSlot(b, n_k, SLT, LD^b_{d^j_{o}})$
+
+            $SLT = FindDepotLoadingSlot(b, SLT, LD^b_{d^j_{o}})$
+
+            $AT =  SLT + LD^b_{d^j{o}} + \alpha_b + t_{bi}$
+
+            $SUT = max\{AT, EDT\}$
+
+            $W^k_{d^j_o}= max\{0,SUT - AT\}$
+
+            $W_{d^j_o}= max\{0,AT - EDT - \lambda_i\}$
+
+            $EUT = SUT + UD_{d^p_{ij}} $
+
+            $WT = WT_k + (EUT + \beta_k - SLT) + t_{n_kb} + t_{i,k}$
+
+            $TC = t_{n_k,b} + t_{b,i}$  %- t_{pos_k,k}$
+
+            \uIf{$Cost(bestSol) < Cost(S) + TC $}{
+                $bestSol = (j,b,k,SLT) \cup (j,b,k,SUT) $
+
+                $Cost(bestSol) = Cost(S) + TC$
+            }
+            }
+            }
+            \KwRet{$bestSol$}
+            }
+        \end{algorithm}
+    }
+}
+
+\subsection{ Local Search}
+
+    % We implement two local search operators for the CDP. The first one is the relocate move commonly used for vehicle routing problems.
+
+   % At the end of each iteration, we expect each customer to have a feasible schedule respecting all constraints.
+For each customer, we combine each scheduling in order to construct a new and possibly better solution.
+
+\section{ Experimental results}
 \label{comp_exp}
 
-In this section, we present and discuss the results of the computational experiments of the GRASP heuristic described in section~\ref{grasp_method}.
-The GRASP algorithm is coded in C++.
-We run our experiments on a dataset extracted from data provided by our industrial partner.
+% In this section, we present and discuss the results obtained in the computational experiments of the GRASP heuristic described in Section~\ref{grasp_method}. The GRASP algorithm was coded in C++.
+% and the branch-and-bound (B\&B) algorithm applied to the model described in Section~\ref{model}.
+The different algorithms were coded in C++ and run on a processor Intel $2.4$ GHz Gold $6148$ Skylake.
+For the exact method, we have used IBM Concert Technology and CPLEX $12.10$ as MIP solver.
+A single thread was used with a time limit of one hour.
+% No benchmark instances for the \dirp exist so we have generated the dataset for the experiments as follows.
 
 \subsection{Generation of instances}
-
-The dataset used in this section was obtained from our partner and contains records representing delivery operations from up to 8 production centers over a period of 50 days.
-We generated 50 instances, with each instance corresponding to a daily operation.
-The instances include informations such as each driver capacity, associated depot, and shift start time.
-For each customer, we have the due time, total demand, and the number of orders (type of concrete) received.
-For each order, we specified its demand and associated production plant.
-Finally, we include loading capacities for each production center.
-The name of an instance has the format $C\_k\_n\_o\_z$. $k$ is the number of available drivers, $n$ the number of customers, $o$ the total number of daily orders, and $z$ the number of plants.
-
-The dataset is subdivised in a small, medium and large sets, according to the number of drivers and total demands served. We summarized both sets in Table~\ref*{}
-or 
-The dataset is subdivised in a small, medium and large sets, according to the number of orders served. We summarized both sets in Table~\ref*{}
-
-% This dataset allows us to evaluate the effectiveness and performance of our GRASP algorithm in optimizing delivery operations in a real-world setting.
-
-\begin{itemize}
-    \item   8 production centers located in greater Montreal
-    \item   42 days of delivery operations from Fall 2020
-    \item   From 7 to 139 orders each day
-    \item   Up to 137 drivers
-    \item   2 types of trucks of capacity 8 and 12 m$^3$
-\end{itemize}
+% Due to the lack of publicly available benchmark data for the CDP, two data sets are created.
+Both sets, as well as code to generate new instances, are available online at [17].
+A summary of the instances is provided in Table 4.DataSetAcontains instances with 10–20 customers,and2–5 vehicles,whereasData Set B contains instances with up to 50 customers and 20 vehicles.
+The demands of customers are always divisible by 5 and are selected at random from the interval [10–75].
+In a similar fashion, the capacities of the vehicles are selected between [10–25].
+It is unrealistic to have a unique capacity for each vehicle, and therefore the number of different vehicles is bounded by the number of vehicle classes.Theprocessing time of a vehicle kAK is set proportional to its capacity, i.e.
+pk ¼qk for all kAK.
+The width of the time window of a customer i AC is computed by λiqi,whereλi is a scalar uniformly selected from the interval ½1:1; 2:1.
+Each instance has up to 4 production stations.
+The locations of the start and end depot, construction sites, and production depots, lay in the euclidean plane.
+The travel time between two locations equals the euclidean distance, rounded upwards.
+For simplicity, the start and end depot are the same.
+The distance between each station and the depot is uniformly selected from the interval [1–30].
+Furthermore, for each customer, there exists a production station within a distance of [1–25].Finally,thetimelagγ is fixed to 5 for all deliveries.
+\subsection{Data sets}
 
@@
 
-- resultats avec les instances où toute la demande est satisfaite. Montrer la distance parcourue, le temps de première livraison, overtime, underwork, temps d'execution.
-
-- resultats avec les instances (8 au total) où uniquement la flotte de Unibéton a éét utilisée. (utiliser les mêmes pour la comparaison avec les données réelles )
-distance, demande servie, first, underwork, overtime.
-- resultats avec les autres instances
-- distance, demande servie, first delivery.
-
-% df = df[df['Name'].isin([20201010,20201021,20201107,20201017,20201031,20201109,20201026,20201121])]
-
-\begin{table}
-    \resizebox*{\textwidth}{!}{
-
-\begin{tabular}{lrrrrrrr}
-    \toprule
-    {} &  satisfiedCost &  driverUsed &  travelCost &  firstDeliveryCost &  underWorkCost &  overTimeCost &  runtime \\
-    Name     &                &             &             &                    &                &               &          \\
-    \midrule
-    20201121 &          226.5 &          13 &     1243.55 &               0.51 &           0.00 &          0.00 &     5.26 \\
-    20201010 &          267.0 &          13 &      745.10 &               0.00 &           0.00 &          0.00 &     5.32 \\
-    20201107 &          333.5 &          18 &      777.25 &               0.00 &         200.16 &          0.00 &     5.19 \\
-    20201031 &          375.0 &          15 &     1017.64 &               0.00 &         216.06 &         28.10 &     5.42 \\
-    20201114 &          388.0 &          19 &     1743.64 &               0.00 &         149.48 &          0.00 &     5.42 \\
-    20201128 &          776.0 &          31 &     2278.18 &               0.00 &           4.19 &        438.14 &     5.16 \\
-    20201024 &          937.5 &          35 &     4403.57 &               0.00 &          58.30 &       1005.24 &     5.72 \\
-    20201021 &         1160.5 &          73 &     6107.22 &               0.00 &        2535.40 &        185.02 &     5.28 \\
-    20201013 &         1565.0 &         102 &     9214.74 &              19.53 &        2955.35 &         95.74 &     5.32 \\
-    20201026 &         1839.5 &         116 &    13585.60 &              44.44 &        1826.70 &        227.56 &     5.41 \\
-    20201007 &         2512.5 &         127 &    15166.00 &               5.09 &         877.68 &        574.76 &     5.67 \\
-    \bottomrule
-    \end{tabular}
-    }
-\end{table}
-
+\subsubsection{ Results for the real-world datasets}
 \subsubsection{ Results for the instances of \cite{kinable2014concrete}}
 
-\cite{kinable2014concrete} provided a benchmark dataset for their variant of the $CDP$.
-Although their primary objective is to maximize the total load delivered each day, without considering loading operations at a plant and ensuring deliveries within specified time windows, we can still use this dataset to analyze and compare the performance of our algorithm under different conditions.
-Table \ref*{tab:cdp_instances} shows the performance of the GRASP heuristic along with the other methods (CP, MIP, SD-heuristic, and Hybrid) used in \cite{kinable2014concrete}.
-For each method, the table shows the average execution time in milliseconds and the corresponding gap between the obtained solution and the upper bound provided by \cite{kinable2014concrete}.
-Overall, the GRASP algorithm shows promise in providing near-optimal solutions in a relatively short time for this benchmark of the CDP.
-It outperformed all methods except CP, achieving an average distance of 5.7\% (13.8\%) with a runtime of 13,186 (574,983) milliseconds for Set A (Set B).
-The SD-heuristic appears to be much faster, but GRASP provides a better balance between solution quality and computation time.
-It should also be noted that the exact MIP, CP, and Hybrid are initialized with the results of the SD-heuristic.
-
-\begin{table}[htbp]
-    \centering
-    \caption{Performance of the GRASP heuristic on the CDP instances}
-    \label{tab:cdp_instances}
-    \resizebox{1\textwidth}{!}{
-    \begin{tabular}{@{}cccccccccccc@{}}
-    \toprule
-     &  & \multicolumn{2}{c}{CP} & \multicolumn{2}{c}{MIP} & \multicolumn{2}{c}{SD-heuristic} & \multicolumn{2}{c}{Hybrid} & \multicolumn{2}{c}{GRASP} \\ \cmidrule(l){3-10} \cmidrule(l){11-12} 
-     &  & Gap (\%) & Time (ms) & Gap (\%) & Time (ms) & Gap (\%) & Time (ms) & Gap (\%) & Time (ms) & Gap (\%) & Time (ms) \\ \cmidrule(l){3-10} \cmidrule(l){11-12} 
-    \multirow{2}{*}{Set A} & Avg & 4.2 & 197,012 & 7.3 & 13,405,798 & 9.1 & 23 & 7.0 & 100,765 & 5.7 & 13,186 \\ \cmidrule(l){2-10} \cmidrule(l){11-12} 
-     & Opt & \multicolumn{2}{c}{40/64} & \multicolumn{2}{c}{37/64} & \multicolumn{2}{c}{30/64} & \multicolumn{2}{c}{35/64} & \multicolumn{2}{c}{35/64} \\ \cmidrule(l){3-10} \cmidrule(l){11-12} 
-     \multirow{2}{*}{Set B} & Avg & 12.1 & 357,313 & - & - & 16.3 & 1,331 & - & - & 13.8 & 574,983\\ \cmidrule(l){2-10} \cmidrule(l){11-12} 
-     & Opt & \multicolumn{2}{c}{55/128} & \multicolumn{2}{c}{-} & \multicolumn{2}{c}{40/128} & \multicolumn{2}{c}{-} & \multicolumn{2}{c}{44/128} \\ \bottomrule
-    \end{tabular}
-    }
-    \end{table}
-
-\subsubsection{Comparisons with the real operations}
-
-- rouler les instances sans shift des livreurs. comparer la distance parcourue, et demande livrée.
-- rouler sans les livreurs externes. regarder la demande servie par jour.
+We run our algorithm on the dataset provided in \cite{kinable2014concrete}.
 

### 2023-08-15 result/cq/results.py

@@ -3,12 +3,10 @@ import numpy as np
 import subprocess
 
 destination_server = 'alious@beluga.calculquebec.ca'
-remote_path = "/home/alious/works/CimentQuebec/result/cdp/"
-local_path = "/Users/aliwa/Documents/Ulaval/thesis/works/cimentQuebec/result/cdp/"
+remote_path = "/home/alious/works/CimentQuebec/result/cq/"
+local_path = "/Users/aliwa/Documents/Ulaval/thesis/works/cimentQuebec/result/cq/"
 path = "/Users/aliwa/Documents/Ulaval/thesis/works/cimentQuebec/"
 
-df_kinableA = pd.read_csv("kinableA.csv", sep=";", header=0)
-df_kinableB = pd.read_csv("kinableB.csv", sep=";", header=0)
 
 def downloadFile(file="results.csv"):
     local_file = f"{local_path}/{file}"
@@ -74,10 +72,6 @@ def getStats(result_file, df_kinable, methods,statFile,instanceName):
             # print(df_better)
 
 methodA= ["UB","CP","MIP","Heuristic","Hybrid"]
-fileA=f"{local_path}/test_A.csv"
-fileA=downloadFile("test_A.csv")
-getStats(fileA,df_kinableA,methodA,"./statsA.csv","A")
-fileB=f"{local_path}/test_B.csv"
-fileB=downloadFile("test_B.csv")
-methodB=["UB","CP","Heuristic"]
-getStats(fileB,df_kinableB,methodB,"./statsB.csv","B")
+# fileA=f"{local_path}/test_A.csv"
+fileA=downloadFile("results.csv")
+# getStats(fileA,df_kinableA,methodA,"./statsA.csv","A")

### 2023-08-15 Data/instances/createInstance/create_instance.py
Save

@@ -35,59 +35,75 @@ def create_instance_CQ(Object,jour)->None:
         sortedMap[ord] = (cur['QUANTITY'].sum(), ordres[ordres['ORDER_ID']==ord]['SHIPTIME (min)'].min(),len(delivered_prod))
 
     sortedMap = sorted(sortedMap.items(), key=lambda x:(x[1][0],x[1][1],x[1][2]))
-    # print(sortedMap)
-    print(totalDemand)
+
     drivers = list(dict.fromkeys(tickets['DRIVER_NBR']))
     dict_driver = {}
+    usedLoc = set()
     for driver in drivers:
         trucks = list(dict.fromkeys(tickets['TRUCK_NBR'][tickets['DRIVER_NBR']==driver]))    
         
         if int(driver) == 1:
             for t in trucks:
                 d = Driver()
+                df_drv = tickets[tickets['TRUCK_NBR']==t]
+                _data = df_drv[["BEGIN_LOAD", "ARRIVE_PLANT"]].copy()
+                _data.sort_values(["BEGIN_LOAD"], inplace=True)
+                begin, end = _data.iloc[0, 0], _data.iloc[-1, 1]
+                a = (begin.hour * 60 + begin.minute - 10)
+                a =  a - a%5
+                d.start_shift = a
+                d.factory_nbr = Counter(df_drv['SHIP_LOC']).most_common(1)[0][0]
                 d.t_nbr = t
                 d.d_nbr = int(driver)
-                _df = tickets[tickets['TRUCK_NBR']==t]
-                _cap = max(_df['QUANTITY'])
+                _cap = max(df_drv['QUANTITY'])
                 if _cap <= 8:
                     d.capacity = 8
                 else:
                     d.capacity = 12
-                d.factory_nbr = Counter(_df['SHIP_LOC']).most_common(1)[0][0]
                 dict_driver[d.t_nbr] = d
 
         elif len(trucks) == 1:
+            d  = Driver()
+            df_drv = tickets[tickets['TRUCK_NBR'].isin(trucks)]
+            _data = df_drv[["BEGIN_LOAD", "ARRIVE_PLANT"]].copy()
+            _data.sort_values(["BEGIN_LOAD"], inplace=True)
+            begin, end = _data.iloc[0, 0], _data.iloc[-1, 1]
+            a = (begin.hour * 60 + begin.minute - 10)
+            a =  a - a%5
             d = Driver()
[…]

### 2023-08-15 Data/instances/createInstance/driver.py

@@ -14,6 +14,7 @@ class Driver:
         self.t_nbr = ""
         self.d_nbr = -1
         self.description = ""
+        self.start_shift = 0
 
     def __str__(self):
         return f"Driver Nbr {self.t_nbr} {self.d_nbr} Rank {self.rank} {self.description} depot n {self.depot_nbr} fac {self.factory_nbr} de capacité {self.capacity}\n"

### 2023-08-15 Data/instances/createInstance/instances.py

@@ -5,28 +5,46 @@ import os
 from driver import Driver
 import datetime
 from collections import Counter
+from create_instance import DATA_FOLDER
 
-INSTANCE_FOLDER = "/".join([os.path.dirname(__file__), "instances/"])
-DATA_FOLDER = "/".join([os.path.dirname(__file__), "data/"])
 
+Instances_with_extern_fleet = set()
+Instances_with_intern_fleet = set()
 
 class Instance:
     def __init__(self) -> None:
-        self.df_orders = pd.read_excel(f"{DATA_FOLDER}/DORDERSTAB_BETON.xlsx", engine='openpyxl', sheet_name="data")
-        self.df_depots = pd.read_excel(f"{DATA_FOLDER}/adresses_complete.xlsx")
-        self.df_tickets = pd.read_excel(f"{DATA_FOLDER}/DTICKETHISTAB.xlsx")
-        self.df_employee = pd.read_excel(f"{DATA_FOLDER}/EMPLOYETAB.xlsx", engine='openpyxl')
-        self.df_depots_loc = pd.read_excel(f"{DATA_FOLDER}/LOCATIONTAB.xlsx", engine='openpyxl')
+        self.df_orders = None
+        self.df_depots = None
+        self.df_tickets = None
+        self.df_employee = None
+        self.df_depots_loc = None
         self.df_driver = None
         self.list_driver_nbr = []
         self.instance_size = dict()
         self.dict_driver = {}
         self.depot_loc_id = {}
-        self.clean_depot()
-        self.clean_employee()
-        self.clean_tickets()
-        self.get_driver_from_ticket()
-
+        #self.clean_depot()
+        #self.clean_employee()
+        #self.clean_tickets()
+        #self.get_driver_from_ticket()
+        
+    
+    def init_order(self)->None:
+        self.df_orders = pd.read_excel(f"{DATA_FOLDER}/DORDERSTAB_BETON.xlsx", engine='openpyxl', sheet_name="data")
+        self.df_orders.sort_values(by='QUANTITY',inplace=True)
+        
+    def init_depot(self):
+        self.df_depots = pd.read_excel(f"{DATA_FOLDER}/adresses_complete.xlsx")
+           
+    def init_tickets(self):
+        self.df_tickets = pd.read_excel(f"{DATA_FOLDER}/DTICKETHISTAB.xlsx")
+    
+    def init_employee(self):
[…]

### 2023-08-15 codeCQ/Data.cpp

@@ -175,6 +175,7 @@ void Data::LoadCQInstance() {
             Driver d;
             inputFile >> d.id >> d.no >> d.truck_nbr >> d.rank >> d.depotID >>
                       d.capacity >> d.start_shift_time;
+//            d.start_shift_time = 260;
             d.id = i;
             if (d.rank == -1) d.rank = MAX_RANK;
             Node d1;
@@ -220,8 +221,8 @@ void Data::LoadCQInstance() {
             n.constID = n.custID;
             n.id = GetNodeCount();
             n.no = n.id + 1;
-//            n.late_tw = n.early_tw + ceil(UnloadingTime(n.demand, n.demand)) + 60;
-            n.late_tw = n.early_tw + Parameters::MAX_LATE_TW;
+            n.late_tw = n.early_tw + ceil(UnloadingTime(n.demand, n.demand)) + 60;
+//            n.late_tw = n.early_tw + Parameters::MAX_LATE_TW;
             n.type = Parameters::CUSTOMER;
             n.siteID = n.id;
             Node d1;

### 2023-08-15 codeCQ/GRASP.cpp

@@ -36,7 +36,7 @@ void GRASP<NodeT, DriverT>::Optimize(
     }
     //TODO SortNode<Node, Driver>::radixSortGreatDemand(list_cust, s.GetData()->GetMaxDemand());
     std::vector<int> SORT_TYPE_VEC{
-            Parameters::SORT::ONE,
+//            Parameters::SORT::ONE,
                             Parameters::SORT::TWO,
 //            Parameters::SORT::THREE,
 //                            Parameters::SORT::FOUR,
@@ -175,6 +175,7 @@ void GRASP<NodeT, DriverT>::Optimize(
             // printf("Best iter %d %2.1lf lateness %d\n", best.iter, s.GetCost().satisfiedCost, s.GetCost().lateDeliveryCost);
             printf("iter_k %d %2.1lf lateness %2.0lf\n", iter_k, s.GetCost().satisfiedCost,
                    s.GetCost().lateDeliveryCost);
+            Prompt::print(best.unscheduledCustomers);
         }
     }
     s.duration_time = _chrono.getTime();

### 2023-08-15 codeCQ/gantt.py

@@ -1,0 +1,83 @@
+import pandas as pd
+import numpy as np
+import plotly.express as px
+
+# Transform minutes to format"hh:mm:ss:"
+def time_to_date(date,t):
+    date=str(date)
+    if len(date.split('-'))==1:
+        date="2022-05-16"
+    else:
+        date=date[:4]+"-"+date[4:6]+"-"+date[6:]
+    min=int(t%60)
+    hour=int(t//60)
+    if min<10:
+        min=f"0{min}"
+    new_date=f"{date} {hour}:{min}:00"
+    if hour>23: 
+        print(new_date,' ',t)
+    return new_date
+
+
+# Creer des diagrammes de Gantt
+# Par ordre: afficher les heures d'arrivée et de départ des différents noeuds
+# J'ai besoin de la date, des heures d'arrivée et de départ.
+def showGanttDel(df):
+    df['task'] = df.apply(lambda x: f" del {x.Del}_{x.Order}_{x.Cust}",axis=1)
+    fig = px.timeline(df, x_start="startUnload", x_end="endUnload", y="task",title="All deliveries",color="Cust")
+    # fig.update_yaxes(autorange="reversed") # otherwise tasks are listed from the bottom up
+    fig.show()
+
+def showGanttDock(df):
+    df['task'] = df.apply(lambda x: f" dock {x.Del}_{x.Order}_{x.Cust}",axis=1)
+    fig = px.timeline(df, x_start="startLoad", x_end="endLoad", y="task",title="All Loading",color="Cust")
+    # fig.update_yaxes(autorange="reversed") # otherwise tasks are listed from the bottom up
+    fig.show()
+
+# Creer des diagrammes de Gantt
+# Par ordre: afficher les heures d'arrivée et de départ des différents noeuds
+# J'ai besoin de la date, des heures d'arrivée et de départ.
+def showGanttOrder(df):
+    orderList=np.unique(df['ordreId'])
+    df['task'] = df[['delId','ordreId']].apply(lambda x: f"del {x.delId}_{x.ordreId}",axis=1)
+    
+    for ord in orderList:
+        df_ordre=df.set_index('ordreId').xs(ord)
+        if type(df_ordre)==pd.core.series.Series:
+            df_ordre = df_ordre.to_frame().T
+        df_ordre.sort_values(by=['ArrivalTimeDel'],inplace=True)
+        fig = px.timeline(df_ordre, x_start="start", x_end="end", y="task",title=f'order {ord}',color="driverId")
[…]

### 2023-08-15 codeCQ/heur/CustInsertion.cpp

@@ -14,13 +14,15 @@ void CustInsertion::Insert(Sol &s) {
     assert(!customersList.empty());
     const int insertion_type = Parameters::LOAD_INSERTION;
     InsertWithBactrack(s, customersList);
-    bool beginLoop = false;
+    bool beginLoop = true;
     while (!removedList.empty() && beginLoop) {
         double demand = s.updateCost.satisfiedCost;
-        customersList = removedList;
+        customersList =   std::vector<Customer*>(removedList.begin(),removedList.end());
+        cout<<" try "<<s.GetLastCost()<<endl;
+        cout<<" try backward "<<s.updateCost<<endl;
+        Prompt::print(customersList);
         Sol cur(s);
         Parameters::LOAD_INSERTION = Parameters::DEPOTINSERTION::BACKWARD;
-
         InsertWithBactrack(cur, customersList);
         if (cur < s) {
             s = cur;
@@ -34,7 +36,6 @@ void CustInsertion::Insert(Sol &s) {
 void CustInsertion::InsertWithBactrack(Sol &s, std::vector<Customer *> &list) {
     if (list.size() > 1) {Sort(s, list, CustInsertion::_k);}
     removedList.clear();
-    removedList.shrink_to_fit();
     listMoves.clear();
     listMoves.shrink_to_fit();
     listMoves.resize(s.GetDeliveryCount());
@@ -53,18 +54,20 @@ void CustInsertion::InsertWithBactrack(Sol &s, std::vector<Customer *> &list) {
         Order *cur_order = s.GetRandomOrder(c, UsedOrder);
         if (cur_order == nullptr) {
             nextCustomer = true;
+            removedList.insert(c);
             s.UnassignCustomer(c);
             continue;
         }
         UsedOrder.insert(cur_order->orderID);
         nextCustomer = Insert(s, c, cur_order);
     }
+    s.Update();
 }
 
 bool CustInsertion::Insert(Sol &s, Customer *c, Order *cur_order) {
     bool nextCustomer = false;
     int depth = 0;
-//    if (c->custID == 5)
+//    if (c->custID == 57)
 //        Parameters::SHOW = true;
     for (int j = 0; j < s.GetDeliveryCount(cur_order);) {
         Delivery *del = s.GetDelivery(cur_order, j);
@@ -80,6 +83,7 @@ bool CustInsertion::Insert(Sol &s, Customer *c, Order *cur_order) {
 
[…]

### 2023-08-15 codeCQ/heur/InsRmvMethodFast.cpp

@@ -147,13 +147,10 @@ void InsRmvMethodFast::ApplyInsertMove(Sol &s, Move<Delivery, Driver, MoveVrp> &
     s.InsertAfter(m.move.dock, m.move.prev, m.to);
     s.InsertAfter(m.n, m.move.dock, m.to);
     s.AssignDeliveryToCustomer(m.n);
-    assert(m.demand > 0);
     s.DeliveryLoad[m.n->delID] = m.demand;
-    assert(s.DeliveryLoad[m.n->delID] > 0);
     s.ArrivalTime[m.move.dock->id] = m.arrival_dock;
     s.ArrivalTime[m.n->id] = m.arrival_del;
     s.DepotAssignTo[m.move.dock->id] = m.move.depot;
-    assert(m.arrival_dock <= m.arrival_del + s.Travel(m.move.dock, m.n));
 }
 
 Move<Delivery, Driver, MoveVrp> InsRmvMethodFast::GetCost(Sol &s, Delivery *n, Driver *d, Cost &solcost, double demand,
@@ -441,6 +438,7 @@ Move<Delivery, Driver, MoveVrp> InsRmvMethodFast::GetCost(Sol &s, Delivery *n, D
         newcost.satisfiedCost = demand;
         newcost.isFeasible = true;
         newcost += s.updateCost;
+        newcost.undeliveredCost -=s.updateCost.undeliveredCost;
 
         newcost.underWorkCost = newcost.underWorkCost - d->underWork + Sol::GetUnderWorkCost(node_arr
                                                                                              + s.Travel(n->distID,

### 2023-08-15 codeCQ/heur/Solver.cpp

@@ -29,14 +29,16 @@ void Solver::run() {
 
     CDPSolver::findCorrelation(dat, listInt, linkedClientSlot, linkedClientDemand, linkedClients, linkedClientsInf,
                                linkedClientSup);
-
     Sol s(&dat);
-    s.keyCustomers = feasibleClients;
+    s.keyCustomers =  feasibleClients;
+
     SolveGrasp(s, dat, linkedClientSlot, Parameters::ITERATION);
     SaveResults(s);
-    s.exportCSVFormat("s1.csv");
+    s.exportCSVFormat(dat.sol_output);
     cout << s.GetLastCost() << endl;
     Prompt::print(s.unscheduledCustomers);
+//    s.ShowDrivers();
+//    s.ShowAllSchedule();
 }
 
 void Solver::SolveGrasp(Sol &s, Data &dat, vector<set<int>> const &linkedClientSlot, int iter) {
@@ -112,28 +114,28 @@ void Solver::SolveGrasp(Sol &s, Data &dat, vector<set<int>> const &linkedClientS
     AllInsertionOperator d2(&driverIns3, driverInfo[2].first, "Builder 3 " + driverInfo[2].second);
 
 
-    grasp.AddInsertOperatorVrp(&c0_1);
+//    grasp.AddInsertOperatorVrp(&c0_1);
     grasp.AddInsertOperatorVrp(&c0_3);
-    grasp.AddInsertOperatorVrp(&c1_1);
-    grasp.AddInsertOperatorVrp(&c1_3);
-    grasp.AddInsertOperatorVrp(&c2_1);
-    grasp.AddInsertOperatorVrp(&c2_3);
-    grasp.AddInsertOperatorVrp(&c3_1);
-    grasp.AddInsertOperatorVrp(&c3_3);
-    grasp.AddInsertOperatorVrp(&c4_1);
-    grasp.AddInsertOperatorVrp(&c4_3);
-    grasp.AddInsertOperatorVrp(&c5_1);
-    grasp.AddInsertOperatorVrp(&c5_3);
+//    grasp.AddInsertOperatorVrp(&c1_1);
+//    grasp.AddInsertOperatorVrp(&c1_3);
+//    grasp.AddInsertOperatorVrp(&c2_1);
+//    grasp.AddInsertOperatorVrp(&c2_3);
[…]

### 2023-08-15 solution/cq/gantt.py

@@ -5,10 +5,8 @@ import plotly.express as px
 # Transform minutes to format"hh:mm:ss:"
 def time_to_date(date,t):
     date=str(date)
-    if len(date.split('-'))==1:
-        date="2022-05-16"
-    else:
-        date=date[:4]+"-"+date[4:6]+"-"+date[6:]
+    # print(date)
+    date=date[:4]+"-"+date[4:6]+"-"+date[6:]
     min=int(t%60)
     hour=int(t//60)
     if min<10:
@@ -21,8 +19,7 @@ def time_to_date(date,t):
 
 def show(df):
     df['task'] = df.apply(lambda x: f" del {x.Del}",axis=1)
-    fig = px.timeline(df, x_start="start", x_end="end", y="Del",title="All deliveries",color="task")
-    # fig = px.timeline(df, x_start="startLoad", x_end="endLoad", y="task",title="All deliveries",color="Cust")
+    fig = px.timeline(df, x_start="start", x_end="end", y="Driver",title="All drivers",color="Driver",template='ggplot2')
     fig.update_yaxes(autorange="reversed") # otherwise tasks are listed from the bottom up
     fig.show()
 
@@ -30,9 +27,10 @@ def show(df):
 # Par ordre: afficher les heures d'arrivée et de départ des différents noeuds
 # J'ai besoin de la date, des heures d'arrivée et de départ.
 def showGanttDel(df):
-    df['task'] = df.apply(lambda x: f" del {x.Del}_{x.Order}_{x.Cust}",axis=1)
-    fig = px.timeline(df, x_start="startUnload", x_end="endUnload", y="task",title="All deliveries",color="Cust")
-    # fig.update_yaxes(autorange="reversed") # otherwise tasks are listed from the bottom up
+    df['task'] = df.apply(lambda x: f" del {x.Order}_{x.Cust}",axis=1)
+    # df['task'] = df.apply(lambda x: f" del {x.Del}",axis=1)
+    fig = px.timeline(df, x_start="start", x_end="end", y="Del",title="All deliveries",color="task")
+    fig.update_yaxes(autorange="reversed") # otherwise tasks are listed from the bottom up
     fig.show()
 
 def showGanttDock(df):
@@ -80,11 +78,12 @@ def showGanttDepot(df):
     depotList=np.unique(df['Depot'])
     df['task'] = df.apply(lambda x: f"del {x.Del}_{x.Order}_{x.Cust}",axis=1)
     
-    for ord in depotList:
[…]

### 2023-08-17 Article/article.tex
save article

@@
 \usepackage{tikz}
+\usepackage{subcaption}
 \usetikzlibrary{automata, topaths, calc, positioning, shapes, backgrounds, fit, matrix}
@@
 The procedure stops and returns the best overall solution when certain stopping conditions, such as reaching a time limit or a maximum number of iterations, are met.
-In our implementation, we also add a path relinking procedure that uses information about previous solutions to improve the search process.
+%In our implementation, we also add a path relinking procedure that uses information about previous solutions to improve the search process.
 
@@
         {\small
-        \begin{algorithm}[hpt]
+        \begin{algorithm}[hbpt]
             \caption{Pseudo-code of the GRASP algorithm }
@@
 
-            $counter \leftarrow  0$
+            % $counter \leftarrow  0$
 
@@
 
-                $T \leftarrow T \cup \{S\}$
+                % $T \leftarrow T \cup \{S\}$
 
@@
 
-                \uIf{$counter \text{ } mod \text{ } \phi == 0$}{
-                    $S^* \leftarrow PathRelinking(T)$
-
-                    $T \leftarrow \emptyset$
-                }
-
-                $counter \leftarrow counter + 1$
+                % \uIf{$counter \text{ } mod \text{ } \phi == 0$}{
+                %     $S^* \leftarrow PathRelinking(T)$
+
+                %     $T \leftarrow \emptyset$
+                % }
+
+                % $counter \leftarrow counter + 1$
 
@@
         %  {\small
-        \begin{algorithm}[]
+        \begin{algorithm}[htpb]
             \caption{Customer-Based insertion algorithm }
@@
         %  {\small
-        \begin{algorithm}[H]
+        \begin{algorithm}[htb]
             \caption{Delivery-based insertion algorithm }
@@
 
-\subsection{Path relinking}
-
-After the construction and local search steps, it is possible that some customers remain unscheduled in one solution while being accommodated in another.
+% \subsection{Path relinking}
+
+% After the construction and local search steps, it is possible that some customers remain unscheduled in one solution while being accommodated in another.
 The goal of the path relinking procedure is to use each customer's scheduling information from existing solutions and combine them effectively to construct a solution that schedules all customers.
@@
 The dataset is divided into small, medium, and large sets according to the total number of daily orders.
-We have summarized the dataset in the Table~\ref{tab:instances}.
+We have summarized the dataset in Table~\ref{tab:instances}.
 
@@
     \label{tab:instances}
+    \small
     \resizebox*{0.8\textwidth}{!}{
@@
         \midrule
-        Small  &     8 &    226.5-937.5 &     7-14 &    4-11 &    13-35 &    1-3 \\
-        Medium &    14 &  1,160.5-2,971.0 &    43-98 &   40-89 &   76-137 &    6-8 \\
-        Large  &    14 &  3,078.5-3,953.5 &  107-136 &  92-127 &  129-150 &    8 \\
+        Small  &     8 &    226.5 - 937.5 &     7 - 14 &    4 - 11 &    13-35 &    1-3 \\
+        Medium &    14 &  1,160.5 - 2,971.0 &    43 - 98 &   40 - 89 &   76-137 &    6-8 \\
+        Large  &    14 &  3,078. - 3,953.5 &  107 - 136 &  92 - 127 &  129-150 &    8 \\
         \bottomrule
@@
     \label{tab:cdp_instances}
+    \scriptsize
     \resizebox{1\textwidth}{!}{
@@
 
-\begin{table}[htpb]
+\begin{table}[hbt]
     \centering
@@
     \label{tab:problem_parameters}
+    \small
 %    \resizebox*{\textwidth}{!}{
@@
         Max delay between two consecutive deliveries of different orders ($\gamma^2)$ & 25 min \\
-    %    $\beta_1$ & 10000  \\
-    %    $\beta_2$ & 10000  \\
-    %    $\beta_3$ & 1000  \\
-    %    $\beta_4$ & 100  \\
+       $\beta_1$ & 10,0000  \\
+       $\beta_2$ & 1,0000  \\
+       $\beta_3$ & 30  \\
+       $\beta_4$ & 20  \\
         \bottomrule
@@
 
-\begin{table}[htbp]
+\begin{table}[htb]
     \centering
@@
     \label{tab:runtime}
-    \resizebox*{0.5\textwidth}{!}{
-        \begin{tabular}{llcccc}
+    \scriptsize
+    \resizebox*{0.8\textwidth}{!}{
+        \begin{tabular}{llccccc}
             \toprule
-            &  &      \multicolumn{4}{c}{Runtime (min)} \\
-            \cmidrule{3-6}
-            & &      5   & 10  &  30  & 60  \\
-            %  & {} &         &         &        &        \\
+            &  &      \multicolumn{5}{c}{Runtime (min)} \\
+            \cmidrule{3-7}
+                  &  &      1  &      5  &      10 &     30 &     60 \\
             \midrule
-            \multirow{1}{*}{Small} & \% Load &  100.00 &  100.00 &  100.0 &  100.0 \\
-                %   & \# instances &    8 &    8 &    8 &    8 \\
-            \cmidrule{2-6}
-                  \multirow{1}{*}{Medium} & \% Load &   99.86 &   99.86 &  100.0 &  100.0 \\
-                %   & \# instances &   13 &   13 &   14 &   14 \\
-                  \cmidrule{2-6}
-                  \multirow{1}{*}{Large} & \% Load &   99.79 &   99.93 &  100.0 &  100.0 \\
-                %   & \# instances &   12 &   13 &   14 &   14 \\
+            \multirow{2}{*}{Small}  & \%Load &  100.00 &  100.00 &  100.00 &  100.0 &  100.0 \\
+                  & \#Instance &    8 &    8 &    8 &    8 &    8 \\
+                  \cmidrule{2-7}
+            \multirow{2}{*}{Medium}  & \%Load &   99.71 &   99.93 &   99.86 &  100.0 &  100.0 \\
+                  & \#Instance &   11 &   13 &   13 &   14 &   14 \\
+                  \cmidrule{2-7}
+            \multirow{2}{*}{Large}  & \%Load &   99.64 &   99.64 &   99.79 &  100.0 &  100.0 \\
+                  & \#Instance &   10 &   10 &   12 &   14 &   14 \\
             \bottomrule
@@
 
-Table \ref*{tab:results_grasp_ls} shows the detailed results for each instance.
+Table \ref*{tab:all_instances} shows the detailed results for each instance.
 The runtime is set to 3600 seconds.
@@
 This tells us that we have a maximum delay that varies between 0 and 12 minutes for small instances, 1 and 60 minutes for medium instances, and 50 to 60 minutes for large instances.
-
-\begin{table}[htbp]
+\begin{table}[htb]
+    \centering
+    \caption{Results with all instances}
+    \label{tab:all_instances}
+    \resizebox*{1.2\textwidth}{!}{
+    \begin{tabular}{llrrrrrrrrrrrrrrrrrr}
+        \toprule
+         &  & \multicolumn{18}{c}{Best} \\
+         &  & Demand & $TC$ & $FDD$ & $mFDD$ & $DUC$ & $DOC$ & $Z$ & \#K & Time & $TC$ & $FDD$ & $mFDD$ & $DUC$ & $DOC$ & Z & \#K & $PUO$ & Time \\
+         &  & (min) & (min) & (min) & (min) & (min) & (min) & (min) & (min) & (min) & (min) & (min) & (min) & (min) & (min) & (min) & (min) & (min) & (min) \\
+        size & Instance &  &  &  &  &  &  &  &  &  &  &  &  &  &  &  &  &  &  \\
+        \midrule
+        \multirow[c]{8}{*}{Small} & C_13_11_12_1 & 226.50 & 1,243.55 & 0.00 & 0.00 & 0.00 & 0.00 & 1,243.55 & 13 & 60.34 & 1,243.55 & 0.00 & 0.00 & 37.68 & 0.00 & 2,373.86 & 13.00 & 0.00 & 60.23 \\
+         & C_13_5_8_1 & 267.00 & 745.10 & 0.00 & 0.00 & 0.00 & 0.00 & 745.10 & 13 & 60.28 & 745.10 & 0.00 & 0.00 & 0.00 & 0.00 & 745.10 & 13.00 & 0.00 & 60.29 \\
+         & C_18_6_11_2 & 333.50 & 753.26 & 0.00 & 0.00 & 125.33 & 0.00 & 4,513.25 & 18 & 60.21 & 753.26 & 0.00 & 0.00 & 194.68 & 0.00 & 6,593.74 & 18.00 & 0.00 & 60.27 \\
+         & C_15_4_7_2 & 375.00 & 1,014.25 & 0.00 & 0.00 & 172.13 & 76.10 & 7,700.13 & 15 & 60.39 & 1,014.93 & 0.00 & 0.00 & 132.77 & 291.65 & 10,830.80 & 15.00 & 0.00 & 60.37 \\
+         & C_19_7_8_2 & 388.00 & 1,745.83 & 0.00 & 0.00 & 224.47 & 87.86 & 10,237.08 & 19 & 60.21 & 1,765.51 & 0.00 & 0.00 & 314.77 & 150.86 & 14,225.62 & 18.60 & 0.00 & 60.21 \\
+         & C_29_10_14_3 & 613.50 & 3,628.79 & 49.18 & 18.77 & 112.94 & 547.16 & 67,145.19 & 29 & 60.88 & 3,741.98 & 48.61 & 21.45 & 210.99 & 735.59 & 73,392.66 & 28.40 & 0.00 & 60.65 \\
+         & C_31_8_11_3 & 776.00 & 2,297.20 & 0.00 & 0.00 & 606.66 & 566.96 & 31,836.07 & 30 & 61.06 & 2,330.55 & 0.00 & 0.00 & 781.61 & 656.53 & 38,909.59 & 29.40 & 0.00 & 60.56 \\
+         & C_35_9_10_3 & 937.50 & 4,174.37 & 0.00 & 0.00 & 221.69 & 1,161.60 & 34,056.98 & 35 & 61.27 & 4,175.30 & 0.00 & 0.00 & 341.84 & 1,329.53 & 41,021.00 & 34.40 & 0.00 & 60.96 \\
+        \multirow[c]{14}{*}{Medium} & C_76_40_43_6 & 1,160.50 & 5,967.04 & 6.18 & 3.48 & 2,961.59 & 644.42 & 113,884.73 & 71 & 60.24 & 6,311.56 & 17.24 & 9.45 & 3,014.61 & 705.95 & 128,106.36 & 67.80 & 0.00 & 60.38 \\
+         & C_104_42_47_8 & 1,565.00 & 9,777.47 & 32.02 & 13.04 & 2,829.97 & 240.64 & 131,507.57 & 98 & 60.30 & 9,358.19 & 34.11 & 20.52 & 3,890.10 & 678.98 & 173,754.55 & 94.40 & 0.00 & 60.26 \\
+         & C_94_63_70_7 & 1,746.50 & 14,428.70 & 78.89 & 19.26 & 1,147.89 & 1,464.35 & 157,038.10 & 94 & 60.52 & 14,369.08 & 90.19 & 24.90 & 1,353.87 & 1,391.38 & 173,007.04 & 93.20 & 0.00 & 60.45 \\
+         & C_116_67_78_8 & 1,839.50 & 13,871.90 & 67.25 & 33.18 & 2,347.71 & 523.28 & 162,021.52 & 114 & 60.28 & 13,602.88 & 67.90 & 32.59 & 2,419.36 & 857.33 & 171,229.15 & 113.60 & 0.00 & 60.36 \\
+         & C_116_71_82_8 & 2,060.00 & 12,089.70 & 147.66 & 27.17 & 1,966.01 & 922.73 & 237,183.56 & 116 & 60.30 & 12,678.16 & 158.64 & 35.84 & 1,991.31 & 1,125.75 & 253,575.07 & 114.20 & 0.00 & 60.26 \\
+         & C_117_57_70_6 & 2,327.50 & 15,153.30 & 114.16 & 22.20 & 1,823.68 & 1,671.48 & 217,448.30 & 117 & 60.45 & 15,495.00 & 142.42 & 34.52 & 1,388.61 & 1,680.45 & 233,181.42 & 117.00 & 0.00 & 60.39 \\
+         & C_137_79_83_7 & 2,425.00 & 18,360.30 & 335.15 & 34.20 & 2,552.72 & 1,794.17 & 465,977.30 & 135 & 60.26 & 18,960.44 & 375.76 & 48.60 & 2,124.67 & 1,857.88 & 495,614.54 & 135.60 & 0.00 & 60.43 \\
+         & C_127_66_80_8 & 2,512.50 & 17,599.10 & 14.37 & 5.00 & 1,046.33 & 1,242.32 & 88,203.50 & 126 & 60.15 & 16,033.26 & 10.42 & 3.60 & 1,579.58 & 1,552.10 & 104,879.09 & 125.40 & 0.00 & 60.26 \\
+         & C_128_68_74_7 & 2,595.00 & 18,728.80 & 191.00 & 42.50 & 1,418.80 & 2,841.15 & 309,116.80 & 127 & 60.36 & 19,774.38 & 361.55 & 49.14 & 944.63 & 2,616.67 & 461,998.19 & 127.80 & 0.00 & 60.61 \\
+         & C_136_85_97_8 & 2,673.00 & 18,894.60 & 137.37 & 48.57 & 1,155.58 & 2,137.00 & 233,669.00 & 134 & 60.71 & 18,223.56 & 188.51 & 46.67 & 2,134.46 & 2,352.19 & 317,811.36 & 131.40 & 0.00 & 60.48 \\
+         & C_128_78_85_8 & 2,685.50 & 18,041.60 & 192.36 & 35.71 & 1,374.29 & 3,234.19 & 316,314.10 & 127 & 60.68 & 18,463.86 & 305.84 & 39.44 & 1,277.34 & 3,156.96 & 425,759.26 & 127.20 & 0.00 & 60.37 \\
+         & C_131_77_85_8 & 2,893.50 & 18,063.00 & 200.24 & 43.22 & 780.69 & 1,539.10 & 272,509.58 & 131 & 60.23 & 18,003.00 & 212.58 & 47.46 & 1,248.71 & 1,716.14 & 302,370.43 & 129.60 & 0.00 & 60.41 \\
+         & C_137_89_97_7 & 2,939.50 & 22,038.10 & 438.27 & 58.56 & 849.81 & 3,461.14 & 555,029.11 & 137 & 60.33 & 23,293.04 & 636.01 & 56.14 & 849.88 & 3,523.86 & 755,280.53 & 137.00 & 0.00 & 60.47 \\
+         & C_133_84_98_7 & 2,971.00 & 20,992.60 & 1,061.17 & 59.39 & 546.11 & 2,742.51 & 1,153,396.07 & 133 & 60.30 & 22,000.05 & 1,157.60 & 59.25 & 805.41 & 2,291.38 & 1,249,585.08 & 133.00 & 18.70 & 60.67 \\
+        \multirow[c]{14}{*}{Large} & C_132_98_109_8 & 3,078.50 & 23,425.80 & 160.94 & 29.44 & 493.20 & 4,426.10 & 287,683.71 & 132 & 60.40 & 23,469.12 & 247.44 & 31.77 & 548.73 & 4,162.33 & 370,622.52 & 131.80 & 0.00 & 60.51 \\
+         & C_129_101_119_8 & 3,229.00 & 19,453.30 & 166.74 & 35.78 & 1,138.10 & 4,719.55 & 314,728.30 & 129 & 60.31 & 19,051.68 & 268.97 & 40.76 & 1,232.06 & 3,820.36 & 401,387.86 & 129.00 & 0.00 & 60.45 \\
+         & C_141_114_136_8 & 3,350.50 & 21,863.20 & 943.94 & 55.35 & 1,177.01 & 2,513.42 & 1,051,386.90 & 141 & 60.36 & 21,505.60 & 1,127.23 & 57.20 & 1,422.68 & 2,339.21 & 1,238,202.03 & 141.00 & 8.70 & 60.58 \\
+         & C_140_114_132_8 & 3,401.50 & 23,969.70 & 988.55 & 54.37 & 981.26 & 3,253.14 & 1,107,019.45 & 140 & 60.55 & 23,938.44 & 1,055.09 & 54.86 & 910.17 & 3,269.09 & 1,171,715.10 & 139.80 & 0.00 & 60.68 \\
+         & C_143_101_123_8 & 3,437.50 & 26,284.90 & 335.48 & 43.51 & 566.06 & 3,912.96 & 457,006.99 & 143 & 60.80 & 25,711.02 & 352.25 & 46.32 & 902.98 & 4,085.07 & 486,754.99 & 143.00 & 0.00 & 60.54 \\
+         & C_137_112_129_8 & 3,471.00 & 22,338.30 & 275.09 & 45.62 & 375.23 & 3,819.75 & 385,076.08 & 137 & 60.76 & 22,562.40 & 340.81 & 49.10 & 402.39 & 3,371.09 & 442,869.08 & 137.00 & 0.00 & 60.82 \\
+         & C_142_114_129_8 & 3,499.50 & 22,387.60 & 654.63 & 55.27 & 1,090.38 & 4,472.55 & 799,184.00 & 142 & 60.58 & 22,478.28 & 800.95 & 57.01 & 1,294.76 & 4,168.21 & 945,633.44 & 141.60 & 0.00 & 61.20 \\
+         & C_149_98_122_8 & 3,513.00 & 21,224.70 & 684.52 & 51.12 & 2,267.63 & 3,046.60 & 834,708.60 & 148 & 61.18 & 20,767.50 & 778.44 & 54.95 & 2,236.93 & 2,852.20 & 923,357.90 & 147.75 & 3.00 & 60.77 \\
+         & C_139_98_108_8 & 3,541.00 & 22,043.20 & 522.32 & 51.62 & 482.79 & 3,648.54 & 631,819.76 & 139 & 61.46 & 21,613.40 & 661.04 & 55.50 & 852.30 & 3,716.15 & 782,546.76 & 138.60 & 0.00 & 61.16 \\
+         & C_144_108_122_8 & 3,670.50 & 25,806.90 & 675.42 & 50.38 & 548.50 & 3,209.16 & 781,866.07 & 144 & 60.38 & 24,738.28 & 746.03 & 56.42 & 664.90 & 3,465.76 & 860,025.70 & 144.00 & 0.00 & 61.69 \\
+         & C_142_92_107_8 & 3,684.50 & 23,586.40 & 182.64 & 24.02 & 802.97 & 4,396.54 & 318,243.24 & 142 & 60.40 & 24,553.30 & 298.65 & 38.73 & 644.19 & 4,423.61 & 431,001.04 & 142.00 & 0.00 & 60.47 \\
+         & C_138_114_136_8 & 3,739.50 & 23,639.00 & 600.38 & 55.18 & 1,230.76 & 5,313.96 & 767,223.00 & 137 & 60.16 & 25,366.94 & 791.53 & 54.14 & 668.63 & 5,501.79 & 946,993.86 & 137.20 & 0.00 & 60.90 \\
+         & C_150_127_136_8 & 3,946.50 & 27,332.90 & 507.30 & 57.10 & 250.32 & 5,295.70 & 648,057.44 & 150 & 60.81 & 25,785.24 & 613.80 & 57.10 & 629.13 & 4,686.38 & 752,189.57 & 149.60 & 0.00 & 60.91 \\
+         & C_148_112_131_8 & 3,953.50 & 25,559.20 & 886.90 & 55.98 & 559.29 & 4,213.16 & 1,013,500.22 & 148 & 60.52 & 26,231.23 & 1,002.04 & 57.21 & 517.76 & 3,467.44 & 1,113,149.29 & 148.00 & 6.90 & 61.01 \\
+        \bottomrule
+        \end{tabular}
+}
+\end{table}
+
+\iffalse
+\begin{table}[!h]
     \caption{Results with all instances}
@@
 \end{table}
-
+\fi
+
+Pour avoir une idée de la file d'attente au niveau des dépôts, traçons le diagramme Gantt des opérations des 3 dépôts utilisés dans l'instance c29,10,14,3
+
+\begin{table}[]
+    \centering
+    \caption{Details of instance $C29\_10\_14\_3$}
+    \label{tab:my-table}
+   \resizebox*{1\textwidth}{!}{
+    \begin{tabular}{@{}c|cccccccccccccc@{}}
+        \toprule
+        Order & 0 & 1 & 2 & 3 & 4 & 5 & 6 & 7 & 8 & 9 & 10 & 11 & 12 & 13 \\ \midrule
+        Customer & 0 & 1 & 2 & 3 & 4 & 5 & 6 & 7 & 8 & 8 & 9 & 9 & 9 & 9 \\
+        \midrule
+        Demand & 3.0 & 6.0 & 6.5 & 9.5 & 10.0 & 15.0 & 17.0 & 118.0 & 1.0 & 210.5 & 1.0 & 3.0 & 19.0 & 194.0 \\
+        \midrule
+        Plant & 1 & 1 & 1 & 1 & 1 & 1 & 1 & 2 & 0 & 0 & 2 & 2 & 2 & 2 \\
+        \midrule
+        TW & 09:00 & 10:00 & 12:00 & 09:15 & 07:00 & 07:00 & 10:30 & 10:00 & 07:00 & 07:00 & 07:25 & 07:25 & 07:25 & 07:25 \\ \bottomrule
+        \end{tabular}
+   }
+\end{table}
+
+\begin{figure}[htb]
+    \centering
+    \begin{subfigure}{0.5\textwidth}
+        \includegraphics[width=\linewidth]{depot_0.pdf}
+        \vspace*{-2em}
+        \caption{Scheduling of tasks at plant 0}
+        \label{fig:plant0}
+    \end{subfigure}%
+    \hfill% Horizontal space between the subfigures
+    \begin{subfigure}{0.5\textwidth}
+        \includegraphics[width=\linewidth]{depot_1.pdf}
+        \vspace*{-2em}
+        \caption{Scheduling of tasks at plant  1}
+        \label{fig:plant1}
+    \end{subfigure}
+    \vspace*{-1mm}
+    \begin{subfigure}{\textwidth}
+        \includegraphics[width=\linewidth]{depot_2.pdf}
+        \vspace*{-4em}
+        \caption{Scheduling of tasks at plant  2}
+        \label{fig:plant2}
+    \end{subfigure}
+
+    \caption{Loading and unloading schedules for the plants of instance $C29\_10\_14\_3$}
+    \label{fig:combined}
+\end{figure}
 We tried to reduce the maximum delay of the first delivery but found that some orders couldn't be fulfilled because of driver shifts and overtime constraints.

### 2023-08-20 Article/article.tex
save article

--- 2023-08-15 Article/article.tex
@@
 \usepackage{breqn}
+
 \tikzset{
@@
 }
+\newcolumntype{H}{>{\setbox0=\hbox\bgroup}c<{\egroup}@{}}
 
@@
         {\small
-        \begin{algorithm}[hpt]
+            \begin{algorithm}[hbpt]
                 \caption{Pseudo-code of the GRASP algorithm }
@@
 
-            $counter \leftarrow  0$
+                % $counter \leftarrow  0$
 
@@
 
-                $T \leftarrow T \cup \{S\}$
+                    % $T \leftarrow T \cup \{S\}$
 
@@
 
-                \uIf{$counter \text{ } mod \text{ } \phi == 0$}{
-                    $S^* \leftarrow PathRelinking(T)$
-
-                    $T \leftarrow \emptyset$
-                }
-
-                $counter \leftarrow counter + 1$
+                    % \uIf{$counter \text{ } mod \text{ } \phi == 0$}{
+                    %     $S^* \leftarrow PathRelinking(T)$
+
+                    %     $T \leftarrow \emptyset$
+                    % }
+
+                    % $counter \leftarrow counter + 1$
 
@@
             \midrule
-        Small  &     8 &    226.5-937.5 &     7-14 &    4-11 &    13-35 &    1-3 \\
-        Medium &    14 &  1,160.5-2,971.0 &    43-98 &   40-89 &   76-137 &    6-8 \\
-        Large  &    14 &  3,078.5-3,953.5 &  107-136 &  92-127 &  129-150 &    8 \\
+            Small  & 8  & 226.5 - 937.5     & 7 - 14    & 4 - 11   & 13-35    & 1-3     \\
+            Medium & 14 & 1,160.5 - 2,971.0 & 43 - 98   & 40 - 89  & 76-137   & 6-8     \\
+            Large  & 14 & 3,078. - 3,953.5  & 107 - 136 & 92 - 127 & 129-150  & 8       \\
             \bottomrule
@@
                                    & Opt & \multicolumn{2}{c}{40/64}  & \multicolumn{2}{c}{37/64} & \multicolumn{2}{c}{30/64}        & \multicolumn{2}{c}{35/64}  & \multicolumn{2}{c}{35/64}                                                            \\ \cmidrule(l){3-10} \cmidrule(l){11-12}
-     \multirow{2}{*}{Set B} & Avg & 12.1 & 357,313 & - & - & 16.3 & 1,331 & - & - & 13.8 & 574,983\\ \cmidrule(l){2-10} \cmidrule(l){11-12} 
+            \multirow{2}{*}{Set B} & Avg & 12.1                       & 357,313                   & -                                & -                          & 16.3                       & 1,331     & -        & -         & 13.8     & 574,983   \\ \cmidrule(l){2-10} \cmidrule(l){11-12}
                                    & Opt & \multicolumn{2}{c}{55/128} & \multicolumn{2}{c}{-}     & \multicolumn{2}{c}{40/128}       & \multicolumn{2}{c}{-}      & \multicolumn{2}{c}{44/128}                                                           \\ \bottomrule
@@
 The computational results of our instances solved with the GRASP heuristic are summarized in the tables below.
-The stopping criterion is set to 20,000 iterations.
+%The stopping criteria are set to 20,000 iterations or 3600 seconds of running time (unless otherwise specified).
 The parameters used in our experiments are listed in Table~\ref*{tab:problem_parameters}.
 
-\begin{table}[htpb]
+\begin{table}[hbt]
     \centering
@@
     \label{tab:problem_parameters}
+    \small
     %    \resizebox*{\textwidth}{!}{
@@
         Max overtime duration ($O_T$)                                                 & 120 min                   \\
-        Max delay between two consecutive deliveries of the same order ($\gamma^1$) & 20 min \\
+        Max delay between two consecutive deliveries of the same order $\gamma^1$)   & 20 min                    \\
         Max delay between two consecutive deliveries of different orders ($\gamma^2)$ & 25 min                    \\
-    %    $\beta_1$ & 10000  \\
-    %    $\beta_2$ & 10000  \\
-    %    $\beta_3$ & 1000  \\
-    %    $\beta_4$ & 100  \\
+        Penalty of unfulfilled orders ($\beta_1$)                                     & 10,0000                   \\
+        Penalty of first delivery delay ($\beta_2$)                                   & 1,0000                    \\
+        Penalty for driver underutilization  ($\beta_3$)                              & 30                        \\
+        Driver overtime penalty $ (\beta_4$)                                          & 20                        \\
         \bottomrule
@@
 \end{table}
-
-Table \ref*{tab:runtime} shows the percentage of concrete delivered by varying the execution times of the algorithm.
-For small instances, all requests are delivered in less than five minutes, while for medium and large instances, over 99\% of requests are delivered.
+%  the travel cost ($TC$), the penalties associated with unfulfilled orders ($PUO$), the sum of the first delivery delays ($FDD$) for each customer, the sum of the driver underutilization costs ($DUC$) and driver overtime costs ($DOC$).
+The table \ref*{tab:all_instances} contains the detailed results for all instances with a stop criterion of 3600 seconds.
+The table shows the demand of each instance, the average and the best values of the objective function when solving each instance five times.
+All daily orders are served by our algorithm within these five replications.
+On average, five instances require multiple runs of the algorithm to successfully serve all customers, as shown in column $UQ$, which is the undelivered quantity.
+The sum of the driver underutilization costs ($DUC$) is high for medium and large instances.
+This indicates the underutilization of the scheduled fleet.
+Some drivers are not scheduled at all or work very little.
+The sum of driver overtime costs ($DOC$) is also high for these instances.
+This can be explained by the fact that the algorithm prefers to put certain drivers on overtime rather than use a driver who has to travel to another plant for a delivery, which would increase the travel costs ($TC$).
+
+$FDD$ is the sum of the delays of the first deliveries.
+For the small instances, all but one of the first deliveries are on time.
+For the other instances, there is at least one customer with a late first delivery.
+To get an idea of the actual delay per customer, we report $mFDD$, which is the maximum delay of all first deliveries.
+This tells us that we have a maximum delay that varies between 0 and 19 minutes for small instances, 3.5 and 60 minutes for medium instances, and 24 to 60 minutes for large instances.
+
+\begin{table}[!ht]
+    \centering
+    \caption{Results with all instances}
+    \label{tab:all_instances}
+    \resizebox{\textwidth}{!}{
+    \begin{tabular}{llccccccccccccc}
+    \toprule
+    &                 &          & \multicolumn{6}{c}{Avg} & \multicolumn{6}{c}{Best}   \\
+    \cmidrule(l){4-9} \cmidrule(l){10-15}
+    &  & Demand   & $UQ$   & $TC$  & $FDD$    & $DUC$    & $DOC$    & \multirow[c]{2}{*}{Z} & $TC$      & $FDD$    & $mFDD$  & $DUC$    & $DOC$    & \multirow[c]{2}{*}{Z} \\
+        & & ($m^3$)  & ($m^3$) & (min)  & (min)    & (min)    & (min)    &     & (min)     & (min)    & (min) & (min)   &(min)& \\
+        \midrule
+        \multirow[c]{8}{*}{Small} & C_13_11_12_1 & 226.5 & 0.0 & 1,243.5 & 0.0 & 37.7 & 0.0 & 2,373.9 & 1,243.5 & 0.0 & 0.0 & 0.0 & 0.0 & 1,243.5 \\
+        & C_13_5_8_1 & 267.0 & 0.0 & 745.1 & 0.0 & 0.0 & 0.0 & 745.1 & 745.1 & 0.0 & 0.0 & 0.0 & 0.0 & 745.1 \\
+        & C_18_6_11_2 & 333.5 & 0.0 & 753.3 & 0.0 & 194.7 & 0.0 & 6,593.7 & 753.3 & 0.0 & 0.0 & 125.3 & 0.0 & 4,513.2 \\
+        & C_15_4_7_2 & 375.0 & 0.0 & 1,014.9 & 0.0 & 132.8 & 291.6 & 10,830.8 & 1,014.2 & 0.0 & 0.0 & 172.1 & 76.1 & 7,700.1 \\
+        & C_19_7_8_2 & 388.0 & 0.0 & 1,765.5 & 0.0 & 314.8 & 150.9 & 14,225.6 & 1,745.8 & 0.0 & 0.0 & 224.5 & 87.9 & 10,237.1 \\
+        & C_29_10_14_3 & 613.5 & 0.0 & 3,742.0 & 48.6 & 211.0 & 735.6 & 73,392.7 & 3,628.8 & 49.2 & 18.8 & 112.9 & 547.2 & 67,145.2 \\
+        & C_31_8_11_3 & 776.0 & 0.0 & 2,330.6 & 0.0 & 781.6 & 656.5 & 38,909.6 & 2,297.2 & 0.0 & 0.0 & 606.7 & 567.0 & 31,836.1 \\
+        & C_35_9_10_3 & 937.5 & 0.0 & 4,175.3 & 0.0 & 341.8 & 1,329.5 & 41,021.0 & 4,174.4 & 0.0 & 0.0 & 221.7 & 1,161.6 & 34,057.0 \\
+        \cmidrule(l){1-15}
+        \multirow[c]{14}{*}{Medium} & C_76_40_43_6 & 1,160.5 & 0.0 & 6,311.6 & 17.2 & 3,014.6 & 706.0 & 128,106.4 & 5,967.0 & 6.2 & 3.5 & 2,961.6 & 644.4 & 113,884.7 \\
+        & C_104_42_47_8 & 1,565.0 & 0.0 & 9,358.2 & 34.1 & 3,890.1 & 679.0 & 173,754.5 & 9,777.5 & 32.0 & 13.0 & 2,830.0 & 240.6 & 131,507.6 \\
+        & C_94_63_70_7 & 1,746.5 & 0.0 & 14,369.1 & 90.2 & 1,353.9 & 1,391.4 & 173,007.0 & 14,428.7 & 78.9 & 19.3 & 1,147.9 & 1,464.3 & 157,038.1 \\
+        & C_116_67_78_8 & 1,839.5 & 0.0 & 13,602.9 & 67.9 & 2,419.4 & 857.3 & 171,229.1 & 13,871.9 & 67.2 & 33.2 & 2,347.7 & 523.3 & 162,021.5 \\
+        & C_116_71_82_8 & 2,060.0 & 0.0 & 12,678.2 & 158.6 & 1,991.3 & 1,125.8 & 253,575.1 & 12,089.7 & 147.7 & 27.2 & 1,966.0 & 922.7 & 237,183.6 \\
+        & C_117_57_70_6 & 2,327.5 & 0.0 & 15,495.0 & 142.4 & 1,388.6 & 1,680.5 & 233,181.4 & 15,153.3 & 114.2 & 22.2 & 1,823.7 & 1,671.5 & 217,448.3 \\
+        & C_137_79_83_7 & 2,425.0 & 0.0 & 18,960.4 & 375.8 & 2,124.7 & 1,857.9 & 495,614.5 & 18,360.3 & 335.1 & 34.2 & 2,552.7 & 1,794.2 & 465,977.3 \\
+        & C_127_66_80_8 & 2,512.5 & 0.0 & 16,033.3 & 10.4 & 1,579.6 & 1,552.1 & 104,879.1 & 17,599.1 & 14.4 & 5.0 & 1,046.3 & 1,242.3 & 88,203.5 \\
+        & C_128_68_74_7 & 2,595.0 & 0.0 & 19,774.4 & 361.6 & 944.6 & 2,616.7 & 461,998.2 & 18,728.8 & 191.0 & 42.5 & 1,418.8 & 2,841.2 & 309,116.8 \\
+        & C_136_85_97_8 & 2,673.0 & 0.0 & 18,223.6 & 188.5 & 2,134.5 & 2,352.2 & 317,811.4 & 18,894.6 & 137.4 & 48.6 & 1,155.6 & 2,137.0 & 233,669.0 \\
+        & C_128_78_85_8 & 2,685.5 & 0.0 & 18,463.9 & 305.8 & 1,277.3 & 3,157.0 & 425,759.3 & 18,041.6 & 192.4 & 35.7 & 1,374.3 & 3,234.2 & 316,314.1 \\
+        & C_131_77_85_8 & 2,893.5 & 0.0 & 18,003.0 & 212.6 & 1,248.7 & 1,716.1 & 302,370.4 & 18,063.0 & 200.2 & 43.2 & 780.7 & 1,539.1 & 272,509.6 \\
+        & C_137_89_97_7 & 2,939.5 & 0.0 & 23,293.0 & 636.0 & 849.9 & 3,523.9 & 755,280.5 & 22,038.1 & 438.3 & 58.6 & 849.8 & 3,461.1 & 555,029.1 \\
+        & C_133_84_98_7 & 2,971.0 & 18.7 & 22,000.0 & 1,157.6 & 805.4 & 2,291.4 & 1,249,585.1 & 20,992.6 & 1,061.2 & 59.4 & 546.1 & 2,742.5 & 1,153,396.1 \\
+        \cmidrule(l){1-15}
+        \multirow[c]{14}{*}{Large} & C_132_98_109_8 & 3,078.5 & 0.0 & 23,469.1 & 247.4 & 548.7 & 4,162.3 & 370,622.5 & 23,425.8 & 160.9 & 29.4 & 493.2 & 4,426.1 & 287,683.7 \\
+        & C_129_101_119_8 & 3,229.0 & 0.0 & 19,051.7 & 269.0 & 1,232.1 & 3,820.4 & 401,387.9 & 19,453.3 & 166.7 & 35.8 & 1,138.1 & 4,719.6 & 314,728.3 \\
+        & C_141_114_136_8 & 3,350.5 & 8.7 & 21,505.6 & 1,127.2 & 1,422.7 & 2,339.2 & 1,238,202.0 & 21,863.2 & 943.9 & 55.4 & 1,177.0 & 2,513.4 & 1,051,386.9 \\
+        & C_140_114_132_8 & 3,401.5 & 0.0 & 23,938.4 & 1,055.1 & 910.2 & 3,269.1 & 1,171,715.1 & 23,969.7 & 988.5 & 54.4 & 981.3 & 3,253.1 & 1,107,019.4 \\
+        & C_143_101_123_8 & 3,437.5 & 0.0 & 25,711.0 & 352.2 & 903.0 & 4,085.1 & 486,755.0 & 26,284.9 & 335.5 & 43.5 & 566.1 & 3,913.0 & 457,007.0 \\
+        & C_137_112_129_8 & 3,471.0 & 0.0 & 22,562.4 & 340.8 & 402.4 & 3,371.1 & 442,869.1 & 22,338.3 & 275.1 & 45.6 & 375.2 & 3,819.8 & 385,076.1 \\
+        & C_142_114_129_8 & 3,499.5 & 0.0 & 22,478.3 & 801.0 & 1,294.8 & 4,168.2 & 945,633.4 & 22,387.6 & 654.6 & 55.3 & 1,090.4 & 4,472.6 & 799,184.0 \\
+        & C_149_98_122_8 & 3,513.0 & 3.0 & 20,767.5 & 778.4 & 2,236.9 & 2,852.2 & 923,357.9 & 21,224.7 & 684.5 & 51.1 & 2,267.6 & 3,046.6 & 834,708.6 \\
+        & C_139_98_108_8 & 3,541.0 & 0.0 & 21,613.4 & 661.0 & 852.3 & 3,716.2 & 782,546.8 & 22,043.2 & 522.3 & 51.6 & 482.8 & 3,648.5 & 631,819.8 \\
+        & C_144_108_122_8 & 3,670.5 & 0.0 & 24,738.3 & 746.0 & 664.9 & 3,465.8 & 860,025.7 & 25,806.9 & 675.4 & 50.4 & 548.5 & 3,209.2 & 781,866.1 \\
+        & C_142_92_107_8 & 3,684.5 & 0.0 & 24,553.3 & 298.6 & 644.2 & 4,423.6 & 431,001.0 & 23,586.4 & 182.6 & 24.0 & 803.0 & 4,396.5 & 318,243.2 \\
+        & C_138_114_136_8 & 3,739.5 & 0.0 & 25,366.9 & 791.5 & 668.6 & 5,501.8 & 946,993.9 & 23,639.0 & 600.4 & 55.2 & 1,230.8 & 5,314.0 & 767,223.0 \\
+        & C_150_127_136_8 & 3,946.5 & 0.0 & 25,785.2 & 613.8 & 629.1 & 4,686.4 & 752,189.6 & 27,332.9 & 507.3 & 57.1 & 250.3 & 5,295.7 & 648,057.4 \\
+        & C_148_112_131_8 & 3,953.5 & 6.9 & 26,231.2 & 1,002.0 & 517.8 & 3,467.4 & 1,113,149.3 & 25,559.2 & 886.9 & 56.0 & 559.3 & 4,213.2 & 1,013,500.2 \\
+        \bottomrule
+    \end{tabular}
+    }
+\end{table}
+
+Table \ref{tab:runtime} shows the percentage of concrete delivered by varying the execution times of the algorithm.
+For small instances, all requests are delivered in less than a minute, while for medium and large instances, over 99\% of requests are delivered.
 For these instances, it takes more than ten minutes to ensure complete delivery of all orders.
@@
     \label{tab:runtime}
-    \resizebox*{0.5\textwidth}{!}{
-        \begin{tabular}{llcccc}
+    \scriptsize
+    \resizebox*{0.7\textwidth}{!}{
+        \begin{tabular}{lcccccc}
             \toprule
-            &  &      \multicolumn{4}{c}{Runtime (min)} \\
-            \cmidrule{3-6}
-            & &      5   & 10  &  30  & 60  \\
-            %  & {} &         &         &        &        \\
+                                    &        & \multicolumn{5}{c}{Runtime (min)}                                   \\
+            \cmidrule{3-7}
+                                    &        & 1                                 & 5      & 10     & 30    & 60    \\
             \midrule
-            \multirow{1}{*}{Small} & \% Load &  100.00 &  100.00 &  100.0 &  100.0 \\
-                %   & \# instances &    8 &    8 &    8 &    8 \\
-            \cmidrule{2-6}
-                  \multirow{1}{*}{Medium} & \% Load &   99.86 &   99.86 &  100.0 &  100.0 \\
-                %   & \# instances &   13 &   13 &   14 &   14 \\
-                  \cmidrule{2-6}
-                  \multirow{1}{*}{Large} & \% Load &   99.79 &   99.93 &  100.0 &  100.0 \\
-                %   & \# instances &   12 &   13 &   14 &   14 \\
+            \multirow{1}{*}{Small}  & \%Load & 100.00                            & 100.00 & 100.00 & 100.0 & 100.0 \\
+            %   & \#Instance &  8 &    8 &    8 &    8 & 8 \\
+            \cmidrule{2-7}
+            \multirow{1}{*}{Medium} & \%Load & 99.71                             & 99.93  & 99.86  & 100.0 & 100.0 \\
+            %   & \#Instance &   11 &   13 &   13 &   14 &   14 \\
+            \cmidrule{2-7}
+            \multirow{1}{*}{Large}  & \%Load & 99.64                             & 99.64  & 99.79  & 100.0 & 100.0 \\
+            %   & \#Instance &   10 &   10 &   12 &   14 &   14 \\
             \bottomrule
@@
 
-Table \ref*{tab:results_grasp_ls} shows the detailed results for each instance.
-The runtime is set to 3600 seconds.
-The table reports the requirements of each instance, the components of the objective function, the number of drivers used, and the execution time in seconds.
-%  the travel cost ($TC$), the penalties associated with unfulfilled orders ($PUO$), the sum of the first delivery delays ($FDD$) for each customer, the sum of the driver underutilization costs ($DUC$) and driver overtime costs ($DOC$).
-
-All daily orders are served with our algorithm, using less drivers than the available fleet.
-For the small size instances, all first deliveries are timely, except for one instance.
-The sum of the driver underutilization costs ($DUC$) is high for medium and large instances.
-This indicates underutilization of the scheduled fleet.
-Some drivers are not scheduled at all or work very little.
-The sum of the driver overtime costs ($DOC$) is also high for these instances.
-This can be explained by the fact that the algorithm prefers to put certain drivers on overtime rather than use a driver who has to travel to another plant for a delivery, which would increase the travel cost ($TC$).
-
-$FDD$ is the sum of the delays of the first deliveries.
-To get an idea of the actuel delay per customer, we report $mFDD$, which is the maximum delay for all first deliveries.
-This tells us that we have a maximum delay that varies between 0 and 12 minutes for small instances, 1 and 60 minutes for medium instances, and 50 to 60 minutes for large instances.
-
-\begin{table}[htbp]
-    \caption{Results with all instances}
-    \label{tab:results_grasp_ls}
-    \resizebox*{\textwidth}{!}{
-        \begin{tabular}{llccccccccc}
+Figure \ref*{fig:plants_schedules} illustrates the Gantt diagrams showing the loading and unloading operations for the orders assigned to each plant in instance $C29\_10\_14\_3$.
+This instance has three plants, $29$ drivers, and two customers wich have multiple orders.
+Customer $8$ has two orders ($o_8$,$o_9$) assigned to plant $0$, and customer $9$ has four orders ($o_{10}$ - - $o_{13}$) assigned to plant $2$.
+All orders requirements are in Table~\ref*{tab:instance_detail}.
+
+\begin{table}[!h]
+    \centering
+    \caption{Informations of instance $C29\_10\_14\_3$}
+    \label{tab:instance_detail}
+    \resizebox*{1\textwidth}{!}{
+        \begin{tabular}{@{}c|cccccccccccccc@{}}
             \toprule
-        &   &  Demand &  $TC$ & $PUO$&  $FDD$ & $mFDD$ &  $DUC$ &  $DOC$ &  \#Drivers &  Time \\
-        & \multicolumn{1}{c}{Instance} &  (m$^3$)  &(min)  &  & (min) &(min) &(min)    & (min) & & (s)\\
+            Order    & 0     & 1     & 2     & 3     & 4     & 5     & 6     & 7     & 8     & 9     & 10    & 11    & 12    & 13    \\ \midrule
+            Customer & 0     & 1     & 2     & 3     & 4     & 5     & 6     & 7     & 8     & 8     & 9     & 9     & 9     & 9     \\
             \midrule
-            \multirow{9}{*}{Small} & C\_13\_11\_12\_1 &          226.5 &     1243.55 &  0.0 &               0.00 &              0.00 &          66.01 &          0.00 &          13 &  3022.06 \\
-                  & C\_13\_5\_8\_1 &          267.0 &      745.10 &  0.0 &               0.00 &              0.00 &           0.00 &          0.00 &          13 &  3027.39 \\
-                  & C\_18\_6\_11\_2 &          333.5 &      753.26 &  0.0 &               0.00 &              0.00 &         653.78 &          0.00 &          17 &  3004.51 \\
-                  & C\_15\_4\_7\_2 &          375.0 &     1013.11 &  0.0 &               0.00 &              0.00 &          56.79 &        244.61 &          15 &  3009.54 \\
-                  & C\_19\_7\_8\_2 &          388.0 &     1714.74 &  0.0 &               0.00 &              0.00 &          97.56 &        283.83 &          19 &  3020.89 \\
-                  & C\_29\_10\_14\_3 &          613.5 &     3724.29 &  0.0 &              12.41 &             12.41 &         246.04 &        454.51 &          28 &  3035.23 \\
-                  & C\_31\_8\_11\_3 &          776.0 &     2297.20 &  0.0 &               0.00 &              0.00 &         627.37 &        539.96 &          30 &  3033.85 \\
-                  & C\_35\_9\_10\_3 &          937.5 &     4077.20 &  0.0 &               0.00 &              0.00 &         720.00 &        866.98 &          31 &  3020.75 \\
-                  \cmidrule{2-11}
-                  \multirow{14}{*}{Medium} & C\_76\_40\_43\_6 &         1160.5 &     5715.53 &  0.0 &               1.00 &              1.00 &        3566.49 &        771.23 &          69 &  3021.92 \\
-                  & C\_104\_42\_47\_8 &         1565.0 &     9114.62 &  0.0 &              23.00 &             16.24 &        4180.95 &        671.94 &          94 &  3047.99 \\
-                  & C\_94\_63\_70\_7 &         1746.5 &    13888.90 &  0.0 &              19.86 &             12.40 &        1086.60 &       1477.76 &          94 &  3025.72 \\
-                  & C\_116\_67\_78\_8 &         1839.5 &    13946.30 &  0.0 &              44.68 &              9.98 &        1578.90 &        639.28 &         116 &  3007.88 \\
-                  & C\_116\_71\_82\_8 &         2060.0 &    12461.20 &  0.0 &              68.38 &             20.06 &        3119.66 &       1478.63 &         107 &  3028.24 \\
-                  & C\_117\_57\_70\_6 &         2327.5 &    15066.80 &  0.0 &              83.79 &             16.44 &        1522.11 &       1892.64 &         117 &  3020.70 \\
-                  & C\_137\_79\_83\_7 &         2425.0 &    18811.40 &  0.0 &             252.94 &             41.65 &        2193.63 &       1963.05 &         136 &  3025.83 \\
-                  & C\_127\_66\_80\_8 &         2512.5 &    15177.80 &  0.0 &               2.00 &              2.00 &        1874.83 &        965.65 &         125 &  3007.70 \\
-                  & C\_128\_68\_74\_7 &         2595.0 &    19133.50 &  0.0 &             251.77 &             54.40 &        1523.19 &       2924.28 &         128 &  3053.96 \\
-                  & C\_136\_85\_97\_8 &         2673.0 &    16832.10 &  0.0 &             154.97 &             50.21 &        3507.40 &       1958.58 &         132 &  3028.15 \\
-                  & C\_128\_78\_85\_8 &         2685.5 &    16927.30 &  0.0 &             202.26 &             50.12 &        1492.50 &       2794.21 &         128 &  3030.32 \\
-                  & C\_131\_77\_85\_8 &         2893.5 &    17572.10 &  0.0 &             187.42 &             49.24 &        1803.22 &       1306.94 &         130 &  3040.74 \\
-                  & C\_137\_89\_97\_7 &         2939.5 &    21934.10 &  0.0 &             697.18 &             59.10 &        1220.15 &       3038.33 &         137 &  3025.45 \\
-                  & C\_133\_84\_98\_7 &         2971.0 &    22399.30 &  0.0 &             867.12 &             59.44 &         550.77 &       2887.51 &         133 &  3137.11 \\
-                  \cmidrule{2-11}
-                  \multirow{14}{*}{Large} & C\_132\_98\_109\_8 &         3078.5 &    23102.40 &  0.0 &             193.01 &             49.44 &         450.57 &       3931.83 &         132 &  3018.05 \\
-                  & C\_129\_101\_119\_8 &         3229.0 &    18978.90 &  0.0 &             467.94 &             59.47 &        1084.78 &       2334.16 &         129 &  3179.81 \\
-                  & C\_141\_114\_136\_8 &         3350.5 &    21760.00 &  0.0 &            1131.39 &             59.58 &        1284.13 &       2360.19 &         141 &  3056.56 \\
-                  & C\_140\_114\_132\_8 &         3401.5 &    23259.00 &  0.0 &            1092.82 &             57.99 &         634.53 &       2705.26 &         140 &  3046.75 \\
-                  & C\_143\_101\_123\_8 &         3437.5 &    24560.70 &  0.0 &             451.36 &             56.61 &        1663.43 &       4359.64 &         142 &  3039.02 \\
-                  & C\_137\_112\_129\_8 &         3471.0 &    21526.10 &  0.0 &             252.62 &             58.09 &         624.87 &       3360.64 &         137 &  3018.76 \\
-                  & C\_142\_114\_129\_8 &         3499.5 &    21991.50 &  0.0 &             898.10 &             55.22 &        1750.60 &       4943.03 &         142 &  3011.54 \\
-                  & C\_149\_98\_122\_8 &         3513.0 &    20696.60 &  0.0 &             839.56 &             52.60 &        2465.87 &       2892.55 &         149 &  3072.58 \\
-                  & C\_139\_98\_108\_8 &         3541.0 &    20550.10 &  0.0 &             566.69 &             57.49 &        1438.11 &       2910.97 &         139 &  3011.84 \\
-                  & C\_144\_108\_122\_8 &         3670.5 &    23401.60 &  0.0 &             668.15 &             59.77 &        1438.60 &       3569.62 &         143 &  3056.51 \\
-                  & C\_142\_92\_107\_8 &         3684.5 &    22700.20 &  0.0 &             147.42 &             48.31 &         899.82 &       3702.01 &         142 &  3019.52 \\
-                  & C\_138\_114\_136\_8 &         3739.5 &    24670.70 &  0.0 &             750.04 &             55.88 &         518.63 &       5156.41 &         137 &  3021.26 \\
-                  & C\_150\_127\_136\_8 &         3946.5 &    25822.20 &  0.0 &             742.75 &             57.10 &         899.92 &       5101.58 &         150 &  3060.79 \\
-                  & C\_148\_112\_131\_8 &         3953.5 &    25455.30 &  0.0 &             868.81 &             49.15 &         500.68 &       3775.57 &         148 &  3005.96 \\
-            \bottomrule
+            Demand   & 3.0   & 6.0   & 6.5   & 9.5   & 10.0  & 15.0  & 17.0  & 118.0 & 1.0   & 210.5 & 1.0   & 3.0   & 19.0  & 194.0 \\
+            \midrule
+            Plant    & 1     & 1     & 1     & 1     & 1     & 1     & 1     & 2     & 0     & 0     & 2     & 2     & 2     & 2     \\
+            \midrule
+            TW       & 09:00 & 10:00 & 12:00 & 09:15 & 07:00 & 07:00 & 10:30 & 10:00 & 07:00 & 07:00 & 07:25 & 07:25 & 07:25 & 07:25 \\ \bottomrule
         \end{tabular}
@@
 
+We can observe that the loading operations for each plant are non-overlapping and the unloading operations occur sequentially, one after the other, at the worksite.
+Plant $0$ is in charge of both orders of customer $8$.
+In figure~\ref*{fig:plant0}, order $o_9$ is served before $o_8$.
+Each row on the figure represents a pair of loading and unloading task for a delivery.
+Figure~\ref*{fig:plant2} shows how the loading bay is alternately assigned to orders $o_{13}$ and $o_7$ between $10:00$ and $15:30$ a.m.
+
+We report for the instance $C29\_10\_14\_3$ a $FDD$ of $49.6$ minutes.
+The gantt chart indicates that customers $4,5$ and $8$ whom need their delivery at $07:00$ a.m.
+have all a first late delivery.
+The reason for these delays is that no
+
+\begin{figure}[htb]
+    \centering
+    \begin{subfigure}{0.5\textwidth}
+        \includegraphics[width=\linewidth]{depot_0.pdf}
+        \vspace*{-2em}
+        \caption{Scheduling of tasks at plant 0}
+        \label{fig:plant0}
+    \end{subfigure}%
+    \hfill% Horizontal space between the subfigures
+    \begin{subfigure}{0.5\textwidth}
+        \includegraphics[width=\linewidth]{depot_1.pdf}
+        \vspace*{-2em}
+        \caption{Scheduling of tasks at plant  1}
+        \label{fig:plant1}
+    \end{subfigure}
+    \vspace*{-1mm}
+    \begin{subfigure}{\textwidth}
+        \includegraphics[width=\linewidth]{depot_2.pdf}
+        \vspace*{-3em}
+        \caption{Scheduling of tasks at plant  2}
+        \label{fig:plant2}
+    \end{subfigure}
+
+    \caption{Loading and unloading schedules for the plants of instance $C29\_10\_14\_3$}
+    \label{fig:plants_schedules}
+\end{figure}
 We tried to reduce the maximum delay of the first delivery but found that some orders couldn't be fulfilled because of driver shifts and overtime constraints.
@@
 
+\begin{table}[htb]
+    \centering
+    \caption{Comparaison with and without shift and overtime constraints}
+    \label{tab:lift_shift_constraints}
+    \begin{tabular}{lrrrrrrrr}
+        \toprule
+               & \multicolumn{8}{c}{D}                                                                 \\
+               & $TC$                  & $PUO$ & $FDD$  & $mFDD$ & $TC$      & $PUO$ & $FDD$  & $mFDD$ \\
+               & (min)                 & (min) & (min)  & (min)  & (min)     & (min) & (min)  & (min)  \\
+        size   &                       &       &        &        &           &       &        &        \\
+        \midrule
+        Small  & 1,971.27              & 0.00  & 6.08   & 2.68   & 1,842.47  & 0.00  & 0.00   & 0.00   \\
+        Medium & 15,922.87             & 1.34  & 228.67 & 35.27  & 14,771.08 & 0.00  & 83.84  & 19.04  \\
+        Large  & 23,424.94             & 1.33  & 621.30 & 50.33  & 22,377.07 & 0.00  & 195.39 & 33.29  \\
+        \bottomrule
+    \end{tabular}
+\end{table}
+
+\iffalse
     \begin{table}[htbp]
@@
                 \toprule
-        {} & \multicolumn{4}{c}{Default parameters} &\multicolumn{4}{c}{$M_T=0$, $N_T,O_T=\infty$, $H_k=0$} \\
+                {}     & \multicolumn{4}{c}{Default parameters} & \multicolumn{4}{c}{$M_T=0$, $N_T,O_T=\infty$, $H_k=0$}                                                          \\
                 \cmidrule(l){2-5} \cmidrule(l){6-9}
                 {}     & Load                                   & $TC$                                                   & $FDD$  & $mFDD$ & Load    & $TC$     & $FDD$  & $mFDD$ \\
-               &($m^3$)  & (min)   &(min)   &(min)  &($m^3$)   &  (min) &(min)  & (min)   \\
+                       & ($m^3$)                                & (min)                                                  & (min)  & (min)  & ($m^3$) & (min)    & (min)  & (min)  \\
                 \midrule
@@
     \end{table}
+\fi
+
+\begin{figure}[htb]
+    \centering
+    \begin{subfigure}{0.5\textwidth}
+        \includegraphics[width=\linewidth]{travel.pdf}
+        \vspace*{-2em}
+        \caption{travel cost}
+        \label{fig:travel}
+    \end{subfigure}%
+    \hfill% Horizontal space between the subfigures
+    \begin{subfigure}{0.5\textwidth}
+        \includegraphics[width=\linewidth]{FDD.pdf}
+        \vspace*{-2em}
+        \caption{first delivevery delay}
+        \label{fig:fdd}
+    \end{subfigure}
+    \vspace*{-1mm}
+
+\end{figure}
+
+\subsection{Sensitivity analyses}
+
+\subsubsection{Influence of $\gamma^1$}
 
@@
 
-\begin{figure}[htbp]
+\begin{figure}[!h]
     \centering
@@
 
+\subsubsection{Influence of the local search step}
 We checked how well the local search part of the GRASP algorithm works by running tests using only constructive heuristics.
@@
 And without local search, most objective metrics experienced a significant increase, except for travel costs.
+
+\begin{table}[htb]
+    \centering
+    \caption{Influence of the local search step on the GRASP algorithm}
+    \label{tab:grasp_ls_nols}
+    \begin{tabular}{lrrrrrrrrrrrrr}
+        \toprule
+               & \multicolumn{13}{c}{LS}                                                                                                                               \\
+               & Load                    & Served_G & $TC$      & $FDD$    & $DUC$    & $DOC$    & Time  & $TC$      & Served_L & $FDD$  & $DUC$    & $DOC$    & Time  \\
+               & (min)                   & (min)    & (min)     & (min)    & (min)    & (min)    & (min) & (min)     & (min)    & (min)  & (min)    & (min)    & (min) \\
+        size   &                         &          &           &          &          &          &       &           &          &        &          &          &       \\
+        \midrule
+        Small  & 3,917.00                & 100.00   & 2,039.77  & 8.04     & 225.15   & 221.71   & 24.22 & 1,965.93  & 100.00   & 5.94   & 250.12   & 328.33   & 60.41 \\
+        Medium & 32,394.00               & 99.82    & 14,118.33 & 358.96   & 2,242.93 & 1,328.91 & 58.32 & 15,896.96 & 100.00   & 239.98 & 1,917.64 & 1,749.73 & 60.48 \\
+        Large  & 49,516.00               & 99.40    & 23,176.19 & 1,008.66 & 839.28   & 3,814.06 & 60.20 & 23,650.53 & 100.00   & 610.78 & 827.73   & 3,797.73 & 60.68 \\
+        Avg    & nan                     & nan      & 12,830.60 & 426.42   & 1,236.43 & 1,659.86 & 48.26 & 15,494.82 & nan      & 321.84 & 1,126.03 & 2,171.76 & 60.54 \\
+        \bottomrule
+    \end{tabular}
+\end{table}
 
@@
             \toprule
-           & &\multicolumn{5}{c}{GRASP$\backslash$LS} & \multicolumn{5}{c}{GRASP}   \\
+                   &         & \multicolumn{5}{c}{GRASP$\backslash$LS} & \multicolumn{5}{c}{GRASP}                                                                                                  \\
             \cmidrule(l){2-14}
-        {size} &Load &&$TC$&   $FDD$ &  $DUC$ &  $DOC$ & Time&&  $TC$ &  $FDD$&  $DUC$ &  $DOC$&Time \\
-        {} & (m$^3$)&\%Served&(min)&   (min) &  (min) &  (min) & (min)&\%Served&  (min) &  (min)&  (min) &  (min)& (min) \\
+            {size} & Load    &                                         & $TC$                      & $FDD$   & $DUC$   & $DOC$   & Time  &          & $TC$     & $FDD$  & $DUC$   & $DOC$   & Time  \\
+            {}     & (m$^3$) & \%Served                                & (min)                     & (min)   & (min)   & (min)   & (min) & \%Served & (min)    & (min)  & (min)   & (min)   & (min) \\
             \cmidrule(l){2-2} \cmidrule(l){3-8} \cmidrule(l){9-14}
-        Small  & 3917.0& 100 &       1998.75 &   64.60 &           231.63 &          215.43 &21.24& 100&      1946.06 &    1.55 &    229.40 &          298.74 &50.34\\
-        Medium & 32394.0& 99.6&     13144.05 &   2528.33 &  2496.47 & 1086.12 &49.41&100&  13564.08 &   75.77 &   2250.66 &         1347.59&50.6 \\
-        Large  & 49516.0& 99.2& 21395.23 & 4885.11 &  1250.14 &  3543.16 & 50.20&100&     21784.95 &  471.62 & 1168.73 &  3640.04&50.74 \\
+            Small  & 3917.0  & 100                                     & 1998.75                   & 64.60   & 231.63  & 215.43  & 21.24 & 100      & 1946.06  & 1.55   & 229.40  & 298.74  & 50.34 \\
+            Medium & 32394.0 & 99.6                                    & 13144.05                  & 2528.33 & 2496.47 & 1086.12 & 49.41 & 100      & 13564.08 & 75.77  & 2250.66 & 1347.59 & 50.6  \\
+            Large  & 49516.0 & 99.2                                    & 21395.23                  & 4885.11 & 1250.14 & 3543.16 & 50.20 & 100      & 21784.95 & 471.62 & 1168.73 & 3640.04 & 50.74 \\
             \bottomrule
@@
 enlver les contraintes de temps. reduire le gamma1
+
+\begin{table}[htb]
+    \centering
+    \caption{Comparaison with partner operations}
+    \label{tab:comp_unibeton}
+    \begin{tabular}{lrrrrrrr}
+        \toprule
+                      & \multicolumn{7}{c}{$\gamma^1$}                                                                                                                               \\
+                      & Demand                         & D         & ('distanceCost', 5) & ('distanceCost', 10) & ('distanceCost', 15) & ('distanceCost', 20) & ('distanceCost', 25) \\
+                      & (min)                          & (min)     & (min)               & (min)                & (min)                & (min)                & (min)                \\
+        Instance      &                                &           &                     &                      &                      &                      &                      \\
+        \midrule
+        C_13_5_8_1    & 267.00                         & 543.05    & 523.46              & 523.46               & 523.46               & 523.46               & 523.46               \\
+        C_76_40_43_6  & 1,160.50                       & 5,244.68  & 4,917.41            & 4,726.81             & 4,662.97             & 4,644.56             & 4,704.28             \\
+        C_116_67_78_8 & 1,839.50                       & 11,664.77 & 12,220.90           & 11,084.60            & 11,019.80            & 10,856.90            & 10,760.60            \\
+        C_15_4_7_2    & 375.00                         & 789.28    & 702.88              & 702.88               & 703.46               & 702.31               & 702.88               \\
+        C_18_6_11_2   & 333.50                         & 555.13    & 471.47              & 471.47               & 471.47               & 471.47               & 471.47               \\
+        C_94_63_70_7  & 1,746.50                       & 12,011.22 & 12,776.30           & 12,201.20            & 11,895.40            & 11,469.30            & 11,440.40            \\
+        \bottomrule
+    \end{tabular}
+\end{table}
 

### 2023-08-20 Data/instances/createInstance/instances.py
Save

@@ -23,12 +23,23 @@ class Instance:
         self.instance_size = dict()
         self.dict_driver = {}
         self.depot_loc_id = {}
+        self.df_temps = None
+        self.temps = None
+        self.df_dist = None
+        self.dist = None
         #self.clean_depot()
         #self.clean_employee()
         #self.clean_tickets()
         #self.get_driver_from_ticket()
-        
     
+    def init_time(self)->None:
+        self.df_temps = pd.read_csv(f"{DATA_FOLDER}/Matrice_de_temps_all.csv",header=None)
+        self.temps = self.df_temps.to_numpy()
+
+    def init_dist(self)->None:
+        self.df_dist = pd.read_csv(f"{DATA_FOLDER}/Matrice_de_distance_all.csv",header=None)
+        self.dist = self.df_dist.to_numpy()
+
     def init_order(self)->None:
         self.df_orders = pd.read_excel(f"{DATA_FOLDER}/DORDERSTAB_BETON.xlsx", engine='openpyxl', sheet_name="data")
         self.df_orders.sort_values(by='QUANTITY',inplace=True)

### 2023-08-20 solution/cq/gantt.py

@@ -1,6 +1,7 @@
 import pandas as pd
 import numpy as np
 import plotly.express as px
+from distinctipy import distinctipy
 
 # Transform minutes to format"hh:mm:ss:"
 def time_to_date(date,t):
@@ -19,8 +20,27 @@ def time_to_date(date,t):
 
 def show(df):
     df['task'] = df.apply(lambda x: f" del {x.Del}",axis=1)
-    fig = px.timeline(df, x_start="start", x_end="end", y="Driver",title="All drivers",color="Driver",template='ggplot2')
-    fig.update_yaxes(autorange="reversed") # otherwise tasks are listed from the bottom up
+    color = [f"rgb{distinctipy.get_rgb256(x)}" for x in distinctipy.get_colors(29)]
+    fig = px.timeline(df, x_start="start", x_end="end", y="Driver",title="All drivers",
+                      color="Driver",template='ggplot2',
+                      color_continuous_scale= color)
+    # fig.update_yaxes(autorange="reversed") # otherwise tasks are listed from the bottom up
+    # fig.update_xaxes(fixedrange=True)
+    # fig.update_xaxes(
+    # tickformatstops=[
+    #     dict(dtickrange=[None, 86400000], value=" %h %m %s \n")
+    # ]
+    # )
+    
+    # fig.update_xaxes(dtick=86400000)
+    fig.update_layout(
+             xaxis=dict(
+        # type="date",           # Set x-axis type to date-based
+        tickmode="auto",       # Automatically choose tick positions
+        # tickformat="%b %d",    # Format of tick labels (Jan 01)
+        nticks=15,  
+    ),
+    )
     fig.show()
 
 # Creer des diagrammes de Gantt
@@ -75,15 +95,56 @@ def showGanttDriver(df):
 # Par ordre: afficher les heures d'arrivée et de départ des différents noeuds
 # J'ai besoin de la date, des heures d'arrivée et de départ.
 def showGanttDepot(df):
+    Prism = [
+   'rgb(0, 255, 0)', 'rgb(255, 0, 255)', 'rgb(0, 128, 255)',
+    'rgb(255, 128, 0)', 'rgb(128, 191, 128)']# 'rgb(74, 17, 157)',
+    # 'rgb(209, 2, 37)', 'rgb(74, 17, 157)', 'rgb(176, 114, 254)',
[…]

### 2023-08-20 Article/article.tex
save

@@
 Figure \ref*{fig:plants_schedules} illustrates the Gantt diagrams showing the loading and unloading operations for the orders assigned to each plant in instance $C29\_10\_14\_3$.
+It also shows the differents assignments of each driver throughout the day.
 This instance has three plants, $29$ drivers, and two customers wich have multiple orders.
@@
 Plant $0$ is in charge of both orders of customer $8$.
-In figure~\ref*{fig:plant0}, order $o_9$ is served before $o_8$.
-Each row on the figure represents a pair of loading and unloading task for a delivery.
-Figure~\ref*{fig:plant2} shows how the loading bay is alternately assigned to orders $o_{13}$ and $o_7$ between $10:00$ and $15:30$ a.m.
-
-We report for the instance $C29\_10\_14\_3$ a $FDD$ of $49.6$ minutes.
-The gantt chart indicates that customers $4,5$ and $8$ whom need their delivery at $07:00$ a.m.
+In figure~\ref*{fig:plant0}, order $o_8$ is served before $o_9$ starting at the due time.
+Each row on the figure represents the pairs of loading and unloading tasks performed by a driver.
+The number near a task references the order.
+Figure~\ref*{fig:plant2} shows how the loading bay is alternately assigned to orders $o_{13}$ and $o_7$ between $10:00$ and $14:00$ a.m.
+Driver $26$ starts its service at plant $0$, serves order $o_9$ four times and travels to plant $2$ to perform the last delivery of order $o_{13}$.
+Driver $25$ also loads at both plants $0$ and $1$.
+
+We report for this solution of instance $C29\_10\_14\_3$ a $FDD$ of $59.5$ minutes.
+The gantt chart indicates that customers $4,5$ whom need their deliveries at $07:00$ a.m.
 have all a first late delivery.
-The reason for these delays is that no
+The reason for these delays is that the drivers assigned to plant $1$ starts their earliest shift at $06:05$, and $06:20$.
+And it takes $59$($63$) minutes travel time to visit customer $4$($5$) from this plant.
+So it appears that the big value of $FDD$ we find in our results may be caused by a bad schedule of the drivers shifts.
 
@@
 \end{figure}
-We tried to reduce the maximum delay of the first delivery but found that some orders couldn't be fulfilled because of driver shifts and overtime constraints.
-We decided to remove these constraints and rerun our computation.
-We found that for small instances, the maximum delay dropped to zero, and for medium and large instances, it was reduced by almost 50\%.
+
+\subsection{Sensitivity analyses}
+
+In this section, we investigate the impact on the solution quality of some constraints and parameters of our problem.
+
+\subsubsection{Influence of the shift and overtime constraints}
+
+We tried to reduce the maximum delay of the first delivery but found that some orders couldn't be fulfilled because of the driver shifts and overtime constraints.
+We lift  these constraints and rerun our computation.
+We found that for small instances, the maximum delay dropped to zero, and for medium and large instances, it was reduced by almost 40\%.
 This suggests that when we input a good work schedule, our algorithm works effectively to ensure that all orders are filled and that delays for first deliveries and travel costs are minimized.
-In Table~\ref*{tab:lift_shift_constraints}, we can see that the average travel time decreases for all instances except the large ones.
-That's because the goals of minimizing travel cost and not delaying the first delivery are sometimes in conflict: to ensure that the first delivery is not late, drivers assigned to one plant often have to travel to another.
+We can see that the average travel time decreases for all instances (Table~\ref*{tab:lift_shift_constraints}).% That's because the goals of minimizing travel cost and not delaying the first delivery are sometimes in conflict: to ensure that the first delivery is not late, drivers assigned to one plant often have to travel to another.
 And since all plants are used in large cases, this can increase travel times.
@@
 
-\begin{table}[htb]
+\begin{table}[!h]
     \centering
     \caption{Comparaison with and without shift and overtime constraints}
+    \scriptsize
     \label{tab:lift_shift_constraints}
+    \resizebox*{0.8\textwidth}{!}{
+
     \begin{tabular}{lrrrrrrrr}
@@
                & (min)                 & (min) & (min)  & (min)  & (min)     & (min) & (min)  & (min)  \\
-        size   &                       &       &        &        &           &       &        &        \\
         \midrule
@@
     \end{tabular}
+    }
 \end{table}
@@
 \fi
-
-\begin{figure}[htb]
-    \centering
-    \begin{subfigure}{0.5\textwidth}
-        \includegraphics[width=\linewidth]{travel.pdf}
-        \vspace*{-2em}
-        \caption{travel cost}
-        \label{fig:travel}
-    \end{subfigure}%
-    \hfill% Horizontal space between the subfigures
-    \begin{subfigure}{0.5\textwidth}
-        \includegraphics[width=\linewidth]{FDD.pdf}
-        \vspace*{-2em}
-        \caption{first delivevery delay}
-        \label{fig:fdd}
-    \end{subfigure}
-    \vspace*{-1mm}
-
-\end{figure}
-
-\subsection{Sensitivity analyses}
 
@@
 
+\subsubsection{Objectives}
+
+Nous avons vérifié l'impact des objectifs de minimisation de la distance et de la première livraison.
+Ces deux objectives sont contradictoires.
+Si on fait diminuer un, l'autre augmente.
+Il faudrait trouver avec le décideur qu'il y ait un trade-off.
+
+\begin{figure}[!ht]
+    \centering
+    \begin{subfigure}{0.5\textwidth}
+        \includegraphics[width=\linewidth]{travel.pdf}
+        \vspace*{-2em}
+        \caption{travel cost}
+        \label{fig:travel}
+    \end{subfigure}%
+    \hfill% Horizontal space between the subfigures
+    \begin{subfigure}{0.5\textwidth}
+        \includegraphics[width=\linewidth]{FDD.pdf}
+        \vspace*{-2em}
+        \caption{first delivevery delay}
+        \label{fig:fdd}
+    \end{subfigure}
+    \vspace*{-1mm}
+
+\end{figure}
+
 \subsubsection{Comparisons with the real operations}
@@
 
-\begin{table}[htb]
+\begin{table}[!ht]
     \centering

### 2023-08-20 solution/cq/gantt.py

@@ -103,7 +103,7 @@ def showGanttDepot(df):
     # 'rgb(250, 81, 134)', 'rgb(255, 128, 0)'
     # ]
     depotList=np.unique(df['Depot'])
-    df['Task'] = df.apply(lambda x: f"del {x.Del}_{x.Order}_{x.Cust}_{x.Driver}",axis=1)
+    df['Task'] = df.apply(lambda x: f"Driver {x.Driver}",axis=1)
     # color = [f"rgb{distinctipy.get_rgb256(x)}" for x in distinctipy.get_colors(14)]
     # print(color)
     for dep in depotList:
@@ -132,7 +132,14 @@ def showGanttDepot(df):
             # yaxis=dict(visible=False),  # Hide the y-axis
             # xaxis=dict(type="date"),     # Use a date-based x-axis
             # title="Timeline",
+            # font=dict(
+            # family="Courier New, monospace",
+            # size=5,  # Set the font size here
+            # color="RebeccaPurple"
             # )
+            # )
+        # fig.update_traces(textposition='inside')
+        fig.update_layout(uniformtext_minsize=6)#, uniformtext_mode='hide')
         fig.update_coloraxes(showscale=False)
         # fig.update_traces(showlegend=False)
         fig.update_xaxes(

### 2023-08-21 Article/article.tex
Article with comparison between LS et noLS

@@
 If the order requires more concrete than a single truck can carry, multiple deliveries are scheduled.
+We will accept a first delivery if it arrives within a time window of $\left[a_i, a_i + \tau^w \right]$, where $\tau^w_i$ is an user-defined parameter.
 
@@
     \label{mod:c9}  & g_{i} \geq v_{d^0_{o_1}} - a_i - M\left(  \sum_{ j \in \delta^{-}_{D}(d^0_{o_1})  }{u_{jd^0_{o_1}}}\right)  & \hspace{-20mm}  \forall  i \in \mathcal{C}, \forall o_1 \in \mathcal{O}_i   \\
+    \label{mod:c91}  & g_{i} - a_i \leq \tau^w  & \hspace{-20mm}  \forall  i \in \mathcal{C}  \\
     \label{mod:c101}       & v_{d^0_{o_1}} \geq w_j - M\left(1- u_{jd^0_{o_1}}\right)   & \hspace{-40mm}    \forall o_1 \in \mathcal{O}_i,  j \in \delta_{D}^{-}(d^0_{o_1})                \\
@@
 
-Constraints (\ref{mod:c8}) - - (\ref{mod:c10}) ensure that the first service of any customer $i$ must start at the due time $a_i$.
+Constraints (\ref{mod:c8}) - - (\ref{mod:c10}) ensure that the first service of any customer $i$ must start after the due time $a_i$.
 This first service may be performed at the first delivery node of any order $\in O_i$, and one order must be completed before another is started.
@@
         \toprule
-        \multicolumn{1}{c}{Parameter}                                                 & \multicolumn{1}{c}{Value} \\ \midrule
+        \multicolumn{1}{c}{Parameter}  &
+        \multicolumn{1}{c}{Value} \\ \midrule
+ Time window ($\tau^w$)  & 60 min    \\
  Unloading duration ($\tau_c^u$)  & 2 min$/m^3$   \\
@@
 The table shows the demand of each instance, the average and the best values of the objective function when solving each instance five times.
-All daily orders are served by our algorithm within these five replications.
-On average, five instances require multiple runs of the algorithm to successfully serve all customers, as shown in column $UQ$, which is the undelivered quantity.
+All daily orders of the small instances are completely served by our algorithm within these five replications, as shown in column $UQ$, which reports the undelivered quantity.
+$99.57\%$ and $99.5\%$ of the total demands are served for the medium and large sets.
 The sum of the driver underutilization costs ($DUC$) is high for medium and large instances.
@@
 For the small instances, all but one of the first deliveries are on time.
-For the other instances, there is at least one customer with a late first delivery.
+For the other instances sets, there is at least one customer with a late first delivery.
 To get an idea of the actual delay per customer, we report $mFDD$, which is the maximum delay of all first deliveries.
-This tells us that we have a maximum delay that varies between 0 and 19 minutes for small instances, 3.5 and 60 minutes for medium instances, and 24 to 60 minutes for large instances.
+This tells us that we have a maximum delay that varies between 0 and $35.7$ minutes for small instances, $16.8$ and $60$ minutes for medium instances, and $56$ to 60 minutes for large instances.
 
@@
     \resizebox{\textwidth}{!}{
-    \begin{tabular}{llccccccccccccc}
+    \begin{tabular}{llcccccccccccccc}
     \toprule
     &                 &          & \multicolumn{6}{c}{Avg} & \multicolumn{6}{c}{Best}  \\
-    \cmidrule(l){4-9} \cmidrule(l){10-15}
-    &  & Demand   & $UQ$   & $TC$  & $FDD$    & $DUC$    & $DOC$    & \multirow[c]{2}{*}{Z} & $TC$      & $FDD$    & $mFDD$  & $DUC$    & $DOC$    & \multirow[c]{2}{*}{Z} \\
-        & & ($m^3$)  & ($m^3$) & (min)  & (min)    & (min)    & (min)    &     & (min)     & (min)    & (min) & (min)   &(min)& \\
+    \cmidrule(l){4-9} \cmidrule(l){10-16}
+     &  & Demand & $TC$ & $UQ$ & $FDD$ & $DUC$ & $DOC$ & \multirow[c]{2}{*}{Z} & $TC$ & $UQ$ & $FDD$ & $mFDD$ & $DUC$ & $DOC$ & \multirow[c]{2}{*}{Z} \\
+     &  & ($m^3$) & (min) & ($m^3$) & (min) & (min) & (min) &  & (min) & ($m^3$) & (min) & (min) & (min) & (min) &  \\
     \midrule
-        \multirow[c]{8}{*}{Small} & C_13_11_12_1 & 226.5 & 0.0 & 1,243.5 & 0.0 & 37.7 & 0.0 & 2,373.9 & 1,243.5 & 0.0 & 0.0 & 0.0 & 0.0 & 1,243.5 \\
-        & C_13_5_8_1 & 267.0 & 0.0 & 745.1 & 0.0 & 0.0 & 0.0 & 745.1 & 745.1 & 0.0 & 0.0 & 0.0 & 0.0 & 745.1 \\
-        & C_18_6_11_2 & 333.5 & 0.0 & 753.3 & 0.0 & 194.7 & 0.0 & 6,593.7 & 753.3 & 0.0 & 0.0 & 125.3 & 0.0 & 4,513.2 \\
-        & C_15_4_7_2 & 375.0 & 0.0 & 1,014.9 & 0.0 & 132.8 & 291.6 & 10,830.8 & 1,014.2 & 0.0 & 0.0 & 172.1 & 76.1 & 7,700.1 \\
-        & C_19_7_8_2 & 388.0 & 0.0 & 1,765.5 & 0.0 & 314.8 & 150.9 & 14,225.6 & 1,745.8 & 0.0 & 0.0 & 224.5 & 87.9 & 10,237.1 \\
-        & C_29_10_14_3 & 613.5 & 0.0 & 3,742.0 & 48.6 & 211.0 & 735.6 & 73,392.7 & 3,628.8 & 49.2 & 18.8 & 112.9 & 547.2 & 67,145.2 \\
-        & C_31_8_11_3 & 776.0 & 0.0 & 2,330.6 & 0.0 & 781.6 & 656.5 & 38,909.6 & 2,297.2 & 0.0 & 0.0 & 606.7 & 567.0 & 31,836.1 \\
-        & C_35_9_10_3 & 937.5 & 0.0 & 4,175.3 & 0.0 & 341.8 & 1,329.5 & 41,021.0 & 4,174.4 & 0.0 & 0.0 & 221.7 & 1,161.6 & 34,057.0 \\
-        \cmidrule(l){1-15}
-        \multirow[c]{14}{*}{Medium} & C_76_40_43_6 & 1,160.5 & 0.0 & 6,311.6 & 17.2 & 3,014.6 & 706.0 & 128,106.4 & 5,967.0 & 6.2 & 3.5 & 2,961.6 & 644.4 & 113,884.7 \\
-        & C_104_42_47_8 & 1,565.0 & 0.0 & 9,358.2 & 34.1 & 3,890.1 & 679.0 & 173,754.5 & 9,777.5 & 32.0 & 13.0 & 2,830.0 & 240.6 & 131,507.6 \\
-        & C_94_63_70_7 & 1,746.5 & 0.0 & 14,369.1 & 90.2 & 1,353.9 & 1,391.4 & 173,007.0 & 14,428.7 & 78.9 & 19.3 & 1,147.9 & 1,464.3 & 157,038.1 \\
-        & C_116_67_78_8 & 1,839.5 & 0.0 & 13,602.9 & 67.9 & 2,419.4 & 857.3 & 171,229.1 & 13,871.9 & 67.2 & 33.2 & 2,347.7 & 523.3 & 162,021.5 \\
-        & C_116_71_82_8 & 2,060.0 & 0.0 & 12,678.2 & 158.6 & 1,991.3 & 1,125.8 & 253,575.1 & 12,089.7 & 147.7 & 27.2 & 1,966.0 & 922.7 & 237,183.6 \\
-        & C_117_57_70_6 & 2,327.5 & 0.0 & 15,495.0 & 142.4 & 1,388.6 & 1,680.5 & 233,181.4 & 15,153.3 & 114.2 & 22.2 & 1,823.7 & 1,671.5 & 217,448.3 \\
-        & C_137_79_83_7 & 2,425.0 & 0.0 & 18,960.4 & 375.8 & 2,124.7 & 1,857.9 & 495,614.5 & 18,360.3 & 335.1 & 34.2 & 2,552.7 & 1,794.2 & 465,977.3 \\
-        & C_127_66_80_8 & 2,512.5 & 0.0 & 16,033.3 & 10.4 & 1,579.6 & 1,552.1 & 104,879.1 & 17,599.1 & 14.4 & 5.0 & 1,046.3 & 1,242.3 & 88,203.5 \\
-        & C_128_68_74_7 & 2,595.0 & 0.0 & 19,774.4 & 361.6 & 944.6 & 2,616.7 & 461,998.2 & 18,728.8 & 191.0 & 42.5 & 1,418.8 & 2,841.2 & 309,116.8 \\
-        & C_136_85_97_8 & 2,673.0 & 0.0 & 18,223.6 & 188.5 & 2,134.5 & 2,352.2 & 317,811.4 & 18,894.6 & 137.4 & 48.6 & 1,155.6 & 2,137.0 & 233,669.0 \\
-        & C_128_78_85_8 & 2,685.5 & 0.0 & 18,463.9 & 305.8 & 1,277.3 & 3,157.0 & 425,759.3 & 18,041.6 & 192.4 & 35.7 & 1,374.3 & 3,234.2 & 316,314.1 \\
-        & C_131_77_85_8 & 2,893.5 & 0.0 & 18,003.0 & 212.6 & 1,248.7 & 1,716.1 & 302,370.4 & 18,063.0 & 200.2 & 43.2 & 780.7 & 1,539.1 & 272,509.6 \\
-        & C_137_89_97_7 & 2,939.5 & 0.0 & 23,293.0 & 636.0 & 849.9 & 3,523.9 & 755,280.5 & 22,038.1 & 438.3 & 58.6 & 849.8 & 3,461.1 & 555,029.1 \\
-        & C_133_84_98_7 & 2,971.0 & 18.7 & 22,000.0 & 1,157.6 & 805.4 & 2,291.4 & 1,249,585.1 & 20,992.6 & 1,061.2 & 59.4 & 546.1 & 2,742.5 & 1,153,396.1 \\
-        \cmidrule(l){1-15}
-        \multirow[c]{14}{*}{Large} & C_132_98_109_8 & 3,078.5 & 0.0 & 23,469.1 & 247.4 & 548.7 & 4,162.3 & 370,622.5 & 23,425.8 & 160.9 & 29.4 & 493.2 & 4,426.1 & 287,683.7 \\
-        & C_129_101_119_8 & 3,229.0 & 0.0 & 19,051.7 & 269.0 & 1,232.1 & 3,820.4 & 401,387.9 & 19,453.3 & 166.7 & 35.8 & 1,138.1 & 4,719.6 & 314,728.3 \\
-        & C_141_114_136_8 & 3,350.5 & 8.7 & 21,505.6 & 1,127.2 & 1,422.7 & 2,339.2 & 1,238,202.0 & 21,863.2 & 943.9 & 55.4 & 1,177.0 & 2,513.4 & 1,051,386.9 \\
-        & C_140_114_132_8 & 3,401.5 & 0.0 & 23,938.4 & 1,055.1 & 910.2 & 3,269.1 & 1,171,715.1 & 23,969.7 & 988.5 & 54.4 & 981.3 & 3,253.1 & 1,107,019.4 \\
-        & C_143_101_123_8 & 3,437.5 & 0.0 & 25,711.0 & 352.2 & 903.0 & 4,085.1 & 486,755.0 & 26,284.9 & 335.5 & 43.5 & 566.1 & 3,913.0 & 457,007.0 \\
-        & C_137_112_129_8 & 3,471.0 & 0.0 & 22,562.4 & 340.8 & 402.4 & 3,371.1 & 442,869.1 & 22,338.3 & 275.1 & 45.6 & 375.2 & 3,819.8 & 385,076.1 \\
-        & C_142_114_129_8 & 3,499.5 & 0.0 & 22,478.3 & 801.0 & 1,294.8 & 4,168.2 & 945,633.4 & 22,387.6 & 654.6 & 55.3 & 1,090.4 & 4,472.6 & 799,184.0 \\
-        & C_149_98_122_8 & 3,513.0 & 3.0 & 20,767.5 & 778.4 & 2,236.9 & 2,852.2 & 923,357.9 & 21,224.7 & 684.5 & 51.1 & 2,267.6 & 3,046.6 & 834,708.6 \\
-        & C_139_98_108_8 & 3,541.0 & 0.0 & 21,613.4 & 661.0 & 852.3 & 3,716.2 & 782,546.8 & 22,043.2 & 522.3 & 51.6 & 482.8 & 3,648.5 & 631,819.8 \\
-        & C_144_108_122_8 & 3,670.5 & 0.0 & 24,738.3 & 746.0 & 664.9 & 3,465.8 & 860,025.7 & 25,806.9 & 675.4 & 50.4 & 548.5 & 3,209.2 & 781,866.1 \\
-        & C_142_92_107_8 & 3,684.5 & 0.0 & 24,553.3 & 298.6 & 644.2 & 4,423.6 & 431,001.0 & 23,586.4 & 182.6 & 24.0 & 803.0 & 4,396.5 & 318,243.2 \\
-        & C_138_114_136_8 & 3,739.5 & 0.0 & 25,366.9 & 791.5 & 668.6 & 5,501.8 & 946,993.9 & 23,639.0 & 600.4 & 55.2 & 1,230.8 & 5,314.0 & 767,223.0 \\
-        & C_150_127_136_8 & 3,946.5 & 0.0 & 25,785.2 & 613.8 & 629.1 & 4,686.4 & 752,189.6 & 27,332.9 & 507.3 & 57.1 & 250.3 & 5,295.7 & 648,057.4 \\
-        & C_148_112_131_8 & 3,953.5 & 6.9 & 26,231.2 & 1,002.0 & 517.8 & 3,467.4 & 1,113,149.3 & 25,559.2 & 886.9 & 56.0 & 559.3 & 4,213.2 & 1,013,500.2 \\
+    \multirow[c]{8}{*}{Small} & C_13_11_12_1 & 226.5 & 1,243.5 & 0.0 & 0.0 & 10.8 & 0.0 & 1,568.3 & 1,243.5 & 0.0 & 0.0 & 0.0 & 0.0 & 0.0 & 1,243.5 \\
+     & C_13_5_8_1 & 267.0 & 745.1 & 0.0 & 0.0 & 0.0 & 0.0 & 745.1 & 745.1 & 0.0 & 0.0 & 0.0 & 0.0 & 0.0 & 745.1 \\
+     & C_18_6_11_2 & 333.5 & 753.3 & 0.0 & 0.0 & 145.9 & 0.0 & 5,130.0 & 753.3 & 0.0 & 0.0 & 0.0 & 118.7 & 0.0 & 4,315.1 \\
+     & C_15_4_7_2 & 375.0 & 1,015.6 & 0.0 & 0.0 & 148.2 & 150.5 & 8,470.9 & 1,015.4 & 0.0 & 0.0 & 0.0 & 123.4 & 110.3 & 6,924.5 \\
+     & C_19_7_8_2 & 388.0 & 1,747.2 & 0.0 & 0.0 & 349.8 & 159.0 & 15,421.6 & 1,781.8 & 0.0 & 0.0 & 0.0 & 186.0 & 0.0 & 7,361.5 \\
+     & C_29_10_14_3 & 613.5 & 3,510.1 & 0.0 & 59.5 & 194.6 & 331.9 & 75,488.1 & 3,493.4 & 0.0 & 59.5 & 35.7 & 120.5 & 60.0 & 67,810.4 \\
+     & C_31_8_11_3 & 776.0 & 2,314.8 & 0.0 & 0.0 & 804.3 & 640.7 & 39,256.8 & 2,278.2 & 0.0 & 0.0 & 0.0 & 503.5 & 617.0 & 29,722.9 \\
+     & C_35_9_10_3 & 937.5 & 4,175.5 & 0.0 & 0.0 & 222.1 & 669.3 & 24,224.7 & 4,163.1 & 0.0 & 0.0 & 0.0 & 243.3 & 570.0 & 22,863.2 \\
+     \cmidrule(l){1-16}
+
+    \multirow[c]{14}{*}{Medium} & C_76_40_43_6 & 1,160.5 & 6,055.8 & 0.0 & 49.1 & 3,286.4 & 990.0 & 173,509.5 & 5,990.6 & 0.0 & 33.9 & 25.4 & 3,061.3 & 788.6 & 147,510.1 \\
+     & C_104_42_47_8 & 1,565.0 & 8,107.1 & 12.0 & 53.3 & 3,637.1 & 119.4 & 1,372,936.3 & 8,301.9 & 12.0 & 54.3 & 16.8 & 3,200.5 & 140.1 & 1,361,393.1 \\
+     & C_94_63_70_7 & 1,746.5 & 13,783.0 & 0.0 & 238.7 & 983.2 & 1,554.6 & 313,077.0 & 13,673.5 & 0.0 & 195.9 & 59.7 & 1,025.2 & 1,407.3 & 268,524.2 \\
+     & C_116_67_78_8 & 1,839.5 & 13,456.8 & 0.0 & 132.7 & 1,869.0 & 614.3 & 214,501.3 & 13,484.8 & 0.0 & 98.0 & 22.7 & 1,792.8 & 622.4 & 177,710.8 \\
+     & C_116_71_82_8 & 2,060.0 & 11,393.8 & 9.0 & 265.1 & 2,714.8 & 1,054.6 & 1,279,068.9 & 10,793.3 & 9.0 & 242.0 & 41.8 & 3,197.9 & 665.2 & 1,262,000.5 \\
+     & C_117_57_70_6 & 2,327.5 & 14,975.0 & 0.0 & 324.1 & 1,005.5 & 1,149.6 & 392,281.0 & 15,303.5 & 0.0 & 295.2 & 54.6 & 659.4 & 1,083.8 & 351,940.4 \\
+     & C_137_79_83_7 & 2,425.0 & 17,644.3 & 0.0 & 553.8 & 2,320.4 & 1,993.3 & 680,888.1 & 17,379.3 & 0.0 & 474.0 & 55.7 & 2,964.4 & 2,016.3 & 620,601.6 \\
+     & C_127_66_80_8 & 2,512.5 & 14,584.3 & 0.0 & 128.3 & 1,600.0 & 1,451.5 & 219,949.8 & 14,499.2 & 0.0 & 127.5 & 47.3 & 1,390.2 & 1,212.8 & 207,962.2 \\
+     & C_128_68_74_7 & 2,595.0 & 18,450.6 & 3.7 & 566.6 & 1,317.5 & 2,706.2 & 1,048,712.1 & 18,261.4 & 0.0 & 473.4 & 56.0 & 1,764.7 & 2,525.6 & 595,143.5 \\
+     & C_136_85_97_8 & 2,673.0 & 16,831.6 & 0.0 & 379.9 & 2,702.4 & 2,368.0 & 525,199.5 & 17,241.3 & 0.0 & 323.7 & 50.2 & 2,190.5 & 2,767.1 & 461,973.3 \\
+     & C_128_78_85_8 & 2,685.5 & 15,679.8 & 63.0 & 474.5 & 1,399.6 & 2,394.4 & 6,880,077.6 & 15,830.6 & 63.0 & 432.7 & 57.8 & 1,206.8 & 2,272.7 & 6,830,154.3 \\
+     & C_131_77_85_8 & 2,893.5 & 16,988.0 & 0.0 & 374.5 & 1,557.9 & 1,996.1 & 478,106.8 & 16,565.6 & 0.0 & 352.1 & 44.6 & 1,624.3 & 1,322.5 & 443,865.0 \\
+     & C_137_89_97_7 & 2,939.5 & 21,716.4 & 9.8 & 924.4 & 821.0 & 3,494.4 & 2,020,622.6 & 21,311.3 & 0.0 & 799.5 & 51.8 & 887.3 & 3,486.7 & 917,162.4 \\
+     & C_133_84_98_7 & 2,971.0 & 20,435.5 & 46.1 & 1,141.3 & 894.1 & 2,127.8 & 5,841,090.1 & 20,520.4 & 9.0 & 1,411.0 & 59.5 & 705.0 & 1,871.9 & 2,390,157.6 \\
+     \cmidrule(l){1-16}
+    \multirow[c]{14}{*}{Large} & C_132_98_109_8 & 3,078.5 & 22,994.6 & 0.0 & 465.5 & 496.5 & 3,804.8 & 579,471.0 & 22,712.7 & 0.0 & 351.5 & 56.2 & 803.2 & 4,011.3 & 478,549.4 \\
+     & C_129_101_119_8 & 3,229.0 & 17,323.1 & 11.0 & 657.8 & 1,577.6 & 2,375.9 & 1,869,937.5 & 17,311.8 & 11.0 & 598.8 & 58.4 & 1,507.7 & 1,948.1 & 1,800,315.9 \\
+     & C_141_114_136_8 & 3,350.5 & 19,891.6 & 39.1 & 1,255.8 & 1,366.2 & 2,251.9 & 5,271,698.5 & 19,556.2 & 18.0 & 1,366.7 & 57.9 & 1,380.0 & 2,695.1 & 3,281,568.6 \\
+     & C_140_114_132_8 & 3,401.5 & 21,917.6 & 23.8 & 1,254.1 & 680.1 & 3,389.9 & 3,744,172.5 & 22,590.2 & 17.5 & 1,248.5 & 58.7 & 766.3 & 3,267.6 & 3,109,441.2 \\
+     & C_143_101_123_8 & 3,437.5 & 23,477.2 & 0.0 & 677.7 & 955.5 & 4,369.4 & 817,244.5 & 23,806.9 & 0.0 & 617.3 & 56.0 & 1,302.2 & 5,147.9 & 783,123.1 \\
+     & C_137_112_129_8 & 3,471.0 & 22,089.2 & 0.0 & 546.6 & 544.8 & 4,187.3 & 668,737.0 & 21,195.2 & 0.0 & 456.6 & 55.0 & 883.9 & 3,532.1 & 574,935.0 \\
+     & C_142_114_129_8 & 3,499.5 & 20,678.2 & 35.7 & 1,007.5 & 1,119.7 & 4,133.1 & 4,714,436.9 & 21,243.7 & 21.0 & 907.0 & 59.8 & 1,038.0 & 4,016.6 & 3,139,750.0 \\
+     & C_149_98_122_8 & 3,513.0 & 19,859.0 & 28.4 & 1,141.7 & 2,154.6 & 2,649.9 & 4,119,224.3 & 19,602.8 & 21.0 & 859.2 & 55.7 & 2,342.9 & 2,728.2 & 3,103,676.6 \\
+     & C_139_98_108_8 & 3,541.0 & 21,060.5 & 0.0 & 966.2 & 684.9 & 3,585.4 & 1,079,477.1 & 21,084.8 & 0.0 & 867.2 & 59.8 & 834.9 & 4,133.9 & 996,046.7 \\
+     & C_144_108_122_8 & 3,670.5 & 23,330.4 & 2.0 & 1,123.4 & 536.2 & 3,251.6 & 1,427,866.8 & 22,882.8 & 0.0 & 969.7 & 57.8 & 649.7 & 3,587.1 & 1,083,850.9 \\
+     & C_142_92_107_8 & 3,684.5 & 23,179.0 & 0.0 & 578.3 & 384.1 & 4,260.4 & 698,184.4 & 23,610.6 & 0.0 & 449.2 & 59.8 & 315.7 & 4,257.2 & 567,403.7 \\
+     & C_138_114_136_8 & 3,739.5 & 23,614.6 & 11.7 & 1,225.1 & 837.4 & 5,668.6 & 2,557,166.5 & 22,859.0 & 7.5 & 1,107.0 & 58.3 & 1,224.2 & 5,078.2 & 2,018,119.1 \\
+     & C_150_127_136_8 & 3,946.5 & 24,940.3 & 0.0 & 949.5 & 775.5 & 5,384.3 & 1,105,391.3 & 24,939.2 & 0.0 & 829.0 & 57.1 & 805.9 & 6,166.7 & 1,001,490.6 \\
+     & C_148_112_131_8 & 3,953.5 & 24,286.0 & 34.0 & 1,247.8 & 657.0 & 3,009.5 & 4,752,026.9 & 24,221.7 & 24.5 & 1,118.5 & 57.1 & 859.7 & 3,483.3 & 3,688,169.1 \\
     \bottomrule
@@
 
-\begin{table}[htbp]
+\begin{table}[!ht]
     \centering
@@
     \resizebox*{0.7\textwidth}{!}{
-        \begin{tabular}{lcccccc}
+    \begin{tabular}{llrrrrr}
     \toprule
@@
     \midrule
-            \multirow{1}{*}{Small}  & \%Load & 100.00                            & 100.00 & 100.00 & 100.0 & 100.0 \\
-            %   & \#Instance &  8 &    8 &    8 &    8 & 8 \\
-            \cmidrule{2-7}
-            \multirow{1}{*}{Medium} & \%Load & 99.71                             & 99.93  & 99.86  & 100.0 & 100.0 \\
-            %   & \#Instance &   11 &   13 &   13 &   14 &   14 \\
-            \cmidrule{2-7}
-            \multirow{1}{*}{Large}  & \%Load & 99.64                             & 99.64  & 99.79  & 100.0 & 100.0 \\
-            %   & \#Instance &   10 &   10 &   12 &   14 &   14 \\
+    Small & \%Load & 100.00 & 100.00 & 100.00 & 100.00 & 100.00 \\
+    Medium & \%Load & 99.00 & 99.29 & 99.29 & 99.43 & 99.57 \\
+    Large & \%Load & 99.00 & 99.21 & 99.36 & 99.43 & 99.50 \\
     \bottomrule
@@
     }
-
 \end{table}
@@
 
-\subsection{Sensitivity analyses}
-
-In this section, we investigate the impact on the solution quality of some constraints and parameters of our problem.
-
-\subsubsection{Influence of the shift and overtime constraints}
-
-We tried to reduce the maximum delay of the first delivery but found that some orders couldn't be fulfilled because of the driver shifts and overtime constraints.
-We lift  these constraints and rerun our computation.
-We found that for small instances, the maximum delay dropped to zero, and for medium and large instances, it was reduced by almost 40\%.
+The medium instance $C\_104\_42\_47\_8$ has an unserved customer with a demand of 12 $m^3$ and a due time of $07:00$.
+Upon investigation, we find that even if we have $104$ drivers in this instance, the number of drivers starting their shift before the due time is not sufficient to serve this customer within the time window of 60 minutes that we define.
+So the value of $UQ$ we report in Table~\ref*{tab:all_instances} seems to be the optimal given the driver shift schedule given to our algorithm.
+We confirmed this observation by lifting the driver shift and overtime constraints and resolving our instances five more times.
+All orders are now completely served, and at each run.
+Table~\ref*{tab:lift_shift_constraints} compare the average values of the results of the tables \ref*{tab:all_instances} and \ref*{tab:all_instances_no_shift}.
+We found that for small instances, the maximum delay dropped to zero, and for medium (large) instances, it was reduced by almost 60\% (40\%) on average.
 This suggests that when we input a good work schedule, our algorithm works effectively to ensure that all orders are filled and that delays for first deliveries and travel costs are minimized.
-We can see that the average travel time decreases for all instances (Table~\ref*{tab:lift_shift_constraints}).% That's because the goals of minimizing travel cost and not delaying the first delivery are sometimes in conflict: to ensure that the first delivery is not late, drivers assigned to one plant often have to travel to another.
-And since all plants are used in large cases, this can increase travel times.
-
-%  Therefore, we can use our algorithm for fleet sizing or as a submodule of a driver scheduling algorithm.
+We can see that the average travel time decreases for all instances, except for the large one where the difference is not notable.
 
@@
     \centering
-    \caption{Comparaison with and without shift and overtime constraints}
-    \scriptsize
+    \small
+    \caption{Comparaison of the average results with and without shift and overtime constraints}
     \label{tab:lift_shift_constraints}
-    \resizebox*{0.8\textwidth}{!}{
-
-    \begin{tabular}{lrrrrrrrr}
+    \resizebox{0.8\textwidth}{!}{
+    \begin{tabular}{lcccccccc}
     \toprule
-               & \multicolumn{8}{c}{D}                                                                 \\
+     & \multicolumn{4}{c}{Default parameters}&\multicolumn{4}{c}{$M_T=0$, $N_T,O_T=\infty$, $H_k=0$} \\
+     \cmidrule(l){2-5} \cmidrule(l){6-9}
      & $TC$ & $PUO$ & $FDD$ & $mFDD$ & $TC$ & $PUO$ & $FDD$ & $mFDD$ \\
-               & (min)                 & (min) & (min)  & (min)  & (min)     & (min) & (min)  & (min)  \\
+     & (min) & ($m^3$) & (min) & (min) & (min) & ($m^3$) & (min) & (min) \\
     \midrule
-        Small  & 1,971.27              & 0.00  & 6.08   & 2.68   & 1,842.47  & 0.00  & 0.00   & 0.00   \\
-        Medium & 15,922.87             & 1.34  & 228.67 & 35.27  & 14,771.08 & 0.00  & 83.84  & 19.04  \\
-        Large  & 23,424.94             & 1.33  & 621.30 & 50.33  & 22,377.07 & 0.00  & 195.39 & 33.29  \\
+    Small & 1,938.13 & 0.00 & 7.44 & 4.47 & 1,842.97 & 0.00 & 0.00 & 0.00 \\
+    Medium & 15,007.28 & 10.26 & 400.45 & 46.31 & 14,879.19 & 0.00 & 74.11 & 17.87 \\
+    Large & 22,045.81 & 13.26 & 935.49 & 56.93 & 22,454.09 & 0.00 & 187.51 & 34.76 \\
     \bottomrule
@@
 
-\iffalse
-    \begin{table}[htbp]
+\begin{table}[!ht]
     \centering
-        \caption{Comparaison with and without shift and overtime constraints}
-        \label{tab:lift_shift_constraints}
-        \resizebox*{0.8\textwidth}{!}{
-            \begin{tabular}{lcccccccc}
+    \caption{Results with all instances without shift constraints}
+    \label{tab:all_instances_no_shift}
+    \resizebox{\textwidth}{!}{
+    \begin{tabular}{llcccccccccc}
     \toprule
-                {}     & \multicolumn{4}{c}{Default parameters} & \multicolumn{4}{c}{$M_T=0$, $N_T,O_T=\infty$, $H_k=0$}                                                          \\
-                \cmidrule(l){2-5} \cmidrule(l){6-9}
-                {}     & Load                                   & $TC$                                                   & $FDD$  & $mFDD$ & Load    & $TC$     & $FDD$  & $mFDD$ \\
-                       & ($m^3$)                                & (min)                                                  & (min)  & (min)  & ($m^3$) & (min)    & (min)  & (min)  \\
+     &  & & \multicolumn{4}{c}{Avg}& \multicolumn{4}{c}{Best}\\
+     \cmidrule(l){4-7} \cmidrule(l){8-12}
+     &  & Demand & $TC$ & $UQ$ & $FDD$ & \multirow[c]{2}{*}{Z} & $TC$ & UQ & $FDD$ & $mFDD$ & \multirow[c]{2}{*}{Z}  \\
+     &  & ($m^3$) & (min) & ($m^3$) & (min) &  & (min) & ($m^3$) & (min) & (min) & \\
     \midrule
-                Small  & 489.62                                 & 1946.06                                                & 1.55   & 1.55   & 489.62  & 1832.90  & 0.00   & 0.00   \\
-                Medium & 2313.86                                & 15641.50                                               & 204.03 & 31.59  & 2313.86 & 14607.66 & 74.76  & 16.56  \\
-                Large  & 3536.86                                & 22748.24                                               & 647.90 & 55.48  & 3536.86 & 22293.04 & 210.14 & 29.00  \\
+    \multirow[c]{8}{*}{Small} & C_13_11_12_1 & 226.5 & 1,243.5 & 0.0 & 0.0 & 1,243.5 & 1,243.5 & 0.0 & 0.0 & 0.0 & 1,243.5 \\
+     & C_13_5_8_1 & 267.0 & 745.1 & 0.0 & 0.0 & 745.1 & 745.1 & 0.0 & 0.0 & 0.0 & 745.1 \\
+     & C_18_6_11_2 & 333.5 & 753.3 & 0.0 & 0.0 & 753.3 & 753.3 & 0.0 & 0.0 & 0.0 & 753.3 \\
+     & C_15_4_7_2 & 375.0 & 1,013.7 & 0.0 & 0.0 & 1,013.7 & 1,013.1 & 0.0 & 0.0 & 0.0 & 1,013.1 \\
+     & C_19_7_8_2 & 388.0 & 1,661.0 & 0.0 & 0.0 & 1,661.0 & 1,658.2 & 0.0 & 0.0 & 0.0 & 1,658.2 \\
+     & C_29_10_14_3 & 613.5 & 3,235.5 & 0.0 & 0.0 & 3,235.5 & 3,198.6 & 0.0 & 0.0 & 0.0 & 3,198.6 \\
+     & C_31_8_11_3 & 776.0 & 2,257.6 & 0.0 & 0.0 & 2,257.6 & 2,219.9 & 0.0 & 0.0 & 0.0 & 2,219.9 \\
+     & C_35_9_10_3 & 937.5 & 3,834.0 & 0.0 & 0.0 & 3,834.0 & 3,808.8 & 0.0 & 0.0 & 0.0 & 3,808.8 \\
+     \cmidrule(l){1-12}
+    \multirow[c]{14}{*}{Medium} & C_76_40_43_6 & 1,160.5 & 5,262.4 & 0.0 & 0.0 & 5,262.4 & 5,187.2 & 0.0 & 0.0 & 0.0 & 5,187.2 \\
+     & C_104_42_47_8 & 1,565.0 & 8,278.0 & 0.0 & 0.0 & 8,278.0 & 8,209.6 & 0.0 & 0.0 & 0.0 & 8,209.6 \\
+     & C_94_63_70_7 & 1,746.5 & 12,730.3 & 0.0 & 0.0 & 12,730.3 & 12,551.9 & 0.0 & 0.0 & 0.0 & 12,551.9 \\
+     & C_116_67_78_8 & 1,839.5 & 11,884.4 & 0.0 & 0.0 & 11,884.4 & 11,641.4 & 0.0 & 0.0 & 0.0 & 11,641.4 \\
+     & C_116_71_82_8 & 2,060.0 & 10,750.9 & 0.0 & 0.0 & 10,750.9 & 10,697.1 & 0.0 & 0.0 & 0.0 & 10,697.1 \\
+     & C_117_57_70_6 & 2,327.5 & 14,547.0 & 0.0 & 25.8 & 40,380.4 & 14,746.8 & 0.0 & 16.1 & 9.1 & 30,816.0 \\
+     & C_137_79_83_7 & 2,425.0 & 17,822.4 & 0.0 & 120.0 & 137,840.9 & 17,982.5 & 0.0 & 92.0 & 17.6 & 110,023.4 \\
+     & C_127_66_80_8 & 2,512.5 & 13,622.2 & 0.0 & 0.0 & 13,622.2 & 13,565.0 & 0.0 & 0.0 & 0.0 & 13,565.0 \\
+     & C_128_68_74_7 & 2,595.0 & 18,641.7 & 0.0 & 74.1 & 92,744.7 & 18,554.8 & 0.0 & 43.4 & 19.0 & 61,987.5 \\
+     & C_136_85_97_8 & 2,673.0 & 17,336.3 & 0.0 & 51.8 & 69,145.9 & 17,610.5 & 0.0 & 23.1 & 16.3 & 40,731.6 \\
+     & C_128_78_85_8 & 2,685.5 & 16,781.2 & 0.0 & 55.5 & 72,294.4 & 16,658.5 & 0.0 & 39.4 & 12.5 & 56,076.4 \\
+     & C_131_77_85_8 & 2,893.5 & 17,789.1 & 0.0 & 22.0 & 39,787.5 & 16,923.5 & 0.0 & 0.0 & 0.0 & 16,923.5 \\
+     & C_137_89_97_7 & 2,939.5 & 21,831.7 & 0.0 & 214.2 & 236,037.5 & 22,630.7 & 0.0 & 157.1 & 48.5 & 179,691.7 \\
+     & C_133_84_98_7 & 2,971.0 & 21,031.1 & 0.0 & 474.0 & 495,073.6 & 21,179.3 & 0.0 & 359.8 & 59.8 & 380,980.3 \\
+     \cmidrule(l){1-12}
+     \multirow[c]{14}{*}{Large} & C_132_98_109_8 & 3,078.5 & 22,272.3 & 0.0 & 7.2 & 29,438.3 & 23,294.5 & 0.0 & 4.0 & 1.8 & 27,252.8 \\
+     & C_129_101_119_8 & 3,229.0 & 18,473.8 & 0.0 & 41.6 & 60,111.3 & 18,595.2 & 0.0 & 28.0 & 15.7 & 46,551.2 \\
+     & C_141_114_136_8 & 3,350.5 & 21,237.4 & 0.0 & 136.7 & 157,923.2 & 20,762.9 & 0.0 & 110.7 & 18.6 & 131,504.9 \\
+     & C_140_114_132_8 & 3,401.5 & 22,518.4 & 0.0 & 174.0 & 196,489.2 & 22,537.0 & 0.0 & 64.8 & 10.6 & 87,390.4 \\
+     & C_143_101_123_8 & 3,437.5 & 23,807.5 & 0.0 & 39.1 & 62,917.8 & 23,249.3 & 0.0 & 30.2 & 13.8 & 53,413.3 \\
+     & C_137_112_129_8 & 3,471.0 & 22,457.5 & 0.0 & 121.0 & 143,439.3 & 21,878.2 & 0.0 & 27.9 & 7.5 & 49,772.5 \\
+     & C_142_114_129_8 & 3,499.5 & 21,574.4 & 0.0 & 362.8 & 384,413.2 & 22,591.1 & 0.0 & 221.0 & 52.9 & 243,564.1 \\
+     & C_149_98_122_8 & 3,513.0 & 20,918.7 & 0.0 & 150.6 & 171,499.5 & 19,503.4 & 0.0 & 94.1 & 28.6 & 113,643.5 \\
+     & C_139_98_108_8 & 3,541.0 & 21,548.1 & 0.0 & 271.5 & 293,062.8 & 22,008.3 & 0.0 & 239.9 & 47.4 & 261,921.3 \\
+     & C_144_108_122_8 & 3,670.5 & 23,604.0 & 0.0 & 169.2 & 192,821.5 & 23,200.9 & 0.0 & 125.9 & 42.5 & 149,122.9 \\
+     & C_142_92_107_8 & 3,684.5 & 21,279.5 & 0.0 & 83.8 & 105,084.6 & 21,244.2 & 0.0 & 36.4 & 14.4 & 57,641.8 \\
+     & C_138_114_136_8 & 3,739.5 & 24,385.3 & 0.0 & 109.8 & 134,180.9 & 25,218.0 & 0.0 & 71.7 & 15.8 & 96,877.5 \\
+     & C_150_127_136_8 & 3,946.5 & 24,434.2 & 0.0 & 391.0 & 415,476.7 & 24,732.2 & 0.0 & 292.5 & 52.8 & 317,191.2 \\
+     & C_148_112_131_8 & 3,953.5 & 25,846.2 & 0.0 & 566.8 & 592,677.4 & 26,245.8 & 0.0 & 383.8 & 52.7 & 410,087.8 \\
     \bottomrule
     \end{tabular}
-
-        }
-
+    }
 \end{table}
-\fi
-
-\subsubsection{Influence of $\gamma^1$}
-
-The influence of the parameter $\gamma^1$ on the solution quality is demonstrated in Figure~\ref*{fig:gamma1_influence}.
-Tests were conducted by varying $\gamma^1$ values between 10 and 25 minutes, without any driver shifts or overtime constraints.
-In small instances, all customers are served regardless of the value of $\gamma^1$.
-For medium (large) instances, all customers are served with $\lambda^1$ starting at 15 (20) minutes.
-The delay of the initial delivery decreases as the value of $\gamma^1$ increases.
-The results indicate that if $\gamma^1$ is high, the plant has the flexibility to delay ongoing deliveries to make room for new ones or make customers wait until the drivers are available.
-By increasing the maximum delay between two successive deliveries from 10 to 20 minutes, the first delivery delay is reduced by over 50\% for the medium and large size instances.
-The average travel time remains constant for all instances, regardless of the value of $\gamma^1$.
+
+\subsection{Sensitivity analyses}
+
+In this section, we analyse the impact on the solution when we change some parameters of our problem, and run the GRASP algorithm with the local search step.
+
+ % That's because the goals of minimizing travel cost and not delaying the first delivery are sometimes in conflict: to ensure that the first delivery is not late, drivers assigned to one plant often have to travel to another.
+And since all plants are used in large cases, this can increase travel times.
+
+%  Therefore, we can use our algorithm for fleet sizing or as a submodule of a driver scheduling algorithm.
+
+\subsubsection{Influence of the maximum time delay $\gamma^1$}
+
+We conducted some tests with different values $\gamma^1$ varying between 5 and 25 minutes, without any shifts or overtime constraints on the drivers.
+The influence of the value of the maximum time delay between two consecutive deliveries of the same order is demonstrated in Figure~\ref*{fig:gamma1_influence}.
+For small instances, all customers are served on time regardless of the value of $\gamma^1$.
+For medium (large) instances, all customers are served with $\gamma^1$ starting at 15 (20) minutes.
+The delay of the initial delivery decreases as $\gamma^1$ increases.
+These results confirm the fact that with a high $\gamma^1$ value, plants have more flexibility to delay ongoing deliveries to make room for new ones or make customers wait until drivers are available.
+By increasing the maximum time delay from 10 to 20 minutes, the first delivery delay is reduced by over 50\% for the medium and large size instances.
+The average travel time does not significantly change regardless of the value of $\gamma^1$.
 
@@
     \includegraphics[width=0.8\textwidth]{gamma1.pdf}
-    \caption{Influence of the maximum time delay between consecutive deliveries of the same order ($\gamma^1$)}
+    \small
+    \caption{Influence of the maximum time delay ($\gamma^1$) between consecutive deliveries of the same order. }
     \label{fig:gamma1_influence}
@@
 And without local search, most objective metrics experienced a significant increase, except for travel costs.
-
-\begin{table}[htb]
-    \centering
-    \caption{Influence of the local search step on the GRASP algorithm}
-    \label{tab:grasp_ls_nols}
-    \begin{tabular}{lrrrrrrrrrrrrr}
-        \toprule
-               & \multicolumn{13}{c}{LS}                                                                                                                               \\
-               & Load                    & Served_G & $TC$      & $FDD$    & $DUC$    & $DOC$    & Time  & $TC$      & Served_L & $FDD$  & $DUC$    & $DOC$    & Time  \\
-               & (min)                   & (min)    & (min)     & (min)    & (min)    & (min)    & (min) & (min)     & (min)    & (min)  & (min)    & (min)    & (min) \\
-        size   &                         &          &           &          &          &          &       &           &          &        &          &          &       \\
-        \midrule
-        Small  & 3,917.00                & 100.00   & 2,039.77  & 8.04     & 225.15   & 221.71   & 24.22 & 1,965.93  & 100.00   & 5.94   & 250.12   & 328.33   & 60.41 \\
-        Medium & 32,394.00               & 99.82    & 14,118.33 & 358.96   & 2,242.93 & 1,328.91 & 58.32 & 15,896.96 & 100.00   & 239.98 & 1,917.64 & 1,749.73 & 60.48 \\
-        Large  & 49,516.00               & 99.40    & 23,176.19 & 1,008.66 & 839.28   & 3,814.06 & 60.20 & 23,650.53 & 100.00   & 610.78 & 827.73   & 3,797.73 & 60.68 \\
-        Avg    & nan                     & nan      & 12,830.60 & 426.42   & 1,236.43 & 1,659.86 & 48.26 & 15,494.82 & nan      & 321.84 & 1,126.03 & 2,171.76 & 60.54 \\
-        \bottomrule
-    \end{tabular}
-\end{table}
 
@@
 
-Il serait intéressant de voir l'effet des paramètres lambda 1 et lambda 2 différents par clients et /ou ordre. Également le temps de service différent par client.
+Il serait intéressant de voir l'effet des paramètres gamma 1 et gamma 2 différents par clients et /ou ordre. Également le temps de service différent par client.
 

### 2023-08-21 Article/article.tex
Article: with Comparisons with the industry partner historical  operation table

@@
 The dataset used in this section was obtained from our partner and contains records representing delivery operations from up to 8 production centers over a period of $\nbInstance$ days.
-We have created $\nbInstance$ instances, where each instance corresponds to a daily operation.
-An instance contains information such as each driver's capacity, associated batch plant, and shift start time.
-For each customer, we have the due time, the total demand, and the number of orders (type of concrete) received.
-For each order, we have its demand and associated production plant.
-Finally, we have the loading capacity for each plant.
+From this historical data, we extract information such as daily orders delivered, plant assignments to orders, and the number of available drivers.
+We created $\nbInstance$ instances from this data, where each instance corresponds to a daily operation.
+To maintain confidentiality, we have removed the GPS coordinates of customers and plants, but have included two matrices containing the distances and driving times between all plants and customers.
+An instance contains information such as each driver's capacity, associated batch plant, and shift start time (which we generate).
+For each customer, we have the due time, the total demand, the number of orders (type of concrete) received, and its index in the time (distance) matrix.
+For each order, we have its demand and the corresponding production plant.
+Finally, we have the loading capacity for each plant, and also its index in the time matrix.
 The name of an instance has the format $C\_k\_n\_o\_z$, where $k$ is the number of available drivers, $n$ is the number of customers, $o$ is the total number of daily orders, and $z$ is the number of plants.
@@
 We have summarized the dataset in Table~\ref{tab:instances}.
+The dataset is available upon request as a contribution to the CDP literature.
 
@@
 
+\subsubsection{Comparisons with the industry partner historical operations}
+
+\begin{table}[htb]
+    \centering
+    \caption{Comparaison with historical operations}
+    \label{tab:comp_unibeton}
+    \begin{tabular}{lccccccc}
+    \toprule
+     & & & \multicolumn{5}{c}{Distance} \\
+     \cmidrule(l){3-3} \cmidrule(l){4-8}
+     & Demand & Distance &  $\gamma^1=5$&  $\gamma^1=10$ & $\gamma^1=15$ & $\gamma^1=20$& $\gamma^1=25$ \\
+     & ($m^3$) & ($Km$)  &($Km$)  & ($Km$) & ($Km$) &($Km$)  &($Km$)  \\
+    \midrule
+    C_13_5_8_1 & 267.0 & 543.1 & 523.5 & 523.5 & 523.5 & 523.5 & 523.5 \\
+    C_76_40_43_6 & 1,160.5 & 5,244.7 & 5,101.2 & 4,706.1 & 4,673.3 & 4,645.6 & 4,732.4 \\
+    C_116_67_78_8 & 1,839.5 & 11,664.8 & 11,895.8 & 11,035.3 & 11,064.2 & 10,928.9 & 10,823.4 \\
+    C_15_4_7_2 & 375.0 & 789.3 & 702.3 & 702.9 & 702.3 & 703.5 & 702.3 \\
+    C_18_6_11_2 & 333.5 & 555.1 & 491.7 & 471.5 & 471.5 & 471.5 & 471.5 \\
+    C_94_63_70_7 & 1,746.5 & 12,011.2 & 12,957.1 & 11,965.5 & 11,915.6 & 11,657.1 & 11,475.7 \\
+    \bottomrule
+    \end{tabular}
+    \end{table}
+
 \subsection{Sensitivity analyses}
 
-In this section, we analyse the impact on the solution when we change some parameters of our problem, and run the GRASP algorithm with the local search step.
+In this section, we analyse the impact on the solution when we change some parameters of our problem, and how some components of the objective function are influencing each other.
 
@@
 
-\subsubsection{Influence of the local search step}
-We checked how well the local search part of the GRASP algorithm works by running tests using only constructive heuristics.
-The results are presented in Table~\ref*{tab:grasp_ls_nols}.
-For medium and large instances, the use of local search was necessary to service all customers within 60 minutes.
-And without local search, most objective metrics experienced a significant increase, except for travel costs.
-
-\begin{table}[htbp]
-    \centering
-    \caption{Influence of the local search step on the  GRASP algorithm}
-    \label{tab:grasp_ls_nols}
-    \resizebox*{1\textwidth}{!}{
-        \begin{tabular}{lccccccccccccc}
-            \toprule
-                   &         & \multicolumn{5}{c}{GRASP$\backslash$LS} & \multicolumn{5}{c}{GRASP}                                                                                                  \\
-            \cmidrule(l){2-14}
-            {size} & Load    &                                         & $TC$                      & $FDD$   & $DUC$   & $DOC$   & Time  &          & $TC$     & $FDD$  & $DUC$   & $DOC$   & Time  \\
-            {}     & (m$^3$) & \%Served                                & (min)                     & (min)   & (min)   & (min)   & (min) & \%Served & (min)    & (min)  & (min)   & (min)   & (min) \\
-            \cmidrule(l){2-2} \cmidrule(l){3-8} \cmidrule(l){9-14}
-            Small  & 3917.0  & 100                                     & 1998.75                   & 64.60   & 231.63  & 215.43  & 21.24 & 100      & 1946.06  & 1.55   & 229.40  & 298.74  & 50.34 \\
-            Medium & 32394.0 & 99.6                                    & 13144.05                  & 2528.33 & 2496.47 & 1086.12 & 49.41 & 100      & 13564.08 & 75.77  & 2250.66 & 1347.59 & 50.6  \\
-            Large  & 49516.0 & 99.2                                    & 21395.23                  & 4885.11 & 1250.14 & 3543.16 & 50.20 & 100      & 21784.95 & 471.62 & 1168.73 & 3640.04 & 50.74 \\
-            \bottomrule
-        \end{tabular}
-    }
-\end{table}
-
-\subsubsection{Objectives}
-
-Nous avons vérifié l'impact des objectifs de minimisation de la distance et de la première livraison.
-Ces deux objectives sont contradictoires.
-Si on fait diminuer un, l'autre augmente.
-Il faudrait trouver avec le décideur qu'il y ait un trade-off.
+\subsubsection{Trade-off of the objective function components}
+
+We ran 5 tests on our data and obtained two solutions: one with the shortest travel time and another with the smallest first delivery delay.
+Figures 1 and 2 illustrate the comparisons between travel times and first delivery delays for these solutions.
+It's clear that as the first delivery delay decreases, the travel time increases, and vice versa.
+This highlights a trade-off between minimizing travel time and minimizing first delivery delay and leads the decision maker to find a balance between reducing operating costs and compromising customer satisfaction.
 
@@
     \centering
+    \resizebox{1.25\textwidth}{!}{
     \begin{subfigure}{0.5\textwidth}
@@
     \vspace*{-1mm}
-
+    }
 \end{figure}
-
-\subsubsection{Comparisons with the real operations}
-
-- resultats avec les instances (8 au total) où uniquement la flotte de Unibéton a éét utilisée.
-enlver les contraintes de temps. reduire le gamma1
-
-\begin{table}[!ht]
-    \centering
-    \caption{Comparaison with partner operations}
-    \label{tab:comp_unibeton}
-    \begin{tabular}{lrrrrrrr}
-        \toprule
-                      & \multicolumn{7}{c}{$\gamma^1$}                                                                                                                               \\
-                      & Demand                         & D         & ('distanceCost', 5) & ('distanceCost', 10) & ('distanceCost', 15) & ('distanceCost', 20) & ('distanceCost', 25) \\
-                      & (min)                          & (min)     & (min)               & (min)                & (min)                & (min)                & (min)                \\
-        Instance      &                                &           &                     &                      &                      &                      &                      \\
-        \midrule
-        C_13_5_8_1    & 267.00                         & 543.05    & 523.46              & 523.46               & 523.46               & 523.46               & 523.46               \\
-        C_76_40_43_6  & 1,160.50                       & 5,244.68  & 4,917.41            & 4,726.81             & 4,662.97             & 4,644.56             & 4,704.28             \\
-        C_116_67_78_8 & 1,839.50                       & 11,664.77 & 12,220.90           & 11,084.60            & 11,019.80            & 10,856.90            & 10,760.60            \\
-        C_15_4_7_2    & 375.00                         & 789.28    & 702.88              & 702.88               & 703.46               & 702.31               & 702.88               \\
-        C_18_6_11_2   & 333.50                         & 555.13    & 471.47              & 471.47               & 471.47               & 471.47               & 471.47               \\
-        C_94_63_70_7  & 1,746.50                       & 12,011.22 & 12,776.30           & 12,201.20            & 11,895.40            & 11,469.30            & 11,440.40            \\
-        \bottomrule
-    \end{tabular}
-\end{table}
-
-(utiliser les mêmes pour la comparaison avec les données réelles )
-distance, demande servie, first, underwork, overtime.
-- resultats avec les autres instances
-- distance, demande servie, first delivery.
-- rouler les instances sans shift des livreurs. comparer la distance parcourue, et demande livrée.
-- rouler sans les livreurs externes. regarder la demande servie par jour.
 

### 2023-08-21 Article/versions/21_08_23/article_CDP.tex
version finale article: 21_08_23

--- 2023-08-21 Article/article.tex
@@
 \begin{abstract}
-    The concrete delivery problem (CDP) is a combinatorial optimization problem that involves scheduling the delivery of ready-mixed concrete to construction sites while balancing the conflicting goals of minimizing transportation costs and maximizing customer satisfaction.
-This paper proposes an exact formulation and a heuristic approach based on the Greedy Randomized Adaptive Search Procedure (GRASP) to tackle a new variant of the CDP that incorporates realistic side constraints, such as drivers' working shifts, minimum working time for each driver, and overtime penalties.
-Additionally, the problem also considers the possibility of customers requiring multiple types of concrete to be delivered within the same time window.
-The performance of the proposed heuristic is evaluated on real-world instances, and it is compared to another variant to demonstrate its effectiveness.
-
+    This paper addresses a novel variant of the Concrete Delivery Problem (CDP), which involves the efficient scheduling of ready-mixed concrete deliveries to construction sites while balancing the conflicting goals of minimizing transportation costs and maximizing customer satisfaction.
+In this study, we propose an exact formulation and a heuristic approach based on the Greedy Randomized Adaptive Search Procedure (GRASP), to tackle this challenging CDP variant.
+This variant introduces realistic side constraints, including driver working shifts, a minimum driver working time, and overtime penalties.
+Additionally, it considers the scenario where customers may request multiple types of concrete delivered within the same time window.
+We assess the performance of our heuristic using new public instances generated for this problem and provide a comparative analysis with another CDP variant to demonstrate its effectiveness.
 \end{abstract}
@@
 However, we include the plant's production rate and driver shift schedule.
-The company uses a centralized dispatcher system to schedule all daily orders, but this system has issues satisfying all daily demands without using an external fleet.
+The company uses a centralized dispatcher system to schedule all daily orders, but this system has issues satisfying all daily demands without using an external fleet or using different plants to serve the same order.
 % Also, as per union rules, they assign a truck driver to a delivery task based on his seniority.
@@
 
-The rest of the paper is structured as follows: Section \ref{lit_review} provides a literature review of previous work related to the CDP.
+The rest of the paper is organized as follows: Section \ref{lit_review} provides a literature review of previous work related to the CDP.
 Section \ref{desc_form} provides a formal description and mathematical model of the problem.
-In Section \ref{grasp_method}, we describe the GRASP algorithm and the constructive heuristics developed to solve this variant of the CDP.
-Section \ref{comp_exp} presents computational experiments to evaluate the proposed approach, and finally, the conclusions are presented in Section \ref{concl}.
+Section \ref{grasp_method} describes the GRASP algorithm and the constructive heuristics developed to solve this variant of the CDP.
+Section \ref{comp_exp} presents computational experiments and sensitivity analyses to evaluate the proposed approach, and finally, section \ref{concl} presents the conclusions.
 
@@
 \cite{mayteekrieangkrai2015optimized} addressed the same problem with the same data using a bee algorithm (BA) and found better solutions than the GA.
-\cite{lu2005optimized} used the same combination of GA and simulation to determine the optimal number of concrete mixers to be deployed and an optimal schedule for batch and delivering concrete.
+\cite{lu2005optimized} used the same combination of GA and simulation to determine the optimal number of concrete mixers to be deployed and an optimal schedule for batching and delivering concrete.
 Their objective was to minimize the idle time of the site crew due to late concrete deliveries and truck queuing time.
@@
 A time-space network is the key component of the real-time DSS developed by \cite{durbin2008or} to solve a dynamic CDP every five minutes.
-The DSS is able to receive new orders, schedule them on the fly, and handle unexpected events such as plant closures, truck breakdowns, and delays in transportation times.
+The DSS can receive new orders, schedule them on the fly, and handle unexpected events such as plant closures, truck breakdowns, and delays in transportation times.
 The authors combined the DSS with a tabu search (TS) heuristic to warm start CPLEX, which made the model performant enough to solve instances with up to 1,500 loads per day with up to 250 trucks.
@@
 If the order requires more concrete than a single truck can carry, multiple deliveries are scheduled.
-We will accept a first delivery if it arrives within a time window of $\left[a_i, a_i + \tau^w \right]$, where $\tau^w_i$ is an user-defined parameter.
+We will accept a first delivery if it arrives within a time window of $\left[a_i, a_i + \tau^w \right]$, where $\tau^w_i$ is a user-defined parameter.
 
@@
 Let $\mathcal{B}$ be the set of batch plants.
-The plants are heterogeneous, as each plant $b$ has its own hourly loading rate, represented by $\tau^l_b$, which affects the duration of the loading process.
+The plants are heterogeneous, as each plant $b$ has its hourly loading rate, represented by $\tau^l_b$, which affects the duration of the loading process.
 After loading the concrete, the driver spends $\alpha_b$ minutes adjusting the concrete in the truck before heading to the customer site.
@@
 Let $w^1_k$ be a continuous variable indicating the difference between the driver's work time and the minimum number of hours to be worked in a day, and $w^2_k$ indicating the difference between the driver's work time and the normal work time.
-Let $g_i$ be the time between due date and first service start for customer $i$.
-
-The objective function minimizes total travel cost ($TC$), penalties associated with unfulfilled orders ($PUO$), first delivery delays ($FDD$), driver underutilization costs ($DUC$), and driver overtime costs ($DOC$).
+Let $g_i$ be the time between the due date and the first service start for customer $i$.
+
+The objective function minimizes total travel cost ($TC$), penalties associated with unfulfilled orders, first delivery delays ($FDD$), driver underutilization costs ($DUC$), and driver overtime costs ($DOC$).
 Together, these components drive the optimization process to find a solution that efficiently balances travel costs, customer satisfaction, on-time delivery, driver utilization, and scheduling constraints.
-
-% The objective function minimizes the total travel cost, the penalty costs incurred when customer demands are not fully met, the lateness of each customer's first delivery, the cost of drivers working less than the minimum hours, and the total overtime cost of drivers working beyond their scheduled hours.
-The mathematical model of this variant of the CDP is as follows:
 
@@
 
-% \paragraph*{}
 Let us consider the solution of a small instance of our problem shown in Figures~\ref{fig_Example} and \ref{fig:ganttExample}.
@@
 As shown in Algorithm~\ref{alg:greedyCustIns}, for each customer, we randomly select the first order to schedule.
-To schedule an order, we iterate over each of its delivery nodes, find a feasible pair of task, and add it to the current solution.
+To schedule an order, we iterate over each of its delivery nodes, find a feasible pair of tasks, and add it to the current solution.
 For a delivery node $d$, we check if the list of operations $M[d]$ is empty.
@@
 We extract from $CL$ the restricted candidate list ($RCL$) of drivers for whom the cost of visiting $d$ falls within a specified range based on the parameter $\theta$.
-If $RCL$ is empty and a visit to $d$ cannot be scheduled due to cold-joint constraints, we use the $DelayPreviousDelivery$ procedure to determine if there is a previous delivery of the same order that we can delay the start of service to make a visit to $d$ feasible.
+If $RCL$ is empty and a visit to $d$ cannot be scheduled due to cold-joint constraints, we use the $DelayPreviousDelivery$ procedure to determine if there is a previous delivery of the same order that we can delay the start of service to visit $d$ feasible.
 This procedure returns the index of the node and $\lambda$, which is the value of the delay.
@@
 Consider the scenario where two customers require their services to start at the same time or within the same time window.
-In this case, we can either finish scheduling one customer before the other, or schedule both simultaneously.
+In this case, we can either finish scheduling one customer before the other or schedule both simultaneously.
 If there are sufficient resources available, either option would be feasible.
@@
 The \textit{Relocate} operator removes a delivery node and places it in a different location.
-The \textit{Consolidate Load} operator assigns an order request initially split between two drivers to a single driver, if the driver's capacity allows.
+The \textit{Consolidate Load} operator assigns an order request initially split between two drivers to a single driver if the driver's capacity allows.
 The \textit{Remove and Reschedule} operator selects an unscheduled customer, denoted $i$, within the current solution.
@@
 The implementation of the GRASP algorithm is coded in C++.
-We run our experiments on the benchmark data used in \cite{kinable2014concrete}, and on a dataset extracted from delivery operations records provided by our industry partner.
+We run our experiments on the benchmark data used in \cite{kinable2014concrete}, and on a new dataset extracted from delivery operations records provided by our industry partner.
 
@@
 \newcommand{\nbInstance}{36}
-The dataset used in this section was obtained from our partner and contains records representing delivery operations from up to 8 production centers over a period of $\nbInstance$ days.
+The dataset used in this section was obtained from our partner and contains records representing delivery operations from up to 8 production centers for $\nbInstance$ days.
 From this historical data, we extract information such as daily orders delivered, plant assignments to orders, and the number of available drivers.
@@
             Small  & 8  & 226.5 - 937.5     & 7 - 14    & 4 - 11   & 13-35    & 1-3     \\
+            % \cmidrule{3-7}
             Medium & 14 & 1,160.5 - 2,971.0 & 43 - 98   & 40 - 89  & 76-137   & 6-8     \\
+            % \cmidrule{3-7}
             Large  & 14 & 3,078. - 3,953.5  & 107 - 136 & 92 - 127 & 129-150  & 8       \\
@@
 
-\subsubsection{Results for the instances of \cite{kinable2014concrete}}
+\subsection{Results for the instances of \cite{kinable2014concrete}}
 
@@
 
-\subsubsection{Results for the generated instances}
+\subsection{Results for the generated instances}
 
@@
     \small
-    %    \resizebox*{\textwidth}{!}{
     \begin{tabularx}{\textwidth}{Xr}
@@
     \end{tabularx}
-    %    }
 \end{table}
-%  the travel cost ($TC$), the penalties associated with unfulfilled orders ($PUO$), the sum of the first delivery delays ($FDD$) for each customer, the sum of the driver underutilization costs ($DUC$) and driver overtime costs ($DOC$).
+
 The table \ref*{tab:all_instances} contains the detailed results for all instances with a stop criterion of 3600 seconds.
-The table shows the demand of each instance, the average and the best values of the objective function when solving each instance five times.
-All daily orders of the small instances are completely served by our algorithm within these five replications, as shown in column $UQ$, which reports the undelivered quantity.
-$99.57\%$ and $99.5\%$ of the total demands are served for the medium and large sets.
-The sum of the driver underutilization costs ($DUC$) is high for medium and large instances.
+The table shows the demand of each instance, the average, and the best values of the objective function when solving each instance five times.
+All daily orders of the small instances are completely served by our algorithm within these five iterations, as shown in column $UQ$, which reports the undelivered quantity.
+For the medium and large instances, $99.57\%$ and $99.5\%$ of the total demands are delivered.
+The sum of driver underutilization costs ($DUC$) is high for medium and large instances.
 This indicates the underutilization of the scheduled fleet.
@@
 The sum of driver overtime costs ($DOC$) is also high for these instances.
-This can be explained by the fact that the algorithm prefers to put certain drivers on overtime rather than use a driver who has to travel to another plant for a delivery, which would increase the travel costs ($TC$).
-
-$FDD$ is the sum of the delays of the first deliveries.
-For the small instances, all but one of the first deliveries are on time.
-For the other instances sets, there is at least one customer with a late first delivery.
-To get an idea of the actual delay per customer, we report $mFDD$, which is the maximum delay of all first deliveries.
-This tells us that we have a maximum delay that varies between 0 and $35.7$ minutes for small instances, $16.8$ and $60$ minutes for medium instances, and $56$ to 60 minutes for large instances.
-
-\begin{table}[!ht]
+This can be explained by the fact that the algorithm prefers to put drivers on overtime rather than use a driver who has to travel to another plant for a delivery, which would increase travel costs ($TC$).
+
+FDD represents the sum of the delays of the first deliveries.
+In the case of small instances, all but one of the first deliveries are on time.
+However, for the other instance sets, there is at least one customer with a late first delivery.
+To provide insight into the actual delay experienced by customers, we report mFDD, which is the maximum delay among all first deliveries.
+This information reveals that for small instances, the maximum delay ranges between 0 and 35.7 minutes.
+For medium instances, it varies from 16.8 to 60 minutes, and for large instances, it falls within the range of 56 to 60 minutes.
+
+\begin{table}[htb]
     \centering
@@
 
-Table \ref{tab:runtime} shows the percentage of concrete delivered by varying the execution times of the algorithm.
-For small instances, all requests are delivered in less than a minute, while for medium and large instances, over 99\% of requests are delivered.
-For these instances, it takes more than ten minutes to ensure complete delivery of all orders.
-
-\begin{table}[!ht]
+Table \ref{tab_runtime} shows that as the algorithm runtime increases, a higher percentage of deliveries are completed.
+For small instances, all requests are delivered in less than a minute, while for medium and large instances, 99\% of requests are delivered.
+
+\begin{table}[htb]
     \centering
-    \caption{RMC delivered within different runtimes}
-    \label{tab:runtime}
+    \caption{RMC delivery completion within different runtimes of the GRASP}
+    \label{tab_runtime}
     \scriptsize
@@
 
-Figure \ref*{fig:plants_schedules} illustrates the Gantt diagrams showing the loading and unloading operations for the orders assigned to each plant in instance $C29\_10\_14\_3$.
-It also shows the differents assignments of each driver throughout the day.
-This instance has three plants, $29$ drivers, and two customers wich have multiple orders.
-Customer $8$ has two orders ($o_8$,$o_9$) assigned to plant $0$, and customer $9$ has four orders ($o_{10}$ - - $o_{13}$) assigned to plant $2$.
-All orders requirements are in Table~\ref*{tab:instance_detail}.
+Figure \ref*{fig:plants_schedules} provides a visual representation using Gantt charts of the loading and unloading operations for the orders assigned to each plant in instance $C29\_10\_14\_3$.
+It also shows the different assignments of each driver throughout the day.
+This instance has three plants, twenty-nine drivers, and ten customers, with two customers having multiple orders.
+Customer $8$ has two orders ($o_8$,$o_9$) assigned to plant $0$ while customer $9$ has four orders ($o_{10}$ - - $o_{13}$) assigned to plant $2$.
+All the order requirements are in the table~\ref*{tab:instance_detail}.
 
@@
 
-We can observe that the loading operations for each plant are non-overlapping and the unloading operations occur sequentially, one after the other, at the worksite.
-Plant $0$ is in charge of both orders of customer $8$.
-In figure~\ref*{fig:plant0}, order $o_8$ is served before $o_9$ starting at the due time.
-Each row on the figure represents the pairs of loading and unloading tasks performed by a driver.
-The number near a task references the order.
-Figure~\ref*{fig:plant2} shows how the loading bay is alternately assigned to orders $o_{13}$ and $o_7$ between $10:00$ and $14:00$ a.m.
-Driver $26$ starts its service at plant $0$, serves order $o_9$ four times and travels to plant $2$ to perform the last delivery of order $o_{13}$.
+We can see that the loading operations for each plant do not overlap, and the unloading operations occur sequentially, one after the other, at each worksite.
+Plant $0$ is responsible for both orders from customer $8$.
+In the figure~\ref*{fig:plant0}, order $o_8$ is served before $o_9$, starting at the due time.
+Each row in the figure represents the pairs of loading and unloading tasks performed by a driver.
+The number next to a task refers to the order.
+Figure~\ref*{fig:plant2} shows how the loading bay is alternately assigned to orders $o_{13}$ and $o_7$ between $10:00$ and $14:00$.
+Driver $26$ starts his service at plant $0$, serves order $o_9$ four times, and travels to plant $2$ to make the final delivery of order $o_{13}$.
 Driver $25$ also loads at both plants $0$ and $1$.
 
-We report for this solution of instance $C29\_10\_14\_3$ a $FDD$ of $59.5$ minutes.
-The gantt chart indicates that customers $4,5$ whom need their deliveries at $07:00$ a.m.
-have all a first late delivery.
-The reason for these delays is that the drivers assigned to plant $1$ starts their earliest shift at $06:05$, and $06:20$.
-And it takes $59$($63$) minutes travel time to visit customer $4$($5$) from this plant.
-So it appears that the big value of $FDD$ we find in our results may be caused by a bad schedule of the drivers shifts.
+We report a $FDD$ of $59.5$ minutes for this solution of instance $C29\_10\_14\_3$.
+The Gantt chart shows that the $4.5$ customers who need their deliveries at $07:00$ all have a late first delivery.
+These delays are due to the work schedules of the drivers assigned to Plant 1.
+These drivers start their shifts at the earliest at $06:05$ and $06:20$, and it takes them $59$ to $63$ minutes to reach the 4 to 5 customers.
+This finding suggests that the relatively high FDD values in our results may be related to suboptimal scheduling of drivers' hours in our instances.
 
@@
 The medium instance $C\_104\_42\_47\_8$ has an unserved customer with a demand of 12 $m^3$ and a due time of $07:00$.
-Upon investigation, we find that even if we have $104$ drivers in this instance, the number of drivers starting their shift before the due time is not sufficient to serve this customer within the time window of 60 minutes that we define.
-So the value of $UQ$ we report in Table~\ref*{tab:all_instances} seems to be the optimal given the driver shift schedule given to our algorithm.
-We confirmed this observation by lifting the driver shift and overtime constraints and resolving our instances five more times.
-All orders are now completely served, and at each run.
-Table~\ref*{tab:lift_shift_constraints} compare the average values of the results of the tables \ref*{tab:all_instances} and \ref*{tab:all_instances_no_shift}.
-We found that for small instances, the maximum delay dropped to zero, and for medium (large) instances, it was reduced by almost 60\% (40\%) on average.
+Upon investigation, we find that even if we have $104$ drivers in this case, the number of drivers who start their shift before t$07:00$ is not enough to serve this customer within the 60-minute window we defined.
+Thus, the value of $UQ$ that we report in the table~\ref*{tab:all_instances} appears to be optimal given the driver shift schedule given as input to our algorithm.
+We confirmed this observation by removing the driver start shift and overtime constraints and resolving our instances five more times.
+All orders are now fully serviced on each run.
+The table~\ref*{tab:lift_shift_constraints} compares the averages of the results from  the tables \ref*{tab:all_instances} and \ref*{tab:all_instances_no_shift}.
+We found that for small instances the maximum delay dropped to zero, and for medium (large) instances it was reduced by almost 60\% (40\%) on average.
 This suggests that when we input a good work schedule, our algorithm works effectively to ensure that all orders are filled and that delays for first deliveries and travel costs are minimized.
-We can see that the average travel time decreases for all instances, except for the large one where the difference is not notable.
+We can see that the average travel time decreases in all cases except the large one, where the difference is not significant.
 
@@
     \small
-    \caption{Comparaison of the average results with and without shift and overtime constraints}
+    \caption{Comparison of the average results with and without shift and overtime constraints}
     \label{tab:lift_shift_constraints}
@@
      \cmidrule(l){2-5} \cmidrule(l){6-9}
-     & $TC$ & $PUO$ & $FDD$ & $mFDD$ & $TC$ & $PUO$ & $FDD$ & $mFDD$ \\
+     & $TC$ & $UQ$ & $FDD$ & $mFDD$ & $TC$ & $UQ$ & $FDD$ & $mFDD$ \\
      & (min) & ($m^3$) & (min) & (min) & (min) & ($m^3$) & (min) & (min) \\
@@
     \centering
-    \caption{Results with all instances without shift constraints}
+    \small
+    \caption{Results with all instances without shift and overtime constraints}
     \label{tab:all_instances_no_shift}
@@
 
-\subsubsection{Comparisons with the industry partner historical operations}
-
-\begin{table}[htb]
-    \centering
-    \caption{Comparaison with historical operations}
-    \label{tab:comp_unibeton}
-    \begin{tabular}{lccccccc}
-    \toprule
-     & & & \multicolumn{5}{c}{Distance} \\
-     \cmidrule(l){3-3} \cmidrule(l){4-8}
-     & Demand & Distance &  $\gamma^1=5$&  $\gamma^1=10$ & $\gamma^1=15$ & $\gamma^1=20$& $\gamma^1=25$ \\
-     & ($m^3$) & ($Km$)  &($Km$)  & ($Km$) & ($Km$) &($Km$)  &($Km$)  \\
-    \midrule
-    C_13_5_8_1 & 267.0 & 543.1 & 523.5 & 523.5 & 523.5 & 523.5 & 523.5 \\
-    C_76_40_43_6 & 1,160.5 & 5,244.7 & 5,101.2 & 4,706.1 & 4,673.3 & 4,645.6 & 4,732.4 \\
-    C_116_67_78_8 & 1,839.5 & 11,664.8 & 11,895.8 & 11,035.3 & 11,064.2 & 10,928.9 & 10,823.4 \\
-    C_15_4_7_2 & 375.0 & 789.3 & 702.3 & 702.9 & 702.3 & 703.5 & 702.3 \\
-    C_18_6_11_2 & 333.5 & 555.1 & 491.7 & 471.5 & 471.5 & 471.5 & 471.5 \\
-    C_94_63_70_7 & 1,746.5 & 12,011.2 & 12,957.1 & 11,965.5 & 11,915.6 & 11,657.1 & 11,475.7 \\
-    \bottomrule
-    \end{tabular}
-    \end{table}
-
 \subsection{Sensitivity analyses}
 
-In this section, we analyse the impact on the solution when we change some parameters of our problem, and how some components of the objective function are influencing each other.
-
- % That's because the goals of minimizing travel cost and not delaying the first delivery are sometimes in conflict: to ensure that the first delivery is not late, drivers assigned to one plant often have to travel to another.
-And since all plants are used in large cases, this can increase travel times.
-
-%  Therefore, we can use our algorithm for fleet sizing or as a submodule of a driver scheduling algorithm.
+In this section, we analyze the effect on the solution when we change some parameters of our problem and how some components of the objective function influence each other.
 
@@
 
-We conducted some tests with different values $\gamma^1$ varying between 5 and 25 minutes, without any shifts or overtime constraints on the drivers.
-The influence of the value of the maximum time delay between two consecutive deliveries of the same order is demonstrated in Figure~\ref*{fig:gamma1_influence}.
+We ran additional tests with different values of $\gamma^1$, varying between 5 and 25 minutes, with no shifts or overtime constraints on the drivers.
+The influence of the value of the maximum time delay between two consecutive deliveries of the same order is shown in figure~\ref*{fig:gamma1_influence}.
 For small instances, all customers are served on time regardless of the value of $\gamma^1$.
 For medium (large) instances, all customers are served with $\gamma^1$ starting at 15 (20) minutes.
-The delay of the initial delivery decreases as $\gamma^1$ increases.
-These results confirm the fact that with a high $\gamma^1$ value, plants have more flexibility to delay ongoing deliveries to make room for new ones or make customers wait until drivers are available.
-By increasing the maximum time delay from 10 to 20 minutes, the first delivery delay is reduced by over 50\% for the medium and large size instances.
-The average travel time does not significantly change regardless of the value of $\gamma^1$.
+The delay of initial delivery decreases as $\gamma^1$ increases.
+These results confirm the fact that when $\gamma^1$ is high, plants have more flexibility to delay current deliveries to make room for new ones or to make customers wait until drivers are available.
+Increasing the maximum delay from 10 to 20 minutes reduces the first delivery delay by over 50\% for the medium and large cases.
+The average delivery time does not change significantly regardless of the value of $\gamma^1$.
 
@@
     \centering
-    \includegraphics[width=0.8\textwidth]{gamma1.pdf}
+    \includegraphics[width=1\textwidth]{gamma1.pdf}
     \small
@@
 
-We ran 5 tests on our data and obtained two solutions: one with the shortest travel time and another with the smallest first delivery delay.
-Figures 1 and 2 illustrate the comparisons between travel times and first delivery delays for these solutions.
-It's clear that as the first delivery delay decreases, the travel time increases, and vice versa.
-This highlights a trade-off between minimizing travel time and minimizing first delivery delay and leads the decision maker to find a balance between reducing operating costs and compromising customer satisfaction.
-
-\begin{figure}[!ht]
+We ran five tests on our dataset and obtained two solutions per instance: one that prioritizes the shortest travel time and another that seeks the smallest first delivery delay.
+Figure~\ref*{fig:trade_off} visually represents the trade-off between travel time and first delivery delay for these solutions.
+It is clear that as the first delivery delay decreases, the travel time increases, and conversely when the travel time is minimized, the first delivery delay tends to be longer.
+This result is expected because drivers assigned to a plant often need to travel to another plant to ensure that the first delivery remains on schedule.
+This trade-off highlights the challenge decision-makers face in finding the right balance between minimizing operational costs and meeting customer satisfaction.
+
+\begin{figure}[!h]
     \centering
-    \resizebox{1.25\textwidth}{!}{
-    \begin{subfigure}{0.5\textwidth}
-        \includegraphics[width=\linewidth]{travel.pdf}
-        \vspace*{-2em}
-        \caption{travel cost}
-        \label{fig:travel}
-    \end{subfigure}%
-    \hfill% Horizontal space between the subfigures
-    \begin{subfigure}{0.5\textwidth}
-        \includegraphics[width=\linewidth]{FDD.pdf}
-        \vspace*{-2em}
-        \caption{first delivevery delay}
-        \label{fig:fdd}
-    \end{subfigure}
-    \vspace*{-1mm}
-    }
+    \includegraphics[width=1\textwidth]{Trav_FDD.pdf}
+    \small
+    \caption{Trade-off between minimizing travel cost VS minimizing first delivery delay }
+    \label{fig:trade_off}
 \end{figure}
@@
 
-Il serait intéressant de voir l'effet des paramètres gamma 1 et gamma 2 différents par clients et /ou ordre. Également le temps de service différent par client.
-
-l'algorithme serait intéressant à utiliser dans le cadre d'un
-
-%  Therefore, we can use our algorithm for fleet sizing or as a submodule of a driver scheduling algorithm.
-
+In this paper, we investigated a variant of the concrete delivery problem, where customers can order different types of concrete from different production centers, which must be delivered at the same time.
+To solve this problem, we introduced a mathematical model and a heuristic solution based on Greedy Randomized Adaptive Search.
+Our analysis, using a dataset derived from our partner's historical data, demonstrated the effectiveness of our algorithm in making good decisions within a reasonable time.
+Furthermore, our model incorporates constraints related to working hours, making our heuristic a valuable tool for designing weekly schedules for company drivers.
+Our GRASP algorithm performed exceptionally well on \cite{kinable2014concrete} datasets, indicating its reliability.
+For future work, it would be interesting to address datasets with different unloading times at construction sites or to explore different scenarios regarding the number of simultaneous deliveries and the maximum time between two consecutive deliveries at a construction site.
+We can also study the problem using a stochastic or robust methodology to account for the uncertainty of the parameters in a real-world setting.
+As a contribution to the literature on CDP, we provide a mathematical formulation for our variant, contribute a dataset, and demonstrate the application of GRASP to solve both our problem and the one studied by \cite{kinable2014concrete}.
 \vspace{0.1in}

### 2023-08-24 Article/article.tex
save

@@
 
-We can observe that the loading operations for each plant are non-overlapping and the unloading operations occur sequentially, one after the other, at the worksite.
-Plant $0$ is in charge of both orders of customer $8$.
-In figure~\ref*{fig:plant0}, order $o_8$ is served before $o_9$ starting at the due time.
-Each row on the figure represents the pairs of loading and unloading tasks performed by a driver.
-The number near a task references the order.
-Figure~\ref*{fig:plant2} shows how the loading bay is alternately assigned to orders $o_{13}$ and $o_7$ between $10:00$ and $14:00$ a.m.
-Driver $26$ starts its service at plant $0$, serves order $o_9$ four times and travels to plant $2$ to perform the last delivery of order $o_{13}$.
+We can see that the loading operations for each plant do not overlap, and the unloading operations occur sequentially, one after the other, at each worksite.
+Plant $0$ is responsible for both orders from customer $8$.
+In the figure~\ref*{fig:plant0}, order $o_8$ is served before $o_9$, starting at the due time.
+Each row in the figure represents the pairs of loading and unloading tasks performed by a driver.
+The number next to a task refers to the order.
+Figure~\ref*{fig:plant2} shows how the loading bay is alternately assigned to orders $o_{13}$ and $o_7$ between $10:00$ and $14:00$.
+Driver $26$ starts his service at plant $0$, serves order $o_9$ four times, and travels to plant $2$ to make the final delivery of order $o_{13}$.
 Driver $25$ also loads at both plants $0$ and $1$.
 
-We report for this solution of instance $C29\_10\_14\_3$ a $FDD$ of $59.5$ minutes.
-The gantt chart indicates that customers $4,5$ whom need their deliveries at $07:00$ a.m.
-have all a first late delivery.
-The reason for these delays is that the drivers assigned to plant $1$ starts their earliest shift at $06:05$, and $06:20$.
-And it takes $59$($63$) minutes travel time to visit customer $4$($5$) from this plant.
-So it appears that the big value of $FDD$ we find in our results may be caused by a bad schedule of the drivers shifts.
+We report a $FDD$ of $59.5$ minutes for this solution of instance $C29\_10\_14\_3$.
+The Gantt chart shows that the customers $4$ and $5$ who need their deliveries at $07:00$ all have a late first delivery.
+These delays are due to the work schedules of the drivers assigned to Plant 1.
+These drivers start their shifts at the earliest at $06:05$ and $06:20$, and it takes them $59$ to $63$ minutes to reach the customers.
+This finding suggests that the relatively high FDD values in our results may be related to suboptimal scheduling of drivers' hours in our instances.
 

### 2023-11-29 Article/cdp_content.tex
Save

--- 2023-08-24 Article/article.tex
@@
-\documentclass{article}
-% \documentclass[review]{elsarticle}
-\usepackage{epsfig}
-\usepackage{amsmath}
-\usepackage{mathtools}
-\usepackage{algorithm}
-\usepackage[algo2e,noend]{algorithm2e}
-\usepackage[compatible,noend]{algpseudocode}
-\usepackage{enumitem}
-\usepackage{xspace}
-\usepackage[table,xcdraw]{xcolor}
-\usepackage{amsfonts}
-\usepackage{pifont}
-\usepackage{bbding}
-\usepackage[english]{babel}
-\usepackage{adjustbox}
-\usepackage{tikz}
-\usepackage{subcaption}
-\usetikzlibrary{automata, topaths, calc, positioning, shapes, backgrounds, fit, matrix}
-\usepackage{caption}
-\usepackage{tabularx}
-\usepackage[T1]{fontenc}
-\usepackage{graphicx}
-\usepackage{natbib}
-\usepackage{datatool}
-\usepackage{array}
-\usepackage{pgfplotstable}
-\pgfplotsset{width=7cm,compat=newest}
-\usepackage{booktabs}
-\usepackage{longtable}
-\usepackage{pdflscape}
-\usepackage{afterpage}
-\usepackage{capt-of}
-\usepackage{multirow}
-\usepackage{float}
-\usepackage[text={15cm,21cm},centering]{geometry}
-\usepackage{setspace}
-\usepackage{array}
-\allowdisplaybreaks
-\usepackage{lineno}
-\usepackage{tkz-graph}
-\usepackage{breqn}
-
-\tikzset{
+\begin{abstract}
+    This paper addresses a novel variant of the Concrete Delivery Problem (CDP), which involves the efficient scheduling of ready-mixed concrete deliveries to construction sites while balancing the conflicting goals of minimizing transportation costs and maximizing customer satisfaction.
+In this study, we propose an exact formulation and a heuristic approach based on the Greedy Randomized Adaptive Search Procedure (GRASP), to tackle this challenging CDP variant.
+This variant introduces realistic side constraints, including driver working shifts, a minimum driver working time, and overtime penalties.
+Additionally, it considers the scenario where customers may request multiple types of concrete delivered within the same time window.
+We assess the performance of our heuristic using new public instances generated for this problem and provide a comparative analysis with another CDP variant to demonstrate its effectiveness.
+\end{abstract}
+
+% \begin{keywords}
+\noindent{\textbf{Keywords:} Vehicle scheduling;  concrete delivery; GRASP; ready-mixed concrete }
+% \end{keywords}
+
+%\end{frontmatter}
+
+\section{Introduction}
+\label{sec:cdp_intro}
+Concrete is a widely used building material in construction projects.
+Its perishable nature is affected by many factors that impact its quality \citep{sinha_quality_2021}, which is crucial for the durability and strength of the final construction.
+Concrete comes in two types: ready-mixed concrete (RMC) and site-mixed concrete (SMC).
+RMC is manufactured in a batch plant and delivered to the construction site, while SMC is produced on-site using raw materials stored on the construction site.
+Using SMC can avoid delays caused by road traffic, but it has a slower and more difficult production process, requires storage for mixing materials and equipment, and is suitable for low amounts of concrete.
+On the other hand, RMC has better quality and benefits from lower production costs \citep{muresan_comparing}.
+However, to take advantage of these benefits, the batch plant manager must ensure efficient and prompt delivery on the construction site, which may require a fleet of high-cost revolving drum trucks (concrete mixers) to dispatch the RMC.
+
+Concrete delivery under the form of RMC is subject to many operational constraints that make the Concrete Delivery Problem (CDP) very challenging.
+In this paper, we study a variant of the CDP to schedule the daily production and dispatching of RMC for a company located in the province of Quebec, Canada.
+This company operates multiple batch plants with varying production rates, using a fleet of concrete mixers of different capacities.
+Each plant has its own fleet of trucks; however, under certain conditions, the trucks can move between plants if necessary.
+The trucks must return to their home plant at the end of the day.
+The company owns two types of trucks with different capacities and can call on an external fleet when needed.
+They serve construction sites from any of their plants, with the first delivery starting at the time specified by the customer.
+The loading and unloading of a concrete mixer depend on the truck capacity, the loading rate at the plant, and the unloading rate at the construction site.
+Drivers are allocated based on their daily work schedules.
+A customer may request several types of concrete to be delivered within the same time window, with no required sequence for the orders, but an order can only start after the completion of the previous order.
+This constraint generalizes the linked order constraints of \cite{durbin2008or} where some customers place two orders for the same day and request that they are linked (the second order begins only after the first order is completed).
+The setting of our study is similar to a variant of the CDP previously studied in \cite{schmid2009hybrid, schmid2010hybridization}.
+However, we include the plant's production rate and driver shift schedule.
+The company uses a centralized dispatcher system to schedule all daily orders, but this system has issues satisfying all daily demands without using an external fleet or using different plants to serve the same order.
+% Also, as per union rules, they assign a truck driver to a delivery task based on his seniority.
+
+According to \cite{blazewicz2019handbook}, the CDP combines vehicle routing with scheduling issues to plan routes to deliver concrete from batch plants (depots) to customers' construction sites.
+RMC is an on-demand product with a short life cycle from production through end use.
+It cannot be stored and cannot stay too long on a truck, or it will harden.
+Hence, concrete mixers must deliver RMC at the planned construction site shortly after its production.
+\cite{tommelein1999just} describes RMC production and delivery as an example of a just-in-time (JIT) production system in construction.
+A customer quantity requirement is often greater than the truck size and must be fulfilled by multiple deliveries.
+In that sense, CDP is similar to the vehicle routing problem with split delivery \citep{archetti2008split}, except that the same truck may visit a customer more than once.
+Concrete hardens quickly, so multiple deliveries must be done back-to-back or at least close in time to avoid the problem of cold joint, which can reduce the strength and durability of the concrete.
+Customers request to be served within a specific time window, which can complicate truck-loading schedules when a plant can only load one truck at a time.
+Similarly, only one truck can unload at a time at a customer location, sometimes leading to concrete mixers queuing and waiting their turn to deliver.
+Furthermore, with trucks of varied sizes, loading, travel, and unloading times that may be uncertain, the CDP is a complex and challenging problem.
+
+In this paper, we propose a mathematical model and a Greedy Randomized Search Procedure (GRASP) heuristic to solve a new variant of the CDP.
+Our model takes into consideration the working shifts of the drivers and the scheduling of multiple orders within the same time window at a construction site.
+To the best of our knowledge, this paper is the first that deals with these specific constraints in the context of the CDP.
+
+The rest of the paper is organized as follows: Section \ref{sec:cdp_lit_review} provides a literature review of previous work related to the CDP.
+Section \ref{sec:cdp_desc_form} provides a formal description and mathematical model of the problem.
+Section \ref{sec:cdp_grasp_method} describes the GRASP algorithm and the constructive heuristics developed to solve this variant of the CDP.
+Section \ref{sec:cdp_comp_exp} presents computational experiments and sensitivity analyses to evaluate the proposed approach, and finally, section \ref{sec:cdp_concl} presents the conclusions.
+
+\section{Literature review}
+\label{sec:cdp_lit_review}
+
+Academic Research on concrete batch and delivery began in the late 1990s.
+\cite{tommelein1999just} described RMC as a prototypical example of a JIT production system in construction and identified two practices for delivering it.
+One approach is for the customer to haul the product from the batch plant with their concrete mixer, while the other is for the batch plant to deliver the concrete directly to the customer's location.
+This latter approach is the one that has been studied in all related papers found in the literature.
+Several works to schedule and dispatch concrete production and delivery have mainly focused on simulation-related methods.
+These methods can be standalone, such as those used by \cite{zayed2001simulation, wang2001scheduling, tian_simulation_based_2010, panas_simulation_based_2013} and \cite{galic2016simulation}, among others.
+Alternatively, these methods can be hybridized with optimization techniques, such as those used by \cite{feng2004optimizing, lu2005optimized}, and \cite{feng_integrating_2006}.
+\cite{wang2001scheduling} developed a simulation model to reveal the effect and value of the concrete mixers' inter-arrival time on the productivity of hired unloading equipment on site.
+\cite{feng2004optimizing} used a combination of genetic algorithm (GA) and simulation process to minimize the total waiting time for trucks at a customer site.
+The study focused on loading trucks with identical capacities at the same batch plant, with fixed loading and unloading durations.
+The GA was used to find the best loading sequence of RMC trucks to be assigned to different construction sites.
+The simulation process determined the loading, arrival, departure, and waiting time of trucks and thus evaluated the cost of each dispatching sequence.
+They evaluated their method using data from a batch plant in Taiwan with up to nine customers served.
+\cite{mayteekrieangkrai2015optimized} addressed the same problem with the same data using a bee algorithm (BA) and found better solutions than the GA.
+\cite{lu2005optimized} used the same combination of GA and simulation to determine the optimal number of concrete mixers to be deployed and an optimal schedule for batching and delivering concrete.
+Their objective was to minimize the idle time of the site crew due to late concrete deliveries and truck queuing time.
+In this setting, it was also necessary to deliver a batch of mortar on-site to lubricate the unloading pump before the concrete delivery.
+As such, the simulation model also included the batch and delivery of mortar.
+Finding the best RMC fleet size was also the purpose of the discrete-event simulation model proposed by \cite{panas_simulation_based_2013}.
+
+In addition to simulation-based methods, several other approaches have been used in the literature to solve the CDP.
+These include metaheuristics \citep{faria2006distributed, misir2011selection, maghrebi2016sequential, yang2022concrete}, exact methods \citep{yan2007optimal, asbach2009analysis, kinable2014concrete}, matheuristics \citep{schmid2009hybrid, schmid2010hybridization}, Benders Decomposition \citep{maghrebi2014benders}, column generation (CG) \citep{maghrebi2014solving, maghrebi2016column}, Lagrangian relaxation \citep{narayanan2015using}, and machine learning approaches such as those used by \cite{graham2006modeling, maghrebi2014exploring, maghrebi2016matching}.
+\cite{matsatsinis2004towards} designed a decision support system (DSS) for the dynamic routing of both concrete and pumps that may be necessary for some construction sites to aid in the unloading of concrete.
+The DSS considered the availability of three plants but stipulated that vehicles fulfilling the same order must all load at the same plant.
+Orders that could not be executed immediately could be postponed for the next day.
+The routing of the pumps was modeled as a multi-depot vehicle routing problem with time windows.
+\cite{naso2007genetic} proposed a sequential GA method combined with constructive heuristics to solve another variant of the CDP.
+In this problem, the plant's production schedule must account for orders for concrete to be delivered to a customer site and orders that customers must pick up themselves.
+The algorithm first schedules the plant loading operations before scheduling truck deliveries.
+The authors also developed a non-linear model that minimizes transportation costs, waiting times, outsourced costs, and overtime work.
+They ran experiments using real-world instances of a concrete supply chain in the Netherlands and found a reduction in the number of outsourced requests.
+\cite{yan2007optimal} also considered overtime considerations in their paper, which focused on scheduling RMC for one batch plant with two loading docks.
+The study took into account that overtime wages are paid for factory and construction site operations after 4 PM.
+They developed a mixed-integer programming (MIP) model on a time-space network to minimize travel times and operating costs at both normal and overtime working hours at the plant and the construction sites.
+They tested the model using real data consisting of three days of operation using a two-stage algorithm.
+First, they solved the MIP relaxation with CPLEX.
+Then, they simplified the original model by fixing some decision variables before solving it.
+The algorithm was found to improve the actual plant operation by 10\%.
+A time-space network is the key component of the real-time DSS developed by \cite{durbin2008or} to solve a dynamic CDP every five minutes.
+The DSS can receive new orders, schedule them on the fly, and handle unexpected events such as plant closures, truck breakdowns, and delays in transportation times.
+The authors combined the DSS with a tabu search (TS) heuristic to warm start CPLEX, which made the model performant enough to solve instances with up to 1,500 loads per day with up to 250 trucks.
+The DSS also considers the case of a customer who places two orders, with the first being completed before the second starts.
+Further insights on the real-time planning and monitoring of CDP are available in \cite{garza2021dynamic}.
+Another variant of the CDP is modeled by \cite{schmid2009hybrid} as an integer multicommodity network flow (MCNF) problem on a time-space network.
+In this paper, concrete is delivered using a heterogeneous fleet of vehicles, and each plant can load an unlimited number of trucks simultaneously.
+Some of the trucks have specialized equipment and must arrive first at certain construction sites to assist in unloading the concrete.
+The objective is to fulfill all orders, minimize the travel cost, and avoid delays between two consecutive unloading operations for an order.
+The model is typically solved using a matheuristic algorithm that combines the MCNF with a variable neighborhood search (VNS) heuristic.
+The method can quickly solve large problem instances with more than 60 orders per day without encountering any memory issues.
+The same problem is addressed by \cite{schmid2010hybridization}, who proposed a MIP model combined with a VNS and a very large neighborhood search (VLNS) to develop two matheuristics approaches.
+Comparisons between both matheuristics and a standalone VNS show that the former methods are much better and suitable for solving larger problem instances.
+These methods also provide better solutions for small to medium instances than the matheuristic used in \cite{schmid2009hybrid}.
+A pure VNS approach with the same problem but without the use of instrumentation has been applied by \cite{payr2009optimizing}.
+
+Regarding objectives, most authors have focused on minimizing travel time and delays between consecutive deliveries.
+However, some authors have been more interested in maximizing customer satisfaction alone.
+We find these situations in the works of \cite{durbin2008or, kinable2014concrete, kinable2014logic, sulaman2017simulated}.
+\cite{kinable2014concrete} introduce a general MIP and constraint programming (CP) models of the CDP reflecting the main constraints commonly found in all CDP works: time lag and no overlapping between consecutive deliveries, covering of all customers' demands, delivery time window, and heterogeneous fleet.
+However, the model did not include constraints limiting the time that concrete may reside in a truck.
+The authors propose a constructive heuristic that schedules the visits to the customers one by one according to the start time of the visit and the truck capacity.
+The procedure is invoked multiple times for different permutations of the customer's order which is determined using the steepest descent (SD) local search procedure.
+One of the paper's main contributions is the creation of the first public test instances for the CDP with up to 50 customers, four batch plants, and 20 concrete mixers.
+They found the CP model to be highly effective in finding high-quality solutions in a relatively short time or improving existing schedules, while the MIP model can be used to compute bounds, as it seems ineffective in solving large problem instances.
+Finally, the heuristic often yields good solutions in less than a second.
+A detailed analysis of the MIP model presented in \cite{kinable2014concrete} and of two more compact models can be found in the thesis of \cite{hernandez_lopez_study_2020}.
+In \cite{kinable2014logic}, we found an attempt to solve the previous problem with a logic-based Benders' approach.
+%\cite{sulaman2017simulated} expand upon the SD-heuristic proposed in \cite{kinable2014concrete}, proposing a simulated annealing (SA) combined with a time-slot Heuristic (SATH).
+This method looks for a slot between existing visits of a truck to schedule a new delivery instead of assigning it to the time slot strictly after the truck's latest assigned delivery.
+The goal is to reduce the large time gaps that can be present in a schedule created with SD due to ignoring the intermediate available time slots.
+Experimental results indicated that SATH outperforms SD in speed and solution quality.
+A generalization of the MIP model of \cite{kinable2014concrete} is addressed in \cite{asbach2009analysis}.
+This model simultaneously minimizes the total sum of travel costs and the penalty costs for customers with unfulfilled demand.
+A customer can request that all concrete deliveries come from the same plant or a subset of plants and that a delivery truck belongs to a subset of the vehicle fleet.
+The MIP model is used in a local search scheme as a black-box solver to reoptimize an incumbent solution in which a neighborhood operator has unfixed some variables.
+\cite{tzanetos2023systematic} provide an overview of the various methods used in the literature to address the CDP and categorizes the problem formulations based on the different concepts used in the literature.
+They also discussed the consistency between industry needs and existing constraints and provided insights into the datasets corresponding to real-world cases, identifying the necessary data for practitioners.
+
+\section{Problem description}
+\label{sec:cdp_desc_form}
+The focus of this paper is on the distribution of RMC from a Canadian company that operates in the greater Montreal area.
+When a customer places an order, it is received at a control center and immediately assigned to one of the company's batch plants.
+These plants produce the concrete and then deliver it to the customer.
+The problem we are examining involves a set of customer orders, a set of concrete-mixer drivers, and a set of batch plants.
+
+A customer $i$ requests one or more types of concrete to be delivered to their construction site on a specific day, with the delivery service starting at the due time $a_i$.
+We call an order $o$ a request for a specific type of concrete.
+$q_i$ is the sum of the demands $q_o$ of each order $o$ placed, $a_i$ is the desired arrival time of the first concrete mixer, and $\tau^u_i$ is the unloading rate for customer $i$.
+If the order requires more concrete than a single truck can carry, multiple deliveries are scheduled.
+We will accept a first delivery if it arrives within a time window of $\left[a_i, a_i + \tau^w \right]$, where $\tau^w_i$ is a user-defined parameter.
+
+Let $O_i$ be the set of all orders requested by customer $i$.
+Each element of $O_i$ must be fully delivered before moving on to another order.
+Exactly one order $o \in O_i$ must have its first delivery start at $a_i$, while the others can start at most $\gamma^2$ time after $o$ is completed.
+A plant is assigned to an order, and it must be the supplier of all subsequent deliveries of that order.
+To avoid cold joint problems with the concrete, subsequent deliveries of the same order must be made in close succession.
+We define a maximum time delay $\gamma^1$ after which no more deliveries are allowed.
+The customer's unloading rate and the quantity to be unloaded give the time required to unload a truckload.
+Let $\mathcal{C}$ be the set of construction sites (customers) with a planned delivery for the day, and $\mathcal{O}=\{{O_i, i \in \mathcal{C}}\}$ be the set of all requested orders for all customers.
+
+The company has two types of concrete mixer trucks with capacities of 8 and 12 cubic meters.
+Each driver $k$ is assigned to a particular batch plant and is responsible for driving a truck with capacity $Q_k$.
+The set of drivers is represented by $K =\cup_{b \in \mathcal{B}}K_b $, where $K_b$ is the set of drivers scheduled to start their shift at batch plant $b$.
+$t_{ij}$ is the known time to travel between any two locations $i$ and $j$.
+A scheduled driver $k$ is required to start his shift at $H_k$, work a minimum of $M_T$ hours and a maximum of $N_T$ hours during regular working hours, with the possibility of overtime of up to $O_T$ hours.
+$\beta_3$ and $\beta_4$ are the penalties for a driver working less than $M_T$ and more than $N_T$.
+
+A driver typically loads RMC at his assigned batch plant but may be required to drive to and load at other plants if needed.
+The batch plant produces concrete on demand using recipes specific to each order.
+This means that a truck can only haul RMC for one order, even if there is spare capacity.
+After unloading the RMC,  driver $k$ takes $\rho$ minutes to clean the concrete mixer before proceeding.
+%When assigning drivers to deliveries, the company prioritizes employees with the highest seniority.
+
+Let $n_o$ be the number of deliveries needed to fulfill the order $o$.
+$n_o$ is not known in advance because we use a fleet of trucks with different capacities.
+However, we can compute its lower ($n_o^{min}$) and upper ($n_o^{max}$) bounds using the capacities of the largest ($Q_{max}$) and smallest ($Q_{min}$) available trucks.
+\begin{alignat}{3}
+    \label{mod:c0}
+    n_o^{min} = \left\lceil \frac{q_o}{Q_{max}} \right\rceil \leq n_o \leq n_o^{max} = \left\lceil \frac{q_o}{Q_{min}} \right\rceil & \text{ } &
+    \forall  o \in \mathcal{O}.
+\end{alignat}
+
+Let $d^j_{o}$ be the $j^{th}$ visit with load $q^j_{o}$ for order $o$.
+We represent the fulfillment of order $o$ by the visits to the ordered set of delivery nodes $\mathcal{D}_o= \left(d^0_{o},d^1_{o},\cdots, d^{n_o}_{o}\right)$.
+The deliveries of customer $i$ are the ordered set $\mathcal{D}_i= (\mathcal{D}_{o_1}, \mathcal{D}_{o_2},\cdots,\mathcal{D}_{o_{|O_i|}})$, where $o_r$ is the $r^{th}$ delivered order.
+We will refer to $d \in \mathcal{D}_i$ ($d \in \mathcal{D}_o$) as the $d^{th}$ potential delivery of customer $i$ (order $o$).
+$\mathcal{D}=\bigcup_{i\in \mathcal{C}} \mathcal{D}_i$ is the union of all delivery nodes.
+
+% \subsection*{Batching plants}
+
+Each of the company's plants has a single loading dock that can accommodate only one truck at a time.
+As a result, trucks often form a queue while waiting for their turn at the loading dock.
+Let $\mathcal{B}$ be the set of batch plants.
+The plants are heterogeneous, as each plant $b$ has its hourly loading rate, represented by $\tau^l_b$, which affects the duration of the loading process.
+After loading the concrete, the driver spends $\alpha_b$ minutes adjusting the concrete in the truck before heading to the customer site.
+Each plant has its own assigned fleet of trucks, but it can borrow trucks from other plants.
+Let $l_{j}$ be the loading dock node associated with delivery node $j$.
+After loading RMC at $l_j$, it must be fully delivered to $j$ at most before $\Delta$ time, which is the concrete lifetime.
+We denote $\mathcal{L}_b$ as the set of loading dock nodes of plant $b$ and $\mathcal{L}$ as the set of all loading docks.
+
+% \subsection*{Model}
+A solution to the problem involves decisions about truck loading schedules, driver assignments to different deliveries, and truck arrival times at construction sites for unloading.
+For a batch plant, the decision involves choosing which driver to load, when to load them, how much to load, and which construction site to deliver.
+For a driver, the decision is to determine the sequence of loading depots and delivery sites.
+And for a construction site, the decision involves determining the arrival times of all scheduled deliveries for the day.
+
+Each driver leaves and returns to their home plant every day.
+We represent the home plant of a driver $k$ with a starting depot $s_k$ and an ending depot $e_k$.
+$S$ and $E$ are the sets of starting and ending depots, respectively.
+
+We define our problem on a complete directed graph where $V=\{ S \cup \mathcal{L} \cup \mathcal{D} \cup E\}$ is the set of nodes.
+The arc sets are $A =  \{(i,j,k) \hspace*{1mm} \vert \hspace*{1mm} i, j \in V \hspace{1mm} k\in K \}$, $A^D = \{(i,j) \hspace*{1mm} \vert \hspace*{1mm} i, j \in \mathcal{D} \hspace{1mm} \}$, and $A^L = \{(i,j) \hspace*{1mm} \vert \hspace*{1mm} i, j \in \mathcal{L} \hspace{1mm} \}$.
+$A$ corresponds to allowed movements of drivers from node $i$ to node $j$. For each driver $k$, the allowed movements are the following:
+
+\begin{itemize}
+    \item From the starting depot $s_k$ to a loading dock $l \in \mathcal{L}$ or to the ending depot $e_k$.
+    \item From a loading dock $l \in \mathcal{L}$ to a delivery node $d \in \mathcal{D}$.
+    \item From a delivery node  $d \in \mathcal{D}$ to a loading dock $l \in \mathcal{L}$ or to the ending depot $e_k$.
+\end{itemize}
+
+For a customer $c$, arcs in $A^D$ link consecutive delivery nodes of the same order   $\lbrace (i,j)\in \mathcal{D}_o, o \in \mathcal{O}_c, i < j  \rbrace$, and pair of delivery nodes of two different orders $ \lbrace (i,d^{0}_{o_2}),  i \in \mathcal{D}_{o_1}, i \geq n^{min}_{o_1}, o_1, o_2 \in \mathcal{O}_c, o_1 \neq o_2 \rbrace $.
+Arcs in $A^L$ link all pairs of loading docks of the same batch plant.
+
+We define $\delta^{+}(i) = \{(i, j,k) \in A \}$ and $\delta^{-}(i) = \{(j, i,k) \in A \}$ as the outcoming and incoming arc sets of node $i \in V$; $\delta^{+}_D(i) = \{(i, j)  \in A^D\}$ and $\delta^{-}_D(i) = \{(j, i) \in A^D \}$ are the outcoming and incoming arc sets of delivery node $i \in \mathcal{D}$.
+Similarly, $\delta^{+}_L(i) = \{(i, j)  \in A^L\}$ and $\delta^{-}_L(i) = \{(j, i) \in A^L \}$ are the outcoming and incoming arc sets of loading node $i \in \mathcal{L}$.
+
+Let the binary variable $x^{k}_{ij}$ be $1$ if driver $k$ travels from node $i$ to $j$.
+Binary variable $y_o$ is $1$ when order $o$ is completely served; $v_i$ and $w_i$ are the start and end of the loading (unloading) operation at node $i \in \mathcal{L}\cup \mathcal{D}$.
+Binary variable $u_{ij}$ is $1$ if node $j$ is served just after $i$, the service being either an unloading or a loading operation.
+% Binary variable $\sigma_{ib}$ is equal to 1 if the orders of customer $i$ are loaded by plant $b$.
+Variable $q^k_j$ is the quantity to be loaded towards $j$ with vehicle $k$.
+Let $w^1_k$ be a continuous variable indicating the difference between the driver's work time and the minimum number of hours to be worked in a day, and $w^2_k$ indicating the difference between the driver's work time and the normal work time.
+Let $g_i$ be the time between the due date and the first service start for customer $i$.
+
+The objective function minimizes total travel cost ($TC$), penalties associated with unfulfilled orders, first delivery delays ($FDD$), driver underutilization costs ($DUC$), and driver overtime costs ($DOC$).
+Together, these components drive the optimization process to find a solution that efficiently balances travel costs, customer satisfaction, on-time delivery, driver utilization, and scheduling constraints.
+
+\begin{alignat}{3}
+    \label{mod:obj}   \min & \smashoperator{ \sum_{ \substack{(i,j,k) \in A} }} {t_{ij}x^{k}_{ij}} + \beta_1\smashoperator{\sum_{ o \in \mathcal{O}}}{  \left(1-y_o\right)} \hspace{-0mm}  + \beta_2 \smashoperator{ \sum_{i \in \mathcal{C} } } { g_i  } +  \sum_{k \in K}\left({\beta_3 w^{1}_k  + \beta_4  w^{2}_k}\right) &                                                                                                  \\
+    \label{mod:c1}         & \sum_{j \in \delta^{+}(i) }{x^k_{ij}} =1                                                                                                                                                                                                                                                         & \forall i \in S, k \in K                                                                         \\
+    \label{mod:c2}         & \sum_{j \in \delta^{-}(i)}{x^k_{ji}} = 1                                                                                                                                                                                                                                                         & \forall i \in E, k \in K                                                                         \\
+    \label{mod:c03}        & v_j \geq  w_i  - M\left(1-x^{k}_{ij}\right)                                                                                                                                                                                                                                                      & \hspace{-20mm}  i \in S, j \in \delta^{+}(i),  k \in K                                           \\
+    \label{mod:c3}         & v_j \geq  w_i + \alpha_{b} + t_{ij} - M\left(1-x^{k}_{ij}\right)                                                                                                                                                                                                                                 & \hspace{-20mm} \forall b \in \mathcal{B}, i \in \mathcal{L}_b, j \in \delta^{+}(i),  k \in K     \\
+    \label{mod:c4}         & v_j \geq  w_i + \rho + t_{ij} - M\left(1-x^{k}_{ij}\right)                                                                                                                                                                                                                                       & \hspace{-20mm} \forall i \in \mathcal{D}  , j \in \delta^{+}(i), k \in K                         \\
+    \label{mod:c5}         & w_{i} \geq v_{i}  + \frac{q^k_j}{\tau^l_b} -M\left(1- x^{k}_{ij}\right)                                                                                                                                                                                                                          & \hspace{-20mm} \forall  b \in \mathcal{B},  i \in \mathcal{L}_{b},  j \in \delta^{+}(i), k \in K \\
+    \label{mod:c6}         & w_{j} \geq v_{j}  + {  \frac{q^k_j}{\tau^u_c} -M\left(1- x^{k}_{ij}\right) }                                                                                                                                                                                                                     & \hspace{-20mm}  \forall c \in  \mathcal{C}, j \in \mathcal{D}_{c}, i \in \delta^{-}(j),  k \in K \\
+    \label{mod:c7}         & w_{j} \leq v_{i}  + \Delta + M\left(1- x^{k}_{ij}\right)                                                                                                                                                                                                                                         & \hspace{-20mm}   j \in \mathcal{D},  i \in \delta^{-}(j),  k \in K                               \\
+    \label{mod:c8}         & v_{d^0_{o}} \geq a_i    & \hspace{-20mm} \forall  i \in \mathcal{C}, \forall o \in \mathcal{O}_i                           \\
+    \label{mod:c9}  & g_{i} \geq v_{d^0_{o_1}} - a_i - M\left(  \sum_{ j \in \delta^{-}_{D}(d^0_{o_1})  }{u_{jd^0_{o_1}}}\right)  & \hspace{-20mm}  \forall  i \in \mathcal{C}, \forall o_1 \in \mathcal{O}_i   \\
+    \label{mod:c91}  & g_{i} - a_i \leq \tau^w  & \hspace{-20mm}  \forall  i \in \mathcal{C}  \\
+    \label{mod:c101}       & v_{d^0_{o_1}} \geq w_j - M\left(1- u_{jd^0_{o_1}}\right)   & \hspace{-40mm}    \forall o_1 \in \mathcal{O}_i,  j \in \delta_{D}^{-}(d^0_{o_1})                \\
+    \label{mod:c10}        & v_{d^0_{o_1}}  \leq  w_j + \gamma^2 + M\left(1- u_{j,d^0_{o_1}}\right)                                                                                                                                                                                                                           & \hspace{-20mm}  \forall  o_1 \in \mathcal{O},  j \in \delta_{D}^{-}(d^0_{o_1})                   \\
+    \label{mod:c11}        & \sum_{ \substack{o_1 \in \mathcal{O}_i }}{ \sum_{ j \in \delta^{-}_{D}(d^0_{o_1})}}{u_{j,d^0_{o_1} } } = |\mathcal{O}_i|-1                                                                                                                                                                       & \hspace{-20mm} \forall  i \in \mathcal{C},|\mathcal{O}_i| > 1                                    \\
+    \label{mod:c12}        & \sum_{ \substack{o_1 \in \mathcal{O}_i }}\sum_{ j \in \delta^{+}_{D}(d^0_{o_1})}{u_{d^0_{o_1},j } } = |\mathcal{O}_i|-1                                                                                                                                                                          & \hspace{-40mm} \forall  i \in \mathcal{C}, |\mathcal{O}_i| > 1                                   \\
+    \label{mod:c13}        & \hspace{3mm} \smashoperator{ \sum_{ j \in \delta^{+}_{D}(d^0_{o_1})}}{u_{d^0_{o_1},j } } \leq 1                                                                                                                                                                                                  & \hspace{-20mm}  \forall o_1 \in \mathcal{O}                                                      \\
+    \label{mod:c14}        & \hspace{2mm} \smashoperator{ \sum_{ j \in \delta^{-}_{D}(d^0_{o_1})}}{u_{j,d^0_{o_1} } } \leq 1                                                                                                                                                                                                  & \hspace{-20mm} \forall o_1 \in \mathcal{O}                                                       \\
+    \label{mod:c15}        & \hspace{2mm} \smashoperator{ \sum_{ j \in \delta^{+}_{D}(d^0_{o_1})}}{u_{d^0_{o_1},j } } +\smashoperator{ \sum_{ j \in \delta^{-}_{D}(d^0_{o_1})}}{u_{j,d^0_{o_1} } } \geq 1                                                                                                                     & \hspace{-20mm} \forall o_1 \in \mathcal{O}                                                       \\
+    \label{mod:c16}        & v_{j} \geq w_{j-1} - M\left(1- u_{j-1,j}\right)                                                                                                                                                                                                                                                  & \hspace{-20mm}  \forall o \in \mathcal{O}, j \in \mathcal{D}_{o}, j \geq 1                       \\
+    \label{mod:c17}        & v_{j}  \leq  w_{j-1} + \gamma^1 + M\left(1- u_{j-1,j}\right)                                                                                                                                                                                                                                     & \hspace{-20mm} \forall i \in \mathcal{C}, \forall o \in O_i, j \in \mathcal{D}_{o}, j \geq 1     \\
+    \label{mod:c18}        & u_{j-1,j} \geq u_{j,j+1}                                                                                                                                                                                                                                                                         & \hspace{-35mm} \forall o_1 \in\mathcal{O},  j \in \mathcal{D}_{o_1}, 1 \leq j \leq n_{o_{1}}-1   \\
+    \label{mod:c19}        & u_{j-1,j} \geq \sum_{l \in \mathcal{L}}{x_{lj}}                                                                                                                                                                                                                                                  & \hspace{-20mm}  \forall o_1 \in \mathcal{O}, j \in \mathcal{D}_{o_1}, j \geq 1                   \\
+    \label{mod:c20}        & v_{j} \geq w_{i} - M\left(1- u_{i,j}\right)                                                                                                                                                                                                                                                      & \hspace{-20mm}  \forall i \in \mathcal{L}, j \in \delta^{+}_{L}(i)                               \\
+    \label{mod:c21}        & \hspace{3mm} \smashoperator{ \sum_{ j \in \delta^{+}_{L}(i)}}{u_{i,j } } \leq 1                                                                                                                                                                                                                  & \hspace{-20mm}  \forall i \in \mathcal{L}                                                        \\
+    \label{mod:c22}        & \hspace{2mm} \smashoperator{ \sum_{ j \in \delta^{-}_{L}(i)}}{u_{j,i } } \leq 1                                                                                                                                                                                                                  & \hspace{-20mm} \forall i \in \mathcal{L}                                                         \\
+    \label{mod:c230}       & \hspace{2mm} \smashoperator{ \sum_{ j \in \delta^{-}_{L}(i)}}{u_{j,i } }  \geq \sum_{k \in K}\sum_{j \in  \delta^{-}(i) }{x^k_{ji}}                                                                                                                                                              & \hspace{-20mm} \forall i \in \mathcal{L}                                                         \\
+    \label{mod:c231}       & \hspace{2mm} \smashoperator{ \sum_{ j \in \delta^{-}_{L}(i)}}{u_{j,i } } \geq \smashoperator{ \sum_{ j \in \delta^{-}_{L}(i+1)}}{u_{j,i+1 } }                                                                                                                                                    & \hspace{-20mm} \forall i \in \mathcal{L}                                                         \\
+    \label{mod:c23}        & \hspace{2mm} \smashoperator{ \sum_{ j \in \delta^{-}_{L}(i)}}{u_{j,i } } + \smashoperator{ \sum_{ j \in \delta^{+}_{L}(i)}}{u_{i,j } } \geq \sum_{k \in K}\sum_{j \in  \delta^{-}(i) }{x^k_{ji}}                                                                                                 & \hspace{-20mm} \forall i \in \mathcal{L}                                                         \\
+    \label{mod:c24}        & \sum_{k \in K}\sum_{j \in \mathcal{D}_o}{q^k_j} = q_o                                                                                                                                                                                                                                            & \hspace{-20mm}  \forall o \in \mathcal{O}                                                        \\
+    \label{mod:c240}       & q^k_j \leq \sum_{i \in  \delta^{-}(j)}{Q^k x^k_{ij}}                                                                                                                                                                                                                                             & \hspace{-20mm}  \forall j \in \mathcal{D}, k \in K                                               \\
+    \label{mod:c25}        & \sum_{k \in K}\sum_{j \in \delta^{+}(i)}{x^{k}_{ij} \leq 1}                                                                                                                                                                                                                                      & i \in \mathcal{L} \cup \mathcal{D}                                                               \\
+    \label{mod:c26}        & \sum_{k \in K}\sum_{j \in \delta^{+}(i)} {x^{k}_{ij} } =  \sum_{k \in K}\sum_{j \in \delta^{-}(i)}{x^{k}_{ji} }                                                                                                                                                                                  & i \in \mathcal{L} \cup \mathcal{D}                                                               \\
+    % \label{mod:c27}        & \sum_{k \in K}  \sum_{i \in \mathcal{D}_c} \sum_{j \in \delta^{-}(i) }  {x^k_{ji}} \leq M\sigma_{c, b}     & \forall c \in \mathcal{C}, b \in \mathcal{B}                                                     \\
+    % \label{mod:c28}        & \sum_{b \in \mathcal{B}} \sigma_{c, b} \leq 1    & \hspace{-20mm}\forall c \in \mathcal{C}, b \in \mathcal{B}                                       \\
+    \label{mod:c29}        & w^1_k \geq  M_T + H_k - v_{e_k}                                                                                                                                                                                                                                                                  & \hspace{-20mm} \forall k \in K                                                                   \\
+    \label{mod:c30}        & w^2_k \geq (v_{e_k} - H_k) - N_T                                                                                                                                                                                                                                                                 & \hspace{-20mm} \forall k \in K                                                                   \\
+    \label{mod:c31}        & w_{e_k} \leq H_k + O_T                                                                                                                                                                                                                                                                           & \hspace{-00mm}  k \in K \hspace{0mm}                                                             \\
+    \label{mod:c32}        & 0 \leq q^k_j \leq Q^k                                                                                                                                                                                                                                                                            & \hspace{-20mm}  j \in \mathcal{D}, k \in K                                                       \\
+    \label{mod:c33}        & x^k_{ij} \in \{0,1\}                                                                                                                                                                                                                                                                             & \hspace{-20mm}  (i,j) \in A , k \in K  \hspace{0mm}                                              \\
+    \label{mod:c34}        & u_{ij} \in \{0,1\}                                                                                                                                                                                                                                                                               & \hspace{-20mm}  (i,j) \in A^D , k \in K  \hspace{0mm}                                            \\
+    % \label{mod:c35}        & \sigma_{ib} \in \{0,1\}    & \hspace{-20mm}  i \in \mathcal{C}, b \in \mathcal{B}                                             \\
+    \label{mod:c36}        & y_o \in \{0,1\}                                                                                                                                                                                                                                                                                  & \hspace{-20mm}  o \in \mathcal{O}                                                                \\
+    \label{mod:c37}        & v_i \geq 0,  w_i \geq 0                                                                                                                                                                                                                                                                          & \hspace{-20mm}  i \in V                                                                          \\
+    \label{mod:c38}        & w^1_{k} \geq 0,  w^2_k \geq 0                                                                                                                                                                                                                                                                    & \hspace{-20mm}  i \in V                                                                          \\
+    \label{mod:cend}       & g_i \geq 0                                                                                                                                                                                                                                                                                       & \hspace{-20mm}  i \in \mathcal{C}.
+\end{alignat}
+
+$\beta_1$ to $\beta_4$ are the penalty coefficients of each component of the objective function (\ref{mod:obj}).
+Constraints (\ref{mod:c1}) and (\ref{mod:c2}) state that driver $k$ leaves his start node exactly once a day and returns to his end node.
+A driver cannot serve a node before the start of his shift with constraints (\ref{mod:c03}).
+Constraints (\ref{mod:c3}) and (\ref{mod:c4}) set a driver to take some time after loading or unloading to adjust the concrete or clean the truck before moving on to the next node.
+The duration of the loading operation depends on the plant's loading rate and the amount $q^k_j$ of RMC loaded (\ref{mod:c5}).
+Similarly, the unloading service depends on the site's rate and $q^k_j$ (\ref{mod:c6}).
+Unloading operations must end at most $\Delta$ minutes after loading begins (\ref{mod:c7}).
+
+Constraints (\ref{mod:c8}) - - (\ref{mod:c10}) ensure that the first service of any customer $i$ must start after the due time $a_i$.
+This first service may be performed at the first delivery node of any order $\in O_i$, and one order must be completed before another is started.
+It also enforces the precedence constraints between the last delivery of an order and the first delivery of the following order.
+Constraints (\ref{mod:c11}) - - (\ref{mod:c15}) find the delivery sequence of all orders $\in \mathcal{O}_i$.
+With constraints (\ref{mod:c16}) - - (\ref{mod:c19}), two trucks cannot unload at the same time for consecutive deliveries of the same order.
+Additionally, these constraints impose a maximum time delay between the two trucks.
+Similarly, constraints (\ref{mod:c20}) - - (\ref{mod:c23}) ensure that two trucks cannot be loaded at the same time at a plant.
+Constraints (\ref{mod:c24}) require that the cumulative load of all concrete mixers serving an order must equal the required quantities, and (\ref{mod:c240}) bound a driver load.
+Constraints (\ref{mod:c25}) require that a driver can only visit a loading/delivery node once.
+Constraints (\ref{mod:c26}) are degree constraints.
+Constraints (\ref{mod:c29}) - - (\ref{mod:c30}) calculate the difference between a driver's hours of service and the minimum and normal hours of service.
+Finally, constraints (\ref{mod:c31}) - - (\ref{mod:cend}) define the nature and bounds of the variables.
+
+Let us consider the solution of a small instance of our problem shown in Figures~\ref{fig_Example} and \ref{fig:ganttExample}.
+Two batch plants $B_1$ and $B_2$ with three drivers serve two construction sites.
+Customer $C_1$ requests one order of $q^1_1=12$ m$^3$.
+Customer $C_2$ requests three orders of $q^1_2=3$, $q^2_2=11$, and $q^3_2=1$ m$^3$ of three different types of concrete.
+$B_1$ has two drivers ($D_1$, $D_2$) using concrete mixers of capacity 8 m$^3$, and $B_2$ has $D_3$ with a capacity 12 m$^3$.
+Figure~\ref{fig_Example} shows the flow of concrete mixers flow in the network.
+We schedule $D_1$ and $D_2$ to deliver 8 and 4 m$^3$ to $C_1$, respectively.
+We select $o^2_2$ as the first order of $C_2$.
+$D_3$ travels to $B_1$ to load $q^2_2$, then visits $C_2$ before returning to his home plant.
+Meanwhile, after their first trip,  $D_1$ and $D_2$ deliver the remaining orders of $C_2$.
+
+Figure~\ref{fig:ganttExample} shows the sequence of loading and unloading operations.
+The first deliveries of $C_1$ and $C_2$ are timely and there is no gap between the two deliveries received by $C_1$.
+There is a delay between the end of the first delivery and the start of the second delivery of $C_2$, however, it is less than the maximal time delay of 20 minutes defined for this example.
+
+\begin{figure}[htbp]
+    % \caption{Solution of an instance with two plants, two construction sites, four orders, and three drivers.}
+    \vspace*{-10mm}
+
+    \centering
+    \begin{adjustbox}{max width=0.85\textwidth}
+
+        \begin{tikzpicture}
+            \SetUpEdge[lw = 1.2pt,
+                labelcolor = white,
+                % labelstyle = {draw}
+            ]
+            \tikzstyle{LabelStyle}=[fill=white]
+
+            \SetGraphUnit{5}
+            \GraphInit[vstyle=Normal]
+            \tikzset{VertexStyle/.style = {shape = circle,text = black,minimum size = 24 pt,draw = black}}
+            \Vertex[L={$B_1$}]{D1}
+            \tikzset{VertexStyle/.style = {shape = circle,text = black,minimum size = 10 pt,draw = blue}}
+            \tikzset{VertexStyle/.style = {shape = rectangle,text = black,minimum size = 10 pt,draw = blue}}
+            \Vertex[x=2,y=5,L={11 m$^3$}]{C21}
+            \EA[unit=2.0,L={3 m$^3$}](C21){C22}
+            \EA[unit=2.0,L={1 m$^3$}](C22){C23}
+            \tikzset{VertexStyle/.style = {shape = rectangle,text = black,minimum size = 10 pt,draw = blue}}
+            \SOEA[L={$12m^3$}](D1){C1}
+            \tikzset{VertexStyle/.style = {shape = circle,text = black,minimum size = 24 pt,draw = black}}
+            \SOEA[L={$B_2$}](C23){D2}
+            % %  \SetVertexArt
+            \tikzset{VertexStyle/.style = {shape = circle,text = black,minimum size = 10 pt}}
+            \SO[unit=0.8,L={$\tau^u_1=$30 $m^3/h$, 07:00 AM}](C1){tau1}
+            \SO[unit=0.5,L={\textbf{$C_1$} }](tau1){o1}
+
+            \NO[unit=0.9,L={\textbf{$C_2$}}](C22){C2}
+            \SOEA[unit=0.4,L={\textcolor{white}{  } $\tau^u_2=30$ $m^3/h$, 07:25 AM}](C2){tau2}
+
+            \tikzset{VertexStyle/.style = {shape = rectangle,text = black,minimum size = 10 pt}}
+            \SOWE[unit=1.,L={$\tau^l_1=$ 70 $m^3/h$}](D1){cap1}
+            \SO[unit=0.5,L={\textbf{\textcolor{brown}{D1: 8 m$^3$}}}](cap1){driver11}
+            \SO[unit=0.5,L={\textbf{\textcolor{orange}{D2: 8 m$^3$}}}](driver11){driver12}
+            \SO[unit=1,L={$\tau^l_2= 87$ $m^3/h$}](D2){cap2}
+            \SO[unit=0.5,L={\textbf{\textcolor{blue}{D3: 12 m$^3$}}}](cap2){driver2}
+            %%%%%%% Edge
+            \tikzset{EdgeStyle/.style={->}}
+            \Edge[](C21)(C22)
+            \Edge[](C22)(C23)
+            \tikzset{EdgeStyle/.style={<->,bend left=0}}
+            \Edge[label=$22$](D1)(C1)
+            \tikzset{EdgeStyle/.style={<->}}
+            \Edge[label=$19$](D2)(C23)
+            \Edge[label=$34$](D2)(C1)
+            \tikzset{EdgeStyle/.style={->}}
+            \tikzset{EdgeStyle/.append style = {bend left = 8}}
+            \Edge[label=$17$](D1)(D2)
+            \Edge[label=$18$](D2)(D1)
+            \tikzset{EdgeStyle/.append style = {bend left = 65}}
+            \Edge[label=$6$](D1)(C21)
+            \tikzset{EdgeStyle/.append style = {bend right = 50}}
+            \Edge[label=$5$](C21)(D1)
+            \tikzset{EdgeStyle/.style = {->,bend left = 20,draw=brown}}
+            \Edge[label=$D1$](D1)(C1)
+            \Edge[label=$D1$](C1)(D1)
+            \tikzset{EdgeStyle/.style = {->,bend left = 37,draw=orange}}
+            \Edge[label=$D2$](C1)(D1)
+            \Edge[label=$D2$](D1)(C1)
+
+            \tikzset{EdgeStyle/.style = {->,bend left= 20,draw=brown}}
+            \Edge[label=$D1$](D1)(C22)
+            \tikzset{EdgeStyle/.style = {->,bend left= 0,draw=brown}}
+            \Edge[label=$D1$](C22)(D1)
+
+            \tikzset{EdgeStyle/.style = {->,bend left = 0,draw=orange}}
+            \Edge[label=$D2$](C23)(D1)
+            \tikzset{EdgeStyle/.style = {->,bend left = -15,draw=orange}}
+            \Edge[label=$D2$](D1)(C23)
+
+            \tikzset{EdgeStyle/.style = {->,draw=blue}}
+            \Edge[label=$D3$](D2)(D1)
+            \tikzset{EdgeStyle/.style = {->,bend left = 15,draw=blue}}
+            \Edge[label=$D3$](D1)(C21)
+            \tikzset{EdgeStyle/.style = {->,bend left = 80,draw=blue}}
+            \Edge[label=$D3$](C21)(D2)
+        \end{tikzpicture}
+    \end{adjustbox}
+    \vspace*{-10mm}
+    % }
+    \caption{Solution of an instance with two plants, two construction sites, four orders, and three drivers.}
+    \label{fig_Example}
+
+\end{figure}
+
+\begin{figure}[htbp]
+    \centering
+    \vspace*{-0mm}
+    \begin{adjustbox}{max width=0.85\textwidth}
+        \begin{tikzpicture}[-latex]
+
+            \matrix (chart)
+            [
             base/.style = {shape=rectangle, 
@@
   inner sep=0em, rounded corners=1mm,bottom color=red!80,},
-  dummy/.style    = {circle,draw}
-}
-\newcolumntype{H}{>{\setbox0=\hbox\bgroup}c<{\egroup}@{}}
-
-\newtheorem{theorem}{Theorem}[section]
-\newtheorem{lemma}[theorem]{Lemma}
-\newtheorem{proposition}[theorem]{Proposition}
-\newtheorem{corollary}[theorem]{Corollary}
-\newtheorem{definition}[theorem]{Definition}
-\newcommand{\pcmaxc}{P$|\mbox{\em cont}|$C$_{\max}$\xspace}
-
-\newcommand{\GRASP}{Greedy randomized adaptive search procedure\xspace}
-\newcommand{\F}{$\mathcal{F}$\xspace}
-\newcommand{\B}{$\mathcal{B}$\xspace}
-\newcommand{\C}{$\mathcal{C}$\xspace}
-\newcommand{\LB}{$\mathcal{L}$\xspace}
-\newcommand{\Lb}{$\mathcal{L}_b$\xspace}
-\newcommand{\D}{$\mathcal{D}$\xspace}
- \newcommand*{\red}{\textcolor{red}}
-
-\renewcommand{\algorithmicforall}{\textbf{for each}}
-\SetKwFor{ForEach}{for each}{}{}%
-\renewcommand{\algorithmiccomment}[2][1\linewidth]{%
-\leavevmode\hfill\makebox[#1][l]{/* \textbf{~#2} */}}
-\renewcommand{\algorithmicrequire}{\textbf{Input:}}
-
-\renewcommand{\tabcolsep}{4pt}
-
-\modulolinenumbers[1]
-\linenumbers
-
-\oddsidemargin=0.5in
-\topmargin=-0.25in
-\textwidth=5.5in
-\textheight=8.25in
-
-%\journal{European Journal of Operational Research}
-
-%\begin{frontmatter}
-
-\title{A GRASP algorithm for the concrete delivery problem. }
-\author{Ousmane Ali$^{(1)}$, Jean-Fran\c cois C\^ot\'e$^{(2)}$, Leandro C.~Coelho$^{(3)}$\\
- $(1)$ {\tt nassoma-wattara-ousmane.ali.1@ulaval.ca}\\
- $(2)$ {\tt Jean-Francois.Cote@fsa.ulaval.ca}\\
- $(3)$ {\tt Leandro.Coelho@fsa.ulaval.ca}\\
-}
-
-% \date{\today}
-
-\begin{document}
-\maketitle
-\begin{abstract}
-    This paper addresses a novel variant of the Concrete Delivery Problem (CDP), which involves the efficient scheduling of ready-mixed concrete deliveries to construction sites while balancing the conflicting goals of minimizing transportation costs and maximizing customer satisfaction.
-In this study, we propose an exact formulation and a heuristic approach based on the Greedy Randomized Adaptive Search Procedure (GRASP), to tackle this challenging CDP variant.
-This variant introduces realistic side constraints, including driver working shifts, a minimum driver working time, and overtime penalties.
-Additionally, it considers the scenario where customers may request multiple types of concrete delivered within the same time window.
-We assess the performance of our heuristic using new public instances generated for this problem and provide a comparative analysis with another CDP variant to demonstrate its effectiveness.
-\end{abstract}
-
-% \begin{keywords}
-\noindent{\textbf{Keywords:} Vehicle scheduling;  concrete delivery; GRASP; ready-mixed concrete }
-% \end{keywords}
-
-%\end{frontmatter}
-
-\section{Introduction}
-\label{sec:intro}
-Concrete is a widely used building material in construction projects.
-Its perishable nature is affected by many factors that impact its quality \citep{sinha_quality_2021}, which is crucial for the durability and strength of the final construction.
-Concrete comes in two types: ready-mixed concrete (RMC) and site-mixed concrete (SMC).
-RMC is manufactured in a batch plant and delivered to the construction site, while SMC is produced on-site using raw materials stored on the construction site.
-Using SMC can avoid delays caused by road traffic, but it has a slower and more difficult production process, requires storage for mixing materials and equipment, and is suitable for low amounts of concrete.
-On the other hand, RMC has better quality and benefits from lower production costs \citep{muresan_comparing}.
-However, to take advantage of these benefits, the batch plant manager must ensure efficient and prompt delivery on the construction site, which may require a fleet of high-cost revolving drum trucks (concrete mixers) to dispatch the RMC.
-
-Concrete delivery under the form of RMC is subject to many operational constraints that make the Concrete Delivery Problem (CDP) very challenging.
-In this paper, we study a variant of the CDP to schedule the daily production and dispatching of RMC for a company located in the province of Quebec, Canada.
-This company operates multiple batch plants with varying production rates, using a fleet of concrete mixers of different capacities.
-Each plant has its own fleet of trucks; however, under certain conditions, the trucks can move between plants if necessary.
-The trucks must return to their home plant at the end of the day.
-The company owns two types of trucks with different capacities and can call on an external fleet when needed.
-They serve construction sites from any of their plants, with the first delivery starting at the time specified by the customer.
-The loading and unloading of a concrete mixer depend on the truck capacity, the loading rate at the plant, and the unloading rate at the construction site.
-Drivers are allocated based on their daily work schedules.
-A customer may request several types of concrete to be delivered within the same time window, with no required sequence for the orders, but an order can only start after the completion of the previous order.
-This constraint generalizes the linked order constraints of \cite{durbin2008or} where some customers place two orders for the same day and request that they are linked (the second order begins only after the first order is completed).
-The setting of our study is similar to a variant of the CDP previously studied in \cite{schmid2009hybrid, schmid2010hybridization}.
-However, we include the plant's production rate and driver shift schedule.
-The company uses a centralized dispatcher system to schedule all daily orders, but this system has issues satisfying all daily demands without using an external fleet or using different plants to serve the same order.
-% Also, as per union rules, they assign a truck driver to a delivery task based on his seniority.
-
-According to \cite{blazewicz2019handbook}, the CDP combines vehicle routing with scheduling issues to plan routes to deliver concrete from batch plants (depots) to customers' construction sites.
-RMC is an on-demand product with a short life cycle from production through end use.
-It cannot be stored and cannot stay too long on a truck, or it will harden.
-Hence, concrete mixers must deliver RMC at the planned construction site shortly after its production.
-\cite{tommelein1999just} describes RMC production and delivery as an example of a just-in-time (JIT) production system in construction.
-A customer quantity requirement is often greater than the truck size and must be fulfilled by multiple deliveries.
-In that sense, CDP is similar to the vehicle routing problem with split delivery \citep{archetti2008split}, except that the same truck may visit a customer more than once.
-Concrete hardens quickly, so multiple deliveries must be done back-to-back or at least close in time to avoid the problem of cold joint, which can reduce the strength and durability of the concrete.
-Customers request to be served within a specific time window, which can complicate truck-loading schedules when a plant can only load one truck at a time.
-Similarly, only one truck can unload at a time at a customer location, sometimes leading to concrete mixers queuing and waiting their turn to deliver.
-Furthermore, with trucks of varied sizes, loading, travel, and unloading times that may be uncertain, the CDP is a complex and challenging problem.
-
-In this paper, we propose a mathematical model and a Greedy Randomized Search Procedure (GRASP) heuristic to solve a new variant of the CDP.
-Our model takes into consideration the working shifts of the drivers and the scheduling of multiple orders within the same time window at a construction site.
-To the best of our knowledge, this paper is the first that deals with these specific constraints in the context of the CDP.
-
-The rest of the paper is organized as follows: Section \ref{lit_review} provides a literature review of previous work related to the CDP.
-Section \ref{desc_form} provides a formal description and mathematical model of the problem.
-Section \ref{grasp_method} describes the GRASP algorithm and the constructive heuristics developed to solve this variant of the CDP.
-Section \ref{comp_exp} presents computational experiments and sensitivity analyses to evaluate the proposed approach, and finally, section \ref{concl} presents the conclusions.
-
-\section{Literature review}
-\label{lit_review}
-
-Academic Research on concrete batch and delivery began in the late 1990s.
-\cite{tommelein1999just} described RMC as a prototypical example of a JIT production system in construction and identified two practices for delivering it.
-One approach is for the customer to haul the product from the batch plant with their concrete mixer, while the other is for the batch plant to deliver the concrete directly to the customer's location.
-This latter approach is the one that has been studied in all related papers found in the literature.
-Several works to schedule and dispatch concrete production and delivery have mainly focused on simulation-related methods.
-These methods can be standalone, such as those used by \cite{zayed2001simulation, wang2001scheduling, tian_simulation_based_2010, panas_simulation_based_2013} and \cite{galic2016simulation}, among others.
-Alternatively, these methods can be hybridized with optimization techniques, such as those used by \cite{feng2004optimizing, lu2005optimized}, and \cite{feng_integrating_2006}.
-\cite{wang2001scheduling} developed a simulation model to reveal the effect and value of the concrete mixers' inter-arrival time on the productivity of hired unloading equipment on site.
-\cite{feng2004optimizing} used a combination of genetic algorithm (GA) and simulation process to minimize the total waiting time for trucks at a customer site.
-The study focused on loading trucks with identical capacities at the same batch plant, with fixed loading and unloading durations.
-The GA was used to find the best loading sequence of RMC trucks to be assigned to different construction sites.
-The simulation process determined the loading, arrival, departure, and waiting time of trucks and thus evaluated the cost of each dispatching sequence.
-They evaluated their method using data from a batch plant in Taiwan with up to nine customers served.
-\cite{mayteekrieangkrai2015optimized} addressed the same problem with the same data using a bee algorithm (BA) and found better solutions than the GA.
-\cite{lu2005optimized} used the same combination of GA and simulation to determine the optimal number of concrete mixers to be deployed and an optimal schedule for batching and delivering concrete.
-Their objective was to minimize the idle time of the site crew due to late concrete deliveries and truck queuing time.
-In this setting, it was also necessary to deliver a batch of mortar on-site to lubricate the unloading pump before the concrete delivery.
-As such, the simulation model also included the batch and delivery of mortar.
-Finding the best RMC fleet size was also the purpose of the discrete-event simulation model proposed by \cite{panas_simulation_based_2013}.
-
-In addition to simulation-based methods, several other approaches have been used in the literature to solve the CDP.
-These include metaheuristics \citep{faria2006distributed, misir2011selection, maghrebi2016sequential, yang2022concrete}, exact methods \citep{yan2007optimal, asbach2009analysis, kinable2014concrete}, matheuristics \citep{schmid2009hybrid, schmid2010hybridization}, Benders Decomposition \citep{maghrebi2014benders}, column generation (CG) \citep{maghrebi2014solving, maghrebi2016column}, Lagrangian relaxation \citep{narayanan2015using}, and machine learning approaches such as those used by \cite{graham2006modeling, maghrebi2014exploring, maghrebi2016matching}.
-\cite{matsatsinis2004towards} designed a decision support system (DSS) for the dynamic routing of both concrete and pumps that may be necessary for some construction sites to aid in the unloading of concrete.
-The DSS considered the availability of three plants but stipulated that vehicles fulfilling the same order must all load at the same plant.
-Orders that could not be executed immediately could be postponed for the next day.
-The routing of the pumps was modeled as a multi-depot vehicle routing problem with time windows.
-\cite{naso2007genetic} proposed a sequential GA method combined with constructive heuristics to solve another variant of the CDP.
-In this problem, the plant's production schedule must account for orders for concrete to be delivered to a customer site and orders that customers must pick up themselves.
-The algorithm first schedules the plant loading operations before scheduling truck deliveries.
-The authors also developed a non-linear model that minimizes transportation costs, waiting times, outsourced costs, and overtime work.
-They ran experiments using real-world instances of a concrete supply chain in the Netherlands and found a reduction in the number of outsourced requests.
-\cite{yan2007optimal} also considered overtime considerations in their paper, which focused on scheduling RMC for one batch plant with two loading docks.
-The study took into account that overtime wages are paid for factory and construction site operations after 4 PM.
-They developed a mixed-integer programming (MIP) model on a time-space network to minimize travel times and operating costs at both normal and overtime working hours at the plant and the construction sites.
-They tested the model using real data consisting of three days of operation using a two-stage algorithm.
-First, they solved the MIP relaxation with CPLEX.
-Then, they simplified the original model by fixing some decision variables before solving it.
-The algorithm was found to improve the actual plant operation by 10\%.
-A time-space network is the key component of the real-time DSS developed by \cite{durbin2008or} to solve a dynamic CDP every five minutes.
-The DSS can receive new orders, schedule them on the fly, and handle unexpected events such as plant closures, truck breakdowns, and delays in transportation times.
-The authors combined the DSS with a tabu search (TS) heuristic to warm start CPLEX, which made the model performant enough to solve instances with up to 1,500 loads per day with up to 250 trucks.
-The DSS also considers the case of a customer who places two orders, with the first being completed before the second starts.
-Further insights on the real-time planning and monitoring of CDP are available in \cite{garza2021dynamic}.
-Another variant of the CDP is modeled by \cite{schmid2009hybrid} as an integer multicommodity network flow (MCNF) problem on a time-space network.
-In this paper, concrete is delivered using a heterogeneous fleet of vehicles, and each plant can load an unlimited number of trucks simultaneously.
-Some of the trucks have specialized equipment and must arrive first at certain construction sites to assist in unloading the concrete.
-The objective is to fulfill all orders, minimize the travel cost, and avoid delays between two consecutive unloading operations for an order.
-The model is typically solved using a matheuristic algorithm that combines the MCNF with a variable neighborhood search (VNS) heuristic.
-The method can quickly solve large problem instances with more than 60 orders per day without encountering any memory issues.
-The same problem is addressed by \cite{schmid2010hybridization}, who proposed a MIP model combined with a VNS and a very large neighborhood search (VLNS) to develop two matheuristics approaches.
-Comparisons between both matheuristics and a standalone VNS show that the former methods are much better and suitable for solving larger problem instances.
-These methods also provide better solutions for small to medium instances than the matheuristic used in \cite{schmid2009hybrid}.
-A pure VNS approach with the same problem but without the use of instrumentation has been applied by \cite{payr2009optimizing}.
-
-Regarding objectives, most authors have focused on minimizing travel time and delays between consecutive deliveries.
-However, some authors have been more interested in maximizing customer satisfaction alone.
-We find these situations in the works of \cite{durbin2008or, kinable2014concrete, kinable2014logic, sulaman2017simulated}.
-\cite{kinable2014concrete} introduce a general MIP and constraint programming (CP) models of the CDP reflecting the main constraints commonly found in all CDP works: time lag and no overlapping between consecutive deliveries, covering of all customers' demands, delivery time window, and heterogeneous fleet.
-However, the model did not include constraints limiting the time that concrete may reside in a truck.
-The authors propose a constructive heuristic that schedules the visits to the customers one by one according to the start time of the visit and the truck capacity.
-The procedure is invoked multiple times for different permutations of the customer's order which is determined using the steepest descent (SD) local search procedure.
-One of the paper's main contributions is the creation of the first public test instances for the CDP with up to 50 customers, four batch plants, and 20 concrete mixers.
-They found the CP model to be highly effective in finding high-quality solutions in a relatively short time or improving existing schedules, while the MIP model can be used to compute bounds, as it seems ineffective in solving large problem instances.
-Finally, the heuristic often yields good solutions in less than a second.
-A detailed analysis of the MIP model presented in \cite{kinable2014concrete} and of two more compact models can be found in the thesis of \cite{hernandez_lopez_study_2020}.
-In \cite{kinable2014logic}, we found an attempt to solve the previous problem with a logic-based Benders' approach.
-%\cite{sulaman2017simulated} expand upon the SD-heuristic proposed in \cite{kinable2014concrete}, proposing a simulated annealing (SA) combined with a time-slot Heuristic (SATH).
-This method looks for a slot between existing visits of a truck to schedule a new delivery instead of assigning it to the time slot strictly after the truck's latest assigned delivery.
-The goal is to reduce the large time gaps that can be present in a schedule created with SD due to ignoring the intermediate available time slots.
-Experimental results indicated that SATH outperforms SD in speed and solution quality.
-A generalization of the MIP model of \cite{kinable2014concrete} is addressed in \cite{asbach2009analysis}.
-This model simultaneously minimizes the total sum of travel costs and the penalty costs for customers with unfulfilled demand.
-A customer can request that all concrete deliveries come from the same plant or a subset of plants and that a delivery truck belongs to a subset of the vehicle fleet.
-The MIP model is used in a local search scheme as a black-box solver to reoptimize an incumbent solution in which a neighborhood operator has unfixed some variables.
-\cite{tzanetos2023systematic} provide an overview of the various methods used in the literature to address the CDP and categorizes the problem formulations based on the different concepts used in the literature.
-They also discussed the consistency between industry needs and existing constraints and provided insights into the datasets corresponding to real-world cases, identifying the necessary data for practitioners.
-
-\section{Problem description}
-\label{desc_form}
-The focus of this paper is on the distribution of RMC from a Canadian company that operates in the greater Montreal area.
-When a customer places an order, it is received at a control center and immediately assigned to one of the company's batch plants.
-These plants produce the concrete and then deliver it to the customer.
-The problem we are examining involves a set of customer orders, a set of concrete-mixer drivers, and a set of batch plants.
-
-% \subsection*{Customer orders}
-
-A customer $i$ requests one or more types of concrete to be delivered to their construction site on a specific day, with the delivery service starting at the due time $a_i$.
-We call an order $o$ a request for a specific type of concrete.
-$q_i$ is the sum of the demands $q_o$ of each order $o$ placed, $a_i$ is the desired arrival time of the first concrete mixer, and $\tau^u_i$ is the unloading rate for customer $i$.
-If the order requires more concrete than a single truck can carry, multiple deliveries are scheduled.
-We will accept a first delivery if it arrives within a time window of $\left[a_i, a_i + \tau^w \right]$, where $\tau^w_i$ is a user-defined parameter.
-
-Let $O_i$ be the set of all orders requested by customer $i$.
-Each element of $O_i$ must be fully delivered before moving on to another order.
-Exactly one order $o \in O_i$ must have its first delivery start at $a_i$, while the others can start at most $\gamma^2$ time after $o$ is completed.
-A plant is assigned to an order, and it must be the supplier of all subsequent deliveries of that order.
-To avoid cold joint problems with the concrete, subsequent deliveries of the same order must be made in close succession.
-We define a maximum time delay $\gamma^1$ after which no more deliveries are allowed.
-The customer's unloading rate and the quantity to be unloaded give the time required to unload a truckload.
-Let $\mathcal{C}$ be the set of construction sites (customers) with a planned delivery for the day, and $\mathcal{O}=\{{O_i, i \in \mathcal{C}}\}$ be the set of all requested orders for all customers.
-
-% \subsection*{Drivers}
-
-The company has two types of concrete mixer trucks with capacities of 8 and 12 cubic meters.
-Each driver $k$ is assigned to a particular batch plant and is responsible for driving a truck with capacity $Q_k$.
-The set of drivers is represented by $K =\cup_{b \in \mathcal{B}}K_b $, where $K_b$ is the set of drivers scheduled to start their shift at batch plant $b$.
-$t_{ij}$ is the known time to travel between any two locations $i$ and $j$.
-A scheduled driver $k$ is required to start his shift at $H_k$, work a minimum of $M_T$ hours and a maximum of $N_T$ hours during regular working hours, with the possibility of overtime of up to $O_T$ hours.
-$\beta_3$ and $\beta_4$ are the penalties for a driver working less than $M_T$ and more than $N_T$.
-
-A driver typically loads RMC at his assigned batch plant but may be required to drive to and load at other plants if needed.
-The batch plant produces concrete on demand using recipes specific to each order.
-This means that a truck can only haul RMC for one order, even if there is spare capacity.
-After unloading the RMC,  driver $k$ takes $\rho$ minutes to clean the concrete mixer before proceeding.
-%When assigning drivers to deliveries, the company prioritizes employees with the highest seniority.
-
-Let $n_o$ be the number of deliveries needed to fulfill the order $o$.
-$n_o$ is not known in advance because we use a fleet of trucks with different capacities.
-However, we can compute its lower ($n_o^{min}$) and upper ($n_o^{max}$) bounds using the capacities of the largest ($Q_{max}$) and smallest ($Q_{min}$) available trucks.
-\begin{alignat}{3}
-    \label{mod:c0}
-    n_o^{min} = \left\lceil \frac{q_o}{Q_{max}} \right\rceil \leq n_o \leq n_o^{max} = \left\lceil \frac{q_o}{Q_{min}} \right\rceil & \text{ } &
-    \forall  o \in \mathcal{O}.
-\end{alignat}
-
-Let $d^j_{o}$ be the $j^{th}$ visit with load $q^j_{o}$ for order $o$.
-We represent the fulfillment of order $o$ by the visits to the ordered set of delivery nodes $\mathcal{D}_o= \left(d^0_{o},d^1_{o},\cdots, d^{n_o}_{o}\right)$.
-The deliveries of customer $i$ are the ordered set $\mathcal{D}_i= (\mathcal{D}_{o_1}, \mathcal{D}_{o_2},\cdots,\mathcal{D}_{o_{|O_i|}})$, where $o_r$ is the $r^{th}$ delivered order.
-We will refer to $d \in \mathcal{D}_i$ ($d \in \mathcal{D}_o$) as the $d^{th}$ potential delivery of customer $i$ (order $o$).
-$\mathcal{D}=\bigcup_{i\in \mathcal{C}} \mathcal{D}_i$ is the union of all delivery nodes.
-
-% \subsection*{Batching plants}
-
-Each of the company's plants has a single loading dock that can accommodate only one truck at a time.
-As a result, trucks often form a queue while waiting for their turn at the loading dock.
-Let $\mathcal{B}$ be the set of batch plants.
-The plants are heterogeneous, as each plant $b$ has its hourly loading rate, represented by $\tau^l_b$, which affects the duration of the loading process.
-After loading the concrete, the driver spends $\alpha_b$ minutes adjusting the concrete in the truck before heading to the customer site.
-Each plant has its own assigned fleet of trucks, but it can borrow trucks from other plants.
-Let $l_{j}$ be the loading dock node associated with delivery node $j$.
-After loading RMC at $l_j$, it must be fully delivered to $j$ at most before $\Delta$ time, which is the concrete lifetime.
-We denote $\mathcal{L}_b$ as the set of loading dock nodes of plant $b$ and $\mathcal{L}$ as the set of all loading docks.
-
-% \subsection*{Model}
-A solution to the problem involves decisions about truck loading schedules, driver assignments to different deliveries, and truck arrival times at construction sites for unloading.
-For a batch plant, the decision involves choosing which driver to load, when to load them, how much to load, and which construction site to deliver.
-For a driver, the decision is to determine the sequence of loading depots and delivery sites.
-And for a construction site, the decision involves determining the arrival times of all scheduled deliveries for the day.
-
-Each driver leaves and returns to their home plant every day.
-We represent the home plant of a driver $k$ with a starting depot $s_k$ and an ending depot $e_k$.
-$S$ and $E$ are the sets of starting and ending depots, respectively.
-
-We define our problem on a complete directed graph where $V=\{ S \cup \mathcal{L} \cup \mathcal{D} \cup E\}$ is the set of nodes.
-The arc sets are $A =  \{(i,j,k) \hspace*{1mm} \vert \hspace*{1mm} i, j \in V \hspace{1mm} k\in K \}$, $A^D = \{(i,j) \hspace*{1mm} \vert \hspace*{1mm} i, j \in \mathcal{D} \hspace{1mm} \}$, and $A^L = \{(i,j) \hspace*{1mm} \vert \hspace*{1mm} i, j \in \mathcal{L} \hspace{1mm} \}$.
-$A$ corresponds to allowed movements of drivers from node $i$ to node $j$. For each driver $k$, the allowed movements are the following:
-
-\begin{itemize}
-    \item From the starting depot $s_k$ to a loading dock $l \in \mathcal{L}$ or to the ending depot $e_k$.
-    \item From a loading dock $l \in \mathcal{L}$ to a delivery node $d \in \mathcal{D}$.
-    \item From a delivery node  $d \in \mathcal{D}$ to a loading dock $l \in \mathcal{L}$ or to the ending depot $e_k$.
-\end{itemize}
-
-For a customer $c$, arcs in $A^D$ link consecutive delivery nodes of the same order   $\lbrace (i,j)\in \mathcal{D}_o, o \in \mathcal{O}_c, i < j  \rbrace$, and pair of delivery nodes of two different orders $ \lbrace (i,d^{0}_{o_2}),  i \in \mathcal{D}_{o_1}, i \geq n^{min}_{o_1}, o_1, o_2 \in \mathcal{O}_c, o_1 \neq o_2 \rbrace $.
-Arcs in $A^L$ link all pairs of loading docks of the same batch plant.
-
-We define $\delta^{+}(i) = \{(i, j,k) \in A \}$ and $\delta^{-}(i) = \{(j, i,k) \in A \}$ as the outcoming and incoming arc sets of node $i \in V$; $\delta^{+}_D(i) = \{(i, j)  \in A^D\}$ and $\delta^{-}_D(i) = \{(j, i) \in A^D \}$ are the outcoming and incoming arc sets of delivery node $i \in \mathcal{D}$.
-Similarly, $\delta^{+}_L(i) = \{(i, j)  \in A^L\}$ and $\delta^{-}_L(i) = \{(j, i) \in A^L \}$ are the outcoming and incoming arc sets of loading node $i \in \mathcal{L}$.
-
-Let the binary variable $x^{k}_{ij}$ be $1$ if driver $k$ travels from node $i$ to $j$.
-Binary variable $y_o$ is $1$ when order $o$ is completely served; $v_i$ and $w_i$ are the start and end of the loading (unloading) operation at node $i \in \mathcal{L}\cup \mathcal{D}$.
-Binary variable $u_{ij}$ is $1$ if node $j$ is served just after $i$, the service being either an unloading or a loading operation.
-% Binary variable $\sigma_{ib}$ is equal to 1 if the orders of customer $i$ are loaded by plant $b$.
-Variable $q^k_j$ is the quantity to be loaded towards $j$ with vehicle $k$.
-Let $w^1_k$ be a continuous variable indicating the difference between the driver's work time and the minimum number of hours to be worked in a day, and $w^2_k$ indicating the difference between the driver's work time and the normal work time.
-Let $g_i$ be the time between the due date and the first service start for customer $i$.
-
-The objective function minimizes total travel cost ($TC$), penalties associated with unfulfilled orders, first delivery delays ($FDD$), driver underutilization costs ($DUC$), and driver overtime costs ($DOC$).
-Together, these components drive the optimization process to find a solution that efficiently balances travel costs, customer satisfaction, on-time delivery, driver utilization, and scheduling constraints.
-
-\begin{alignat}{3}
-    \label{mod:obj}   \min & \smashoperator{ \sum_{ \substack{(i,j,k) \in A} }} {t_{ij}x^{k}_{ij}} + \beta_1\smashoperator{\sum_{ o \in \mathcal{O}}}{  \left(1-y_o\right)} \hspace{-0mm}  + \beta_2 \smashoperator{ \sum_{i \in \mathcal{C} } } { g_i  } +  \sum_{k \in K}\left({\beta_3 w^{1}_k  + \beta_4  w^{2}_k}\right) &                                                                                                  \\
-    \label{mod:c1}         & \sum_{j \in \delta^{+}(i) }{x^k_{ij}} =1                                                                                                                                                                                                                                                         & \forall i \in S, k \in K                                                                         \\
-    \label{mod:c2}         & \sum_{j \in \delta^{-}(i)}{x^k_{ji}} = 1                                                                                                                                                                                                                                                         & \forall i \in E, k \in K                                                                         \\
-    \label{mod:c03}        & v_j \geq  w_i  - M\left(1-x^{k}_{ij}\right)                                                                                                                                                                                                                                                      & \hspace{-20mm}  i \in S, j \in \delta^{+}(i),  k \in K                                           \\
-    \label{mod:c3}         & v_j \geq  w_i + \alpha_{b} + t_{ij} - M\left(1-x^{k}_{ij}\right)                                                                                                                                                                                                                                 & \hspace{-20mm} \forall b \in \mathcal{B}, i \in \mathcal{L}_b, j \in \delta^{+}(i),  k \in K     \\
-    \label{mod:c4}         & v_j \geq  w_i + \rho + t_{ij} - M\left(1-x^{k}_{ij}\right)                                                                                                                                                                                                                                       & \hspace{-20mm} \forall i \in \mathcal{D}  , j \in \delta^{+}(i), k \in K                         \\
-    \label{mod:c5}         & w_{i} \geq v_{i}  + \frac{q^k_j}{\tau^l_b} -M\left(1- x^{k}_{ij}\right)                                                                                                                                                                                                                          & \hspace{-20mm} \forall  b \in \mathcal{B},  i \in \mathcal{L}_{b},  j \in \delta^{+}(i), k \in K \\
-    \label{mod:c6}         & w_{j} \geq v_{j}  + {  \frac{q^k_j}{\tau^u_c} -M\left(1- x^{k}_{ij}\right) }                                                                                                                                                                                                                     & \hspace{-20mm}  \forall c \in  \mathcal{C}, j \in \mathcal{D}_{c}, i \in \delta^{-}(j),  k \in K \\
-    \label{mod:c7}         & w_{j} \leq v_{i}  + \Delta + M\left(1- x^{k}_{ij}\right)                                                                                                                                                                                                                                         & \hspace{-20mm}   j \in \mathcal{D},  i \in \delta^{-}(j),  k \in K                               \\
-    \label{mod:c8}         & v_{d^0_{o}} \geq a_i    & \hspace{-20mm} \forall  i \in \mathcal{C}, \forall o \in \mathcal{O}_i                           \\
-    \label{mod:c9}  & g_{i} \geq v_{d^0_{o_1}} - a_i - M\left(  \sum_{ j \in \delta^{-}_{D}(d^0_{o_1})  }{u_{jd^0_{o_1}}}\right)  & \hspace{-20mm}  \forall  i \in \mathcal{C}, \forall o_1 \in \mathcal{O}_i   \\
-    \label{mod:c91}  & g_{i} - a_i \leq \tau^w  & \hspace{-20mm}  \forall  i \in \mathcal{C}  \\
-    \label{mod:c101}       & v_{d^0_{o_1}} \geq w_j - M\left(1- u_{jd^0_{o_1}}\right)   & \hspace{-40mm}    \forall o_1 \in \mathcal{O}_i,  j \in \delta_{D}^{-}(d^0_{o_1})                \\
-    \label{mod:c10}        & v_{d^0_{o_1}}  \leq  w_j + \gamma^2 + M\left(1- u_{j,d^0_{o_1}}\right)                                                                                                                                                                                                                           & \hspace{-20mm}  \forall  o_1 \in \mathcal{O},  j \in \delta_{D}^{-}(d^0_{o_1})                   \\
-    \label{mod:c11}        & \sum_{ \substack{o_1 \in \mathcal{O}_i }}{ \sum_{ j \in \delta^{-}_{D}(d^0_{o_1})}}{u_{j,d^0_{o_1} } } = |\mathcal{O}_i|-1                                                                                                                                                                       & \hspace{-20mm} \forall  i \in \mathcal{C},|\mathcal{O}_i| > 1                                    \\
-    \label{mod:c12}        & \sum_{ \substack{o_1 \in \mathcal{O}_i }}\sum_{ j \in \delta^{+}_{D}(d^0_{o_1})}{u_{d^0_{o_1},j } } = |\mathcal{O}_i|-1                                                                                                                                                                          & \hspace{-40mm} \forall  i \in \mathcal{C}, |\mathcal{O}_i| > 1                                   \\
-    \label{mod:c13}        & \hspace{3mm} \smashoperator{ \sum_{ j \in \delta^{+}_{D}(d^0_{o_1})}}{u_{d^0_{o_1},j } } \leq 1                                                                                                                                                                                                  & \hspace{-20mm}  \forall o_1 \in \mathcal{O}                                                      \\
-    \label{mod:c14}        & \hspace{2mm} \smashoperator{ \sum_{ j \in \delta^{-}_{D}(d^0_{o_1})}}{u_{j,d^0_{o_1} } } \leq 1                                                                                                                                                                                                  & \hspace{-20mm} \forall o_1 \in \mathcal{O}                                                       \\
-    \label{mod:c15}        & \hspace{2mm} \smashoperator{ \sum_{ j \in \delta^{+}_{D}(d^0_{o_1})}}{u_{d^0_{o_1},j } } +\smashoperator{ \sum_{ j \in \delta^{-}_{D}(d^0_{o_1})}}{u_{j,d^0_{o_1} } } \geq 1                                                                                                                     & \hspace{-20mm} \forall o_1 \in \mathcal{O}                                                       \\
-    \label{mod:c16}        & v_{j} \geq w_{j-1} - M\left(1- u_{j-1,j}\right)                                                                                                                                                                                                                                                  & \hspace{-20mm}  \forall o \in \mathcal{O}, j \in \mathcal{D}_{o}, j \geq 1                       \\
-    \label{mod:c17}        & v_{j}  \leq  w_{j-1} + \gamma^1 + M\left(1- u_{j-1,j}\right)                                                                                                                                                                                                                                     & \hspace{-20mm} \forall i \in \mathcal{C}, \forall o \in O_i, j \in \mathcal{D}_{o}, j \geq 1     \\
-    \label{mod:c18}        & u_{j-1,j} \geq u_{j,j+1}                                                                                                                                                                                                                                                                         & \hspace{-35mm} \forall o_1 \in\mathcal{O},  j \in \mathcal{D}_{o_1}, 1 \leq j \leq n_{o_{1}}-1   \\
-    \label{mod:c19}        & u_{j-1,j} \geq \sum_{l \in \mathcal{L}}{x_{lj}}                                                                                                                                                                                                                                                  & \hspace{-20mm}  \forall o_1 \in \mathcal{O}, j \in \mathcal{D}_{o_1}, j \geq 1                   \\
-    \label{mod:c20}        & v_{j} \geq w_{i} - M\left(1- u_{i,j}\right)                                                                                                                                                                                                                                                      & \hspace{-20mm}  \forall i \in \mathcal{L}, j \in \delta^{+}_{L}(i)                               \\
-    \label{mod:c21}        & \hspace{3mm} \smashoperator{ \sum_{ j \in \delta^{+}_{L}(i)}}{u_{i,j } } \leq 1                                                                                                                                                                                                                  & \hspace{-20mm}  \forall i \in \mathcal{L}                                                        \\
-    \label{mod:c22}        & \hspace{2mm} \smashoperator{ \sum_{ j \in \delta^{-}_{L}(i)}}{u_{j,i } } \leq 1                                                                                                                                                                                                                  & \hspace{-20mm} \forall i \in \mathcal{L}                                                         \\
-    \label{mod:c230}       & \hspace{2mm} \smashoperator{ \sum_{ j \in \delta^{-}_{L}(i)}}{u_{j,i } }  \geq \sum_{k \in K}\sum_{j \in  \delta^{-}(i) }{x^k_{ji}}                                                                                                                                                              & \hspace{-20mm} \forall i \in \mathcal{L}                                                         \\
-    \label{mod:c231}       & \hspace{2mm} \smashoperator{ \sum_{ j \in \delta^{-}_{L}(i)}}{u_{j,i } } \geq \smashoperator{ \sum_{ j \in \delta^{-}_{L}(i+1)}}{u_{j,i+1 } }                                                                                                                                                    & \hspace{-20mm} \forall i \in \mathcal{L}                                                         \\
-    \label{mod:c23}        & \hspace{2mm} \smashoperator{ \sum_{ j \in \delta^{-}_{L}(i)}}{u_{j,i } } + \smashoperator{ \sum_{ j \in \delta^{+}_{L}(i)}}{u_{i,j } } \geq \sum_{k \in K}\sum_{j \in  \delta^{-}(i) }{x^k_{ji}}                                                                                                 & \hspace{-20mm} \forall i \in \mathcal{L}                                                         \\
-    \label{mod:c24}        & \sum_{k \in K}\sum_{j \in \mathcal{D}_o}{q^k_j} = q_o                                                                                                                                                                                                                                            & \hspace{-20mm}  \forall o \in \mathcal{O}                                                        \\
-    \label{mod:c240}       & q^k_j \leq \sum_{i \in  \delta^{-}(j)}{Q^k x^k_{ij}}                                                                                                                                                                                                                                             & \hspace{-20mm}  \forall j \in \mathcal{D}, k \in K                                               \\
-    \label{mod:c25}        & \sum_{k \in K}\sum_{j \in \delta^{+}(i)}{x^{k}_{ij} \leq 1}                                                                                                                                                                                                                                      & i \in \mathcal{L} \cup \mathcal{D}                                                               \\
-    \label{mod:c26}        & \sum_{k \in K}\sum_{j \in \delta^{+}(i)} {x^{k}_{ij} } =  \sum_{k \in K}\sum_{j \in \delta^{-}(i)}{x^{k}_{ji} }                                                                                                                                                                                  & i \in \mathcal{L} \cup \mathcal{D}                                                               \\
-    % \label{mod:c27}        & \sum_{k \in K}  \sum_{i \in \mathcal{D}_c} \sum_{j \in \delta^{-}(i) }  {x^k_{ji}} \leq M\sigma_{c, b}     & \forall c \in \mathcal{C}, b \in \mathcal{B}                                                     \\
-    % \label{mod:c28}        & \sum_{b \in \mathcal{B}} \sigma_{c, b} \leq 1    & \hspace{-20mm}\forall c \in \mathcal{C}, b \in \mathcal{B}                                       \\
-    \label{mod:c29}        & w^1_k \geq  M_T + H_k - v_{e_k}                                                                                                                                                                                                                                                                  & \hspace{-20mm} \forall k \in K                                                                   \\
-    \label{mod:c30}        & w^2_k \geq (v_{e_k} - H_k) - N_T                                                                                                                                                                                                                                                                 & \hspace{-20mm} \forall k \in K                                                                   \\
-    \label{mod:c31}        & w_{e_k} \leq H_k + O_T                                                                                                                                                                                                                                                                           & \hspace{-00mm}  k \in K \hspace{0mm}                                                             \\
-    \label{mod:c32}        & 0 \leq q^k_j \leq Q^k                                                                                                                                                                                                                                                                            & \hspace{-20mm}  j \in \mathcal{D}, k \in K                                                       \\
-    \label{mod:c33}        & x^k_{ij} \in \{0,1\}                                                                                                                                                                                                                                                                             & \hspace{-20mm}  (i,j) \in A , k \in K  \hspace{0mm}                                              \\
-    \label{mod:c34}        & u_{ij} \in \{0,1\}                                                                                                                                                                                                                                                                               & \hspace{-20mm}  (i,j) \in A^D , k \in K  \hspace{0mm}                                            \\
-    % \label{mod:c35}        & \sigma_{ib} \in \{0,1\}    & \hspace{-20mm}  i \in \mathcal{C}, b \in \mathcal{B}                                             \\
-    \label{mod:c36}        & y_o \in \{0,1\}                                                                                                                                                                                                                                                                                  & \hspace{-20mm}  o \in \mathcal{O}                                                                \\
-    \label{mod:c37}        & v_i \geq 0,  w_i \geq 0                                                                                                                                                                                                                                                                          & \hspace{-20mm}  i \in V                                                                          \\
-    \label{mod:c38}        & w^1_{k} \geq 0,  w^2_k \geq 0                                                                                                                                                                                                                                                                    & \hspace{-20mm}  i \in V                                                                          \\
-    \label{mod:cend}       & g_i \geq 0                                                                                                                                                                                                                                                                                       & \hspace{-20mm}  i \in \mathcal{C}.
-\end{alignat}
-
-$\beta_1$ to $\beta_4$ are the penalty coefficients of each component of the objective function (\ref{mod:obj}).
-Constraints (\ref{mod:c1}) and (\ref{mod:c2}) state that driver $k$ leaves his start node exactly once a day and returns to his end node.
-A driver cannot serve a node before the start of his shift with constraints (\ref{mod:c03}).
-Constraints (\ref{mod:c3}) and (\ref{mod:c4}) set a driver to take some time after loading or unloading to adjust the concrete or clean the truck before moving on to the next node.
-The duration of the loading operation depends on the plant's loading rate and the amount $q^k_j$ of RMC loaded (\ref{mod:c5}).
-Similarly, the unloading service depends on the site's rate and $q^k_j$ (\ref{mod:c6}).
-Unloading operations must end at most $\Delta$ minutes after loading begins (\ref{mod:c7}).
-
-Constraints (\ref{mod:c8}) - - (\ref{mod:c10}) ensure that the first service of any customer $i$ must start after the due time $a_i$.
-This first service may be performed at the first delivery node of any order $\in O_i$, and one order must be completed before another is started.
-It also enforces the precedence constraints between the last delivery of an order and the first delivery of the following order.
-Constraints (\ref{mod:c11}) - - (\ref{mod:c15}) find the delivery sequence of all orders $\in \mathcal{O}_i$.
-With constraints (\ref{mod:c16}) - - (\ref{mod:c19}), two trucks cannot unload at the same time for consecutive deliveries of the same order.
-Additionally, these constraints impose a maximum time delay between the two trucks.
-Similarly, constraints (\ref{mod:c20}) - - (\ref{mod:c23}) ensure that two trucks cannot be loaded at the same time at a plant.
-Constraints (\ref{mod:c24}) require that the cumulative load of all concrete mixers serving an order must equal the required quantities, and (\ref{mod:c240}) bound a driver load.
-Constraints (\ref{mod:c25}) require that a driver can only visit a loading/delivery node once.
-Constraints (\ref{mod:c26}) are degree constraints.
-Constraints (\ref{mod:c29}) - - (\ref{mod:c30}) calculate the difference between a driver's hours of service and the minimum and normal hours of service.
-Finally, constraints (\ref{mod:c31}) - - (\ref{mod:cend}) define the nature and bounds of the variables.
-
-Let us consider the solution of a small instance of our problem shown in Figures~\ref{fig_Example} and \ref{fig:ganttExample}.
-Two batch plants $B_1$ and $B_2$ with three drivers serve two construction sites.
-Customer $C_1$ requests one order of $q^1_1=12$ m$^3$.
-Customer $C_2$ requests three orders of $q^1_2=3$, $q^2_2=11$, and $q^3_2=1$ m$^3$ of three different types of concrete.
-$B_1$ has two drivers ($D_1$, $D_2$) using concrete mixers of capacity 8 m$^3$, and $B_2$ has $D_3$ with a capacity 12 m$^3$.
-Figure~\ref{fig_Example} shows the flow of concrete mixers flow in the network.
-We schedule $D_1$ and $D_2$ to deliver 8 and 4 m$^3$ to $C_1$, respectively.
-We select $o^2_2$ as the first order of $C_2$.
-$D_3$ travels to $B_1$ to load $q^2_2$, then visits $C_2$ before returning to his home plant.
-Meanwhile, after their first trip,  $D_1$ and $D_2$ deliver the remaining orders of $C_2$.
-
-Figure~\ref{fig:ganttExample} shows the sequence of loading and unloading operations.
-The first deliveries of $C_1$ and $C_2$ are timely and there is no gap between the two deliveries received by $C_1$.
-There is a delay between the end of the first delivery and the start of the second delivery of $C_2$, however, it is less than the maximal time delay of 20 minutes defined for this example.
-
-\begin{figure}[htbp]
-    % \caption{Solution of an instance with two plants, two construction sites, four orders, and three drivers.}
-    \vspace*{-10mm}
-
-    \centering
-    \begin{adjustbox}{max width=0.85\textwidth}
-
-        \begin{tikzpicture}
-            \SetUpEdge[lw = 1.2pt,
-                labelcolor = white,
-                % labelstyle = {draw}
-            ]
-            \tikzstyle{LabelStyle}=[fill=white]
-
-            \SetGraphUnit{5}
-            \GraphInit[vstyle=Normal]
-            \tikzset{VertexStyle/.style = {shape = circle,text = black,minimum size = 24 pt,draw = black}}
-            \Vertex[L={$B_1$}]{D1}
-            \tikzset{VertexStyle/.style = {shape = circle,text = black,minimum size = 10 pt,draw = blue}}
-            \tikzset{VertexStyle/.style = {shape = rectangle,text = black,minimum size = 10 pt,draw = blue}}
-            \Vertex[x=2,y=5,L={11 m$^3$}]{C21}
-            \EA[unit=2.0,L={3 m$^3$}](C21){C22}
-            \EA[unit=2.0,L={1 m$^3$}](C22){C23}
-            \tikzset{VertexStyle/.style = {shape = rectangle,text = black,minimum size = 10 pt,draw = blue}}
-            \SOEA[L={$12m^3$}](D1){C1}
-            \tikzset{VertexStyle/.style = {shape = circle,text = black,minimum size = 24 pt,draw = black}}
-            \SOEA[L={$B_2$}](C23){D2}
-            % %  \SetVertexArt
-            \tikzset{VertexStyle/.style = {shape = circle,text = black,minimum size = 10 pt}}
-            \SO[unit=0.8,L={$\tau^u_1=$30 $m^3/h$, 07:00 AM}](C1){tau1}
-            \SO[unit=0.5,L={\textbf{$C_1$} }](tau1){o1}
-
-            \NO[unit=0.9,L={\textbf{$C_2$}}](C22){C2}
-            \SOEA[unit=0.4,L={\textcolor{white}{  } $\tau^u_2=30$ $m^3/h$, 07:25 AM}](C2){tau2}
-
-            \tikzset{VertexStyle/.style = {shape = rectangle,text = black,minimum size = 10 pt}}
-            \SOWE[unit=1.,L={$\tau^l_1=$ 70 $m^3/h$}](D1){cap1}
-            \SO[unit=0.5,L={\textbf{\textcolor{brown}{D1: 8 m$^3$}}}](cap1){driver11}
-            \SO[unit=0.5,L={\textbf{\textcolor{orange}{D2: 8 m$^3$}}}](driver11){driver12}
-            \SO[unit=1,L={$\tau^l_2= 87$ $m^3/h$}](D2){cap2}
-            \SO[unit=0.5,L={\textbf{\textcolor{blue}{D3: 12 m$^3$}}}](cap2){driver2}
-            %%%%%%% Edge
-            \tikzset{EdgeStyle/.style={->}}
-            \Edge[](C21)(C22)
-            \Edge[](C22)(C23)
-            \tikzset{EdgeStyle/.style={<->,bend left=0}}
-            \Edge[label=$22$](D1)(C1)
-            \tikzset{EdgeStyle/.style={<->}}
-            \Edge[label=$19$](D2)(C23)
-            \Edge[label=$34$](D2)(C1)
-            \tikzset{EdgeStyle/.style={->}}
-            \tikzset{EdgeStyle/.append style = {bend left = 8}}
-            \Edge[label=$17$](D1)(D2)
-            \Edge[label=$18$](D2)(D1)
-            \tikzset{EdgeStyle/.append style = {bend left = 65}}
-            \Edge[label=$6$](D1)(C21)
-            \tikzset{EdgeStyle/.append style = {bend right = 50}}
-            \Edge[label=$5$](C21)(D1)
-            \tikzset{EdgeStyle/.style = {->,bend left = 20,draw=brown}}
-            \Edge[label=$D1$](D1)(C1)
-            \Edge[label=$D1$](C1)(D1)
-            \tikzset{EdgeStyle/.style = {->,bend left = 37,draw=orange}}
-            \Edge[label=$D2$](C1)(D1)
-            \Edge[label=$D2$](D1)(C1)
-
-            \tikzset{EdgeStyle/.style = {->,bend left= 20,draw=brown}}
-            \Edge[label=$D1$](D1)(C22)
-            \tikzset{EdgeStyle/.style = {->,bend left= 0,draw=brown}}
-            \Edge[label=$D1$](C22)(D1)
-
-            \tikzset{EdgeStyle/.style = {->,bend left = 0,draw=orange}}
-            \Edge[label=$D2$](C23)(D1)
-            \tikzset{EdgeStyle/.style = {->,bend left = -15,draw=orange}}
-            \Edge[label=$D2$](D1)(C23)
-
-            \tikzset{EdgeStyle/.style = {->,,draw=blue}}
-            \Edge[label=$D3$](D2)(D1)
-            \tikzset{EdgeStyle/.style = {->,bend left = 15,draw=blue}}
-            \Edge[label=$D3$](D1)(C21)
-            \tikzset{EdgeStyle/.style = {->,bend left = 80,draw=blue}}
-            \Edge[label=$D3$](C21)(D2)
-        \end{tikzpicture}
-    \end{adjustbox}
-    \vspace*{-10mm}
-    % }
-    \caption{Solution of an instance with two plants, two construction sites, four orders, and three drivers.}
-    \label{fig_Example}
-
-\end{figure}
-
-\begin{figure}[htbp]
-    \centering
-
-    \vspace*{-0mm}
-    \begin{adjustbox}{max width=0.85\textwidth}
-        \begin{tikzpicture}[-latex]
-            \matrix (chart)
-            [
+  dummy/.style    = {circle,draw},
             matrix of nodes, nodes in empty cells,
@@
 \section{Constructive heuristics and GRASP}
-\label{grasp_method}
+\label{sec:cdp_grasp_method}
 
@@
 \section{Computational experiments}
-\label{comp_exp}
-
-In this section, we present and discuss the results of computational experiments conducted using the GRASP heuristic described in section~\ref{grasp_method}.
+\label{sec:cdp_comp_exp}
+
+In this section, we present and discuss the results of computational experiments conducted using the GRASP heuristic described in section~\ref{sec:cdp_grasp_method}.
 The implementation of the GRASP algorithm is coded in C++.
@@
 The dataset is divided into small, medium, and large sets according to the total number of daily orders.
-We have summarized the dataset in Table~\ref{tab:instances}.
+We have summarized the dataset in Table~\ref{tab:cdp_instances}.
 The dataset is available upon request as a contribution to the CDP literature.
@@
     \caption{Instances summary}
-    \label{tab:instances}
+    \label{tab:cdp_instances}
     \small
@@
 Although their primary objective is to maximize the total load delivered each day, without considering loading operations at a plant and ensuring deliveries within specified time windows, we still use this dataset to analyze and compare the performance of our algorithm under different conditions.
-Table \ref*{tab:cdp_instances} shows the performance of the GRASP heuristic along with the other methods (CP, MIP, SD-heuristic, and Hybrid) used in \cite{kinable2014concrete}.
+Table \ref{tab:cdp_lit_comp} shows the performance of the GRASP heuristic along with the other methods (CP, MIP, SD-heuristic, and Hybrid) used in \cite{kinable2014concrete}.
 For each method, the table shows the average execution time in milliseconds and the corresponding gap between the obtained solution and the upper bound provided by \cite{kinable2014concrete}.
@@
     \caption{Performance of the GRASP heuristic on the CDP benchmark instances}
-    \label{tab:cdp_instances}
+    \label{tab:cdp_lit_comp}
     \scriptsize
@@
 %The stopping criteria are set to 20,000 iterations or 3600 seconds of running time (unless otherwise specified).
-The parameters used in our experiments are listed in Table~\ref*{tab:problem_parameters}.
+The parameters used in our experiments are listed in Table~\ref{tab:cdp_problem_parameters}.
 
@@
     \caption{Parameters of the real-world instances}
-    \label{tab:problem_parameters}
+    \label{tab:cdp_problem_parameters}
     \small
@@
 
-The table \ref*{tab:all_instances} contains the detailed results for all instances with a stop criterion of 3600 seconds.
+The table \ref{tab:cdp_all_instances} contains the detailed results for all instances with a stop criterion of 3600 seconds.
 The table shows the demand of each instance, the average, and the best values of the objective function when solving each instance five times.
@@
     \caption{Results with all instances}
-    \label{tab:all_instances}
+    \label{tab:cdp_all_instances}
     \resizebox{\textwidth}{!}{
@@
     \midrule
-    \multirow[c]{8}{*}{Small} & C_13_11_12_1 & 226.5 & 1,243.5 & 0.0 & 0.0 & 10.8 & 0.0 & 1,568.3 & 1,243.5 & 0.0 & 0.0 & 0.0 & 0.0 & 0.0 & 1,243.5 \\
-     & C_13_5_8_1 & 267.0 & 745.1 & 0.0 & 0.0 & 0.0 & 0.0 & 745.1 & 745.1 & 0.0 & 0.0 & 0.0 & 0.0 & 0.0 & 745.1 \\
-     & C_18_6_11_2 & 333.5 & 753.3 & 0.0 & 0.0 & 145.9 & 0.0 & 5,130.0 & 753.3 & 0.0 & 0.0 & 0.0 & 118.7 & 0.0 & 4,315.1 \\
-     & C_15_4_7_2 & 375.0 & 1,015.6 & 0.0 & 0.0 & 148.2 & 150.5 & 8,470.9 & 1,015.4 & 0.0 & 0.0 & 0.0 & 123.4 & 110.3 & 6,924.5 \\
-     & C_19_7_8_2 & 388.0 & 1,747.2 & 0.0 & 0.0 & 349.8 & 159.0 & 15,421.6 & 1,781.8 & 0.0 & 0.0 & 0.0 & 186.0 & 0.0 & 7,361.5 \\
-     & C_29_10_14_3 & 613.5 & 3,510.1 & 0.0 & 59.5 & 194.6 & 331.9 & 75,488.1 & 3,493.4 & 0.0 & 59.5 & 35.7 & 120.5 & 60.0 & 67,810.4 \\
-     & C_31_8_11_3 & 776.0 & 2,314.8 & 0.0 & 0.0 & 804.3 & 640.7 & 39,256.8 & 2,278.2 & 0.0 & 0.0 & 0.0 & 503.5 & 617.0 & 29,722.9 \\
-     & C_35_9_10_3 & 937.5 & 4,175.5 & 0.0 & 0.0 & 222.1 & 669.3 & 24,224.7 & 4,163.1 & 0.0 & 0.0 & 0.0 & 243.3 & 570.0 & 22,863.2 \\
+    \multirow[c]{8}{*}{Small} & C\_13\_11\_12\_1 & 226.5 & 1,243.5 & 0.0 & 0.0 & 10.8 & 0.0 & 1,568.3 & 1,243.5 & 0.0 & 0.0 & 0.0 & 0.0 & 0.0 & 1,243.5 \\
+     & C\_13\_5\_8\_1 & 267.0 & 745.1 & 0.0 & 0.0 & 0.0 & 0.0 & 745.1 & 745.1 & 0.0 & 0.0 & 0.0 & 0.0 & 0.0 & 745.1 \\
+     & C\_18\_6\_11\_2 & 333.5 & 753.3 & 0.0 & 0.0 & 145.9 & 0.0 & 5,130.0 & 753.3 & 0.0 & 0.0 & 0.0 & 118.7 & 0.0 & 4,315.1 \\
+     & C\_15\_4\_7\_2 & 375.0 & 1,015.6 & 0.0 & 0.0 & 148.2 & 150.5 & 8,470.9 & 1,015.4 & 0.0 & 0.0 & 0.0 & 123.4 & 110.3 & 6,924.5 \\
+     & C\_19\_7\_8\_2 & 388.0 & 1,747.2 & 0.0 & 0.0 & 349.8 & 159.0 & 15,421.6 & 1,781.8 & 0.0 & 0.0 & 0.0 & 186.0 & 0.0 & 7,361.5 \\
+     & C\_29\_10\_14\_3 & 613.5 & 3,510.1 & 0.0 & 59.5 & 194.6 & 331.9 & 75,488.1 & 3,493.4 & 0.0 & 59.5 & 35.7 & 120.5 & 60.0 & 67,810.4 \\
+     & C\_31\_8\_11\_3 & 776.0 & 2,314.8 & 0.0 & 0.0 & 804.3 & 640.7 & 39,256.8 & 2,278.2 & 0.0 & 0.0 & 0.0 & 503.5 & 617.0 & 29,722.9 \\
+     & C\_35\_9\_10\_3 & 937.5 & 4,175.5 & 0.0 & 0.0 & 222.1 & 669.3 & 24,224.7 & 4,163.1 & 0.0 & 0.0 & 0.0 & 243.3 & 570.0 & 22,863.2 \\
      \cmidrule(l){1-16}
 
-    \multirow[c]{14}{*}{Medium} & C_76_40_43_6 & 1,160.5 & 6,055.8 & 0.0 & 49.1 & 3,286.4 & 990.0 & 173,509.5 & 5,990.6 & 0.0 & 33.9 & 25.4 & 3,061.3 & 788.6 & 147,510.1 \\
-     & C_104_42_47_8 & 1,565.0 & 8,107.1 & 12.0 & 53.3 & 3,637.1 & 119.4 & 1,372,936.3 & 8,301.9 & 12.0 & 54.3 & 16.8 & 3,200.5 & 140.1 & 1,361,393.1 \\
-     & C_94_63_70_7 & 1,746.5 & 13,783.0 & 0.0 & 238.7 & 983.2 & 1,554.6 & 313,077.0 & 13,673.5 & 0.0 & 195.9 & 59.7 & 1,025.2 & 1,407.3 & 268,524.2 \\
-     & C_116_67_78_8 & 1,839.5 & 13,456.8 & 0.0 & 132.7 & 1,869.0 & 614.3 & 214,501.3 & 13,484.8 & 0.0 & 98.0 & 22.7 & 1,792.8 & 622.4 & 177,710.8 \\
-     & C_116_71_82_8 & 2,060.0 & 11,393.8 & 9.0 & 265.1 & 2,714.8 & 1,054.6 & 1,279,068.9 & 10,793.3 & 9.0 & 242.0 & 41.8 & 3,197.9 & 665.2 & 1,262,000.5 \\
-     & C_117_57_70_6 & 2,327.5 & 14,975.0 & 0.0 & 324.1 & 1,005.5 & 1,149.6 & 392,281.0 & 15,303.5 & 0.0 & 295.2 & 54.6 & 659.4 & 1,083.8 & 351,940.4 \\
-     & C_137_79_83_7 & 2,425.0 & 17,644.3 & 0.0 & 553.8 & 2,320.4 & 1,993.3 & 680,888.1 & 17,379.3 & 0.0 & 474.0 & 55.7 & 2,964.4 & 2,016.3 & 620,601.6 \\
-     & C_127_66_80_8 & 2,512.5 & 14,584.3 & 0.0 & 128.3 & 1,600.0 & 1,451.5 & 219,949.8 & 14,499.2 & 0.0 & 127.5 & 47.3 & 1,390.2 & 1,212.8 & 207,962.2 \\
-     & C_128_68_74_7 & 2,595.0 & 18,450.6 & 3.7 & 566.6 & 1,317.5 & 2,706.2 & 1,048,712.1 & 18,261.4 & 0.0 & 473.4 & 56.0 & 1,764.7 & 2,525.6 & 595,143.5 \\
-     & C_136_85_97_8 & 2,673.0 & 16,831.6 & 0.0 & 379.9 & 2,702.4 & 2,368.0 & 525,199.5 & 17,241.3 & 0.0 & 323.7 & 50.2 & 2,190.5 & 2,767.1 & 461,973.3 \\
-     & C_128_78_85_8 & 2,685.5 & 15,679.8 & 63.0 & 474.5 & 1,399.6 & 2,394.4 & 6,880,077.6 & 15,830.6 & 63.0 & 432.7 & 57.8 & 1,206.8 & 2,272.7 & 6,830,154.3 \\
-     & C_131_77_85_8 & 2,893.5 & 16,988.0 & 0.0 & 374.5 & 1,557.9 & 1,996.1 & 478,106.8 & 16,565.6 & 0.0 & 352.1 & 44.6 & 1,624.3 & 1,322.5 & 443,865.0 \\
-     & C_137_89_97_7 & 2,939.5 & 21,716.4 & 9.8 & 924.4 & 821.0 & 3,494.4 & 2,020,622.6 & 21,311.3 & 0.0 & 799.5 & 51.8 & 887.3 & 3,486.7 & 917,162.4 \\
-     & C_133_84_98_7 & 2,971.0 & 20,435.5 & 46.1 & 1,141.3 & 894.1 & 2,127.8 & 5,841,090.1 & 20,520.4 & 9.0 & 1,411.0 & 59.5 & 705.0 & 1,871.9 & 2,390,157.6 \\
+    \multirow[c]{14}{*}{Medium} & C\_76\_40\_43\_6 & 1,160.5 & 6,055.8 & 0.0 & 49.1 & 3,286.4 & 990.0 & 173,509.5 & 5,990.6 & 0.0 & 33.9 & 25.4 & 3,061.3 & 788.6 & 147,510.1 \\
+     & C\_104\_42\_47\_8 & 1,565.0 & 8,107.1 & 12.0 & 53.3 & 3,637.1 & 119.4 & 1,372,936.3 & 8,301.9 & 12.0 & 54.3 & 16.8 & 3,200.5 & 140.1 & 1,361,393.1 \\
+     & C\_94\_63\_70\_7 & 1,746.5 & 13,783.0 & 0.0 & 238.7 & 983.2 & 1,554.6 & 313,077.0 & 13,673.5 & 0.0 & 195.9 & 59.7 & 1,025.2 & 1,407.3 & 268,524.2 \\
+     & C\_116\_67\_78\_8 & 1,839.5 & 13,456.8 & 0.0 & 132.7 & 1,869.0 & 614.3 & 214,501.3 & 13,484.8 & 0.0 & 98.0 & 22.7 & 1,792.8 & 622.4 & 177,710.8 \\
+     & C\_116\_71\_82\_8 & 2,060.0 & 11,393.8 & 9.0 & 265.1 & 2,714.8 & 1,054.6 & 1,279,068.9 & 10,793.3 & 9.0 & 242.0 & 41.8 & 3,197.9 & 665.2 & 1,262,000.5 \\
+     & C\_117\_57\_70\_6 & 2,327.5 & 14,975.0 & 0.0 & 324.1 & 1,005.5 & 1,149.6 & 392,281.0 & 15,303.5 & 0.0 & 295.2 & 54.6 & 659.4 & 1,083.8 & 351,940.4 \\
+     & C\_137\_79\_83\_7 & 2,425.0 & 17,644.3 & 0.0 & 553.8 & 2,320.4 & 1,993.3 & 680,888.1 & 17,379.3 & 0.0 & 474.0 & 55.7 & 2,964.4 & 2,016.3 & 620,601.6 \\
+     & C\_127\_66\_80\_8 & 2,512.5 & 14,584.3 & 0.0 & 128.3 & 1,600.0 & 1,451.5 & 219,949.8 & 14,499.2 & 0.0 & 127.5 & 47.3 & 1,390.2 & 1,212.8 & 207,962.2 \\
+     & C\_128\_68\_74\_7 & 2,595.0 & 18,450.6 & 3.7 & 566.6 & 1,317.5 & 2,706.2 & 1,048,712.1 & 18,261.4 & 0.0 & 473.4 & 56.0 & 1,764.7 & 2,525.6 & 595,143.5 \\
+     & C\_136\_85\_97\_8 & 2,673.0 & 16,831.6 & 0.0 & 379.9 & 2,702.4 & 2,368.0 & 525,199.5 & 17,241.3 & 0.0 & 323.7 & 50.2 & 2,190.5 & 2,767.1 & 461,973.3 \\
+     & C\_128\_78\_85\_8 & 2,685.5 & 15,679.8 & 63.0 & 474.5 & 1,399.6 & 2,394.4 & 6,880,077.6 & 15,830.6 & 63.0 & 432.7 & 57.8 & 1,206.8 & 2,272.7 & 6,830,154.3 \\
+     & C\_131\_77\_85\_8 & 2,893.5 & 16,988.0 & 0.0 & 374.5 & 1,557.9 & 1,996.1 & 478,106.8 & 16,565.6 & 0.0 & 352.1 & 44.6 & 1,624.3 & 1,322.5 & 443,865.0 \\
+     & C\_137\_89\_97\_7 & 2,939.5 & 21,716.4 & 9.8 & 924.4 & 821.0 & 3,494.4 & 2,020,622.6 & 21,311.3 & 0.0 & 799.5 & 51.8 & 887.3 & 3,486.7 & 917,162.4 \\
+     & C\_133\_84\_98\_7 & 2,971.0 & 20,435.5 & 46.1 & 1,141.3 & 894.1 & 2,127.8 & 5,841,090.1 & 20,520.4 & 9.0 & 1,411.0 & 59.5 & 705.0 & 1,871.9 & 2,390,157.6 \\
      \cmidrule(l){1-16}
-    \multirow[c]{14}{*}{Large} & C_132_98_109_8 & 3,078.5 & 22,994.6 & 0.0 & 465.5 & 496.5 & 3,804.8 & 579,471.0 & 22,712.7 & 0.0 & 351.5 & 56.2 & 803.2 & 4,011.3 & 478,549.4 \\
-     & C_129_101_119_8 & 3,229.0 & 17,323.1 & 11.0 & 657.8 & 1,577.6 & 2,375.9 & 1,869,937.5 & 17,311.8 & 11.0 & 598.8 & 58.4 & 1,507.7 & 1,948.1 & 1,800,315.9 \\
-     & C_141_114_136_8 & 3,350.5 & 19,891.6 & 39.1 & 1,255.8 & 1,366.2 & 2,251.9 & 5,271,698.5 & 19,556.2 & 18.0 & 1,366.7 & 57.9 & 1,380.0 & 2,695.1 & 3,281,568.6 \\
-     & C_140_114_132_8 & 3,401.5 & 21,917.6 & 23.8 & 1,254.1 & 680.1 & 3,389.9 & 3,744,172.5 & 22,590.2 & 17.5 & 1,248.5 & 58.7 & 766.3 & 3,267.6 & 3,109,441.2 \\
-     & C_143_101_123_8 & 3,437.5 & 23,477.2 & 0.0 & 677.7 & 955.5 & 4,369.4 & 817,244.5 & 23,806.9 & 0.0 & 617.3 & 56.0 & 1,302.2 & 5,147.9 & 783,123.1 \\
-     & C_137_112_129_8 & 3,471.0 & 22,089.2 & 0.0 & 546.6 & 544.8 & 4,187.3 & 668,737.0 & 21,195.2 & 0.0 & 456.6 & 55.0 & 883.9 & 3,532.1 & 574,935.0 \\
-     & C_142_114_129_8 & 3,499.5 & 20,678.2 & 35.7 & 1,007.5 & 1,119.7 & 4,133.1 & 4,714,436.9 & 21,243.7 & 21.0 & 907.0 & 59.8 & 1,038.0 & 4,016.6 & 3,139,750.0 \\
-     & C_149_98_122_8 & 3,513.0 & 19,859.0 & 28.4 & 1,141.7 & 2,154.6 & 2,649.9 & 4,119,224.3 & 19,602.8 & 21.0 & 859.2 & 55.7 & 2,342.9 & 2,728.2 & 3,103,676.6 \\
-     & C_139_98_108_8 & 3,541.0 & 21,060.5 & 0.0 & 966.2 & 684.9 & 3,585.4 & 1,079,477.1 & 21,084.8 & 0.0 & 867.2 & 59.8 & 834.9 & 4,133.9 & 996,046.7 \\
-     & C_144_108_122_8 & 3,670.5 & 23,330.4 & 2.0 & 1,123.4 & 536.2 & 3,251.6 & 1,427,866.8 & 22,882.8 & 0.0 & 969.7 & 57.8 & 649.7 & 3,587.1 & 1,083,850.9 \\
-     & C_142_92_107_8 & 3,684.5 & 23,179.0 & 0.0 & 578.3 & 384.1 & 4,260.4 & 698,184.4 & 23,610.6 & 0.0 & 449.2 & 59.8 & 315.7 & 4,257.2 & 567,403.7 \\
-     & C_138_114_136_8 & 3,739.5 & 23,614.6 & 11.7 & 1,225.1 & 837.4 & 5,668.6 & 2,557,166.5 & 22,859.0 & 7.5 & 1,107.0 & 58.3 & 1,224.2 & 5,078.2 & 2,018,119.1 \\
-     & C_150_127_136_8 & 3,946.5 & 24,940.3 & 0.0 & 949.5 & 775.5 & 5,384.3 & 1,105,391.3 & 24,939.2 & 0.0 & 829.0 & 57.1 & 805.9 & 6,166.7 & 1,001,490.6 \\
-     & C_148_112_131_8 & 3,953.5 & 24,286.0 & 34.0 & 1,247.8 & 657.0 & 3,009.5 & 4,752,026.9 & 24,221.7 & 24.5 & 1,118.5 & 57.1 & 859.7 & 3,483.3 & 3,688,169.1 \\
+    \multirow[c]{14}{*}{Large} & C\_132\_98\_109\_8 & 3,078.5 & 22,994.6 & 0.0 & 465.5 & 496.5 & 3,804.8 & 579,471.0 & 22,712.7 & 0.0 & 351.5 & 56.2 & 803.2 & 4,011.3 & 478,549.4 \\
+     & C\_129\_101\_119\_8 & 3,229.0 & 17,323.1 & 11.0 & 657.8 & 1,577.6 & 2,375.9 & 1,869,937.5 & 17,311.8 & 11.0 & 598.8 & 58.4 & 1,507.7 & 1,948.1 & 1,800,315.9 \\
+     & C\_141\_114\_136\_8 & 3,350.5 & 19,891.6 & 39.1 & 1,255.8 & 1,366.2 & 2,251.9 & 5,271,698.5 & 19,556.2 & 18.0 & 1,366.7 & 57.9 & 1,380.0 & 2,695.1 & 3,281,568.6 \\
+     & C\_140\_114\_132\_8 & 3,401.5 & 21,917.6 & 23.8 & 1,254.1 & 680.1 & 3,389.9 & 3,744,172.5 & 22,590.2 & 17.5 & 1,248.5 & 58.7 & 766.3 & 3,267.6 & 3,109,441.2 \\
+     & C\_143\_101\_123\_8 & 3,437.5 & 23,477.2 & 0.0 & 677.7 & 955.5 & 4,369.4 & 817,244.5 & 23,806.9 & 0.0 & 617.3 & 56.0 & 1,302.2 & 5,147.9 & 783,123.1 \\
+     & C\_137\_112\_129\_8 & 3,471.0 & 22,089.2 & 0.0 & 546.6 & 544.8 & 4,187.3 & 668,737.0 & 21,195.2 & 0.0 & 456.6 & 55.0 & 883.9 & 3,532.1 & 574,935.0 \\
+     & C\_142\_114\_129\_8 & 3,499.5 & 20,678.2 & 35.7 & 1,007.5 & 1,119.7 & 4,133.1 & 4,714,436.9 & 21,243.7 & 21.0 & 907.0 & 59.8 & 1,038.0 & 4,016.6 & 3,139,750.0 \\
+     & C\_149\_98\_122\_8 & 3,513.0 & 19,859.0 & 28.4 & 1,141.7 & 2,154.6 & 2,649.9 & 4,119,224.3 & 19,602.8 & 21.0 & 859.2 & 55.7 & 2,342.9 & 2,728.2 & 3,103,676.6 \\
+     & C\_139\_98\_108\_8 & 3,541.0 & 21,060.5 & 0.0 & 966.2 & 684.9 & 3,585.4 & 1,079,477.1 & 21,084.8 & 0.0 & 867.2 & 59.8 & 834.9 & 4,133.9 & 996,046.7 \\
+     & C\_144\_108\_122\_8 & 3,670.5 & 23,330.4 & 2.0 & 1,123.4 & 536.2 & 3,251.6 & 1,427,866.8 & 22,882.8 & 0.0 & 969.7 & 57.8 & 649.7 & 3,587.1 & 1,083,850.9 \\
+     & C\_142\_92\_107\_8 & 3,684.5 & 23,179.0 & 0.0 & 578.3 & 384.1 & 4,260.4 & 698,184.4 & 23,610.6 & 0.0 & 449.2 & 59.8 & 315.7 & 4,257.2 & 567,403.7 \\
+     & C\_138\_114\_136\_8 & 3,739.5 & 23,614.6 & 11.7 & 1,225.1 & 837.4 & 5,668.6 & 2,557,166.5 & 22,859.0 & 7.5 & 1,107.0 & 58.3 & 1,224.2 & 5,078.2 & 2,018,119.1 \\
+     & C\_150\_127\_136\_8 & 3,946.5 & 24,940.3 & 0.0 & 949.5 & 775.5 & 5,384.3 & 1,105,391.3 & 24,939.2 & 0.0 & 829.0 & 57.1 & 805.9 & 6,166.7 & 1,001,490.6 \\
+     & C\_148\_112\_131\_8 & 3,953.5 & 24,286.0 & 34.0 & 1,247.8 & 657.0 & 3,009.5 & 4,752,026.9 & 24,221.7 & 24.5 & 1,118.5 & 57.1 & 859.7 & 3,483.3 & 3,688,169.1 \\
     \bottomrule
@@
 
-Table \ref{tab_runtime} shows that as the algorithm runtime increases, a higher percentage of deliveries are completed.
+Table \ref{tab_cdp_runtime} shows that as the algorithm runtime increases, a higher percentage of deliveries are completed.
 For small instances, all requests are delivered in less than a minute, while for medium and large instances, 99\% of requests are delivered.
@@
     \caption{RMC delivery completion within different runtimes of the GRASP}
-    \label{tab_runtime}
+    \label{tab_cdp_runtime}
     \scriptsize
@@
 
-Figure \ref*{fig:plants_schedules} provides a visual representation using Gantt charts of the loading and unloading operations for the orders assigned to each plant in instance $C29\_10\_14\_3$.
+Figure \ref{fig:plants_schedules} provides a visual representation using Gantt charts of the loading and unloading operations for the orders assigned to each plant in instance $C29\_10\_14\_3$.
 It also shows the different assignments of each driver throughout the day.
@@
 Customer $8$ has two orders ($o_8$,$o_9$) assigned to plant $0$ while customer $9$ has four orders ($o_{10}$ - - $o_{13}$) assigned to plant $2$.
-All the order requirements are in the table~\ref*{tab:instance_detail}.
-
-\begin{table}[!h]
+All the order requirements are in the table~\ref{tab:instance_detail}.
+
+\begin{table}[!ht]
     \centering
@@
 Plant $0$ is responsible for both orders from customer $8$.
-In the figure~\ref*{fig:plant0}, order $o_8$ is served before $o_9$, starting at the due time.
+In the figure~\ref{fig:plant0}, order $o_8$ is served before $o_9$, starting at the due time.
 Each row in the figure represents the pairs of loading and unloading tasks performed by a driver.
 The number next to a task refers to the order.
-Figure~\ref*{fig:plant2} shows how the loading bay is alternately assigned to orders $o_{13}$ and $o_7$ between $10:00$ and $14:00$.
+Figure~\ref{fig:plant2} shows how the loading bay is alternately assigned to orders $o_{13}$ and $o_7$ between 10:00 and 14:00.
 Driver $26$ starts his service at plant $0$, serves order $o_9$ four times, and travels to plant $2$ to make the final delivery of order $o_{13}$.
@@
 We report a $FDD$ of $59.5$ minutes for this solution of instance $C29\_10\_14\_3$.
-The Gantt chart shows that the customers $4$ and $5$ who need their deliveries at $07:00$ all have a late first delivery.
+The Gantt chart shows that the customers $4$ and $5$, who need their deliveries at 07:00, have a late first delivery.
 These delays are due to the work schedules of the drivers assigned to Plant 1.
-These drivers start their shifts at the earliest at $06:05$ and $06:20$, and it takes them $59$ to $63$ minutes to reach the customers.
+These drivers start their shifts at the earliest at 06:05 and $06:20$, and it takes them $59$ to $63$ minutes to reach the customers.
 This finding suggests that the relatively high FDD values in our results may be related to suboptimal scheduling of drivers' hours in our instances.
@@
 Upon investigation, we find that even if we have $104$ drivers in this case, the number of drivers who start their shift before t$07:00$ is not enough to serve this customer within the 60-minute window we defined.
-Thus, the value of $UQ$ that we report in the table~\ref*{tab:all_instances} appears to be optimal given the driver shift schedule given as input to our algorithm.
+Thus, the value of $UQ$ that we report in the table~\ref{tab:cdp_all_instances} appears to be optimal given the driver shift schedule given as input to our algorithm.
 We confirmed this observation by removing the driver start shift and overtime constraints and resolving our instances five more times.
 All orders are now fully serviced on each run.
-The table~\ref*{tab:lift_shift_constraints} compares the averages of the results from  the tables \ref*{tab:all_instances} and \ref*{tab:all_instances_no_shift}.
+The table~\ref{tab:lift_shift_constraints} compares the averages of the results from  the tables \ref{tab:cdp_all_instances} and \ref{tab:cdp_all_instances_no_shift}.
 We found that for small instances the maximum delay dropped to zero, and for medium (large) instances it was reduced by almost 60\% (40\%) on average.
@@
 
-\begin{table}[!h]
+\begin{table}[!ht]
     \centering
@@
     \caption{Results with all instances without shift and overtime constraints}
-    \label{tab:all_instances_no_shift}
+    \label{tab:cdp_all_instances_no_shift}
     \resizebox{\textwidth}{!}{
@@
     \midrule
-    \multirow[c]{8}{*}{Small} & C_13_11_12_1 & 226.5 & 1,243.5 & 0.0 & 0.0 & 1,243.5 & 1,243.5 & 0.0 & 0.0 & 0.0 & 1,243.5 \\
-     & C_13_5_8_1 & 267.0 & 745.1 & 0.0 & 0.0 & 745.1 & 745.1 & 0.0 & 0.0 & 0.0 & 745.1 \\
-     & C_18_6_11_2 & 333.5 & 753.3 & 0.0 & 0.0 & 753.3 & 753.3 & 0.0 & 0.0 & 0.0 & 753.3 \\
-     & C_15_4_7_2 & 375.0 & 1,013.7 & 0.0 & 0.0 & 1,013.7 & 1,013.1 & 0.0 & 0.0 & 0.0 & 1,013.1 \\
-     & C_19_7_8_2 & 388.0 & 1,661.0 & 0.0 & 0.0 & 1,661.0 & 1,658.2 & 0.0 & 0.0 & 0.0 & 1,658.2 \\
-     & C_29_10_14_3 & 613.5 & 3,235.5 & 0.0 & 0.0 & 3,235.5 & 3,198.6 & 0.0 & 0.0 & 0.0 & 3,198.6 \\
-     & C_31_8_11_3 & 776.0 & 2,257.6 & 0.0 & 0.0 & 2,257.6 & 2,219.9 & 0.0 & 0.0 & 0.0 & 2,219.9 \\
-     & C_35_9_10_3 & 937.5 & 3,834.0 & 0.0 & 0.0 & 3,834.0 & 3,808.8 & 0.0 & 0.0 & 0.0 & 3,808.8 \\
+    \multirow[c]{8}{*}{Small} & C\_13\_11\_12\_1 & 226.5 & 1,243.5 & 0.0 & 0.0 & 1,243.5 & 1,243.5 & 0.0 & 0.0 & 0.0 & 1,243.5 \\
+     & C\_13\_5\_8\_1 & 267.0 & 745.1 & 0.0 & 0.0 & 745.1 & 745.1 & 0.0 & 0.0 & 0.0 & 745.1 \\
+     & C\_18\_6\_11\_2 & 333.5 & 753.3 & 0.0 & 0.0 & 753.3 & 753.3 & 0.0 & 0.0 & 0.0 & 753.3 \\
+     & C\_15\_4\_7\_2 & 375.0 & 1,013.7 & 0.0 & 0.0 & 1,013.7 & 1,013.1 & 0.0 & 0.0 & 0.0 & 1,013.1 \\
+     & C\_19\_7\_8\_2 & 388.0 & 1,661.0 & 0.0 & 0.0 & 1,661.0 & 1,658.2 & 0.0 & 0.0 & 0.0 & 1,658.2 \\
+     & C\_29\_10\_14\_3 & 613.5 & 3,235.5 & 0.0 & 0.0 & 3,235.5 & 3,198.6 & 0.0 & 0.0 & 0.0 & 3,198.6 \\
+     & C\_31\_8\_11\_3 & 776.0 & 2,257.6 & 0.0 & 0.0 & 2,257.6 & 2,219.9 & 0.0 & 0.0 & 0.0 & 2,219.9 \\
+     & C\_35\_9\_10\_3 & 937.5 & 3,834.0 & 0.0 & 0.0 & 3,834.0 & 3,808.8 & 0.0 & 0.0 & 0.0 & 3,808.8 \\
      \cmidrule(l){1-12}
-    \multirow[c]{14}{*}{Medium} & C_76_40_43_6 & 1,160.5 & 5,262.4 & 0.0 & 0.0 & 5,262.4 & 5,187.2 & 0.0 & 0.0 & 0.0 & 5,187.2 \\
-     & C_104_42_47_8 & 1,565.0 & 8,278.0 & 0.0 & 0.0 & 8,278.0 & 8,209.6 & 0.0 & 0.0 & 0.0 & 8,209.6 \\
-     & C_94_63_70_7 & 1,746.5 & 12,730.3 & 0.0 & 0.0 & 12,730.3 & 12,551.9 & 0.0 & 0.0 & 0.0 & 12,551.9 \\
-     & C_116_67_78_8 & 1,839.5 & 11,884.4 & 0.0 & 0.0 & 11,884.4 & 11,641.4 & 0.0 & 0.0 & 0.0 & 11,641.4 \\
-     & C_116_71_82_8 & 2,060.0 & 10,750.9 & 0.0 & 0.0 & 10,750.9 & 10,697.1 & 0.0 & 0.0 & 0.0 & 10,697.1 \\
-     & C_117_57_70_6 & 2,327.5 & 14,547.0 & 0.0 & 25.8 & 40,380.4 & 14,746.8 & 0.0 & 16.1 & 9.1 & 30,816.0 \\
-     & C_137_79_83_7 & 2,425.0 & 17,822.4 & 0.0 & 120.0 & 137,840.9 & 17,982.5 & 0.0 & 92.0 & 17.6 & 110,023.4 \\
-     & C_127_66_80_8 & 2,512.5 & 13,622.2 & 0.0 & 0.0 & 13,622.2 & 13,565.0 & 0.0 & 0.0 & 0.0 & 13,565.0 \\
-     & C_128_68_74_7 & 2,595.0 & 18,641.7 & 0.0 & 74.1 & 92,744.7 & 18,554.8 & 0.0 & 43.4 & 19.0 & 61,987.5 \\
-     & C_136_85_97_8 & 2,673.0 & 17,336.3 & 0.0 & 51.8 & 69,145.9 & 17,610.5 & 0.0 & 23.1 & 16.3 & 40,731.6 \\
-     & C_128_78_85_8 & 2,685.5 & 16,781.2 & 0.0 & 55.5 & 72,294.4 & 16,658.5 & 0.0 & 39.4 & 12.5 & 56,076.4 \\
-     & C_131_77_85_8 & 2,893.5 & 17,789.1 & 0.0 & 22.0 & 39,787.5 & 16,923.5 & 0.0 & 0.0 & 0.0 & 16,923.5 \\
-     & C_137_89_97_7 & 2,939.5 & 21,831.7 & 0.0 & 214.2 & 236,037.5 & 22,630.7 & 0.0 & 157.1 & 48.5 & 179,691.7 \\
-     & C_133_84_98_7 & 2,971.0 & 21,031.1 & 0.0 & 474.0 & 495,073.6 & 21,179.3 & 0.0 & 359.8 & 59.8 & 380,980.3 \\
+    \multirow[c]{14}{*}{Medium} & C\_76\_40\_43\_6 & 1,160.5 & 5,262.4 & 0.0 & 0.0 & 5,262.4 & 5,187.2 & 0.0 & 0.0 & 0.0 & 5,187.2 \\
+     & C\_104\_42\_47\_8 & 1,565.0 & 8,278.0 & 0.0 & 0.0 & 8,278.0 & 8,209.6 & 0.0 & 0.0 & 0.0 & 8,209.6 \\
+     & C\_94\_63\_70\_7 & 1,746.5 & 12,730.3 & 0.0 & 0.0 & 12,730.3 & 12,551.9 & 0.0 & 0.0 & 0.0 & 12,551.9 \\
+     & C\_116\_67\_78\_8 & 1,839.5 & 11,884.4 & 0.0 & 0.0 & 11,884.4 & 11,641.4 & 0.0 & 0.0 & 0.0 & 11,641.4 \\
+     & C\_116\_71\_82\_8 & 2,060.0 & 10,750.9 & 0.0 & 0.0 & 10,750.9 & 10,697.1 & 0.0 & 0.0 & 0.0 & 10,697.1 \\
+     & C\_117\_57\_70\_6 & 2,327.5 & 14,547.0 & 0.0 & 25.8 & 40,380.4 & 14,746.8 & 0.0 & 16.1 & 9.1 & 30,816.0 \\
+     & C\_137\_79\_83\_7 & 2,425.0 & 17,822.4 & 0.0 & 120.0 & 137,840.9 & 17,982.5 & 0.0 & 92.0 & 17.6 & 110,023.4 \\
+     & C\_127\_66\_80\_8 & 2,512.5 & 13,622.2 & 0.0 & 0.0 & 13,622.2 & 13,565.0 & 0.0 & 0.0 & 0.0 & 13,565.0 \\
+     & C\_128\_68\_74\_7 & 2,595.0 & 18,641.7 & 0.0 & 74.1 & 92,744.7 & 18,554.8 & 0.0 & 43.4 & 19.0 & 61,987.5 \\
+     & C\_136\_85\_97\_8 & 2,673.0 & 17,336.3 & 0.0 & 51.8 & 69,145.9 & 17,610.5 & 0.0 & 23.1 & 16.3 & 40,731.6 \\
+     & C\_128\_78\_85\_8 & 2,685.5 & 16,781.2 & 0.0 & 55.5 & 72,294.4 & 16,658.5 & 0.0 & 39.4 & 12.5 & 56,076.4 \\
+     & C\_131\_77\_85\_8 & 2,893.5 & 17,789.1 & 0.0 & 22.0 & 39,787.5 & 16,923.5 & 0.0 & 0.0 & 0.0 & 16,923.5 \\
+     & C\_137\_89\_97\_7 & 2,939.5 & 21,831.7 & 0.0 & 214.2 & 236,037.5 & 22,630.7 & 0.0 & 157.1 & 48.5 & 179,691.7 \\
+     & C\_133\_84\_98\_7 & 2,971.0 & 21,031.1 & 0.0 & 474.0 & 495,073.6 & 21,179.3 & 0.0 & 359.8 & 59.8 & 380,980.3 \\
      \cmidrule(l){1-12}
-     \multirow[c]{14}{*}{Large} & C_132_98_109_8 & 3,078.5 & 22,272.3 & 0.0 & 7.2 & 29,438.3 & 23,294.5 & 0.0 & 4.0 & 1.8 & 27,252.8 \\
-     & C_129_101_119_8 & 3,229.0 & 18,473.8 & 0.0 & 41.6 & 60,111.3 & 18,595.2 & 0.0 & 28.0 & 15.7 & 46,551.2 \\
-     & C_141_114_136_8 & 3,350.5 & 21,237.4 & 0.0 & 136.7 & 157,923.2 & 20,762.9 & 0.0 & 110.7 & 18.6 & 131,504.9 \\
-     & C_140_114_132_8 & 3,401.5 & 22,518.4 & 0.0 & 174.0 & 196,489.2 & 22,537.0 & 0.0 & 64.8 & 10.6 & 87,390.4 \\
-     & C_143_101_123_8 & 3,437.5 & 23,807.5 & 0.0 & 39.1 & 62,917.8 & 23,249.3 & 0.0 & 30.2 & 13.8 & 53,413.3 \\
-     & C_137_112_129_8 & 3,471.0 & 22,457.5 & 0.0 & 121.0 & 143,439.3 & 21,878.2 & 0.0 & 27.9 & 7.5 & 49,772.5 \\
-     & C_142_114_129_8 & 3,499.5 & 21,574.4 & 0.0 & 362.8 & 384,413.2 & 22,591.1 & 0.0 & 221.0 & 52.9 & 243,564.1 \\
-     & C_149_98_122_8 & 3,513.0 & 20,918.7 & 0.0 & 150.6 & 171,499.5 & 19,503.4 & 0.0 & 94.1 & 28.6 & 113,643.5 \\
-     & C_139_98_108_8 & 3,541.0 & 21,548.1 & 0.0 & 271.5 & 293,062.8 & 22,008.3 & 0.0 & 239.9 & 47.4 & 261,921.3 \\
-     & C_144_108_122_8 & 3,670.5 & 23,604.0 & 0.0 & 169.2 & 192,821.5 & 23,200.9 & 0.0 & 125.9 & 42.5 & 149,122.9 \\
-     & C_142_92_107_8 & 3,684.5 & 21,279.5 & 0.0 & 83.8 & 105,084.6 & 21,244.2 & 0.0 & 36.4 & 14.4 & 57,641.8 \\
-     & C_138_114_136_8 & 3,739.5 & 24,385.3 & 0.0 & 109.8 & 134,180.9 & 25,218.0 & 0.0 & 71.7 & 15.8 & 96,877.5 \\
-     & C_150_127_136_8 & 3,946.5 & 24,434.2 & 0.0 & 391.0 & 415,476.7 & 24,732.2 & 0.0 & 292.5 & 52.8 & 317,191.2 \\
-     & C_148_112_131_8 & 3,953.5 & 25,846.2 & 0.0 & 566.8 & 592,677.4 & 26,245.8 & 0.0 & 383.8 & 52.7 & 410,087.8 \\
+     \multirow[c]{14}{*}{Large} & C\_132\_98\_109\_8 & 3,078.5 & 22,272.3 & 0.0 & 7.2 & 29,438.3 & 23,294.5 & 0.0 & 4.0 & 1.8 & 27,252.8 \\
+     & C\_129\_101\_119\_8 & 3,229.0 & 18,473.8 & 0.0 & 41.6 & 60,111.3 & 18,595.2 & 0.0 & 28.0 & 15.7 & 46,551.2 \\
+     & C\_141\_114\_136\_8 & 3,350.5 & 21,237.4 & 0.0 & 136.7 & 157,923.2 & 20,762.9 & 0.0 & 110.7 & 18.6 & 131,504.9 \\
+     & C\_140\_114\_132\_8 & 3,401.5 & 22,518.4 & 0.0 & 174.0 & 196,489.2 & 22,537.0 & 0.0 & 64.8 & 10.6 & 87,390.4 \\
+     & C\_143\_101\_123\_8 & 3,437.5 & 23,807.5 & 0.0 & 39.1 & 62,917.8 & 23,249.3 & 0.0 & 30.2 & 13.8 & 53,413.3 \\
+     & C\_137\_112\_129\_8 & 3,471.0 & 22,457.5 & 0.0 & 121.0 & 143,439.3 & 21,878.2 & 0.0 & 27.9 & 7.5 & 49,772.5 \\
+     & C\_142\_114\_129\_8 & 3,499.5 & 21,574.4 & 0.0 & 362.8 & 384,413.2 & 22,591.1 & 0.0 & 221.0 & 52.9 & 243,564.1 \\
+     & C\_149\_98\_122\_8 & 3,513.0 & 20,918.7 & 0.0 & 150.6 & 171,499.5 & 19,503.4 & 0.0 & 94.1 & 28.6 & 113,643.5 \\
+     & C\_139\_98\_108\_8 & 3,541.0 & 21,548.1 & 0.0 & 271.5 & 293,062.8 & 22,008.3 & 0.0 & 239.9 & 47.4 & 261,921.3 \\
+     & C\_144\_108\_122\_8 & 3,670.5 & 23,604.0 & 0.0 & 169.2 & 192,821.5 & 23,200.9 & 0.0 & 125.9 & 42.5 & 149,122.9 \\
+     & C\_142\_92\_107\_8 & 3,684.5 & 21,279.5 & 0.0 & 83.8 & 105,084.6 & 21,244.2 & 0.0 & 36.4 & 14.4 & 57,641.8 \\
+     & C\_138\_114\_136\_8 & 3,739.5 & 24,385.3 & 0.0 & 109.8 & 134,180.9 & 25,218.0 & 0.0 & 71.7 & 15.8 & 96,877.5 \\
+     & C\_150\_127\_136\_8 & 3,946.5 & 24,434.2 & 0.0 & 391.0 & 415,476.7 & 24,732.2 & 0.0 & 292.5 & 52.8 & 317,191.2 \\
+     & C\_148\_112\_131\_8 & 3,953.5 & 25,846.2 & 0.0 & 566.8 & 592,677.4 & 26,245.8 & 0.0 & 383.8 & 52.7 & 410,087.8 \\
     \bottomrule
@@
 We ran additional tests with different values of $\gamma^1$, varying between 5 and 25 minutes, with no shifts or overtime constraints on the drivers.
-The influence of the value of the maximum time delay between two consecutive deliveries of the same order is shown in figure~\ref*{fig:gamma1_influence}.
+The influence of the value of the maximum time delay between two consecutive deliveries of the same order is shown in figure~\ref{fig:gamma1_influence}.
 For small instances, all customers are served on time regardless of the value of $\gamma^1$.
@@
 
-\begin{figure}[!h]
+\begin{figure}[!ht]
     \centering
@@
 We ran five tests on our dataset and obtained two solutions per instance: one that prioritizes the shortest travel time and another that seeks the smallest first delivery delay.
-Figure~\ref*{fig:trade_off} visually represents the trade-off between travel time and first delivery delay for these solutions.
+Figure~\ref{fig:trade_off} visually represents the trade-off between travel time and first delivery delay for these solutions.
 It is clear that as the first delivery delay decreases, the travel time increases, and conversely when the travel time is minimized, the first delivery delay tends to be longer.
@@
 \section{Conclusion}
-\label{concl}
+\label{sec:cdp_concl}
 
@@
 \bibliography{References}
-
-\end{document}

### 2023-12-19 codeCQ/Combinatorial.h
save

@@ -7,6 +7,7 @@
 #include "Parameters.h"
 #include <unordered_map>
 #include <tuple>
+#include <algorithm>
 
 class Combinatorial {
 public:

### 2023-12-19 codeCQ/Data.cpp

@@ -14,8 +14,8 @@ Data::Data()
           maxDriverCap(0), minDriverCap(0), _deliveries(0), _drivers(0), _orders_deliveries(0), _distances(0),
           _times(0), _docks(0),
           _index_orders(0),
-//          path(std::filesystem::current_path().parent_path().parent_path()),
-          path(std::filesystem::current_path().parent_path()),
+          path(std::filesystem::current_path().parent_path().parent_path()),
+//          path(std::filesystem::current_path().parent_path()),
           instance_name(),
           problem_name(), sol_output(), result_file() {
     time_mat_file = path + "/matrices/Matrice_de_temps_all.csv";

### 2023-12-19 codeCQ/Solution.h

@@ -12,6 +12,7 @@
 #include "Node.h"
 #include <utility>
 #include <vector>
+#include <algorithm>
 #include <set>
 #include <cassert>
 #include <iostream>

### 2023-12-19 codeCQ/heur/CDPSolver.cpp

@@ -79,7 +79,7 @@ void CDPSolver::SolveCDP(Sol &s, Data &dat, int iter, bool restart) {
     vector<AllInsertionOperator> AllInsertionOp;
     AllInsertionOp.reserve(3 * (custInfo.size() + priorityInfo.size() + driverInfo.size()));
     for (const auto &val: custInfo) {
-       AllInsertionOp.emplace_back(&custIns1, val.first, "Builder 1 " + val.second);
+        AllInsertionOp.emplace_back(&custIns1, val.first, "Builder 1 " + val.second);
         AllInsertionOp.emplace_back(&custIns3, val.first, "Builder 3 " + val.second);
     }
     for (const auto &val: priorityInfo) {
@@ -210,7 +210,7 @@ void CDPSolver::RecursiveSols(Sol &s, Delivery *del, std::unordered_map<std::str
 void CDPSolver::LearnParameters(Sol &s, std::set<int> const &customer) {
     for (auto i: customer) {
         Customer *c = s.GetCustomer(i);
-        unordered_map<string, Sol, MyHashFunction> myMap;
+        unordered_map < string, Sol, MyHashFunction > myMap;
         Sol cur1(s.GetData());
         find_all_routes(cur1, c, &myMap);
         int count = 0;

### 2023-12-19 codeCQ/heur/CustInsertion.cpp

@@ -69,9 +69,7 @@ void CustInsertion::InsertWithBactrack(Sol &s, std::vector<Customer *> &list) {
 
 bool CustInsertion::Insert(Sol &s, Customer *c, Order *cur_order) {
     bool nextCustomer = false;
-    int depth = 0;
-//    if (c->custID == 57)
-//        Parameters::SHOW = true;
+    int depth = 0; // backtracking depth
     for (int j = 0; j < s.GetDeliveryCount(cur_order);) {
         Delivery *del = s.GetDelivery(cur_order, j);
         if (del == nullptr) break;

### 2023-12-19 codeCQ/heur/CustInsertionBacktrack.cpp

@@ -50,7 +50,7 @@ void CustInsertionBacktrack::Insert(Sol &s) {
 }
 
 void CustInsertionBacktrack::Insert(Sol &s, std::vector<Customer *> &list,
-                           std::vector<int> &list_ID) {
+                                    std::vector<int> &list_ID) {
     if (list.size() > 1) {
         Sort(s, list, list_ID, CustInsertionBacktrack::_k);
 //        Prompt::print(list,",");
@@ -64,7 +64,7 @@ void CustInsertionBacktrack::Insert(Sol &s, std::vector<Customer *> &list,
     std::shuffle(driversList.begin(), driversList.end(), Parameters::RANDOM_GEN);
     Customer *c = list[0];
 //    c->early_tw=144;
-     for (int i = 0; i < list.size();) {
+    for (int i = 0; i < list.size();) {
 //        cout << *c << "--" << std::endl;
         bool custom_client = false;
         if (s.isClientSatisfied(c)) {
@@ -75,13 +75,13 @@ void CustInsertionBacktrack::Insert(Sol &s, std::vector<Customer *> &list,
         }
         Order *cur_order = s.GetRandomOrder(c);
 
-         int iter_count = 0;
+        int iter_count = 0;
 //        if (cur_order->custID == 18)
 //         Parameters::SHOW = true;
-         for (int j = 0; j < s.GetDeliveryCount(cur_order);) {
-             iter_count++;
+        for (int j = 0; j < s.GetDeliveryCount(cur_order);) {
+            iter_count++;
 
-             if (iter_count >= 100 and j == 0) {
+            if (iter_count >= 100 and j == 0) {
 //                 cout << iter_count << endl;
 //                 s.ShowSchedule(c);
 //                 cout << *c << endl;
@@ -89,7 +89,7 @@ void CustInsertionBacktrack::Insert(Sol &s, std::vector<Customer *> &list,
 
 //                exit(1);
 
-             }
+            }
             Delivery *del = s.GetDelivery(cur_order, j);
             if (Parameters::SHOW) {
                 Prompt::print(
@@ -105,7 +105,7 @@ void CustInsertionBacktrack::Insert(Sol &s, std::vector<Customer *> &list,
                         Prompt::print({"Push", to_string(del->id), "by", to_string(Sol::minDelay[del->id])});
                     }
                     ListMoveVrp temp_moves;
-                    if(del->rank==0){
+                    if (del->rank == 0) {
[…]

### 2023-12-19 codeCQ/heur/InsRmvMethodFast.cpp

@@ -373,8 +373,6 @@ Move<Delivery, Driver, MoveVrp> InsRmvMethodFast::GetCost(Sol &s, Delivery *n, D
         TimeSlot intv = TimeSlot(loadSlot.lower, node_arr);
         auto it = s.driverWorkingIntervals[d->id].find(intv);
         if (it != s.driverWorkingIntervals[d->id].end()) {
-            Depot *temp_dep = s.GetDepot(it->n.depotID);
-
             Customer *it_cust = s.GetCustomer(dynamic_cast<Delivery *>(s.GetNode(it->n.id))->custID);
             if (Parameters::PENALTY_COST) {
                 newcost.lateDeliveryCost +=
@@ -512,9 +510,7 @@ void InsRmvMethodFast::repairSolution(Sol &s, Order *o) {
     s.Update(o);
 }
 
-void InsRmvMethodFast::repairSolution(Sol &s, Customer *c) {
-
-}
+void InsRmvMethodFast::repairSolution(Sol &s, Customer *c) {}
 
 void InsRmvMethodFast::FillStructures(Sol &s, std::vector<Customer *> &customersList,
                                       std::vector<Driver *> &driversList) {

### 2023-12-19 codeCQ/heur/RechercheLocale.cpp

@@ -49,10 +49,10 @@ void RechercheLocale::Run(Sol &s) {
 				entree = true;
 				continue;
 			}
-			if (RelocateDriver(s)) {
-				entree = true;
-				continue;
-			}
+//			if (RelocateDriver(s)) {
+//				entree = true;
+//				continue;
+//			}
 		}
 	}
 	RunAllFeasible(s);

### 2023-12-19 codeCQ/heur/RechercheLocale.h

@@ -5,6 +5,7 @@
 #include "CDPSolver.h"
 #include <vector>
 #include <iostream>
+#include <iterator>
 #include<set>
 
 class RechercheLocale {

### 2023-12-19 codeCQ/heur/Solver.cpp

@@ -147,6 +147,7 @@ void Solver::SolveGrasp(Sol &s, Data &dat, vector<set<int>> const &linkedClientS
     loc_search.LinkedClientSlot = linkedClientSlot;
     if (Parameters::LOCAL_SEARCH) {
         grasp.Optimize(s, nullptr, nullptr, &loc_search, false);
+//        grasp.Optimize(s, nullptr, nullptr, nullptr, false);
     } else {
         grasp.Optimize(s, nullptr, nullptr, nullptr, false);
     }

### 2023-12-19 solution/cq/gantt.py

@@ -2,28 +2,37 @@ import pandas as pd
 import numpy as np
 import plotly.express as px
 from distinctipy import distinctipy
+from datetime import datetime
 
 # Transform minutes to format"hh:mm:ss:"
-def time_to_date(date,t):
-    date=str(date)
+
+
+def time_to_date(date, t):
+    date = str(date)
     # print(date)
-    date=date[:4]+"-"+date[4:6]+"-"+date[6:]
-    min=int(t%60)
-    hour=int(t//60)
-    if min<10:
-        min=f"0{min}"
-    new_date=f"{date} {hour}:{min}:00"
-    if hour>23: 
-        print(new_date,' ',t)
-    return new_date
+    date = date[:4]+"-"+date[4:6]+"-"+date[6:]
+    min = int(t % 60)
+    hour = int(t//60)
+    if min < 10:
+        min = f"0{min}"
+    new_date = f"{date} {hour}:{min}:00"
+    if hour > 23:
+        print(new_date, ' ', t)
+
+    # datetime_string = "2023-12-18 15:30:00"
+    datetime_format = "%Y-%m-%d %H:%M:%S"
+    datetime_object = datetime.strptime(new_date, datetime_format)
+    # print(datetime_object)
+    return datetime_object
 
 
 def show(df):
-    df['task'] = df.apply(lambda x: f" del {x.Del}",axis=1)
-    color = [f"rgb{distinctipy.get_rgb256(x)}" for x in distinctipy.get_colors(29)]
-    fig = px.timeline(df, x_start="start", x_end="end", y="Driver",title="All drivers",
-                      color="Driver",template='ggplot2',
-                      color_continuous_scale= color)
+    df['task'] = df.apply(lambda x: f" del {x.Del}", axis=1)
+    color = [
+        f"rgb{distinctipy.get_rgb256(x)}" for x in distinctipy.get_colors(29)]
+    fig = px.timeline(df, x_start="start", x_end="end", y="Driver", title="All drivers",
+                      color="Driver", template='ggplot2',
+                      color_continuous_scale=color)
     # fig.update_yaxes(autorange="reversed") # otherwise tasks are listed from the bottom up
     # fig.update_xaxes(fixedrange=True)
     # fig.update_xaxes(
@@ -31,127 +40,280 @@ def show(df):
     #     dict(dtickrange=[None, 86400000], value=" %h %m %s \n")
     # ]
     # )
-    
+
     # fig.update_xaxes(dtick=86400000)
     fig.update_layout(
-             xaxis=dict(
[…]

### 2024-03-09 ArticleCDP/articleCDP.tex
save

--- 2023-11-29 Article/cdp_content.tex
@@
+\documentclass{article}
+% \documentclass[review]{elsarticle}
+\usepackage{epsfig}
+\usepackage{amsmath}
+\usepackage{mathtools}
+\usepackage{algorithm}
+\usepackage[algo2e,noend]{algorithm2e}
+\usepackage[compatible,noend]{algpseudocode}
+\usepackage{enumitem}
+\usepackage{xspace}
+\usepackage[table,xcdraw]{xcolor}
+\usepackage{amsfonts}
+\usepackage{pifont}
+\usepackage{bbding}
+\usepackage[english]{babel}
+\usepackage{adjustbox}
+\usepackage{tikz}
+\usepackage{subcaption}
+\usetikzlibrary{automata, topaths, calc, positioning, shapes, backgrounds, fit, matrix}
+\usepackage{caption}
+\usepackage{tabularx}
+\usepackage[T1]{fontenc}
+\usepackage{graphicx}
+\usepackage{natbib}
+\usepackage{datatool}
+\usepackage{array}
+\usepackage{pgfplotstable}
+\pgfplotsset{width=7cm,compat=newest}
+\usepackage{booktabs}
+\usepackage{longtable}
+\usepackage{pdflscape}
+\usepackage{afterpage}
+\usepackage{capt-of}
+\usepackage{multirow}
+\usepackage{float}
+\usepackage[text={15cm,21cm},centering]{geometry}
+\usepackage{setspace}
+\usepackage{array}
+\allowdisplaybreaks
+\usepackage{lineno}
+\usepackage{breqn}
+\usepackage{tkz-graph}
+
+\newcolumntype{H}{>{\setbox0=\hbox\bgroup}c<{\egroup}@{}}
+
+\newtheorem{theorem}{Theorem}[section]
+\newtheorem{lemma}[theorem]{Lemma}
+\newtheorem{proposition}[theorem]{Proposition}
+\newtheorem{corollary}[theorem]{Corollary}
+\newtheorem{definition}[theorem]{Definition}
+\newcommand{\pcmaxc}{P$|\mbox{\em cont}|$C$_{\max}$\xspace}
+
+\newcommand{\GRASP}{Greedy randomized adaptive search procedure\xspace}
+\newcommand{\F}{$\mathcal{F}$\xspace}
+\newcommand{\B}{$\mathcal{B}$\xspace}
+\newcommand{\C}{$\mathcal{C}$\xspace}
+\newcommand{\LB}{$\mathcal{L}$\xspace}
+\newcommand{\Lb}{$\mathcal{L}_b$\xspace}
+\newcommand{\D}{$\mathcal{D}$\xspace}
+ \newcommand*{\red}{\textcolor{red}}
+
+\renewcommand{\algorithmicforall}{\textbf{for each}}
+\SetKwFor{ForEach}{for each}{}{}%
+\renewcommand{\algorithmiccomment}[2][1\linewidth]{%
+\leavevmode\hfill\makebox[#1][l]{/* \textbf{~#2} */}}
+\renewcommand{\algorithmicrequire}{\textbf{Input:}}
+
+\renewcommand{\tabcolsep}{4pt}
+
+\modulolinenumbers[1]
+\linenumbers
+
+\oddsidemargin=0.5in
+\topmargin=-0.25in
+\textwidth=5.5in
+\textheight=8.25in
+
+%\journal{European Journal of Operational Research}
+
+%\begin{frontmatter}
+
+\title{A GRASP algorithm for the concrete delivery problem. }
+\author{Ousmane Ali$^{(1)}$, Jean-Fran\c cois C\^ot\'e$^{(2)}$, Leandro C.~Coelho$^{(3)}$\\
+ $(1)$ {\tt nassoma-wattara-ousmane.ali.1@ulaval.ca}\\
+ $(2)$ {\tt Jean-Francois.Cote@fsa.ulaval.ca}\\
+ $(3)$ {\tt Leandro.Coelho@fsa.ulaval.ca}\\
+}
+
+% \date{\today}
+
+\begin{document}
+\maketitle
 \begin{abstract}
@@
 
-Financial support for this work was provided by the Canadian Natural Sciences and Engineering Research Council (NSERC) under grants 2015-04893 and 2019-00094. This support is gratefully acknowledged.
+Financial support for this work was provided by the Canadian Natural Sciences and Engineering Research Council (NSERC) under grants 2019-00094 and 2021-04037.
+This support is gratefully acknowledged.
+We thank the Digital Research Alliance of Canada for providing high-performance computing facilities.
 
@@
 \bibliography{References}
+
+\end{document}

### 2024-03-09 ArticleCDP/original/articleCDP.tex
Save

\documentclass{article}
% \documentclass[review]{elsarticle}
\usepackage{epsfig}
\usepackage{amsmath}
\usepackage{mathtools}
\usepackage{algorithm}
\usepackage[algo2e,noend]{algorithm2e}
\usepackage[compatible,noend]{algpseudocode}
\usepackage{enumitem}
\usepackage{xspace}
\usepackage[table,xcdraw]{xcolor}
\usepackage{amsfonts}
\usepackage{pifont}
\usepackage{bbding}
\usepackage[english]{babel}
\usepackage{adjustbox}
\usepackage{tikz}
\usepackage{subcaption}
\usetikzlibrary{automata, topaths, calc, positioning, shapes, backgrounds, fit, matrix}
\usepackage{caption}
\usepackage{tabularx}
\usepackage[T1]{fontenc}
\usepackage{graphicx}
\usepackage{natbib}
\usepackage{datatool}
\usepackage{array}
\usepackage{pgfplotstable}
\pgfplotsset{width=7cm,compat=newest}
\usepackage{booktabs}
\usepackage{longtable}
\usepackage{pdflscape}
\usepackage{afterpage}
\usepackage{capt-of}
\usepackage{multirow}
\usepackage{float}
\usepackage[text={15cm,21cm},centering]{geometry}
\usepackage{setspace}
\usepackage{array}
\usepackage{lineno}
\usepackage{breqn}
\usepackage{tkz-graph}
\usepackage{varwidth}
\usepackage{url}
\allowdisplaybreaks
\newcolumntype{H}{>{\setbox0=\hbox\bgroup}c<{\egroup}@{}}

\newtheorem{theorem}{Theorem}[section]
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{proposition}[theorem]{Proposition}
\newtheorem{corollary}[theorem]{Corollary}
\newtheorem{definition}[theorem]{Definition}
\newcommand{\pcmaxc}{P$|\mbox{\em cont}|$C$_{\max}$\xspace}

\newcommand{\GRASP}{Greedy randomized adaptive search procedure\xspace}
\newcommand{\F}{$\mathcal{F}$\xspace}
\newcommand{\B}{$\mathcal{B}$\xspace}
\newcommand{\C}{$\mathcal{C}$\xspace}
\newcommand{\LB}{$\mathcal{L}$\xspace}
\newcommand{\Lb}{$\mathcal{L}_b$\xspace}
\newcommand{\D}{$\mathcal{D}$\xspace}
 \newcommand*{\red}{\textcolor{red}}

\renewcommand{\algorithmicforall}{\textbf{for each}}
\SetKwFor{ForEach}{for each}{}{}%
\renewcommand{\algorithmiccomment}[2][1\linewidth]{%
\leavevmode\hfill\makebox[#1][l]{/* \textbf{~#2} */}}
\renewcommand{\algorithmicrequire}{\textbf{Input:}}

\renewcommand{\tabcolsep}{4pt}

\modulolinenumbers[1]
% \linenumbers

\oddsidemargin=0.5in
\topmargin=-0.25in
\textwidth=5.5in
\textheight=8.25in

%\journal{European Journal of Operational Research}

%\begin{frontmatter}

\title{A GRASP algorithm for the concrete delivery problem}
\author{Ousmane Ali$^{(1)}$, Jean-Fran\c cois C\^ot\'e$^{(2)}$, Leandro C.~Coelho$^{(3)}$\\
 $(1)$ {\tt nassoma-wattara-ousmane.ali.1@ulaval.ca}\\
 $(2)$ {\tt Jean-Francois.Cote@fsa.ulaval.ca}\\
 $(3)$ {\tt Leandro.Coelho@fsa.ulaval.ca}\\
}

\date{\today}

\begin{document}
\maketitle
\begin{abstract}
  This paper addresses a novel variant of the Concrete Delivery Problem (CDP), which involves the efficient scheduling of ready-mixed concrete deliveries to construction sites while balancing the conflicting goals of minimizing transportation costs and maximizing customer satisfaction. In this study, we propose an exact formulation and a heuristic approach based on the Greedy Randomized Adaptive Search Procedure (GRASP) to tackle this challenging CDP variant. This variant introduces realistic side constraints, including driver working shifts, a minimum driver working time, and overtime penalties. Additionally, it considers the case where customers may request multiple types of concrete delivered within the same time window. We assess the performance of our heuristic using new instances generated for this problem and provide a comparative analysis with another CDP variant from the literature to demonstrate its effectiveness.
\end{abstract}

% \begin{keywords}
\noindent{\textbf{Keywords:} Vehicle scheduling;  concrete delivery; GRASP; ready-mixed concrete}
% \end{keywords}

\section{Introduction}
\label{sec:cdp_intro}
Concrete is a widely used building material in construction projects. Its perishable nature is affected by many factors that impact its quality \citep{sinha_quality_2021}, which is crucial for the durability and strength of the final construction. Concrete comes in two types: ready-mixed concrete (RMC) and site-mixed concrete (SMC). RMC is manufactured in a batch plant and delivered to the construction site, while SMC is produced on-site using raw materials stored on the construction site. Using SMC can avoid delays caused by road traffic, but it has a slower and more difficult production process, requires storage for mixing materials and equipment, and is suitable for low amounts of concrete. On the other hand, RMC has better quality and benefits from lower production costs \citep{muresan_comparing}. However, to take advantage of these benefits, the batch plant manager must ensure efficient and prompt delivery to the construction site, which may require a fleet of high-cost revolving drum trucks (concrete mixers) to dispatch the RMC.

Concrete delivery under the form of RMC is subject to many operational constraints that make the Concrete Delivery Problem (CDP) very challenging. In this paper, we study a variant of the CDP to schedule the daily production and dispatching of RMC for a company located in the province of Quebec, Canada. This company operates multiple batch plants with varying production rates, using a fleet of concrete mixers of different capacities. Each plant has its own fleet of trucks; however, under certain conditions, the trucks can move between plants if necessary. The trucks must return to their home plant at the end of the day. The company owns two types of trucks with different capacities and can call on an external fleet when needed. They serve construction sites from any of their plants, with the first delivery starting at the time specified by the customer. The loading and unloading of a concrete mixer depend on the truck capacity, the loading rate at the plant, and the unloading rate at the construction site. Drivers are allocated based on their daily work schedules. A customer may request several types of concrete to be delivered within the same time window, with no required sequence for the orders, but an order can only start after the completion of the previous order. This constraint generalizes the linked order constraints of \cite{durbin2008or} where some customers place two orders for the same day and request that they are linked (the second order begins only after the first order is completed). The setting of our study is similar to a variant of the CDP previously studied in \cite{schmid2009hybrid, schmid2010hybridization}. However, we include the plant's production rate and driver shift schedule. The company uses a centralized dispatcher system to schedule all daily orders, but this system has issues satisfying all daily demands without using an external fleet or using different plants to serve the same order.

According to \cite{blazewicz2019handbook}, the CDP combines vehicle routing with scheduling issues to plan routes to deliver concrete from batch plants (depots) to customers' construction sites. RMC is an on-demand product with a short life cycle from production through end use. It cannot be stored and cannot stay too long on a truck, or it will harden. Hence, concrete mixers must deliver RMC at the planned construction site shortly after its production. A customer quantity requirement is often greater than the truck size and must be fulfilled by multiple deliveries. In that sense, CDP is similar to the vehicle routing problem with split delivery \citep{archetti2008split}, except that the same truck may visit a customer more than once. Concrete hardens quickly, so multiple deliveries must be done back-to-back or at least close in time to avoid the problem of cold joint, which can reduce the strength and durability of the concrete. Customers request to be served within a specific time window, which can complicate truck-loading schedules when a plant can only load one truck at a time. Similarly, only one truck can unload at a time at a customer location, sometimes leading to concrete mixers queuing and waiting their turn to deliver. Furthermore, with trucks of varied sizes, loading, travel, and unloading times that may be uncertain, the CDP is a complex and challenging problem.

In this paper, we propose a mathematical model and a Greedy Randomized Search Procedure (GRASP) heuristic to solve a new variant of the CDP. Our model takes into consideration the working shifts of the drivers and the scheduling of multiple orders within the same time window at a construction site. To the best of our knowledge, this paper is the first that deals with these specific constraints in the context of the CDP.

The rest of the paper is organized as follows: Section \ref{sec:cdp_lit_review} provides a literature review of previous works related to the CDP. Section~\ref{sec:cdp_desc_form} provides a formal description and mathematical model of the problem. Section~\ref{sec:cdp_grasp_method} describes the GRASP algorithm and the constructive heuristics developed to solve this variant of the CDP. Section~\ref{sec:cdp_comp_exp} presents computational experiments and sensitivity analyses to evaluate the proposed approach, and finally, we conclude in  Section~\ref{sec:cdp_concl}.

\section{Literature review}
\label{sec:cdp_lit_review}

Academic research on concrete batch and delivery began in the late 1990s. \cite{tommelein1999just} described RMC as a prototypical example of a JIT production system in construction and identified two practices for delivering it. One approach is for the customer to haul the product from the batch plant with their concrete mixer, while the other is for the batch plant to deliver the concrete directly to the customer's location. This latter approach is the one that has been studied in all related papers found in the literature. Several works to schedule and dispatch concrete production and delivery have mainly focused on simulation-related methods. These methods can be standalone, such as those used by \cite{zayed2001simulation, wang2001scheduling} and  \cite{panas_simulation_based_2013}, among others. Alternatively, these methods can be hybridized with optimization techniques, such as those used by \cite{feng2004optimizing} and \cite{lu2005optimized}. \cite{wang2001scheduling} developed a simulation model to reveal the effect and value of the concrete mixers' inter-arrival time on the productivity of hired unloading equipment on site. \cite{feng2004optimizing} used a combination of genetic algorithm (GA) and simulation process to minimize the total waiting time for trucks at a customer site. The study focused on loading homogeneous trucks at the same batch plant, with fixed loading and unloading durations. The GA was used to find the best loading sequence of RMC trucks to be assigned to different construction sites. The simulation process determined the loading, arrival, departure, and waiting time of trucks and thus evaluated the cost of each dispatching sequence. They evaluated their method using data from a batch plant in Taiwan with up to nine customers served. \cite{lu2005optimized} used the same combination of GA and simulation to determine the optimal number of concrete mixers to be deployed and an optimal schedule for batching and delivering concrete. Their objective was to minimize the idle time of the site crew due to late concrete deliveries and truck queuing time. In this setting, it was also necessary to deliver a batch of mortar on-site to lubricate the unloading pump before the concrete delivery. As such, the simulation model also included the batch and delivery of mortar. Finding the best RMC fleet size was also the purpose of the discrete-event simulation model proposed by \cite{panas_simulation_based_2013}.

In addition to simulation-based methods, several other approaches have been used in the literature to solve the CDP. These include metaheuristics \citep{misir2011selection, maghrebi2016sequential, yang2022concrete}, exact methods \citep{asbach2009analysis, kinable2014concrete}, matheuristics \citep{schmid2009hybrid, schmid2010hybridization}, Benders decomposition \citep{maghrebi2014benders}, column generation (CG) \citep{maghrebi2014solving}, Lagrangian relaxation \citep{narayanan2015using}, and machine learning approaches \citep{graham2006modeling}. \cite{matsatsinis2004towards} designed a decision support system (DSS) for the dynamic routing of both concrete and pumps that may be necessary for some construction sites to aid in the unloading of concrete. The DSS considered the availability of three plants but stipulated that vehicles fulfilling the same order must all load at the same plant. Orders that could not be executed immediately could be postponed for the next day. The routing of the pumps was modeled as a multi-depot vehicle routing problem with time windows. \cite{naso2007genetic} proposed a sequential GA method combined with constructive heuristics to solve another variant of the CDP. In this problem, the plant's production schedule must consider orders that must be delivered to a customer's site as well as orders that must be picked up by the customer. The algorithm schedules the plant loading operations first before scheduling the truck deliveries. The authors also developed a non-linear model that minimizes transportation costs, waiting times, outsourced costs, and overtime work. They ran experiments using real-world instances of a concrete supply chain in the Netherlands and found a reduction in the number of outsourced requests. \cite{yan2007optimal} also considered overtime considerations in their paper, which focused on scheduling RMC for one batch plant with two loading docks. The study took into account that overtime wages are paid for factory and construction site operations after 4 PM. They developed a mixed-integer programming (MIP) model on a time-space network to minimize travel times and operating costs at both normal and overtime working hours at the plant and the construction sites. They tested the model using real data consisting of three days of operation using a two-stage algorithm. First, they solved the MIP relaxation with CPLEX. Then, they simplified the original model by fixing some decision variables before solving it. The algorithm was found to improve the actual plant operation by 10\%. A time-space network is the key component of the real-time DSS developed by \cite{durbin2008or} to solve a dynamic CDP every five minutes. The DSS can receive new orders, schedule them on the fly, and handle unexpected events such as plant closures, truck breakdowns, and delays in transportation times. The authors combined the DSS with a tabu search (TS) heuristic to warm start CPLEX, which made the model performant enough to solve instances with up to 1,500 loads per day with up to 250 trucks. The DSS also considers the case of a customer who places two orders, with the first being completed before the second starts. Further insights on the real-time planning and monitoring of CDP are available in \cite{garza2021dynamic}. Another variant of the CDP is modeled by \cite{schmid2009hybrid} as an integer multicommodity network flow (MCNF) problem on a time-space network. In this paper, concrete is delivered using a heterogeneous fleet of vehicles, and each plant can load an unlimited number of trucks simultaneously. Some of the trucks have specialized equipment and must arrive first at certain construction sites to assist in unloading the concrete. The objective is to fulfill all orders, minimize the travel cost, and avoid delays between two consecutive unloading operations for an order. The model is typically solved using a matheuristic algorithm that combines the MCNF with a variable neighborhood search (VNS) heuristic. The method can quickly solve large problem instances with more than 60 orders per day. The same problem is addressed by \cite{schmid2010hybridization}, who proposed a MIP model combined with a VNS and a very large neighborhood search (VLNS) to develop two matheuristics approaches. Comparisons between both matheuristics and a standalone VNS show that the former methods are much better and suitable for solving larger problem instances. These methods also provide better solutions for small to medium instances than the matheuristic used in \cite{schmid2009hybrid}. A pure VNS approach with the same problem but without the use of instrumentation has been applied by \cite{payr2009optimizing}.

Regarding objectives, most authors have focused on minimizing travel time and delays between consecutive deliveries. However, some authors have been more interested in maximizing customer satisfaction alone. We find these situations in the works of \cite{durbin2008or} and \cite{kinable2014concrete}. \cite{kinable2014concrete} introduce a general MIP and constraint programming (CP) models of the CDP reflecting the main constraints commonly found in all CDP works: time lag and no overlapping between consecutive deliveries, covering of all customers' demands, delivery time window, and heterogeneous fleet. However, the model did not include constraints limiting the time that concrete may stay on a truck. The authors propose a constructive heuristic that schedules the visits to the customers one by one according to the start time of the visit and the truck capacity. The procedure is invoked multiple times for different permutations of the customer's order which is determined using the steepest descent local search procedure. One of the paper's main contributions is the creation of the first public test instances for the CDP with up to 50 customers, four batch plants, and 20 concrete mixers. They found the CP model to be highly effective in finding high-quality solutions in a relatively short time or improving existing schedules, while the MIP model can be used to compute bounds, as it seems ineffective in solving large problem instances. Finally, the heuristic often yields good solutions in less than a second. A generalization of the MIP model of \cite{kinable2014concrete} is addressed in \cite{asbach2009analysis}. This model simultaneously minimizes the total sum of travel costs and the penalty costs for customers with unfulfilled demand. A customer can request that all concrete deliveries come from the same plant or a subset of plants and that a delivery truck belongs to a subset of the vehicle fleet. The MIP model is used in a local search scheme as a black-box solver to reoptimize an incumbent solution in which a neighborhood operator has unfixed some variables. \cite{tzanetos2023systematic} provide an overview of the various methods used in the literature to address the CDP and categorizes the problem formulations based on the different concepts used in the literature. They also discussed the consistency between industry needs and existing constraints and provided insights into the datasets corresponding to real-world cases, identifying the necessary data for practitioners.

\section{Problem description}
\label{sec:cdp_desc_form}
The focus of this paper is on the distribution of RMC from a Canadian company that operates in the greater Montreal area. When a customer places an order, it is received at a control center and immediately assigned to one of the company's batch plants. These plants produce the concrete and then deliver it to the customer. The problem involves a set of batch plants, a set of concrete-mixer drivers, and a set of customer orders.

Each of the company's plants has a single loading dock that can accommodate only one truck at a time. As a result, trucks often form a queue while waiting for their turn at the loading dock. Let $B$ be the set of batch plants. The plants are heterogeneous, as each plant $b$ has its hourly loading rate, represented by $\tau^l_b$, which affects the duration of the loading process. After loading the concrete, the driver spends $\alpha_b$ minutes adjusting the concrete in the truck before heading to the customer site. Each plant has its own assigned fleet of trucks, but it can borrow trucks from other plants. Let $l_{j}$ be the loading dock node associated with delivery node $j$. After loading RMC at $l_j$, it must be fully delivered to $j$ at most before $\Delta$ time, which is the concrete lifetime. We denote $L_b$ as the set of loading dock nodes of plant $b$ and $L$ as the set of all loading docks.

The company has two types of concrete mixer trucks with 8 and 12 cubic meters capacities. Each driver $k$ is assigned to a particular batch plant and is responsible for driving a truck with capacity $Q_k$. The set of drivers is represented by $K =\bigcup_{}K_b $, where $K_b$ is the set of drivers scheduled to start their shift at batch plant $b$. A driver $k$ is required to start his shift at $h_k$, work a minimum of $\mu$ hours and a maximum of $\eta$ hours during regular working hours, with the possibility of overtime of up to $\varLambda$ hours.

A driver typically loads RMC at his assigned batch plant but may be required to drive to and load at other plants if needed. The batch plant produces concrete on demand using recipes specific to each order. This means that a truck can only haul RMC for one order, even if there is spare capacity. After unloading the RMC, the driver takes $\rho$ minutes to clean the concrete mixer before proceeding.

A customer $c$ requests one or more types of concrete to be delivered to their construction site on a specific day, with the delivery service starting at the due time $a_c$. The customer's unloading rate $\tau^u_c$ and the quantity to be unloaded give the time required to unload a truckload. We call an order $o$ a request for a specific type of concrete. Let $q_c$ be the sum of the demands $q_o$ of each order $o$ placed and $a_c$ be the desired arrival time of the first concrete mixer. If an order requires more concrete than a single truck can carry, multiple deliveries are scheduled. To avoid cold joint problems with the concrete, subsequent deliveries of the same order must be made in close succession. We define a maximum time delay $\gamma^1$, after which no more deliveries are allowed for the same order. Let $C$ be the set of customers, $O_c$ be the set of all orders requested by customer $c$, and $O=\{O_c, c \in C\}$ be the set of all requested orders for all customers.

Each element of $O_c$ must be entirely delivered before moving on to another order. The first order $o \in O_c$ to complete must have its first delivery start between the time windows $\left[a_c, a_c + \tau^w_c \right]$, while the subsequent orders can start their deliveries at most $\gamma^2$ time units after their preceding orders are completed. Here, $\tau^w_c$ is a user-defined parameter representing the time window duration, and $\gamma^2$ is the maximum time delay between the completion of an order and the start of the next order. A plant is assigned to an order, and it must supply all of its subsequent deliveries.

Let $n_o$ be the number of deliveries needed to fulfill the order $o$. $n_o$ is not known in advance because the trucks have different capacities. However, we can compute lower $n_o^{min}$ and upper $n_o^{max}$ bounds using the capacities of the largest $Q_{max}$ and smallest $Q_{min}$ available trucks.
\begin{alignat}{3}
    \label{mod:c0}
    n_o^{min} = \left\lceil \frac{q_o}{Q_{max}} \right\rceil \leq n_o \leq n_o^{max} = \left\lceil \frac{q_o}{Q_{min}} \right\rceil & \text{ } &
    \forall  o \in O.
\end{alignat}

Let $d^j_{o}$ be the $j^{th}$ visit with load $q^j_{o}$ for order $o$. We represent the fulfillment of order $o$ by the visits to the ordered set of delivery nodes $D_o= \left(d^1_{o},d^2_{o},\cdots, d^{n_o}_{o}\right)$. The deliveries of customer $c$ are the ordered set $D_c= (D_{o_1}, D_{o_2},\cdots,D_{o_{|O_c|}})$, where $o_r$ is the $r^{th}$ delivered order. We will refer to $d \in D_c$ ($d \in D_o$) as the $d^{th}$ potential delivery of customer $c$ (order $o$). $D=\bigcup_{c\in C} D_c$ is the union of all delivery nodes.

A solution to the problem involves decisions about truck loading schedules, driver assignments to different deliveries, and truck arrival times at construction sites for unloading. For a batch plant, the decision involves choosing which driver to load, when to load them, how much to load, and which construction site to deliver. For a driver, the decision is to determine the sequence of loading depots and delivery sites. And for a construction site, the decision involves determining the arrival times of all scheduled deliveries for the day.

Each driver leaves and returns to their home plant every day. We represent the home plant of a driver $k$ with a starting depot $s_k$ and an ending depot $e_k$. Let $S$ and $E$ be the sets of starting and ending depots, respectively.

We define our problem on a directed graph where $V=\{ S \cup L \cup D \cup E\}$ is the set of nodes. The arc sets are $A =  \{(i,j) \hspace*{1mm} \vert \hspace*{1mm} i, j \in V \hspace{1mm}\}$, $A^D = \{(i,j) \hspace*{1mm} \vert \hspace*{1mm} i, j \in D \hspace{1mm} \}$, and $A^L = \{(i,j) \hspace*{1mm} \vert \hspace*{1mm} i, j \in L \hspace{1mm} \}$. $A$ corresponds to allowed movements of drivers from node $i$ to node $j$. For each driver $k$, the allowed movements are the following:
\begin{itemize}
    \item From the starting depot $s_k$ to a loading dock $l \in L$ or to the ending depot $e_k$.
    \item From a loading dock $l \in L$ to a delivery node $d \in D$.
    \item From a delivery node  $d \in D$ to a loading dock $l \in L$ or to the ending depot $e_k$.
\end{itemize}
For a customer $c$, arcs in $A^D$ link consecutive delivery nodes of the same order   $\lbrace (i,j)\in D_o, o \in O_c, i < j  \rbrace$, and pair of delivery nodes of two different orders $ \lbrace (i,d^{0}_{o_2}),  i \in D_{o_1}, i \geq n^{min}_{o_1}, o_1, o_2 \in O_c, o_1 \neq o_2 \rbrace $. Arcs in $A^L$ link all pairs of loading docks of the same batch plant.

We define $\delta^{+}(i) = \{(i, j) \in A \}$ and $\delta^{-}(i) = \{(j, i) \in A \}$ as the outcoming and incoming arc sets of node $i \in V$; $\delta^{+}_D(i) = \{(i, j)  \in A^D\}$ and $\delta^{-}_D(i) = \{(j, i) \in A^D \}$ are the outcoming and incoming arc sets of delivery node $i \in D$.  Similarly, $\delta^{+}_L(i) = \{(i, j)  \in A^L\}$ and $\delta^{-}_L(i) = \{(j, i) \in A^L \}$ are the outcoming and incoming arc sets of loading node $i \in L$. $t_{ij}$ is the time to travel from $i$ to $j$.

Let the binary variable $x^{k}_{ij}$ be $1$ if driver $k$ travels from node $i$ to $j$. Binary variable $y_o$ is $1$ when order $o$ is completely served; $v_i$ and $w_i$ are the start and end of the loading (unloading) operation at node $i \in L\cup D$. Binary variable $u_{ij}$ is $1$ if node $j$ is served just after $i$, the service being either an unloading or a loading operation. Variable $q^k_j$ is the quantity to be loaded towards $j$ with vehicle $k$. Let $w^1_k$ be a continuous variable indicating the difference between the driver's work time and the minimum number of hours to be worked in a day, and $w^2_k$ indicating the difference between the driver's work time and the normal work time. Let $g_i$ be the time between the due date and the first service start for customer $i$.

The objective function minimizes total travel cost ($TC$), penalties associated with unfulfilled orders, first delivery delays ($FDD$), driver underutilization costs ($DUC$), and driver overtime costs ($DOC$). Together, these components drive the optimization process to find a solution that efficiently balances travel costs, customer satisfaction, on-time delivery, driver utilization, and scheduling constraints.

\begin{alignat}{3}
    \label{mod:obj}   \min & \sum_{k \in K}{{ \sum_{ \substack{(i,j) \in A} }} {t_{ij}x^{k}_{ij}}} + \beta_1\smashoperator{\sum_{ o \in O}}{  \left(1-y_o\right)} \hspace{-0mm}  + \beta_2 \smashoperator{ \sum_{i \in C } } { g_i  } +  \sum_{k \in K}\left({\beta_3 w^{1}_k  + \beta_4  w^{2}_k}\right) &                                                                                                  \\
    \label{mod:c1}         & \sum_{j \in \delta^{+}(i) }{x^k_{ij}} =1     & \forall i \in S, k \in K                                                                      \\
    \label{mod:c2}         & \sum_{j \in \delta^{-}(i)}{x^k_{ji}} = 1     & \forall i \in E, k \in K                                                                         \\
    \label{mod:c03}        & v_j \geq  w_i  - M\left(1-x^{k}_{ij}\right)   & \hspace{-20mm}  i \in S, j \in \delta^{+}(i),  k \in K      \\
    \label{mod:c3}         & v_j \geq  w_i + \alpha_{b} + t_{ij} - M\left(1-x^{k}_{ij}\right) & \hspace{-25mm} \forall b \in B, i \in L_b, j \in \delta^{+}(i),  k \in K     \\
    \label{mod:c4}         & v_j \geq  w_i + \rho + t_{ij} - M\left(1-x^{k}_{ij}\right)                                                                       & \hspace{-20mm} \forall i \in D  , j \in \delta^{+}(i), k \in K                         \\
    \label{mod:c5}         & w_{i} \geq v_{i}  + \frac{q^k_j}{\tau^l_b} -M\left(1- x^{k}_{ij}\right)   & \hspace{-25mm} \forall  b \in B,  i \in L_{b},  j \in \delta^{+}(i), k \in K \\
    \label{mod:c6}         & w_{j} \geq v_{j}  + {  \frac{q^k_j}{\tau^u_c} -M\left(1- x^{k}_{ij}\right) }     & \hspace{-25mm}  \forall c \in  C, j \in D_{c}, i \in \delta^{-}(j),  k \in K \\
    \label{mod:c7}         & w_{j} \leq v_{i}  + \Delta + M\left(1- x^{k}_{ij}\right)  & \hspace*{-40mm} j \in D,  i \in \delta^{-}(j),  k \in K \text{ }  \\
    \label{mod:c8}         & v_{d^0_{o}} \geq a_c    & \hspace{-40mm} \forall  c \in C, \forall o \in O_c \text{ }                           \\
    \label{mod:c9}          & g_{c} \geq v_{d^0_{o_1}} - a_c - M\left(  \sum_{ j \in \delta^{-}_{D}(d^0_{o_1})  }{u_{jd^0_{o_1}}}\right)  & \hspace{-40mm}  \forall  c \in C, \forall o_1 \in O_c \text{ }  \\
    \label{mod:c91}  & g_{c} - a_c \leq \tau^w_c  & \hspace{-20mm}  \forall  c \in C \text{ }   \\
    \label{mod:c101}       & v_{d^0_{o_1}} \geq w_j - M\left(1- u_{jd^0_{o_1}}\right)   & \hspace{-50mm}    \forall o_1 \in O_i,  j \in \delta_{D}^{-}(d^0_{o_1}) \text{ }                \\
    \label{mod:c10}        & v_{d^0_{o_1}}  \leq  w_j + \gamma^2 + M\left(1- u_{j,d^0_{o_1}}\right)     & \hspace{-40mm}  \forall  o_1 \in O,  j \in \delta_{D}^{-}(d^0_{o_1}) \text{ }                    \\
    \label{mod:c11}        & \sum_{ \substack{o_1 \in O_c }}{ \sum_{ j \in \delta^{-}_{D}(d^0_{o_1})}}{u_{j,d^0_{o_1} } } = |O_c|-1      & \hspace{-40mm} \forall  c \in C,|O_c| > 1   \text{ }   \\
    \label{mod:c12}        & \sum_{ \substack{o_1 \in O_c }}\sum_{ j \in \delta^{+}_{D}(d^0_{o_1})}{u_{d^0_{o_1},j } } = |O_c|-1      & \hspace{-40mm} \forall  c \in C, |O_c| > 1  \text{ }          \\
    \label{mod:c13}    & \hspace{3mm} \smashoperator{ \sum_{ j \in \delta^{+}_{D}(d^0_{o_1})}}{u_{d^0_{o_1},j } } \leq 1    & \hspace{-20mm}  \forall o_1 \in O \text{ }         \\
    \label{mod:c14}        & \hspace{2mm} \smashoperator{ \sum_{ j \in \delta^{-}_{D}(d^0_{o_1})}}{u_{j,d^0_{o_1} } } \leq 1      & \hspace{-20mm} \forall o_1 \in O \text{ }        \\
    \label{mod:c15}        & \hspace{2mm} \smashoperator{ \sum_{ j \in \delta^{+}_{D}(d^0_{o_1})}}{u_{d^0_{o_1},j } } +\smashoperator{ \sum_{ j \in \delta^{-}_{D}(d^0_{o_1})}}{u_{j,d^0_{o_1} } } \geq 1    & \hspace{-20mm} \forall o_1 \in O \text{ }    \\
    \label{mod:c16}        & v_{j} \geq w_{j-1} - M\left(1- u_{j-1,j}\right)   & \hspace{-40mm}  \forall o \in O, j \in D_{o}, j \geq 1  \text{ }    \\
    \label{mod:c17}        & v_{j}  \leq  w_{j-1} + \gamma^1 + M\left(1- u_{j-1,j}\right)    & \hspace{-50mm} \forall c \in C, \forall o \in O_c, j \in D_{o}, j \geq 1 \text{ }     \\
    \label{mod:c18}        & u_{j-1,j} \geq u_{j,j+1}      & \hspace{-60mm} \forall o_1 \in O,  j \in D_{o_1}, 1 \leq j \leq n_{o_{1}}-1 \text{ }   \\
    \label{mod:c19}        & u_{j-1,j} \geq \sum_{l \in L}{x_{lj}}   & \hspace{-40mm}  \forall o_1 \in O, j \in D_{o_1}, j \geq 1 \text{ }                   \\
    \label{mod:c20}        & v_{j} \geq w_{i} - M\left(1- u_{i,j}\right)   & \hspace{-40mm}  \forall i \in L, j \in \delta^{+}_{L}(i) \text{ }  \\
    \label{mod:c21}        & \hspace{3mm} \smashoperator{ \sum_{ j \in \delta^{+}_{L}(i)}}{u_{i,j } } \leq 1     & \hspace{-40mm}  \forall i \in L \text{ }         \\
    \label{mod:c22}        & \hspace{2mm} \smashoperator{ \sum_{ j \in \delta^{-}_{L}(i)}}{u_{j,i } } \leq 1                                      & \hspace{-40mm} \forall i \in L           \text{ }       \\
    \label{mod:c230}       & \hspace{2mm} \smashoperator{ \sum_{ j \in \delta^{-}_{L}(i)}}{u_{j,i } }  \geq \sum_{k \in K}\sum_{j \in  \delta^{-}(i) }{x^k_{ji}}    & \hspace{-40mm} \forall i \in L \text{ }                                                      \\
    \label{mod:c231}       & \hspace{2mm} \smashoperator{ \sum_{ j \in \delta^{-}_{L}(i)}}{u_{j,i } } \geq \smashoperator{ \sum_{ j \in \delta^{-}_{L}(i+1)}}{u_{j,i+1 } }   & \hspace{-20mm} \forall i \in L  \text{ }                                                        \\
    \label{mod:c23}        & \hspace{2mm} \smashoperator{ \sum_{ j \in \delta^{-}_{L}(i)}}{u_{j,i } } + \smashoperator{ \sum_{ j \in \delta^{+}_{L}(i)}}{u_{i,j } } \geq \sum_{k \in K}\sum_{j \in  \delta^{-}(i) }{x^k_{ji}}    & \hspace{-20mm} \forall i \in L  \text{ }                                                        \\
    \label{mod:c24}        & \sum_{k \in K}\sum_{j \in D_o}{q^k_j} = q_o            & \hspace{-40mm}  \forall o \in O  \text{ }           \\
    \label{mod:c240}       & q^k_j \leq \sum_{i \in  \delta^{-}(j)}{Q^k x^k_{ij}}   & \hspace{-40mm}  \forall j \in D, k \in K  \text{ }                                              \\
    \label{mod:c25}        & \sum_{k \in K}\sum_{j \in \delta^{+}(i)}{x^{k}_{ij} \leq 1}     & \hspace{-40mm} i \in L \cup D       \text{ }          \\
    \label{mod:c26}        & \sum_{k \in K}\sum_{j \in \delta^{+}(i)} {x^{k}_{ij} } =  \sum_{k \in K}\sum_{j \in \delta^{-}(i)}{x^{k}_{ji} }        & \hspace{-40mm}  i \in L \cup D  \text{ }       \\
    \label{mod:c29}        & w^1_k \geq  \mu + h_k - v_{e_k}            & \hspace{-40mm} \forall k \in K   \text{ }         \\
    \label{mod:c30}        & w^2_k \geq (v_{e_k} - h_k) - \eta   & \hspace{-40mm} \forall k \in K         \text{ }         \\
    \label{mod:c31}   & w_{e_k} \leq h_k + \varLambda  & \hspace{-20mm}  k \in K \hspace{0mm}       \text{ }                                            \\
    \label{mod:c32}        & 0 \leq q^k_j \leq Q^k   & \hspace{-40mm}  j \in D, k \in K      \text{ }                                                   \\
    \label{mod:c33}        & x^k_{ij} \in \{0,1\}      & \hspace{-40mm}  (i,j) \in A , k \in K   \text{ }                                            \\
    \label{mod:c34}        & u_{ij} \in \{0,1\}     & \hspace{-40mm}  (i,j) \in A^D \hspace{0mm}   \text{ }                                           \\
    \label{mod:c36}        & y_o \in \{0,1\}     & \hspace{-20mm}  o \in O            \text{ }                     \\
    \label{mod:c37}        & v_i \geq 0,  w_i \geq 0     & \hspace{-20mm}  i \in V         \text{ }                \\
    \label{mod:c38}        & w^1_{k} \geq 0,  w^2_k \geq 0  & \hspace{-20mm}  i \in V         \text{ }        \\
    \label{mod:cend}       & g_c \geq 0    & \hspace{-20mm}  c \in C.  \text{ }
\end{alignat}

$\beta_1$ to $\beta_4$ are the penalty coefficients of each component of the objective function (\ref{mod:obj}). Constraints (\ref{mod:c1}) and (\ref{mod:c2}) state that driver $k$ leaves his start node exactly once a day and returns to his end node. A driver cannot serve a node before the start of his shift with constraints (\ref{mod:c03}). Constraints (\ref{mod:c3}) and (\ref{mod:c4}) set a driver to take some time after loading or unloading to adjust the concrete or clean the truck before moving on to the next node. The duration of the loading operation depends on the plant's loading rate and the amount $q^k_j$ of RMC loaded (\ref{mod:c5}). Similarly, the unloading service depends on the site's rate and $q^k_j$ (\ref{mod:c6}). Unloading operations must end at most $\Delta$ minutes after loading begins (\ref{mod:c7}).

Constraints (\ref{mod:c8}) to (\ref{mod:c10}) ensure that the first service of any customer $c$ must start after the due time $a_c$. This first service may be performed at the first delivery node of any order, and one order must be completed before another is started. It also enforces the precedence constraints between the last delivery of an order and the first delivery of the following order. Constraints (\ref{mod:c11}) to (\ref{mod:c15}) find the delivery sequence of all orders. With constraints (\ref{mod:c16})--(\ref{mod:c19}), two trucks cannot unload at the same time for consecutive deliveries of the same order. Additionally, these constraints impose a maximum time delay between the two trucks. Similarly, constraints (\ref{mod:c20})--(\ref{mod:c23}) ensure that two trucks cannot be loaded at the same time at a plant. Constraints (\ref{mod:c24}) require that the cumulative load of all concrete mixers serving an order must equal the required quantities, and (\ref{mod:c240}) bound a driver load. Constraints (\ref{mod:c25}) require a driver to only visit a loading/delivery node once. Constraints (\ref{mod:c26}) are degree constraints. Constraints (\ref{mod:c29})--(\ref{mod:c30}) calculate the difference between a driver's hours of service and the minimum and normal hours of service. Finally, constraints (\ref{mod:c31}) to (\ref{mod:cend}) define the nature and bounds of the variables.

Figure~\ref{fig:plants_schedules} presents a solution for an instance. The instance has two plants, nine drivers, and five customers, with two customers having multiple orders. Customer $4$ has two orders ($o_4$--$o_5$) assigned to plant $1$, and customer $5$ has three orders ($o_{6}$--$o_{8}$) assigned to plant $2$. All the order requirements are in the Table~\ref{tab:instance_detail}. Figure \ref{fig:plants_schedules} provides a visual representation using a Gantt chart of the loading and unloading operations performed by the drivers at each plant. The light gray, dark gray and black colors represent a customer's first, second and third order. We can see that the loading operations for each plant do not overlap, and the unloading operations occur sequentially, one after the other, at each worksite. At plant $1$, loading operations for $C3$ are performed first by drivers $3,5,4,2$ followed by loadings for  $C1$. The second order of $C5$ is delivered, followed by the second, and finally, the first order. All first deliveries start at their due time, except for $C2$, which is 20 minutes late. There is a delay between each delivery of $C2$. However, it is less than the maximal time delay of 25 minutes defined for this example. We can see that some drivers load at both plants. For example, driver $5$ loads at plant $1$ to perform the second delivery of $C3$ then drives to plant $2$ for the first delivery of $C2$. Driver $1$ is originally assigned to plant $1$ but starts his first service at $2$ before returning to his home plant. The loadings at plant $2$ show how the loading dock is alternately assigned for orders of customers $2,4$ and $5$ between $06\text{:}25$ and $10\text{:}00$. We obtain this solution using a heuristic solution approach that we describe next.

\begin{table}[!ht]
    \centering
    \caption{Information of the illustration instance}
    \label{tab:instance_detail}
    \resizebox*{0.8\textwidth}{!}{
        \begin{tabular}{@{}c|cccccccccccccc@{}}
            \toprule
            Order         & 1     & 2     & 3     & 4     & 5     & 6     & 7     & 8       \\ \midrule
            Customer      & 1     & 2     & 3     & 4     & 4     & 5     & 5     & 5       \\
            \midrule
            Demand   & 24   & 32   & 36   & 40   & 16  & 20  & 32  & 8 \\
            \midrule
            Plant    & 1     & 2     & 1     & 2     & 2    & 2  &  2 & 2     \\
            \midrule
            TW       & 08:45 & 08:00 & 07:00 & 07:00 & 07:00 & 06:50 & 06:50 & 06:50  \\ \bottomrule
        \end{tabular}
    }
\end{table}

\begin{figure}[htb]
    \centering
    \caption{Gantt chart of the loading and unloading operations of the illustration instance}
    \label{fig:plants_schedules}
        \includegraphics[width=\linewidth]{exemple.pdf}
\end{figure}

\section{Constructive heuristics and GRASP}
\label{sec:cdp_grasp_method}

This section presents the heuristic approach to solving the Concrete Delivery Problem. Our method consists of constructing feasible solutions of the CDP with randomized greedy heuristics and iteratively invoking these heuristics in the GRASP metaheuristic. From now on, we represent the scheduling of a delivery node $d$ as the pair ($L^k_{d}$,$U^k_{d}$) of loading and unloading tasks. $L^k_{d} = \left(b,q_d,v_b,w_b\right)$ indicates that driver $k$ will load $q_d$ $m^3$ of RMC at plant $b$ between timeslots $\left[v_b,w_b\right]$ to serve $d$. $U^k_{d} = \left(k,v_d, w_d\right)$ indicates that driver $k$ will start unloading at time $v_d$ and finish at time $w_d$. We then call a solution $Sol =\left\lbrace \cup _{1 \leq j \leq |D|} (L^k_{j}, U^k_{j}), k \in K \right\rbrace$ the set of all pairs of loading and unloading tasks performed on a given day.

\subsection{GRASP algorithm}

The Greedy Randomized Search Procedure (GRASP) algorithm is a multi-start metaheuristic for combinatorial optimization problems proposed by \cite{feo1989probabilistic}. It is a two-phase iterative sequence of constructive and local search algorithms. A feasible solution is generated in the construction phase using a greedy randomized algorithm. Then, a local search phase tries to improve it. Algorithm~\ref{alg_grasp} outlines the GRASP algorithm. We start with a list of greedy randomized heuristics and iterate on each to construct an incumbent solution $Sol$ on which we apply a local search procedure to explore its neighborhood. The best solution found so far is then stored. To avoid redundant computations, a hash table $T$ ensures that the local search is not applied twice to the same solution. The procedure stops and returns the best overall solution when certain stopping conditions are met, such as reaching a time limit or a maximum number of iterations. 
    \begin{algorithm}[hbpt]
        \caption{Pseudo-code of the GRASP algorithm }
        \label{alg_grasp}
        \DontPrintSemicolon
        \LinesNumbered
        \setcounter{AlgoLine}{0}
        \KwIn{ \textit{H}: List of constructive greedy randomized heuristics }
        $Sol^{*} \leftarrow \emptyset$    \hspace{2mm}        $Cost(Sol^*) \leftarrow \infty$

        Initialize an empty hash table $T \leftarrow \emptyset$

        % $counter \leftarrow  0$

        \While{conditions not met}{

            \ForEach{$heuristic \in H$}{
                $Sol \leftarrow heuristic()$

                \uIf{ $Sol \notin T$}{
                    $Sol \leftarrow LocalSearch(Sol)$

                    Add $Sol$ to $T$

                    \uIf{$Cost(Sol)< Cost(Sol^*)$}{
                    $Sol^* \leftarrow Sol$
                    }
                }
            }
        }

        \KwRet{$S^{*}$}
    \end{algorithm}

For the construction phase, we developed two greedy randomized heuristics: the customer-based insertion with backtracking, which schedules customers one at a time, and the delivery-based insertion.%, which does not necessarily complete the scheduling of one customer before starting to schedule another.
% The greedy aspect of GRASP is the creation of the set $RCL$, the probabilistic aspect is the random selection in $RCL$, and the adaptive aspect is the updating of $CL$ and the re-evaluation of the incremental costs.

\subsection{Customer-based insertion with backtracking heuristic}

The customer-based insertion heuristic builds a solution from scratch by iteratively scheduling one customer at a time. When a customer is selected, all deliveries needed to serve that customer are scheduled before moving on to the next customer. As shown in Algorithm~\ref{alg:greedyCustIns}, between lines $4$ and $38$, we iterate over a shuffled list of orders for each customer.

To schedule an order $o$, we begin with its first delivery node. Once this node is scheduled, we proceed to schedule the subsequent delivery nodes until $o$ is completely served. For each delivery node $d$, the algorithm identifies between lines $9$ and $12$ each driver capable of visiting this node using the $SimulateVisit$ procedure. For all drivers able to serve $d$, at line $12$ we store the details about the pair of loading and unloading tasks and its cost in $M[d]$.

If at least one driver can serve the node (line $13$), we randomly select at line $16$  a pair of loading and unloading tasks whose cost falls within a specified range based on a parameter $\theta$ and insert this pair into $Sol$ at line $18$. Then we update the order remaining demand and proceed with the next delivery node at line $19$.

In cases where no driver can visit the node $d$, the algorithm employs a backtracking scheme to address two distinct scenarios. First, if the visit to $d$ cannot be scheduled due to cold-joint constraints (i.e., the time between consecutive deliveries exceeds $\gamma^1$), the $DelayPreviousDelivery$ procedure is employed. This procedure determines if a previous delivery $D^{j'}_o$ of the same order exists such that $d$ can be successfully scheduled in the current solution by delaying the start of service of $D^{j'}_o$ by $\lambda$. If $D^{j'}_o$ exists, the schedules of all delivery nodes between $D^{j'}_o$ and $d$ are removed, and the scheduling of order $o$ resumes from $D^{j'}_o$.

The procedure $DelayPreviousDelivery$ computes the difference $\lambda$ between the earliest time a driver can visit $D^{j}_o$ and the end of service of its immediate predecessor node $D^{j-1}_o$. Then, it tries to find one of the precedent nodes of $D^{j}_o$ for which we can delay the start of service of at least $\lambda = min(\lambda,\gamma^1)$. If such a node exists, the procedure returns it and the value of the delay. Then we backtrack to this node and restart the scheduling process.

In cases where infeasibility arises from other violated constraints, the algorithm attempts to backtrack to a precedent node and retry the scheduling with another driver. At line $31$, the algorithm searches for the first precedent node $D^{j'}_o$ with remaining pair of tasks in $M[D^{j'}_o]$. If found, the schedules of all delivery nodes between $D^{j'}_o$ and $d$ are removed, and backtracking occurs to $D^{j'}_o$. In this iteration, since $M[D^{j'}_o]$ is not empty, the $SimulateVisit$ procedure is not called again; instead the stored elements of $M[D^{j'}_o]$ are utilized.

If no precedent node suitable for backtracking is found, it indicates that the current customer cannot be scheduled. Consequently, all previously scheduled visits to its delivery nodes are removed before advancing to the next customer (as described in lines $23$--$25$ and lines $32$--$34$). The procedure stops when all customers are evaluated and returns the constructed solution $Sol$ at the end.

The $SimulateVisit$ procedure takes a partial solution, a delivery node $d$, a driver $k$ and the delay value $\lambda$ as parameters. Its goal is to find the earliest time when the loading service for $d$ can start at its home plant $b$. When planning to visit delivery node $d$ of order $o$ and customer $c$ with driver $k$, we first compute his delivery time $v_d$ as follows. If $d$ is the first  visited node of customer $c$, then we expect its service to start at $v_d = a_c + \lambda$. Otherwise, if the previous node $j$ visited before $d$ belongs to the same order $o$, then the service at $d$ may begin at the random value $w_j + rand(\lambda, \gamma^1)$ to respect the allowed time between two consecutive deliveries of the same order. We randomize $v_d$ to allow more diversity in our solution space. Finally, if $j$ and $d$ belong to different orders, then $v_d = w_j + rand(0,\gamma^2)$.

Let $\bar{q}_o$ be the remaining demand of order $o$. Driver $k$ can supply $d$ with the quantity $q^k_d= \min(\bar{q}_o,Q^k)$. We determine the loading start time $v_l$ of the corresponding dock node $l$ by subtracting the travel time, adjustment time, and loading duration from $v_d$: $w_l = v_d - t_{dl} - \alpha_b - q^k_{d}/\tau^l{_b} $. If $j$ is the last node where $k$ was unloading with a service duration $s_j$, then $w_l= \max(w_l, v_j + s_{j} + t_{jl})$. Now, we have the earliest time $k$ can start loading at plant $b$, but if the loading dock is occupied at that time, $k$ will have to wait. To find the real loading time, we use a data structure that stores all timeslots when a plant is occupied. Then, starting from $w_l$, we search forward for the first timeslot with duration $q^k_{d}/\tau^l{_b}$ where we can insert this loading operation and update $w_l$ accordingly. Changing $w_l$ can cause violations of the time window or cold-joint constraints, especially for an order's subsequent deliveries, and thus make the schedule infeasible. So with this new value of $w_l$, we recompute $v_l$ and check for constraint violations.

Algorithm~\ref{alg:greedyCustIns} is designed with a list of customers as input. Initial experimentation has revealed that the sequence of this list affects the final solution. Within the GRASP framework, we invoke the customer-based heuristic multiple times with various customer orderings. These include sorting by decreasing demand and time window, sorting by increasing demand and time window, and finally, with a randomized list.
    \begin{algorithm}[!ht]
        \caption{Customer-based insertion algorithm }
        \label{alg:greedyCustIns}
        \DontPrintSemicolon
        \LinesNumbered
        \setcounter{AlgoLine}{0}
        \KwIn{  $Sol$: empty solution, $K$: vehicles set, $\Pi$: list of customers }

        $M \leftarrow \emptyset $ \tcp{List of pairs of loading -- unloading tasks for each delivery node}

        \ForEach{$c \in \Pi$}{

            Shuffle $ O_c$

            \ForEach{$o \in O_c$}{

                $\lambda \leftarrow 0$, $\bar{q}_o\leftarrow q_o$, $j \leftarrow 1$

            \While{$\bar{q}_o>0$}{

                $d \leftarrow D^j_{o}$

                \uIf{($M[d] = \emptyset $) or ($\lambda >0$)}{

                    \ForEach{ $k \in K$  }{

                        $(L^k_{d}$, $U^k_{d},cost,f)$ $\leftarrow $ SimulateVisit($Sol$,$d$,$k$,$\lambda$)

                        \uIf{f=1}{
                                $M[d] \leftarrow M[d] \cup \{(L^k_{d}, U^k_{d},cost,k)\}$ 
                            }
                    }
                }

                \uIf{$M[d]\neq \emptyset$}{

    $c_{min} \leftarrow \min\left\{ cost \ | \ (L^k_{d}, U^k_{d},cost,k) \in M[d]   \right\} $

    $c_{max} \leftarrow \max\left\{ cost \ | \ (L^k_{d}, U^k_{d},cost,k) \in M[d]   \right\} $

    Randomly choose $(L^k_{d}, U^k_{d},cost,k)$ in $M[d]$ such that $cost \leq c_{min} + \theta (c_{max}-c_{min}) $

    $M[d]\leftarrow M[d] \backslash \{(L^k_{d}, U^k_{d},cost,k)\} $

    $Sol \leftarrow Sol \cup \{(L^k_{d}$, $U^k_{d})\}  $

        $\bar{q}_o \leftarrow \bar{q}_o-q^k_{d}$, $\lambda \leftarrow 0$, $j \leftarrow j+1$

    }
    \uElse{
        \uIf{cold-joint constraints}{
            $(D^{j'}_o,delay)$ = $DelayPreviousDelivery$($Sol$,$d$) \tcp{$j'<j$}

        \uIf{no $D^{j'}_o$ exists}{

            $Sol \leftarrow Sol \backslash \{(L^k_{i}, U^k_{i}), \forall i \in D_c\}$

            Go to line 2 \tcp{continue with next customer}
}

$Sol \leftarrow Sol \backslash \{(L^k_{i}, U^k_{i}), \forall i \in (D^{j'}_{o},..,D^{j-1}_{o})  \}$

Update $\bar{q}_o$

$j \leftarrow j'$, $\lambda \leftarrow delay$

    Go to line 6 \tcp{backtrack to $j'$ position and retry to insert $D^{j'}_{o}$}
    }
    \uElse{
     Search for a precedent delivery node $D^{j'}_o, j'<j$ with $M[D^{j'}_o]\neq \emptyset$

    \uIf{ no $D^{j'}_o$ exists}{

    $Sol \leftarrow Sol \backslash \{(L^k_{i}, U^k_{i}), \forall i \in D_c\}$

    Go to line 2 \tcp{continue with next customer}
    }

    $Sol \leftarrow Sol \backslash \{(L^k_{i}, U^k_{i}), \forall i \in (D^{j'}_{o},..,D^{j-1}_{o})  \}$

    Update $\bar{q}_o$

     $j \leftarrow j'$

    Go to line 6 \tcp{backtrack to precedent delivery node $D^{j'}_o$ }
                        }

                    }

                }
            }
        }
        \KwRet{$Sol$}
    \end{algorithm}

When confronted with scenarios where several customers require their services to start simultaneously or within the same time window, we observed that with Algorithm~\ref{alg:greedyCustIns}, some customers remain unscheduled when resources are limited. To address this limitation, we introduce the delivery-based insertion heuristic.

\subsection{Delivery-based insertion heuristic}

The delivery-based insertion heuristic (Algorithm~\ref{alg:delbasedIns}) iteratively constructs a solution from scratch by scheduling one delivery node at a time. The algorithm starts with an empty solution $Sol$ and initially fills a candidate list $CL$ with the first node of a randomly selected order for each customer. Within each iteration of the $while$ loop, the algorithm filters $CL$ to obtain a reduced candidate list $CL'$ of deliveries based on a specified criterion, such as prioritizing deliveries with the highest (lowest) demands, earliest (latest) due dates, or lowest order numbers. This filtering is performed by the $Filter$ procedure, allowing the algorithm to explore different neighborhoods. Thus, within the GRASP framework, we can invoke the delivery-based heuristic multiple times with different filtering procedures which evaluate different neighborhoods.

Between lines $14$ and $18$, for each delivery $d$ in $CL'$, the algorithm determines the best available driver $k$ who can visit it. Then, the corresponding pair of loading and unloading tasks $(L^k_d, U^k_d)$, along with its costs, are stored in $M[d]$. This is done using the $SimulateVisit$ procedure described earlier. Then, the algorithm constructs at line $20$ a list $T$ containing all pairs of loading and unloading tasks for the delivery nodes in $CL'$.

If $T$ is empty, the algorithm assumes that the customers associated with the nodes in $CL'$ cannot be scheduled, and it removes those customers' existing schedules from the solution. It then excludes $CL'$ from $CL$ at line $25$ and proceeds to the next iteration.

Otherwise, from line $26$ to line $29$, the algorithm randomly selects from $T$ a pair of loading and unloading tasks whose cost falls within a specified range determined by the parameter $\theta$ and inserts this pair into $Sol$. The algorithm then updates the remaining orders and customer quantities and excludes the scheduled delivery node from $CL$. If an order $o$ is not completely fulfilled after visiting $d^j_{o}$, the next delivery node $d^{j+1}_{o}$ is added to $CL$.  Otherwise, if the customer has remaining orders, the first delivery node $d^{1}_{o'}$ of another randomly selected order $o'$ is added to $CL$.

Unlike the customer-based insertion, the delivery-based insertion allows each customer to have one of its delivery nodes considered for scheduling at each iteration.
%In these two heuristics, the greedy aspect of GRASP refers to the creation of the set $RCL$, the probabilistic aspect is the random selection in $RCL$, and the adaptive part is the updating of $CL$ and the re-evaluation of the costs.

\begin{algorithm}[!ht]
    \caption{Delivery-based insertion algorithm }
    \label{alg:delbasedIns}
    \DontPrintSemicolon
    \LinesNumbered
    \setcounter{AlgoLine}{0}
    \KwIn{  $Sol$: empty solution, $\Pi$: list of customers, $Filter$: filtering procedure}

    $CL \leftarrow \emptyset$

    $M \leftarrow \emptyset $ \tcp{List of pairs of loading -- unloading tasks for each delivery node}

    $\bar{q} \leftarrow \emptyset$ \tcp{List of remaining demands}

    \ForEach{$c \in \Pi$}{
        $\bar{q}_c \leftarrow q_c$

        Select a random order $o \in O_c$

        $CL \leftarrow CL \cup \{d^1_{o}\}$

        $\bar{q}_o \leftarrow q_o$
    }

    \While{$CL \neq \emptyset$}{

        $CL' \leftarrow \text{} Filter (CL) $

        $T \leftarrow \emptyset $ \tcp{List of loading and unloading tasks}

        \ForEach{ $d \in CL'$}{

            $best\_cost \leftarrow +\infty$

            \ForEach{ $k \in K$  }{

                $(L^k_{d}$, $U^k_{d},cost,f)$ $\leftarrow $ SimulateVisit($Sol$,$d$,$k$,$\lambda$)

                \uIf{f=1 and $cost <       best\_cost$}{
                    $M[d] \leftarrow \{(L^k_{d}, U^k_{d},cost,k)\}$

                    $best\_cost \leftarrow c$
                }
            }

            \uIf{$M[d]\neq \emptyset$}{
                $T \leftarrow T \cup M[d]$    
            }
        }
        \uIf{$T = \emptyset$}{
            \ForEach{$d \in CL'$}{

           $c \leftarrow $ customer associated with $d$

            $Sol \leftarrow Sol \backslash \{(L^k_j,U^k_j), \forall j \in D_c \}$
           }
           $CL \leftarrow CL\backslash CL'$ 
        }
        \uElse{
            $c_{min} \leftarrow \min\left\{ cost \ | \ (L^k_{d}, U^k_{d},cost,k) \in T   \right\} $

            $c_{max} \leftarrow \max\left\{ cost \ | \ (L^k_{d}, U^k_{d},cost,k) \in T   \right\} $

            Randomly choose $(L^k_{d^j_o}, U^k_{d^j_o},cost,k) \in T$ such that $cost \leq c_{min} + \theta (c_{max}-c_{min}) $

            $Sol \leftarrow Sol \cup \{(L^k_{d^j_o}$, $U^k_{d^j_o})\} $

            $\bar{q}_o \leftarrow \bar{q}_o-q^k_{d^j_o}$, $\bar{q}_c \leftarrow \bar{q}_c-q^k_{d^j_o}$ \tcp{$o \in O_c$}

            $CL \leftarrow CL \backslash \{d^{j}_o\}$

            \uIf {$\bar{q}_{o} \neq 0$ }{

                $CL \leftarrow CL \cup \{d^{j+1}_{o}\}$
            }
            \uElseIf{ $\bar{q}_c \neq 0$}{

                Select another order $o' \in O_c$

                $CL \leftarrow CL \cup \{d^{1}_{o'}\} $
            }

        }      
    }
        \KwRet{$Sol$}
\end{algorithm}

\subsection{Local Search}

After the construction phase, the local search phase invokes local search operators to either repair or improve the incumbent. One operator is used to repair a solution where not all customers are scheduled. Then four operators are used to improve the incumbent using a first-improvement strategy. The local search continues until no more improvements can be made or a timeout is reached.

\subsubsection{Repair operator}

The Remove and Reschedule operator is used to address situations where a solution contains unscheduled customers. This operator first selects an unscheduled customer $c$ from the solution and identifies a set $\Pi_c$ containing scheduled customers sharing the same time slot with $c$. The schedules of all elements of $\Pi_c$ are then removed from the solution and retried to be scheduled again along with $c$.

\subsubsection{Improvement operators}

The first improvement or Load Backward operator is used when dealing with solutions containing customers whose first delivery is late. In such cases, the goal is to schedule the loading start time for the first delivery node for these customers as early as possible. This operator is implemented by iterating backward through the timeslots data structure when calculating the loading start time within the $SimulateVisit$ procedure. By doing so, the operator aims to find a loading start time that allows for earlier delivery, thus reducing the late first delivery cost. The other improvement operators are based on the swap and relocate moves.

We have implemented two moves based on the swap operator. The Swap Driver operator explores alternative driver assignments by swapping drivers assigned to two delivery nodes. This allows for the assessment of different driver configurations, potentially leading to improvements in solution quality. The Swap Load operator exchanges loads between a delivery node and the last delivery node of the same order. In our greedy randomized heuristics, all deliveries of an order are typically filled to capacity, except for the last one.  This operator investigates the effect of delivering this last load much earlier in the delivery sequence. Delivering the final load earlier in the sequence can improve the loading operations, as it may be easier to find an empty timeslot for inserting a loading task with a smaller load.

We also implemented two moves based on the relocate operator. The Consolidate Load operator is applied on an order with a remaining demand between $]Q_{min},Q_{max}]$. If this remaining demand is served with two drivers with capacity $Q_{min}$, then the operator tries to consolidate it to be served by a single driver with a capacity of $Q_{max}$. Next, the Relocate Driver operator removes a delivery node from its current route and inserts it in another driver's route. This enables the redistribution of delivery nodes among drivers, potentially reducing drivers underutilization and overtime costs.

\section{Computational experiments}
\label{sec:cdp_comp_exp}

In this section, we present and discuss the results of computational experiments conducted using the GRASP heuristic described in Section~\ref{sec:cdp_grasp_method}. The implementation of the GRASP algorithm is coded in C++. We run our experiments on the benchmark data used in \cite{kinable2014concrete}, and on a new dataset extracted from delivery operations records provided by our industry partner.

\subsection{Generation of instances}
\newcommand{\nbInstance}{35}
The dataset used in this section was obtained from our partner and contains records representing delivery operations from up to 8 plants for $\nbInstance$ days. From this historical data, we extracted information such as daily orders delivered, plant assignments to orders, and the number of available drivers. We created $\nbInstance$ instances from this data, where each instance corresponds to a daily operation. To maintain confidentiality, we have removed the GPS coordinates of customers and plants but have included two matrices containing the distances and driving times between all plants and customers. An instance contains information such as each driver's capacity, associated batch plant, and shift start time (which we generate). For each customer, we have the due time, the total demand, the number of orders (type of concrete) received, and its index in the time (distance) matrix. For each order, we have its demand and the corresponding customer and production plant. Finally, we have the loading capacity for each plant, as well as its index in the time matrix. The name of an instance has the format $C\_n\_o\_k\_p$, where $n$ is the number of customers, $o$ is the total number of orders, $k$ is the number of available drivers, and $p$ is the number of plants. The dataset is divided into small, medium, and large sets according to the total number of daily orders. We have summarized the dataset in Table~\ref{tab:cdp_instances}. The small dataset has eight instances with four to eleven customers whose demands vary between 226.5 and 937.5 m$^3$. These customers request between 7 and 14 orders, which are served from one to three depots with a fleet of 13 to 35 concrete mixers. The dataset is available online at \url{https://sites.google.com/view/jfcote/} as a contribution to the CDP literature.
\begin{table}[htpb]
    \centering
    \caption{Instances summary}
    \label{tab:cdp_instances}
    \resizebox*{0.9\textwidth}{!}{
        \begin{tabular}{lcccccc}
            \toprule
            {}     & \# & Demand (m$^3$)           & \#Orders   & \#Customers & \#Drivers & \#Depots \\
            \midrule
            Small  & 8  & 226.5 -- 937.5     & 7 -- 14    & 4 -- 11   & 13 -- 35    & 1 -- 3     \\
            Medium & 14 & 1,160.5 -- 2,971.0 & 43 -- 98   & 40 -- 89  & 76 -- 137   & 6 -- 8     \\
            Large  & 13 & 3,078.5 -- 3,953.5  & 107 -- 136 & 92 -- 114 & 129 -- 149  & 8       \\
            \bottomrule
        \end{tabular}
    }
\end{table}

\subsection{Results for the instances of \cite{kinable2014concrete}}

\cite{kinable2014concrete} provided a benchmark dataset for their variant of the CDP. Although their primary objective is to maximize the total load delivered each day without considering loading operations at a plant and ensuring deliveries within specified time windows, we still use this dataset to analyze and compare the performance of our algorithm under different conditions. \cite{kinable2014concrete} solved their problem with a CP algorithm, MIP, steepest descent heuristic (SD), and fix-and-optimize MIP heuristic (MIP fix \& opt), which is a hybridization of the MIP and SD. They also provide dual bounds to help assess the optimality of the instances. Table \ref{tab:cdp_lit_comp} shows the performance of the GRASP heuristic against the methods mentioned above. For each method, the table shows the average execution time in seconds and the corresponding gap between the obtained solution and the dual bound provided by \cite{kinable2014concrete}. The row Opt reports the number of optimal solutions found by each method.

\begin{table}[htbp]
    \centering
    \caption{Performance of the GRASP heuristic on the CDP benchmark instances}
    \label{tab:cdp_lit_comp}
    \scriptsize
    \resizebox{1\textwidth}{!}{
        \begin{tabular}{@{}cccccccccc|cc@{}}
            \toprule
            & &  \multicolumn{8}{c}{\cite{kinable2014concrete}} & \multicolumn{2}{|c}{Our paper} \\
            \cmidrule(l){3-10} \cmidrule(r){11-12}
             &     & \multicolumn{2}{c}{CP}     & \multicolumn{2}{c}{MIP}   & \multicolumn{2}{c}{SD-heuristic} & \multicolumn{2}{c}{MIP fix \& opt} & \multicolumn{2}{|c}{GRASP}                                                 \\ \cmidrule(l){3-10} \cmidrule(r){11-12}
               &     & Gap (\%)                   & Time (s)                 & Gap (\%)                         & Time (s)                  & Gap (\%)                   & Time (s) & Gap (\%) & Time (s) & Gap (\%) & Time (s) \\ \cmidrule(l){3-10} \cmidrule(l){11-12}
            \multirow{2}{*}{Set A} & Avg & 4.2                        & 196.95                  & 7.3                              & 149.14                 & 9.1                        & 0.02        & 7.0      & 100.70   & 5.7      & 13.18    \\ \cmidrule(l){2-10} \cmidrule(r){11-12}
                & Opt & \multicolumn{2}{c}{40/64}  & \multicolumn{2}{c}{37/64} & \multicolumn{2}{c}{30/64}        & \multicolumn{2}{c|}{35/64}  & \multicolumn{2}{c}{35/64}                                                            \\ \cmidrule(l){3-10} \cmidrule(r){11-12}
            \multirow{2}{*}{Set B} & Avg & 12.1                       & 355.95                   & -                                & -                          & 16.3                       & 1.33     & -        &  \multicolumn{1}{c|}{-}         & 13.8     & 574.98   \\ \cmidrule(l){2-10} \cmidrule(r){11-12}
              & Opt & \multicolumn{2}{c}{55/128} & \multicolumn{2}{c}{-}     & \multicolumn{2}{c}{40/128}       & \multicolumn{2}{c|}{-}      & \multicolumn{2}{c}{44/128}                                                           \\ \bottomrule
        \end{tabular}
    }
\end{table}

Overall, our GRASP algorithm shows its effectiveness in providing good and competitive solutions in a relatively short time for this benchmark of the CDP, making it a suitable candidate for further analysis and real-world application.  It outperformed all methods except CP, achieving an average gap of 5.7\% (13.8\%) with a runtime of 13 (575) seconds for Set A (Set B). The SD-heuristic appears to be much faster (23 milliseconds for set A), but GRASP provides a better balance between solution quality and computation time. It should also be noted that the exact MIP, CP, and MIP fix \& opt are initialized with the results of the SD-heuristic.

\subsection{Results for the generated instances}

The computational results of our instances solved with the GRASP heuristic are summarized in the tables below. The parameters used in our experiments are listed in Table~\ref{tab:cdp_problem_parameters}.

\begin{table}[!ht]
    \centering
    \caption{Parameters of the real-world instances}
    \label{tab:cdp_problem_parameters}
    \small
    \begin{tabularx}{\textwidth}{Xr}
        \toprule
        \multicolumn{1}{c}{Parameter}  &
        \multicolumn{1}{c}{Value} \\ \midrule
 Time window ($\tau^w$)  & 60 min    \\
 Unloading duration ($\tau_c^u$)  & 2 min$/m^3$   \\
 Cleaning duration ($\rho$)    & 10 min  \\
        Adjustment duration ($\alpha$)                                                & 10 min                    \\
        Max travel time ($\Delta$)                                                    & 120 min                   \\
        Min working time ($\mu$)                                                      & 180 min                   \\
        Max working time ($\eta$)                                                      & 480 min                   \\
        Max overtime duration ($\varLambda$)                                                 & 120 min                   \\
        Max delay between two consecutive deliveries of the same order $\gamma^1$)   & 20 min                    \\
        Max delay between two consecutive deliveries of different orders ($\gamma^2)$ & 25 min                    \\
        Penalty of unfulfilled orders ($\beta_1$)                                     & 100,000                   \\
        Penalty of first delivery delay ($\beta_2$)                                   & 10,000                    \\
        Penalty for driver underutilization  ($\beta_3$)                              & 30                        \\
        Driver overtime penalty $ (\beta_4$)                                          & 20                        \\
        \bottomrule
    \end{tabularx}
\end{table}

Table \ref{tab:cdp_all_instances} contains the detailed results for all instances with a stop criterion of 3600 seconds. The table shows the demand of each instance, the average, and the best values of the objective function when solving each instance five times. All daily orders of the small instances are completely served by our algorithm within these five iterations, as shown in column $UQ$, which reports the undelivered quantities. For the medium and large instances, $99.5\%$ and $99.6\%$ of the total demands are delivered. The sum of driver underutilization costs ($DUC$) is high for medium and large instances. This indicates the underutilization of the scheduled fleet. Some drivers are not scheduled at all or work very little. The sum of driver overtime costs ($DOC$) is also high for these instances. This can be explained by the fact that the algorithm prioritizes drivers on overtime rather than using a driver who has to travel to another plant for a delivery, which would increase travel costs ($TC$).

$FDD$ represents the sum of the delays of the first deliveries. In the case of small instances, all but one of the first deliveries are on time. However, for the other instance sets, there is at least one customer with a late first delivery. To provide insight into the actual delay experienced by customers, we report $mFDD$, which is the maximum delay among all first deliveries. This information reveals that for small instances, the maximum delay ranges between 0 and 35.7 minutes. For medium instances, it varies from 32.7 to 60 minutes, and for large instances, it falls within the range of 56 to 60 minutes.

\begin{table}[ht!]
    \centering
    \caption{Results with all instances}
    \label{tab:cdp_all_instances}
    \resizebox{\textwidth}{!}{
    \begin{tabular}{lcccccrrrrrrrrrrrrr}
    \toprule
     &  &  &  & & & \multicolumn{6}{c}{Avg} & \multicolumn{7}{c}{Best}  \\
    \cmidrule(l){7-12} \cmidrule(l){13-19}
     &  &  &  &  & \multicolumn{1}{c}{Demand}  & \multicolumn{1}{c}{$TC$}& \multicolumn{1}{c}{$UQ$} & \multicolumn{1}{c}{$FDD$} & \multicolumn{1}{c}{$DUC$} & \multicolumn{1}{c}{$DOC$} & \multicolumn{1}{c}{\multirow[c]{2}{*}{Z}}  & \multicolumn{1}{c}{($TC$)} & \multicolumn{1}{c}{$UQ$} & \multicolumn{1}{c}{$FDD$} & \multicolumn{1}{c}{$mFDD$} & \multicolumn{1}{c}{$DUC$} & \multicolumn{1}{c}{$DOC$} & \multicolumn{1}{c}{\multirow[c]{2}{*}{Z}}  \\
      & |$C$| & $|O|$ & $|K|$ & $|B|$ & \multicolumn{1}{c}{(m$^3$)}& \multicolumn{1}{c}{(min)} & \multicolumn{1}{c}{(m$^3$)} & \multicolumn{1}{c}{(min)} & \multicolumn{1}{c}{(min)} & \multicolumn{1}{c}{(min)} &  & \multicolumn{1}{c}{(min)} & \multicolumn{1}{c}{(m$^3$)} & \multicolumn{1}{c}{(min)} & \multicolumn{1}{c}{(min)} & \multicolumn{1}{c}{(min)} & \multicolumn{1}{c}{(min)} &  \\
    \midrule
    \multirow[c]{8}{*}{Small} & 11 & 12 & 13 & 1 & 226.5 & 1,243.5 & 0.0 & 0.0 & 18.8 & 0.0 & 1,807.7 & 1,243.5 & 0.0 & 0.0 & 0.0 & 0.0 & 0.0 & 1,243.5 \\
    & 5 & 8 & 13 & 1 & 267.0 & 745.1 & 0.0 & 0.0 & 0.0 & 0.0 & 745.1 & 745.1 & 0.0 & 0.0 & 0.0 & 0.0 & 0.0 & 745.1 \\
    & 6 & 11 & 18 & 2 & 333.5 & 753.3 & 0.0 & 0.0 & 255.9 & 0.0 & 8,430.6 & 753.3 & 0.0 & 0.0 & 0.0 & 176.0 & 0.0 & 6,034.2 \\
    & 4 & 7 & 15 & 2 & 375.0 & 1,017.5 & 0.0 & 0.0 & 218.2 & 131.3 & 10,190.9 & 1,017.6 & 0.0 & 0.0 & 0.0 & 82.8 & 33.9 & 4,179.8 \\
    & 7 & 8 & 19 & 2 & 388.0 & 1,790.5 & 0.0 & 0.0 & 318.4 & 137.8 & 14,119.2 & 1,817.0 & 0.0 & 0.0 & 0.0 & 52.6 & 98.7 & 5,368.6 \\
    & 10 & 14 & 29 & 3 & 613.5 & 3,482.1 & 0.0 & 59.5 & 239.4 & 373.5 & 77,635.1 & 3,513.2 & 0.0 & 59.5 & 35.7 & 147.2 & 304.0 & 73,510.2 \\
    & 8 & 11 & 31 & 3 & 776.0 & 2,325.7 & 0.0 & 0.0 & 796.3 & 690.1 & 40,015.9 & 2,311.7 & 0.0 & 0.0 & 0.0 & 593.7 & 601.0 & 32,141.2 \\
    & 9 & 10 & 35 & 3 & 937.5 & 4,110.9 & 0.0 & 0.0 & 447.7 & 712.0 & 31,782.1 & 4,139.9 & 0.0 & 0.0 & 0.0 & 243.3 & 550.0 & 22,439.9 \\
     \cmidrule(l){1-12} \cmidrule(l){13-19}
     \multicolumn{5}{r}{Average} & \multicolumn{1}{c}{-} & 1,933.6 & 0.0 & 7.4 & 286.8 & 255.6 & 23,090.8& 1942.7&0.0&7.4&4.5&161.9&198.5&18,207.8\\
     \cmidrule(l){1-12} \cmidrule(l){13-19}
     \multirow[c]{14}{*}{Medium} & 40 & 43 & 76 & 6 & 1,160.5 & 6,085.6 & 0.0 & 44.8 & 3,017.6 & 1,140.3 & 164,186.8 & 6,451.5 & 0.0 & 32.7 & 8.5 & 2,714.0 & 899.1 & 138,530.5 \\
 & 42 & 47 & 104 & 8 & 1,565.0 & 8,024.3 & 12.0 & 55.5 & 3,563.0 & 160.2 & 1,373,593.7 & 8,106.6 & 12.0 & 54.7 & 16.8 & 3,358.8 & 138.7 & 1,366,363.9 \\
 & 63 & 70 & 94 & 7 & 1,746.5 & 13,871.4 & 0.0 & 228.2 & 1,250.1 & 1,398.4 & 307,504.3 & 13,902.5 & 0.0 & 195.3 & 58.9 & 1,465.1 & 1,206.7 & 277,317.6 \\
 & 67 & 78 & 116 & 8 & 1,839.5 & 13,225.9 & 0.0 & 142.1 & 1,931.4 & 751.0 & 228,250.5 & 13,460.9 & 0.0 & 139.3 & 33.8 & 935.5 & 962.9 & 200,057.2 \\
 & 71 & 82 & 116 & 8 & 2,060.0 & 11,158.1 & 9.0 & 263.3 & 2,797.8 & 982.1 & 1,277,996.2 & 10,748.5 & 9.0 & 243.3 & 45.9 & 2,218.9 & 594.8 & 1,232,488.4 \\
 & 57 & 70 & 117 & 6 & 2,327.5 & 14,551.7 & 0.0 & 322.2 & 1,654.3 & 1,566.4 & 417,729.4 & 14,732.0 & 0.0 & 252.5 & 49.5 & 1,438.4 & 1,551.1 & 341,397.0 \\
 & 79 & 83 & 137 & 7 & 2,425.0 & 17,886.5 & 0.0 & 663.2 & 2,132.2 & 2,031.5 & 785,718.4 & 17,555.1 & 0.0 & 559.6 & 58.6 & 1,707.6 & 2,159.2 & 671,538.1 \\
 & 66 & 80 & 127 & 8 & 2,512.5 & 15,108.4 & 0.0 & 118.4 & 1,600.3 & 1,596.7 & 213,496.4 & 14,567.4 & 0.0 & 114.1 & 34.3 & 1,349.9 & 1,143.2 & 191,985.5 \\
 & 68 & 74 & 128 & 7 & 2,595.0 & 18,744.1 & 1.6 & 547.5 & 1,238.4 & 2,827.7 & 822,423.5 & 18,276.6 & 0.0 & 355.6 & 55.8 & 1,206.4 & 3,061.4 & 471,318.0 \\
 & 85 & 97 & 136 & 8 & 2,673.0 & 17,252.5 & 0.0 & 427.7 & 2,085.1 & 2,355.1 & 554,640.8 & 18,164.5 & 0.0 & 361.0 & 53.3 & 1,447.8 & 2,081.4 & 464,253.2 \\
 & 78 & 85 & 128 & 8 & 2,685.5 & 15,931.9 & 63.0 & 461.4 & 1,243.9 & 2,783.1 & 6,870,259.9 & 16,164.5 & 63.0 & 391.1 & 57.8 & 1,553.3 & 2,371.0 & 6,801,308.3 \\
 & 77 & 85 & 131 & 8 & 2,893.5 & 17,168.9 & 0.0 & 410.7 & 1,360.1 & 1,756.7 & 503,850.9 & 17,620.2 & 0.0 & 388.1 & 57.5 & 963.2 & 2,149.1 & 477,607.4 \\
 & 89 & 97 & 137 & 7 & 2,939.5 & 21,350.0 & 13.4 & 865.0 & 1,074.2 & 3,247.2 & 2,327,223.6 & 21,551.0 & 0.0 & 880.5 & 47.7 & 1,318.9 & 3,375.7 & 1,009,098.9 \\
 & 84 & 98 & 133 & 7 & 2,971.0 & 20,620.2 & 39.2 & 1,091.8 & 736.8 & 2,666.4 & 5,106,602.3 & 20,578.3 & 31.5 & 887.0 & 59.1 & 839.1 & 2,910.7 & 4,141,019.3 \\
     \cmidrule(l){1-12} \cmidrule(l){13-19}
     \multicolumn{5}{r}{Average} & \multicolumn{1}{c}{-} & 15,070.0 & 9.9 & 403.0 & 1,834.7 & 1,804.5 & 1,496,676.9& 15,134.3&8.2&346.8&45.5&1,608.4&1,757.5&1,270,305.9\\
    \cmidrule(l){1-12} \cmidrule(l){13-19}
    \multirow[c]{14}{*}{Large} & 98 & 109 & 132 & 8 & 3,078.5 & 22,498.9 & 0.0 & 456.2 & 457.4 & 3,961.9 & 571,626.0 & 22,359.3 & 0.0 & 394.4 & 56.2 & 463.1 & 3,905.6 & 508,725.2 \\
 & 101 & 119 & 129 & 8 & 3,229.0 & 17,814.1 & 9.6 & 647.3 & 1,342.6 & 2,674.1 & 1,721,369.9 & 18,811.0 & 0.0 & 608.8 & 57.5 & 969.3 & 2,715.1 & 710,970.8 \\
 & 114 & 136 & 141 & 8 & 3,350.5 & 20,596.8 & 44.7 & 1,346.5 & 1,048.3 & 2,672.2 & 5,920,713.6 & 20,811.8 & 33.5 & 1,453.0 & 56.6 & 616.4 & 1,788.8 & 4,878,101.5 \\
 &114  & 132 & 140 & 8 & 3,401.5 & 22,016.2 & 27.7 & 1,320.5 & 810.1 & 3,021.1 & 4,196,040.1 & 22,693.0 & 14.5 & 1,487.4 & 58.6 & 525.9 & 2,616.0 & 3,028,191.3 \\
 & 101 & 123 & 143 & 8 & 3,437.5 & 23,804.3 & 0.0 & 649.5 & 942.7 & 4,558.3 & 792,752.6 & 22,514.1 & 0.0 & 447.0 & 59.1 & 1,561.5 & 3,605.0 & 588,441.4 \\
 & 112 & 129 & 137 & 8 & 3,471.0 & 21,873.0 & 0.0 & 520.3 & 613.1 & 3,935.4 & 639,262.4 & 21,352.6 & 0.0 & 473.6 & 52.4 & 958.4 & 3,604.6 & 595,823.5 \\
 & 114 & 129 & 142 & 8 & 3,499.5 & 21,173.3 & 28.1 & 1,119.6 & 994.0 & 4,116.9 & 4,065,444.5 & 21,178.8 & 8.0 & 1,042.9 & 58.1 & 757.2 & 4,141.9 & 1,969,653.9 \\
 & 98 & 122 & 149 & 8 & 3,513.0 & 19,739.2 & 35.7 & 1,141.4 & 1,842.7 & 2,570.9 & 4,836,612.0 & 19,413.8 & 18.5 & 1,233.5 & 59.2 & 2,192.8 & 2,399.4 & 3,216,666.5 \\
 &98  & 108 & 139 & 8 & 3,541.0 & 20,862.3 & 0.0 & 941.6 & 628.4 & 3,526.0 & 1,051,844.1 & 20,241.0 & 0.0 & 847.2 & 59.1 & 1,106.1 & 2,790.2 & 956,432.6 \\
 & 108 & 122 & 144 & 8 & 3,670.5 & 22,605.8 & 2.1 & 1,127.8 & 877.3 & 2,790.1 & 1,445,018.0 & 22,369.7 & 0.0 & 873.3 & 56.0 & 1,126.8 & 2,922.1 & 987,898.0 \\
 & 92 & 107 & 142 & 8 & 3,684.5 & 22,877.9 & 0.0 & 584.4 & 638.0 & 4,754.3 & 721,474.4 & 23,511.5 & 0.0 & 458.5 & 53.4 & 593.9 & 4,715.3 & 594,093.8 \\
 & 114 & 136 & 138 & 8 & 3,739.5 & 23,361.3 & 8.5 & 1,178.1 & 853.7 & 5,271.4 & 2,182,537.5 & 23,633.6 & 0.0 & 961.3 & 56.1 & 766.6 & 5,434.9 & 1,116,585.2 \\
 & 112 & 131 & 148 & 8 & 3,953.5 & 24,313.5 & 34.9 & 1,251.9 & 524.6 & 2,961.1 & 4,838,688.0 & 25,301.2 & 13.5 & 1,694.5 & 59.4 & 766.0 & 2,944.6 & 3,151,661.8 \\
     \cmidrule(l){1-12} \cmidrule(l){13-19}
     \multicolumn{5}{r}{Average} &\multicolumn{1}{c}{-} & 21,810.5 & 14.7 & 945.0 & 890.2 & 3,601.0 & 2,537,183.3& 21,860.9&6.8&921.2&57.1&954.2&3,352.6&1,715,634.3 \\
    \bottomrule
    \end{tabular}
    }
    \begin{minipage}{12cm}
    \vspace{0.1cm}
    {\footnotesize $TC$: travel cost, $UQ$: undelivered quantity, $FDD$: first delivery delay, $mFDD$: maximum of first delivery delays, $DUC$: driver underutilization cost, $DOC$: driver overtime cost.}        
    \end{minipage}
\end{table}

Table \ref{tab_cdp_runtime} shows that a higher percentage of deliveries are completed as the algorithm runtime increases. For small instances, all requests are scheduled in less than a minute of runtime, while for medium and large instances, 99\% of requests are scheduled to be delivered with only a minute of runtime.

\begin{table}[htb]
    \centering
    \caption{RMC delivery completion within different runtimes of the GRASP}
    \label{tab_cdp_runtime}
    \scriptsize
    \resizebox*{0.7\textwidth}{!}{
    \begin{tabular}{llrrrrr}
    \toprule
     &  & \multicolumn{5}{c}{Runtime (min)} \\
     \cmidrule{3-7}
     &  & 1 & 5 & 10 & 30 & 60 \\
    \midrule
    Small & \%Load & 100.0 & 100.0 & 100.0 & 100.0 & 100.0 \\
Medium & \%Load & 98.9 & 99.3 & 99.4 & 99.5 & 99.5 \\
Large & \%Load & 98.8 & 99.3 & 99.4 & 99.6 & 99.6 \\
    \bottomrule
    \end{tabular}
    }
\end{table}

The medium instance $C\_42\_47\_104\_8$ has an unserved customer with a demand of 12 $m^3$ and a due time of $07\text{:}00$. Upon investigation, we find that even though we have $104$ drivers in this instance, the number of drivers who start their shift before $07\text{:}00$ is insufficient to serve this customer within the 60-minute window we defined. Thus, the value of $UQ$ that we report in Table~\ref{tab:cdp_all_instances} appears optimal given the driver shift schedule given as input to our algorithm. We confirmed this observation by removing the driver start shift and overtime constraints and resolving our instances five more times. All orders are now fully serviced on each run, as shown in Table~\ref{tab:cdp_all_instances_no_shift}. Table~\ref{tab:cdp_lift_shift_constraints} compares the averages of the results from Tables \ref{tab:cdp_all_instances} and \ref{tab:cdp_all_instances_no_shift}. We found that the maximum delay dropped to zero for small instances, and for medium (large) instances, it was reduced by almost 60\% (40\%) on average. This suggests that when we input a good work schedule, our algorithm works effectively to ensure that all orders are filled and that delays for first deliveries and travel costs are minimized. We can see that the average travel time decreases in all cases except the large one, where the difference is not significant.

\begin{table}[ht!]
    \centering
    \caption{Results with all instances without shift and overtime constraints}
    \label{tab:cdp_all_instances_no_shift}
    \resizebox{\textwidth}{!}{
    \begin{tabular}{lcccccrrrrrrrrr}
    \toprule
    &  &  &  & & & \multicolumn{4}{c}{Avg} & \multicolumn{5}{c}{Best}  \\
     \cmidrule(l){7-10} \cmidrule(l){11-15}
     &  &  &  &  & Demand & \multicolumn{1}{c}{$TC$}& \multicolumn{1}{c}{$UQ$} & \multicolumn{1}{c}{$FDD$} & \multicolumn{1}{c}{\multirow[c]{2}{*}{Z}} &  \multicolumn{1}{c}{$TC$}& \multicolumn{1}{c}{$UQ$} & \multicolumn{1}{c}{$FDD$}& \multicolumn{1}{c}{$mFDD$} & \multicolumn{1}{c}{\multirow[c]{2}{*}{Z}} \\
     & |$C$| & $|O|$ & $|K|$ & $|B|$ & (m$^3$) & \multicolumn{1}{c}{(min)} & \multicolumn{1}{c}{(m$^3$)} &\multicolumn{1}{c}{(min)} &  & \multicolumn{1}{c}{(min)} & \multicolumn{1}{c}{(m$^3$)} & \multicolumn{1}{c}{(min)} & \multicolumn{1}{c}{(min)}&  \\
    \midrule
    \multirow[c]{8}{*}{Small} & 11 & 12 & 13 & 1 & 226.5 & 1,243.5 & 0.0 & 0.0 & 1,243.5 & 1,243.5 & 0.0 & 0.0 & 0.0 & 1,243.5 \\
    & 5 & 8 & 13 & 1 & 267.0 & 745.1 & 0.0 & 0.0 & 745.1 & 745.1 & 0.0 & 0.0 & 0.0 & 745.1 \\
    & 6 & 11 & 18 & 2 & 333.5 & 753.3 & 0.0 & 0.0 & 753.3 & 753.3 & 0.0 & 0.0 & 0.0 & 753.3 \\
    & 4 & 7 & 15 & 2 & 375.0 & 1,017.6 & 0.0 & 0.0 & 1,017.6 & 1,016.5 & 0.0 & 0.0 & 0.0 & 1,016.5 \\
    & 7 & 8 & 19 & 2 & 388.0 & 1,662.7 & 0.0 & 0.0 & 1,662.7 & 1,651.8 & 0.0 & 0.0 & 0.0 & 1,651.8 \\
    & 10 & 14 & 29 & 3 & 613.5 & 3,259.0 & 0.0 & 0.0 & 3,259.0 & 3,200.9 & 0.0 & 0.0 & 0.0 & 3,200.9 \\
    & 8 & 11 & 31 & 3 & 776.0 & 2,222.0 & 0.0 & 0.0 & 2,222.0 & 2,208.9 & 0.0 & 0.0 & 0.0 & 2,208.9 \\
    & 9 & 10 & 35 & 3 & 937.5 & 3,839.6 & 0.0 & 0.0 & 3,839.6 & 3,750.5 & 0.0 & 0.0 & 0.0 & 3,750.5 \\
     \cmidrule(l){1-10} \cmidrule(l){11-15}
     \multicolumn{5}{r}{Average} & - & 1,822.7 & 0.0 & 0.0 & 1,822.7  & 1,821.3 & 0.0 & 0.0 & 0.0 & 1,821.3 \\
     \cmidrule(l){1-10} \cmidrule(l){11-15}
     \multirow[c]{14}{*}{Medium} & 40 & 43 & 76 & 6 & 1,160.5 & 5,242.1 & 0.0 & 0.0 & 5,242.1 & 5,213.5 & 0.0 & 0.0 & 0.0 & 5,213.5 \\
     & 42 & 47 & 104 & 8 & 1,565.0 & 8,291.5 & 0.0 & 0.0 & 8,291.5 & 7,993.7 & 0.0 & 0.0 & 0.0 & 7,993.7 \\
     & 63 & 70 & 94 & 7 & 1,746.5 & 12,744.2 & 0.0 & 0.0 & 12,744.2 & 12,638.2 & 0.0 & 0.0 & 0.0 & 12,638.2 \\
     & 67 & 78 & 116 & 8 & 1,839.5 & 11,918.0 & 0.0 & 0.0 & 11,919.8 & 11,793.3 & 0.0 & 0.0 & 0.0 & 11,793.3 \\
     & 71 & 82 & 116 & 8 & 2,060.0 & 10,612.5 & 0.0 & 0.0 & 10,612.5 & 10,449.9 & 0.0 & 0.0 & 0.0 & 10,449.9 \\
     & 57 & 70 & 117 & 6 & 2,327.5 & 14,781.6 & 0.0 & 15.7 & 30,457.6 & 14,391.0 & 0.0 & 2.4 & 1.4 & 16,820.6 \\
     & 79 & 83 & 137 & 7 & 2,425.0 & 17,637.5 & 0.0 & 106.5 & 124,143.7 & 17,504.0 & 0.0 & 66.7 & 21.7 & 84,156.1 \\
     & 66 & 80 & 127 & 8 & 2,512.5 & 13,676.8 & 0.0 & 0.0 & 13,676.8 & 13,595.0 & 0.0 & 0.0 & 0.0 & 13,595.0 \\
     & 68 & 74 & 128 & 7 & 2,595.0 & 18,565.9 & 0.0 & 97.4 & 115,988.9 & 18,484.6 & 0.0 & 72.3 & 29.0 & 90,740.3 \\
     & 85 & 97 & 136 & 8 & 2,673.0 & 15,741.9 & 0.0 & 75.7 & 91,395.7 & 15,779.3 & 0.0 & 48.2 & 18.6 & 64,002.6 \\
     & 78 & 85 & 128 & 8 & 2,685.5 & 17,097.4 & 0.0 & 68.0 & 85,088.9 & 16,326.4 & 0.0 & 55.6 & 14.0 & 71,906.4 \\
     & 77 & 85 & 131 & 8 & 2,893.5 & 17,807.8 & 0.0 & 10.7 & 28,544.5 & 17,994.5 & 0.0 & 9.4 & 6.1 & 27,404.5 \\
     & 89 & 97 & 137 & 7 & 2,939.5 & 20,810.3 & 0.0 & 203.9 & 224,680.9 & 20,829.3 & 0.0 & 150.0 & 42.0 & 170,861.3 \\
     & 84 & 98 & 133 & 7 & 2,971.0 & 21,142.9 & 0.0 & 380.8 & 401,955.9 & 21,206.1 & 0.0 & 252.7 & 55.6 & 273,892.1 \\
     \cmidrule(l){1-10} \cmidrule(l){11-15}
     \multicolumn{5}{r}{Average}  & - & 14,719.3 & 0.0 & 68.5 & 83,195.9 & 14,585.6 & 0.0 & 46.9 & 13.5 & 61,533.4 \\
     \cmidrule(l){1-10} \cmidrule(l){11-15}
     \multirow[c]{14}{*}{Large} & 98 & 109 & 132 & 8 & 3,078.5 & 22,186.9 & 0.0 & 15.2 & 37,379.4 & 22,061.8 & 0.0 & 9.9 & 6.3 & 31,964.6 \\
 & 101 & 119 & 129 & 8 & 3,229.0 & 18,290.6 & 0.0 & 40.5 & 58,817.5 & 17,111.2 & 0.0 & 10.5 & 5.7 & 27,595.2 \\
 & \multirow[c]{2}{*}{114} & 136 & 141 & 8 & 3,350.5 & 21,244.6 & 0.0 & 122.4 & 143,618.2 & 21,702.8 & 0.0 & 90.0 & 58.9 & 111,699.9 \\
 &  & 132 & 140 & 8 & 3,401.5 & 22,888.6 & 0.0 & 119.9 & 142,811.7 & 23,079.3 & 0.0 & 78.7 & 16.9 & 101,807.4 \\
 & 101 & 123 & 143 & 8 & 3,437.5 & 24,070.1 & 0.0 & 44.1 & 68,139.4 & 23,863.3 & 0.0 & 13.7 & 11.9 & 37,577.6 \\
 & 112 & 129 & 137 & 8 & 3,471.0 & 22,497.5 & 0.0 & 120.4 & 142,915.9 & 21,867.0 & 0.0 & 104.6 & 16.4 & 126,500.0 \\
 & 114 & 129 & 142 & 8 & 3,499.5 & 21,545.5 & 0.0 & 461.2 & 482,716.9 & 20,476.5 & 0.0 & 236.3 & 53.1 & 256,802.5 \\
 & \multirow[c]{2}{*}{98} & 122 & 149 & 8 & 3,513.0 & 20,370.9 & 0.0 & 150.5 & 170,865.5 & 19,725.6 & 0.0 & 106.5 & 46.3 & 126,172.6 \\
 &  & 108 & 139 & 8 & 3,541.0 & 21,538.2 & 0.0 & 339.1 & 360,638.8 & 21,205.7 & 0.0 & 201.3 & 59.5 & 222,493.7 \\
 & 108 & 122 & 144 & 8 & 3,670.5 & 23,462.8 & 0.0 & 253.3 & 276,751.0 & 23,827.4 & 0.0 & 144.9 & 31.9 & 168,731.4 \\
 & 92 & 107 & 142 & 8 & 3,684.5 & 22,799.8 & 0.0 & 52.4 & 75,149.3 & 24,071.4 & 0.0 & 30.9 & 10.2 & 54,975.8 \\
 & 114 & 136 & 138 & 8 & 3,739.5 & 23,491.1 & 0.0 & 86.7 & 110,190.5 & 23,242.2 & 0.0 & 46.7 & 32.5 & 69,982.2 \\
 & 112 & 131 & 148 & 8 & 3,953.5 & 25,300.6 & 0.0 & 387.0 & 412,296.4 & 26,296.6 & 0.0 & 247.8 & 32.5 & 274,119.6 \\
     \cmidrule(l){1-10} \cmidrule(l){11-15}
     \multicolumn{5}{r}{Average} & - &  22,454.1 & 0.0 & 187.5 & 209,966.8  & 22,504.4 & 0.0 & 123.6 & 26.8 & 146,138.2  \\
    \bottomrule
    \end{tabular}
    }
    \begin{minipage}{12cm}
        \vspace{0.1cm}
        {\footnotesize $TC$: travel cost, $UQ$: undelivered quantity, $FDD$: first delivery delay, $mFDD$: maximum of first delivery delays.}        
        \end{minipage}
\end{table}

\begin{table}[!ht]
    \centering
    \small
    \caption{Comparison of the average results with and without shift and overtime constraints}
    \label{tab:cdp_lift_shift_constraints}
    \resizebox{0.8\textwidth}{!}{
    \begin{tabular}{lrrrrrrrr}
    \toprule
     & \multicolumn{4}{c}{Default parameters}&\multicolumn{4}{c}{$\mu=0$, $\eta,\varLambda=\infty$, $h_k=0$} \\
     \cmidrule(l){2-5} \cmidrule(l){6-9}
    &   \multicolumn{1}{c}{$TC$}& \multicolumn{1}{c}{$UQ$} & \multicolumn{1}{c}{$FDD$} & \multicolumn{1}{c}{$mFDD$}  & \multicolumn{1}{c}{($TC$)} & \multicolumn{1}{c}{$UQ$} & \multicolumn{1}{c}{$FDD$} & \multicolumn{1}{c}{$mFDD$}  \\
    & \multicolumn{1}{c}{(min)} & \multicolumn{1}{c}{(m$^3$)} & \multicolumn{1}{c}{(min)} & \multicolumn{1}{c}{(min)} & \multicolumn{1}{c}{(min)} & \multicolumn{1}{c}{(m$^3$)} & (min) & \multicolumn{1}{c}{(min)} \\
    \midrule
    Small & 1,933.6 & 0.0 & 7.4 & 4.5 & 1,822.7 & 0.0 & 0.0 & 0.0 \\
Medium & 15,070.0 & 9.9 & 403.0 & 46.9 & 14,719.3 & 0.0 & 68.5 & 18.4 \\
Large & 21,810.5 & 14.7 & 945.0 & 56.4 & 22,283.6 & 0.0 & 168.7 & 31.0 \\
    \bottomrule
    \end{tabular}
    }
    \begin{minipage}{12cm}
        \vspace{0.1cm}
        {\footnotesize $TC$: travel cost, $UQ$: undelivered quantity, $FDD$: first delivery delay, $mFDD$: maximum of first delivery delays.}        
        \end{minipage}
\end{table}

\subsection{Sensitivity analyses}

In this section, we analyze the effect on the solution when we change some parameters of our problem and how some objective function components influence each other.

\subsubsection{Influence of the maximum time delay $\gamma^1$}

We ran additional tests with different values of $\gamma^1$, varying between 5 and 25 minutes, with no shifts or overtime constraints on the drivers. The influence of the value of the maximum time delay between two consecutive deliveries of the same order is shown in Figure~\ref{fig:gamma1_influence}. For small instances, all customers are served on time regardless of the value of $\gamma^1$. For medium (large) instances, all customers are served with $\gamma^1$ starting at 15 (20) minutes. The delay of initial delivery decreases as $\gamma^1$ increases. These results confirm the fact that when the maximum time delay $\gamma^1$ is high, plants have more flexibility to delay current deliveries to make room for new ones or to make customers wait until drivers are available. Increasing the maximum delay from 10 to 25 minutes reduces the first delivery delay by over 50\% for the medium and large cases. The average delivery time does not change significantly regardless of the value of $\gamma^1$.

\begin{figure}[!ht]
    \centering
    \includegraphics[width=1\textwidth]{gamma1.pdf}
    \small
    \caption{Influence of the maximum time delay ($\gamma^1$) between consecutive deliveries of the same order. }
    \label{fig:gamma1_influence}
\end{figure}

\subsubsection{Trade-off of the objective function components}

We ran five tests on our dataset and obtained two solutions per instance: one that minimizes the travel time and another that minimizes the first delivery delay. Table~\ref{tab:cdp_tradeoff} reports the trade-off between the travel time and the first delivery delay for these solutions. The difference is not noticeable for small instances. However, for medium and large instances, it is clear that as the first delivery delay decreases, the travel time increases, and conversely, when we minimize the travel time, the first delivery delay tends to be longer. This result is expected because drivers assigned to one plant often must travel to another one to ensure that the first delivery remains on schedule. This trade-off highlights the challenge decision-makers face in finding the right balance between minimizing operational costs and meeting customer satisfaction with high service levels.

\begin{table}[!ht]
    \centering
    \caption{Trade-off between travel cost and first delivery delay}
    \label{tab:cdp_tradeoff}
    \begin{adjustbox}{max width=0.8\textwidth}
    \begin{tabular}{lcccccc}
    \toprule
     & \multicolumn{3}{c}{Minimize TC} & \multicolumn{3}{c}{Minimize $FDD$} \\
     \cmidrule(l){2-4} \cmidrule(l){5-7}
     & $TC$ & $FDD$ & $mFDD$ & $TC$ & $FDD$ & $mFDD$ \\
     & (min) & (min) & (min) & (min) & (min) & (min) \\
    \midrule
    Small & 1,821.3 & 0.0 & 0.0 & 1,831.6 & 0.0 & 0.0 \\
    Medium & 14,311.5 & 89.6 & 23.0 & 14,679.3 & 46.9 & 13.4 \\
    Large & 21,403.8 & 170.5 & 31.4 & 22,194.7 & 101.7 & 29.4 \\
    \bottomrule
    \end{tabular}
\end{adjustbox}

\end{table}

\section{Conclusion}
\label{sec:cdp_concl}

In this paper, we investigated a variant of the concrete delivery problem, where customers can order different types of concrete from different production centers, which must be delivered during the same time window. To solve this problem, we introduced a mathematical model and a heuristic solution based on Greedy Randomized Adaptive Search. Our analysis, using a dataset derived from our partner's historical data, demonstrated the effectiveness of our algorithm in providing good solutions within a reasonable time. Furthermore, our model incorporates constraints related to working hours, making our heuristic a valuable tool for designing weekly schedules for company drivers. It can also better reassign drivers to their home plants, as these assignments affect travel costs to ensure timely deliveries. Our GRASP algorithm performed exceptionally well on the \cite{kinable2014concrete} datasets, indicating its reliability. For future work, it would be interesting to address datasets with different unloading times at construction sites or to explore different scenarios regarding the number of simultaneous deliveries and the maximum time between two consecutive orders at a construction site. One can also study the problem using a stochastic or robust methodology to account for the uncertainty of the parameters in a real-world setting. As a contribution to the literature on CDP, we provide a mathematical formulation for our variant, contribute with a dataset, and demonstrate the application of GRASP to solve both our real problem and the one studied by \cite{kinable2014concrete}.

\vspace{1.5cm} \noindent \textbf{Acknowledgments}

Financial support for this work was provided by the Canadian Natural Sciences and Engineering Research Council (NSERC) under grants 2019-00094 and 2021-04037. This support is gratefully acknowledged. We thank the Digital Research Alliance of Canada for providing high-performance computing facilities.

\bibliographystyle{plainnat}
\bibliography{references}

\end{document}

### 2024-03-09 ArticleCDP/original/draft.tex
Save

\documentclass{article}

\usepackage{epsfig}
\usepackage{amsmath}
\usepackage{algorithm}
\usepackage[algo2e]{algorithm2e}
\usepackage[compatible,noend]{algpseudocode}
\usepackage{enumitem}
\usepackage{xspace}
\usepackage[table,xcdraw]{xcolor}
\usepackage{amsfonts}
\usepackage{pifont}
\usepackage{bbding}
\usepackage{tikz}
\usetikzlibrary{automata,topaths,calc,positioning,shapes,backgrounds}
\usepackage{caption}
\usepackage{tabularx}
\usepackage[T1]{fontenc}
\usepackage{natbib}
\usepackage{datatool}
\usepackage{array}
\usepackage{pgfplotstable}
\pgfplotsset{width=7cm,compat=newest}
\usepackage{booktabs}
\usepackage{longtable}
\usepackage{pdflscape}
\usepackage{afterpage}
\usepackage{capt-of}
\usepackage{adjustbox}
\usepackage{multirow}
\usepackage{float}
\usepackage[text={15cm,21cm},centering]{geometry}
\usepackage{setspace}
\usepackage{array}
\allowdisplaybreaks
\usepackage{lipsum} 
\newtheorem{theorem}{Theorem}[section]
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{proposition}[theorem]{Proposition}
\newtheorem{corollary}[theorem]{Corollary}
\newtheorem{definition}[theorem]{Definition}
\newcommand{\pcmaxc}{P$|\mbox{\em cont}|$C$_{\max}$\xspace}
\newcommand{\alns}{adaptive large neighborhood search\xspace}
\newcommand{\Alns}{Adaptive large neighborhood search\xspace}
\renewcommand{\algorithmicforall}{\textbf{for each}}
\SetKwFor{ForEach}{for each}{}{}%
\renewcommand{\algorithmiccomment}[2][1\linewidth]{%
\leavevmode\hfill\makebox[#1][l]{/* \textbf{~#2} */}}
\renewcommand{\algorithmicrequire}{\textbf{Input:}}
\newcommand{\F}{$\mathcal{F}$\xspace}
\newcommand{\B}{$\mathcal{B}$\xspace}
\newcommand{\C}{$\mathcal{C}$\xspace}
\renewcommand{\tabcolsep}{4pt}

\oddsidemargin=0.5in
\topmargin=-0.25in
\textwidth=5.5in
\textheight=8.25in

%\journal{European Journal of Operational Research}

%\begin{frontmatter}

\title{A GRASP algorithm for the concrete delivery problem. }
\author{Ousmane Ali$^{(1)}$, Jean-Fran\c cois C\^ot\'e$^{(2)}$, Leandro C.~Coelho$^{(3)}$\\
 $(1)$ {\tt nassoma-wattara-ousmane.ali.1@ulaval.ca}\\
 $(2)$ {\tt Jean-Francois.Cote@fsa.ulaval.ca}\\
 $(3)$ {\tt Leandro.Coelho@fsa.ulaval.ca}\\
}

% \date{\today}

\begin{document}
\maketitle
\begin{abstract}
	%We study the daily scheduling of concrete delivery. Truck drivers are dispatched based on assignment priority and remaining working time over the horizon planning. A customer may be served from several available production plants. Due to the heterogeneity of the fleet, the number of visits to a customer location is not known ahead of time. The objective is to schedule drivers while minimizing travel times, waiting duration, idle time, and overtime. We solve this problem using a greedy randomized adaptive search procedure. Computational experiments with real data show the performance of our methods over the approach used by a local company. 
\end{abstract}

% \begin{keyword}
\noindent{\textbf{Keywords:} Vehicle routing; scheduling; ready-mixed concrete; concrete delivery
% \end{keyword}

%\end{frontmatter}

\section{Introduction}
\label{sec:intro}

Construction projects involve a huge movement of equipment, people, and materials. Among the latter, concrete is one of the most used. It is a perishable product with many factors affecting its quality \citep{sinha_quality_2021}. It comes in two types: ready-mixed concrete (RMC) and site-mixed concrete (SMC). As its name implies, SMC is produced on the spot using raw materials (water, aggregates, and cement) stored on the construction site, while RMC is manufactured in a batch plant and delivered to the construction site. SMC can avoid delays caused by road traffic, but has a slower and more difficult production process, requires storage for mixing materials and equipment, and is suitable for low amounts of concrete. It is quite the opposite for the RMC, which has better quality and benefits from lower production cost \citep{muresan_comparing_nodate}. Nevertheless, the batch plant manager must employ a fleet of high-cost revolving drum trucks (concrete mixers) to dispatch the ready-mixed concrete. Hence, the key to reaping the RMC benefits is to ensure efficient and prompt delivery on the construction site.

 Concrete delivery under the form of RMC is subject to many operational constraints that make it a challenging problem met in Operations Research and addressed by the Concrete Delivery Problem (CDP). In this paper, we study a variant of the Concrete Delivery Problem to schedule the daily production and dispatching of RMC for a company located in the province of Quebec, Canada. This company produces RMC in multiple batching plants with different production rates, using a fleet of concrete mixers of various capacities. Each plant has its specific fleet, but a truck can load elsewhere if deemed necessary, except it must return to its home plant at the end of the day. They own two types of trucks with different capacities but can call on an external fleet when needed. They serve a construction site from any batch plant, with the first delivery starting at the time specified by the customer.  Loading and unloading a concrete mixer depends on the truck capacity, a plant's loading, and a construction site's unloading rate, respectively. The problem is like the one addressed in \cite {schmid2009hybrid, schmid2010hybridization}, except for the plant's production rate and the lack of unloading instrumentation in our case. Also, as per union rules, they assign a truck driver to a delivery task based on his seniority. The company uses a centralized dispatcher system to schedule all daily orders, but this system has issues satisfying all daily demands without using an external fleet.

According to \cite{blazewicz2019handbook}, the concrete delivery problem combines vehicle routing with scheduling issues to plan routes to deliver concrete from batch plants (depots) to customers' construction sites. Ready-mixed concrete is an on-demand product with a short life cycle from production through end use. It cannot be stored and cannot stay too long in a truck, or it will harden. Hence, concrete mixers must deliver RMC at the planned construction site shortly after its production. RMC production and delivery is then an example of a Just-In-Time (JIT) production system in construction \citep{tommelein1999just}. A customer quantity requirement is often greater than the truck size and must be fulfilled by multiple deliveries. In that sense, CDP is like the Vehicle Routing Problem (VRP) with split delivery (VRPSD) \citep{archetti2008split}, except that the same truck may visit a customer more than once.

Since concrete hardens quickly, in case of multiple deliveries, back-to-back deliveries must be continuous or at least close in time to avoid the problem of cold joint that negatively affects the quality of the concrete. Customers request to be served within a specific time window, complicating truck loading schedules when a plant can only load one truck at a time. Likewise, often only one truck can unload at a time at a customer location, sometimes leading to concrete mixers queuing and waiting their turn to deliver. Furthermore, with trucks of varied sizes, loading, travel, and unloading times that may be uncertain, the Concrete Delivery Problem is a complex problem to solve and has been proven by \cite{asbach2009analysis} to be NP-hard. With a single batching plant, the objective is to maximize the service level (serve all customers and avoid delays between subsequent deliveries and truck queuing). With additional production centers to choose from, a minimization of the total travel time is added to the objective. %However, these objectives can be conflictive.

The remainder of the paper is organized as follows. Section \ref{lit_review} presents an overview of the CDP related literature.  Section \ref{desc_form} provides a formal description of the problem. Then in Section \ref{method}, we describe the Greedy Randomized Search Procedure (GRASP) algorithm and the constructive heuristics we design to solve this variant of the CDP. Computational experiments and our conclusions follow in Sections \ref{comp_exp} and \ref{concl}, respectively.

\section{Literature review}
\label{lit_review}

A text mining approach for reviewing the ready-mixed concrete literature \citep{maghrebi2015text} showed that concrete technology and material science are the main core of research in this area. The first academic research on concrete batching and delivery began in the late 1990s. \cite{tommelein1999just} describe RMC as a prototypical example of a Just-In-Time production system in construction and identify two practices occurring when delivering it. An alternative is for the customer to haul the product from the batch plant with their concrete mixer. The other approach is where the batch plant delivers the concrete to the customer's location. This latter approach is the one studied in all related papers found in the literature.

Several works on the CDP mainly focussed on Simulation related methods, either standalone \citep{zayed2001simulation, wang2001scheduling, tian_simulation_based_2010, panas_simulation_based_2013, galic2016simulation}, etc. or hybridized with optimization methods \citep{feng2004optimizing, lu2005optimized, feng_integrating_2006}, etc. to schedule and dispatch concrete production and delivery. \cite{wang2001scheduling} developed a simulation model to reveal the effect and value of the concrete mixers' inter-arrival time on the productivity of hired unloading equipment on site. With a combination of Genetic Algorithm (GA) and simulation process, \cite{feng2004optimizing} tried to minimize the total waiting time for trucks at a customer site. Loading trucks with identical capacities is carried out at the same batch plant. Loading and unloading durations are fixed. GA is used to find the most efficient and effective loading sequence of RMC trucks to be assigned to different construction sites. The simulation process determines the loading, arrival, departure and waiting time of trucks and thus evaluates the cost of each dispatching sequence. They evaluated the method using data from a batch plant in Taiwan with up to nine customers served. \cite{mayteekrieangkrai2015optimized} solved the same problem with the same data using a bee algorithm (BA) and found better solutions than the GA. \cite{lu2005optimized} used the same combination of GA and simulation to decide on the optimal number of concrete mixers to be deployed together with an optimal schedule for batching and delivering concrete. Their objective is to minimize the site crew idle times due to late concrete deliveries plus truck queuing time. The particularity of this setting is that mortar must be delivered on-site before concrete to lubricate an unloading pump. Therefore, mortar batching and delivery are also modelled in the simulation. Finding the best RMC truck size was also the purpose of the discrete-event simulation model proposed by \cite{panas_simulation_based_2013}.

Besides simulation-based methods, we find in the literature works using metaheuristics \citep{faria2006distributed, misir2011selection, maghrebi2016sequential, yang2022concrete}, exact methods \citep{yan2007optimal, kinable2014concrete}, matheuristics \citep{schmid2009hybrid, schmid2010hybridization}, Benders Decomposition \citep{maghrebi2014benders}, Column Generation (CG) \citep{maghrebi2014solving, maghrebi2016column}, Lagrangian relaxation \citep{narayanan2015using}, and machine learning approach \citep{graham2006modeling, maghrebi2014exploring, maghrebi2016matching}. \cite{matsatsinis2004towards} designed a decision support system (DSS) for the dynamic routing of both concrete and pumps that may be necessary for some sites to help the concrete unloading. Three plants are available, but vehicles fulfilling the same order must all load at the same plant. An order that cannot be executed may be postponed for the next day. The routing of the pumps is modelled as a multi-depot VRP with time windows. \cite{naso2007genetic} proposed a sequential GA method combined with constructive heuristics to solve a more general variant of the CDP. In this problem, besides batching concrete delivered to a customer site, the plant's production schedule must include orders picked up by the customers themselves. The algorithm first schedules the plant loading operations before scheduling truck job deliveries. A non-linear model minimizing transportation costs, waiting times, outsourced costs and overtime work is also developed. The authors ran experiments using real-world instances of a concrete supply chain in the Netherlands. They found a reduction in the number of requests redirected to external companies. \cite{yan2007optimal} also emphasized overtime considerations in their paper scheduling RMC for one batching plant with two loading docks. After 4 PM, operations at plant and construction sites are in overtime with a higher hourly wage. A mixed integer programming (MIP) model on a time-space network is developed to minimize travel times and operating costs at both normal and overtime working hours at the plant and the construction sites. Real data consisting of 3 days of operation is tested using a two-stage algorithm. First, they solved the MIP relaxation with CPLEX. Then they simplify the original model by fixing some decision variables before solving it. This algorithm seems to improve the actual plant operation by 10\%. A time-space network is the principal component of the real-time DSS developed by \cite{durbin2008or} to solve a dynamic CDP every five minutes. The DSS can receive new orders, schedule them on the fly, and deal with unexpected events such as plant closures, truck breakdowns, delays in transportation times, etc. Combined with a Tabu Search (TS) heuristic to warm start CPLEX, the model is performant enough to solve instances with up to 1,500 loads per day with up to 250 trucks. The real-time planning and monitoring of CDP are studied more in detail by \cite{garza2021dynamic} in his thesis. Another variant of the CDP is modelled by \cite{schmid2009hybrid} as an integer multicommodity network flow (MCNF) problem on a time-space network. In this paper, concrete is delivered using a heterogeneous fleet of vehicles, and each plant can load an illimited number of trucks simultaneously. Among the trucks, some have specialized equipment and must arrive first at certain construction sites to assist concrete-mixers unloading. The objective is to fulfill all orders, minimize the travel cost, and avoid delays between two consecutive unloading operations for an order. The model is typically solved using a matheuristic algorithm that combines the MCNF with a variable neighbourhood search (VNS) heuristic. The method can solve large problem instances with more than 60 orders per day quickly without encountering any memory issues. The same problem is addressed by \cite{schmid2010hybridization}. The authors proposed a MIP model combined with a VNS and a very large neighbourhood search (VLNS) to develop two matheuristics approaches. Comparisons between both matheuristics and a standalone VNS show that the former methods are much better and are suitable for solving larger problem instances. These methods also provide better solutions for small to medium instances than the matheuristic used in \cite{schmid2009hybrid}. A pure VNS approach with the same problem but without the use of instrumentation has been applied by \cite{payr2009optimizing}.

Regarding objectives, most authors have been interested in minimizing travel time and delays between consecutive deliveries. Some authors, however, were more interested in maximizing only customer satisfaction. We find these situations in the works of \cite{durbin2008or, kinable2014concrete, kinable2014logic, sulaman2017simulated}. \cite{kinable2014concrete} introduce a more general MIP and constraint programming (CP) models of the CDP reflecting the main constraints commonly found in all CDP works: time lag and no overlapping between consecutive deliveries, covering of all customers' demands, delivery time window, and heterogeneous fleet. However, the model did not include constraints limiting the time that concrete may reside in a truck. The authors also propose a constructive heuristic that schedules the visits to the customers one by one according to the start time of the visit and the truck capacity. The procedure is invoked multiple times for different permutations of the customer's order which is determined using a steepest descent (SD) local search procedure. One of the paper's main contributions is the creation of the first public test instances for the CDP with up to 50 customers, four batching plants, and 20 concrete mixers.  They found the CP model to be highly effective in finding high-quality solutions in relatively little time or improving existing schedules, while the MIP model can be used to compute bounds, as it seems ineffective in solving large problem instances. Finally, the heuristic often yields good solutions in less than a second. A detailed analysis of the MIP model presented in cite{kinable2014concrete}and of two more compact models can be found in the thesis of \cite{hernandez_lopez_study_2020}. In \cite{kinable2014logic}, we find an attempt to solve the previous problem with a Logic Based Benders' approach. \cite{sulaman2017simulated} expand upon the SD heuristic proposed in \cite{kinable2014concrete}, proposing a Simulated Annealing (SA) with a Time-Slot Heuristic (TH). TH mechanism is to look for a slot between existing visits of a truck to schedule a new delivery instead of assigning it to the time slot strictly after the truck's latest assigned delivery. The goal is to reduce the large time gaps that can be present in a schedule created with SD due to ignoring the intermediate available time slots. Experimental results indicated that SATH outperforms SD in speed and solution quality. A generalization of the MIP model of \cite{kinable2014concrete} is addressed in \cite{asbach2009analysis}. This model simultaneously minimizes the total sum of travel costs and the penalty costs for customers with unfulfilled demand. A customer can request that all concrete deliveries come from the same plant or a subset of plants and that a delivery truck belongs to a subset of the vehicle fleet. The MIP model is used in a local search scheme as a black-box solver to reoptimize an incumbent solution in which a neighbourhood operator has unfixed some variables.

%We provide in the Table an extensive summary of the available literature for the Concrete Delivery Problem.

\section{Problem description}
\label{desc_form}
The problem studied within this paper is about the distribution of ready-mixed concrete from a Canadian company operating in the greater Montreal area. Concrete order from a customer arrives at a central center and is assigned to one of the batching plants from where the product will be produced and delivered to the customer. We define an instance of our problem on a set of batching plants, a set of customer orders, and a set of drivers.

\subsection*{Batching plant}

The company serves its customers from 8 concrete batching plants geographically scattered. Each plant has only one loading bay and can load one truck at a time, leading to trucks lining up. The plants are heterogeneous, as each has its hourly loading rate. A plant's rate and truck capacity influence the time needed for loading. After loading, a driver adjusts the concrete in his truck before departing to the customer site. Any batch plant can serve a construction site if the travelled time between them is less than the concrete lifespan. Each plant has its assigned fleet, but it can get a vehicle from another plant or even hire an external fleet.

 For a batching plant, the decision is to determine which driver must load at a given time.

\subsection*{Customer orders}

A constructor requests one or several types of concrete to be delivered to his site. When placing an order, he specifies the quantity of concrete of each type, the arrival time of the first concrete mixer, and his unloading rate. If the amount of concrete usually exceeds any truck capacity, several deliveries are scheduled to satisfy the customer. The arrival time of the first delivery must be respected, and to prevent cold joint in concrete, subsequent deliveries must be in just in time, or at least close in time. We define a maximum time lag beyond which no next delivery should be allowed. The construction site unloading rate gives the time necessary to discharge a truckload. When a construction site requires several types of concrete, all quantities of each type must be thoroughly delivered before delivering the other types. Once a plant is chosen to produce a customer's first order, it must be the provider of all subsequent deliveries of this customer.

\subsection*{Drivers}

A driver is an important entity here, as we must incorporate some union constraints into our model. The company has two types of trucks with a capacity of 8 $m^3$ of 12 $m^3$. Each driver is assigned to one of these trucks. A driver is assigned to a plant from where he starts and ends his shift day. When a driver is scheduled for a day, he must work a minimum of four hours for a maximum of 8 hours in normal time plus two hours of overtime. A driver mainly loads RMC at his home plant but can still drive at other plants and load there if needed. Batching plant produces concrete on-demand with recipes specific to customers, which means that truck cannot hold orders for more than one client, even if it has spare capacity. Thus, a driver must refill at a plant between two consecutive deliveries. After unloading the RMC, a driver must clean the concrete mixer before travelling to his next loading plant. When assigning a driver to a delivery, priority is given to the employees with the highest seniority.

Let us consider a small example to illustrate our problem. 
\iffalse

% We now introduce a mathematical model to our variant of the CDP. The model we present here is inspired from the works of \citet{asbach2009analysis}. Let $I =\{i_1,...,i_n\}$ be the set of $n$ customers, $J = \{j_1,...,j_m\}$ be the set of $m$ concrete production facilities (depots), and $K=\{k_1,...,k_p\}$ the set of $p$ vehicles used to deliver concrete from the depots to the customers.

To each vehicle $k$ is assigned a start ($o_k$) and end ($f_k$) depot. The sets $O$ and $F$ define the $p$ starting depots $o_k$ and ending depots $f_k$, respectively.

A customer $i \in I$ requires $q_i$ amount of concrete with an incurred penalty of $\beta_i$ if we do not deliver all demands within the working horizon. %Service times are mainly determined by the customer while different vehicles have no significant effect.

A vehicle has to start delivery at customer $i$ within the time window $\left[a_i,b_i\right]$. A first delivery deadline may be required by the customer, and hence, the first delivery of the day must not start later than at time $b^{'}_i$. Let consider the subset of the depots $J_i \subset J$ from which customer $i$ may only be supplied. $K_i \in K$ is also the subset of vehicles than can serve customer $i$. The total demand $q_i$ may be scheduled into $n_i=\left\lceil {\frac{q_i}{K^i_{min}}} \right\rceil$  maximal number of deliveries chronologically ordered in non-decreasing times, with $K^i_{min}$ being the minimum size of a vehicle in $K_i$. Thus, we represent each customer $i \in I$ by the set $C_i = \{c^i_{1},...,c^i_{n_i}\}$ customers nodes. To prevent the concrete becoming solid at customer $i$ location before the arrival of subsequent deliveries, a parameter maximum time lag $maxtl_i$ is used to enforce that two consecutive deliveries are not too far in time. Furthermore, as only one vehicle can unload concrete at the same time, the parameter minimum time lag $mintl_i$ is used to ensure that two consecutive deliveries are not too close in time.

Similar to the customers, a depot $j \in J$ also has service time $s_j$ for operations such as parking the vehicles at the depot, loading the concrete, etc. A time window $[a_j,b_j]$ is also assigned to a depot to restrict the times of reloading. Each depot $j$ is represented by the set $ D_j = \{d^j_{1},...,D^j_{n_j}\}$ depots nodes, where $n_j=\left\lceil {\frac{b_j-a_j}{mintl_j}+1} \right\rceil$ is an upper bound for the maximum number of possible reloads of vehicles and $mintl_j$ is the minimum time lag between two consecutive reloading at depot $j$.

A vehicle $k$ has capacity $q^k$ which specify how many units of concrete a vehicle may deliver. Using a particular vehicle $k$ for delivering concrete during the working day accumulates vehicle usage costs $\alpha(k)$. Each vehicle $k$ is also given a time window $[a_k,b_k]$ in which it may supply customers, reload at depots, start its tour at $o_k$ and finally finish its tour at $f_k$. Finally, the parameter $\gamma$ represents the maximum time that concrete may reside in the vehicles before hardening.

We define our problem on a complete directed graph $G = (V, E)$ where $V=\{ I \cup J \cup O \cup F\}$ is the set of nodes. The arc sets are $E = \{(i,j,k) \hspace*{1mm} \vert \hspace*{1mm} i, j \in V \hspace{1mm} k\in K \}$ which corresponds to a movement of vehicle $k$ from node $i$ to node $j$. The outcoming arc sets of node $i$ using vehicle $k$ are $\delta_k^{+}(i) = \{(i, j,k) \hspace*{1mm} \vert \hspace*{1mm}  j \in V \}$ while the incoming arc sets of node $i$ using $k$ are $\delta_k^{-}(i) = \{(j, i,k) \hspace*{1mm} \vert \hspace*{1mm}  j \in V \}$.  For each arc $(i,j,k)$, a cost $c_{ijk}$ and a travel time $t_{ijk}$ are assumed.

% We define the directed, twofold weighted, multi-graph $G=\{V,E,t,z\}$ as follows. The edges of graph G have the form $(u, v, k)$, where $u$ and $v$ are nodes of $V$ and $k$ is a vehicle from the vehicle fleet $K$. An edge $(u, v, k) \in E$ corresponds to a movement of vehicle $k$ from node $u$ to node $v$. First, for every vehicle $k \in K$, there is an edge $(O_k, F_k, k) \in E$.Second, for every vehicle $k \in K$, there is an edge $(O_k, d, k) \in E$ to every depot node $d \in D$. These edges correspond to the first trip of a used vehicle to a depot to be filled up with concrete. An operating vehicle uses exactly one of these edges. Third, $E$ comprises for every vehicle $k \in K$, for every depot node $D_{i,n_i}$ and for every customer node $C_{j,n_j}$ one edge $(D_{i,n_i},C_{j,n_j},k)$, which relates to a loaded trip of vehicle $k$ from depot $D_i$ to customer $C_j$.  Fourth, $E$ contains for every vehicle $k \in K$, for every customer node $C_{i,n_i}$ and for every depot node $D_{j,n_j}$ one edge $(C_{i,n_i},D_{j,n_j},k)$, which corresponds to an unloaded trip of vehicle $k$ from customer $C_i$ to depot $D_j$. Finally, we have for every vehicle $k \in K$ and for every customer node $C_{i,n_i}$ one edge $e =(C_{i,n_i},F_k,k)$. Edge $e$ corresponds to the last movement of vehicle $k$ from customer $C_{i}$ to the ending location $F_k$.

We use the following decision variables. Let binary variable $x_{ijk}$ be equal to 1 if vehicle $k$ travels from $i$ to $j$, and 0 otherwise. For each node $i$, the continuous decision time variable $w_i$ determine which routes the vehicles take and at which times vehicles either supply, reload, begin or end their route at $i$. The binary decision variable $y_i$ for customer $i$ indicates whether his demand is satisfied in the current schedule. We define the boolean decision variables $\sigma_{ij}$ that determine which depot $j$ produces concrete for customer $i$. The objective function of the problem is to minimize the total travel costs of all routes while ensuring that all customers are fully satisfied. The mathematical model is as follows:

\begin{align}
	\label{obj:sch}  \min &  \sum_{(i,j,k) \in E} c_{ijk}x_{ijk} +  \sum_{i \in I}(1-y_i)\beta_i  &   \\ 
	\label{sch_1}   & \sum_{j \in \delta^{+}_{k}(o_k)}x_{o_kjk} = 1 &  k \in K   \\
	\label{sch_2}   & \sum_{i \in \delta^{-}_{k}(f_k)}x_{if_kk} = 1 &  k \in K  \\
	\label{sch_3}   & \sum_{i \in \delta^{-}_{k}(j)}x_{ijk} - \sum_{i \in \delta^{+}_{k}(j)}x_{jik} = 0 &  k \in K, j \in I \cup J   \\
	\label{sch_4}   &  \sum_{k \in K}\sum_{j \in \delta^{+}_{k}(i)}x_{ijk} \leq 1 &   i \in I    \\
	\label{sch_5}   &  \sum_{k \in K} \sum_{v \in \delta^{+}_{k}(c^{i}_{j+1})}\hspace{-0mm}x_{c^{i}_{j+1}vk} - \sum_{k \in K}\sum_{v \in \delta^{+}_{k}(c^{i}_{j})}\hspace{-0mm}x_{c^{i}_{j}vk}\leq 0 &  \\ \nonumber
      & i \in C, j \in \{1,..n_i-1\} &   \\
    \label{sch_6}   &  \sum_{u \in C_i}\sum_{k \in K}\sum_{j \in \delta^{+}_{k}(u)}q^k x_{ujk}  \geq q_iy_{i} &   i \in C   \\
	\label{sch_7}   &  \sum_{k \in K}\sum_{i \in \delta^{+}_{k}(j)}x_{jik} \leq 1 &   j \in J    \\
	\label{sch_8}   & s_i + t_{ijk} - M(1-x_{ijk})  \leq \omega_j - \omega_i  & \hspace{-20mm}   (i,j,k) \in E    \\
	\label{sch_9}   & \gamma + s_i +  M(1-x_{ijk})  \geq \omega_j - \omega_i  & \hspace{-20mm}  (i,j,k) \in E, i \in D, j \in I    \\
	% \label{sch_10}   & \omega_i \geq  a_i - M\left(1-\sum_{k \in K}\sum_{j \in \delta^+(i)}x_{ijk}\right)   &   i \in V    \\
	\label{sch_10}   & \omega_i \geq  a_i   &   i \in V    \\
	% \label{sch_11}   &  \omega_i \leq b_i + M\left(1-\sum_{k \in K}\sum_{j \in \delta^{-}(i)}x_{jik}\right)   &   i \in V    \\
	\label{sch_11}   &  \omega_i \leq b_i   &   i \in V    \\
	\label{sch_12}   &  \omega_{c^{i}_{1}} \leq b^{'}_{i} &   i \in C    \\
	\label{sch_13}   &  \omega_{c^{i}_{j+1}} - \omega_{c^{i}_{j}} \geq mintl(i) &  \hspace{-20mm}    i \in I, j \in \{1,...n_i-1\}    \\
	\label{sch_14}   &  \omega_{d^{i}_{j+1}} - \omega_{d^{i}_{j}} \geq mintl(i) &   \hspace{-20mm}  i \in J, j \in \{1,...n_i-1\}    \\
	\label{sch_15}   &  \omega_{c^{i}_{j+1}} - \omega_{c^{i}_{j}} \leq maxtl(i) &   \hspace{-20mm}  i \in I, j \in \{1,...n_i-1\}    \\
    \label{sch_16}   & \sum_{k \in K} \sum_{u \in D_j}\sum_{v \in C_i}x_{uvk}  \leq M\sigma_{ij} &   i \in I, j \in J   \\
    \label{sch_17}   &  \sum_{j \in J}\sigma_{ij} \leq 1 &   i \in I   \\
    \label{sch_18}   & x_{ijk} \in \{0,1\} &   (i,j,k) \in E   \\
    \label{sch_19}   &  \omega_i \geq 0 &   i \in V   \\
    \label{sch_20}   & y_{i} \in \{0,1\} &   i \in I   \\
    \label{sch_21}   &  \sigma_{ij} \in \{0,1\} &   i \in I, j \in J.   
\end{align}

\label{sec_formulation}

% The objective function (\ref{obj:sch}) minimizes simultaneously the total sum of travel costs of edges used by any vehicle and penalty costs $ \beta_c $ for customers $c$ whose demand is not fully satisfied. Constraints (\ref{sch_1})--(\ref{sch_2}) ensure that each vehicle $k$ leaves( comes back) to its starting (terminal) location exactly once.  Constraints (\ref{sch_3}) are flow conservation constraints for all nodes except the technical sources and sinks $O(k$ and $F_k$.  Constraints (\ref{sch_4}) state that each customer node may be used at most once. Constraints (\ref{sch_5}) ensure that a customer node $C_{i,j}$ is served before node $C_{i,j'}$ if $j < j'$.  Constraints (\ref{sch_6}) make sure that the $y_c$ variables indicate whether the demand of customer $c$ is satisfied. constraints  (\ref{sch_7}) ensure that no depot node is used more than once.
% Constraints (\ref{sch_10}) and (\ref{sch_11}) are time windows constraints.

\fi

\iffalse
We now introduce a formal definition for the problem. It is defined on a complete directed graph $G = (V, A)$ where $V = \{D \cup P \cup O \cup F\}$ is the set of nodes.  $D = \{1,\dots, n\}$ is the set of customer nodes and $D^I \subseteq D$ is the subset of all customers requiring having a demand greater than the lowest vehicle capacity.

The arc sets are $A = \{(i, j) \hspace*{1mm} \vert \hspace*{1mm} i, j \in V \}$ and $A^I = \{(i, j) \hspace*{1mm} \vert \hspace*{1mm} i, j \in D^I \cup P\}$. The outcoming arc sets of node $i$ are $\delta^{+}(i) = \{(i, j) \hspace*{1mm} \vert \hspace*{1mm}  j \in D \}$ and $\delta^{+}_I(i) = \{(i, j) \hspace*{1mm} \vert \hspace*{1mm}  j \in D^I \}$. The incoming arc sets of node $i$ are $\delta^{-}(i) = \{(j, i) \hspace*{1mm} \vert \hspace*{1mm}  j \in D \}$ and $\delta^{-}_I(i) = \{(j, i) \hspace*{1mm} \vert \hspace*{1mm} j \in D^I \}$. A time window $[a_{i},b_{i}]$ is associated to each node $i$, with $a_{i}$ and $b_{i}$ being respectively the earliest and latest time the service must begin at node $i$.

Let $K$ be the set of heterogeneous delivery vehicles partitioned into the subset. For each arc $(i,j)$, a cost $c_{ijk}$ and a time $t_{ijk}$ are known. They are respectively the transportation cost and the travelling time from node $i$ to node $j$ by vehicle $k$.  The capacity of each delivery vehicle $k$ is $Q^k$.

Let $P_i$ be the set of products ordered by the customer $i$. A customer can order between one to three product categories: furniture, home appliance and electronic products. Each delivery node $i$ has a total demand $q_{i}$ which is the sum of the specific demand $q^{p}_{i} $ of each product of type $p$ that the customer will receive. The service time $s_{i}^{D}$ for each delivery node $i$ is standard for all delivery vehicles. The service time of an installation node $i$ depends on the delivered products and the driver that performs it. Knowing that vehicle $k \in K^D \cup K^I$ performs the installation of product $p$, the service time will be $s_{ikp}^{I}$. All products of a customer must be installed by the same worker. Thus, the total installation time is $s_{ik}^{I} = \sum_{p \in P_i} s_{ikp}^{I}$ for vehicle $k$ and node $i \in D^I$.

We formulate the problem with the following decision variables. Let binary variable $x_{ijk}$ be equal to 1 if a deliveryman travels from $i$ to $j$ using vehicle $k$, and 0 otherwise.

$z_{ik}$ is equal to 1 if the delivery service at customer $i$ is done by driver $k$ and 0 otherwise.

$z_{ikp}$ equals 1 if the product $p$ of customer $i$ is installed by the driver $k$, 0 otherwise.

Binary variable $y_{ijk}$ equals 1 if an installer travels from $i$ to $j$ using vehicle $k$, and 0 otherwise. Binary variable $z_{i}$ equals 1 if the installation service at customer $i$ is done by a deliveryman, and 0 if it is done by an installer. The use of $z_i$ allows our formulation to be flexible; by setting $z_i=1$, we have a VRPTW model with flexible service time which we can relate to the VRPTWDST. With $z_i=0$, we can relate to a VRPTW model with temporal precedence and synchronization constraints similar to the VRPMS. Finally, let $S_{i}$ be the beginning time of delivery service at node $i \in D$ and $S^I_{i}$ be the beginning time of installation service at node $i \in D^I$. The objective function of the DIRP is to minimize the total transportation costs of all the routes traveled by the vehicle fleets. The mathematical model of the DIRP is as follows:
\fi

% In model (1)--(\ref{cend}), we assumed that only one worker installs at a customer location. But, as different product categories are being dealt with and a delivery crew may be versed in assembling a product, but less so in installing another product, we wanted to investigate the effect of allowing more workers to install the products of a customer. To the decision variables of model (1)--(\ref{cend}), we added the binary variable $z_{ik}$ and $z_{ikp}$. $z_{ik}$ is equal to 1 if the installation service at customer $i$ is done by driver $k$ and 0 otherwise. $z_{ikp}$ equals 1 if the product $p$ of customer $i$ is installed by the driver $k$, 0 otherwise. The mathematical formulation of this new model ($\mathcal{M}_3$) is derived from model (1)--(\ref{cend}) by replacing constraints (\ref{c7}) by constraints (\ref{c47})--(\ref{c51}), constraints (\ref{c9}) by (\ref{c53}) and constraints (\ref{c10}) by (\ref{c54}). These new constraints allow a deliveryman to share the installation of the products of a client with an installer. We solved this new formulation with CPLEX using our instances and show the results in . The column $\%Inst$ in the table denotes the average ratio of the number of products installed by the deliverymen over the total number of products to install.
\iffalse
\begin{align}
\label{model:obj} \min  &  \sum_{k \in K^{D}}\sum_{(i,j) \in A} c_{ijk}x_{ijk} + \sum_{k \in K^{I}}\sum_{(i,j) \in A^I} \hspace{-2mm}c_{ijk}y_{ijk}  &\\
\label{c1} &\hspace{-10mm} \sum_{j \in \delta^{+}(0)}x_{0jk}\ \leq 1 & \hspace{-40mm}  k \in K^{D} \hspace{0mm}\\
\label{c2} &\hspace{-10mm} \sum_{j \in \delta^{+}_I(0)}y_{0jk}\ \leq 1 & \hspace{-40mm}  k \in K^{I} \hspace{0mm}  \\
\label{c3} & \hspace{-10mm}  \sum_{k \in K^{D}}\sum_{j \in \delta^{+}(i)}x_{ijk}\ = 1 & \hspace{-40mm}i \in D \hspace{0mm}  \\
\label{c4} & \hspace{-10mm}  \sum_{j \in \delta^{-}(i)}x_{jik} - \sum_{j \in \delta^{+}(i)}x_{ijk} = 0 &\hspace{-40mm}i \in D ,\ k \in K^{D} \hspace{0mm} \\
\label{c5} & \hspace{-10mm}  \sum_{j \in \delta^{-}_I(i)}y_{jik} - \sum_{j \in \delta^{+}_I(i)}y_{ijk} = 0 & \hspace{-40mm}  i \in D^I ,\ k \in K^{I} \hspace{0mm}  \\
\label{c6} & \hspace{-10mm}  \sum_{i \in D}q_{i} \sum_{j \in \delta^{+}(i)}x_{ijk}\ \leq Q & \hspace{-40mm}k \in K^{D} \hspace{0mm}  \\
\label{c7} & \hspace{-10mm}  \sum_{k \in K^{I}}\sum_{j \in \delta^{+}_I(i)}y_{ijk}\ = 1 - z_{i} & \hspace{-40mm}  i \in D^I \hspace{0mm} \\
\label{c8} &\hspace{-10mm}  S_j \geq \hspace{-0mm} S_i +  s_i^D + \sum_{k \in K^{D}} t_{ijk} x_{ijk} - M\left(1-\sum_{k \in K^{D}} x_{ijk}\right) & \hspace{-15mm}  (i,j) \in A, i \not\in D^I \hspace{0mm} \\
\label{c9} & \hspace{-10mm}  S_j \hspace{-0mm} \geq \hspace{-0mm} S_i \hspace{-0mm}+ s^D_i \hspace{-0mm}+ \hspace{-0mm}\sum_{k \in K^{D}}\hspace{-1mm}(s_{ik}^I + t_{ijk} )x_{ijk} \hspace{-0mm}- \hspace{-0mm}M\hspace{-0mm}\left(\hspace{-0mm}2-\hspace{-3mm}\sum_{k \in K^{D}} \hspace{-2mm}x_{ijk}-z_{i}\hspace{-1mm}\right)  &  \hspace{-4mm}(i,j) \in A, i \in D^I \hspace{0mm}  \\
\label{c10} & \hspace{-10mm} S^I_j \geq \hspace{-0mm} S^I_i +  \sum_{k \in K^{I}}(s_{ik}^I + t_{ijk} )y_{ijk} - M\left(1-\sum_{k \in K^{I}} y_{ijk}\right)  & \hspace{-40mm}  (i,j) \in A^I \hspace{0mm} \\
\label{c11} & \hspace{-10mm} S^I_{i} \geq S_{i} + s^{D}_{i} & \hspace{-40mm}i \in D^I \hspace{0mm}  \\
\label{c12}  & \hspace{-10mm} a_i \leq S_i \leq b_{i} & \hspace{-40mm}  i \in D \hspace{0mm} \\
\label{c13}  &  \hspace{-10mm} a_i \leq S^I_i \leq b_{i} & \hspace{-40mm}  i \in D^I \hspace{0mm}  \\
\label{c14} & \hspace{-10mm} x_{ijk} \in \{0,1\} & \hspace{-50mm}  (i,j) \in A , k \in K^D  \hspace{0mm} \\
\label{c15} & \hspace{-10mm} y_{ijk} \in \{0,1\} & \hspace{-50mm} (i,j) \in A^I , k \in K^I  \hspace{0mm} \\
\label{cend} & \hspace{-10mm} z_{i} \in \{0,1\} & \hspace{-40mm}  i \in D^I. \hspace{0mm}
\end{align}
\fi

\iffalse
\begin{align}
    \label{model:obj2} \min  &  \sum_{k \in K^{D}}\sum_{(i,j) \in A} c_{ijk}x_{ijk} + \sum_{k \in K^{I}}\sum_{(i,j) \in A^I} \hspace{-2mm}c_{ijk}y_{ijk}  &\\
    \label{mod1:c1} & s.t. \hspace{2mm} (\ref{c1})\text{--}(\ref{c6}), \hspace{2mm} (\ref{c8}), \hspace{2mm} (\ref{c11})\text{--}(\ref{cend}) & \\
    \label{c47} & \hspace{2mm}  \sum_{j \in \delta^{+}_I(i)}y_{ijk} = z_{ik} & \hspace{-40mm}  i \in {D^I},k \in K^{I} \hspace{5mm} \\
    \label{c470}  & \hspace{2mm} \sum_{j \in \delta^{+}_I(i)}x_{ijk} \geq z_{ik} & \hspace{-40mm}  i \in D^I,k \in K^D \hspace{5mm} \\
    \label{c471} & \hspace{2mm} \sum_{k \in K^D} z_{ik} = z_i & \hspace{-40mm}  i \in {D^I} \hspace{5mm} \\
    \label{c48} & \hspace{2mm} \sum_{k \in K^I} z_{ik} \geq 1-z_i & \hspace{-40mm} i \in {D^I} \hspace{5mm} \\
    \label{c49} & \hspace{2mm} \sum_{k \in K^D}\sum_{p \in P_i} z_{ikp} \geq z_i & \hspace{-40mm} i \in {D^I} \hspace{5mm} \\
    \label{c50}& \hspace{2mm} \sum_{k \in K} z_{ikp} = 1 & \hspace{-40mm} i \in {D^I}, p \in P_i \hspace{5mm} \\
    \label{c51} & \hspace{2mm} z_{ikp} \leq z_{ik} & \hspace{-40mm} i \in {D^I}, p \in P_i, k \in K \hspace{5mm} \\
    \label{c53} & \hspace{2mm}  S_j \hspace{-0mm} \geq S_i + s^D_i + \hspace{-2mm} \sum_{k \in K^D} \hspace{-1mm} \left( \sum_{p \in P_i}s_{ipk}^Iz_{ikp} +\left( t_{ijk} + M \right) x_{ijk} \right) - M(2-z_i)   & \hspace{-3mm} (i,j) \in {A}, i \in D^I \hspace{5mm} \\
    \label{c54} & \hspace{2mm} S^I_j \geq \hspace{-0mm} S^I_{i} + \sum_{p \in P_i}s_{ipk}^I z_{ikp} + t_{ijk}y_{ijk} - M(1-y_{ijk})  & \hspace{-30mm}  (i,j) \in A^I, k \in K^{I} \hspace{5mm} \\
    \label{c61} & \hspace{2mm} z_{ik} \in \{0,1\} & \hspace{-50mm}  i \in D , k \in K  \hspace{5mm} \\
    \label{cend2} & \hspace{2mm} z_{ikp} \in \{0,1\} & \hspace{-50mm} i \in D^I , k \in K, p \in P_i.  \hspace{5mm}
\end{align}
\fi

\section{A GRASP heuristic}
\label{method}

\subsection{Priority insertion greedy heuristic}
This algorithm (PIGH) uses a minimum priority queue data structure to schedule the loading and unloading operations. We define a priority queue \F that contains an estimated delivery time for each delivery node. We start the algorithm outlined in Algorithm~\ref{alg:prioIns} by initializing \F with a big value $M$ for all nodes. Then we select for each customer $i$ a random order $o^p_i$. We update in the priority queue the value of the first delivery node $d^p_{0i}$ of each customer with $a_i$. Next, we repeat in a loop the instructions between lines 6 and 16 as long as \F is not empty. At each iteration, we first retrieve the node with the earliest delivery time and we schedule its loading and unloading tasks with the procedure described in Algorithm~\ref{alg_ScheduleTask}. Next, we update the solution and the remaining quantities that the order and customer must receive. If an order $o^p$ is not yet fulfilled after the visiting $d^p_{ij}$, we update in \F the next delivery node $d^p_{ij+1}$ with the expected start time plus a random value. Otherwise, if the customer has remaining orders, we select another random order $p'$ and update the value of its first delivery node in \F with also its expected start delivery time plus a random value. We add this random value to ensure the randomization of the algorithm.

{\setstretch{1}
{\small
    % \vspace{-4mm}
    \begin{algorithm}[hbt]         
        \caption{Delivery priority insertion heuristic }
        \label{alg:prioIns}
        \DontPrintSemicolon
        \LinesNumbered
        \setcounter{AlgoLine}{0}
        \KwIn{  $S$: empty solution $S$}

        \F: priority queue

        \ForEach(delivery node j){}{

            \F[$j$]=$M$ // \textit{M: big value}

        }

        \ForEach(customer i){}{
        Select a random order $o^p_i$

        Push $d^p_{i0}$ into \F: \F[$d^p_{i0}$]=$a_i$
        }

        \While{\F $\neq \emptyset$}
        {
        Pop $d^p_{ij}$ with the earliest delivery time from \F

        $l^p_{ij}$, $u^p_{ij}$ = ScheduleTasks($d^p_{ij},S$)

        $S \leftarrow S \cup (l^p_{ij}$, $u^p_{ij})  $

        $q_i \leftarrow q_i-q^p_{ij}$;  $q^p_i \leftarrow q^p_i-q^p_{ij}$

        \uIf {$q^p_i \neq 0$ }{

        Push $d^p_{ij+1}$  into \F

        \F$[d^p_{ij+1}]=$ $ES_{d^p_{ij+1}]} \pm$ rand().
        }
        \uElseIf{ $q_i \neq 0$}{

        Select another order $o^{p'}_i$

        Push $d^{p'}_{i0}$ into \F: \F[$d^{p'}_{i0}$]=$ES_{d^{p'}_{i0}} \pm$ rand().
        }
    }

        \KwRet{$S$}
    \end{algorithm}
}
}

\section{Computational experiments}
\label{comp_exp}
\iffalse
\begin{table}
    \begin{tabular}[]{cc}
        \toprule
        Depot &  Capacity \\
        \hline
        005 & 90    \\
        010 & 80    \\
        020 & 80	\\
        \bottomrule
    \end{tabular}
\end{table}
\fi
\section{Conclusion}
\label{concl}

\vspace{0.1in}

\vspace{1.5cm} \noindent \textbf{Acknowledgments}

Financial support for this work was provided by the Canadian Natural Sciences and Engineering Research Council (NSERC) under grants 2015-04893 and 2019-00094. This support is gratefully acknowledged.

\bibliographystyle{plainnat}
\bibliography{References}

\end{document}

### 2024-03-09 ArticleCDP/original/versions/13_01_23/cdp_ali.tex
Save

--- 2023-08-15 Article/versions/04-03-23/article_04_03_23.tex
@@
 \newcommand{\pcmaxc}{P$|\mbox{\em cont}|$C$_{\max}$\xspace}
-
-\newcommand{\GRASP}{Greedy randomized adaptive search procedure\xspace}
+\newcommand{\alns}{adaptive large neighborhood search\xspace}
+\newcommand{\Alns}{Adaptive large neighborhood search\xspace}
 \newcommand{\F}{$\mathcal{F}$\xspace}
@@
 \newcommand{\C}{$\mathcal{C}$\xspace}
-\newcommand{\LB}{$\mathcal{L}$\xspace}
-\newcommand{\Lb}{$\mathcal{L}_b$\xspace}
+\newcommand{\Lb}{$\mathcal{L}$\xspace}
 \newcommand{\D}{$\mathcal{D}$\xspace}
- \newcommand*{\red}{\textcolor{red}}
 
@@
 Concrete is a widely used building material in construction projects.
-Its perishable nature is affected by many factors that impact its quality \citep{sinha_quality_2021}, which is crucial for the durability and strength of the final construction.
+It is a perishable product that can be affected by many different factors that impact its quality \citep{sinha_quality_2021} which is crucial for the durability and strength of the final construction.
+
 Concrete comes in two types: ready-mixed concrete (RMC) and site-mixed concrete (SMC).
 RMC is manufactured in a batch plant and delivered to the construction site, while SMC is produced on-site using raw materials stored on the construction site.
-Using SMC can avoid delays caused by road traffic, but it has a slower and more difficult production process, requires storage for mixing materials and equipment, and is suitable for low amounts of concrete.
+Site-mixed concrete can avoid delays caused by road traffic, but it has a slower and more difficult production process, requires storage for mixing materials and equipment, and is suitable for low amounts of concrete.
 On the other hand, RMC has better quality and benefits from lower production costs \citep{muresan_comparing}.
-However, to take advantage of these benefits, the batch plant manager must ensure efficient and prompt delivery on the construction site, which may require a fleet of high-cost revolving drum trucks (concrete mixers) to dispatch the RMC.
+However, to take advantage of these benefits, the batch plant manager must ensure efficient and prompt delivery on the construction site, which may require a fleet of high-cost revolving drum trucks (concrete mixers) to dispatch the ready-mixed concrete.
 
-Concrete delivery under the form of RMC is subject to many operational constraints that make the Concrete Delivery Problem (CDP) very challenging.
-In this paper, we study a variant of the CDP to schedule the daily production and dispatching of RMC for a company located in the province of Quebec, Canada.
-This company operates multiple batching plants with varying production rates, using a fleet of concrete mixers of different capacities.
-Each plant has its own fleet of trucks; however, under certain conditions, the trucks can move between plants if necessary.
+Concrete delivery under the form of RMC is subject to many operational constraints that make it a challenging problem met in Operations Research and addressed by the Concrete Delivery Problem (CDP).
+In this paper, we study a variant of the Concrete Delivery Problem to schedule the daily production and dispatching of RMC for a company located in the province of Quebec, Canada.
+This company operates multiple batching plants with varying production rates, using a fleet of concrete mixers with different capacities.
+Each plant has its own fleet of trucks, however, the trucks are able to move between plants if necessary.
 The trucks must return to their home plant at the end of the day.
-The company owns two types of trucks with different capacities and can call on an external fleet when needed.
-They serve construction sites from any of their plants, with the first delivery starting at the time specified by the customer.
+The company owns two types of trucks with different capacities but also has the option to call on an external fleet when needed.
+They serve construction sites from any of their batch plants, with the first delivery starting at the time specified by the customer.
 The loading and unloading of a concrete mixer depend on the truck capacity, the loading rate at the plant, and the unloading rate at the construction site.
@@
 This constraint generalizes the linked order constraints of \cite{durbin2008or} where some customers place two orders for the same day and request that they are linked (the second order begins only after the first order is completed).
-The setting of our study is similar to a variant of the CDP previously studied in \cite{schmid2009hybrid, schmid2010hybridization}.
-However, we include the plant's production rate and driver shift schedule.
+The setting of this study is similar to a variant of the Concrete Delivery Problem previously studied in \cite{schmid2009hybrid, schmid2010hybridization}.
+However, it includes the plant's production rate and driver shift schedule and does not consider unloading instrumentation in contrast to the previous research.
 The company uses a centralized dispatcher system to schedule all daily orders, but this system has issues satisfying all daily demands without using an external fleet.
@@
 
-According to \cite{blazewicz2019handbook}, the CDP combines vehicle routing with scheduling issues to plan routes to deliver concrete from batch plants (depots) to customers' construction sites.
-RMC is an on-demand product with a short life cycle from production through end use.
-It cannot be stored and cannot stay too long on a truck, or it will harden.
+According to \cite{blazewicz2019handbook}, the concrete delivery problem combines vehicle routing with scheduling issues to plan routes to deliver concrete from batch plants (depots) to customers' construction sites.
+Ready-mixed concrete is an on-demand product with a short life cycle from production through end use.
+It cannot be stored and cannot stay too long in a truck, or it will harden.
 Hence, concrete mixers must deliver RMC at the planned construction site shortly after its production.
-\cite{tommelein1999just} describes RMC production and delivery as an example of a just-in-time (JIT) production system in construction.
+RMC production and delivery is described by \cite{tommelein1999just} as an example of a Just-In-Time (JIT) production system in construction.
 A customer quantity requirement is often greater than the truck size and must be fulfilled by multiple deliveries.
-In that sense, CDP is similar to the vehicle routing problem with split delivery \citep{archetti2008split}, except that the same truck may visit a customer more than once.
-Concrete hardens quickly, so multiple deliveries must be done back-to-back or at least close in time to avoid the problem of cold joint, which can reduce the strength and durability of the concrete.
+In that sense, CDP is like the Vehicle Routing Problem (VRP) with split delivery (VRPSD) \citep{archetti2008split}, except that the same truck may visit a customer more than once.
+Concrete hardens quickly, so when there are multiple deliveries, they must be done back-to-back or at least close in time to avoid the problem of cold joint, which can reduce the strength and durability of the concrete.
 Customers request to be served within a specific time window, which can complicate truck-loading schedules when a plant can only load one truck at a time.
 Similarly, only one truck can unload at a time at a customer location, sometimes leading to concrete mixers queuing and waiting their turn to deliver.
-Furthermore, with trucks of varied sizes, loading, travel, and unloading times that may be uncertain, the CDP is a complex and challenging problem.
+Furthermore, with trucks of varied sizes, loading, travel, and unloading times that may be uncertain, the Concrete Delivery Problem is a complex problem to solve and has been proven to be NP-hard \citep{asbach2009analysis,kinable2014concrete}.
 
-In this paper, we propose a mathematical model and a Greedy Randomized Search Procedure (GRASP) heuristic solution to solve a new variant of the CDP.
+In this paper, we propose a mathematical model and a Greedy Randomized Search Procedure (GRASP) heuristic solution to solve a new variant for the CDP.
 Our model takes into consideration the working shifts of the drivers and the scheduling of multiple orders within the same time window at a construction site.
@@
 Section \ref{desc_form} provides a formal description and mathematical model of the problem.
-In Section \ref{grasp_method}, we describe the GRASP algorithm and the constructive heuristics developed to solve this variant of the CDP.
+In Section \ref{method}, we describe the Greedy Randomized Search Procedure algorithm and the constructive heuristics developed to solve this variant of the CDP.
 Section \ref{comp_exp}  presents computational experiments to evaluate the proposed approach, and finally, the conclusions are presented in Section \ref{concl}.
@@
 
-Academic Research on concrete batching and delivery began in the late 1990s.
-\cite{tommelein1999just} described RMC as a prototypical example of a JIT production system in construction and identified two practices for delivering it.
-One approach is for the customer to haul the product from the batch plant with their concrete mixer, while the other is for the batch plant to deliver the concrete directly to the customer's location.
+A text mining approach by \cite{maghrebi2015text} for reviewing the ready-mixed concrete literature showed that concrete technology and material science are the main core of research in this area.
+Academic research on concrete batching and delivery began in the late 1990s.
+\cite{tommelein1999just} described ready-mixed concrete (RMC) as a prototypical example of a Just-In-Time production system in construction and identified two practices for delivering it.
+One approach is for the customer to haul the product from the batch plant with their concrete mixer, while the other approach is for the batch plant to deliver the concrete directly to the customer's location.
 This latter approach is the one that has been studied in all related papers found in the literature.
+
 Several works to schedule and dispatch concrete production and delivery have mainly focused on simulation-related methods.
 These methods can be standalone, such as those used by \cite{zayed2001simulation, wang2001scheduling, tian_simulation_based_2010, panas_simulation_based_2013} and \cite{galic2016simulation}, among others.
-Alternatively, these methods can be hybridized with optimization techniques, such as those used by \cite{feng2004optimizing, lu2005optimized}, and \cite{feng_integrating_2006}.
+Alternatively, these methods can be hybridized with optimization techniques, such as those used by \cite{feng2004optimizing, lu2005optimized, feng_integrating_2006}, etc.
 \cite{wang2001scheduling} developed a simulation model to reveal the effect and value of the concrete mixers' inter-arrival time on the productivity of hired unloading equipment on site.
-\cite{feng2004optimizing} used a combination of genetic algorithm (GA) and simulation process to minimize the total waiting time for trucks at a customer site.
+\cite{feng2004optimizing} used a combination of Genetic Algorithm (GA) and simulation process to minimize the total waiting time for trucks at a customer site.
 The study focused on loading trucks with identical capacities at the same batch plant, with fixed loading and unloading durations.
@@
 As such, the simulation model also included the batching and delivery of mortar.
-Finding the best RMC fleet size was also the purpose of the discrete-event simulation model proposed by \cite{panas_simulation_based_2013}.
+Finding the best RMC truck size was also the purpose of the discrete-event simulation model proposed by \cite{panas_simulation_based_2013}.
 
 In addition to simulation-based methods, several other approaches have been used in the literature to solve the CDP.
-These include metaheuristics \citep{faria2006distributed, misir2011selection, maghrebi2016sequential, yang2022concrete}, exact methods \citep{yan2007optimal, asbach2009analysis, kinable2014concrete}, matheuristics \citep{schmid2009hybrid, schmid2010hybridization}, Benders Decomposition \citep{maghrebi2014benders}, column generation (CG) \citep{maghrebi2014solving, maghrebi2016column}, Lagrangian relaxation \citep{narayanan2015using}, and machine learning approaches such as those used by \cite{graham2006modeling, maghrebi2014exploring, maghrebi2016matching}.
+These include metaheuristics \citep{faria2006distributed, misir2011selection, maghrebi2016sequential, yang2022concrete}, exact methods \citep{yan2007optimal, asbach2009analysis, kinable2014concrete}, matheuristics \citep{schmid2009hybrid, schmid2010hybridization}, Benders Decomposition \citep{maghrebi2014benders}, Column Generation (CG) \citep{maghrebi2014solving, maghrebi2016column}, Lagrangian relaxation \citep{narayanan2015using}, and machine learning approaches such as those used by \cite{graham2006modeling, maghrebi2014exploring, maghrebi2016matching}.
 \cite{matsatsinis2004towards} designed a decision support system (DSS) for the dynamic routing of both concrete and pumps that may be necessary for some construction sites to aid in the unloading of concrete.
@@
 Orders that could not be executed immediately could be postponed for the next day.
-The routing of the pumps was modeled as a multi-depot vehicle routing problem with time windows.
+The routing of the pumps was modelled as a multi-depot vehicle routing problem with time windows.
 \cite{naso2007genetic} proposed a sequential GA method combined with constructive heuristics to solve another variant of the CDP.
@@
 The authors also developed a non-linear model that minimizes transportation costs, waiting times, outsourced costs, and overtime work.
-They ran experiments using real-world instances of a concrete supply chain in the Netherlands and found a reduction in the number of outsourced requests.
+They ran experiments using real-world instances of a concrete supply chain in the Netherlands and found a reduction in the number of requests redirected to external companies.
 \cite{yan2007optimal} also considered overtime considerations in their paper, which focused on scheduling RMC for one batching plant with two loading docks.
 The study took into account that overtime wages are paid for factory and construction site operations after 4 PM.
-They developed a mixed-integer programming (MIP) model on a time-space network to minimize travel times and operating costs at both normal and overtime working hours at the plant and the construction sites.
-They tested the model using real data consisting of three days of operation using a two-stage algorithm.
+They developed a mixed integer programming (MIP) model on a time-space network to minimize travel times and operating costs at both normal and overtime working hours at the plant and the construction sites.
+They tested the model using real-data consisting of 3 days of operation using a two-stage algorithm.
 First, they solved the MIP relaxation with CPLEX.
@@
 The algorithm was found to improve the actual plant operation by 10\%.
-A time-space network is the key component of the real-time DSS developed by \cite{durbin2008or} to solve a dynamic CDP every five minutes.
+A time-space network is the key component of the real-time decision support system (DSS) developed by \cite{durbin2008or} to solve a dynamic CDP problem every five minutes.
 The DSS is able to receive new orders, schedule them on the fly, and handle unexpected events such as plant closures, truck breakdowns, and delays in transportation times.
-The authors combined the DSS with a tabu search (TS) heuristic to warm start CPLEX, which made the model performant enough to solve instances with up to 1,500 loads per day with up to 250 trucks.
+The authors combined the DSS with a Tabu Search (TS) heuristic to warm start CPLEX, which made the model performant enough to solve instances with up to 1,500 loads per day with up to 250 trucks.
 The DSS also considers the case of a customer who places two orders, with the first being completed before the second starts.
@@
 The method can quickly solve large problem instances with more than 60 orders per day without encountering any memory issues.
-The same problem is addressed by \cite{schmid2010hybridization}, who proposed a MIP model combined with a VNS and a very large neighborhood search (VLNS) to develop two matheuristics approaches.
+The same problem is addressed by \cite{schmid2010hybridization} who proposed a mixed integer programming (MIP) model combined with a VNS and a very large neighborhood search (VLNS) to develop two matheuristics approaches.
 Comparisons between both matheuristics and a standalone VNS show that the former methods are much better and suitable for solving larger problem instances.
@@
 One of the paper's main contributions is the creation of the first public test instances for the CDP with up to 50 customers, four batching plants, and 20 concrete mixers.
-They found the CP model to be highly effective in finding high-quality solutions in a relatively short time or improving existing schedules, while the MIP model can be used to compute bounds, as it seems ineffective in solving large problem instances.
+They found the CP model to be highly effective in finding high-quality solutions in relatively little time or improving existing schedules, while the MIP model can be used to compute bounds, as it seems ineffective in solving large problem instances.
 Finally, the heuristic often yields good solutions in less than a second.
-A detailed analysis of the MIP model presented in \cite{kinable2014concrete} and of two more compact models can be found in the thesis of \cite{hernandez_lopez_study_2020}.
-In \cite{kinable2014logic}, we found an attempt to solve the previous problem with a logic-based Benders' approach.
-\cite{sulaman2017simulated} expand upon the SD heuristic proposed in \cite{kinable2014concrete}, proposing a simulated annealing (SA) combined with a time-slot Heuristic (SATH).
-This method looks for a slot between existing visits of a truck to schedule a new delivery instead of assigning it to the time slot strictly after the truck's latest assigned delivery.
+A detailed analysis of the MIP model presented in cite{kinable2014concrete}and of two more compact models can be found in the thesis of \cite{hernandez_lopez_study_2020}.
+In \cite{kinable2014logic}, we found an attempt to solve the previous problem with a Logic Based Benders' approach.
+\cite{sulaman2017simulated} expand upon the SD heuristic proposed in \cite{kinable2014concrete}, proposing a Simulated Annealing (SA) with a Time-Slot Heuristic (TH).
+TH mechanism is to look for a slot between existing visits of a truck to schedule a new delivery instead of assigning it to the time slot strictly after the truck's latest assigned delivery.
 The goal is to reduce the large time gaps that can be present in a schedule created with SD due to ignoring the intermediate available time slots.
@@
 The MIP model is used in a local search scheme as a black-box solver to reoptimize an incumbent solution in which a neighborhood operator has unfixed some variables.
-\cite{tzanetos2023systematic} provide an overview of the various methods used in the literature to address the CDP and categorizes the problem formulations based on the different concepts used in the literature.
+\cite{tzanetos2023systematic} provides an overview of the various methods used in the literature to address the Concrete Delivery Problem and categorizes the problem formulations based on the different concepts used in the literature, such as Integer Programming (IP), time-space network, and job shop model.
 They also discussed the consistency between industry needs and existing constraints and provided insights into the datasets corresponding to real-world cases, identifying the necessary data for practitioners.
@@
 \label{desc_form}
-The focus of this paper is the distribution of RMC from a Canadian company that operates in the greater Montreal area.
-When a customer places an order, it is received at a central center and assigned to one of the company's batching plants.
+The focus of this paper is the distribution of ready-mixed concrete from a Canadian company that operates in the greater Montreal area.
+When a customer places an order, it is received at a central center and then assigned to one of the company's batching plants.
 These plants produce the concrete and then deliver it to the customer.
-The problem we are examining involves a set of customer orders, a set of concrete-mixer drivers, and a set of batching plants.
+The problem we are examining involves a set of batching plants, a set of customer orders, and a set of concrete-mixer drivers.
+
+% \subsection*{Batching plants}
+
+The company owns a total of eight batching plants, located in various geographical regions.
+Each plant has a loading bay that can only accommodate one truck at a time, which leads to trucks lining up.
+Let $\mathcal{B}$ be the set of batching plants.
+The plants are heterogeneous, as each plant has its own hourly loading rate, represented by $\tau^l_b$, which affects the duration of the loading process.
+After the concrete has been loaded, the driver spends $\alpha_b$ minutes adjusting the concrete in the truck before heading to the customer site.
+A batching plant can serve a construction site as long as the travel time between the two is less than the concrete's lifespan, which is represented by $\Delta$.
+Each plant has its own assigned fleet of trucks, but it can also borrow trucks from other plants or hire external fleets if necessary.
 
@@
 
-A customer $i$ requests one or more types of concrete to be delivered to his construction site on a particular day, starting at time $a_i$.
-We call an order $o$ a request for a specific type of concrete.
-$q_i$ is the sum of the demand $q_o$ of each order $o$ placed, $a_i$ is the desired arrival time of the first concrete mixer, and $\tau^u_i$ is the unloading rate for customer $i$.
+A customer $i$ requests a total of $q_i$ of one or several types of concrete to be delivered to his construction site on a particular day.
+We call an order a request for a specific type of concrete.
+When placing an order $o$, the customer $i$ specifies the required quantity, $q_o$, the desired arrival time of the first concrete mixer, $a_i$, and the unloading rate at their construction site, $\tau^u_i$.
 If the order requires more concrete than a single truck can carry, multiple deliveries are scheduled.
-
-Let $O_i$ be the set of all orders requested by customer $i$.
-Each element of $O_i$ must be fully delivered before moving on to another order.
-Exactly one order $o$ of $O_i$ must have its first delivery start at $a_i$, while the others can start at any time after $o$ is completed.
-Once a plant is selected to produce a customer's first delivery, it must be the supplier of all subsequent deliveries from that customer.
-To avoid cold joint problems with the concrete, the subsequent deliveries must be made in close succession.
-We define a maximum time delay $\gamma_i$ after which no more deliveries will be allowed.
-The customer unloading rate and the quantity to be unloaded give the time required to unload a truckload.
-Let $\mathcal{C}$ be the set of construction sites (customers) with a planned delivery for the day, and $\mathcal{O}=\{{O_i, i \in \mathcal{C}}\}$ be the set of all requested orders for all customers.
+Let $O_i$ be the unordered set of all orders requested by customer $i$.
+Each element of $O_i$ must be thoroughly delivered before moving on to another order.
+One element $o$ of $O_i$ must have its first delivery start at $a_i$, while the others can start any time after all deliveries of $o$.
+Once a plant is chosen to produce a customer's first delivery, it must be the provider of all subsequent deliveries of this customer.
+To prevent cold joint issues with the concrete, subsequent deliveries must be made just in time, or at least close in time.
+We define a maximum time lag $\gamma_i$ beyond which no next unloading operation should be allowed.
+The construction site unloading rate and the quantity to unload give the time necessary to discharge a truckload.
+$\mathcal{C}$ is the set of construction sites (customers) with a planned delivery for the day.
+$\mathcal{O}$ is the set of all requested orders.
 
@@
 The company has two types of concrete mixer trucks with capacities of 8 and 12 cubic meters.
-Each driver $k$ is assigned to a particular batch plant and is responsible for driving a truck with capacity $Q_k$.
-The set of drivers is represented by $K =\cup_{b \in \mathcal{B}}K_b $, where $K_b$ is the set of drivers scheduled to start their shift at batch plant $b$.
+Each driver $k$ is assigned to a specific batching plant and is responsible for driving a truck with a specific capacity $Q_k$.
+The set of drivers is represented by $K =\cup_{b \in \mathcal{B}}K_b $, $K_b$ being the set of drivers scheduled to start their shift at the batching plant $b$.
 $t_{ij}$ is the known time to travel between any two locations $i$ and $j$.
-A scheduled driver $k$ is required to start his shift at $H_k$, work a minimum of $M_T$ hours and a maximum of $N_T$ hours during regular working hours, with the possibility of overtime of up to $O_T$ hours.
-$\beta_3$ and $\beta_4$ are the penalties incurred if a driver works less than $M_t$ and more than $N_t$.
+A scheduled driver $k$ is required to begin their shift at $H_k$, work for a minimum of $M_T$ hours and a maximum of $N_T$ hours during regular working hours, with the possibility of overtime of up to $O_T$ hours.
+%Let the continuous variable $WT_k$ be the working time of each driver $k$ during a day ($WT_k \leq O_T$).
+$\Gamma_1$ and $\Gamma_2$ are the incurred penalties when a driver works less than $M_t$ and more than $N_t$.
 
-A driver typically loads RMC at his assigned batch plant but may be required to drive to and load at other plants if needed.
-The batch plant produces concrete on demand using recipes specific to each order.
-This means that a truck can only haul RMC for one order, even if there is spare capacity.
-To fill multiple orders, a driver must restock at a plant between deliveries.
-After unloading the RMC, a driver takes $\beta_k$ minutes to clean the concrete mixer before proceeding.
+A driver typically loads RMC at their assigned batching plant, but may also be required to drive to and load at other plants if needed.
+The batching plant produces concrete on-demand with recipes specific to each customer.
+This means that a truck cannot hold orders for more than one customer, even if there is spare capacity.
+To fulfill multiple orders, a driver must refill at a plant between deliveries.
+After unloading the RMC, a driver takes $\beta_k$ minutes to clean the concrete mixer before travelling to his next loading plant.
 %When assigning drivers to deliveries, the company prioritizes employees with the highest seniority.
 
-Let $n_o$ be the number of deliveries needed to fulfill the order $o$.
-$n_o$ is not known in advance because we use a fleet of trucks with different capacities.
-However, we can compute its lower ($n_o^{min}$) and upper ($n_o^{max}$) bounds using the capacities of the largest ($Q_{max}$) and smallest ($Q_{min}$) available trucks.
-The lower bound is the number of deliveries required if we only use trucks with capacity $Q_{max}$, while the upper bound is the number of deliveries required if we only use trucks with capacity $Q_{min}$.
-\begin{alignat}{3}
-    \label{mod:c0}
-       n_o^{min} = \left\lceil \frac{q_o}{Q_{max}} \right\rceil \leq n_o \leq n_o^{max} = \left\lceil \frac{q_o}{Q_{min}} \right\rceil  & \text{ } & 
- \forall  o \in \mathcal{O}.
-\end{alignat}
+% \subsection*{Model}
 
-Let $d^j_{o}$ be the $j_{th}$ visit with load $q^j_{o}$ for order $o$.
-We represent the fulfillment of order $o$ by the visits to the ordered set of delivery nodes $\mathcal{D}_o= \left(d^0_{o},d^1_{o},\cdots, d^{n_o}_{o}\right)$.
-The deliveries of customer $i$ are the ordered set $\mathcal{D}_i= (\mathcal{D}_{o_1}, \mathcal{D}_{o_2},\cdots,\mathcal{D}_{o_{|O_i|}})$, where $o_r$ is the $rth$ delivered order.
-We will refer to $d \in \mathcal{D}_i$ ($d \in \mathcal{D}_o$) as the $dth$ potential delivery of customer $i$ (order $o$).
-$\mathcal{D}=\bigcup_{i\in \mathcal{C}} \mathcal{D}_i$ is the union of all delivery nodes.
-
-% \subsection*{Batching plants}
-
-The company owns a total of eight batching plants located in various geographical regions.
-Each plant has a loading dock that can only accommodate only one truck at a time, which leads to trucks lining up.
-Let $\mathcal{B}$ be the set of batching plants.
-The plants are heterogeneous, as each plant $b$ has its own hourly loading rate, represented by $\tau^l_b$, which affects the duration of the loading process.
-After loading the concrete, the driver spends $\alpha_b$ minutes adjusting the concrete in the truck before heading to the customer site.
-Each plant has its own assigned fleet of trucks, but it can borrow trucks from other plants or hire external vehicles if necessary.
-A batching plant can serve a construction site as long as the travel time between the two is less than the concrete's lifespan, which is represented by $\Delta$.
-Let $l_{b,j}$ be the loading dock node associated with delivery node $j$ at plant $b$.
-We represent each plant $b$ by the set $\mathcal{L}_b = \{l_{b,j}, t_{bj}\leq  \Delta, j \in \mathcal{D} \}$  of loading docks nodes.
-$\mathcal{L}=\bigcup_{b\in B} \mathcal{L}_b$ is the union of all loading docks.
-
-% \subsection*{Model}
 A solution to the problem involves making decisions about truck loading schedules, driver assignments to different deliveries, and truck arrival times at construction sites for unloading.
@@
 
-Each driver leaves and returns to his home plant every day.
-We represent the home plant of a driver $k$ with a starting depot $s_k$ and an ending depot $e_k$.
-$S$ and $E$ are the sets of starting and ending depots, respectively.
+We now present a formal definition of our problem.
+Let $n_o$ be the number of deliveries needed to fulfill order $o$.
+$n_o$ is not known in advance, as we use a fleet of trucks with various capacities.
+However, we can compute its lower ($n_o^{min}$) and upper ($n_o^{max}$) bounds using the capacities of the highest ($Q_{max}$) and smallest ($Q_{min}$) available trucks.
+The lower bound is the number of deliveries needed if we only use trucks of capacity $Q_{max}$, while the upper bound is the number of deliveries needed if we only use trucks of capacity $Q_{min}$.
+
+\begin{alignat}{3}
+    \label{mod:c1}
+       \left\lceil \frac{q_o}{Q_{max}} \right\rceil \leq n_o \leq \left\lceil \frac{q_o}{Q_{min}} \right\rceil  & \text{ } & 
+ \forall  o \in \mathcal{O}.
+\end{alignat}
+
+Let $d^j_{o}$ be the $j_{th}$ visit with load $q^j_{o}$ for order $o$.
+We represent the delivery of order $o$ by the visits to the ordered set of nodes $\mathcal{D}_o= \left(d^0_{o},d^1_{o},\cdots, d^{n_o}_{o}\right)$.
+We represent the deliveries of customer $i$ by the ordered set $\mathcal{D}_i= (\mathcal{D}_{o_1}, \mathcal{D}_{o_2},\cdots,\mathcal{D}_{o_{|O_i|}})$, where $o_r$ is the $rth$ order delivered.
+We will refer to $d \in \mathcal{D}_i$ ($d \in \mathcal{D}_o$) as the $dth$ potential delivery of customer $i$ (order $o$).
+$\mathcal{D}=\bigcup_{i\in \mathcal{C}} \mathcal{D}_i$ is the union of all delivery nodes.
+
+We represent each plant $b$ by the ordered set $L_b=\{L_{b,1}, L_{b,2},\cdots, L_{b,n(b)}\}$ of loading docks nodes, with $n(b)$ being the number of delivery nodes at a maximal travel time of $\Delta$ of $b$.
+We will refer to $L_{b,l}$ or  $l \in L_b$  as the $lth$ potential loading operation at $b$.
+$\mathcal{L}=\bigcup_{b\in B} L_b$ is the union of all loading docks.
+
+Each driver leaves and returns to his home plant each day. We represent a driver $k$ home plant with a starting $s_k$ and ending $e_k$ depot. $S$ and $E$ are the sets of starting and ending depot.
 
 We define our problem on a complete directed graph where $V=\{ S \cup \mathcal{L} \cup \mathcal{D} \cup E\}$ is the set of nodes.
-The arc sets are $A =  \{(i,j,k) \hspace*{1mm} \vert \hspace*{1mm} i, j \in V \hspace{1mm} k\in K \}$, $A^D = \{(i,j) \hspace*{1mm} \vert \hspace*{1mm} i, j \in \mathcal{D} \hspace{1mm} \}$, and $A^L = \{(i,j) \hspace*{1mm} \vert \hspace*{1mm} i, j \in \mathcal{L} \hspace{1mm} \}$.
+The arc sets are $A =  \{(i,j,k) \hspace*{1mm} \vert \hspace*{1mm} i, j \in V \hspace{1mm} k\in K \}$, and $A^D = \{(i,j) \hspace*{1mm} \vert \hspace*{1mm} i, j \in \mathcal{D} \hspace{1mm} \}$.
 $A$ corresponds to allowed movements of drivers from node $i$ to node $j$. For each driver $k$, the allowed movements are the following:
@@
 For a customer $c$, arcs in $A^D$ link consecutive delivery nodes of the same order   $\lbrace (i,j)\in \mathcal{D}_o, o \in \mathcal{O}_c, i < j  \rbrace$, and pair of delivery nodes of two different orders $ \lbrace (i,d^{0}_{o_2}),  i \in \mathcal{D}_{o_1}, i \geq n^{min}_{o_1}, o_1, o_2 \in \mathcal{O}_c, o_1 \neq o_2 \rbrace $.
-Arcs in $A^L$ link all pairs of loading docks of the same batching plant.
 
 $\delta^{+}(i) = \{(i, j,k) \in A \}$, and $\delta^{-}(i) = \{(j, i,k) \in A \}$ are the outcoming and incoming arc sets of any node $i \in V$.
-$\delta^{+}_D(i) = \{(i, j)  \in A^D\}$, and $\delta^{-}_D(i) = \{(j, i) \in A^D \}$ are the outcoming and incoming arc sets of delivery node $i \in \mathcal{D}$.
-Similarly, $\delta^{+}_L(i) = \{(i, j)  \in A^L\}$, and $\delta^{-}_L(i) = \{(j, i) \in A^L \}$ are the outcoming and incoming arc sets of loading node $i \in \mathcal{D}$.
+Similarly, $\delta^{+}_D(i) = \{(i, j)  \in A^D\}$, and $\delta^{-}_D(i) = \{(j, i) \in A^D \}$ are the outcoming and incoming arc sets of delivery node $i \in \mathcal{D}$.
 
-Let the binary variable $x^{k}_{ij}$ be $1$ if the driver $k$ travels from node $i$ to $j$.
-The binary variable $y_o$ is $1$ when the order $o$ is completely served.
-$v_i$ and $w_i$ are the start and end of the loading (unloading) operation at node $i \in \mathcal{L} \cup \mathcal{D}$.
-The binary variable $u_{i,j}$ is $1$ if node $j$ is served just after $i$, the service being either an unloading or a loading operation.
-The binary variable $\sigma_{ib}$ is equal to 1 if the orders of customer $i$ are loaded by plant $b$.
-Variable $q^k_j$ is the quantity to be loaded towards $j$ with vehicle $k$.
-Let $w^1_k$ be a continuous variable indicating the difference between the driver's work time and the minimum number of hours to be worked in a day.
-$w^2_k$ is a continuous variable indicating the difference between the driver's work time and the normal work time.
-Let $g_i$ be the time between due date and first service start for customer $i$.
+Let the binary variable $x^{k}_{ij}$ be $1$ if driver $k$ travels from node $i$ to $j$.
+Binary variable $y_i$ equals $1$ if customer $i$ is completely served.
+$v_i$ is the start of the loading (unloading) operation at node $i \in L \cup D$.
+$w_i$ is the end of the loading (unloading) operation at node $i \in L\cup D$.
+The binary variable $u_{i,j}$ equals $1$ if delivery node $j$ is served just after delivery $i$.
+Binary variable $\sigma_{ib}$ equals 1 if customer $i$ orders' are loaded from plant $b$.
 % Binary variable $u_{o_{1},o_{2}}$ equals $1$ if order $o_{2}$ is served directly after order $o_{1}$.
 Binary variables $z_{o,d}$ ($z_{i,d}$) indicates whether or not delivery $d \in  \mathcal{D}_o$ ($d \in  \mathcal{D}_i$) is supposed to be executed.
-$w_i$ is the end of the loading (unloading) operation at node $i \in L\cup D$.
 
-The objective function minimizes the travel cost, the penalty costs incurred when customer demands are not fully met, the lateness of each customer's first delivery, the cost of drivers working less than the minimum hours, and the total overtime cost of drivers working beyond their scheduled hours.
-The mathematical model of this variant of the CDP is as follows:
+A driver $k$ leaves his starting node exactly once a day and travels to either a loading dock or his ending node.
+When loading at a loading dock $i$, he takes some time to adjust the concrete in the truck before travelling to delivery node $j$ to unload.
+The loading operation duration depends on the plant loading rate and the quantity $q^k_j$ of RMC to load towards $j$.
+After completing his last unloading operation of the day, he returns to his end node.
 
+\begin{alignat}{3}
+\label{mod:c2} \sum_{j \in \delta^{+}(s_k) }{x^k_{s_kj}} = 1 & \text{ } &\forall k \in K \\
+\label{mod:c3}   v_j \geq  w_i + \alpha_{i} + t_{ij} - M(1-x^{k}_{ij}) & \text{ }& \forall i \in \mathcal{L}, j \in \mathcal{D},  k \in K \\
+\label{mod:c4}  w_{i} \geq v_{i}  + \frac{q^k_j}{\tau^l_b} -M(1- x^{k}_{ij}) &  \hspace{5mm} &\forall  b \in \mathcal{B},  i \in L_{b}, j \in \mathcal{D}, k \in K \\
+\label{mod:c5} \sum_{j \in \delta^{-}(e_k)}{x^k_{je_k}} = 1 & \text{ }   &  \forall k \in K
+\end{alignat}
+
+Constraints (\ref{mod:c4}) ensure that two trucks cannot be loaded simultaneously in a plant.
+
+\begin{alignat}{3}
+    \label{mod:c6}       w_{l-1} \leq v_{l} & \hspace{5mm} &\forall  b \in B,  l \in L_{b}, l \geq 1 
+\end{alignat}
+
+After completing a delivery  at node $i$, a driver must clean the truck, and travel to the next loading depot or end node.
+The unloading service depends on the construction site's rate and the received quantity.
+Unloading operations must finish $\Delta$ minutes after the beginning of loading.
+\begin{alignat}{3}
+    \label{mod:c7}     v_j \geq  w_i + \beta_{k} + t_{ij} - M\left(1-x^{k}_{ij}\right) & \text{ }& \forall i \in \mathcal{D}, j \in \mathcal{L} \cup \{e_k\},   k \in K \\
+    \label{mod:c8}          w_{j} \geq v_{j}  + {  \frac{q^k_j}{\tau^u_c} -M\left(1- x^{k}_{ij}\right) } & \text{ }&  \forall c \in  \mathcal{C}, j \in \mathcal{D}_{c}, i \in \mathcal{L},  k \in K \\
+    \label{mod:c31}          w_{j} \leq v_{i}  + \Delta + M\left(1- x^{k}_{ij}\right)&  \text{ }&  \forall c \in  \mathcal{C}, j \in \mathcal{D}_{c}, i \in \mathcal{L},  k \in K
+    % \label{mod:c31}      w_{d-1}  \leq v_{d} & \hspace{5mm} & \forall i \in \mathcal{C}, o \in O_i, d \in \mathcal{D}_{o}, d \geq 1 \\
+    % \label{mod:c10}      z_{o,d} \leq z_{o,d-1} & \hspace{5mm} & \forall i \in \mathcal{C}, o \in O_i, d \in \mathcal{D}_{o}, d \geq 1\\
+    % \label{mod:c11}  v_{d} - w_{d-1} \leq \gamma_i   & \hspace{5mm}&  \forall i \in \mathcal{C}, o \in O_i, d \in \mathcal{D}_{o}, d \geq 1
+\end{alignat}
+
+The first service of any customer $i$ has to start at the due time $a_i$.
+Let $g_i$ be the gap between the due time and the start of the first service at $i$.
+This first service may be performed at the first delivery node of any order $\in O_i$.
+Constraints (\ref{mod:c11}--\ref{mod:c13}) are used to find the delivery sequence of the first delivery node of all orders $\in \mathcal{O}_i$.
+
+\begin{alignat}{3}
+    \label{mod:c9}   v_{d^0_{o}} \geq a_i & & \forall  i \in \mathcal{C}, \forall o \in \mathcal{O}_i\\
+    \label{mod:c10}  g_{i} \geq v_{d} - a_i - M\left(  \sum_{ j \in \delta^{-}_{D}(d)  }{u_{j,d}}\right) & & \forall  i \in \mathcal{C}, \forall o_1 \in \mathcal{O}_i, d=d^0_{o_1} \\
+    \label{mod:c11}   \sum_{ \substack{o_1 \in \mathcal{O}_i }}{ \sum_{ j \in \delta^{-}_{D}(d^0_{o_1})}}{u_{j,d^0_{o_1} } } = |\mathcal{O}_i|-1 & \text{ }& \forall  i \in \mathcal{C},|\mathcal{O}_i| > 1 \\
+    \label{mod:c12}   \sum_{ \substack{o_1 \in \mathcal{O}_i }}\sum_{ j \in \delta^{+}_{D}(d^0_{o_1})}{u_{d^0_{o_1},j } } = |\mathcal{O}_i|-1 & \text{ }& \forall  i \in \mathcal{C}, |\mathcal{O}_i| > 1 \\
+    \label{mod:c13}  \hspace{3mm} \smashoperator{ \sum_{ j \in \delta^{+}_{D}(d^0_{o_1})}}{u_{d^0_{o_1},j } } \leq 1, \hspace{2mm} \smashoperator{ \sum_{ j \in \delta^{-}_{D}(d^0_{o_1})}}{u_{j,d^0_{o_1} } } \leq 1, \hspace{2mm} \smashoperator{ \sum_{ j \in \delta^{+}_{D}(d^0_{o_1})}}{u_{d^0_{o_1},j } } +\smashoperator{ \sum_{ j \in \delta^{-}_{D}(d^0_{o_1})}}{u_{j,d^0_{o_1} } } \geq1         & \text{ }& \forall  i \in \mathcal{C}, \forall o_1 \in \mathcal{O}_i
+\end{alignat}
+
+For a customer, an order must be completed before starting another.
+Constraints (\ref{mod:c14}--\ref{mod:c15}) enforce the precedence constraints between the last delivery of an order and the first delivery of the subsequent order.
+
+\begin{alignat}{3}
+  \label{mod:c14} & v_{d^0_{o_2}} \geq w_j - M(1- u_{j,d^0_{o_2}}) & \text{} & \forall i \in \mathcal{C}, \forall o_1, o_2 \in O_i, j \in \mathcal{D}_{o_1}, (j,d^0_{o_2}) \in \delta_{D}^{-}(d^0_{o_2})  \\
+  \label{mod:c15} &  v_{d^0_{o_2}}  \leq  w_j + \gamma_i + M(1- u_{j,d^0_{o_2}}) & \hspace{1mm} & \forall i \in \mathcal{C}, \forall o_1, o_2 \in O_i, j \in \mathcal{D}_{o_1}, (j,d^0_{o_2}) \in \delta_{D}^{-}(d^0_{o_2})
+%   \label{mod:c12} & v_{j} \geq w_d - M(1- u_{d,j}) & \text{} & \forall i \in \mathcal{C}, \forall o_1 \in O_i, d= d_{o_1,0}, j \in \mathcal{D}_{i}, j \neq  d\\
+%   \label{mod:c112} &  v_{j}  \leq  w_d + \gamma_i + M(1- u_{d,j}) & \text{ } & \forall i \in \mathcal{C}, j \in \mathcal{D}_{i}, o_1 \in O_i, d= d_{o_1,0},    j \neq  d\\
+%   \label{mod:c112} &  \smashoperator{\sum \limits_{j \in \mathcal{D}_i, j\neq d}}{u_{d,j}} \leq 1 & \text{ } & \forall i \in \mathcal{C}, o \in O_i,  d = d_{o,0} \\
+%   \label{mod:c113} &  \smashoperator{\sum \limits_{j \in \mathcal{D}_i, j\neq d}}{u_{j,d}} \leq 1 & \text{ } & \forall i \in \mathcal{C}, o \in O_i,  d = d_{o,0} \\
+%   \label{mod:c111} &  \smashoperator{\sum \limits_{j \in \mathcal{D}_i, j\neq d}}{u_{d,j}}+ \smashoperator{\sum_{j \in \mathcal{D}_i, j\neq d}}{u_{j,d}} \geq 1 & \text{ } & \forall i \in \mathcal{C}, o \in O_i,  d = d_{o,0}
+\end{alignat}
+
+Constraints (\ref{mod:c29}--\ref{mod:c30}) enforce the precedence constraints between two consecutive delivery nodes of the same order.
+
+\begin{alignat}{3}
+\label{mod:c29}   &   u_{j-1,j} \geq u_{j,j+1} & \hspace{5mm} & \forall i \in \mathcal{C}, \forall o \in O_i,  j \in \mathcal{D}_{o}, 1 \leq j \leq n_o-1\\
+  \label{mod:c30} &  u_{j-1,j} \geq \sum_{l \in \mathcal{L}}{x_{lj}} & \text{} & \forall i \in \mathcal{C}, \forall o \in O_i, j \in \mathcal{D}_{o}, j \geq 1 
+\end{alignat}
+
+Constraints (\ref{mod:c17}--\ref{mod:c18}) state that it is not possible for two trucks serving consecutive deliveries to be unloaded at the same time and there is a maximum time lag between them.
+
+\begin{alignat}{3}
+  \label{mod:c17} & v_{j} \geq w_{j-1} - M(1- u_{j-1,j}) & \text{} & \forall i \in \mathcal{C}, \forall o \in O_i, j \in \mathcal{D}_{o}, j \geq 1 \\
+  \label{mod:c18} &  v_{j}  \leq  w_{j-1} + \gamma_i + M(1- u_{j-1,j}) & \text{ } & \forall i \in \mathcal{C}, \forall o \in O_i, j \in \mathcal{D}_{o}, j \geq 1
+\end{alignat}
+
+With constraints (\ref{mod:c19}) and (\ref{mod:c20}), we respectively impose that the cumulative capacity of all concrete mixers serving an order and a customer may exceed the required quantities.
+
+\begin{alignat}{1}
+    \label{mod:c19} \sum_{d \in D_o}\sum_{k \in K}{q^k} \sum_{j \in \mathcal{L}}{x^k_{jd}}\geq q_o, \forall o \in \mathcal{O} \\
+    \label{mod:c20} \sum_{d \in D_i}\sum_{k \in K}{q^k} \sum_{j \in \mathcal{L}}{x^k_{jd}}\geq q_i, \forall i \in \mathcal{C}
+\end{alignat}
+
+Constraints (\ref{mod:c21}) and (\ref{mod:c22}) impose that a loading or delivery node may only be visited once by one driver. Constraints (\ref{mod:c23}) and (\ref{mod:c24}) are degree constraints.
+
+\begin{alignat}{3}
+\label{mod:c21} &  \sum_{k \in K}\sum_{j \in \mathcal{D}}{x^{k}_{ij} \leq 1} & \text{ } & i \in \mathcal{L}\\
+  \label{mod:c22}  &  \sum_{k \in K}\sum_{i \in \mathcal{L}}{x^{k}_{ij} \leq 1} & \text{ } & j \in \mathcal{D} \\
+  \label{mod:c23}  &  \sum_{k \in K}\sum_{i \in \mathcal{L}}{x^{k}_{ij} } =  \sum_{k \in K}\sum_{i \in \mathcal{L}}{x^{k}_{ji} } & \text{ } & j \in \mathcal{D} \\
+  \label{mod:c24}  &  \sum_{k \in K}\sum_{i \in \mathcal{D}}{x^{k}_{ij} } =  \sum_{k \in K}\sum_{i \in \mathcal{D}}{x^{k}_{ji} } & \text{ } & j \in \mathcal{L}
+\end{alignat}
+
+ Constraints (\ref{mod:c25}) and (\ref{mod:c26}) impose that all orders of customer $i$ must be loaded from the same plant.
+
+\begin{alignat}{2}
+    \label{mod:c25} & \sum_{k \in K} \sum_{l \in \mathcal{L}_b} \sum_{d \in \mathcal{D}}{x^k_{ld}} \leq M*\sigma_{i, b} & \forall i \in \mathcal{C}, b \in \mathcal{B} \\
+    \label{mod:c26} & \sum_{b \in \mathcal{B}} \sigma_{i, b} \leq 1 & \text{ } \forall i \in \mathcal{C}, b \in \mathcal{B} 
+\end{alignat}
+
+Let $w^1_k$ be the binary variables indicating if the driver $k$ has reached the minimum number of hours worked in the day.
+$w^2_k$ is a continuous variable indicating the gap between the driver working time and the normal working time.
+
+\begin{alignat}{3}
+    \label{mod:c27}  M_T*w^1_k \leq v_{e_k} - H_k & \hspace{5mm} &\forall k \in K \\
+    \label{mod:c28}  w^2_k \geq (v_{e_k} - H_k) - N_T & \hspace{5mm} &\forall k \in K.
+\end{alignat}
+
+The objective function (\ref{mod:obj}) minimizes the travelled distance, the penalty costs incurred when customers' demands are not fully satisfied, the lateness of each customer first delivery, the number of drivers who work under the minimum working time and the total overtime cost when drivers work beyond their scheduled hours.
 %The objective function (\ref{mod:obj}) is designed to minimize multiple factors that impact the overall cost and efficiency of the concrete delivery process.
@@
 The penalty costs $\beta_1$, which are incurred when customers' demands are not fully satisfied.The lateness of each customer's first delivery, which can impact customer satisfaction and lead to additional penalty costs.The number of drivers who work under the minimum working time, which can lead to increased labor costs and decreased productivity.The total overtime cost, which is incurred when drivers work beyond their scheduled hours.All these factors are taken into consideration in the objective function to minimize the cost and maximize the satisfaction of the customers.
+\begin{align}
+\label{mod:obj}
+\hspace{-20mm}
+    \begin{split}
+      \min  \smashoperator{\sum_{k \in K}} \sum_{ \substack{i \in \mathcal{L}} } \sum_{ j \in \delta^{+}{(i)} } {t_{ij}x^{k}_{ij}} + 
+      \sum_{k \in K} \sum_{ \substack{i \in \mathcal{D}}} \sum_{j \in \delta^{+}{(i)} } {t_{ij}x^{k}_{ij}} + 
+      \smashoperator{\sum_{ i \in C}}{ \beta_1 \left(1-y_i\right)} \hspace{-0mm} \\ +
+      \beta_2 \smashoperator{ \sum_{i \in \mathcal{C} } } { g_i  } 
+    %   + \beta_3 \smashoperator{\sum_{i \in \mathcal{C}}}{  {\sum_{d \in D_i}}{\left(v_d - w_{d-1} - \gamma_i\right)} } 
+      +  \sum_{k \in K}{\beta_3* w^{1}_k } + \beta_4 *{ w^{2}_k }
+    \end{split}
+\end{align}
 
-\begin{alignat}{3}
-\label{mod:obj}   \min &  \smashoperator{ \sum_{ \substack{(i,j,k) \in A} }} {t_{ij}x^{k}_{ij}} + \beta_1\smashoperator{\sum_{ o \in \mathcal{O}}}{  \left(1-y_o\right)} \hspace{-0mm}  + \beta_2 \smashoperator{ \sum_{i \in \mathcal{C} } } { g_i  } +  \sum_{k \in K}{\beta_3* w^{1}_k } + \beta_4 *{ w^{2}_k }&\\
-    \label{mod:c1} &\sum_{j \in \delta^{+}(s_k) }{x^k_{s_kj}} = 1   &  \forall k \in K  \\
-    \label{mod:c2} &\sum_{j \in \delta^{-}(e_k)}{x^k_{je_k}} = 1   &  \forall k \in K\\
-    \label{mod:c3} & v_j \geq  w_i + \alpha_{i} + t_{ij} - M\left(1-x^{k}_{ij}\right)  & \hspace{-20mm} \forall i \in \mathcal{L}, j \in \delta^{+}(i),  k \in K \\
-    \label{mod:c4}  & v_j \geq  w_i + \beta_{k} + t_{ij} - M\left(1-x^{k}_{ij}\right) & \hspace{-20mm} \forall i \in \mathcal{D}, j \in \delta^{+}(i), k \in K \\
-    \label{mod:c5} & w_{i} \geq v_{i}  + \frac{q^k_j}{\tau^l_b} -M\left(1- x^{k}_{ij}\right) &  \hspace{-20mm} \forall  b \in \mathcal{B},  i \in \mathcal{L}_{b},  j \in \delta^{+}(i), k \in K \\
-    \label{mod:c6} &          w_{j} \geq v_{j}  + {  \frac{q^k_j}{\tau^u_c} -M\left(1- x^{k}_{ij}\right) } & \hspace{-20mm}  \forall c \in  \mathcal{C}, j \in \mathcal{D}_{c}, i \in \delta^{-}(j),  k \in K \\
-    \label{mod:c7} &         w_{j} \leq v_{i}  + \Delta + M\left(1- x^{k}_{ij}\right) & \hspace{-20mm}   j \in \mathcal{D},  i \in \delta^{-}(j),  k \in K \\
-    \label{mod:c8}  &  v_{d^0_{o}} \geq a_i  & \hspace{-20mm} \forall  i \in \mathcal{C}, \forall o \in \mathcal{O}_i\\
-    \label{mod:c9} & g_{i} \geq v_{d^0_{o_1}} - a_i - M\left(  \sum_{ j \in \delta^{-}_{D}(d^0_{o_1})  }{u_{j,d^0_{o_1}}}\right)  & \hspace{-20mm}  \forall  i \in \mathcal{C}, \forall o_1 \in \mathcal{O}_i, d=d^0_{o_1} \\
-     \label{mod:c101} & v_{d^0_{o_1}} \geq w_j - M\left(1- u_{j,d^0_{o_1}}\right)  & \hspace{-40mm}    \forall o_1 \in \mathcal{O}_i,  j \in \delta_{D}^{-}(d^0_{o_1})  \\
-  \label{mod:c10} &  v_{d^0_{o_1}}  \leq  w_j + \gamma_i + M\left(1- u_{j,d^0_{o_1}}\right) &  \hspace{-20mm}  \forall  o_1 \in \mathcal{O},  j \in \delta_{D}^{-}(d^0_{o_1}) \\
-    \label{mod:c11} &  \sum_{ \substack{o_1 \in \mathcal{O}_i }}{ \sum_{ j \in \delta^{-}_{D}(d^0_{o_1})}}{u_{j,d^0_{o_1} } } = |\mathcal{O}_i|-1 & \hspace{-20mm} \forall  i \in \mathcal{C},|\mathcal{O}_i| > 1 \\
-    \label{mod:c12}&   \sum_{ \substack{o_1 \in \mathcal{O}_i }}\sum_{ j \in \delta^{+}_{D}(d^0_{o_1})}{u_{d^0_{o_1},j } } = |\mathcal{O}_i|-1 & \hspace{-40mm} \forall  i \in \mathcal{C}, |\mathcal{O}_i| > 1 \\
-    \label{mod:c13} & \hspace{3mm} \smashoperator{ \sum_{ j \in \delta^{+}_{D}(d^0_{o_1})}}{u_{d^0_{o_1},j } } \leq 1 & \hspace{-20mm}  \forall o_1 \in \mathcal{O} \\
-    \label{mod:c14} & \hspace{2mm} \smashoperator{ \sum_{ j \in \delta^{-}_{D}(d^0_{o_1})}}{u_{j,d^0_{o_1} } } \leq 1 & \hspace{-20mm} \forall o_1 \in \mathcal{O} \\
-    \label{mod:c15} & \hspace{2mm} \smashoperator{ \sum_{ j \in \delta^{+}_{D}(d^0_{o_1})}}{u_{d^0_{o_1},j } } +\smashoperator{ \sum_{ j \in \delta^{-}_{D}(d^0_{o_1})}}{u_{j,d^0_{o_1} } } \geq 1         & \hspace{-20mm} \forall o_1 \in \mathcal{O} \\
-    \label{mod:c16} & v_{j} \geq w_{j-1} - M\left(1- u_{j-1,j}\right) & \hspace{-20mm}  \forall o \in \mathcal{O}, j \in \mathcal{D}_{o}, j \geq 1 \\
-  \label{mod:c17} &  v_{j}  \leq  w_{j-1} + \gamma_i + M\left(1- u_{j-1,j}\right) & \hspace{-20mm} \forall i \in \mathcal{C}, \forall o \in O_i, j \in \mathcal{D}_{o}, j \geq 1 \\
-   \label{mod:c18}   &   u_{j-1,j} \geq u_{j,j+1} & \hspace{-35mm} \forall o_1 \in\mathcal{O},  j \in \mathcal{D}_{o_1}, 1 \leq j \leq n_{o_{1}}-1\\
-  \label{mod:c19} &  u_{j-1,j} \geq \sum_{l \in \mathcal{L}}{x_{lj}} &\hspace{-20mm}  \forall o_1 \in \mathcal{O}, j \in \mathcal{D}_{o_1}, j \geq 1 \\
-    \label{mod:c20} & v_{j} \geq w_{i} - M\left(1- u_{i,j}\right) & \hspace{-20mm}  \forall i \in \mathcal{L}, j \in \delta^{+}_{L}(i) \\
-   \label{mod:c21} & \hspace{3mm} \smashoperator{ \sum_{ j \in \delta^{+}_{L}(i)}}{u_{i,j } } \leq 1 & \hspace{-20mm}  \forall i \in \mathcal{L} \\
-    \label{mod:c22} & \hspace{2mm} \smashoperator{ \sum_{ j \in \delta^{-}_{L}(i)}}{u_{j,i } } \leq 1 & \hspace{-20mm} \forall i \in \mathcal{L} \\
-    \label{mod:c23} & \hspace{2mm} \smashoperator{ \sum_{ j \in \delta^{-}_{L}(i)}}{u_{j,i } } + \smashoperator{ \sum_{ j \in \delta^{+}_{L}(i)}}{u_{i,j } } \geq \sum_{k \in K}\sum_{j \in  \delta^{-}(i) }{x^k_{ji}} & \hspace{-20mm} \forall i \in \mathcal{L} \\
-    \label{mod:c24} & \sum_{k \in K}\sum_{j \in \mathcal{D}_o}{q^k_j} = q_o & \hspace{-20mm}  \forall o \in \mathcal{O} \\
-    \label{mod:c25} &  \sum_{k \in K}\sum_{j \in \delta^{+}(i)}{x^{k}_{ij} \leq 1} &  i \in \mathcal{L} \cup \mathcal{D} \\
-  \label{mod:c26}  &  \sum_{k \in K}\sum_{j \in \delta^{+}(i)} {x^{k}_{ij} } =  \sum_{k \in K}\sum_{j \in \delta^{-}(i)}{x^{k}_{ji} } &  i \in \mathcal{L} \cup \mathcal{D} \\
-  \label{mod:c27} & \sum_{k \in K}  \sum_{i \in \mathcal{D}_c} \sum_{j \in \delta^{-}(i) }  {x^k_{ji}} \leq M*\sigma_{c, b} & \forall c \in \mathcal{C}, b \in \mathcal{B} \\
-    \label{mod:c28} & \sum_{b \in \mathcal{B}} \sigma_{c, b} \leq 1 & \hspace{-20mm}\forall c \in \mathcal{C}, b \in \mathcal{B} \\
-    \label{mod:c29} & w^1_k \geq  M_T + H_k - v_{e_k}  & \hspace{-20mm} \forall k \in K \\
-    \label{mod:c30} & w^2_k \geq (v_{e_k} - H_k) - N_T & \hspace{-20mm} \forall k \in K \\
-    \label{mod:c31}  & w_{e_k} \leq H_k + O_T  & \hspace{-00mm}  k \in K \hspace{0mm} \\
-    \label{mod:c32} &  0 \leq q^k_j \leq Q^k  & \hspace{-20mm}  j \in \mathcal{D}, k \in K \\
-    \label{mod:c33} &  x^k_{ij} \in \{0,1\} & \hspace{-20mm}  (i,j) \in A , k \in K  \hspace{0mm} \\
-    \label{mod:c34} &  u_{ij} \in \{0,1\} & \hspace{-20mm}  (i,j) \in A^D , k \in K  \hspace{0mm} \\
-    \label{mod:c35} &  \sigma_{ib} \in \{0,1\} & \hspace{-20mm}  i \in \mathcal{C}, b \in \mathcal{B} \\
-    \label{mod:c36} &  y_o \in \{0,1\} & \hspace{-20mm}  o \in \mathcal{O} \\
-    \label{mod:c37} &  v_i \geq 0,  w_i \geq 0 & \hspace{-20mm}  i \in V \\\
-    \label{mod:c38} &  w^1_{k} \geq 0,  w^2_k \geq 0  & \hspace{-20mm}  i \in V \\
-    \label{mod:cend} &  g_i \geq 0   & \hspace{-20mm}  i \in \mathcal{C}.
-\end{alignat}
-
-$\beta_1$ to $\beta_4$ are the penalty coefficients of each component of the objective function (\ref{mod:obj}).
-Constraints (\ref{mod:c1}--\ref{mod:c2}) state that a driver $k$ leaves his start node exactly once a day and returns to his end node after completing his last unloading operation of the day.
-Constraints (\ref{mod:c3}--\ref{mod:c4}) set a driver to take some time after loading or unloading to adjust the concrete or clean the truck before moving on to the next node.
-The duration of the loading operation depends on the plant's loading rate and the amount $q^k_j$ of RMC loaded (Eq.
-\ref{mod:c5}).
-Similarly, the unloading service depends on the site's rate and $q^k_j$ (Eq.
-\ref{mod:c6}).
-Unloading operations must end $\Delta$ minutes after loading begins (Eq.
-\ref{mod:c7}).
-
-Constraints (\ref{mod:c8}--\ref{mod:c10}) ensure that the first service of any customer $i$ must start at the due time $a_i$.
-This first service may be performed at the first delivery node of any order $\in O_i$, and one order must be completed before another is started.
-It also enforces the precedence constraints between the last delivery of an order and the first delivery of the following order.
-Constraints (\ref{mod:c11}--\ref{mod:c15}) find the delivery sequence of all orders $\in \mathcal{O}_i$.
-Constraints (\ref{mod:c16}--\ref{mod:c19}) prohibit two trucks serving consecutive deliveries of the same order from unloading at the same time and enforce a maximum time delay between them.
-Similarly, constraints (\ref{mod:c20}--\ref{mod:c23}) ensure that two trucks cannot be loaded at the same time at a plant.
-Constraints (\ref{mod:c24}) require that the cumulative load of all concrete mixers serving an order must equal the required quantities.
-Constraints (\ref{mod:c25}) require that a driver can only visit a loading/delivery node once.
-Constraints (\ref{mod:c26}) are degree constraints.
-Constraints (\ref{mod:c27}) and (\ref{mod:c28}) require that all orders from customer $i$ must be loaded from the same plant.
-Constraints (\ref{mod:c29}--\ref{mod:c30}) calculate the difference between a driver's hours of service and the minimum and normal hours of service.
-Finally, constraints (\ref{mod:c31})--(\ref{mod:cend}) define the nature and bounds of the variables.
-
-% \paragraph*{}
-Let us consider the solution of a small instance of our problem shown in Figures~\ref{fig_Example} and \ref{fig:ganttExample}.
+\paragraph*{}
+Let us consider the solution of a small instance of our problem illustrated in Figures~\ref{fig_Example} and \ref{fig:ganttExample}.
 Two batching plants $B_1$ and $B_2$ with three drivers serve two construction sites.
@@
 Customer $C_2$ requests three orders of $q^1_2=3$, $q^2_2=11$, and $q^3_2=1$ m$^3$ of three different types of concrete.
-$B_1$ has two drivers ($D_1$, $D_2$) using concrete mixers of capacity 8 m$^3$, and $B_2$ has $D_3$ with a capacity 12 m$^3$.
-Figure~\ref{fig_Example} shows the flow of concrete mixers flow in the network.
-We schedule $D_1$ and $D_2$ to deliver 8 and 4 m$^3$ to $C_1$, respectively.
+$B_1$ has two drivers ($D_1$, $D_2$) using concrete mixers of capacity 8 m$^3$, and $B_2$ has $D_3$ of capacity 12 m$^3$.
+Figure~\ref{fig_Example} depicts the concrete mixers flow in the network.
+We schedule $D_1$ and $D_2$ to respectively deliver 8 and 4 m$^3$ to $C_1$.
 We select $o^2_2$ as the first order of $C_2$ to deliver.
@@
 The first deliveries of $C_1$ and $C_2$ are timely and there is no gap between the two deliveries received by $C_1$.
-There is a delay between the end of the first delivery and the start of the second delivery of $C_2$, however, it is less than the maximal time delay of 20 minutes defined for this example.
+There is a delay between the end of the first delivery and the start of the second delivery of $C_2$, however, it is less than the maximal time lag of 20 minutes defined for this example.
 
@@
 \section{Constructive heuristics and GRASP}
-\label{grasp_method}
+\label{method}
 
 We now present the heuristic approach we implement to solve this variant of the Concrete Delivery Problem.
-Our method consists of constructing feasible solutions to the CDP with randomized heuristics and iteratively invoking these heuristics in the GRASP metaheuristic.
+Our method consists in constructing feasible solutions for the CDP with randomized heuristics and iteratively calling these heuristics in the greedy randomized adaptive search procedure (GRASP) metaheuristic.
 
-In this section, we represent the scheduling of a delivery node $d$ as a pair of loading ($L_{d}$) and unloading ($U_{d}$) tasks.
-$L_{d} = \left(b,k,q_d,v_b\right) $ indicates that driver $k$ starts loading $q_d$ $m^3$ of RMC at plant $b$ at time $v_b$.
-$U_{d} = \left(k,v_d, w_d\right)$ indicates that driver $k$ is serving $d$ between $v_d$ and $w_d$.
-We then call a solution $ S =\left\lbrace \cup _{1 \leq j \leq n_o} (L_{d^j_{o}}, U_{d^j_{o}}), \forall o \in \mathcal{O} \right\rbrace$ the set of all loading and unloading tasks performed on a given day.
+In this section, we represent the scheduling of a delivery node $d$ as a pair of loading $L_{d}$ and unloading $U_{d}$ tasks.
+We then refer to a solution $S$ as a set of loading and unloading tasks performed by a set of drivers during a particular day.
+\begin{alignat*}{2}
+    L_{d} &= \left\lbrace (b,k,q_d,v_b), b \in \mathcal{B}, k \in K \right\rbrace \\
+    U_{d} & =  (k,v_d, w_d), k \in K  \} \\
+    S &=\left\lbrace \cup _{1 \leq j \leq n_o}  (L_{d^j_{o}},U_{d^j_{o}}), \forall o \in \mathcal{O} \right\rbrace
+\end{alignat*}
 
@@
 The GRASP algorithm is an iterative suite of constructive and local search algorithms introduced by \cite{feo1995greedy}.
-A feasible solution $S$ is constructed using a greedy randomized algorithm at each iteration.
-Then a local search algorithm explores the neighborhood of $S$ to find a local optimum.
-The pseudocode of the algorithm~\ref{alg_grasp} shows that the procedure returns the best overall solution (for a minimization problem) after some stopping conditions (time limit or a maximum number of iterations) are met.
-In our implementation, we introduce at line 13 an additional local search step using the solutions found by all heuristics in the current iteration.
+For each iteration, a feasible solution $S$ is built using a greedy randomized algorithm.
+Then a local search algorithm investigates the neighborhood of  $S$ to find a local optimum.
+The pseudo-code of Algorithm~\ref{alg_grasp} shows that the procedure returns the best overall solution (for a minimization problem) after some stopping conditions (time limit or a maximal number of iterations) are met.
 
@@
             \KwIn{ \textit{H}: List of constructive randomized heuristics }
-            $S^* \leftarrow \emptyset$    \hspace{2mm} $T \leftarrow \emptyset$       $Cost(S^*) \leftarrow \infty$
+            $S^* \leftarrow \emptyset$    \hspace{2mm}       $Cost(S^*) \leftarrow \infty$
 
@@
                 $Cost(S') \leftarrow \infty$
-
-                $T \leftarrow \emptyset$
 
@@
                     $S \leftarrow LocalSearch(S)$
-
-                    $T \leftarrow T \cup \{S\}$
 
@@
                 }
-                $S^* \leftarrow LocalSearch(T)$
             }
@@
 
-As mentioned in \cite{resende2019greedy}, the general outline of a greedy randomized algorithm used in the GRASP framework works as described in Algorithm~\ref{alg:greedyIns}.
-Let's consider the set $CL$ of candidate delivery nodes that have not yet been scheduled.
-We first initialize $CL$ with the first delivery node of a randomly selected order, for each customer.
-
-At each iteration, we create the set $CL'$ with elements of $CL$ from a given neighborhood.
-We then evaluate the increase in the cost function of incorporating each $d \in CL'$ into the incumbent solution $S$ and construct a restricted candidate list $RCL$ consisting of nodes whose incremental cost is less than a given threshold.
-From $RCL$, we randomly select the next delivery node to be incorporated into the incumbent solution and determine its loading and unloading schedule using the procedure described in Algorithm~\ref{alg_ScheduleTask}.
-Next, we update $S$ and the remaining order and customer quantities.
-Then we add the next candidate delivery node to $CL$.
+As referred to in \cite{resende2019greedy}, the general outline of a greedy randomized algorithm used in the GRASP framework works as described in Algorithm~\ref{alg:greedyIns}.
+At each iteration, let's consider the set $CL$ of candidate delivery nodes that are not yet scheduled.
+We note that, for our problem, not all delivery nodes are candidates for insertion at each iteration, since deliveries of the same order are ordered, any order of a customer may be chosen, and it must be satisfied before switching to another order, and the number of visits to a customer is not known beforehand.
+We first initialize $CL$ with the first delivery of a random order for each customer.
+We create the set $CL'$ with elements of $CL$ we deem promising for a better solution.
+We evaluate the increase in the cost function when incorporating each $d \in CL'$ into the incumbent solution $S$ and create a restricted candidate list $RCL$ formed by nodes whose incremental costs are less than a defined threshold.
+From $RCL$, we randomly select the next delivery node to be incorporated into the incumbent solution and we determine its loading and unloading schedule with the procedure described in Algorithm~\ref{alg_ScheduleTask}.
+Next, we update $S$ and the remaining quantities of the order and customer.
+Then we add to $CL$ the next delivery node candidate.
 If an order $o$ is not yet fulfilled after visiting $d^j_{o}$, the candidate is the next delivery node $d^{j+1}_{o}$.
 Otherwise, if the customer has remaining orders, the candidate is the first delivery node $d^{0}_{o'}$ of another randomly selected order $o'$.
-Note that for our problem, not all unscheduled delivery nodes are candidates at each iteration, since deliveries of the same order are ordered, any order of a customer can be chosen and must be satisfied before switching to another order, and the number of visits to a customer is not known in advance.
 
-Consider two customers whose services are required to start at the same time or in the same time slot.
-We could either finish scheduling one before the other, or schedule both at the same time.
-Obviously, each of these choices would result in a different solution.
-Similarly, in some cases, it would be better to schedule a customer that has a high demand and therefore requires more resources first.
-That's why we added the filtering part to impose some criteria on the algorithm.
-We can limit $CL'$ to delivery nodes with the same demand, the same due date, or from the same customer.
-We can also decide if we want to schedule customers with the highest (or lowest) demand or the earliest (latest) due time first.
-In fact, this helps us to intensify the search in specific neighborhoods of the solution.
-To diversify the search, we simply remove the filter component.
-The greedy aspect of GRASP is the creation of the set $RCL$, the probabilistic aspect is the random selection in $RCL$, and the adaptive aspect is the updating of $CL$ and the re-evaluation of the incremental costs.
-
-%Also, using $CL'$ helps reduce the complexity of the algorithm by reducing the number of nodes to evaluate.  %We thus obtain four greedy randomized insertion heuristics for our GRASP framework.
+The greedy aspect of the GRASP is the creation of the $RCL$ set, the probabilistic aspect is the random selection in $RCL$, and the adaptive aspect is the update of $CL$ and the reevaluating of the incremental costs.
+We use the additional set $CL'$ because we noticed that restricting $CL$ to delivery nodes with the same demand, delivery due time, or overlapping unloading timeslots helps us intensify the search.
+Using $CL'$ also helps improve the algorithm complexity by reducing the number of nodes to evaluate.
+To diversify the search we simply remove the filter component.
+We thus obtain four greedy randomized insertion heuristics for our GRASP framework.
 
@@
 
-            $RCL \leftarrow \left\lbrace d \in CL', \text{ } C(d) \leq C_{min} + \alpha (C_{max}-C_{min}) \right\rbrace $
+            $RCL \leftarrow \left\lbrace d \in CL, \text{ } C(d) \leq C_{min} + \alpha (C_{max}-C_{min}) \right\rbrace $
 
@@
  Algorithm~\ref{alg_ScheduleTask} takes as parameters a delivery node $d^j_{o}$ of the customer $i$ and a partial solution and looks for a plant and a driver available for loading and unloading operations.
-For each driver $k$ and plant $b$ we simulate the loading and unloading operations while checking the concrete lifespan and cold joint constraints, and the plant and driver availability.
+For each driver $k$ and plant $b$ we simulate the loading and unloading operations while checking the concrete lifespan constraints, and the plant and driver availability.
 For each node, we first determine the start loading time $SLT$ by subtracting $t_{bi}$, $\alpha_b$, and the loading duration $LD^b_{d^j_{o}}$ to the expected delivery time $EDT$.
@@
 
-\subsection{ Local Search}
-
-    % We implement two local search operators for the CDP. The first one is the relocate move commonly used for vehicle routing problems.
-
-   % At the end of each iteration, we expect each customer to have a feasible schedule respecting all constraints.
-For each customer, we combine each scheduling in order to construct a new and possibly better solution.
+\subsection*{ Local Search}
 
@@
 
-% In this section, we present and discuss the results obtained in the computational experiments of the GRASP heuristic described in Section~\ref{grasp_method}. The GRASP algorithm was coded in C++.
-% and the branch-and-bound (B\&B) algorithm applied to the model described in Section~\ref{model}.
-The different algorithms were coded in C++ and run on a processor Intel $2.4$ GHz Gold $6148$ Skylake.
-For the exact method, we have used IBM Concert Technology and CPLEX $12.10$ as MIP solver.
-A single thread was used with a time limit of one hour.
-% No benchmark instances for the \dirp exist so we have generated the dataset for the experiments as follows.
-
-\subsection{Generation of instances}
-% Due to the lack of publicly available benchmark data for the CDP, two data sets are created.
-Both sets, as well as code to generate new instances, are available online at [17].
-A summary of the instances is provided in Table 4.DataSetAcontains instances with 10–20 customers,and2–5 vehicles,whereasData Set B contains instances with up to 50 customers and 20 vehicles.
-The demands of customers are always divisible by 5 and are selected at random from the interval [10–75].
-In a similar fashion, the capacities of the vehicles are selected between [10–25].
-It is unrealistic to have a unique capacity for each vehicle, and therefore the number of different vehicles is bounded by the number of vehicle classes.Theprocessing time of a vehicle kAK is set proportional to its capacity, i.e.
-pk ¼qk for all kAK.
-The width of the time window of a customer i AC is computed by λiqi,whereλi is a scalar uniformly selected from the interval ½1:1; 2:1.
-Each instance has up to 4 production stations.
-The locations of the start and end depot, construction sites, and production depots, lay in the euclidean plane.
-The travel time between two locations equals the euclidean distance, rounded upwards.
-For simplicity, the start and end depot are the same.
-The distance between each station and the depot is uniformly selected from the interval [1–30].
-Furthermore, for each customer, there exists a production station within a distance of [1–25].Finally,thetimelagγ is fixed to 5 for all deliveries.
 \subsection{Data sets}
-
-\subsubsection{ Results for the generated instances}
-
-\subsubsection{ Results for the real-world datasets}
-\subsubsection{ Results for the instances of \cite{kinable2014concrete}}
-
-We run our algorithm on the dataset provided in \cite{kinable2014concrete}.
 

### 2024-03-09 ArticleCDP/original/versions/23_02_24/articleCDP.tex
Save

--- 2024-03-09 ArticleCDP/original/articleCDP.tex
@@
 \usepackage{array}
+\allowdisplaybreaks
 \usepackage{lineno}
@@
 \usepackage{tkz-graph}
-\usepackage{varwidth}
-\usepackage{url}
-\allowdisplaybreaks
+
 \newcolumntype{H}{>{\setbox0=\hbox\bgroup}c<{\egroup}@{}}
@@
 
-\title{A GRASP algorithm for the concrete delivery problem}
+\title{A GRASP algorithm for the concrete delivery problem. }
 \author{Ousmane Ali$^{(1)}$, Jean-Fran\c cois C\^ot\'e$^{(2)}$, Leandro C.~Coelho$^{(3)}$\\
@@
 
-\date{\today}
+% \date{\today}
 
@@
   This paper addresses a novel variant of the Concrete Delivery Problem (CDP), which involves the efficient scheduling of ready-mixed concrete deliveries to construction sites while balancing the conflicting goals of minimizing transportation costs and maximizing customer satisfaction.
-In this study, we propose an exact formulation and a heuristic approach based on the Greedy Randomized Adaptive Search Procedure (GRASP) to tackle this challenging CDP variant.
+In this study, we propose an exact formulation and a heuristic approach based on the Greedy Randomized Adaptive Search Procedure (GRASP), to tackle this challenging CDP variant.
 This variant introduces realistic side constraints, including driver working shifts, a minimum driver working time, and overtime penalties.
 Additionally, it considers the case where customers may request multiple types of concrete delivered within the same time window.
-We assess the performance of our heuristic using new instances generated for this problem and provide a comparative analysis with another CDP variant from the literature to demonstrate its effectiveness.
+We assess the performance of our heuristic using new instances generated for this problem and provide a comparative analysis with another CDP variant to demonstrate its effectiveness.
 \end{abstract}
@@
 On the other hand, RMC has better quality and benefits from lower production costs \citep{muresan_comparing}.
-However, to take advantage of these benefits, the batch plant manager must ensure efficient and prompt delivery to the construction site, which may require a fleet of high-cost revolving drum trucks (concrete mixers) to dispatch the RMC.
+However, to take advantage of these benefits, the batch plant manager must ensure efficient and prompt delivery on the construction site, which may require a fleet of high-cost revolving drum trucks (concrete mixers) to dispatch the RMC.
 
@@
 
-Academic research on concrete batch and delivery began in the late 1990s.
+Academic Research on concrete batch and delivery began in the late 1990s.
 \cite{tommelein1999just} described RMC as a prototypical example of a JIT production system in construction and identified two practices for delivering it.
@@
 \cite{feng2004optimizing} used a combination of genetic algorithm (GA) and simulation process to minimize the total waiting time for trucks at a customer site.
-The study focused on loading homogeneous trucks at the same batch plant, with fixed loading and unloading durations.
+The study focused on loading trucks with identical capacities at the same batch plant, with fixed loading and unloading durations.
 The GA was used to find the best loading sequence of RMC trucks to be assigned to different construction sites.
@@
 In addition to simulation-based methods, several other approaches have been used in the literature to solve the CDP.
-These include metaheuristics \citep{misir2011selection, maghrebi2016sequential, yang2022concrete}, exact methods \citep{asbach2009analysis, kinable2014concrete}, matheuristics \citep{schmid2009hybrid, schmid2010hybridization}, Benders decomposition \citep{maghrebi2014benders}, column generation (CG) \citep{maghrebi2014solving}, Lagrangian relaxation \citep{narayanan2015using}, and machine learning approaches \citep{graham2006modeling}.
+These include metaheuristics \citep{misir2011selection, maghrebi2016sequential, yang2022concrete}, exact methods \citep{asbach2009analysis, kinable2014concrete}, matheuristics \citep{schmid2009hybrid, schmid2010hybridization}, Benders Decomposition \citep{maghrebi2014benders}, column generation (CG) \citep{maghrebi2014solving, maghrebi2016column}, Lagrangian relaxation \citep{narayanan2015using}, and machine learning approaches \citep{graham2006modeling, maghrebi2016matching}.
 \cite{matsatsinis2004towards} designed a decision support system (DSS) for the dynamic routing of both concrete and pumps that may be necessary for some construction sites to aid in the unloading of concrete.
@@
 \cite{kinable2014concrete} introduce a general MIP and constraint programming (CP) models of the CDP reflecting the main constraints commonly found in all CDP works: time lag and no overlapping between consecutive deliveries, covering of all customers' demands, delivery time window, and heterogeneous fleet.
-However, the model did not include constraints limiting the time that concrete may stay on a truck.
+However, the model did not include constraints limiting the time that concrete may reside in a truck.
 The authors propose a constructive heuristic that schedules the visits to the customers one by one according to the start time of the visit and the truck capacity.
@@
 A driver $k$ is required to start his shift at $h_k$, work a minimum of $\mu$ hours and a maximum of $\eta$ hours during regular working hours, with the possibility of overtime of up to $\varLambda$ hours.
+$\beta_3$ and $\beta_4$ are the penalties for a driver working less than $\mu$ and more than $\eta$.
 
@@
 This means that a truck can only haul RMC for one order, even if there is spare capacity.
-After unloading the RMC, the driver takes $\rho$ minutes to clean the concrete mixer before proceeding.
+After unloading the RMC,  driver $k$ takes $\rho$ minutes to clean the concrete mixer before proceeding.
 
@@
 If an order requires more concrete than a single truck can carry, multiple deliveries are scheduled.
+We will accept a first delivery if it arrives within a time window of $\left[a_c, a_c + \tau^w_c \right]$, where $\tau^w_c$ is a user-defined parameter.
+Let $C$ be the set of customers, $O_c$ be the set of all orders requested by customer $c$, and $O=\{O_c, c \in C\}$ be the set of all requested orders for all customers.
+
+Each element of $O_c$ must be entirely delivered before moving on to another order.
+Exactly one order $o \in O_c$ must have its first delivery start between $\left[a_c, a_c + \tau^w_c \right]$, while the others can start at most $\gamma^2$ time after $o$ is completed.
+A plant is assigned to an order, and it must supply all of its subsequent deliveries.
 To avoid cold joint problems with the concrete, subsequent deliveries of the same order must be made in close succession.
 We define a maximum time delay $\gamma^1$, after which no more deliveries are allowed for the same order.
-Let $C$ be the set of customers, $O_c$ be the set of all orders requested by customer $c$, and $O=\{O_c, c \in C\}$ be the set of all requested orders for all customers.
-
-Each element of $O_c$ must be entirely delivered before moving on to another order.
-The first order $o \in O_c$ to complete must have its first delivery start between the time windows $\left[a_c, a_c + \tau^w_c \right]$, while the subsequent orders can start their deliveries at most $\gamma^2$ time units after their preceding orders are completed.
-Here, $\tau^w_c$ is a user-defined parameter representing the time window duration, and $\gamma^2$ is the maximum time delay between the completion of an order and the start of the next order.
-A plant is assigned to an order, and it must supply all of its subsequent deliveries.
 
@@
 $n_o$ is not known in advance because the trucks have different capacities.
-However, we can compute lower $n_o^{min}$ and upper $n_o^{max}$ bounds using the capacities of the largest $Q_{max}$ and smallest $Q_{min}$ available trucks.
+However, we can compute a lower $n_o^{min}$ and upper $n_o^{max}$ bounds using the capacities of the largest $Q_{max}$ and smallest $Q_{min}$ available trucks.
 \begin{alignat}{3}
@@
 
-We define our problem on a directed graph where $V=\{ S \cup L \cup D \cup E\}$ is the set of nodes.
+We define our problem on a complete directed graph where $V=\{ S \cup L \cup D \cup E\}$ is the set of nodes.
 The arc sets are $A =  \{(i,j) \hspace*{1mm} \vert \hspace*{1mm} i, j \in V \hspace{1mm}\}$, $A^D = \{(i,j) \hspace*{1mm} \vert \hspace*{1mm} i, j \in D \hspace{1mm} \}$, and $A^L = \{(i,j) \hspace*{1mm} \vert \hspace*{1mm} i, j \in L \hspace{1mm} \}$.
@@
 Constraints (\ref{mod:c8}) to (\ref{mod:c10}) ensure that the first service of any customer $c$ must start after the due time $a_c$.
-This first service may be performed at the first delivery node of any order, and one order must be completed before another is started.
+This first service may be performed at the first delivery node of any order $\in O_c$, and one order must be completed before another is started.
 It also enforces the precedence constraints between the last delivery of an order and the first delivery of the following order.
-Constraints (\ref{mod:c11}) to (\ref{mod:c15}) find the delivery sequence of all orders.
-With constraints (\ref{mod:c16})--(\ref{mod:c19}), two trucks cannot unload at the same time for consecutive deliveries of the same order.
+Constraints (\ref{mod:c11}) to (\ref{mod:c15}) find the delivery sequence of all orders $\in O_c$.
+With constraints (\ref{mod:c16}) -- (\ref{mod:c19}), two trucks cannot unload at the same time for consecutive deliveries of the same order.
 Additionally, these constraints impose a maximum time delay between the two trucks.
-Similarly, constraints (\ref{mod:c20})--(\ref{mod:c23}) ensure that two trucks cannot be loaded at the same time at a plant.
+Similarly, constraints (\ref{mod:c20}) -- (\ref{mod:c23}) ensure that two trucks cannot be loaded at the same time at a plant.
 Constraints (\ref{mod:c24}) require that the cumulative load of all concrete mixers serving an order must equal the required quantities, and (\ref{mod:c240}) bound a driver load.
@@
 Constraints (\ref{mod:c26}) are degree constraints.
-Constraints (\ref{mod:c29})--(\ref{mod:c30}) calculate the difference between a driver's hours of service and the minimum and normal hours of service.
+Constraints (\ref{mod:c29}) -- (\ref{mod:c30}) calculate the difference between a driver's hours of service and the minimum and normal hours of service.
 Finally, constraints (\ref{mod:c31}) to (\ref{mod:cend}) define the nature and bounds of the variables.
@@
     \centering
-    \caption{Information of the illustration instance}
+    \caption{Informations of the illustration instance}
     \label{tab:instance_detail}
@@
 From now on, we represent the scheduling of a delivery node $d$ as the pair ($L^k_{d}$,$U^k_{d}$) of loading and unloading tasks.
-$L^k_{d} = \left(b,q_d,v_b,w_b\right)$ indicates that driver $k$ will load $q_d$ $m^3$ of RMC at plant $b$ between timeslots $\left[v_b,w_b\right]$ to serve $d$.
+$L^k_{d} = \left(b,q_d,v_b,w_b\right)$ indicates that driver $k$ will load $q_d$ $m^3$ of RMC at plant $b$ between timeslot $\left[v_b,w_b\right]$ to serve $d$.
 $U^k_{d} = \left(k,v_d, w_d\right)$ indicates that driver $k$ will start unloading at time $v_d$ and finish at time $w_d$.
@@
 A feasible solution is generated in the construction phase using a greedy randomized algorithm.
-Then, a local search phase tries to improve it.
+Then, the local search phase tries to improve it.
 Algorithm~\ref{alg_grasp} outlines the GRASP algorithm.
@@
 In cases where no driver can visit the node $d$, the algorithm employs a backtracking scheme to address two distinct scenarios.
-First, if the visit to $d$ cannot be scheduled due to cold-joint constraints (i.e., the time between consecutive deliveries exceeds $\gamma^1$), the $DelayPreviousDelivery$ procedure is employed.
+First, if the visit to $d$ cannot be scheduled due to cold-joint constraints (i.e.
+the time between consecutive deliveries exceeds $\gamma^1$), the $DelayPreviousDelivery$ procedure is employed.
 This procedure determines if a previous delivery $D^{j'}_o$ of the same order exists such that $d$ can be successfully scheduled in the current solution by delaying the start of service of $D^{j'}_o$ by $\lambda$.
 If $D^{j'}_o$ exists, the schedules of all delivery nodes between $D^{j'}_o$ and $d$ are removed, and the scheduling of order $o$ resumes from $D^{j'}_o$.
+
+In cases where infeasibility arises from other violated constraints, the algorithm attempts to backtrack to a precedent node and retry the scheduling with another driver.
+At line $31$, the algorithm searches for the first precedent node $D^{j'}_o$ with remaining pair of tasks in $M[D^{j'}_o]$.
+If found, the schedules of all delivery nodes between $D^{j'}_o$ and $d$ are removed, and backtracking occurs to $D^{j'}_o$.
+In this iteration, since $M[D^{j'}_o]$ is not empty, the $SimulateVisit$ procedure is not called again; instead the stored elements of $M[D^{j'}_o]$ are utilized.
+
+If no precedent node suitable for backtracking is found, it indicates that the current customer cannot be scheduled.
+Consequently, all previously scheduled visits to its delivery nodes are removed before advancing to the next customer (as described in lines $23$--$25$ and lines $32$--$34$).
+The procedure stops when all customers are evaluated and returns the constructed solution $Sol$ at the end.
 
@@
 Then we backtrack to this node and restart the scheduling process.
-
-In cases where infeasibility arises from other violated constraints, the algorithm attempts to backtrack to a precedent node and retry the scheduling with another driver.
-At line $31$, the algorithm searches for the first precedent node $D^{j'}_o$ with remaining pair of tasks in $M[D^{j'}_o]$.
-If found, the schedules of all delivery nodes between $D^{j'}_o$ and $d$ are removed, and backtracking occurs to $D^{j'}_o$.
-In this iteration, since $M[D^{j'}_o]$ is not empty, the $SimulateVisit$ procedure is not called again; instead the stored elements of $M[D^{j'}_o]$ are utilized.
-
-If no precedent node suitable for backtracking is found, it indicates that the current customer cannot be scheduled.
-Consequently, all previously scheduled visits to its delivery nodes are removed before advancing to the next customer (as described in lines $23$--$25$ and lines $32$--$34$).
-The procedure stops when all customers are evaluated and returns the constructed solution $Sol$ at the end.
 
@@
 
-        $M \leftarrow \emptyset $ \tcp{List of pairs of loading -- unloading tasks for each delivery node}
+        $M \leftarrow \emptyset $ \tcp{List of pairs of loading and unloading tasks for each delivery node}
 
@@
 
-When confronted with scenarios where several customers require their services to start simultaneously or within the same time window, we observed that with Algorithm~\ref{alg:greedyCustIns}, some customers remain unscheduled when resources are limited.
+When confronted with scenarios where several customers require their services to start simultaneously or within the same time window, we observed that Algorithm~\ref{alg:greedyCustIns} may leave some customers remain unscheduled when resources are limited.
 To address this limitation, we introduce the delivery-based insertion heuristic.
@@
 
-    $M \leftarrow \emptyset $ \tcp{List of pairs of loading -- unloading tasks for each delivery node}
+    $M \leftarrow \emptyset $ \tcp{List of pairs of loading and unloading tasks for each delivery node}
 
@@
 \subsection{Generation of instances}
-\newcommand{\nbInstance}{35}
-The dataset used in this section was obtained from our partner and contains records representing delivery operations from up to 8 plants for $\nbInstance$ days.
+\newcommand{\nbInstance}{36}
+The dataset used in this section was obtained from our partner and contains records representing delivery operations from up to 8 production centers for $\nbInstance$ days.
 From this historical data, we extracted information such as daily orders delivered, plant assignments to orders, and the number of available drivers.
@@
 For each customer, we have the due time, the total demand, the number of orders (type of concrete) received, and its index in the time (distance) matrix.
-For each order, we have its demand and the corresponding customer and production plant.
+For each order, we have its demand and the corresponding production plant.
 Finally, we have the loading capacity for each plant, as well as its index in the time matrix.
-The name of an instance has the format $C\_n\_o\_k\_p$, where $n$ is the number of customers, $o$ is the total number of orders, $k$ is the number of available drivers, and $p$ is the number of plants.
+The name of an instance has the format $C\_n\_o\_k\_z$, where $n$ is the number of customers, $o$ is the total number of daily orders, $k$ is the number of available drivers, and $z$ is the number of plants.
 The dataset is divided into small, medium, and large sets according to the total number of daily orders.
 We have summarized the dataset in Table~\ref{tab:cdp_instances}.
-The small dataset has eight instances with four to eleven customers whose demands vary between 226.5 and 937.5 m$^3$.
+The small dataset has eight files with four to eleven customers whose demands vary between 226.5 and 937.5 m$^3$.
 These customers request between 7 and 14 orders, which are served from one to three depots with a fleet of 13 to 35 concrete mixers.
-The dataset is available online at \url{https://sites.google.com/view/jfcote/} as a contribution to the CDP literature.
+The dataset is available upon request as a contribution to the CDP literature.
 \begin{table}[htpb]
@@
     \label{tab:cdp_instances}
+    % \small
     \resizebox*{0.9\textwidth}{!}{
@@
             \toprule
-            {}     & \# & Demand (m$^3$)           & \#Orders   & \#Customers & \#Drivers & \#Depots \\
+            {}     & \# & Demand (m$^3$)           & \#Order   & \#Client & \#Driver & \#Depot \\
             \midrule
-            Small  & 8  & 226.5 -- 937.5     & 7 -- 14    & 4 -- 11   & 13 -- 35    & 1 -- 3     \\
-            Medium & 14 & 1,160.5 -- 2,971.0 & 43 -- 98   & 40 -- 89  & 76 -- 137   & 6 -- 8     \\
-            Large  & 13 & 3,078.5 -- 3,953.5  & 107 -- 136 & 92 -- 114 & 129 -- 149  & 8       \\
+            Small  & 8  & 226.5 - 937.5     & 7 - 14    & 4 - 11   & 13-35    & 1-3     \\
+            Medium & 14 & 1,160.5 - 2,971.0 & 43 - 98   & 40 - 89  & 76-137   & 6-8     \\
+            Large  & 14 & 3,078.5 - 3,953.5  & 107 - 136 & 92 - 127 & 129-150  & 8       \\
             \bottomrule
@@
 \cite{kinable2014concrete} solved their problem with a CP algorithm, MIP, steepest descent heuristic (SD), and fix-and-optimize MIP heuristic (MIP fix \& opt), which is a hybridization of the MIP and SD.
-They also provide dual bounds to help assess the optimality of the instances.
+They also provide upper bounds to help assess the optimality of the instances.
 Table \ref{tab:cdp_lit_comp} shows the performance of the GRASP heuristic against the methods mentioned above.
-For each method, the table shows the average execution time in seconds and the corresponding gap between the obtained solution and the dual bound provided by \cite{kinable2014concrete}.
-The row Opt reports the number of optimal solutions found by each method.
+For each method, the table shows the average execution time in seconds and the corresponding gap between the obtained solution and the upper bound provided by \cite{kinable2014concrete}.
+The row Opt reports the number of optimal solutions found by each methods.
 
@@
         \begin{tabular}{@{}cccccccccc|cc@{}}
-            \toprule
-            & &  \multicolumn{8}{c}{\cite{kinable2014concrete}} & \multicolumn{2}{|c}{Our paper} \\
-            \cmidrule(l){3-10} \cmidrule(r){11-12}
-             &     & \multicolumn{2}{c}{CP}     & \multicolumn{2}{c}{MIP}   & \multicolumn{2}{c}{SD-heuristic} & \multicolumn{2}{c}{MIP fix \& opt} & \multicolumn{2}{|c}{GRASP}                                                 \\ \cmidrule(l){3-10} \cmidrule(r){11-12}
+            \toprule &     & \multicolumn{2}{c}{CP}     & \multicolumn{2}{c}{MIP}   & \multicolumn{2}{c}{SD-heuristic} & \multicolumn{2}{c}{MIP fix \& opt} & \multicolumn{2}{|c}{GRASP}                                                 \\ \cmidrule(l){3-10} \cmidrule(r){11-12}
                &     & Gap (\%)                   & Time (s)                 & Gap (\%)                         & Time (s)                  & Gap (\%)                   & Time (s) & Gap (\%) & Time (s) & Gap (\%) & Time (s) \\ \cmidrule(l){3-10} \cmidrule(l){11-12}
@@
 
-Overall, our GRASP algorithm shows its effectiveness in providing good and competitive solutions in a relatively short time for this benchmark of the CDP, making it a suitable candidate for further analysis and real-world application.
+Overall, the GRASP algorithm shows its effectiveness in providing near-optimal solutions in a relatively short time for this benchmark of the CDP, making it a suitable candidate for further analysis and real-world application.
 It outperformed all methods except CP, achieving an average gap of 5.7\% (13.8\%) with a runtime of 13 (575) seconds for Set A (Set B).
@@
 All daily orders of the small instances are completely served by our algorithm within these five iterations, as shown in column $UQ$, which reports the undelivered quantities.
-For the medium and large instances, $99.5\%$ and $99.6\%$ of the total demands are delivered.
+For the medium and large instances, $99.57\%$ and $99.5\%$ of the total demands are delivered.
 The sum of driver underutilization costs ($DUC$) is high for medium and large instances.
@@
 The sum of driver overtime costs ($DOC$) is also high for these instances.
-This can be explained by the fact that the algorithm prioritizes drivers on overtime rather than using a driver who has to travel to another plant for a delivery, which would increase travel costs ($TC$).
+This can be explained by the fact that the algorithm prefers to put drivers on overtime rather than use a driver who has to travel to another plant for a delivery, which would increase travel costs ($TC$).
 
@@
 This information reveals that for small instances, the maximum delay ranges between 0 and 35.7 minutes.
-For medium instances, it varies from 32.7 to 60 minutes, and for large instances, it falls within the range of 56 to 60 minutes.
+For medium instances, it varies from 16.8 to 60 minutes, and for large instances, it falls within the range of 56 to 60 minutes.
 
@@
     \resizebox{\textwidth}{!}{
-    \begin{tabular}{lcccccrrrrrrrrrrrrr}
+    \begin{tabular}{lcccccccccccccccccc}
     \toprule
-     &  &  &  & & & \multicolumn{6}{c}{Avg} & \multicolumn{7}{c}{Best}  \\
+     &  &  &  &  & \multicolumn{6}{c}{Avg} & \multicolumn{6}{c}{Best}  \\
     \cmidrule(l){7-12} \cmidrule(l){13-19}
-     &  &  &  &  & \multicolumn{1}{c}{Demand}  & \multicolumn{1}{c}{$TC$}& \multicolumn{1}{c}{$UQ$} & \multicolumn{1}{c}{$FDD$} & \multicolumn{1}{c}{$DUC$} & \multicolumn{1}{c}{$DOC$} & \multicolumn{1}{c}{\multirow[c]{2}{*}{Z}}  & \multicolumn{1}{c}{($TC$)} & \multicolumn{1}{c}{$UQ$} & \multicolumn{1}{c}{$FDD$} & \multicolumn{1}{c}{$mFDD$} & \multicolumn{1}{c}{$DUC$} & \multicolumn{1}{c}{$DOC$} & \multicolumn{1}{c}{\multirow[c]{2}{*}{Z}}  \\
-      & |$C$| & $|O|$ & $|K|$ & $|B|$ & \multicolumn{1}{c}{(m$^3$)}& \multicolumn{1}{c}{(min)} & \multicolumn{1}{c}{(m$^3$)} & \multicolumn{1}{c}{(min)} & \multicolumn{1}{c}{(min)} & \multicolumn{1}{c}{(min)} &  & \multicolumn{1}{c}{(min)} & \multicolumn{1}{c}{(m$^3$)} & \multicolumn{1}{c}{(min)} & \multicolumn{1}{c}{(min)} & \multicolumn{1}{c}{(min)} & \multicolumn{1}{c}{(min)} &  \\
+     &  &  &  &  & Demand & $TC$ & $UQ$ & $FDD$ & $DUC$ & $DOC$ & \multirow[c]{2}{*}{Z}  & $TC$ & $UQ$ & $FDD$ & $mFDD$ & $DUC$ & $DOC$ & \multirow[c]{2}{*}{Z}  \\
+      & |$C$| & $|O|$ & $|K|$ & $|B|$ & (m$^3$) & (min) & (m$^3$) & (min) & (min) & (min) &  & (min) & (m$^3$) & (min) & (min) & (min) & (min) &  \\
     \midrule
-    \multirow[c]{8}{*}{Small} & 11 & 12 & 13 & 1 & 226.5 & 1,243.5 & 0.0 & 0.0 & 18.8 & 0.0 & 1,807.7 & 1,243.5 & 0.0 & 0.0 & 0.0 & 0.0 & 0.0 & 1,243.5 \\
+    \multirow[c]{8}{*}{Small} & 11 & 12 & 13 & 1 & 226.5 & 1,243.5 & 0.0 & 0.0 & 10.8 & 0.0 & 1,568.3 & 1,243.5 & 0.0 & 0.0 & 0.0 & 0.0 & 0.0 & 1,243.5 \\
      & 5 & 8 & 13 & 1 & 267.0 & 745.1 & 0.0 & 0.0 & 0.0 & 0.0 & 745.1 & 745.1 & 0.0 & 0.0 & 0.0 & 0.0 & 0.0 & 745.1 \\
-    & 6 & 11 & 18 & 2 & 333.5 & 753.3 & 0.0 & 0.0 & 255.9 & 0.0 & 8,430.6 & 753.3 & 0.0 & 0.0 & 0.0 & 176.0 & 0.0 & 6,034.2 \\
-    & 4 & 7 & 15 & 2 & 375.0 & 1,017.5 & 0.0 & 0.0 & 218.2 & 131.3 & 10,190.9 & 1,017.6 & 0.0 & 0.0 & 0.0 & 82.8 & 33.9 & 4,179.8 \\
-    & 7 & 8 & 19 & 2 & 388.0 & 1,790.5 & 0.0 & 0.0 & 318.4 & 137.8 & 14,119.2 & 1,817.0 & 0.0 & 0.0 & 0.0 & 52.6 & 98.7 & 5,368.6 \\
-    & 10 & 14 & 29 & 3 & 613.5 & 3,482.1 & 0.0 & 59.5 & 239.4 & 373.5 & 77,635.1 & 3,513.2 & 0.0 & 59.5 & 35.7 & 147.2 & 304.0 & 73,510.2 \\
-    & 8 & 11 & 31 & 3 & 776.0 & 2,325.7 & 0.0 & 0.0 & 796.3 & 690.1 & 40,015.9 & 2,311.7 & 0.0 & 0.0 & 0.0 & 593.7 & 601.0 & 32,141.2 \\
-    & 9 & 10 & 35 & 3 & 937.5 & 4,110.9 & 0.0 & 0.0 & 447.7 & 712.0 & 31,782.1 & 4,139.9 & 0.0 & 0.0 & 0.0 & 243.3 & 550.0 & 22,439.9 \\
+     & 6 & 11 & 18 & 2 & 333.5 & 753.3 & 0.0 & 0.0 & 145.9 & 0.0 & 5,130.0 & 753.3 & 0.0 & 0.0 & 0.0 & 118.7 & 0.0 & 4,315.1 \\
+     & 4 & 7 & 15 & 2 & 375.0 & 1,015.6 & 0.0 & 0.0 & 148.2 & 150.5 & 8,470.9 & 1,015.4 & 0.0 & 0.0 & 0.0 & 123.4 & 110.3 & 6,924.5 \\
+     & 7 & 8 & 19 & 2 & 388.0 & 1,747.2 & 0.0 & 0.0 & 349.8 & 159.0 & 15,421.6 & 1,781.8 & 0.0 & 0.0 & 0.0 & 186.0 & 0.0 & 7,361.5 \\
+     & 10 & 14 & 29 & 3 & 613.5 & 3,510.1 & 0.0 & 59.5 & 194.6 & 331.9 & 75,488.1 & 3,493.4 & 0.0 & 59.5 & 35.7 & 120.5 & 60.0 & 67,810.4 \\
+     & 8 & 11 & 31 & 3 & 776.0 & 2,314.8 & 0.0 & 0.0 & 804.3 & 640.7 & 39,256.8 & 2,278.2 & 0.0 & 0.0 & 0.0 & 503.5 & 617.0 & 29,722.9 \\
+     & 9 & 10 & 35 & 3 & 937.5 & 4,175.5 & 0.0 & 0.0 & 222.1 & 669.3 & 24,224.7 & 4,163.1 & 0.0 & 0.0 & 0.0 & 243.3 & 570.0 & 22,863.2 \\
      \cmidrule(l){1-12} \cmidrule(l){13-19}
-     \multicolumn{5}{r}{Average} & \multicolumn{1}{c}{-} & 1,933.6 & 0.0 & 7.4 & 286.8 & 255.6 & 23,090.8& 1942.7&0.0&7.4&4.5&161.9&198.5&18,207.8\\
+     \multicolumn{5}{r}{Average} & - & 1,938.1 & 0.0 & 7.4 & 234.4 & 243.9 & 21,288.2 & 1,934.2 & 0.0 & 7.4 & 4.5 & 161.9 & 169.6 & 17,623.3 \\
      \cmidrule(l){1-12} \cmidrule(l){13-19}
-     \multirow[c]{14}{*}{Medium} & 40 & 43 & 76 & 6 & 1,160.5 & 6,085.6 & 0.0 & 44.8 & 3,017.6 & 1,140.3 & 164,186.8 & 6,451.5 & 0.0 & 32.7 & 8.5 & 2,714.0 & 899.1 & 138,530.5 \\
- & 42 & 47 & 104 & 8 & 1,565.0 & 8,024.3 & 12.0 & 55.5 & 3,563.0 & 160.2 & 1,373,593.7 & 8,106.6 & 12.0 & 54.7 & 16.8 & 3,358.8 & 138.7 & 1,366,363.9 \\
- & 63 & 70 & 94 & 7 & 1,746.5 & 13,871.4 & 0.0 & 228.2 & 1,250.1 & 1,398.4 & 307,504.3 & 13,902.5 & 0.0 & 195.3 & 58.9 & 1,465.1 & 1,206.7 & 277,317.6 \\
- & 67 & 78 & 116 & 8 & 1,839.5 & 13,225.9 & 0.0 & 142.1 & 1,931.4 & 751.0 & 228,250.5 & 13,460.9 & 0.0 & 139.3 & 33.8 & 935.5 & 962.9 & 200,057.2 \\
- & 71 & 82 & 116 & 8 & 2,060.0 & 11,158.1 & 9.0 & 263.3 & 2,797.8 & 982.1 & 1,277,996.2 & 10,748.5 & 9.0 & 243.3 & 45.9 & 2,218.9 & 594.8 & 1,232,488.4 \\
- & 57 & 70 & 117 & 6 & 2,327.5 & 14,551.7 & 0.0 & 322.2 & 1,654.3 & 1,566.4 & 417,729.4 & 14,732.0 & 0.0 & 252.5 & 49.5 & 1,438.4 & 1,551.1 & 341,397.0 \\
- & 79 & 83 & 137 & 7 & 2,425.0 & 17,886.5 & 0.0 & 663.2 & 2,132.2 & 2,031.5 & 785,718.4 & 17,555.1 & 0.0 & 559.6 & 58.6 & 1,707.6 & 2,159.2 & 671,538.1 \\
- & 66 & 80 & 127 & 8 & 2,512.5 & 15,108.4 & 0.0 & 118.4 & 1,600.3 & 1,596.7 & 213,496.4 & 14,567.4 & 0.0 & 114.1 & 34.3 & 1,349.9 & 1,143.2 & 191,985.5 \\
- & 68 & 74 & 128 & 7 & 2,595.0 & 18,744.1 & 1.6 & 547.5 & 1,238.4 & 2,827.7 & 822,423.5 & 18,276.6 & 0.0 & 355.6 & 55.8 & 1,206.4 & 3,061.4 & 471,318.0 \\
- & 85 & 97 & 136 & 8 & 2,673.0 & 17,252.5 & 0.0 & 427.7 & 2,085.1 & 2,355.1 & 554,640.8 & 18,164.5 & 0.0 & 361.0 & 53.3 & 1,447.8 & 2,081.4 & 464,253.2 \\
- & 78 & 85 & 128 & 8 & 2,685.5 & 15,931.9 & 63.0 & 461.4 & 1,243.9 & 2,783.1 & 6,870,259.9 & 16,164.5 & 63.0 & 391.1 & 57.8 & 1,553.3 & 2,371.0 & 6,801,308.3 \\
- & 77 & 85 & 131 & 8 & 2,893.5 & 17,168.9 & 0.0 & 410.7 & 1,360.1 & 1,756.7 & 503,850.9 & 17,620.2 & 0.0 & 388.1 & 57.5 & 963.2 & 2,149.1 & 477,607.4 \\
- & 89 & 97 & 137 & 7 & 2,939.5 & 21,350.0 & 13.4 & 865.0 & 1,074.2 & 3,247.2 & 2,327,223.6 & 21,551.0 & 0.0 & 880.5 & 47.7 & 1,318.9 & 3,375.7 & 1,009,098.9 \\
- & 84 & 98 & 133 & 7 & 2,971.0 & 20,620.2 & 39.2 & 1,091.8 & 736.8 & 2,666.4 & 5,106,602.3 & 20,578.3 & 31.5 & 887.0 & 59.1 & 839.1 & 2,910.7 & 4,141,019.3 \\
+     \multirow[c]{14}{*}{Medium} & 40 & 43 & 76 & 6 & 1,160.5 & 6,055.8 & 0.0 & 49.1 & 3,286.4 & 990.0 & 173,509.5 & 5,990.6 & 0.0 & 33.9 & 25.4 & 3,061.3 & 788.6 & 147,510.1 \\
+     & 42 & 47 & 104 & 8 & 1,565.0 & 8,107.1 & 12.0 & 53.3 & 3,637.1 & 119.4 & 1,372,936.3 & 8,301.9 & 12.0 & 54.3 & 16.8 & 3,200.5 & 140.1 & 1,361,393.1 \\
+     & 63 & 70 & 94 & 7 & 1,746.5 & 13,783.0 & 0.0 & 238.7 & 983.2 & 1,554.6 & 313,077.0 & 13,673.5 & 0.0 & 195.9 & 59.7 & 1,025.2 & 1,407.3 & 268,524.2 \\
+     & 67 & 78 & 116 & 8 & 1,839.5 & 13,456.8 & 0.0 & 132.7 & 1,869.0 & 614.3 & 214,501.3 & 13,484.8 & 0.0 & 98.0 & 22.7 & 1,792.8 & 622.4 & 177,710.8 \\
+     & 71 & 82 & 116 & 8 & 2,060.0 & 11,393.8 & 9.0 & 265.1 & 2,714.8 & 1,054.6 & 1,279,068.9 & 10,793.3 & 9.0 & 242.0 & 41.8 & 3,197.9 & 665.2 & 1,262,000.5 \\
+     & 57 & 70 & 117 & 6 & 2,327.5 & 14,975.0 & 0.0 & 324.1 & 1,005.5 & 1,149.6 & 392,281.0 & 15,303.5 & 0.0 & 295.2 & 54.6 & 659.4 & 1,083.8 & 351,940.4 \\
+     & 79 & 83 & 137 & 7 & 2,425.0 & 17,644.3 & 0.0 & 553.8 & 2,320.4 & 1,993.3 & 680,888.1 & 17,379.3 & 0.0 & 474.0 & 55.7 & 2,964.4 & 2,016.3 & 620,601.6 \\
+     & 66 & 80 & 127 & 8 & 2,512.5 & 14,584.3 & 0.0 & 128.3 & 1,600.0 & 1,451.5 & 219,949.8 & 14,499.2 & 0.0 & 127.5 & 47.3 & 1,390.2 & 1,212.8 & 207,962.2 \\
+     & 68 & 74 & 128 & 7 & 2,595.0 & 18,450.6 & 3.7 & 566.6 & 1,317.5 & 2,706.2 & 1,048,712.1 & 18,261.4 & 0.0 & 473.4 & 56.0 & 1,764.7 & 2,525.6 & 595,143.5 \\
+     & 85 & 97 & 136 & 8 & 2,673.0 & 16,831.6 & 0.0 & 379.9 & 2,702.4 & 2,368.0 & 525,199.5 & 17,241.3 & 0.0 & 323.7 & 50.2 & 2,190.5 & 2,767.1 & 461,973.3 \\
+     & 78 & 85 & 128 & 8 & 2,685.5 & 15,679.8 & 63.0 & 474.5 & 1,399.6 & 2,394.4 & 6,880,077.6 & 15,830.6 & 63.0 & 432.7 & 57.8 & 1,206.8 & 2,272.7 & 6,830,154.3 \\
+     & 77 & 85 & 131 & 8 & 2,893.5 & 16,988.0 & 0.0 & 374.5 & 1,557.9 & 1,996.1 & 478,106.8 & 16,565.6 & 0.0 & 352.1 & 44.6 & 1,624.3 & 1,322.5 & 443,865.0 \\
+     & 89 & 97 & 137 & 7 & 2,939.5 & 21,716.4 & 9.8 & 924.4 & 821.0 & 3,494.4 & 2,020,622.6 & 21,311.3 & 0.0 & 799.5 & 51.8 & 887.3 & 3,486.7 & 917,162.4 \\
+     & 84 & 98 & 133 & 7 & 2,971.0 & 20,435.5 & 46.1 & 1,141.3 & 894.1 & 2,127.8 & 5,841,090.1 & 20,520.4 & 9.0 & 1,411.0 & 59.5 & 705.0 & 1,871.9 & 2,390,157.6 \\
      \cmidrule(l){1-12} \cmidrule(l){13-19}
-     \multicolumn{5}{r}{Average} & \multicolumn{1}{c}{-} & 15,070.0 & 9.9 & 403.0 & 1,834.7 & 1,804.5 & 1,496,676.9& 15,134.3&8.2&346.8&45.5&1,608.4&1,757.5&1,270,305.9\\
+     \multicolumn{5}{r}{Average} & - & 15,007.3 & 10.2 & 400.4 & 1,864.9 & 1,715.3 & 1,531,430 & 14,939.8 & 6.6 & 379.5 & 46.0 & 1,833.6 & 1,584.5 & 1,145,436 \\
     \cmidrule(l){1-12} \cmidrule(l){13-19}
-    \multirow[c]{14}{*}{Large} & 98 & 109 & 132 & 8 & 3,078.5 & 22,498.9 & 0.0 & 456.2 & 457.4 & 3,961.9 & 571,626.0 & 22,359.3 & 0.0 & 394.4 & 56.2 & 463.1 & 3,905.6 & 508,725.2 \\
- & 101 & 119 & 129 & 8 & 3,229.0 & 17,814.1 & 9.6 & 647.3 & 1,342.6 & 2,674.1 & 1,721,369.9 & 18,811.0 & 0.0 & 608.8 & 57.5 & 969.3 & 2,715.1 & 710,970.8 \\
- & 114 & 136 & 141 & 8 & 3,350.5 & 20,596.8 & 44.7 & 1,346.5 & 1,048.3 & 2,672.2 & 5,920,713.6 & 20,811.8 & 33.5 & 1,453.0 & 56.6 & 616.4 & 1,788.8 & 4,878,101.5 \\
- &114  & 132 & 140 & 8 & 3,401.5 & 22,016.2 & 27.7 & 1,320.5 & 810.1 & 3,021.1 & 4,196,040.1 & 22,693.0 & 14.5 & 1,487.4 & 58.6 & 525.9 & 2,616.0 & 3,028,191.3 \\
- & 101 & 123 & 143 & 8 & 3,437.5 & 23,804.3 & 0.0 & 649.5 & 942.7 & 4,558.3 & 792,752.6 & 22,514.1 & 0.0 & 447.0 & 59.1 & 1,561.5 & 3,605.0 & 588,441.4 \\
- & 112 & 129 & 137 & 8 & 3,471.0 & 21,873.0 & 0.0 & 520.3 & 613.1 & 3,935.4 & 639,262.4 & 21,352.6 & 0.0 & 473.6 & 52.4 & 958.4 & 3,604.6 & 595,823.5 \\
- & 114 & 129 & 142 & 8 & 3,499.5 & 21,173.3 & 28.1 & 1,119.6 & 994.0 & 4,116.9 & 4,065,444.5 & 21,178.8 & 8.0 & 1,042.9 & 58.1 & 757.2 & 4,141.9 & 1,969,653.9 \\
- & 98 & 122 & 149 & 8 & 3,513.0 & 19,739.2 & 35.7 & 1,141.4 & 1,842.7 & 2,570.9 & 4,836,612.0 & 19,413.8 & 18.5 & 1,233.5 & 59.2 & 2,192.8 & 2,399.4 & 3,216,666.5 \\
- &98  & 108 & 139 & 8 & 3,541.0 & 20,862.3 & 0.0 & 941.6 & 628.4 & 3,526.0 & 1,051,844.1 & 20,241.0 & 0.0 & 847.2 & 59.1 & 1,106.1 & 2,790.2 & 956,432.6 \\
- & 108 & 122 & 144 & 8 & 3,670.5 & 22,605.8 & 2.1 & 1,127.8 & 877.3 & 2,790.1 & 1,445,018.0 & 22,369.7 & 0.0 & 873.3 & 56.0 & 1,126.8 & 2,922.1 & 987,898.0 \\
- & 92 & 107 & 142 & 8 & 3,684.5 & 22,877.9 & 0.0 & 584.4 & 638.0 & 4,754.3 & 721,474.4 & 23,511.5 & 0.0 & 458.5 & 53.4 & 593.9 & 4,715.3 & 594,093.8 \\
- & 114 & 136 & 138 & 8 & 3,739.5 & 23,361.3 & 8.5 & 1,178.1 & 853.7 & 5,271.4 & 2,182,537.5 & 23,633.6 & 0.0 & 961.3 & 56.1 & 766.6 & 5,434.9 & 1,116,585.2 \\
- & 112 & 131 & 148 & 8 & 3,953.5 & 24,313.5 & 34.9 & 1,251.9 & 524.6 & 2,961.1 & 4,838,688.0 & 25,301.2 & 13.5 & 1,694.5 & 59.4 & 766.0 & 2,944.6 & 3,151,661.8 \\
+    \multirow[c]{14}{*}{Large} & 98 & 109 & 132 & 8 & 3,078.5 & 22,994.6 & 0.0 & 465.5 & 496.5 & 3,804.8 & 579,471.0 & 22,712.7 & 0.0 & 351.5 & 56.2 & 803.2 & 4,011.3 & 478,549.4 \\
+     & 101 & 119 & 129 & 8 & 3,229.0 & 17,323.1 & 11.0 & 657.8 & 1,577.6 & 2,375.9 & 1,869,937.5 & 17,311.8 & 11.0 & 598.8 & 58.4 & 1,507.7 & 1,948.1 & 1,800,315.9 \\
+     & 114 & 136 & 141 & 8 & 3,350.5 & 19,891.6 & 39.1 & 1,255.8 & 1,366.2 & 2,251.9 & 5,271,698.5 & 19,556.2 & 18.0 & 1,366.7 & 57.9 & 1,380.0 & 2,695.1 & 3,281,568.6 \\
+     &  114& 132 & 140 & 8 & 3,401.5 & 21,917.6 & 23.8 & 1,254.1 & 680.1 & 3,389.9 & 3,744,172.5 & 22,590.2 & 17.5 & 1,248.5 & 58.7 & 766.3 & 3,267.6 & 3,109,441.2 \\
+     & 101 & 123 & 143 & 8 & 3,437.5 & 23,477.2 & 0.0 & 677.7 & 955.5 & 4,369.4 & 817,244.5 & 23,806.9 & 0.0 & 617.3 & 56.0 & 1,302.2 & 5,147.9 & 783,123.1 \\
+     & 112 & 129 & 137 & 8 & 3,471.0 & 22,089.2 & 0.0 & 546.6 & 544.8 & 4,187.3 & 668,737.0 & 21,195.2 & 0.0 & 456.6 & 55.0 & 883.9 & 3,532.1 & 574,935.0 \\
+     & 114 & 129 & 142 & 8 & 3,499.5 & 20,678.2 & 35.7 & 1,007.5 & 1,119.7 & 4,133.1 & 4,714,436.9 & 21,243.7 & 21.0 & 907.0 & 59.8 & 1,038.0 & 4,016.6 & 3,139,750.0 \\
+     & 98 & 122 & 149 & 8 & 3,513.0 & 19,859.0 & 28.4 & 1,141.7 & 2,154.6 & 2,649.9 & 4,119,224.3 & 19,602.8 & 21.0 & 859.2 & 55.7 & 2,342.9 & 2,728.2 & 3,103,676.6 \\
+     & 98 & 108 & 139 & 8 & 3,541.0 & 21,060.5 & 0.0 & 966.2 & 684.9 & 3,585.4 & 1,079,477.1 & 21,084.8 & 0.0 & 867.2 & 59.8 & 834.9 & 4,133.9 & 996,046.7 \\
+     & 108 & 122 & 144 & 8 & 3,670.5 & 23,330.4 & 2.0 & 1,123.4 & 536.2 & 3,251.6 & 1,427,866.8 & 22,882.8 & 0.0 & 969.7 & 57.8 & 649.7 & 3,587.1 & 1,083,850.9 \\
+     & 92 & 107 & 142 & 8 & 3,684.5 & 23,179.0 & 0.0 & 578.3 & 384.1 & 4,260.4 & 698,184.4 & 23,610.6 & 0.0 & 449.2 & 59.8 & 315.7 & 4,257.2 & 567,403.7 \\
+     & 114 & 136 & 138 & 8 & 3,739.5 & 23,614.6 & 11.7 & 1,225.1 & 837.4 & 5,668.6 & 2,557,166.5 & 22,859.0 & 7.5 & 1,107.0 & 58.3 & 1,224.2 & 5,078.2 & 2,018,119.1 \\
+     & 127 & 136 & 150 & 8 & 3,946.5 & 24,940.3 & 0.0 & 949.5 & 775.5 & 5,384.3 & 1,105,391.3 & 24,939.2 & 0.0 & 829.0 & 57.1 & 805.9 & 6,166.7 & 1,001,490.6 \\
+     & 112 & 131 & 148 & 8 & 3,953.5 & 24,286.0 & 34.0 & 1,247.8 & 657.0 & 3,009.5 & 4,752,026.9 & 24,221.7 & 24.5 & 1,118.5 & 57.1 & 859.7 & 3,483.3 & 3,688,169.1 \\
      \cmidrule(l){1-12} \cmidrule(l){13-19}
-     \multicolumn{5}{r}{Average} &\multicolumn{1}{c}{-} & 21,810.5 & 14.7 & 945.0 & 890.2 & 3,601.0 & 2,537,183.3& 21,860.9&6.8&921.2&57.1&954.2&3,352.6&1,715,634.3 \\
+     \multicolumn{5}{r}{Average} &- & 22,045.8&        13.3& 935.5&      912.1&   3,737.3 & 2,386,074 & 21,972.7 & 8.6 & 839.0 & 57.7 & 1,051.0&3,860.9 & 1,830,460  \\
     \bottomrule
@@
     }
-    \begin{minipage}{12cm}
-    \vspace{0.1cm}
-    {\footnotesize $TC$: travel cost, $UQ$: undelivered quantity, $FDD$: first delivery delay, $mFDD$: maximum of first delivery delays, $DUC$: driver underutilization cost, $DOC$: driver overtime cost.}
-    \end{minipage}
 \end{table}
@@
 Table \ref{tab_cdp_runtime} shows that a higher percentage of deliveries are completed as the algorithm runtime increases.
-For small instances, all requests are scheduled in less than a minute of runtime, while for medium and large instances, 99\% of requests are scheduled to be delivered with only a minute of runtime.
+For small instances, all requests are delivered in less than a minute, while for medium and large instances, 99\% of requests are delivered.
 
@@
     \midrule
-    Small & \%Load & 100.0 & 100.0 & 100.0 & 100.0 & 100.0 \\
-Medium & \%Load & 98.9 & 99.3 & 99.4 & 99.5 & 99.5 \\
-Large & \%Load & 98.8 & 99.3 & 99.4 & 99.6 & 99.6 \\
+    Small & \%Load & 100.00 & 100.00 & 100.00 & 100.00 & 100.00 \\
+    Medium & \%Load & 99.00 & 99.29 & 99.29 & 99.43 & 99.57 \\
+    Large & \%Load & 99.00 & 99.21 & 99.36 & 99.43 & 99.50 \\
     \bottomrule
@@
 All orders are now fully serviced on each run, as shown in Table~\ref{tab:cdp_all_instances_no_shift}.
-Table~\ref{tab:cdp_lift_shift_constraints} compares the averages of the results from Tables \ref{tab:cdp_all_instances} and \ref{tab:cdp_all_instances_no_shift}.
+Table~\ref{tab:lift_shift_constraints} compares the averages of the results from Tables \ref{tab:cdp_all_instances} and \ref{tab:cdp_all_instances_no_shift}.
 We found that the maximum delay dropped to zero for small instances, and for medium (large) instances, it was reduced by almost 60\% (40\%) on average.
@@
     \resizebox{\textwidth}{!}{
-    \begin{tabular}{lcccccrrrrrrrrr}
+    \begin{tabular}{lcccccccccccccc}
     \toprule
-    &  &  &  & & & \multicolumn{4}{c}{Avg} & \multicolumn{5}{c}{Best}  \\
+    &  &  &  &  & \multicolumn{4}{c}{Avg} & \multicolumn{5}{c}{Best}  \\
      \cmidrule(l){7-10} \cmidrule(l){11-15}
-     &  &  &  &  & Demand & \multicolumn{1}{c}{$TC$}& \multicolumn{1}{c}{$UQ$} & \multicolumn{1}{c}{$FDD$} & \multicolumn{1}{c}{\multirow[c]{2}{*}{Z}} &  \multicolumn{1}{c}{$TC$}& \multicolumn{1}{c}{$UQ$} & \multicolumn{1}{c}{$FDD$}& \multicolumn{1}{c}{$mFDD$} & \multicolumn{1}{c}{\multirow[c]{2}{*}{Z}} \\
-     & |$C$| & $|O|$ & $|K|$ & $|B|$ & (m$^3$) & \multicolumn{1}{c}{(min)} & \multicolumn{1}{c}{(m$^3$)} &\multicolumn{1}{c}{(min)} &  & \multicolumn{1}{c}{(min)} & \multicolumn{1}{c}{(m$^3$)} & \multicolumn{1}{c}{(min)} & \multicolumn{1}{c}{(min)}&  \\
+     &  &  &  &  & Demand & $TC$ & $UQ$ & $FDD$ & \multirow[c]{2}{*}{Z} & $TC$ & $UQ$ & $FDD$ & $mFDD$ & \multirow[c]{2}{*}{Z} \\
+     & |$C$| & $|O|$ & $|K|$ & $|B|$ & (m$^3$) & (min) & (m$^3$) & (min) &  & (min) & (m$^3$)& (min) & (min) &  \\
     \midrule
@@
      & 6 & 11 & 18 & 2 & 333.5 & 753.3 & 0.0 & 0.0 & 753.3 & 753.3 & 0.0 & 0.0 & 0.0 & 753.3 \\
-    & 4 & 7 & 15 & 2 & 375.0 & 1,017.6 & 0.0 & 0.0 & 1,017.6 & 1,016.5 & 0.0 & 0.0 & 0.0 & 1,016.5 \\
-    & 7 & 8 & 19 & 2 & 388.0 & 1,662.7 & 0.0 & 0.0 & 1,662.7 & 1,651.8 & 0.0 & 0.0 & 0.0 & 1,651.8 \\
-    & 10 & 14 & 29 & 3 & 613.5 & 3,259.0 & 0.0 & 0.0 & 3,259.0 & 3,200.9 & 0.0 & 0.0 & 0.0 & 3,200.9 \\
-    & 8 & 11 & 31 & 3 & 776.0 & 2,222.0 & 0.0 & 0.0 & 2,222.0 & 2,208.9 & 0.0 & 0.0 & 0.0 & 2,208.9 \\
-    & 9 & 10 & 35 & 3 & 937.5 & 3,839.6 & 0.0 & 0.0 & 3,839.6 & 3,750.5 & 0.0 & 0.0 & 0.0 & 3,750.5 \\
+     & 4 & 7 & 15 & 2 & 375.0 & 1,013.7 & 0.0 & 0.0 & 1,013.7 & 1,013.1 & 0.0 & 0.0 & 0.0 & 1,013.1 \\
+     & 7 & 8 & 19 & 2 & 388.0 & 1,661.0 & 0.0 & 0.0 & 1,661.0 & 1,658.2 & 0.0 & 0.0 & 0.0 & 1,658.2 \\
+     & 10 & 14 & 29 & 3 & 613.5 & 3,235.5 & 0.0 & 0.0 & 3,235.5 & 3,198.6 & 0.0 & 0.0 & 0.0 & 3,198.6 \\
+     & 8 & 11 & 31 & 3 & 776.0 & 2,257.6 & 0.0 & 0.0 & 2,257.6 & 2,219.9 & 0.0 & 0.0 & 0.0 & 2,219.9 \\
+     & 9 & 10 & 35 & 3 & 937.5 & 3,834.0 & 0.0 & 0.0 & 3,834.0 & 3,808.8 & 0.0 & 0.0 & 0.0 & 3,808.8 \\
      \cmidrule(l){1-10} \cmidrule(l){11-15}
-     \multicolumn{5}{r}{Average} & - & 1,822.7 & 0.0 & 0.0 & 1,822.7  & 1,821.3 & 0.0 & 0.0 & 0.0 & 1,821.3 \\
+     \multicolumn{5}{r}{Average} & - & 1,843.0 & 0.0 & 0.0 & 1,843.0 & 1,830.1 & 0.0 & 0.0 & 0.0 & 1,830.1 \\
      \cmidrule(l){1-10} \cmidrule(l){11-15}
-     \multirow[c]{14}{*}{Medium} & 40 & 43 & 76 & 6 & 1,160.5 & 5,242.1 & 0.0 & 0.0 & 5,242.1 & 5,213.5 & 0.0 & 0.0 & 0.0 & 5,213.5 \\
-     & 42 & 47 & 104 & 8 & 1,565.0 & 8,291.5 & 0.0 & 0.0 & 8,291.5 & 7,993.7 & 0.0 & 0.0 & 0.0 & 7,993.7 \\
-     & 63 & 70 & 94 & 7 & 1,746.5 & 12,744.2 & 0.0 & 0.0 & 12,744.2 & 12,638.2 & 0.0 & 0.0 & 0.0 & 12,638.2 \\
-     & 67 & 78 & 116 & 8 & 1,839.5 & 11,918.0 & 0.0 & 0.0 & 11,919.8 & 11,793.3 & 0.0 & 0.0 & 0.0 & 11,793.3 \\
-     & 71 & 82 & 116 & 8 & 2,060.0 & 10,612.5 & 0.0 & 0.0 & 10,612.5 & 10,449.9 & 0.0 & 0.0 & 0.0 & 10,449.9 \\
-     & 57 & 70 & 117 & 6 & 2,327.5 & 14,781.6 & 0.0 & 15.7 & 30,457.6 & 14,391.0 & 0.0 & 2.4 & 1.4 & 16,820.6 \\
-     & 79 & 83 & 137 & 7 & 2,425.0 & 17,637.5 & 0.0 & 106.5 & 124,143.7 & 17,504.0 & 0.0 & 66.7 & 21.7 & 84,156.1 \\
-     & 66 & 80 & 127 & 8 & 2,512.5 & 13,676.8 & 0.0 & 0.0 & 13,676.8 & 13,595.0 & 0.0 & 0.0 & 0.0 & 13,595.0 \\
-     & 68 & 74 & 128 & 7 & 2,595.0 & 18,565.9 & 0.0 & 97.4 & 115,988.9 & 18,484.6 & 0.0 & 72.3 & 29.0 & 90,740.3 \\
-     & 85 & 97 & 136 & 8 & 2,673.0 & 15,741.9 & 0.0 & 75.7 & 91,395.7 & 15,779.3 & 0.0 & 48.2 & 18.6 & 64,002.6 \\
-     & 78 & 85 & 128 & 8 & 2,685.5 & 17,097.4 & 0.0 & 68.0 & 85,088.9 & 16,326.4 & 0.0 & 55.6 & 14.0 & 71,906.4 \\
-     & 77 & 85 & 131 & 8 & 2,893.5 & 17,807.8 & 0.0 & 10.7 & 28,544.5 & 17,994.5 & 0.0 & 9.4 & 6.1 & 27,404.5 \\
-     & 89 & 97 & 137 & 7 & 2,939.5 & 20,810.3 & 0.0 & 203.9 & 224,680.9 & 20,829.3 & 0.0 & 150.0 & 42.0 & 170,861.3 \\
-     & 84 & 98 & 133 & 7 & 2,971.0 & 21,142.9 & 0.0 & 380.8 & 401,955.9 & 21,206.1 & 0.0 & 252.7 & 55.6 & 273,892.1 \\
+    \multirow[c]{14}{*}{Medium} & 40 & 43 & 76 & 6 & 1,160.5 & 5,262.4 & 0.0 & 0.0 & 5,262.4 & 5,187.2 & 0.0 & 0.0 & 0.0 & 5,187.2 \\
+     & 42 & 47 & 104 & 8 & 1,565.0 & 8,278.0 & 0.0 & 0.0 & 8,278.0 & 8,209.6 & 0.0 & 0.0 & 0.0 & 8,209.6 \\
+     & 63 & 70 & 94 & 7 & 1,746.5 & 12,730.3 & 0.0 & 0.0 & 12,730.3 & 12,551.9 & 0.0 & 0.0 & 0.0 & 12,551.9 \\
+     & 67 & 78 & 116 & 8 & 1,839.5 & 11,884.4 & 0.0 & 0.0 & 11,884.4 & 11,641.4 & 0.0 & 0.0 & 0.0 & 11,641.4 \\
+     & 71 & 82 & 116 & 8 & 2,060.0 & 10,750.9 & 0.0 & 0.0 & 10,750.9 & 10,697.1 & 0.0 & 0.0 & 0.0 & 10,697.1 \\
+     & 57 & 70 & 117 & 6 & 2,327.5 & 14,547.0 & 0.0 & 25.8 & 40,380.4 & 14,746.8 & 0.0 & 16.1 & 9.1 & 30,816.0 \\
+     & 79 & 83 & 137 & 7 & 2,425.0 & 17,822.4 & 0.0 & 120.0 & 137,840.9 & 17,982.5 & 0.0 & 92.0 & 17.6 & 110,023.4 \\
+     & 66 & 80 & 127 & 8 & 2,512.5 & 13,622.2 & 0.0 & 0.0 & 13,622.2 & 13,565.0 & 0.0 & 0.0 & 0.0 & 13,565.0 \\
+     & 68 & 74 & 128 & 7 & 2,595.0 & 18,641.7 & 0.0 & 74.1 & 92,744.7 & 18,554.8 & 0.0 & 43.4 & 19.0 & 61,987.5 \\
+     & 85 & 97 & 136 & 8 & 2,673.0 & 17,336.3 & 0.0 & 51.8 & 69,145.9 & 17,610.5 & 0.0 & 23.1 & 16.3 & 40,731.6 \\
+     & 78 & 85 & 128 & 8 & 2,685.5 & 16,781.2 & 0.0 & 55.5 & 72,294.4 & 16,658.5 & 0.0 & 39.4 & 12.5 & 56,076.4 \\
+     & 77 & 85 & 131 & 8 & 2,893.5 & 17,789.1 & 0.0 & 22.0 & 39,787.5 & 16,923.5 & 0.0 & 0.0 & 0.0 & 16,923.5 \\
+     & 89 & 97 & 137 & 7 & 2,939.5 & 21,831.7 & 0.0 & 214.2 & 236,037.5 & 22,630.7 & 0.0 & 157.1 & 48.5 & 179,691.7 \\
+     & 84 & 98 & 133 & 7 & 2,971.0 & 21,031.1 & 0.0 & 474.0 & 495,073.6 & 21,179.3 & 0.0 & 359.8 & 59.8 & 380,980.3 \\
      \cmidrule(l){1-10} \cmidrule(l){11-15}
-     \multicolumn{5}{r}{Average}  & - & 14,719.3 & 0.0 & 68.5 & 83,195.9 & 14,585.6 & 0.0 & 46.9 & 13.5 & 61,533.4 \\
+     \multicolumn{5}{r}{Average}  & - & 14,879.2 & 0.0 & 74.1 & 88,988.1 & 14,867.1 & 0.0 & 52.2 & 13.1 & 67,077.3 \\
      \cmidrule(l){1-10} \cmidrule(l){11-15}
-     \multirow[c]{14}{*}{Large} & 98 & 109 & 132 & 8 & 3,078.5 & 22,186.9 & 0.0 & 15.2 & 37,379.4 & 22,061.8 & 0.0 & 9.9 & 6.3 & 31,964.6 \\
- & 101 & 119 & 129 & 8 & 3,229.0 & 18,290.6 & 0.0 & 40.5 & 58,817.5 & 17,111.2 & 0.0 & 10.5 & 5.7 & 27,595.2 \\
- & \multirow[c]{2}{*}{114} & 136 & 141 & 8 & 3,350.5 & 21,244.6 & 0.0 & 122.4 & 143,618.2 & 21,702.8 & 0.0 & 90.0 & 58.9 & 111,699.9 \\
- &  & 132 & 140 & 8 & 3,401.5 & 22,888.6 & 0.0 & 119.9 & 142,811.7 & 23,079.3 & 0.0 & 78.7 & 16.9 & 101,807.4 \\
- & 101 & 123 & 143 & 8 & 3,437.5 & 24,070.1 & 0.0 & 44.1 & 68,139.4 & 23,863.3 & 0.0 & 13.7 & 11.9 & 37,577.6 \\
- & 112 & 129 & 137 & 8 & 3,471.0 & 22,497.5 & 0.0 & 120.4 & 142,915.9 & 21,867.0 & 0.0 & 104.6 & 16.4 & 126,500.0 \\
- & 114 & 129 & 142 & 8 & 3,499.5 & 21,545.5 & 0.0 & 461.2 & 482,716.9 & 20,476.5 & 0.0 & 236.3 & 53.1 & 256,802.5 \\
- & \multirow[c]{2}{*}{98} & 122 & 149 & 8 & 3,513.0 & 20,370.9 & 0.0 & 150.5 & 170,865.5 & 19,725.6 & 0.0 & 106.5 & 46.3 & 126,172.6 \\
- &  & 108 & 139 & 8 & 3,541.0 & 21,538.2 & 0.0 & 339.1 & 360,638.8 & 21,205.7 & 0.0 & 201.3 & 59.5 & 222,493.7 \\
- & 108 & 122 & 144 & 8 & 3,670.5 & 23,462.8 & 0.0 & 253.3 & 276,751.0 & 23,827.4 & 0.0 & 144.9 & 31.9 & 168,731.4 \\
- & 92 & 107 & 142 & 8 & 3,684.5 & 22,799.8 & 0.0 & 52.4 & 75,149.3 & 24,071.4 & 0.0 & 30.9 & 10.2 & 54,975.8 \\
- & 114 & 136 & 138 & 8 & 3,739.5 & 23,491.1 & 0.0 & 86.7 & 110,190.5 & 23,242.2 & 0.0 & 46.7 & 32.5 & 69,982.2 \\
- & 112 & 131 & 148 & 8 & 3,953.5 & 25,300.6 & 0.0 & 387.0 & 412,296.4 & 26,296.6 & 0.0 & 247.8 & 32.5 & 274,119.6 \\
+     \multirow[c]{14}{*}{Large} & 98 & 109 & 132 & 8 & 3,078.5 & 22,272.3 & 0.0 & 7.2 & 29,438.3 & 23,294.5 & 0.0 & 4.0 & 1.8 & 27,252.8 \\
+     & 101 & 119 & 129 & 8 & 3,229.0 & 18,473.8 & 0.0 & 41.6 & 60,111.3 & 18,595.2 & 0.0 & 28.0 & 15.7 & 46,551.2 \\
+     & 114 & 136 & 141 & 8 & 3,350.5 & 21,237.4 & 0.0 & 136.7 & 157,923.2 & 20,762.9 & 0.0 & 110.7 & 18.6 & 131,504.9 \\
+     & 114 & 132 & 140 & 8 & 3,401.5 & 22,518.4 & 0.0 & 174.0 & 196,489.2 & 22,537.0 & 0.0 & 64.8 & 10.6 & 87,390.4 \\
+     & 101 & 123 & 143 & 8 & 3,437.5 & 23,807.5 & 0.0 & 39.1 & 62,917.8 & 23,249.3 & 0.0 & 30.2 & 13.8 & 53,413.3 \\
+     & 112 & 129 & 137 & 8 & 3,471.0 & 22,457.5 & 0.0 & 121.0 & 143,439.3 & 21,878.2 & 0.0 & 27.9 & 7.5 & 49,772.5 \\
+     & 114 & 129 & 142 & 8 & 3,499.5 & 21,574.4 & 0.0 & 362.8 & 384,413.2 & 22,591.1 & 0.0 & 221.0 & 52.9 & 243,564.1 \\
+     & 98 & 122 & 149 & 8 & 3,513.0 & 20,918.7 & 0.0 & 150.6 & 171,499.5 & 19,503.4 & 0.0 & 94.1 & 28.6 & 113,643.5 \\
+     & 98 & 108 & 139 & 8 & 3,541.0 & 21,548.1 & 0.0 & 271.5 & 293,062.8 & 22,008.3 & 0.0 & 239.9 & 47.4 & 261,921.3 \\
+     & 108 & 122 & 144 & 8 & 3,670.5 & 23,604.0 & 0.0 & 169.2 & 192,821.5 & 23,200.9 & 0.0 & 125.9 & 42.5 & 149,122.9 \\
+     & 92 & 107 & 142 & 8 & 3,684.5 & 21,279.5 & 0.0 & 83.8 & 105,084.6 & 21,244.2 & 0.0 & 36.4 & 14.4 & 57,641.8 \\
+     & 114 & 136 & 138 & 8 & 3,739.5 & 24,385.3 & 0.0 & 109.8 & 134,180.9 & 25,218.0 & 0.0 & 71.7 & 15.8 & 96,877.5 \\
+     & 127 & 136 & 150 & 8 & 3,946.5 & 24,434.2 & 0.0 & 391.0 & 415,476.7 & 24,732.2 & 0.0 & 292.5 & 52.8 & 317,191.2 \\
+     & 112 & 131 & 148 & 8 & 3,953.5 & 25,846.2 & 0.0 & 566.8 & 592,677.4 & 26,245.8 & 0.0 & 383.8 & 52.7 & 410,087.8 \\
      \cmidrule(l){1-10} \cmidrule(l){11-15}
@@
     }
-    \begin{minipage}{12cm}
-        \vspace{0.1cm}
-        {\footnotesize $TC$: travel cost, $UQ$: undelivered quantity, $FDD$: first delivery delay, $mFDD$: maximum of first delivery delays.}        
-        \end{minipage}
 \end{table}
@@
     \caption{Comparison of the average results with and without shift and overtime constraints}
-    \label{tab:cdp_lift_shift_constraints}
+    \label{tab:lift_shift_constraints}
     \resizebox{0.8\textwidth}{!}{
-    \begin{tabular}{lrrrrrrrr}
+    \begin{tabular}{lcccccccc}
     \toprule
@@
      \cmidrule(l){2-5} \cmidrule(l){6-9}
-    &   \multicolumn{1}{c}{$TC$}& \multicolumn{1}{c}{$UQ$} & \multicolumn{1}{c}{$FDD$} & \multicolumn{1}{c}{$mFDD$}  & \multicolumn{1}{c}{($TC$)} & \multicolumn{1}{c}{$UQ$} & \multicolumn{1}{c}{$FDD$} & \multicolumn{1}{c}{$mFDD$}  \\
-    & \multicolumn{1}{c}{(min)} & \multicolumn{1}{c}{(m$^3$)} & \multicolumn{1}{c}{(min)} & \multicolumn{1}{c}{(min)} & \multicolumn{1}{c}{(min)} & \multicolumn{1}{c}{(m$^3$)} & (min) & \multicolumn{1}{c}{(min)} \\
+     & $TC$ & $UQ$ & $FDD$ & $mFDD$ & $TC$ & $UQ$ & $FDD$ & $mFDD$ \\
+     & (min) & ($m^3$) & (min) & (min) & (min) & ($m^3$) & (min) & (min) \\
     \midrule
-    Small & 1,933.6 & 0.0 & 7.4 & 4.5 & 1,822.7 & 0.0 & 0.0 & 0.0 \\
-Medium & 15,070.0 & 9.9 & 403.0 & 46.9 & 14,719.3 & 0.0 & 68.5 & 18.4 \\
-Large & 21,810.5 & 14.7 & 945.0 & 56.4 & 22,283.6 & 0.0 & 168.7 & 31.0 \\
+    Small & 1,938.13 & 0.00 & 7.44 & 4.47 & 1,842.97 & 0.00 & 0.00 & 0.00 \\
+    Medium & 15,007.28 & 10.26 & 400.45 & 46.31 & 14,879.19 & 0.00 & 74.11 & 17.87 \\
+    Large & 22,045.81 & 13.26 & 935.49 & 56.93 & 22,454.09 & 0.00 & 187.51 & 34.76 \\
     \bottomrule
@@
 The delay of initial delivery decreases as $\gamma^1$ increases.
-These results confirm the fact that when the maximum time delay $\gamma^1$ is high, plants have more flexibility to delay current deliveries to make room for new ones or to make customers wait until drivers are available.
-Increasing the maximum delay from 10 to 25 minutes reduces the first delivery delay by over 50\% for the medium and large cases.
+These results confirm the fact that when $\gamma^1$ is high, plants have more flexibility to delay current deliveries to make room for new ones or to make customers wait until drivers are available.
+Increasing the maximum delay from 10 to 20 minutes reduces the first delivery delay by over 50\% for the medium and large cases.
 The average delivery time does not change significantly regardless of the value of $\gamma^1$.
@@
 Table~\ref{tab:cdp_tradeoff} reports the trade-off between the travel time and the first delivery delay for these solutions.
-The difference is not noticeable for small instances.
+The difference is not notable for small instances.
 However, for medium and large instances, it is clear that as the first delivery delay decreases, the travel time increases, and conversely, when we minimize the travel time, the first delivery delay tends to be longer.
-This result is expected because drivers assigned to one plant often must travel to another one to ensure that the first delivery remains on schedule.
-This trade-off highlights the challenge decision-makers face in finding the right balance between minimizing operational costs and meeting customer satisfaction with high service levels.
+This result is expected because drivers assigned to one plant often must travel to another to ensure that the first delivery remains on schedule.
+This trade-off highlights the challenge decision-makers face in finding the right balance between minimizing operational costs and meeting customer satisfaction.
 
@@
     \midrule
-    Small & 1,821.3 & 0.0 & 0.0 & 1,831.6 & 0.0 & 0.0 \\
-    Medium & 14,311.5 & 89.6 & 23.0 & 14,679.3 & 46.9 & 13.4 \\
-    Large & 21,403.8 & 170.5 & 31.4 & 22,194.7 & 101.7 & 29.4 \\
+    Small & 1,837.8 & 0.0 & 0.0 & 1,841.9 & 0.0 & 0.0 \\
+    Medium & 14,390.1 & 60.6 & 17.2 & 14,687.1 & 46.5 & 16.4 \\
+    Large & 21,753.1 & 180.0 & 33.7 & 22,316.4 & 128.1 & 37.7 \\
     \bottomrule
@@
 To solve this problem, we introduced a mathematical model and a heuristic solution based on Greedy Randomized Adaptive Search.
-Our analysis, using a dataset derived from our partner's historical data, demonstrated the effectiveness of our algorithm in providing good solutions within a reasonable time.
+Our analysis, using a dataset derived from our partner's historical data, demonstrated the effectiveness of our algorithm in making good decisions within a reasonable time.
 Furthermore, our model incorporates constraints related to working hours, making our heuristic a valuable tool for designing weekly schedules for company drivers.
 It can also better reassign drivers to their home plants, as these assignments affect travel costs to ensure timely deliveries.
-Our GRASP algorithm performed exceptionally well on the \cite{kinable2014concrete} datasets, indicating its reliability.
-For future work, it would be interesting to address datasets with different unloading times at construction sites or to explore different scenarios regarding the number of simultaneous deliveries and the maximum time between two consecutive orders at a construction site.
-One can also study the problem using a stochastic or robust methodology to account for the uncertainty of the parameters in a real-world setting.
-As a contribution to the literature on CDP, we provide a mathematical formulation for our variant, contribute with a dataset, and demonstrate the application of GRASP to solve both our real problem and the one studied by \cite{kinable2014concrete}.
+Our GRASP algorithm performed exceptionally well on \cite{kinable2014concrete} datasets, indicating its reliability.
+For future work, it would be interesting to address datasets with different unloading times at construction sites or to explore different scenarios regarding the number of simultaneous deliveries and the maximum time between two consecutive deliveries at a construction site.
+We can also study the problem using a stochastic or robust methodology to account for the uncertainty of the parameters in a real-world setting.
+As a contribution to the literature on CDP, we provide a mathematical formulation for our variant, contribute a dataset, and demonstrate the application of GRASP to solve both our problem and the one studied by \cite{kinable2014concrete}.
 

### 2024-03-09 instances/literature/script.py
Create instance to share with literature

@@ -1,0 +1,109 @@
+#%%
+import os
+import pandas as pd
+
+#%%
+data = [
+    {"Name": "20201121", "Instance": "C_13_11_12_1", "folder": "Small"},
+    {"Name": "20201010", "Instance": "C_13_5_8_1", "folder": "Small"},
+    {"Name": "20201107", "Instance": "C_18_6_11_2", "folder": "Small"},
+    {"Name": "20201031", "Instance": "C_15_4_7_2", "folder": "Small"},
+    {"Name": "20201114", "Instance": "C_19_7_8_2", "folder": "Small"},
+    {"Name": "20201003", "Instance": "C_29_10_14_3", "folder": "Small"},
+    {"Name": "20201128", "Instance": "C_31_8_11_3", "folder": "Small"},
+    {"Name": "20201024", "Instance": "C_35_9_10_3", "folder": "Small"},
+    {"Name": "20201021", "Instance": "C_76_40_43_6", "folder": "Medium"},
+    {"Name": "20201013", "Instance": "C_104_42_47_8", "folder": "Medium"},
+    {"Name": "20201109", "Instance": "C_94_63_70_7", "folder": "Medium"},
+    {"Name": "20201026", "Instance": "C_116_67_78_8", "folder": "Medium"},
+    {"Name": "20201016", "Instance": "C_116_71_82_8", "folder": "Medium"},
+    {"Name": "20201130", "Instance": "C_117_57_70_6", "folder": "Medium"},
+    {"Name": "20201125", "Instance": "C_137_79_83_7", "folder": "Medium"},
+    {"Name": "20201007", "Instance": "C_127_66_80_8", "folder": "Medium"},
+    {"Name": "20201123", "Instance": "C_128_68_74_7", "folder": "Medium"},
+    {"Name": "20201020", "Instance": "C_136_85_97_8", "folder": "Medium"},
+    {"Name": "20201124", "Instance": "C_128_78_85_8", "folder": "Medium"},
+    {"Name": "20201002", "Instance": "C_131_77_85_8", "folder": "Medium"},
+    {"Name": "20201116", "Instance": "C_137_89_97_7", "folder": "Medium"},
+    {"Name": "20201126", "Instance": "C_133_84_98_7", "folder": "Medium"},
+    {"Name": "20201102", "Instance": "C_132_98_109_8", "folder": "Large"},
+    {"Name": "20201009", "Instance": "C_129_101_119_8", "folder": "Large"},
[…]

### 2024-03-09 draft/modele/conflict_refiner.py
remove draft folder

@@ -1,23 +1,0 @@
-import cplex
-
-# Create a new modeler object
-modeler = cplex.Cplex()
-
-# Read the .lp file
-modeler.read("filename.lp")
-
-# Extract the model from the modeler
-model = modeler.get_model()
-
-# Set up the conflict refiner
-conflict = model.conflict()
-conflict.set_level(2)  # Set the level of detail to 2
-
-# Solve the model
-model.solve()
-
-# Check if the model is infeasible
-if model.solution.is_infeasible():
-  # Identify the minimal conflicting set of constraints or variables
-  minimal_conflict = conflict.refine()
-  print("Minimal conflicting set:", minimal_conflict)

### 2024-03-25 ArticleCDP/journals/elsevier/articleCDP.tex
article with elsevier template

--- 2024-03-09 ArticleCDP/original/articleCDP.tex
@@
-\documentclass{article}
-% \documentclass[review]{elsarticle}
+\documentclass[authoryear,preprint,review,11pt]{elsarticle}
+
+%% Use the option review to obtain double line spacing
+%% \documentclass[authoryear,preprint,review,12pt]{elsarticle}
+
 \usepackage{epsfig}
@@
 \usepackage{graphicx}
-\usepackage{natbib}
+% \usepackage{natbib}
 \usepackage{datatool}
@@
 \usepackage{pgfplotstable}
-\pgfplotsset{width=7cm,compat=newest}
+\pgfplotsset{compat=newest}
 \usepackage{booktabs}
@@
 \usepackage{float}
-\usepackage[text={15cm,21cm},centering]{geometry}
+\usepackage[text={16cm,21cm},centering]{geometry}
 \usepackage{setspace}
@@
 
-\newtheorem{theorem}{Theorem}[section]
-\newtheorem{lemma}[theorem]{Lemma}
-\newtheorem{proposition}[theorem]{Proposition}
-\newtheorem{corollary}[theorem]{Corollary}
-\newtheorem{definition}[theorem]{Definition}
-\newcommand{\pcmaxc}{P$|\mbox{\em cont}|$C$_{\max}$\xspace}
-
 \newcommand{\GRASP}{Greedy randomized adaptive search procedure\xspace}
-\newcommand{\F}{$\mathcal{F}$\xspace}
-\newcommand{\B}{$\mathcal{B}$\xspace}
-\newcommand{\C}{$\mathcal{C}$\xspace}
-\newcommand{\LB}{$\mathcal{L}$\xspace}
-\newcommand{\Lb}{$\mathcal{L}_b$\xspace}
-\newcommand{\D}{$\mathcal{D}$\xspace}
+
  \newcommand*{\red}{\textcolor{red}}
@@
 
-\renewcommand{\tabcolsep}{4pt}
-
-\modulolinenumbers[1]
+% \renewcommand{\tabcolsep}{4pt}
+
+% \modulolinenumbers[1]
 % \linenumbers
 
-\oddsidemargin=0.5in
-\topmargin=-0.25in
-\textwidth=5.5in
-\textheight=8.25in
-
-%\journal{European Journal of Operational Research}
-
-%\begin{frontmatter}
+\journal{Computers \& Operations Research}
+
+\begin{document}
+
+\begin{frontmatter}
+
+%% Title, authors and addresses
+
+%% use the tnoteref command within \title for footnotes;
+%% use the tnotetext command for theassociated footnote;
+%% use the fnref command within \author or \address for footnotes;
+%% use the fntext command for theassociated footnote;
+%% use the corref command within \author for corresponding author footnotes;
+%% use the cortext command for theassociated footnote;
+%% use the ead command for the email address,
+%% and the form \ead[url] for the home page:
+%% \title{Title\tnoteref{label1}}
+%% \tnotetext[label1]{}
+%% \author{Name\corref{cor1}\fnref{label2}}
+%% \ead{email address}
+%% \ead[url]{home page}
+%% \fntext[label2]{}
+%% \cortext[cor1]{}
+%% \affiliation{organization={},
+%%             addressline={},
+%%             city={},
+%%             postcode={},
+%%             state={},
+%%             country={}}
+%% \fntext[label3]{}
 
 \title{A GRASP algorithm for the concrete delivery problem}
-\author{Ousmane Ali$^{(1)}$, Jean-Fran\c cois C\^ot\'e$^{(2)}$, Leandro C.~Coelho$^{(3)}$\\
- $(1)$ {\tt nassoma-wattara-ousmane.ali.1@ulaval.ca}\\
- $(2)$ {\tt Jean-Francois.Cote@fsa.ulaval.ca}\\
- $(3)$ {\tt Leandro.Coelho@fsa.ulaval.ca}\\
-}
-
-\date{\today}
-
-\begin{document}
-\maketitle
+
+%% use optional labels to link authors explicitly to addresses:
+%% \author[label1,label2]{}
+%% \affiliation[label1]{organization={},
+%%             addressline={},
+%%             city={},
+%%             postcode={},
+%%             state={},
+%%             country={}}
+%%
+%% \affiliation[label2]{organization={},
+%%             addressline={},
+%%             city={},
+%%             postcode={},
+%%             state={},
+%%             country={}}
+
+\author[inst1]{Ousmane Ali}
+
+\affiliation[inst1]{organization={CIRRELT, Université LAVAL},%Department and Organization
+            % addressline={Address One}, 
+            % city={Québec},
+            % postcode={00000}, 
+            % state={Qu\'ebec},
+            country={Canada}}
+
+\author[inst1]{Jean-Fran\c cois C\^ot\'e}
+\author[inst1,inst2]{Leandro C.~Coelho}
+
+\affiliation[inst2]{organization={Canada research chair in integrated logistics},%Department and Organization
+            % addressline={Address Two}, 
+            % city={City Two},
+            % postcode={22222}, 
+            % state={Québec},
+            country={Canada}}
+
 \begin{abstract}
@@
 
-% \begin{keywords}
-\noindent{\textbf{Keywords:} Vehicle scheduling;  concrete delivery; GRASP; ready-mixed concrete}
-% \end{keywords}
+\begin{keyword}
+Vehicle scheduling \sep  concrete delivery \sep GRASP \sep ready-mixed concrete
+\end{keyword}
+
+\end{frontmatter}
+
+% \linenumbers
 
@@
 Several works to schedule and dispatch concrete production and delivery have mainly focused on simulation-related methods.
-These methods can be standalone, such as those used by \cite{zayed2001simulation, wang2001scheduling} and  \cite{panas_simulation_based_2013}, among others.
+These methods can be standalone, such as those used by \cite{zayed2001simulation} and \cite{wang2001scheduling}, among others.
 Alternatively, these methods can be hybridized with optimization techniques, such as those used by \cite{feng2004optimizing} and \cite{lu2005optimized}.
@@
 As such, the simulation model also included the batch and delivery of mortar.
-Finding the best RMC fleet size was also the purpose of the discrete-event simulation model proposed by \cite{panas_simulation_based_2013}.
 
 In addition to simulation-based methods, several other approaches have been used in the literature to solve the CDP.
-These include metaheuristics \citep{misir2011selection, maghrebi2016sequential, yang2022concrete}, exact methods \citep{asbach2009analysis, kinable2014concrete}, matheuristics \citep{schmid2009hybrid, schmid2010hybridization}, Benders decomposition \citep{maghrebi2014benders}, column generation (CG) \citep{maghrebi2014solving}, Lagrangian relaxation \citep{narayanan2015using}, and machine learning approaches \citep{graham2006modeling}.
+These include metaheuristics \citep{ maghrebi2016sequential, yang2022concrete}, exact methods \citep{asbach2009analysis, kinable2014concrete}, matheuristics \citep{schmid2009hybrid, schmid2010hybridization}, Benders decomposition \citep{maghrebi2014benders}, column generation (CG) \citep{maghrebi2014solving}, Lagrangian relaxation \citep{narayanan2015using}, and machine learning approaches \citep{graham2006modeling}.
 \cite{matsatsinis2004towards} designed a decision support system (DSS) for the dynamic routing of both concrete and pumps that may be necessary for some construction sites to aid in the unloading of concrete.
@@
 For a driver, the decision is to determine the sequence of loading depots and delivery sites.
-And for a construction site, the decision involves determining the arrival times of all scheduled deliveries for the day.
+For a construction site, the decision involves determining the arrival times of all scheduled deliveries for the day.
 
@@
 Together, these components drive the optimization process to find a solution that efficiently balances travel costs, customer satisfaction, on-time delivery, driver utilization, and scheduling constraints.
-
+{
+\small
 \begin{alignat}{3}
     \label{mod:obj}   \min & \sum_{k \in K}{{ \sum_{ \substack{(i,j) \in A} }} {t_{ij}x^{k}_{ij}}} + \beta_1\smashoperator{\sum_{ o \in O}}{  \left(1-y_o\right)} \hspace{-0mm}  + \beta_2 \smashoperator{ \sum_{i \in C } } { g_i  } +  \sum_{k \in K}\left({\beta_3 w^{1}_k  + \beta_4  w^{2}_k}\right) &                             \\
-    \label{mod:c1}         & \sum_{j \in \delta^{+}(i) }{x^k_{ij}} =1     & \forall i \in S, k \in K                                                                      \\
-    \label{mod:c2}         & \sum_{j \in \delta^{-}(i)}{x^k_{ji}} = 1     & \forall i \in E, k \in K                                                                         \\
-    \label{mod:c03}        & v_j \geq  w_i  - M\left(1-x^{k}_{ij}\right)   & \hspace{-20mm}  i \in S, j \in \delta^{+}(i),  k \in K      \\
-    \label{mod:c3}         & v_j \geq  w_i + \alpha_{b} + t_{ij} - M\left(1-x^{k}_{ij}\right) & \hspace{-25mm} \forall b \in B, i \in L_b, j \in \delta^{+}(i),  k \in K     \\
-    \label{mod:c4}         & v_j \geq  w_i + \rho + t_{ij} - M\left(1-x^{k}_{ij}\right)                                                                       & \hspace{-20mm} \forall i \in D  , j \in \delta^{+}(i), k \in K                         \\
-    \label{mod:c5}         & w_{i} \geq v_{i}  + \frac{q^k_j}{\tau^l_b} -M\left(1- x^{k}_{ij}\right)   & \hspace{-25mm} \forall  b \in B,  i \in L_{b},  j \in \delta^{+}(i), k \in K \\
-    \label{mod:c6}         & w_{j} \geq v_{j}  + {  \frac{q^k_j}{\tau^u_c} -M\left(1- x^{k}_{ij}\right) }     & \hspace{-25mm}  \forall c \in  C, j \in D_{c}, i \in \delta^{-}(j),  k \in K \\
-    \label{mod:c7}         & w_{j} \leq v_{i}  + \Delta + M\left(1- x^{k}_{ij}\right)  & \hspace*{-40mm} j \in D,  i \in \delta^{-}(j),  k \in K \text{ }  \\
-    \label{mod:c8}         & v_{d^0_{o}} \geq a_c    & \hspace{-40mm} \forall  c \in C, \forall o \in O_c \text{ }                           \\
-    \label{mod:c9}          & g_{c} \geq v_{d^0_{o_1}} - a_c - M\left(  \sum_{ j \in \delta^{-}_{D}(d^0_{o_1})  }{u_{jd^0_{o_1}}}\right)  & \hspace{-40mm}  \forall  c \in C, \forall o_1 \in O_c \text{ }  \\
+    \label{mod:c1}         & \sum_{j \in \delta^{+}(i) }{x^k_{ij}} =1   & \hspace{-10mm}  \forall i \in S, k \in K   \\
+    \label{mod:c2}         & \sum_{j \in \delta^{-}(i)}{x^k_{ji}} = 1     & \hspace{-10mm} \forall i \in E, k \in K\\
+    \label{mod:c03}        & v_j \geq  w_i  - M\left(1-x^{k}_{ij}\right)   & \hspace{-30mm}  i \in S, j \in \delta^{+}(i),  k \in K      \\
+    \label{mod:c3}         & v_j \geq  w_i + \alpha_{b} + t_{ij} - M\left(1-x^{k}_{ij}\right) & \hspace{-40mm} \forall b \in B, i \in L_b, j \in \delta^{+}(i),  k \in K     \\
+    \label{mod:c4}         & v_j \geq  w_i + \rho + t_{ij} - M\left(1-x^{k}_{ij}\right)                                                                       & \hspace{-40mm} \forall i \in D  , j \in \delta^{+}(i), k \in K                         \\
+    \label{mod:c5}         & w_{i} \geq v_{i}  + \frac{q^k_j}{\tau^l_b} -M\left(1- x^{k}_{ij}\right)   & \hspace{-40mm} \forall  b \in B,  i \in L_{b},  j \in \delta^{+}(i), k \in K \\
+    \label{mod:c6}         & w_{j} \geq v_{j}  + {  \frac{q^k_j}{\tau^u_c} -M\left(1- x^{k}_{ij}\right) }     & \hspace{-40mm}  \forall c \in  C, j \in D_{c}, i \in \delta^{-}(j),  k \in K \\
+    \label{mod:c7}         & w_{j} \leq v_{i}  + \Delta + M\left(1- x^{k}_{ij}\right)  & \hspace*{-50mm} j \in D,  i \in \delta^{-}(j),  k \in K \text{ }  \\
+    \label{mod:c8}         & v_{d^0_{o}} \geq a_c    & \hspace{-50mm} \forall  c \in C, \forall o \in O_c \text{ }                           \\
+    \label{mod:c9}          & g_{c} \geq v_{d^0_{o_1}} - a_c - M\left(  \sum_{ j \in \delta^{-}_{D}(d^0_{o_1})  }{u_{jd^0_{o_1}}}\right)  & \hspace{-50mm}  \forall  c \in C, \forall o_1 \in O_c \text{ }  \\
     \label{mod:c91}  & g_{c} - a_c \leq \tau^w_c  & \hspace{-20mm}  \forall  c \in C \text{ }   \\
@@
     \label{mod:c14}        & \hspace{2mm} \smashoperator{ \sum_{ j \in \delta^{-}_{D}(d^0_{o_1})}}{u_{j,d^0_{o_1} } } \leq 1      & \hspace{-20mm} \forall o_1 \in O \text{ }        \\
-    \label{mod:c15}        & \hspace{2mm} \smashoperator{ \sum_{ j \in \delta^{+}_{D}(d^0_{o_1})}}{u_{d^0_{o_1},j } } +\smashoperator{ \sum_{ j \in \delta^{-}_{D}(d^0_{o_1})}}{u_{j,d^0_{o_1} } } \geq 1    & \hspace{-20mm} \forall o_1 \in O \text{ }    \\
-    \label{mod:c16}        & v_{j} \geq w_{j-1} - M\left(1- u_{j-1,j}\right)   & \hspace{-40mm}  \forall o \in O, j \in D_{o}, j \geq 1  \text{ }    \\
-    \label{mod:c17}        & v_{j}  \leq  w_{j-1} + \gamma^1 + M\left(1- u_{j-1,j}\right)    & \hspace{-50mm} \forall c \in C, \forall o \in O_c, j \in D_{o}, j \geq 1 \text{ }     \\
-    \label{mod:c18}        & u_{j-1,j} \geq u_{j,j+1}      & \hspace{-60mm} \forall o_1 \in O,  j \in D_{o_1}, 1 \leq j \leq n_{o_{1}}-1 \text{ }   \\
-    \label{mod:c19}        & u_{j-1,j} \geq \sum_{l \in L}{x_{lj}}   & \hspace{-40mm}  \forall o_1 \in O, j \in D_{o_1}, j \geq 1 \text{ }                   \\
+    \label{mod:c15}        & \hspace{2mm} \smashoperator{ \sum_{ j \in \delta^{+}_{D}(d^0_{o_1})}}{u_{d^0_{o_1},j } } +\smashoperator{ \sum_{ j \in \delta^{-}_{D}(d^0_{o_1})}}{u_{j,d^0_{o_1} } } \geq 1    & \hspace{-50mm} \forall o_1 \in O \text{ }    \\
+    \label{mod:c16}        & v_{j+1} \geq w_{j} - M\left(1- u_{j,j+1}\right)   & \hspace{-50mm}  \forall o \in O, j \in D_{o}, j \geq 1  \text{ }    \\
+    \label{mod:c17}        & v_{j+1}  \leq  w_{j} + \gamma^1 + M\left(1- u_{j,j+1}\right)    & \hspace{-60mm} \forall c \in C, \forall o \in O_c, j \in D_{o}, j \geq 1 \text{ }     \\
+    \label{mod:c18}        & u_{j,j+1} \geq u_{j+1,j+2}      & \hspace{-70mm} \forall o_1 \in O,  j \in D_{o_1}, 1 \leq j \leq n_{o_{1}}-2 \text{ }   \\
+    \label{mod:c19}        & u_{j,j+1} \geq \sum_{l \in L}{x_{lj}}   & \hspace{-60mm}  \forall o_1 \in O, j \in D_{o_1}, j \geq 1 \text{ }                   \\
     \label{mod:c20}        & v_{j} \geq w_{i} - M\left(1- u_{i,j}\right)   & \hspace{-40mm}  \forall i \in L, j \in \delta^{+}_{L}(i) \text{ }  \\
@@
 \end{alignat}
-
+}
 $\beta_1$ to $\beta_4$ are the penalty coefficients of each component of the objective function (\ref{mod:obj}).
@@
 
+{
+    % \small
 \begin{table}[!ht]
@@
 \end{table}
-
+}
 \begin{figure}[htb]
@@
     \label{fig:plants_schedules}
-        \includegraphics[width=\linewidth]{exemple.pdf}
+        \includegraphics[width=\linewidth]{gant_example.pdf}
 \end{figure}
@@
 $U^k_{d} = \left(k,v_d, w_d\right)$ indicates that driver $k$ will start unloading at time $v_d$ and finish at time $w_d$.
-We then call a solution $Sol =\left\lbrace \cup _{1 \leq j \leq |D|} (L^k_{j}, U^k_{j}), k \in K \right\rbrace$ the set of all pairs of loading and unloading tasks performed on a given day.
+We then call a solution $Sol =\left\lbrace \cup _{1 \leq j \leq |D|} (L^k_{j}, U^k_{j}), k \in K \right\rbrace$ the set of all pairs of loading and unloading tasks performed.
 
@@
 The Greedy Randomized Search Procedure (GRASP) algorithm is a multi-start metaheuristic for combinatorial optimization problems proposed by \cite{feo1989probabilistic}.
-It is a two-phase iterative sequence of constructive and local search algorithms.
+It has been successfully applied to solving scheduling problems \citep{bamoumen2023efficient,yepes2023flowshop} and transportation-related problems such as the prisoner transportation problem \citep{ferone2023efficient} and the truck and trailer routing problem \citep{villegas2011grasp}.
+The GRASP is a two-phase iterative sequence of constructive and local search algorithms.
 A feasible solution is generated in the construction phase using a greedy randomized algorithm.
@@
 These include sorting by decreasing demand and time window, sorting by increasing demand and time window, and finally, with a randomized list.
+{
+\footnotesize
     \begin{algorithm}[!ht]
@@
 
-        $M \leftarrow \emptyset $ \tcp{List of pairs of loading -- unloading tasks for each delivery node}
+        $M \leftarrow \emptyset $ \small \tcp{List of pairs of loading -- unloading tasks for each delivery node}
 
@@
 
-    Go to line 6 \tcp{backtrack to $j'$ position and retry to insert $D^{j'}_{o}$}
+    Go to line 6 \small \tcp{backtrack and retry to insert $D^{j'}_{o}$}
     }
@@
     \end{algorithm}
-
+}
 When confronted with scenarios where several customers require their services to start simultaneously or within the same time window, we observed that with Algorithm~\ref{alg:greedyCustIns}, some customers remain unscheduled when resources are limited.
-To address this limitation, we introduce the delivery-based insertion heuristic.
+To address this limitation, we introduce the delivery-based insertion heuristic, which allows each customer to have one of its delivery nodes considered for scheduling at each iteration.
 
@@
 Then, the algorithm constructs at line $20$ a list $T$ containing all pairs of loading and unloading tasks for the delivery nodes in $CL'$.
-
 If $T$ is empty, the algorithm assumes that the customers associated with the nodes in $CL'$ cannot be scheduled, and it removes those customers' existing schedules from the solution.
@@
 Otherwise, if the customer has remaining orders, the first delivery node $d^{1}_{o'}$ of another randomly selected order $o'$ is added to $CL$.
-
-Unlike the customer-based insertion, the delivery-based insertion allows each customer to have one of its delivery nodes considered for scheduling at each iteration.
 %In these two heuristics, the greedy aspect of GRASP refers to the creation of the set $RCL$, the probabilistic aspect is the random selection in $RCL$, and the adaptive part is the updating of $CL$ and the re-evaluation of the costs.
@@
 
-    $M \leftarrow \emptyset $ \tcp{List of pairs of loading -- unloading tasks for each delivery node}
+    $M \hspace{-2mm} \leftarrow \emptyset $ \small \tcp{List of pairs of loading -- unloading tasks for each delivery node}
 
@@
 
-                    $best\_cost \leftarrow c$
+                    $best\_cost \leftarrow cost$
                 }
@@
         \uElse{
-            $c_{min} \leftarrow \min\left\{ cost \ | \ (L^k_{d}, U^k_{d},cost,k) \in T   \right\} $
-
-            $c_{max} \leftarrow \max\left\{ cost \ | \ (L^k_{d}, U^k_{d},cost,k) \in T   \right\} $
+            $c_{min} \leftarrow \min\left\{ cost \ | \ (L^k_{d}, U^k_{d},cost,k) \in T   \right\} $, $c_{max} \leftarrow \max\left\{ cost \ | \ (L^k_{d}, U^k_{d},cost,k) \in T   \right\} $
 
@@
 This enables the redistribution of delivery nodes among drivers, potentially reducing drivers underutilization and overtime costs.
-
+% \clearpage
 \section{Computational experiments}
@@
     }
-    \begin{minipage}{12cm}
+    \begin{minipage}{\textwidth}
     \vspace{0.1cm}
-    {\footnotesize $TC$: travel cost, $UQ$: undelivered quantity, $FDD$: first delivery delay, $mFDD$: maximum of first delivery delays, $DUC$: driver underutilization cost, $DOC$: driver overtime cost.}
+    {\scriptsize \hspace*{-5mm} $TC$: travel cost, $UQ$: undelivered quantity, $FDD$: first delivery delay, $mFDD$: maximum of first delivery delays, $DUC$: driver underutilization cost, $DOC$: driver overtime cost.}
     \end{minipage}
@@
 
-\begin{table}[htb]
+\begin{table}[ht!]
     \centering
@@
     \toprule
-     &  & \multicolumn{5}{c}{Runtime (min)} \\
-     \cmidrule{3-7}
-     &  & 1 & 5 & 10 & 30 & 60 \\
+    %  &  & \multicolumn{5}{c}{Runtime (min)} \\
+    %  \cmidrule{3-7}
+     &  & \multicolumn{1}{c}{1 min} & \multicolumn{1}{c}{5 min} & \multicolumn{1}{c}{10 min} & \multicolumn{1}{c}{30 min} &\multicolumn{1}{c}{60 min} \\
     \midrule
@@
 Upon investigation, we find that even though we have $104$ drivers in this instance, the number of drivers who start their shift before $07\text{:}00$ is insufficient to serve this customer within the 60-minute window we defined.
-Thus, the value of $UQ$ that we report in Table~\ref{tab:cdp_all_instances} appears optimal given the driver shift schedule given as input to our algorithm.
+Thus, the value of $UQ$ that we report in Table~\ref{tab:cdp_all_instances} appears optimal given the driver shift schedule provided as input to our algorithm.
 We confirmed this observation by removing the driver start shift and overtime constraints and resolving our instances five more times.
@@
     }
-    \begin{minipage}{12cm}
+    \begin{minipage}{\textwidth}
         \vspace{0.1cm}
-        {\footnotesize $TC$: travel cost, $UQ$: undelivered quantity, $FDD$: first delivery delay, $mFDD$: maximum of first delivery delays.}        
+        {\scriptsize $TC$: travel cost, $UQ$: undelivered quantity, $FDD$: first delivery delay, $mFDD$: maximum of first delivery delays.}        
         \end{minipage}
@@
     \label{tab:cdp_tradeoff}
-    \begin{adjustbox}{max width=0.8\textwidth}
+    \begin{adjustbox}{max width=0.9\textwidth}
     \begin{tabular}{lcccccc}
@@
      \cmidrule(l){2-4} \cmidrule(l){5-7}
-     & $TC$ & $FDD$ & $mFDD$ & $TC$ & $FDD$ & $mFDD$ \\
-     & (min) & (min) & (min) & (min) & (min) & (min) \\
+     & \multicolumn{1}{c}{$TC$ (min)} & \multicolumn{1}{c}{$FDD$ (min)} & \multicolumn{1}{c}{$mFDD$ (min)} & \multicolumn{1}{c}{$TC$ (min)} & \multicolumn{1}{c}{$FDD$ (min)} & \multicolumn{1}{c}{$mFDD$ (min)}\\
+    %  & \multicolumn{1}{c}{(min)} & \multicolumn{1}{c}{(min)}  & \multicolumn{1}{c}{(min)} & \multicolumn{1}{c}{(min)}  & \multicolumn{1}{c}{(min)} & \multicolumn{1}{c}{(min)}  \\
     \midrule
@@
 
-\vspace{1.5cm} \noindent \textbf{Acknowledgments}
+\vspace{1.cm} \noindent \textbf{Acknowledgments}
 
@@
 
-\bibliographystyle{plainnat}
-\bibliography{references}
+ \bibliographystyle{elsarticle-num-names} 
+ \bibliography{References}
 

### 2024-03-25 ArticleCDP/original/articleCDP.tex
correct indices in model
last version

@@
-\documentclass{article}
-% \documentclass[review]{elsarticle}
+\documentclass[]{article}
 \usepackage{epsfig}
@@
     \label{fig:plants_schedules}
-        \includegraphics[width=\linewidth]{exemple.pdf}
+        \includegraphics[width=\linewidth]{gantt_example.pdf}
 \end{figure}
@@
  \bibliographystyle{plainnat} 
-\bibliography{references}
+ \bibliography{References}
 

### 2024-04-16 ArticleCDP/journals/elsevier/articleCDP.tex
Article soumis

@@
 Several works to schedule and dispatch concrete production and delivery have mainly focused on simulation-related methods.
-These methods can be standalone, such as those used by \cite{zayed2001simulation} and \cite{wang2001scheduling}, among others.
-Alternatively, these methods can be hybridized with optimization techniques, such as those used by \cite{feng2004optimizing} and \cite{lu2005optimized}.
-\cite{wang2001scheduling} developed a simulation model to reveal the effect and value of the concrete mixers' inter-arrival time on the productivity of hired unloading equipment on site.
+These methods can be standalone \citep{zayed2001simulation} or can be hybridized with optimization techniques, such as those used by \cite{feng2004optimizing} and \cite{lu2005optimized}.
 \cite{feng2004optimizing} used a combination of genetic algorithm (GA) and simulation process to minimize the total waiting time for trucks at a customer site.
@@
     \label{mod:c0}
-    n_o^{min} = \left\lceil \frac{q_o}{Q_{max}} \right\rceil \leq n_o \leq n_o^{max} = \left\lceil \frac{q_o}{Q_{min}} \right\rceil & \text{ } &
-    \forall  o \in O.
+    n_o^{min} = \left\lceil \frac{q_o}{Q_{max}} \right\rceil \leq n_o \leq n_o^{max} = \left\lceil \frac{q_o}{Q_{min}} \right\rceil.
 \end{alignat}
@@
     \label{mod:obj}   \min & \sum_{k \in K}{{ \sum_{ \substack{(i,j) \in A} }} {t_{ij}x^{k}_{ij}}} + \beta_1\smashoperator{\sum_{ o \in O}}{  \left(1-y_o\right)} \hspace{-0mm}  + \beta_2 \smashoperator{ \sum_{i \in C } } { g_i  } +  \sum_{k \in K}\left({\beta_3 w^{1}_k  + \beta_4  w^{2}_k}\right) &                             \\
-    \label{mod:c1}         & \sum_{j \in \delta^{+}(i) }{x^k_{ij}} =1   & \hspace{-10mm}  \forall i \in S, k \in K   \\
-    \label{mod:c2}         & \sum_{j \in \delta^{-}(i)}{x^k_{ji}} = 1     & \hspace{-10mm} \forall i \in E, k \in K\\
+    \label{mod:c1}         & \sum_{j \in \delta^{+}(i) }{x^k_{ij}} =1   & \hspace{-10mm}  i \in S, k \in K   \\
+    \label{mod:c2}         & \sum_{j \in \delta^{-}(i)}{x^k_{ji}} = 1     & \hspace{-10mm} i \in E, k \in K\\
     \label{mod:c03}        & v_j \geq  w_i  - M\left(1-x^{k}_{ij}\right)   & \hspace{-30mm}  i \in S, j \in \delta^{+}(i),  k \in K      \\
-    \label{mod:c3}         & v_j \geq  w_i + \alpha_{b} + t_{ij} - M\left(1-x^{k}_{ij}\right) & \hspace{-40mm} \forall b \in B, i \in L_b, j \in \delta^{+}(i),  k \in K     \\
-    \label{mod:c4}         & v_j \geq  w_i + \rho + t_{ij} - M\left(1-x^{k}_{ij}\right)                                                                       & \hspace{-40mm} \forall i \in D  , j \in \delta^{+}(i), k \in K                         \\
-    \label{mod:c5}         & w_{i} \geq v_{i}  + \frac{q^k_j}{\tau^l_b} -M\left(1- x^{k}_{ij}\right)   & \hspace{-40mm} \forall  b \in B,  i \in L_{b},  j \in \delta^{+}(i), k \in K \\
-    \label{mod:c6}         & w_{j} \geq v_{j}  + {  \frac{q^k_j}{\tau^u_c} -M\left(1- x^{k}_{ij}\right) }     & \hspace{-40mm}  \forall c \in  C, j \in D_{c}, i \in \delta^{-}(j),  k \in K \\
+    \label{mod:c3}         & v_j \geq  w_i + \alpha_{b} + t_{ij} - M\left(1-x^{k}_{ij}\right) & \hspace{-40mm} b \in B, i \in L_b, j \in \delta^{+}(i),  k \in K     \\
+    \label{mod:c4}         & v_j \geq  w_i + \rho + t_{ij} - M\left(1-x^{k}_{ij}\right)                                                                       & \hspace{-40mm} i \in D  , j \in \delta^{+}(i), k \in K                         \\
+    \label{mod:c5}         & w_{i} \geq v_{i}  + \frac{q^k_j}{\tau^l_b} -M\left(1- x^{k}_{ij}\right)   & \hspace{-40mm}  b \in B,  i \in L_{b},  j \in \delta^{+}(i), k \in K \\
+    \label{mod:c6}         & w_{j} \geq v_{j}  + {  \frac{q^k_j}{\tau^u_c} -M\left(1- x^{k}_{ij}\right) }     & \hspace{-40mm}  c \in  C, j \in D_{c}, i \in \delta^{-}(j),  k \in K \\
     \label{mod:c7}         & w_{j} \leq v_{i}  + \Delta + M\left(1- x^{k}_{ij}\right)  & \hspace*{-50mm} j \in D,  i \in \delta^{-}(j),  k \in K \text{ }  \\
-    \label{mod:c8}         & v_{d^0_{o}} \geq a_c    & \hspace{-50mm} \forall  c \in C, \forall o \in O_c \text{ }                           \\
-    \label{mod:c9}          & g_{c} \geq v_{d^0_{o_1}} - a_c - M\left(  \sum_{ j \in \delta^{-}_{D}(d^0_{o_1})  }{u_{jd^0_{o_1}}}\right)  & \hspace{-50mm}  \forall  c \in C, \forall o_1 \in O_c \text{ }  \\
-    \label{mod:c91}  & g_{c} - a_c \leq \tau^w_c  & \hspace{-20mm}  \forall  c \in C \text{ }   \\
-    \label{mod:c101}       & v_{d^0_{o_1}} \geq w_j - M\left(1- u_{jd^0_{o_1}}\right)   & \hspace{-50mm}    \forall o_1 \in O_i,  j \in \delta_{D}^{-}(d^0_{o_1}) \text{ }                \\
-    \label{mod:c10}        & v_{d^0_{o_1}}  \leq  w_j + \gamma^2 + M\left(1- u_{j,d^0_{o_1}}\right)     & \hspace{-40mm}  \forall  o_1 \in O,  j \in \delta_{D}^{-}(d^0_{o_1}) \text{ }                    \\
-    \label{mod:c11}        & \sum_{ \substack{o_1 \in O_c }}{ \sum_{ j \in \delta^{-}_{D}(d^0_{o_1})}}{u_{j,d^0_{o_1} } } = |O_c|-1      & \hspace{-40mm} \forall  c \in C,|O_c| > 1   \text{ }   \\
-    \label{mod:c12}        & \sum_{ \substack{o_1 \in O_c }}\sum_{ j \in \delta^{+}_{D}(d^0_{o_1})}{u_{d^0_{o_1},j } } = |O_c|-1      & \hspace{-40mm} \forall  c \in C, |O_c| > 1  \text{ }          \\
-    \label{mod:c13}    & \hspace{3mm} \smashoperator{ \sum_{ j \in \delta^{+}_{D}(d^0_{o_1})}}{u_{d^0_{o_1},j } } \leq 1    & \hspace{-20mm}  \forall o_1 \in O \text{ }         \\
-    \label{mod:c14}        & \hspace{2mm} \smashoperator{ \sum_{ j \in \delta^{-}_{D}(d^0_{o_1})}}{u_{j,d^0_{o_1} } } \leq 1      & \hspace{-20mm} \forall o_1 \in O \text{ }        \\
-    \label{mod:c15}        & \hspace{2mm} \smashoperator{ \sum_{ j \in \delta^{+}_{D}(d^0_{o_1})}}{u_{d^0_{o_1},j } } +\smashoperator{ \sum_{ j \in \delta^{-}_{D}(d^0_{o_1})}}{u_{j,d^0_{o_1} } } \geq 1    & \hspace{-50mm} \forall o_1 \in O \text{ }    \\
-    \label{mod:c16}        & v_{j+1} \geq w_{j} - M\left(1- u_{j,j+1}\right)   & \hspace{-50mm}  \forall o \in O, j \in D_{o}, j \geq 1  \text{ }    \\
-    \label{mod:c17}        & v_{j+1}  \leq  w_{j} + \gamma^1 + M\left(1- u_{j,j+1}\right)    & \hspace{-60mm} \forall c \in C, \forall o \in O_c, j \in D_{o}, j \geq 1 \text{ }     \\
-    \label{mod:c18}        & u_{j,j+1} \geq u_{j+1,j+2}      & \hspace{-70mm} \forall o_1 \in O,  j \in D_{o_1}, 1 \leq j \leq n_{o_{1}}-2 \text{ }   \\
-    \label{mod:c19}        & u_{j,j+1} \geq \sum_{l \in L}{x_{lj}}   & \hspace{-60mm}  \forall o_1 \in O, j \in D_{o_1}, j \geq 1 \text{ }                   \\
-    \label{mod:c20}        & v_{j} \geq w_{i} - M\left(1- u_{i,j}\right)   & \hspace{-40mm}  \forall i \in L, j \in \delta^{+}_{L}(i) \text{ }  \\
-    \label{mod:c21}        & \hspace{3mm} \smashoperator{ \sum_{ j \in \delta^{+}_{L}(i)}}{u_{i,j } } \leq 1     & \hspace{-40mm}  \forall i \in L \text{ }         \\
-    \label{mod:c22}        & \hspace{2mm} \smashoperator{ \sum_{ j \in \delta^{-}_{L}(i)}}{u_{j,i } } \leq 1                                      & \hspace{-40mm} \forall i \in L           \text{ }       \\
-    \label{mod:c230}       & \hspace{2mm} \smashoperator{ \sum_{ j \in \delta^{-}_{L}(i)}}{u_{j,i } }  \geq \sum_{k \in K}\sum_{j \in  \delta^{-}(i) }{x^k_{ji}}    & \hspace{-40mm} \forall i \in L \text{ }                                                      \\
-    \label{mod:c231}       & \hspace{2mm} \smashoperator{ \sum_{ j \in \delta^{-}_{L}(i)}}{u_{j,i } } \geq \smashoperator{ \sum_{ j \in \delta^{-}_{L}(i+1)}}{u_{j,i+1 } }   & \hspace{-20mm} \forall i \in L  \text{ }                                                        \\
-    \label{mod:c23}        & \hspace{2mm} \smashoperator{ \sum_{ j \in \delta^{-}_{L}(i)}}{u_{j,i } } + \smashoperator{ \sum_{ j \in \delta^{+}_{L}(i)}}{u_{i,j } } \geq \sum_{k \in K}\sum_{j \in  \delta^{-}(i) }{x^k_{ji}}    & \hspace{-20mm} \forall i \in L  \text{ }                                                        \\
-    \label{mod:c24}        & \sum_{k \in K}\sum_{j \in D_o}{q^k_j} = q_o            & \hspace{-40mm}  \forall o \in O  \text{ }           \\
-    \label{mod:c240}       & q^k_j \leq \sum_{i \in  \delta^{-}(j)}{Q^k x^k_{ij}}   & \hspace{-40mm}  \forall j \in D, k \in K  \text{ }                                              \\
+    \label{mod:c8}         & v_{d^0_{o}} \geq a_c    & \hspace{-50mm}  c \in C, o \in O_c \text{ }                           \\
+    \label{mod:c9}          & g_{c} \geq v_{d^0_{o_1}} - a_c - M\left(  \sum_{ j \in \delta^{-}_{D}(d^0_{o_1})  }{u_{jd^0_{o_1}}}\right)  & \hspace{-50mm}   c \in C, o_1 \in O_c \text{ }  \\
+    \label{mod:c91}  & g_{c} - a_c \leq \tau^w_c  & \hspace{-20mm}   c \in C \text{ }   \\
+    \label{mod:c101}       & v_{d^0_{o_1}} \geq w_j - M\left(1- u_{jd^0_{o_1}}\right)   & \hspace{-50mm}    o_1 \in O_i,  j \in \delta_{D}^{-}(d^0_{o_1}) \text{ }                \\
+    \label{mod:c10}        & v_{d^0_{o_1}}  \leq  w_j + \gamma^2 + M\left(1- u_{j,d^0_{o_1}}\right)     & \hspace{-40mm}   o_1 \in O,  j \in \delta_{D}^{-}(d^0_{o_1}) \text{ }                    \\
+    \label{mod:c11}        & \sum_{ \substack{o_1 \in O_c }}{ \sum_{ j \in \delta^{-}_{D}(d^0_{o_1})}}{u_{j,d^0_{o_1} } } = |O_c|-1      & \hspace{-40mm}  c \in C,|O_c| > 1   \text{ }   \\
+    \label{mod:c12}        & \sum_{ \substack{o_1 \in O_c }}\sum_{ j \in \delta^{+}_{D}(d^0_{o_1})}{u_{d^0_{o_1},j } } = |O_c|-1      & \hspace{-40mm}  c \in C, |O_c| > 1  \text{ }          \\
+    \label{mod:c13}    & \hspace{3mm} \smashoperator{ \sum_{ j \in \delta^{+}_{D}(d^0_{o_1})}}{u_{d^0_{o_1},j } } \leq 1    & \hspace{-20mm}  o_1 \in O \text{ }         \\
+    \label{mod:c14}        & \hspace{2mm} \smashoperator{ \sum_{ j \in \delta^{-}_{D}(d^0_{o_1})}}{u_{j,d^0_{o_1} } } \leq 1      & \hspace{-20mm} o_1 \in O \text{ }        \\
+    \label{mod:c15}        & \hspace{2mm} \smashoperator{ \sum_{ j \in \delta^{+}_{D}(d^0_{o_1})}}{u_{d^0_{o_1},j } } +\smashoperator{ \sum_{ j \in \delta^{-}_{D}(d^0_{o_1})}}{u_{j,d^0_{o_1} } } \geq 1    & \hspace{-50mm} o_1 \in O \text{ }    \\
+    \label{mod:c16}        & v_{j+1} \geq w_{j} - M\left(1- u_{j,j+1}\right)   & \hspace{-50mm}  o \in O, j \in D_{o}, j \geq 1  \text{ }    \\
+    \label{mod:c17}        & v_{j+1}  \leq  w_{j} + \gamma^1 + M\left(1- u_{j,j+1}\right)    & \hspace{-60mm} c \in C, o \in O_c, j \in D_{o}, j \geq 1 \text{ }     \\
+    \label{mod:c18}        & u_{j,j+1} \geq u_{j+1,j+2}      & \hspace{-70mm} o_1 \in O,  j \in D_{o_1}, 1 \leq j \leq n_{o_{1}}-2 \text{ }   \\
+    \label{mod:c19}        & u_{j,j+1} \geq \sum_{l \in L}{x_{lj}}   & \hspace{-60mm}  o_1 \in O, j \in D_{o_1}, j \geq 1 \text{ }                   \\
+    \label{mod:c20}        & v_{j} \geq w_{i} - M\left(1- u_{i,j}\right)   & \hspace{-40mm}  i \in L, j \in \delta^{+}_{L}(i) \text{ }  \\
+    \label{mod:c21}        & \hspace{3mm} \smashoperator{ \sum_{ j \in \delta^{+}_{L}(i)}}{u_{i,j } } \leq 1     & \hspace{-40mm}  i \in L \text{ }         \\
+    \label{mod:c22}        & \hspace{2mm} \smashoperator{ \sum_{ j \in \delta^{-}_{L}(i)}}{u_{j,i } } \leq 1                                      & \hspace{-40mm} i \in L           \text{ }       \\
+    \label{mod:c230}       & \hspace{2mm} \smashoperator{ \sum_{ j \in \delta^{-}_{L}(i)}}{u_{j,i } }  \geq \sum_{k \in K}\sum_{j \in  \delta^{-}(i) }{x^k_{ji}}    & \hspace{-40mm} i \in L \text{ }                                                      \\
+    \label{mod:c231}       & \hspace{2mm} \smashoperator{ \sum_{ j \in \delta^{-}_{L}(i)}}{u_{j,i } } \geq \smashoperator{ \sum_{ j \in \delta^{-}_{L}(i+1)}}{u_{j,i+1 } }   & \hspace{-20mm} i \in L  \text{ }                                                        \\
+    \label{mod:c23}        & \hspace{2mm} \smashoperator{ \sum_{ j \in \delta^{-}_{L}(i)}}{u_{j,i } } + \smashoperator{ \sum_{ j \in \delta^{+}_{L}(i)}}{u_{i,j } } \geq \sum_{k \in K}\sum_{j \in  \delta^{-}(i) }{x^k_{ji}}    & \hspace{-20mm} i \in L  \text{ }                                                        \\
+    \label{mod:c24}        & \sum_{k \in K}\sum_{j \in D_o}{q^k_j} = q_o            & \hspace{-40mm}  o \in O  \text{ }           \\
+    \label{mod:c240}       & q^k_j \leq \sum_{i \in  \delta^{-}(j)}{Q^k x^k_{ij}}   & \hspace{-40mm}  j \in D, k \in K  \text{ }                                              \\
     \label{mod:c25}        & \sum_{k \in K}\sum_{j \in \delta^{+}(i)}{x^{k}_{ij} \leq 1}     & \hspace{-40mm} i \in L \cup D       \text{ }          \\
     \label{mod:c26}        & \sum_{k \in K}\sum_{j \in \delta^{+}(i)} {x^{k}_{ij} } =  \sum_{k \in K}\sum_{j \in \delta^{-}(i)}{x^{k}_{ji} }        & \hspace{-40mm}  i \in L \cup D  \text{ }       \\
-    \label{mod:c29}        & w^1_k \geq  \mu + h_k - v_{e_k}            & \hspace{-40mm} \forall k \in K   \text{ }         \\
-    \label{mod:c30}        & w^2_k \geq (v_{e_k} - h_k) - \eta   & \hspace{-40mm} \forall k \in K         \text{ }         \\
+    \label{mod:c29}        & w^1_k \geq  \mu + h_k - v_{e_k}            & \hspace{-40mm} k \in K   \text{ }         \\
+    \label{mod:c30}        & w^2_k \geq (v_{e_k} - h_k) - \eta   & \hspace{-40mm} k \in K         \text{ }         \\
     \label{mod:c31}   & w_{e_k} \leq h_k + \varLambda  & \hspace{-20mm}  k \in K \hspace{0mm}       \text{ }                                            \\
@@
 
-            $Sol \leftarrow Sol \backslash \{(L^k_{i}, U^k_{i}), \forall i \in D_c\}$
+            $Sol \leftarrow Sol \backslash \{(L^k_{i}, U^k_{i}), i \in D_c\}$
 
@@
 
-$Sol \leftarrow Sol \backslash \{(L^k_{i}, U^k_{i}), \forall i \in (D^{j'}_{o},..,D^{j-1}_{o})  \}$
+$Sol \leftarrow Sol \backslash \{(L^k_{i}, U^k_{i}), i \in (D^{j'}_{o},..,D^{j-1}_{o})  \}$
 
@@
 
-    $Sol \leftarrow Sol \backslash \{(L^k_{i}, U^k_{i}), \forall i \in D_c\}$
+    $Sol \leftarrow Sol \backslash \{(L^k_{i}, U^k_{i}), i \in D_c\}$
 
@@
 
-    $Sol \leftarrow Sol \backslash \{(L^k_{i}, U^k_{i}), \forall i \in (D^{j'}_{o},..,D^{j-1}_{o})  \}$
+    $Sol \leftarrow Sol \backslash \{(L^k_{i}, U^k_{i}), i \in (D^{j'}_{o},..,D^{j-1}_{o})  \}$
 
@@
 
-            $Sol \leftarrow Sol \backslash \{(L^k_j,U^k_j), \forall j \in D_c \}$
+            $Sol \leftarrow Sol \backslash \{(L^k_j,U^k_j), j \in D_c \}$
            }
@@
 % \clearpage
+
 \section{Computational experiments}
